### 2017-01-16 Geological Inference based on Kinematic Structural Models/Experiment 1/PGM1_clean.tex
initial commit alex thesis

% Default to the notebook output style

% Inherit from the specified cell style.

\documentclass{article}

    \usepackage{graphicx} % Used to insert images
    \usepackage{adjustbox} % Used to constrain images to a maximum size 
    \usepackage{color} % Allow colors to be defined
    \usepackage{enumerate} % Needed for markdown enumerations to work
    \usepackage{geometry} % Used to adjust the document margins
    \usepackage{amsmath} % Equations
    \usepackage{amssymb} % Equations
    \usepackage{eurosym} % defines \euro
    \usepackage[mathletters]{ucs} % Extended unicode (utf-8) support
    \usepackage[utf8x]{inputenc} % Allow utf-8 characters in the tex document
    \usepackage{fancyvrb} % verbatim replacement that allows latex
    \usepackage{grffile} % extends the file name processing of package graphics 
                         % to support a larger range 
    % The hyperref package gives us a pdf with properly built
    % internal navigation ('pdf bookmarks' for the table of contents,
    % internal cross-reference links, web links for URLs, etc.)
    \usepackage{hyperref}
    \usepackage{longtable} % longtable support required by pandoc >1.10
    \usepackage{booktabs}  % table support for pandoc > 1.12.2
    \usepackage{ulem} % ulem is needed to support strikethroughs (\sout)

    \definecolor{orange}{cmyk}{0,0.4,0.8,0.2}
    \definecolor{darkorange}{rgb}{.71,0.21,0.01}
    \definecolor{darkgreen}{rgb}{.12,.54,.11}
    \definecolor{myteal}{rgb}{.26, .44, .56}
    \definecolor{gray}{gray}{0.45}
    \definecolor{lightgray}{gray}{.95}
    \definecolor{mediumgray}{gray}{.8}
    \definecolor{inputbackground}{rgb}{.95, .95, .85}
    \definecolor{outputbackground}{rgb}{.95, .95, .95}
    \definecolor{traceback}{rgb}{1, .95, .95}
    % ansi colors
    \definecolor{red}{rgb}{.6,0,0}
    \definecolor{green}{rgb}{0,.65,0}
    \definecolor{brown}{rgb}{0.6,0.6,0}
    \definecolor{blue}{rgb}{0,.145,.698}
    \definecolor{purple}{rgb}{.698,.145,.698}
    \definecolor{cyan}{rgb}{0,.698,.698}
    \definecolor{lightgray}{gray}{0.5}

    % bright ansi colors
    \definecolor{darkgray}{gray}{0.25}
    \definecolor{lightred}{rgb}{1.0,0.39,0.28}
    \definecolor{lightgreen}{rgb}{0.48,0.99,0.0}
    \definecolor{lightblue}{rgb}{0.53,0.81,0.92}
    \definecolor{lightpurple}{rgb}{0.87,0.63,0.87}
    \definecolor{lightcyan}{rgb}{0.5,1.0,0.83}

    % commands and environments needed by pandoc snippets
    % extracted from the output of `pandoc -s`
    \providecommand{\tightlist}{%
      \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
    \DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
    % Add ',fontsize=\small' for more characters per line
    \newenvironment{Shaded}{}{}
    \newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
    \newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{{#1}}}
    \newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{{#1}}}}
    \newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{{#1}}}
    \newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
    \newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{{#1}}}
    \newcommand{\RegionMarkerTok}[1]{{#1}}
    \newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
    \newcommand{\NormalTok}[1]{{#1}}

    % Additional commands for more recent versions of Pandoc
    \newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{{#1}}}
    \newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{{#1}}}
    \newcommand{\ImportTok}[1]{{#1}}
    \newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{{#1}}}}
    \newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
    \newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
    \newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{{#1}}}
    \newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
    \newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{{#1}}}
    \newcommand{\BuiltInTok}[1]{{#1}}
    \newcommand{\ExtensionTok}[1]{{#1}}
    \newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{{#1}}}
    \newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{{#1}}}
    \newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
    \newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}

    % Define a nice break command that doesn't care if a line doesn't already
    % exist.
    \def\br{\hspace*{\fill} \\* }
    % Math Jax compatability definitions
    \def\gt{>}
    \def\lt{<}
    % Document parameters
    \title{PGM1\_clean}

    % Pygments definitions

\makeatletter
\def\PY@reset{\let\PY@it=\relax \let\PY@bf=\relax%
    \let\PY@ul=\relax \let\PY@tc=\relax%
    \let\PY@bc=\relax \let\PY@ff=\relax}
\def\PY@tok#1{\csname PY@tok@#1\endcsname}
\def\PY@toks#1+{\ifx\relax#1\empty\else%
    \PY@tok{#1}\expandafter\PY@toks\fi}
\def\PY@do#1{\PY@bc{\PY@tc{\PY@ul{%
    \PY@it{\PY@bf{\PY@ff{#1}}}}}}}
\def\PY#1#2{\PY@reset\PY@toks#1+\relax+\PY@do{#2}}

\expandafter\def\csname PY@tok@gd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@gu\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.50,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@gt\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.27,0.87}{##1}}}
\expandafter\def\csname PY@tok@gs\endcsname{\let\PY@bf=\textbf}
\expandafter\def\csname PY@tok@gr\endcsname{\def\PY@tc##1{\textcolor[rgb]{1.00,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@cm\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@vg\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@vi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@mh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@cs\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@ge\endcsname{\let\PY@it=\textit}
\expandafter\def\csname PY@tok@vc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@il\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@go\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.53,0.53}{##1}}}
\expandafter\def\csname PY@tok@cp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.74,0.48,0.00}{##1}}}
\expandafter\def\csname PY@tok@gi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.63,0.00}{##1}}}
\expandafter\def\csname PY@tok@gh\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@ni\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.60,0.60,0.60}{##1}}}
\expandafter\def\csname PY@tok@nl\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.63,0.00}{##1}}}
\expandafter\def\csname PY@tok@nn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@no\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@na\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.49,0.56,0.16}{##1}}}
\expandafter\def\csname PY@tok@nb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@nc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@nd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
\expandafter\def\csname PY@tok@ne\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.82,0.25,0.23}{##1}}}
\expandafter\def\csname PY@tok@nf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@si\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
\expandafter\def\csname PY@tok@s2\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@nt\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@nv\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@s1\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@ch\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@m\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@gp\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@sh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@ow\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
\expandafter\def\csname PY@tok@sx\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@bp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@c1\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@o\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@kc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@c\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@mf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@err\endcsname{\def\PY@bc##1{\setlength{\fboxsep}{0pt}\fcolorbox[rgb]{1.00,0.00,0.00}{1,1,1}{\strut ##1}}}
\expandafter\def\csname PY@tok@mb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@ss\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@sr\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
\expandafter\def\csname PY@tok@mo\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@kd\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@mi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@kn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@cpf\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@kr\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@s\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@kp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@w\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.73,0.73}{##1}}}
\expandafter\def\csname PY@tok@kt\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.69,0.00,0.25}{##1}}}
\expandafter\def\csname PY@tok@sc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@sb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@k\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@se\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.13}{##1}}}
\expandafter\def\csname PY@tok@sd\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}

\def\PYZbs{\char`\\}
\def\PYZus{\char`\_}
\def\PYZob{\char`\{}
\def\PYZcb{\char`\}}
\def\PYZca{\char`\^}
\def\PYZam{\char`\&}
\def\PYZlt{\char`\<}
\def\PYZgt{\char`\>}
\def\PYZsh{\char`\#}
\def\PYZpc{\char`\%}
\def\PYZdl{\char`\$}
\def\PYZhy{\char`\-}
\def\PYZsq{\char`\'}
\def\PYZdq{\char`\"}
\def\PYZti{\char`\~}
% for compatibility with earlier versions
\def\PYZat{@}
\def\PYZlb{[}
\def\PYZrb{]}
\makeatother

    % Exact colors from NB
    \definecolor{incolor}{rgb}{0.0, 0.0, 0.5}
    \definecolor{outcolor}{rgb}{0.545, 0.0, 0.0}

    % Prevent overflowing lines due to hard-to-break entities
    \sloppy 
    % Setup hyperref package
    \hypersetup{
      breaklinks=true,  % so long urls are correctly broken across lines
      colorlinks=true,
      urlcolor=blue,
      linkcolor=darkorange,
      citecolor=darkgreen,
      }
    % Slightly bigger margins than the latex defaults

    \geometry{verbose,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}

    \begin{document}

    \maketitle

    \section{Experiment 1}\label{experiment-1}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor} }]:} \PY{k+kn}{import} \PY{n+nn}{numpy} \PY{k+kn}{as} \PY{n+nn}{np}
        \PY{k+kn}{import} \PY{n+nn}{pymc}
        \PY{k+kn}{import} \PY{n+nn}{json}
        \PY{k+kn}{import} \PY{n+nn}{matplotlib.pyplot} \PY{k+kn}{as} \PY{n+nn}{plt}
        \PY{k+kn}{import} \PY{n+nn}{matplotlib} \PY{k+kn}{as} \PY{n+nn}{mpl}
        \PY{o}{\PYZpc{}}\PY{k}{matplotlib} inline
        \PY{n}{plt}\PY{o}{.}\PY{n}{style}\PY{o}{.}\PY{n}{use}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{seaborn\PYZhy{}colorblind}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}

        \PY{k+kn}{import} \PY{n+nn}{sys}\PY{o}{,} \PY{n+nn}{os}
        \PY{n}{sys}\PY{o}{.}\PY{n}{path}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{C:/Users/lebobcrash/Documents/GitHub/pynoddy/}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
        \PY{n}{sys}\PY{o}{.}\PY{n}{path}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{../pynoddy/}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
        \PY{k+kn}{import} \PY{n+nn}{pynoddy.experiment}
        \PY{k+kn}{import} \PY{n+nn}{pynoddy.history}
        \PY{k+kn}{import} \PY{n+nn}{pynoddy.output}

        \PY{k+kn}{import} \PY{n+nn}{seaborn} \PY{k+kn}{as} \PY{n+nn}{sns}
        \PY{n}{sns}\PY{o}{.}\PY{n}{set\PYZus{}style}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{white}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
        \PY{n}{sns}\PY{o}{.}\PY{n}{set\PYZus{}context}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{paper}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{font\PYZus{}scale}\PY{o}{=}\PY{l+m+mf}{1.3}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor} }]:} \PY{n}{initial\PYZus{}history} \PY{o}{=} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{init.his}\PY{l+s+s2}{\PYZdq{}} \PY{c+c1}{\PYZsh{} set initial noddy history file}
        \PY{n}{cs} \PY{o}{=} \PY{l+m+mi}{10} \PY{c+c1}{\PYZsh{} cube size}
        \PY{n}{ws} \PY{o}{=} \PY{l+m+mi}{900} \PY{c+c1}{\PYZsh{} well site (x\PYZhy{}coord)}
\end{Verbatim}

    Create prior pymc Distributions for the fold amplitude and wavelength:

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor} }]:} \PY{n}{fold\PYZus{}amplitude} \PY{o}{=} \PY{n}{pymc}\PY{o}{.}\PY{n}{Normal}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{fold\PYZus{}amplitude}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{l+m+mf}{350.}\PY{p}{,} \PY{l+m+mf}{1.}\PY{o}{/}\PY{n}{np}\PY{o}{.}\PY{n}{square}\PY{p}{(}\PY{l+m+mf}{25.}\PY{p}{)}\PY{p}{)}
        \PY{n}{fold\PYZus{}wavelength} \PY{o}{=} \PY{n}{pymc}\PY{o}{.}\PY{n}{Normal}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{fold\PYZus{}wavelength}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{l+m+mf}{2500.}\PY{p}{,} \PY{l+m+mf}{1.}\PY{o}{/}\PY{n}{np}\PY{o}{.}\PY{n}{square}\PY{p}{(}\PY{l+m+mf}{140.}\PY{p}{)}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor} }]:} \PY{n}{fold\PYZus{}amplitude2} \PY{o}{=} \PY{n}{pymc}\PY{o}{.}\PY{n}{Normal}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{fold\PYZus{}amplitude}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{l+m+mf}{350.}\PY{p}{,} \PY{l+m+mf}{1.}\PY{o}{/}\PY{n}{np}\PY{o}{.}\PY{n}{square}\PY{p}{(}\PY{l+m+mf}{25.}\PY{p}{)}\PY{p}{)}
        \PY{n}{fold\PYZus{}wavelength2} \PY{o}{=} \PY{n}{pymc}\PY{o}{.}\PY{n}{Normal}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{fold\PYZus{}wavelength}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{l+m+mf}{2500.}\PY{p}{,} \PY{l+m+mf}{1.}\PY{o}{/}\PY{n}{np}\PY{o}{.}\PY{n}{square}\PY{p}{(}\PY{l+m+mf}{140.}\PY{p}{)}\PY{p}{)}
\end{Verbatim}

    Create the deterministic pymc function, which takes the prior
distributions to create the model:

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor} }]:} \PY{n}{ex\PYZus{}pymc} \PY{o}{=} \PY{n}{pynoddy}\PY{o}{.}\PY{n}{experiment}\PY{o}{.}\PY{n}{Experiment}\PY{p}{(}\PY{n}{initial\PYZus{}history}\PY{p}{)}
\end{Verbatim}

    \subsection{PYMC Functions}\label{pymc-functions}

\subsubsection{Deterministic Modelling
Function}\label{deterministic-modelling-function}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor} }]:} \PY{n+nd}{@pymc.deterministic}
        \PY{k}{def} \PY{n+nf}{pynoddy\PYZus{}model}\PY{p}{(}\PY{n}{value}\PY{o}{=}\PY{l+m+mi}{0}\PY{p}{,}
                          \PY{n}{fold\PYZus{}amplitude}\PY{o}{=}\PY{n}{fold\PYZus{}amplitude}\PY{p}{,}
                          \PY{n}{fold\PYZus{}wavelength}\PY{o}{=}\PY{n}{fold\PYZus{}wavelength}\PY{p}{,} 
                          \PY{n}{ex}\PY{o}{=}\PY{n}{ex\PYZus{}pymc}\PY{p}{,} \PY{n}{ws}\PY{o}{=}\PY{n}{ws}\PY{p}{,} \PY{n}{cs}\PY{o}{=}\PY{n}{cs}\PY{p}{)}\PY{p}{:}
            \PY{c+c1}{\PYZsh{} set fold event properties to prior parameter draws:}
            \PY{n}{ex}\PY{o}{.}\PY{n}{events}\PY{p}{[}\PY{l+m+mi}{2}\PY{p}{]}\PY{o}{.}\PY{n}{properties}\PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Amplitude}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]} \PY{o}{=} \PY{n}{fold\PYZus{}amplitude}
            \PY{n}{ex}\PY{o}{.}\PY{n}{events}\PY{p}{[}\PY{l+m+mi}{2}\PY{p}{]}\PY{o}{.}\PY{n}{properties}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Wavelength}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]} \PY{o}{=} \PY{n}{fold\PYZus{}wavelength}
            \PY{c+c1}{\PYZsh{} drillhole extraction (1\PYZhy{}D model for faster simulation)}
            \PY{n}{well} \PY{o}{=} \PY{n}{ex}\PY{o}{.}\PY{n}{get\PYZus{}drillhole\PYZus{}data}\PY{p}{(}\PY{n}{ws}\PY{p}{,}\PY{l+m+mi}{0}\PY{p}{)} 
            \PY{k}{return} \PY{n}{well}
\end{Verbatim}

    \subsubsection{Stochastic Likelihood
Function}\label{stochastic-likelihood-function}

Create the stochastic likelihood function for the z-position of layer 2:

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor} }]:} \PY{n+nd}{@pymc.stochastic}
        \PY{k}{def} \PY{n+nf}{like\PYZus{}layer2\PYZus{}height}\PY{p}{(}\PY{n}{value}\PY{o}{=}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{well}\PY{o}{=}\PY{n}{pynoddy\PYZus{}model}\PY{p}{)}\PY{p}{:}
            \PY{k}{if} \PY{n+nb}{len}\PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{where}\PY{p}{(}\PY{n}{well}\PY{o}{==}\PY{l+m+mi}{2}\PY{p}{)}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{)} \PY{o}{\PYZlt{}} \PY{l+m+mi}{1}\PY{p}{:}
                \PY{n}{layer2\PYZus{}height} \PY{o}{=} \PY{o}{\PYZhy{}}\PY{l+m+mi}{9999}
            \PY{k}{else}\PY{p}{:}
                \PY{n}{layer2\PYZus{}height} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{where}\PY{p}{(}\PY{n}{well}\PY{o}{==}\PY{l+m+mi}{2}\PY{p}{)}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}
            \PY{k}{return} \PY{n}{pymc}\PY{o}{.}\PY{n}{normal\PYZus{}like}\PY{p}{(}\PY{n}{layer2\PYZus{}height}\PY{p}{,} \PY{l+m+mf}{550.}\PY{p}{,} \PY{l+m+mf}{1.}\PY{o}{/}\PY{n}{np}\PY{o}{.}\PY{n}{square}\PY{p}{(}\PY{l+m+mf}{25.}\PY{p}{)}\PY{p}{)}
\end{Verbatim}

    \subsection{Simulation}\label{simulation}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor} }]:} \PY{n}{iterations} \PY{o}{=} \PY{l+m+mi}{30000}
\end{Verbatim}

    \subsubsection{Bayesian Inference}\label{bayesian-inference}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor} }]:} \PY{n}{model} \PY{o}{=} \PY{n}{pymc}\PY{o}{.}\PY{n}{Model}\PY{p}{(}\PY{p}{[}\PY{n}{fold\PYZus{}wavelength}\PY{p}{,} \PY{n}{layer\PYZus{}height}\PY{p}{,}
                            \PY{n}{pynoddy\PYZus{}model}\PY{p}{,} \PY{n}{like\PYZus{}layer2\PYZus{}height}\PY{p}{]}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor} }]:} \PY{n}{BI\PYZus{}RUN} \PY{o}{=} \PY{n}{pymc}\PY{o}{.}\PY{n}{MCMC}\PY{p}{(}\PY{n}{model}\PY{p}{,} \PY{n}{db}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{hdf5}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{name}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{BI\PYZus{}database\PYZus{}name}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
        \PY{n}{BI\PYZus{}RUN}\PY{o}{.}\PY{n}{sample}\PY{p}{(}\PY{n+nb}{iter}\PY{o}{=}\PY{n}{iterations}\PY{p}{)}
\end{Verbatim}

    \subsubsection{Monte Carlo Forward
Simulation}\label{monte-carlo-forward-simulation}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor} }]:} \PY{n}{model\PYZus{}mcfs} \PY{o}{=} \PY{n}{pymc}\PY{o}{.}\PY{n}{Model}\PY{p}{(}\PY{p}{[}\PY{n}{layer\PYZus{}height2}\PY{p}{,} \PY{n}{fold\PYZus{}wavelength2}\PY{p}{,} 
                                 \PY{n}{pynoddy\PYZus{}model}\PY{p}{]}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor} }]:} \PY{n}{MCFS\PYZus{}RUN} \PY{o}{=} \PY{n}{pymc}\PY{o}{.}\PY{n}{MCMC}\PY{p}{(}\PY{n}{model\PYZus{}mcfs}\PY{p}{,} \PY{n}{db}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{hdf5}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{name}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{MCFS\PYZus{}database\PYZus{}name}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
        \PY{n}{MCFS\PYZus{}RUN}\PY{o}{.}\PY{n}{sample}\PY{p}{(}\PY{n+nb}{iter}\PY{o}{=}\PY{n}{iterations}\PY{p}{)}
\end{Verbatim}

    \subsection{Results}\label{results}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor} }]:} \PY{n}{fig}\PY{p}{,} \PY{n}{ax} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{subplots}\PY{p}{(}\PY{n}{nrows}\PY{o}{=}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{ncols}\PY{o}{=}\PY{l+m+mi}{2}\PY{p}{,} \PY{n}{figsize}\PY{o}{=}\PY{p}{(}\PY{l+m+mi}{9}\PY{p}{,}\PY{l+m+mi}{4}\PY{p}{)}\PY{p}{,} \PY{n}{sharey}\PY{o}{=}\PY{n+nb+bp}{True}\PY{p}{)}

        \PY{n}{ax}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{o}{.}\PY{n}{set\PYZus{}xlim}\PY{p}{(}\PY{l+m+mi}{1400}\PY{p}{,}\PY{l+m+mi}{3200}\PY{p}{)}
        \PY{n}{ax}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{o}{.}\PY{n}{set\PYZus{}xlim}\PY{p}{(}\PY{l+m+mi}{1400}\PY{p}{,}\PY{l+m+mi}{3200}\PY{p}{)}
        \PY{n}{ax}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{o}{.}\PY{n}{set\PYZus{}xticks}\PY{p}{(}\PY{p}{[}\PY{l+m+mi}{1400}\PY{p}{,}\PY{l+m+mi}{1850}\PY{p}{,}\PY{l+m+mi}{2300}\PY{p}{,}\PY{l+m+mi}{2750}\PY{p}{,}\PY{l+m+mi}{3200}\PY{p}{]}\PY{p}{)}
        \PY{n}{ax}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{o}{.}\PY{n}{set\PYZus{}xticks}\PY{p}{(}\PY{p}{[}\PY{l+m+mi}{1400}\PY{p}{,}\PY{l+m+mi}{1850}\PY{p}{,}\PY{l+m+mi}{2300}\PY{p}{,}\PY{l+m+mi}{2750}\PY{p}{,}\PY{l+m+mi}{3200}\PY{p}{]}\PY{p}{)}

        \PY{n}{sns}\PY{o}{.}\PY{n}{regplot}\PY{p}{(}\PY{n}{MCFS\PYZus{}RUN}\PY{o}{.}\PY{n}{trace}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{fold\PYZus{}wavelength}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{[}\PY{p}{:}\PY{p}{]}\PY{p}{,}
                    \PY{n}{MCFS\PYZus{}RUN}\PY{o}{.}\PY{n}{trace}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{fold\PYZus{}amplitude}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{[}\PY{p}{:}\PY{p}{]}\PY{p}{,} 
                    \PY{n}{ax}\PY{o}{=}\PY{n}{ax}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{)}

        \PY{n}{sns}\PY{o}{.}\PY{n}{regplot}\PY{p}{(}\PY{n}{BI\PYZus{}RUN}\PY{o}{.}\PY{n}{trace}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{fold\PYZus{}wavelength}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{[}\PY{p}{:}\PY{p}{]}\PY{p}{,}
                    \PY{n}{BI\PYZus{}RUN}\PY{o}{.}\PY{n}{trace}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{fold\PYZus{}amplitude}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{p}{[}\PY{p}{:}\PY{p}{]}\PY{p}{,}
                    \PY{n}{color}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{darkorange}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{ax}\PY{o}{=}\PY{n}{ax}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{)}

        \PY{n}{ax}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{o}{.}\PY{n}{set\PYZus{}title}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Monte Carlo Forward Simulation}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
        \PY{n}{ax}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{o}{.}\PY{n}{set\PYZus{}title}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Bayesian Inference}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
        \PY{n}{ax}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{o}{.}\PY{n}{set\PYZus{}ylabel}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Fold Amplitude [m]}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
        \PY{n}{ax}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{o}{.}\PY{n}{set\PYZus{}xlabel}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Fold Wavelength [m]}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
        \PY{n}{ax}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{o}{.}\PY{n}{set\PYZus{}xlabel}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Fold Wavelength [m]}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
        \PY{n}{plt}\PY{o}{.}\PY{n}{tight\PYZus{}layout}\PY{p}{(}\PY{p}{)}
\end{Verbatim}

    % Add a bibliography block to the postdoc

    \end{document}

### 2017-01-16 Geological Inference based on Kinematic Structural Models/Experiment 2/PosteriorAnalysis/Posteriors2.py

@@ -1,0 +1,291 @@
+# Posteriors Class to handle Posterior Outputs
+import numpy as np
+import matplotlib.pyplot as plt
+
+from sklearn.neighbors import KernelDensity
+from sklearn.grid_search import GridSearchCV
+from scipy.stats.kde import gaussian_kde
+
+import json
+import pymc
+
+import sys
+sys.path.append("C:/Users/lebobcrash/Documents/GitHub/pynoddy/")
+
+import pynoddy.history
+import pynoddy.output
+import pynoddy.experiment
+
+class Posteriors:
+    """Handles PYMC run posteriors for pynoddy graben simulations."""
+    def __init__(self, parameters, sim_type):
+        self.parameters_filename = parameters
+        self.sim_type = sim_type # pynoddy or gempy
+
+        # load parameter list
+        self.parameters = []
+        file = open(self.parameters_filename+'.txt', 'r')
+        for line in file:
+            key = line.rstrip('\n')
+            self.parameters.append(key)
+            #= np.load(self.name+"_posterior_"+key+".npy")
+        
+        # load posteriors
+        self.posteriors = {}
+        for entry in self.parameters:
+            self.posteriors[entry] = np.load(entry+"_posterior.npy")
+
+        self._layers = np.unique(self.posteriors[self.sim_type+"_model"][1])
+        self._n_layers = len(self._layers)
+        self._n_iter = len(self.posteriors[self.sim_type+"_model"])
+        self.x_extent = len(self.posteriors[self.sim_type+"_model"][0,:,0])
+        self.z_extent = len(self.posteriors[self.sim_type+"_model"][0,0,:])
+        self._NaN = -9999
+     
+    def plot_section(self, n, ax=None):
+        if ax == None:
+            ax = plt.axes()
+
+        ax.imshow(self.posteriors[self.sim_type+"_model"][n,:,:])
+
+    def posteriors_browse_sections(self):
+        from IPython.html.widgets import interact
+        interact(self.posteriors_plot_section, n=(0,self._n_iter - 1,1), cube_size=(10,50,5))
+
+    def posteriors_plot_fault_dip(self, nbins=24):
+        plt.hist(self.posteriors["fault_E_dip"], label="Fault E", alpha=0.65, normed=True, bins=nbins)
+        plt.hist(self.posteriors["fault_W_dip"], label="Fault W", alpha=0.65, normed=True, bins=nbins)
+        plt.legend()
[…]

### 2017-01-16 Geological Inference based on Kinematic Structural Models/Experiment 2/PosteriorAnalysis/__init__.py

@@ -1,0 +1,202 @@
+# Posteriors Class to handle Posterior Outputs
+import numpy as np
+import matplotlib.pyplot as plt
+
+from sklearn.neighbors import KernelDensity
+from sklearn.grid_search import GridSearchCV
+from scipy.stats.kde import gaussian_kde
+
+import json
+import pymc
+
+import sys
+sys.path.append("C:/Users/lebobcrash/Documents/GitHub/pynoddy/")
+
+import pynoddy.history
+import pynoddy.output
+import pynoddy.experiment
+
+class PosteriorAnalysis:
+    """Handles PYMC run posteriors for pynoddy graben simulations."""
+    def __init__(self, parameters, sim_type):
+        self.parameters_filename = parameters
+        self.sim_type = sim_type # pynoddy or gempy
+
+        # load parameter list
+        self.parameters = []
+        file = open(self.parameters_filename+'.txt', 'r')
+        for line in file:
+            key = line.rstrip('\n')
+            self.parameters.append(key)
+            #= np.load(self.name+"_posterior_"+key+".npy")
+        
+        # load posteriors
+        self.posteriors = {}
+        for entry in self.parameters:
+            self.posteriors[entry] = np.load(entry+"_posterior.npy")
+
+        self._layers = np.unique(self.posteriors[self.sim_type+"_model"][1])
+        self._n_layers = len(self._layers)
+        self._n_iter = len(self.posteriors[self.sim_type+"_model"])
+        self.x_extent = len(self.posteriors[self.sim_type+"_model"][0,:,0])
+        self.z_extent = len(self.posteriors[self.sim_type+"_model"][0,0,:])
+        self._NaN = -9999
+     
+    def plot_section(self, n):
+        ax = None
+        if ax == None:
+            ax = plt.axes()
+
+        ax.imshow(self.posteriors[self.sim_type+"_model"][n,:,:].T, origin="lower")
+
+    def browse_sections(self):
+        from IPython.html.widgets import interact
+        interact(self.plot_section, n=(0,self._n_iter - 1,1))
+
+    def compute_entropy(self):
+        """
+        Calculates the per-voxel entropy of the model section.
+        Return: numpy ndarray
+        """
+        lith_count = np.zeros_like(self.posteriors[self.sim_type+"_model"][0:self._n_layers])
+
+        for x in range(self.x_extent):
+            for z in range(self.z_extent):
+                for layer in range(self._n_layers):
[…]

### 2017-01-16 Geological Inference based on Kinematic Structural Models/Experiment 3/pymc_noddy_output_p.pl

@@ -1,0 +1,35 @@
+GOCAD PLine 1
+HEADER {name: pymc_noddy_output_p
+}
+PROPERTIES DiffType DiffAge Num 
+PROPERTY_CLASSES DiffType DiffAge Num
+
+ILINE
+PVRTX 1 467.204487 10.000000 889.719586 3 3 692
+PVRTX 2 181.465753 10.000000 622.630137 3 3 692
+SEG 1 2
+ILINE
+PVRTX 1 181.465753 10.000000 622.630137 0 0 796
+PVRTX 2 1011.984193 10.000000 591.955075 0 0 796
+SEG 1 2
+ILINE
+PVRTX 1 1011.984193 10.000000 591.955075 0 0 2808
+PVRTX 2 994.720786 10.000000 422.844881 0 0 2808
+SEG 1 2
+ILINE
+PVRTX 1 994.720786 10.000000 422.844881 0 0 2624
+PVRTX 2 895.159814 10.000000 205.858998 0 0 2624
+SEG 1 2
+ILINE
+PVRTX 1 467.204487 10.000000 889.719586 3 3 596
+PVRTX 2 1011.984193 10.000000 591.955075 3 3 596
+SEG 1 2
+ILINE
+PVRTX 1 467.204487 10.000000 889.719586 3 3 496
+PVRTX 2 1663.529412 10.000000 749.865084 3 3 496
+SEG 1 2
+ILINE
+PVRTX 1 1011.984193 10.000000 591.955075 0 0 1412
+PVRTX 2 1663.529412 10.000000 749.865084 0 0 1412
+SEG 1 2
+END

### 2017-01-16 Geological Inference based on Kinematic Structural Models/PosteriorAnalysis/Utility.py

@@ -1,0 +1,148 @@
+import sys; sys.path.append("C:/Users/lebobcrash/Documents/GitHub/pynoddy/")
+import numpy as np
+import json
+import pynoddy.history
+import pynoddy.output
+import pynoddy.experiment
+import pandas as pn
+
+
+def calc_fucking_foliations(interface_df):
+    test_sorted = interface_df.sort_values("formation")
+    _foliation_col_names = ["X", "Y", "Z", "azimuth", "dip", "polarity", "formation"]
+    foliations = pn.DataFrame(columns=_foliation_col_names)
+
+    for i,row in enumerate(test_sorted.values):
+        #print i,row
+        if row[3] == "Layer3" or row[3] == "Layer5":
+            if i != len(test_sorted.values)-1:
+                if row[3] == test_sorted.values[i+1][3]:
+                    dx = row[0] - test_sorted.values[i+1][0]
+                    dz = row[2] - test_sorted.values[i+1][2]
+
+                    dip = np.rad2deg(np.arctan(dz/float(dx)))
+                    #print dip
+                    foliations.loc[len(foliations)] = [np.abs((row[0] + test_sorted.values[i+1][0])/2),
+                                                       0,
+                                                         np.abs((row[2] + test_sorted.values[i+1][2]) / 2),
+                                                         -90,
+                                                         dip,
+                                                         1.0,
+                                                         row[3]]
+    return foliations
+
+
+def noddy_extract_lh(block, x_pos, y_pos, layers):
+    storage = []
+    for x in x_pos:
+        for y in y_pos:
+            for layer in np.unique(block[x,y,:]):
+                if int(layer) in layers:
+                    z = np.where(block[x, y, :] == layer)[0][0]
+                    if int(z) != 0:
+                        storage.append([x, y, z, "Layer" + str(int(layer))])
+                else:
+                    pass
+    return storage
+
+
+def pynoddy_assign_priors(prior_dict, ex, verbose=None):
+    """
+    Function to assign prior draws to respective pynoddy event properties automatically.
+
+    :param prior_dict: dictionary of prior parameters {parameter_key: {mean: m, stdev: sigma, pymc: pymc.Distr object}}
+    :param ex: pynoddy.experiment object
+    :return: Nothing, sets prior parameters in pynoddy.experiments
+    """
+    # TODO: Add Rotation event support
+    # iterate over every event in the experiment
+    for event in ex.events:
+        if verbose == 1:
+            print event, ex.events[event]
[…]

### 2017-01-16 Geological Inference based on Kinematic Structural Models/PosteriorAnalysis/__init__.py

@@ -1,0 +1,244 @@
+# Posteriors Class to handle Posterior Outputs
+import numpy as np
+import matplotlib.pyplot as plt
+
+from sklearn.neighbors import KernelDensity
+from sklearn.grid_search import GridSearchCV
+from scipy.stats.kde import gaussian_kde
+
+import json
+import pymc
+
+import sys
+
+sys.path.append("C:/Users/lebobcrash/Documents/GitHub/pynoddy/")
+
+import pynoddy.history
+import pynoddy.output
+import pynoddy.experiment
+
+
+class PosteriorAnalysis:
+    """Handles PYMC run posteriors for pynoddy graben simulations."""
+
+    def __init__(self, db_name, sim_type):
+        # self.parameters_filename = parameters
+        self.sim_type = sim_type  # pynoddy or gempy
+        self.db = pymc.database.hdf5.load(db_name)
+
+        # get trace names
+        self.trace_names = self.db.trace_names[0]
+
+        self.prior_trace_names = []
+        for entry in self.trace_names:
+            if "Metropolis" not in entry and "_model" not in entry and "deviance" not in entry and "like" not in entry:
+                self.prior_trace_names.append(entry)
+
+
+        if sim_type is "gempy":
+            self.blocks = self.db.gempy_model.gettrace()
+        elif sim_type is "pynoddy":
+            self.blocks = self.db.pynoddy_model.gettrace()
+        #elif sim_type is "both":
+        #    self.gempy_sections = self.db.gempy_model.gettrace()
+        #    self.pynoddy_sections = self.db.pynoddy_model.gettrace()
+
+        if len(np.shape(self.blocks)) is 3:
+            self.blocks = np.expand_dims(self.blocks, axis=2)
+
+        # load posteriors
+        # self.posteriors = {}
+        # for entry in self.parameters:
+        #    self.posteriors[entry] = np.load(entry+"_posterior.npy")
+
+        self._layers = np.unique(self.blocks[:])
+        self._n_layers = len(self._layers)
+        self._n_iter = len(self.blocks)
+        self.x_extent = len(self.blocks[0, :, 0, 0])
+        self.y_extent = len(self.blocks[0, 0, :, 0])
+        self.z_extent = len(self.blocks[0, 0, 0, :])
+        self._NaN = -9999
+
+    #def noddy_plot_average_trace_model(self):
+
+
+
+    def plot_traces(self):
[…]

### 2017-01-20 Risk analysis/Text/Layout/Layout/MSC_layout_orig.tex
Intento 3

%% THIS IS THE MAIN FILE %%
% Here the main arrangement of your thesis is determined. 
% In this file you read in all chapters (as separate .tex files)
% This is also the file you 'Build' with a LaTeX editor, like TeXmaker.
% The .cls file (mscThesis.cls) contains all details regarding copyrights and so on. Take a look and adjust where applicable.
% This MSc thesis standard layout is optimized for double sided printing on A4 format paper.
% Good luck and have fun with your Master Research in Applied Geophysics! Kind regards, Niels Grobbe
\documentclass[a4paper,11pt]{mscThesis}
%
%\usepackage{pifont}
\mscName{YourName}
\mscDate{\today}
\mscTitle{Test}
\mscSubTitle{Subtitle}
\mscKeyWords{thesis, msc, subject}
%
%\mscBackPicture{2660PF3}    % eps of 21 * 29.7 cm
\mscReaderOne{Name of first supervisor}
\mscReaderTwo{Name of second supervisor}
\mscReaderThree{Name of third supervisor or committee member}
%\mscReaderFour{}
%
\setThesisInfo
%
\begin{document}
%
%============================= Front matter ========================================
\frontmatter %
%
% Make a hell of a lot of title pages
    \maketitle
%
% Abstract
    \nonumchap{Abstract}

    This is the shortest abstract you have ever seen.
    \cleardoublepage
%
% Acknowledgements
    \nonumchap{Acknowledgements}%
    First of all I want to thank all the people who have participated in this project ..
    Remember, often more people have contributed to your final thesis than you initially would think of.
    \vspace*{15mm}

    \noindent
    Delft, University of Technology \hfill \mscname\\ % change to the university where you carried out your final MSc thesis research
    \mscdate

%
% table of contents, (\toc of \toclof of \tocloflot )
    \tocloflot
%
% Nomenclature
    \printnomencl %
%
% Acronyms
    \nonumchap{Acronyms} %
    %
    \begin{acronym}%
        \acro{DUT}{Delft University of Technology}%
    \end{acronym}%
    %
    \cleardoublepage%
%
%
%============================= Main matter =========================================
%
\mainmatter
%
% Introduction
\include{chapter1} % notice how the reference to the introduction takes place.. It refers to the name.tex
\cleardoublepage

%
% First Part
    \part{First Part} % you can divide your thesis into different parts, for example Theory&Modeling in part 1 and real data examples in part 2.
		\include{chapter2}

%
% Biblio
    \bibliographystyle{apalike}
    \printbib{MyBib} % in MyBib.bib you add all your reference information, following the correct format. Sometimes, the bib file needs to be built several times, as well as the main file, before all references occur correctly in your PDF.

%============================= Back matter =========================================
\appendix

    \chapter{The back of the thesis}

    \section{An appendix section}

    \subsection{An appendix subsection with C++ Lisitng}

    \lstset{language=C++}
    \lstinputlisting{test.c}

    \subsection{A \matlab $ $ Listing}

    \lstset{language=matlab}
    \lstinputlisting{test.m}

    \chapter{Yet another appendix}

    \section{Another test section}

    Ok, all is well.

% Index
    \printindex%
    \cleardoublepage%

\end{document}

### 2017-01-20 Risk analysis/Text/Layout/Layout/MSc_layout.tex
Intento 3

--- 2017-01-20 Risk analysis/Text/Layout/Layout/MSC_layout_orig.tex
@@
 % Here the main arrangement of your thesis is determined. 
-% In this file you read in all chapters (as separate .tex files)
+% In this file you read-in all chapters (as separate .tex files)
 % This is also the file you 'Build' with a LaTeX editor, like TeXmaker.
-% The .cls file (mscThesis.cls) contains all details regarding copyrights and so on. Take a look and adjust where applicable.
+% The .cls file (MScThesis.cls) contains all details regarding copyrights and so on.
+Take a look and adjust where applicable.
+But be careful changing this file: it determines the whole lay-out of your thesis!
 % This MSc thesis standard layout is optimized for double sided printing on A4 format paper.
 % Good luck and have fun with your Master Research in Applied Geophysics! Kind regards, Niels Grobbe
-\documentclass[a4paper,11pt]{mscThesis}
+
+\documentclass[a4paper,11pt]{MScThesis}
 %
 %\usepackage{pifont}
-\mscName{YourName}
+\mscName{Fabian Antonio Stamm}
 \mscDate{\today}
@@
 %\mscBackPicture{2660PF3}    % eps of 21 * 29.7 cm
-\mscReaderOne{Name of first supervisor}
-\mscReaderTwo{Name of second supervisor}
-\mscReaderThree{Name of third supervisor or committee member}
+\mscReaderOne{Prof. Florian Wellmann, Ph.D.}
+\mscReaderTwo{Prof. Dr. Janos Urai}
+\mscReaderThree{Miguel de la Varga, M.Sc.}
 %\mscReaderFour{}
@@
 
-    This is the shortest abstract you have ever seen.
+Please pay particular attention to the preparation of your abstract; use this text as a guide.
+Every master thesis report must be accompanied by an informative abstract of no more than one paragraph (max 300 words).
+The abstract should be self-contained.
+No references, figures, tables, or equations are allowed in an abstract.
+Do not use new terminology in an abstract unless it is defined or is well-known from the literature.
+The abstract must not simply list the topics covered in the paper but should (1) state the scope and principal objectives of the research, (2) describe the methods used, (3) summarize the results, and (4) state the principal conclusions.
+Do not refer to the master thesis report itself in the abstract.
+For example, do not say, "In this thesis we will discuss".
+Furthermore the abstract must stand alone as a very short version of the master thesis report rather than as a description of the contents.
+Remember that the abstract will be the first and most widely read portion of the master thesis report.
+Readers will be influenced by the abstract to the point that they decide to read the master thesis report or not.
+
     \cleardoublepage
@@
     First of all I want to thank all the people who have participated in this project ..
-    Remember, often more people have contributed to your final thesis than you initially would think of.
+    Remember, often more people have (in some way) contributed to your final thesis than you would initially think of....
     \vspace*{15mm}
@@
     \noindent
-    Delft, University of Technology \hfill \mscname\\ % change to the university where you carried out your final MSc thesis research
+    Delft University of Technology \hfill \mscname\\ % choose the university where you carried out your final MSc thesis research
+    Swiss Federal Institute of Technology \hfill \mscname\\ % choose the university where you carried out your final MSc thesis research
+    RWTH Aachen University \hfill \mscname\\ % choose the university where you carried out your final MSc thesis research
     \mscdate
@@
 %
-% table of contents, (\toc of \toclof of \tocloflot )
+% table of contents, (\toc or \toclof or \tocloflot )
     \tocloflot
@@
         \acro{DUT}{Delft University of Technology}%
+        \acro{ETH}{Swiss Federal Institute of Technology}%
+        \acro{RWTH}{Aachen University}%
     \end{acronym}%
@@
 \include{chapter1} % notice how the reference to the introduction takes place.. It refers to the name.tex
+
 \cleardoublepage
@@
 %
-% Biblio
+% Bibliography
     \bibliographystyle{apalike}
     \printbib{MyBib} % in MyBib.bib you add all your reference information, following the correct format.
+    % The *.bin file needs to be compiled using the Bib-compile option of your LaTeX text editor.
 Sometimes, the bib file needs to be built several times, as well as the main file, before all references occur correctly in your PDF.
@@
 % Index
-    \printindex%
-    \cleardoublepage%
+    \printindex
+    \cleardoublepage
 

### 2017-01-20 Risk analysis/Text/Layout/Layout/chapter1.tex
Intento 3

\chapter{Introduction} \label{chap::intro}

    Welcome to the standard layout for your IDEA LEAGUE MSc thesis written in \LaTeX. \LaTeX\  has a variety of advantages over conventional/ standard text editing programs, which you will soon enough discover yourself. \LaTeX\  almost forms a standard in the Scientific Community, especially due to its effective and straightforward mathematical capabilities.\\
    This is Chapter\ \ref{chap::intro}. If you want to know more about \LaTeX\ you better read
    \cite{texbook} or use the extensive help available on the internet. \index{LaTeX}. This 'hidden' index command helps you making an index at the end of your thesis. You can add this flag anywhere you want to make an index hit. You can see here also how to use acronyms, like \ac{DUT}. The acronyms are automatically listed in the corresponding section. Also, hyperlinks are created automatically with the developed class file, such that your digital PDF version of your thesis can be read dynamically.
 Have fun with \LaTeX\ and your M.Sc. research project and good luck! \\ \\

The purpose of the introduction is to tell readers why they should want to read what follows the introduction. This chapter should provide sufficient background information to allow readers to understand the context and significance of the problem. This does not mean, however, that authors should use the introduction to rederive established results or to indulge in other needless repetition. The introduction should (1) present the nature and scope of the problem; (2) review the pertinent literature, within reason; (3) state the objectives; (4) describe the method of investigation; and (5) describe the principal results of the investigation.

### 2017-01-20 Risk analysis/Text/Layout/Layout/chapter2.tex
Intento 3

    \chapter{First Real Chapter}

    This is a demonstration chapter. I will explain some of the possibilities of \LaTeX. Here something will be shown of control theory, 'the transfer function' \lsymb{$H(s)$}{Transfer function}. Subscripts and superscripts can be put in the nomenclature \index{nomenclature} list. \supers{max}{Maximum} \subs{min}{Minimum} Other things can also be added to the nomenclature list, like explanations of symbols being used throughout the thesis. \others{[kts]}{Knots} \others{$^{\circ}$, [deg]}{Degrees}

        \section{First section}

        This is the section. Referring to equations, figures and tables can easily be done by the commands \verb"\eqnref{}",
        \verb"\figref{}" and \verb"\tabref{}".
        \begin{equation}\label{eq:First}
        H(s) = \frac{1}{s+2}
        \end{equation}

        You see? Refer to equations like this \eqnref{eq:First}, i.e. the name of the label you have given the specific equation, figure or table.

        \subsection{The first subsection}

        Now I demonstrate, numbering equations, using subequations:
	  \begin{subequations}
		\begin{eqnarray}
    \label{2eq1d1}
	  \nabla\times\mathbf{L}  &=& \frac{\partial\mathbf{G}}{\partial t} \\
    \label{2eq1d2}
  	\nabla\times\mathbf{G}  &=& \frac{\partial\mathbf{L}}{\partial t} + \mathbf{J} \\
    \label{2eq1d3}
    \mathbf{G}              &=& \sigma\mathbf{J}
		\end{eqnarray}
			  \end{subequations}

				Or we can make matrices:
				\begin{equation}
				\mathbf{Q}_{12}=\left[\begin{array}{ccc}
		          0  &     1          &  0 \\
	            1  &     0          &  1 \\
	            0  &     1          &  0 \\
       	\end{array}\right]\quad
        \nonumber
		    \label{2eq1cf}
		    \end{equation}
		    This can also be done using the \verb"\align{}" command. Equation arrays are also possible:
     		\begin{eqnarray}
    		\label{2eq1e1}
	  \nabla\times\mathbf{L}  &=& \frac{\partial\mathbf{G}}{\partial t} \\
	    	\label{2eq1e2}
  	\nabla\times\mathbf{G}  &=& \frac{\partial\mathbf{L}}{\partial t} + \mathbf{J} \\
		    \label{2eq1e3}
    \mathbf{G}              &=& \sigma\mathbf{J}
		  \end{eqnarray}

        \subsubsection[Subsection Short Title]{The first sub-subsection with a very very very long title, but in the table of contents one can only see the short title in square brackets}

                Impressed by the capabilities? \index{Nicecapabilities}
                If you want to know more about the capabilities of \LaTeX, take a look at the "\textbf{The Not So Short Introduction to \LaTeXe}", which can be found on the internet.

    \paragraph{Next paragraph.}
    \begin{figure}[htbp]  % fig 6.3 page 186 Turbulence and Remous
    \centering
    \includegraphics[width=11cm]{vertstab.eps}
    \caption{Stability conditions for the vertical stability of saturated and unsaturated air.} %this is how it shows up in the List of Figures
    \label{f:verticalstab}
    \end{figure}
.\\    
And finally I end this example file with a table which will be centered in the middle of the following page
\begin{table}[hp]
\centering
\renewcommand{\arraystretch}{1.5}
\begin{tabular}{|ll|}
\hline
\multicolumn{2}{|c|}{\bf\sffamily {\it Data} files listed in a table }\\
\hline\hline
\multicolumn{2}{|l|}{\bf\sffamily First part} \\
Fabracadabra.m	& - Saturation computation\\
Fobracadabra.m	& - Pressure computation\\
Fibricadibri.m	& - Permeability computation\\
&\\
\multicolumn{2}{|l|}{\bf\sffamily Structural rock model} \\
struct.m	& - Rock structural data using symmetric boundary condition \\
bstruct.m	& - Rock structural data using anti-symmetric boundary condition \\
\hline
\end{tabular}
\caption{Good-looking program data deck files.} %this is how it shows up in the List of Tables
\label{tbl:tbl}
\end{table}

### 2017-03-22 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Started writing methods part in texStudio

    \chapter{Methods}

    The methods utilized in this work are presented in the following chapter. This includes introductions into Bayesian analysis and decision theory. In particular Bayesian inference, estimation of uncertain values and how loss function can be used in this context. These methods are incorporated in the context of numerical modeling of structural geological settings. This is done by programming in a python environment. Central tools for model construction and conduction of the statistical evaluation are Geomodeler3D, GemPy and PyMC in particular and are also presented in this chapter.

        \section{Bayesian analysis and decision theory}
        Introduction into section: What is decision theory, what is Bayesian analysis? Why use it here?
        What is a decision here and in relation to geoscientific cases and models?

        \subsection{Bayesian inference}
        Bayesian inference is most importantly characterized by its preservation of uncertainty, in contrast to standard statistical inference \cite{davidson2015}. Probability is seen as a measure of belief for an event to occur. It has been argued by \cite{davidson2015}  that this Bayesian approach is intuitive and inherent in the natural human perspective. These beliefs can furthermore be assigned to individuals \cite{davidson2015}. Thus, different and even contradicting beliefs about the probability of an event might be held by different individuals, based on variations and disparities in the information available to each one individual \cite{davidson2015}.
        The initial belief or guess about an event A can be denoted as P(A). This is used as the so-called prior probability on which Bayesian updating is based. The beliefs about the occurrence of an event are revalued in the presence of additional information, i.e. the observation of new evidence which can be denominated as X. These observations are included as likelihoods P(X$|$A). This process of updating results in a posterior probability P(A$|$X) \cite{davidson2015}. It is important to note that the prior is not simply discarded but re-weighted by Bayesian updating. It was also pointed out by \cite{davidson2015} that by utilizing an uncertain prior, the potential for wrongfulness of the initial guess is already included. This means that Bayesian updating is about reducing uncertainty in a belief and reaching a guess that is less wrong \cite{davidson2015}.
        Bayesian updating is defined by and conducted via the following equation, called the Bayes' Theorem:

        \begin{equation}\label{eq:BayesTheorem}
        P(A|X) = \frac{P(X|A)P(A)}{P(X)}
        \propto P(X|A)P(A)
        \end{equation}

        -Law of Large Numbers!

        \subsection{Estimation}
        The resulting posterior distribution can be used to acquire point estimates for the true state of nature $\theta$. Common and simple examples for such estimators are the mode (i.e. the generalized maximum likelihood estimate), the mean and the median of a distribution \cite{berger2013stat}. The presentation of a point estimate should usually come with a measure for its estimation error. According to \cite{berger2013stat}, the posterior variance is most commonly used as an indication for estimate accuracy. However, it is argued by \cite{davidson2015} that by using pure accuracy metrics, while this technique is objective, it ignores the original intention of conducting the statistical inference in cases, in which payoffs of decisions are valued more than their accuracies. A more appropriate approach can be seen in the introduction of loss and the use of loss functions \cite{davidson2015}.

        \subsection{Expected loss and loss functions} 
        Loss is a statistical measure of how bad an estimate is, i.e. how much is lost by making a certain decision. Gains are considered by statisticians as negative losses \cite{davidson2015}.
        The magnitude of an estimate's loss is defined by a loss function, which is a function of the estimate of the parameter and the true value of the parameter \cite{davidson2015}:

        \begin{equation}\label{eq:LossFunction}
        L(\theta,\hat{\theta}) = f(\theta,\hat{\theta})
        \end{equation}

        So, "how bad" a current estimate is, depends on the way a loss function weights accuracy errors and returns respective losses. Two standard loss functions are the absolute-error and the squared-error loss function. Both are simple to understand and commonly used \cite{davidson2015}.

        As implied by its name, the absolute-error loss function returns loss as the absolute error, i.e. difference between the estimate and the true parameter \cite{davidson2015}:

        \begin{equation}\label{eq:LossFunction}
        L(\theta,\hat{\theta}) = |\theta - \hat{\theta}|
        \end{equation}

        Accordingly, losses increasing linearly with the distance to the true value are returned for respective estimates.

        Using the squared-error loss function returns losses that increase quadratically with distance of the estimator to the true parameter value \cite{davidson2015}:

        \begin{equation}\label{eq:LossFunction}
        L(\theta,\hat{\theta}) = |\theta - \hat{\theta}|^2
        \end{equation}

        This exponential growth of loss also means that large errors are weighted much stronger than small errors. This might come with overvaluation of distant outliers and misrepresentation of magnitudes in distance. Regarding this, the absolute-error loss function can be seen as more robust \cite{davidson2015}.

        - each give mean or median --> estimator with accuracy metric? objective and symmetric! aim at high precision.

        -shift in prespective. look at outcome instead of estimation precision!

        \cite{davidson2015} and (LF DESIGN PAPER?) propose that it might be useful to move away from these type of objective loss functions to the design of new loss functions that are specifically customized to an reflect an individual's, i.e. the decision maker's objectives, preferences and outcomes.

        The standard loss function defined above are symmetric, but can easily be adapted to be asymmetric, for example by weighting on the negative side stronger than those on the positive side. Preference over estimates larger than the true value is thus incorporated in an uncomplicated way \cite{davidson2015}. Much more complicated designs of loss functions are possible, depending on purpose, objective and application \cite{davidson2015}. Some case-specific loss functions are designed in the following chapters of this work.

        - Cite from paper about the design of loss functions! Subjectivity of customized loss functions!

        The presence of uncertainty during decision making implies that the truly incurred loss $L(\theta,a)$ cannot be known at the time of making the decision \cite{berger2013stat}. A suitable alternative is to consider a decision's expected loss and to decide on a decision that is optimal in relation to this expected loss \cite{berger2013stat}.

        - Averaging of EL over prob. distribution

        - Bayesian point estimate

        - Law of Large Numbers

        - MAP?

        - Implementation of risk as done by \cite{davidson2015}

        \section{Application in structural geological modeling}

        \section{Example section}

        This is the section. Referring to equations, figures and tables can easily be done by the commands \verb"\eqnref{}",
        \verb"\figref{}" and \verb"\tabref{}".
        \begin{equation}\label{eq:First}
        H(s) = \frac{1}{s+2}
        \end{equation}

        You see? Refer to equations like this \eqnref{eq:First}, i.e. the name of the label you have given the specific equation, figure or table.

        \subsection{The first subsection}

        Now I demonstrate, numbering equations, using subequations:
	  \begin{subequations}
		\begin{eqnarray}
    \label{2eq1d1}
	  \nabla\times\mathbf{L}  &=& \frac{\partial\mathbf{G}}{\partial t} \\
    \label{2eq1d2}
  	\nabla\times\mathbf{G}  &=& \frac{\partial\mathbf{L}}{\partial t} + \mathbf{J} \\
    \label{2eq1d3}
    \mathbf{G}              &=& \sigma\mathbf{J}
		\end{eqnarray}
			  \end{subequations}

				Or we can make matrices:
				\begin{equation}
				\mathbf{Q}_{12}=\left[\begin{array}{ccc}
		          0  &     1          &  0 \\
	            1  &     0          &  1 \\
	            0  &     1          &  0 \\
       	\end{array}\right]\quad
        \nonumber
		    \label{2eq1cf}
		    \end{equation}
		    This can also be done using the \verb"\align{}" command. Equation arrays are also possible:
     		\begin{eqnarray}
    		\label{2eq1e1}
	  \nabla\times\mathbf{L}  &=& \frac{\partial\mathbf{G}}{\partial t} \\
	    	\label{2eq1e2}
  	\nabla\times\mathbf{G}  &=& \frac{\partial\mathbf{L}}{\partial t} + \mathbf{J} \\
		    \label{2eq1e3}
    \mathbf{G}              &=& \sigma\mathbf{J}
		  \end{eqnarray}

        \subsubsection[Subsection Short Title]{The first sub-subsection with a very very very long title, but in the table of contents one can only see the short title in square brackets}

                Impressed by the capabilities? \index{Nicecapabilities}
                If you want to know more about the capabilities of \LaTeX, take a look at the "\textbf{The Not So Short Introduction to \LaTeXe}", which can be found on the internet.

    \paragraph{Next paragraph.}
    \begin{figure}[htbp]  % fig 6.3 page 186 Turbulence and Remous
    \centering
    \includegraphics[width=11cm]{vertstab.eps}
    \caption{Stability conditions for the vertical stability of saturated and unsaturated air.} %this is how it shows up in the List of Figures
    \label{f:verticalstab}
    \end{figure}
.\\    
And finally I end this example file with a table which will be centered in the middle of the following page
\begin{table}[hp]
\centering
\renewcommand{\arraystretch}{1.5}
\begin{tabular}{|ll|}
\hline
\multicolumn{2}{|c|}{\bf\sffamily {\it Data} files listed in a table }\\
\hline\hline
\multicolumn{2}{|l|}{\bf\sffamily First part} \\
Fabracadabra.m	& - Saturation computation\\
Fobracadabra.m	& - Pressure computation\\
Fibricadibri.m	& - Permeability computation\\
&\\
\multicolumn{2}{|l|}{\bf\sffamily Structural rock model} \\
struct.m	& - Rock structural data using symmetric boundary condition \\
bstruct.m	& - Rock structural data using anti-symmetric boundary condition \\
\hline
\end{tabular}
\caption{Good-looking program data deck files.} %this is how it shows up in the List of Tables
\label{tbl:tbl}
\end{table}

### 2017-03-28 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Text update (methods) and export of 3D model

@@
         \section{Bayesian analysis and decision theory}
-        Introduction into section: What is decision theory, what is Bayesian analysis? Why use it here?
-        What is a decision here and in relation to geoscientific cases and models?
+	    Problems and reasoning behind decision making are examined in the field of decision theory, as implied by the name \cite{berger2013stat}.
+Such decision problems are commonly influenced by parameters that are uncertain.
+In statistical decision theory, the presence of statistical knowledge is used to gain information on the nature of these uncertainties.
+(Such uncertain parameters can be considered as numerical quantities.) In order to find the best decision to a problem, it is possible to combine sample information with other aspects such as the possible consequences of decision making and the availability of prior information on our uncertainties.
+Decision consequences are expressed as gains in economic decision theory and as losses, which equal negative gains, in statistics.
+Prior information might be given for example due to experience from previous similar problems or from expert knowledge (see Batvold and Begg).
+The approach of utilizing such as priors is known as Bayesian analysis and is explained in the following \cite{berger2013stat}.
+(It goes well with decision theory.)
+
+	    \section{Basic elements}
+	    Some basic elements are to be defined.
+The unknown (uncertain?) quantity influencing decision making is usually named denominated as $\theta$, the state of nature.
+Given statistical information on $\theta$ in the form of probability distributions, $\theta$ is called the parameter.
+	    Decisions are also referred to as actions $a$.
+	    The outcome of statistical tests in form of information or statistical evidence is denoted as $X$.	    
+	    Loss is defined as $L(\theta,a)$, so $L(\theta_1,a-1)$ is the actual loss incurred when action $a_1$ is taken while the true state of nature is $\theta_1$ \cite{berger2013stat}.
+Loss, expected loss and loss functions are further explained below.
 
@@
 
-        The presence of uncertainty during decision making implies that the truly incurred loss $L(\theta,a)$ cannot be known at the time of making the decision \cite{berger2013stat}.
-A suitable alternative is to consider a decision's expected loss and to decide on a decision that is optimal in relation to this expected loss \cite{berger2013stat}.
+        The presence of uncertainty during decision making implies that the true parameter is unknown and thus the truly incurred loss $L(\theta,a)$ cannot be known at the time of making the decision \cite{berger2013stat, davidson2015}.
+The Bayesian perspective considers unknown parameters as random variables and samples that are drawn from the posterior distribution as possible realizations of the unknown parameter, i.e.
+all possible true values are represented by this distribution \cite{davidson2015}.
+A suitable alternative to the actual loss is to consider a decision's expected loss and to make a decision that is optimal in relation to this expected loss \cite{berger2013stat}.
 
-        - Averaging of EL over prob. distribution
+        Given a posterior distribution $P(\theta|X)$, the expected loss of choosing an estimate $\hat{\theta}$ over the true parameter $\theta$ (after evidence X has been observed) is defined by the function below \cite{davidson2015}:
 
-        - Bayesian point estimate
+        \begin{equation}\label{eq:ExpectedLoss}
+        l(\hat{\theta}) = E_{\theta}[L(\theta,\hat{\theta})]
+        \end{equation}
 
-        - Law of Large Numbers
+        The expectation symbol $E$ is subscripted with $\theta$, by which it is indicated that $\theta$ is the respective unknown variable.
+This expected loss as defined above, is also referred to the risk of estimate $\hat{\theta}$ \cite{davidson2015}.
+
+        By the Law of Large Numbers, the expected loss of $\hat{\theta}$ can be approximated drawing $N$ samples from the posterior distribution, respectively applying a loss function $L$ and averaging of the number of samples \cite{davidson2015}:
+
+        \begin{equation}\label{eq:ExpectedLoss2}
+        \frac{1}{N}\sum_{i=1}^{N} L(\theta_i,\hat{\theta}) \approx E_{\theta}[L(\theta,\hat{\theta})] = l(\hat{\theta})
+        \end{equation}
+
+        Minimization of \ref{eq:ExpectedLoss} or \ref{eq:ExpectedLoss2} returns a Bayesian point estimate known as Bayes action $\delta^P(X)$ (???), which is the estimate, action or decision with the least expected loss according to the loss function \cite{berger2013stat}.
+For a unimodal and symmetric absolute-error loss function, the Bayes action is simply the median of the posterior distribution, while squared-error loss it is the mean \cite{davidson2015, berger2013stat}.
+The MAP estimate is returned when using zero-one loss \cite{davidson2015}  (WRITE THIS?).
+The possibility of more than on minimum also implies that several Bayes actions can exist for one problem \cite{berger2013stat}.
 
@@
 
-        - Implementation of risk as done by \cite{davidson2015}
+        \cite{davidson2015} implemented different risk affinities by simply  introducing a risk parameter into the loss function.
+By using different values for this parameter, it can be represented how comfortable an individual is with being wrong and furthermore which "side of wrong" is preferred by this decision maker \cite{davidson2015}.
+In the chapter below, different risk parameters are introduced to alter the weighting in the loss functions (???).
+
+        \subsection{Value of information}
+        - considering the change in gain (or loss) after regarding new information or evidence, the value of this information can be calculated
+        - in the presence of uncertainty we look at expected value of information
+        - this can be used as a measure, to assess if the effort to attain new evidence or information is worth it
+        - also as a measure to see how much additional information is worth to different actors with different risk affinities
+        - here we use the comparison between Bayes actions before and after Bayesian updating
 
         \section{Application in structural geological modeling}
+
+        - applying these statistical theories in the context of geological modeling 
+        - as done by De la Varga (2015)
+
+        - equivalence of basic elements in geologic models, some examples
+
+        - sampling methods
+
+        - numerical implementation in python using pyMC
+
+        - construction of 1D model in python and pyMC
+
+        - construction of 3D model using GeoModeller3D and GeMPy additionally
+
+        - what about theory of petroleum reservoir and volumetric calculation? explain this in later chapter of 3D model???
 

### 2017-04-04 Risk analysis/Text/Layout/Layout/MSc_layout_Fabian.tex
Progressed with methods text (PyMC3 text). Still heavy work needed in: Model construction, sampling (MCMC), structural geologic modelling, etc. Also added text from 1D case notebook - requires updating!

--- 2017-01-20 Risk analysis/Text/Layout/Layout/MSc_layout.tex
@@
     \noindent 
-    Delft University of Technology \hfill \mscname\\ % choose the university where you carried out your final MSc thesis research
-    Swiss Federal Institute of Technology \hfill \mscname\\ % choose the university where you carried out your final MSc thesis research
     RWTH Aachen University \hfill \mscname\\ % choose the university where you carried out your final MSc thesis research
@@
     \begin{acronym}%
-        \acro{DUT}{Delft University of Technology}%
-        \acro{ETH}{Swiss Federal Institute of Technology}%
         \acro{RWTH}{Aachen University}%
@@
     \part{First Part} % you can divide your thesis into different parts, for example Theory&Modeling in part 1 and real data examples in part 2.
-		\include{chapter2}
+		\include{chapter2_methods}
+		\include{chapter3_1Dcase}
 
@@
     \bibliographystyle{apalike}
-    \printbib{MyBib} % in MyBib.bib you add all your reference information, following the correct format.
+    \printbib{complete_biblio.bib} % in MyBib.bib you add all your reference information, following the correct format.
     % The *.bin file needs to be compiled using the Bib-compile option of your LaTeX text editor.

### 2017-04-04 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Progressed with methods text (PyMC3 text). Still heavy work needed in: Model construction, sampling (MCMC), structural geologic modelling, etc. Also added text from 1D case notebook - requires updating!

@@
 This is done by programming in a python environment.
-Central tools for model construction and conduction of the statistical evaluation are Geomodeler3D, GemPy and PyMC in particular and are also presented in this chapter.
+Central tools for model construction and conduction of the statistical evaluation are Geomodeler3D, GemPy and PyMC3 in particular and are also presented in this chapter.
 
@@
 Such decision problems are commonly influenced by parameters that are uncertain.
-In statistical decision theory, the presence of statistical knowledge is used to gain information on the nature of these uncertainties.
+In statistical decision theory, available statistical knowledge is used to gain information on the nature of these uncertainties.
 (Such uncertain parameters can be considered as numerical quantities.) In order to find the best decision to a problem, it is possible to combine sample information with other aspects such as the possible consequences of decision making and the availability of prior information on our uncertainties.
@@
 Prior information might be given for example due to experience from previous similar problems or from expert knowledge (see Batvold and Begg).
-The approach of utilizing such as priors is known as Bayesian analysis and is explained in the following \cite{berger2013stat}.
+The approach of utilizing priors is known as Bayesian analysis and is explained in the following \cite{berger2013stat}.
 (It goes well with decision theory.)
@@
 	    Some basic elements are to be defined.
-The unknown (uncertain?) quantity influencing decision making is usually named denominated as $\theta$, the state of nature.
+The unknown (uncertain) quantity influencing decision making is usually denominated as $\theta$, the state of nature \cite{berger2013stat}.
 Given statistical information on $\theta$ in the form of probability distributions, $\theta$ is called the parameter.
@@
 It has been argued by \cite{davidson2015}  that this Bayesian approach is intuitive and inherent in the natural human perspective.
-These beliefs can furthermore be assigned to individuals \cite{davidson2015}.
+These beliefs can be assigned to individuals \cite{davidson2015}.
 Thus, different and even contradicting beliefs about the probability of an event might be held by different individuals, based on variations and disparities in the information available to each one individual \cite{davidson2015}.
-        The initial belief or guess about an event A can be denoted as P(A).
+
+        The initial belief or guess about an event A can be denoted as P(A) \cite{davidson2015}.
 This is used as the so-called prior probability on which Bayesian updating is based.
 The beliefs about the occurrence of an event are revalued in the presence of additional information, i.e.
-the observation of new evidence which can be denominated as X.
+the observation of new evidence X.
 These observations are included as likelihoods P(X$|$A).
@@
         -Law of Large Numbers!
+
+        -MCMC sampling methods!
 
@@
 
-        \section{Example section}
+		\subsection{Implementing Bayesian analysis numerically with Python and PyMC3}
+		Bayesian analysis can be conducted using probabilistic programming \cite{salvatier2016pymc3}.
+For doings this, the programming language of choice in this work is Python.
+The merits of Python have been pointed out by \cite{behnel2010, salvatier2016pymc3, Langtangen2008}.
+Development is facilitated by an expressive but concise and clean syntax that is easy to learn.
+Python is dynamic, compatible with multiple platforms and offers good support for numerical computing.
+Integration of other scientific libraries and extension via C, C++, Fortran or Cython is easily possible \cite{behnel2010, salvatier2016pymc3, Langtangen2008}.
+Python is thus a straightforward tool for the implementation of central components of Bayesian analysis, such as custom statistical distributions and samplers \cite{salvatier2016pymc3}.
 
-        This is the section. Referring to equations, figures and tables can easily be done by the commands \verb"\eqnref{}",
-        \verb"\figref{}" and \verb"\tabref{}".
-        \begin{equation}\label{eq:First}
-        H(s) = \frac{1}{s+2}
-        \end{equation}
+		In probabilistic programming, programming variables are used as components to build probabilistic models \cite{davidson2015}.
 
-        You see? Refer to equations like this \eqnref{eq:First}, i.e. the name of the label you have given the specific equation, figure or table.
-
-        \subsection{The first subsection}
-
-        Now I demonstrate, numbering equations, using subequations:
-	  \begin{subequations}
-		\begin{eqnarray}
-    \label{2eq1d1}
-	  \nabla\times\mathbf{L}  &=& \frac{\partial\mathbf{G}}{\partial t} \\
-    \label{2eq1d2}
-  	\nabla\times\mathbf{G}  &=& \frac{\partial\mathbf{L}}{\partial t} + \mathbf{J} \\
-    \label{2eq1d3}
-    \mathbf{G}              &=& \sigma\mathbf{J}
-		\end{eqnarray}
-			  \end{subequations}
-
-				Or we can make matrices:
-				\begin{equation}
-				\mathbf{Q}_{12}=\left[\begin{array}{ccc}
-		          0  &     1          &  0 \\
-	            1  &     0          &  1 \\
-	            0  &     1          &  0 \\
-       	\end{array}\right]\quad
-        \nonumber
-		    \label{2eq1cf}
-		    \end{equation}
-		    This can also be done using the \verb"\align{}" command. Equation arrays are also possible:
-     		\begin{eqnarray}
-    		\label{2eq1e1}
-	  \nabla\times\mathbf{L}  &=& \frac{\partial\mathbf{G}}{\partial t} \\
-	    	\label{2eq1e2}
-  	\nabla\times\mathbf{G}  &=& \frac{\partial\mathbf{L}}{\partial t} + \mathbf{J} \\
-		    \label{2eq1e3}
-    \mathbf{G}              &=& \sigma\mathbf{J}
-		  \end{eqnarray}
-
-        \subsubsection[Subsection Short Title]{The first sub-subsection with a very very very long title, but in the table of contents one can only see the short title in square brackets}
-
-                Impressed by the capabilities? \index{Nicecapabilities}
-                If you want to know more about the capabilities of \LaTeX, take a look at the "\textbf{The Not So Short Introduction to \LaTeXe}", which can be found on the internet.
-
-    \paragraph{Next paragraph.}
-    \begin{figure}[htbp]  % fig 6.3 page 186 Turbulence and Remous
-    \centering
-    \includegraphics[width=11cm]{vertstab.eps}
-    \caption{Stability conditions for the vertical stability of saturated and unsaturated air.} %this is how it shows up in the List of Figures
-    \label{f:verticalstab}
-    \end{figure}
-.\\    
-And finally I end this example file with a table which will be centered in the middle of the following page
-\begin{table}[hp]
-\centering
-\renewcommand{\arraystretch}{1.5}
-\begin{tabular}{|ll|}
-\hline
-\multicolumn{2}{|c|}{\bf\sffamily {\it Data} files listed in a table }\\
-\hline\hline
-\multicolumn{2}{|l|}{\bf\sffamily First part} \\
-Fabracadabra.m	& - Saturation computation\\
-Fobracadabra.m	& - Pressure computation\\
-Fibricadibri.m	& - Permeability computation\\
-&\\
-\multicolumn{2}{|l|}{\bf\sffamily Structural rock model} \\
-struct.m	& - Rock structural data using symmetric boundary condition \\
-bstruct.m	& - Rock structural data using anti-symmetric boundary condition \\
-\hline
-\end{tabular}
-\caption{Good-looking program data deck files.} %this is how it shows up in the List of Tables
-\label{tbl:tbl}
-\end{table}
+		The numerical implementation of Bayesian analysis is further aided by the use of the Python library PyMC3, which was developed for conducting Bayesian inference and prediction problems in an open-source probabilistic programming framework \cite{davidson2015, salvatier2016pymc3}.
+Different model fitting techniques are provided in PyMC3, such as the \textit{maximum a posteriori} (MAP) method and several \textit{Monte Carlo Markov Chain} (MCMC) sampling methods \cite{salvatier2016pymc3}.
+		Name examples for state-of-the-art samplers included.
+		\cite{salvatier2016pymc3} point out that the development of PyMC3 continuing, as the inclusion of further tools is planned for future updates.

### 2017-04-04 Risk analysis/Text/Layout/Layout/chapter3_1Dcase.tex
Progressed with methods text (PyMC3 text). Still heavy work needed in: Model construction, sampling (MCMC), structural geologic modelling, etc. Also added text from 1D case notebook - requires updating!

\chapter{One-dimensional reservoir case}

An abstract one-dimensional case of a petroleum reservoir is presented in this chapter. The underlying model and basic approach are inherited from De la Varga (2016). Parameters were adapted to better represent a reasonable geological petroleum system consisting of a reservoir with overlying seal in the subsurface. This way, a certain economic significance is ascribed to the model and further relevant questions can be derived. Regarding the petroleum sector, the problem of interest is most commonly one of estimating the monetary value contained in an oil or gas reservoir. Limiting the model to only one dimension and a small number of uncertain parameters allows for a relatively straightforward and simplified approach to assessing an abstract type of value for a reservoir and designing a respective loss function for value estimation. A step by step derivation of such as case follows below.

	\section{Constructing the one-dimensional model}

	De la Varga (2016) constructed a simple geological model using three uncertain locations in one-dimensional space, which mark the boundaries of layers. The probability of positions of these points are defined by sampling from normal distributions. Standard deviations of these distributions increase with depth, representing an increase in uncertainty. For an approximate representation of a petroleum system, the distribution means were set to depths of 2000 (seal top), 2050 (reservoir top) and 2200 (reservoir bottom). These points confine two layers in the middle, from which the upper one can be labeled as seal and the lower one as reservoir. The resulting model with its possible layer boundary locations is illustrated in Figure 1.

	\section{Assessing reservoir quality using scores}

	The next step is to find a way to assess the quality of the reservoir from a petroleum industry perspective in such a simplified model. What can be deduced from the distributions of layer boundary locations, is the thickness of the seal and the reservoir, as well as the depth of both, again as probability distributions. To assess the reservoir quality in an abstract way, it can be assigned with a score. This score is made dependant on the three uncertain parameters (1) reservoir thickness, (2) reservoir top depth and (3) seal thickness (see Figure 2).

	\subsection{Two parameters scoring}

	Assuming that reservoir thickness is a simplified indicator for the extractable oil or gas and thus value in place, a gain in score can be correlated with increase in thickness. Here, two score points are assigned to one meter of thickness. Increasing costs of drilling are indicated by increasing depth of the reservoir top. Consequently, one negative score point is ascribed to every meter in depth. Samples from the probability distributions of these two parameters are drawn to model the true score of the reservoir (depth scores are subtracted from reservoir thickness scores). For the data used here, it can be seen in Figure 3 that the results of this two parameter score modelling are represented by an approximately normal distribution. The score is negative in about 17\% of the cases. Mean and median are about the same.

	\subsection{Three parameters scoring}

	Influence by the seal thickness is added in the following. Score points are not added or subtracted by this parameter directly. Instead, a threshold for seal reliability is defined beforehand. Here it is set at 20 m thickness. If the seal thickness falls below this threshold, it is assumed that the seal fails completely and thus all the potential value of the reservoir is lost, while costs of depth remain. Thus, a condition to check whether the seal is reliable is now included in the model. Results are visualized in Figure 4. The main distribution was not changed significantly, except for a striking peak of probability that emerged for the possibility of a score of -2000. Furthermore, mean and median have been shifted to lower values and are now found further apart.

	\section{Designing a case-specific loss function}

	Now that a distribution of reservoir score probabilities has been modelled, a loss function for estimation of the true score value can be developed.

	\subsection{Testing standard loss functions}

	Some standard loss function were presented in chapter X (theory of loss fucntions to be explained in foregone chapter). For this case, the absolute-error loss and the squared-error loss function were considered as starting points, to design a more case-specific loss function. Their general forms for certain determined values are illustrated in Figure 5.
	As explained before in chapter X, the expected loss for an estimate can be approximated by calculating the arithmetic mean of the losses compared to all samples in the modelled score distribution.

	1N∑i=1NL(θi,θ̂ )≈Eθ[L(θ,θ̂ )]=l(θ̂ )

	Calculating the expected loss for estimates ranging from -3000 to 6000 using these two standard loss function results in the graphs depicted in Figure 6. As expected (find proper source!), the median of the distribution coincides with the minimum of the absolute-error and the median with the minimum of the squared-error expected loss (Bayes action/estimator).

	\subsection{Adaptions to design a customized loss function}

	These standard loss functions provide objectively good estimators minimizing expected loss. Due to their symmetric properties, both will always give the median or mean of the underlying distribution as minimizing estimator respectively. However, assigning an economic notion to our model and assuming the case of an actor or decision maker in any field, naturally necessitates the consideration of preferences, interests and the overall subjective perspective such an individual or company might have. Further constraints and influences can also me specific to the field, industry or generally to the problem at hand. Consequently, the design of a more specific non-standard and possibly asymmetric loss function might be required, so that an adapted Bayesian estimator can be found. One that includes subjective aspects and difference in weighting of particular gains or losses, arising from an actors preferences and the environment in which the actor has to estimate or make a decision. In the face of several uncertain parameters, a perfectly true estimate is virtually unattainable. However, the attempt can be made, to design a customized loss function that returns a Bayesian estimator involving the least bad consequences for an actor in a specific environment. Regarding the petroleum system case modelled above, such an attempt is made and explained step by step in the following.

	Text NOT up-to-date!!!

	For the purpose of estimation, it makes sense that a standard loss function is chosen as a basis. The absolute-error loss seems most approriate for this case of petroleum reservoir value estimation. As stated above, the reservoir score we defined is an abstract and simplified way to reflect a value contained in the reservoir. An actor would wish to extract this, should it be positive (and feasible, but this would require the inclusion of numerous more factors). Ideally, an actor would like to know the exact true score, so that investments or resources can appropriately be allocated. This allocation is the decision to be made or action to be taken. Deviations from the unknown true score in the form of over- and underestimation bring about an error and loss accordingly. In principle, there is no reason for loss to increase exponentially with distance from the true value. Allocation of investments would not increase exponentially with linear increase of the value of the resource. For this reason, the absolute-error loss function is favored over the squared-error loss function in this case. Some adaptions are made below, based on mostly logical case-specific assumptions.

	    1st Adaption: For any negative estimate, an actor will take the same action as for a zero score, i.e. not to invest at all, since there is no positive value present. If the true score is zero or negative, this means that over- and underestimating, while the estimate is also negative or zero, is irrelevant an results in:

	L(θ,θ̂ )=0

	In other words: There is nothing to win and nothing is lost, since no action is taken.

	    2nd Adaption: Considering the development of a petroleum reservoir, one might also assume that overinvesting is worse than underinvesting. Overestimating the size of a reservoir might for example lead to the installation of equipment or facilites that are actually not needed. This comes with additional unrecoverable expenditures. Consequences from underestimating, however, may presumably be easier to resolve. Additional equipment can often be installed later afterwards. Hence, overestimation is weighted stronger in our loss function by multiplying the error with an overstimation factor a = 1.25:

	L(θ,θ̂ )=|(θ−θ̂ )|∗a

	    3d Adaption: The worst case for any project would be, to start and progress its development, only to discover later that the value in the reservoir does not cover the costs of realizing the project. A petroleum system might also turn out to be a complete failure, although the actor's estimate indicated the opposite. This is refered to here as a worst case or fatal overestimation, when a positive score is estimated, but the true score is zero or negative. It is included in the loss function with a second weighting factor b = 2:

	L(θ,θ̂ )=|(θ−θ̂ )|∗b

	In other words: Worst case or fatal overestimation is twice as bad as simply underestimating.

	    4th Adaption: A worst case or fatal underestimation can also be derived from the idea of estimating a zero or negative score, when the true score is actually positive. This basically reflects the opportunity costs of completely discarding a reservoir with a true score of θ:

	L(θ,θ̂ )=θ

	A step by step realization of these four adaptions is depicted in the plots in Figure 7. The first adaption is visualized in plot IIa, reflecting the irrelevance of negative estimates. Assignment of a stronger weight to overestimation, has led to a shift of the Bayesian action to a lower value in IIb. This shift has been in reinforced in plott IIc, due to adopting the concept of fatal overestimation. Finally, the inclusion of worst case underestimation raises the baseline for expected loss of not developing the reservoir (estimating zero or negative score).

	The implementation of this customized loss function using single deterministed values for the true score is plotted in Figure 8.

	It has to be emphasized that this is just one possible proposal for customization. There exists not one perfect loss function design for such a case. Other designs might differ slightly, for example by choosing different values for the same factors, to fundamentally by basing it on a significantly different mathematical structure. Loss functions are customized according to the subjective needs and objectives of an actor. Thus, they are defined by the actor expressing his perspective and might change, should the actor change his mind. Especially considering individual persons as actors, even psychological aspects may play a significant role.

	(Estimate or prediction? True value only known if a "yes"-action (development) is taken.)

	\subsection{Including different risk-affinities in the loss function}

	TEXT NOT UP TO DATE!!!

	One can assume that several actors in one sector or decision environment may have the same general loss functions, but different affinities concerning risks. This might be based on psychological factors or economic philosophies followed by a company. It might also be based on the budget and options such an actor has available. An intuitive example is the comparison of a small and a large company. A certain false estimate or error might have a significantly stronger impact on a company which has a a generally lower market share and only few projects, than on a larger company which might possess a higher financial flexibility and for which one project is only one of many development options.

	In the following, the loss function is further adapted to consider different risk-affinities of different actors. Representing risk behavior in a loss function can also be done in different ways. Here, it is simply included as a the factor r that is multiplied with the overestimation weighting factor a and b:

	L(θ,θ̂ )=|(θ−θ̂ )|∗(a∗r)

	L(θ,θ̂ )=|(θ−θ̂ )|∗(b∗r)

	Thus, the risk-neutral loss function is returned for r = 1. For r < 1, the weight of overestimating is reduced. This represents a more risk-friendly actor that is willing to bid on a higher estimate to attain a higher return in value. For r > 1, the overestimation weight is increased and respectively more risk-averse actors are prompted to bid on lower estimates. They prefer a lower but safer return value, due to the higher loss of wrong overestimation. The implementation of a risk-affinity factor r leads to different steepnesses of the plotted curves of the loss functions on the positive side of estimates (see Figures 9 and 10). The same three given deterministic values are used as true scores in the three plots in Figure 9. For certain determined values, the steepnesses of the functions vary, but their loss minima are found at the same location. However, in Figure 10 it can be seen that the minima for expected loss, i.e. the Bayesian estimators for differently risk-affine actors, are located at different estimates. Mean and median are clearly surpassed by the best estimate of the most risk-fiendly actor, while for the most risk-averse actor the Bayes action equals a zero estimate (and thus the decision to take on action).

	\section{Updating the model with additional information in the form of thickness likelihoods}

	De la Varga (2016) made use of Bayesian inference to reduce the uncertainty in this type of one-dimensional model. The same is done here. The probability distributions for the location of the layer boundaries are treated as priors. Likelihood functions for the thicknesses of the layers are introduced as new information in the model. For the seal, a normal distribution with a mean of 25 m and a standard devation of 20 m is chosen to reflect the likelihood. For the reservoir, a normal distribution with a mean of 180 m and a standard deviation of 60 m is chosen. Implementing these likelihoods in the model, the uncertainty in the posterior probability distributions for the layer boundary depths is reduced (see Figure 11).

	Modelling of a reservoir score is conducted as above, now based on these new distributions. In Figure 12 it can be recognized that the bulk of the distribution was shifted to the positive side and the peak at -2000 was raised. The probability of scores between -2000 and 0 decreased significantly. An interpretation of this would be that the true score is most likely either positive or -2000 if it is negative.

	The customized loss function is now applied on this updated distribution of the reservoir score. This is visualized in Figure 13 in which the expected losses are compared before and after uncertainty reduction. The risk-neutral loss function is depicted in plots A1 and A2, while in plots B1 and B2 different risk-affinities are included. It is observable, that after adding information about layer thickness likelihoods, Bayesian estimators shift relative to the nature of the information. In this case, the added data generally reinforces the likelihoods of the reservoir to be significantly thick and for the seal to be appropriately reliable. Thus, the minima of expected loss are found at higher estimates for all actors.

	\section{Quantifying the value of adding information and reducing uncertainty}

	TEXT NOT UP TO DATE!!!

	This shift in Bayesian estimator can be quantified and shows in this case, that the risk-neutral actor experienced the greatest shift and actors most different in risk-affinity were were influenced the least, but still significantly.

	\section{Case results}

	Bla Bla

### 2017-04-18 Risk analysis/Text/Layout/Layout/MSc_layout.tex
Big update: Chapter 3! Some figures in notebooks. New model exports.

@@
     \part{First Part} % you can divide your thesis into different parts, for example Theory&Modeling in part 1 and real data examples in part 2.
-		\include{chapter2}
+		\include{chapter2_methods}
 
@@
     \bibliographystyle{apalike}
-    \printbib{MyBib} % in MyBib.bib you add all your reference information, following the correct format.
+    \printbib{complete_bib} % in MyBib.bib you add all your reference information, following the correct format.
     % The *.bin file needs to be compiled using the Bib-compile option of your LaTeX text editor.

### 2017-04-18 Risk analysis/Text/Layout/Layout/chapter1_orig.tex
Big update: Chapter 3! Some figures in notebooks. New model exports.

--- 2017-01-20 Risk analysis/Text/Layout/Layout/chapter1.tex
@@
     This is Chapter\ \ref{chap::intro}. If you want to know more about \LaTeX\ you better read
-    \cite{texbook} or use the extensive help available on the internet.
-\index{LaTeX}.
+    \cite{texbook} or use the extensive help available on the internet.\index{LaTeX}.
 This 'hidden' index command helps you making an index at the end of your thesis.
@@
 Also, hyperlinks are created automatically with the developed class file, such that your digital PDF version of your thesis can be read dynamically.
- Have fun with \LaTeX\ and your M.Sc. research project and good luck! \\ \\
-
-The purpose of the introduction is to tell readers why they should want to read what follows the introduction.
-This chapter should provide sufficient background information to allow readers to understand the context and significance of the problem.
-This does not mean, however, that authors should use the introduction to rederive established results or to indulge in other needless repetition.
-The introduction should (1) present the nature and scope of the problem; (2) review the pertinent literature, within reason; (3) state the objectives; (4) describe the method of investigation; and (5) describe the principal results of the investigation.
+ Have fun with \LaTeX\ and your M.Sc. research project and good luck!

### 2017-04-18 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Big update: Chapter 3! Some figures in notebooks. New model exports.

@@
-    \chapter{Methods}
+    \chapter{Methods}\label{cha:met}
 
@@
         \section{Bayesian analysis and decision theory}
-	    Problems and reasoning behind decision making are examined in the field of decision theory, as implied by the name \cite{berger2013stat}.
+	    Problems and reasoning behind decision making are examined in the field of decision theory, as implied by the name \citep{berger2013stat}.
 Such decision problems are commonly influenced by parameters that are uncertain.
@@
 Prior information might be given for example due to experience from previous similar problems or from expert knowledge (see Batvold and Begg).
-The approach of utilizing priors is known as Bayesian analysis and is explained in the following \cite{berger2013stat}.
+The approach of utilizing priors is known as Bayesian analysis and is explained in the following \citep{berger2013stat}.
 (It goes well with decision theory.)
@@
 	    Some basic elements are to be defined.
-The unknown (uncertain) quantity influencing decision making is usually denominated as $\theta$, the state of nature \cite{berger2013stat}.
+The unknown (uncertain) quantity influencing decision making is usually denominated as $\theta$, the state of nature \citep{berger2013stat}.
 Given statistical information on $\theta$ in the form of probability distributions, $\theta$ is called the parameter.
@@
 	    The outcome of statistical tests in form of information or statistical evidence is denoted as $X$.	    
-	    Loss is defined as $L(\theta,a)$, so $L(\theta_1,a-1)$ is the actual loss incurred when action $a_1$ is taken while the true state of nature is $\theta_1$ \cite{berger2013stat}.
+	    Loss is defined as $L(\theta,a)$, so $L(\theta_1,a-1)$ is the actual loss incurred when action $a_1$ is taken while the true state of nature is $\theta_1$ \cite{pberger2013stat}.
 Loss, expected loss and loss functions are further explained below.
@@
         \subsection{Bayesian inference}
-        Bayesian inference is most importantly characterized by its preservation of uncertainty, in contrast to standard statistical inference \cite{davidson2015}.
+        Bayesian inference is most importantly characterized by its preservation of uncertainty, in contrast to standard statistical inference \citep{davidson2015}.
 Probability is seen as a measure of belief for an event to occur.
-It has been argued by \cite{davidson2015}  that this Bayesian approach is intuitive and inherent in the natural human perspective.
-These beliefs can be assigned to individuals \cite{davidson2015}.
-Thus, different and even contradicting beliefs about the probability of an event might be held by different individuals, based on variations and disparities in the information available to each one individual \cite{davidson2015}.
+It has been argued by \citep{davidson2015}  that this Bayesian approach is intuitive and inherent in the natural human perspective.
+These beliefs can be assigned to individuals \citep{davidson2015}.
+Thus, different and even contradicting beliefs about the probability of an event might be held by different individuals, based on variations and disparities in the information available to each one individual \citep{davidson2015}.
 
-        The initial belief or guess about an event A can be denoted as P(A) \cite{davidson2015}.
+        The initial belief or guess about an event A can be denoted as P(A) \citep{davidson2015}.
 This is used as the so-called prior probability on which Bayesian updating is based.
@@
 These observations are included as likelihoods P(X$|$A).
-This process of updating results in a posterior probability P(A$|$X) \cite{davidson2015}.
+This process of updating results in a posterior probability P(A$|$X) \citep{davidson2015}.
 It is important to note that the prior is not simply discarded but re-weighted by Bayesian updating.
-It was also pointed out by \cite{davidson2015} that by utilizing an uncertain prior, the potential for wrongfulness of the initial guess is already included.
-This means that Bayesian updating is about reducing uncertainty in a belief and reaching a guess that is less wrong \cite{davidson2015}.
+It was also pointed out by \citet{davidson2015} that by utilizing an uncertain prior, the potential for wrongfulness of the initial guess is already included.
+This means that Bayesian updating is about reducing uncertainty in a belief and reaching a guess that is less wrong \citep{davidson2015}.
         Bayesian updating is defined by and conducted via the following equation, called the Bayes' Theorem:
@@
 Common and simple examples for such estimators are the mode (i.e.
-the generalized maximum likelihood estimate), the mean and the median of a distribution \cite{berger2013stat}.
+the generalized maximum likelihood estimate), the mean and the median of a distribution \citep{berger2013stat}.
 The presentation of a point estimate should usually come with a measure for its estimation error.
-According to \cite{berger2013stat}, the posterior variance is most commonly used as an indication for estimate accuracy.
-However, it is argued by \cite{davidson2015} that by using pure accuracy metrics, while this technique is objective, it ignores the original intention of conducting the statistical inference in cases, in which payoffs of decisions are valued more than their accuracies.
-A more appropriate approach can be seen in the introduction of loss and the use of loss functions \cite{davidson2015}.
+According to \citet{berger2013stat}, the posterior variance is most commonly used as an indication for estimate accuracy.
+However, it is argued by \citet{davidson2015} that by using pure accuracy metrics, while this technique is objective, it ignores the original intention of conducting the statistical inference in cases, in which payoffs of decisions are valued more than their accuracies.
+A more appropriate approach can be seen in the introduction of loss and the use of loss functions \citep{davidson2015}.
 
-        \subsection{Expected loss and loss functions} 
-        Loss is a statistical measure of how bad an estimate is, i.e. how much is lost by making a certain decision. Gains are considered by statisticians as negative losses \cite{davidson2015}.
-        The magnitude of an estimate's loss is defined by a loss function, which is a function of the estimate of the parameter and the true value of the parameter \cite{davidson2015}:
+        \subsection{Expected loss and loss functions}\label{sec:loss} 
+        Loss is a statistical measure of how bad an estimate is, i.e. how much is lost by making a certain decision. Gains are considered by statisticians as negative losses \citep{davidson2015}.
+        The magnitude of an estimate's loss is defined by a loss function, which is a function of the estimate of the parameter and the true value of the parameter \citep{davidson2015}:
 
@@
 Two standard loss functions are the absolute-error and the squared-error loss function.
-Both are simple to understand and commonly used \cite{davidson2015}.
+Both are simple to understand and commonly used \citep{davidson2015}.
 
-        As implied by its name, the absolute-error loss function returns loss as the absolute error, i.e. difference between the estimate and the true parameter \cite{davidson2015}:
+        As implied by its name, the absolute-error loss function returns loss as the absolute error, i.e. difference between the estimate and the true parameter \citep{davidson2015}:
 
-        \begin{equation}\label{eq:LossFunction}
+        \begin{equation}\label{eq:AbsLossFunction}
         L(\theta,\hat{\theta}) = |\theta - \hat{\theta}|
@@
 
-        Using the squared-error loss function returns losses that increase quadratically with distance of the estimator to the true parameter value \cite{davidson2015}:
+        Using the squared-error loss function returns losses that increase quadratically with distance of the estimator to the true parameter value \citep{davidson2015}:
 
-        \begin{equation}\label{eq:LossFunction}
+        \begin{equation}\label{eq:SqrLossFunction}
         L(\theta,\hat{\theta}) = |\theta - \hat{\theta}|^2
@@
         This exponential growth of loss also means that large errors are weighted much stronger than small errors.
-This might come with overvaluation of distant outliers and misrepresentation of magnitudes in distance.
-Regarding this, the absolute-error loss function can be seen as more robust \cite{davidson2015}.
+This might come with over-valuation of distant outliers and misrepresentation of magnitudes in distance.
+Regarding this, the absolute-error loss function can be seen as more robust \citep{davidson2015}.
 
@@
 
-        -shift in prespective. look at outcome instead of estimation precision!
+        -shift in perspective. look at outcome instead of estimation precision!
 
-        \cite{davidson2015} and (LF DESIGN PAPER?) propose that it might be useful to move away from these type of objective loss functions to the design of new loss functions that are specifically customized to an reflect an individual's, i.e.
+        \citet{davidson2015} and (LF DESIGN PAPER?) propose that it might be useful to move away from these type of objective loss functions to the design of new loss functions that are specifically customized to an reflect an individual's, i.e.
 the decision maker's objectives, preferences and outcomes.
@@
         The standard loss function defined above are symmetric, but can easily be adapted to be asymmetric, for example by weighting on the negative side stronger than those on the positive side.
-Preference over estimates larger than the true value is thus incorporated in an uncomplicated way \cite{davidson2015}.
-Much more complicated designs of loss functions are possible, depending on purpose, objective and application \cite{davidson2015}.
+Preference over estimates larger than the true value is thus incorporated in an uncomplicated way \citep{davidson2015}.
+Much more complicated designs of loss functions are possible, depending on purpose, objective and application \citep{davidson2015}.
 Some case-specific loss functions are designed in the following chapters of this work.
@@
 
-        The presence of uncertainty during decision making implies that the true parameter is unknown and thus the truly incurred loss $L(\theta,a)$ cannot be known at the time of making the decision \cite{berger2013stat, davidson2015}.
+        The presence of uncertainty during decision making implies that the true parameter is unknown and thus the truly incurred loss $L(\theta,a)$ cannot be known at the time of making the decision \citep{berger2013stat, davidson2015}.
 The Bayesian perspective considers unknown parameters as random variables and samples that are drawn from the posterior distribution as possible realizations of the unknown parameter, i.e.
-all possible true values are represented by this distribution \cite{davidson2015}.
-A suitable alternative to the actual loss is to consider a decision's expected loss and to make a decision that is optimal in relation to this expected loss \cite{berger2013stat}.
+all possible true values are represented by this distribution \citep{davidson2015}.
+A suitable alternative to the actual loss is to consider a decision's expected loss and to make a decision that is optimal in relation to this expected loss \citep{berger2013stat}.
 
-        Given a posterior distribution $P(\theta|X)$, the expected loss of choosing an estimate $\hat{\theta}$ over the true parameter $\theta$ (after evidence X has been observed) is defined by the function below \cite{davidson2015}:
+        Given a posterior distribution $P(\theta|X)$, the expected loss of choosing an estimate $\hat{\theta}$ over the true parameter $\theta$ (after evidence X has been observed) is defined by the function below \citep{davidson2015}:
 
@@
         The expectation symbol $E$ is subscripted with $\theta$, by which it is indicated that $\theta$ is the respective unknown variable.
-This expected loss as defined above, is also referred to the risk of estimate $\hat{\theta}$ \cite{davidson2015}.
+This expected loss as defined above, is also referred to the risk of estimate $\hat{\theta}$ \citep{davidson2015}.
 
-        By the Law of Large Numbers, the expected loss of $\hat{\theta}$ can be approximated drawing $N$ samples from the posterior distribution, respectively applying a loss function $L$ and averaging of the number of samples \cite{davidson2015}:
+        By the Law of Large Numbers, the expected loss of $\hat{\theta}$ can be approximated drawing $N$ samples from the posterior distribution, respectively applying a loss function $L$ and averaging of the number of samples \citep{davidson2015}:
 
@@
 
-        Minimization of \ref{eq:ExpectedLoss} or \ref{eq:ExpectedLoss2} returns a Bayesian point estimate known as Bayes action $\delta^P(X)$ (???), which is the estimate, action or decision with the least expected loss according to the loss function \cite{berger2013stat}.
-For a unimodal and symmetric absolute-error loss function, the Bayes action is simply the median of the posterior distribution, while squared-error loss it is the mean \cite{davidson2015, berger2013stat}.
-The MAP estimate is returned when using zero-one loss \cite{davidson2015}  (WRITE THIS?).
-The possibility of more than on minimum also implies that several Bayes actions can exist for one problem \cite{berger2013stat}.
+        Minimization of \ref{eq:ExpectedLoss} or \ref{eq:ExpectedLoss2} returns a Bayesian point estimate known as Bayes action $\delta^P(X)$ (???), which is the estimate, action or decision with the least expected loss according to the loss function \citep{berger2013stat}.
+For a unimodal and symmetric absolute-error loss function, the Bayes action is simply the median of the posterior distribution, while squared-error loss it is the mean \citep{davidson2015, berger2013stat}.
+The MAP estimate is returned when using zero-one loss \citep{davidson2015}  (WRITE THIS?).
+The possibility of more than on minimum also implies that several Bayes actions can exist for one problem \citep{berger2013stat}.
 
@@
 
-        \cite{davidson2015} implemented different risk affinities by simply  introducing a risk parameter into the loss function.
-By using different values for this parameter, it can be represented how comfortable an individual is with being wrong and furthermore which "side of wrong" is preferred by this decision maker \cite{davidson2015}.
+        \citet{davidson2015} implemented different risk affinities by simply  introducing a risk parameter into the loss function.
+By using different values for this parameter, it can be represented how comfortable an individual is with being wrong and furthermore which "side of wrong" is preferred by this decision maker \citep{davidson2015}.
 In the chapter below, different risk parameters are introduced to alter the weighting in the loss functions (???).
@@
         - applying these statistical theories in the context of geological modeling 
-        - as done by De la Varga (2015)
+        - as done by De la Varga (2016)
 
@@
 		\subsection{Implementing Bayesian analysis numerically with Python and PyMC3}
-		Bayesian analysis can be conducted using probabilistic programming \cite{salvatier2016pymc3}.
+		Bayesian analysis can be conducted using probabilistic programming \citep{salvatier2016pymc3}.
 For doings this, the programming language of choice in this work is Python.
-The merits of Python have been pointed out by \cite{behnel2010, salvatier2016pymc3, Langtangen2008}.
+The merits of Python have been pointed out by \citet{behnel2010, salvatier2016pymc3, Langtangen2008}.
 Development is facilitated by an expressive but concise and clean syntax that is easy to learn.
 Python is dynamic, compatible with multiple platforms and offers good support for numerical computing.
-Integration of other scientific libraries and extension via C, C++, Fortran or Cython is easily possible \cite{behnel2010, salvatier2016pymc3, Langtangen2008}.
-Python is thus a straightforward tool for the implementation of central components of Bayesian analysis, such as custom statistical distributions and samplers \cite{salvatier2016pymc3}.
+Integration of other scientific libraries and extension via C, C++, Fortran or Cython is easily possible \citep{behnel2010, salvatier2016pymc3, Langtangen2008}.
+Python is thus a straightforward tool for the implementation of central components of Bayesian analysis, such as custom statistical distributions and samplers \citep{salvatier2016pymc3}.
 
-		In probabilistic programming, programming variables are used as components to build probabilistic models \cite{davidson2015}.
+		In probabilistic programming, programming variables are used as components to build probabilistic models \citep{davidson2015}.
 
-		The numerical implementation of Bayesian analysis is further aided by the use of the Python library PyMC3, which was developed for conducting Bayesian inference and prediction problems in an open-source probabilistic programming framework \cite{davidson2015, salvatier2016pymc3}.
-Different model fitting techniques are provided in PyMC3, such as the \textit{maximum a posteriori} (MAP) method and several \textit{Monte Carlo Markov Chain} (MCMC) sampling methods \cite{salvatier2016pymc3}.
+		The numerical implementation of Bayesian analysis is further aided by the use of the Python library PyMC3, which was developed for conducting Bayesian inference and prediction problems in an open-source probabilistic programming framework \citep{davidson2015, salvatier2016pymc3}.
+Different model fitting techniques are provided in PyMC3, such as the \textit{maximum a posteriori} (MAP) method and several \textit{Monte Carlo Markov Chain} (MCMC) sampling methods \citep{salvatier2016pymc3}.
 		Name examples for state-of-the-art samplers included.
-		\cite{salvatier2016pymc3} point out that the development of PyMC3 continuing, as the inclusion of further tools is planned for future updates.
+		\citet{salvatier2016pymc3} point out that the development of PyMC3 continuing, as the inclusion of further tools is planned for future updates.

### 2017-04-18 Risk analysis/Text/Layout/Layout/chapter3_1Dcase.tex
Big update: Chapter 3! Some figures in notebooks. New model exports.

@@
 
-An abstract one-dimensional case of a petroleum reservoir is presented in this chapter.
-The underlying model and basic approach are inherited from De la Varga (2016).
-Parameters were adapted to better represent a reasonable geological petroleum system consisting of a reservoir with overlying seal in the subsurface.
-This way, a certain economic significance is ascribed to the model and further relevant questions can be derived.
-Regarding the petroleum sector, the problem of interest is most commonly one of estimating the monetary value contained in an oil or gas reservoir.
+For simple understanding and working out the application of the methods, first, an abstract one-dimensional reservoir case is presented in this chapter.
+The underlying model and basic approach are inherited from \citet{delaVarga2016}.
+Parameters were adapted to more appropriately represent a reasonable geological petroleum system, consisting of a reservoir with overlying seal in the subsurface.
+In doing this, a certain economic significance is ascribed to the model and further relevant questions can be derived.
+Regarding the petroleum sector, the problem of interest is most commonly one of estimating the volume of resources or monetary value contained in an oil or gas reservoir.
 Limiting the model to only one dimension and a small number of uncertain parameters allows for a relatively straightforward and simplified approach to assessing an abstract type of value for a reservoir and designing a respective loss function for value estimation.
@@
 
-	\section{Constructing the one-dimensional model}
+	\section{Constructing the one-dimensional model}\label{sec:1D_construction}
 
-	De la Varga (2016) constructed a simple geological model using three uncertain locations in one-dimensional space, which mark the boundaries of layers.
-The probability of positions of these points are defined by sampling from normal distributions.
+	\citet{delaVarga2016} constructed a simple geological model using three uncertain positions in vertical one-dimensional space, marking hypothetical boundaries of layers in a subsurface column.
+The location probabilities for these points are defined by sampling from normal distributions.
 Standard deviations of these distributions increase with depth, representing an increase in uncertainty.
-For an approximate representation of a petroleum system, the distribution means were set to depths of 2000 (seal top), 2050 (reservoir top) and 2200 (reservoir bottom).
+For an approximate representation of a petroleum system, the distribution means were set to depths of 2000 m (seal top), 2050 m (reservoir top) and 2200 m (reservoir bottom).
 These points confine two layers in the middle, from which the upper one can be labeled as seal and the lower one as reservoir.
-The resulting model with its possible layer boundary locations is illustrated in Figure 1.
+The resulting model with its possible layer boundary locations is illustrated in Figure \ref{fig:1D_model}.
 
-	\section{Assessing reservoir quality using scores}
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=1\textwidth]{Figures/1D_model.png}
+		\caption{Probability distributions for positions of layer boundaries in the subsurface and a respective representation using lines.}\label{fig:1D_model}
+	\end{figure}
+
+	\section{Assessing reservoir quality using scores}\label{sec:scoring}
 
@@
 What can be deduced from the distributions of layer boundary locations, is the thickness of the seal and the reservoir, as well as the depth of both, again as probability distributions.
-To assess the reservoir quality in an abstract way, it can be assigned with a score.
-This score is made dependant on the three uncertain parameters (1) reservoir thickness, (2) reservoir top depth and (3) seal thickness (see Figure 2).
+To assess the reservoir quality in an abstract way, it can be assigned a score.
+This score is made dependent on the three uncertain parameters (1) reservoir thickness, (2) reservoir top depth and (3) seal thickness (see Figure \ref{fig:3_parameters}).
+
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=1\textwidth]{Figures/3_parameters.png}
+		\caption{Probability distributions for the three parameters used for modeling reservoir score.}\label{fig:3_parameters}
+	\end{figure}
 
@@
 Samples from the probability distributions of these two parameters are drawn to model the true score of the reservoir (depth scores are subtracted from reservoir thickness scores).
-For the data used here, it can be seen in Figure 3 that the results of this two parameter score modelling are represented by an approximately normal distribution.
+For the data used here, it can be seen in Figure \ref{fig:score_results1}-A, that the results of modeling with only these two parameters, are represented by an approximately normal distribution.
 The score is negative in about 17\% of the cases.
@@
 
-	Influence by the seal thickness is added in the following.
+	Next, influence by the third parameter seal thickness is included.
 Score points are not added or subtracted by this parameter directly.
@@
 Here it is set at 20 m thickness.
-If the seal thickness falls below this threshold, it is assumed that the seal fails completely and thus all the potential value of the reservoir is lost, while costs of depth remain.
+If the seal thickness falls below this threshold, it is assumed that the seal fails completely and thus all the potential value (positive score) of the reservoir is lost, while costs of depth (negative score) remain.
 Thus, a condition to check whether the seal is reliable is now included in the model.
-Results are visualized in Figure 4.
-The main distribution was not changed significantly, except for a striking peak of probability that emerged for the possibility of a score of -2000.
+Results are visualized in Figure \ref{fig:score_results1}-B.
+The main distribution is not changed significantly, except for a striking peak of probability at the possibility of a score of -2000.
 Furthermore, mean and median have been shifted to lower values and are now found further apart.
+
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=1\textwidth]{Figures/score_results1.png}
+		\caption{Posterior probability distributions from modeling scores using two (A) and three parameters (B).}\label{fig:score_results1}
+	\end{figure}
 
@@
 
-	Now that a distribution of reservoir score probabilities has been modelled, a loss function for estimation of the true score value can be developed.
+	Now that a distribution of reservoir score probabilities has been modeled, a loss function for estimation of the true score value can be developed.
 
@@
 
-	Some standard loss function were presented in chapter X (theory of loss fucntions to be explained in foregone chapter).
+	Some standard loss function were presented in chapter \ref{sec:loss}.
 For this case, the absolute-error loss and the squared-error loss function were considered as starting points, to design a more case-specific loss function.
-Their general forms for certain determined values are illustrated in Figure 5.
-	As explained before in chapter X, the expected loss for an estimate can be approximated by calculating the arithmetic mean of the losses compared to all samples in the modelled score distribution.
+	Calculating the expected loss for estimates ranging from -3000 to 6000 based on these two standard loss function results in the graphs depicted in Figure \ref{fig:abs_sqr_LF}.
+It can be observed that the median of the distribution coincides with the minimum of the absolute-error and the mean with the minimum of the squared-error expected loss.
 
-	1N∑i=1NL(θi,θ̂ )≈Eθ[L(θ,θ̂ )]=l(θ̂ )
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=1\textwidth]{Figures/abs_sqr_LF.png}
+		\caption{Expected loss based on the standard absolute-error (A) and squared-error (B) loss functions applied on the probability distribution resulting from \ref{sec:scoring} and for estimates ranging from -3000 to 6000.}\label{fig:abs_sqr_LF}
+	\end{figure}
 
-	Calculating the expected loss for estimates ranging from -3000 to 6000 using these two standard loss function results in the graphs depicted in Figure 6.
-As expected (find proper source!), the median of the distribution coincides with the minimum of the absolute-error and the median with the minimum of the squared-error expected loss (Bayes action/estimator).
-
-	\subsection{Adaptions to design a customized loss function}
+	\subsection{Adaptions for loss function customization}
 
@@
 Due to their symmetric properties, both will always give the median or mean of the underlying distribution as minimizing estimator respectively.
-However, assigning an economic notion to our model and assuming the case of an actor or decision maker in any field, naturally necessitates the consideration of preferences, interests and the overall subjective perspective such an individual or company might have.
-Further constraints and influences can also me specific to the field, industry or generally to the problem at hand.
+However, assigning an economic notion to our model and assuming the case of an actor or decision maker in any field, naturally necessitates the consideration of preferences, interests and the overall subjective perspective such an individual or for example a company might have.
+Further constraints, properties and factors can also be specific to the field, industry or generally to the problem at hand.
 Consequently, the design of a more specific non-standard and possibly asymmetric loss function might be required, so that an adapted Bayesian estimator can be found.
-One that includes subjective aspects and difference in weighting of particular gains or losses, arising from an actors preferences and the environment in which the actor has to estimate or make a decision.
+One that includes subjective aspects and difference in weighting of particular gains or losses, arising from an actor's inherent preferences and the environment in which the actor has to estimate or make a decision.
 In the face of several uncertain parameters, a perfectly true estimate is virtually unattainable.
-However, the attempt can be made, to design a customized loss function that returns a Bayesian estimator involving the least bad consequences for an actor in a specific environment.
-Regarding the petroleum system case modelled above, such an attempt is made and explained step by step in the following.
+However, an attempt can be made, to design a customized loss function that returns a Bayesian estimator involving the least bad consequences for an individual in a specific environment.
+Regarding the petroleum system case modeled above, such an attempt is made and explained step by step in the following.
 
-	Text NOT up-to-date!!!
-
-	For the purpose of estimation, it makes sense that a standard loss function is chosen as a basis.
-The absolute-error loss seems most approriate for this case of petroleum reservoir value estimation.
-As stated above, the reservoir score we defined is an abstract and simplified way to reflect a value contained in the reservoir.
-An actor would wish to extract this, should it be positive (and feasible, but this would require the inclusion of numerous more factors).
-Ideally, an actor would like to know the exact true score, so that investments or resources can appropriately be allocated.
+	For the purpose of estimation, it makes sense that one of the standard loss functions is chosen as a basis and a customized loss function is developed from there.
+The absolute-error loss seems most appropriate for this case of petroleum reservoir quality estimation.
+As stated above, the reservoir score is an abstract and simplified way to reflect a value contained in the reservoir.
+Ideally, an actor would like to know the exact true score, so that investments or resources can be allocated appropriately, in order to acquire economic gains.
 This allocation is the decision to be made or action to be taken.
 Deviations from the unknown true score in the form of over- and underestimation bring about an error and loss accordingly.
-In principle, there is no reason for loss to increase exponentially with distance from the true value.
-Allocation of investments would not increase exponentially with linear increase of the value of the resource.
+In this case, there is no reason for loss to increase exponentially with distance from the true value.
+It is assumed that investments increase linearly with linear growth in the value of the resource.
 For this reason, the absolute-error loss function is favored over the squared-error loss function in this case.
-Some adaptions are made below, based on mostly logical case-specific assumptions.
+Some customization steps are taken below, based on mostly logical case-specific assumptions.
 
-	    1st Adaption: For any negative estimate, an actor will take the same action as for a zero score, i.e.
-not to invest at all, since there is no positive value present.
-If the true score is zero or negative, this means that over- and underestimating, while the estimate is also negative or zero, is irrelevant an results in:
+	\begin{enumerate}
+		\item Step I: The standard symmetrical absolute-error loss function is chosen as a starting point for further customization steps:
 
-	L(θ,θ̂ )=0
+		\begin{equation}
+		L(\theta,\hat{\theta}) = |\theta - \hat{\theta}|
+		\end{equation}
 
-	In other words: There is nothing to win and nothing is lost, since no action is taken.
+		\item Step II: Considering the development of a petroleum reservoir, it can be assumed that over-investing is worse than under-investing.
+Overestimating the size of a reservoir might for example lead to the installation of equipment or facilites that are actually redundant or unnecessary.
+This would come with additional unrecoverable expenditures.
+Consequences from underestimating, however, may presumably be easier to resolve.
+Additional equipment can often be installed later on.
+Hence, overestimation is weighted stronger in our loss function by multiplying the error with an overestimation factor \textit{a} (= 1.25):
 
-	    2nd Adaption: Considering the development of a petroleum reservoir, one might also assume that overinvesting is worse than underinvesting.
-Overestimating the size of a reservoir might for example lead to the installation of equipment or facilites that are actually not needed.
-This comes with additional unrecoverable expenditures.
-Consequences from underestimating, however, may presumably be easier to resolve.
-Additional equipment can often be installed later afterwards.
-Hence, overestimation is weighted stronger in our loss function by multiplying the error with an overstimation factor a = 1.25:
+		\begin{equation}\label{eq:LF_II}
+		L(\theta,\hat{\theta}) = |(\theta-\hat{\theta})|*a
+		\end{equation}
 
-	L(θ,θ̂ )=|(θ−θ̂ )|∗a
+		\item Step III: The worst case for any project would be that its development is set into motion, expecting a gain, only to discover later that the value in the reservoir does not cover the costs of realizing the project, resulting in an overall loss.
+A petroleum system might also turn out to be a complete failure, containing no value at all, although the actor's estimate indicated the opposite.
+Here, this is referred to as a worst case or fatal overestimation.
+A positive score is estimated, but the true score is zero or negative.
+This is worse than the "normal" non-fatal overestimation, where both values are positive and a net gain is still achieved, which is only smaller then the best possible gain of expecting the true score.
+Fatal overestimation is included in the loss function by using another weighting factor \textit{b} (= 2) that replaces \textit{a}:
 
-	    3d Adaption: The worst case for any project would be, to start and progress its development, only to discover later that the value in the reservoir does not cover the costs of realizing the project.
-A petroleum system might also turn out to be a complete failure, although the actor's estimate indicated the opposite.
-This is refered to here as a worst case or fatal overestimation, when a positive score is estimated, but the true score is zero or negative.
-It is included in the loss function with a second weighting factor b = 2:
+		\begin{equation}\label{eq:LF_III}
+		L(\theta,\hat{\theta}) = |(\theta-\hat{\theta})|*b
+		\end{equation}
 
-	L(θ,θ̂ )=|(θ−θ̂ )|∗b
+		(In other words: Worst case or fatal overestimation is twice as bad as simple underestimation.)
 
-	In other words: Worst case or fatal overestimation is twice as bad as simply underestimating.
+		\item Step IV: A worst case or fatal underestimation can also be derived from the idea of estimating a zero or negative score, when the true score is actually positive.
+This is assumed to be worse than non-fatal overestimation, but clearly better than fatal overestimation.
+No already owned resources are wasted, it is only the potential value that is lost, i.e.
+opportunity costs that arise from completely discarding a reservoir with a potential gain equal to the positive true score.
+Fatal underestimation is weighted using a third factor \textit{c}:
 
-	    4th Adaption: A worst case or fatal underestimation can also be derived from the idea of estimating a zero or negative score, when the true score is actually positive.
-This basically reflects the opportunity costs of completely discarding a reservoir with a true score of θ:
+		\begin{equation}\label{eq:LF_IV}
+		L(\theta,\hat{\theta}) = |(\theta-\hat{\theta})|*c
+		\end{equation}
 
-	L(θ,θ̂ )=θ
+	\end{enumerate}
 
-	A step by step realization of these four adaptions is depicted in the plots in Figure 7.
-The first adaption is visualized in plot IIa, reflecting the irrelevance of negative estimates.
-Assignment of a stronger weight to overestimation, has led to a shift of the Bayesian action to a lower value in IIb.
-This shift has been in reinforced in plott IIc, due to adopting the concept of fatal overestimation.
-Finally, the inclusion of worst case underestimation raises the baseline for expected loss of not developing the reservoir (estimating zero or negative score).
+	Combining these adaption steps and the conditions defined in them, results in the following customized loss function:
 
-	The implementation of this customized loss function using single deterministed values for the true score is plotted in Figure 8.
+	\begin{equation}\label{eq:LF_final}
+	L(\theta,\hat{\theta}) =
+	\begin{cases}
+	|\theta - \hat{\theta}|, & \text{for } 0<\hat{\theta}<\theta  \\
+	|\theta-\hat{\theta}|*a, & \text{for } 0<\theta<\hat{\theta} \\
+	|\theta-\hat{\theta}|*b, & \text{for } \theta\leq0<\hat{\theta} \\
+	|\theta-\hat{\theta}|*c, & \text{for } \hat{\theta}\leq0<\theta 
+	\end{cases},
+	\text{ with } a,b,c \in \mathbb{Q}
+	\end{equation}
 
-	It has to be emphasized that this is just one possible proposal for customization.
-There exists not one perfect loss function design for such a case.
-Other designs might differ slightly, for example by choosing different values for the same factors, to fundamentally by basing it on a significantly different mathematical structure.
-Loss functions are customized according to the subjective needs and objectives of an actor.
-Thus, they are defined by the actor expressing his perspective and might change, should the actor change his mind.
-Especially considering individual persons as actors, even psychological aspects may play a significant role.
+	Realizations of the four adaption steps are depicted in the plots in Figure \ref{fig:LF_4steps}.
+It is important to note that the weighting factors defined above can take basically any numerical values but should be chosen in a way that they appropriately represent the framework conditions of the problem.
+Here, for example, it is assumed that normal overestimation is 25\% (\textit{a} = 1.25), fatal overestimation 100\% (\textit{b} = 2) and fatal underestimation 50\% (\textit{c} = 1.5) worse than normal underestimation.
+
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=1\textwidth]{Figures/LF_4steps.png}
+		\caption{The single steps of customizing the loss function are depicted in plots I - IV.}\label{fig:LF_4steps}
+	\end{figure}
+
+	In Figure \ref{fig:LF_4steps} it can be observed, that assigning a stronger weight on overestimation (plot II) steepens the curve on the right hand side and shifts the minimum to the left, i.e.
+to a lower estimate.
+Using \textit{a} = 1.25, the Bayes action changes from the median, to a value close to the mean of the distribution.
+The shift and steepening are significantly reinforced by the introduction of fatal overestimation (plot III).
+With \textit{b} = 2, the Bayes action drops to the zero score estimate.
+It can also be noted, that by defining a condition dependent on the algebraic sign of the values, according to which only losses for positive estimates are multiplied by \textit{b}, a distinct jump appears at the zero score boundary.
+Due to a similar condition, the same effect is observed on the negative side of estimate values, where the curve has also been steepened, after including fatal underestimation (plot IV).
+This comes with a shift of the minimum towards positive values.
+It is also to be noted that with every customization step, the overall expected loss is increased.
+
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=1\textwidth]{Figures/LF4_det_values.png}
+		\caption{Loss based on the customized loss function (IV) for determined true scores of -750, 0 and 750.}\label{fig:LF4_det_values}
+	\end{figure}
+
+	The implementation of this customized loss function (IV) using single determined values for the true score is plotted in Figure \ref{fig:LF4_det_values}.
+This helps to clarify the way real losses result for each guess, relative to given true score values.
+The expected loss is acquired by arithmetically averaging such loss realizations based on the true score probability distribution by using formula \ref{eq:ExpectedLoss2}.
+
+	It has to be emphasized that this is just one possible proposal for loss function customization.
+There exists not one perfect design for such a case.
+Slight to strong changes can already be implemented by simply varying the values of the weighting factors \textit{a, b} and \textit{c}.
+Fundamentally different loss functions can also be based on a significantly different mathematical structure.
+Loss functions are customized regarding the problem environment and according to the to the subjective needs and objectives of an actor.
+Thus, they are mostly defined by the actor expressing his perspective.
+Changes in the individual's perception and viewpoint might lead to further customization needs even later on.
+(Especially considering individual persons as actors, psychological aspects may play a significant role.)
 
@@
 
-	TEXT NOT UP TO DATE!!!
+	One can assume that several actors in one sector or decision environment may have the same general loss function, but different affinities concerning risks.
+This might be based for example on the different psychological factors or economic philosophies followed by companies.
+It might also be based on budgets and options such actors have available.
+An intuitive example is the comparison of a small and a large company.
+A certain false estimate or error might have a significantly stronger impact on a company which has a a generally lower market share and only few projects, than on a larger company which might possess a higher financial flexibility and for which one project is only one of many development options in a portfolio.
 
-	One can assume that several actors in one sector or decision environment may have the same general loss functions, but different affinities concerning risks.
-This might be based on psychological factors or economic philosophies followed by a company.
-It might also be based on the budget and options such an actor has available.
-An intuitive example is the comparison of a small and a large company.
-A certain false estimate or error might have a significantly stronger impact on a company which has a a generally lower market share and only few projects, than on a larger company which might possess a higher financial flexibility and for which one project is only one of many development options.
+	In the following, the loss function is further adapted to consider different risk-affinities of several actors.
+Representing risk behavior in a loss function can also be done in different ways and regarding different types of risks.
+Here, bidding lower is considered the cautious, risk-averse option, as smaller losses can be expected from underestimating.
+Guessing higher is deemed riskier.
+Losses from overestimation are greater.
+However, bidding correctly on a higher value, will also return a greater gain.
+It is assumed that risk-friendly actors care less about fatal underestimation, i.e.
+they will rather develop a project than discard it.
+In the loss function, risk is simply included using a risk factor \textit{r} which alters the weighting factors \textit{a, b} and \textit{c} respectively:
 
-	In the following, the loss function is further adapted to consider different risk-affinities of different actors.
-Representing risk behavior in a loss function can also be done in different ways.
-Here, it is simply included as a the factor r that is multiplied with the overestimation weighting factor a and b:
+	\begin{equation}\label{eq:LFR_final}
+	L(\theta,\hat{\theta}) =
+	\begin{cases}
+	|\theta - \hat{\theta}|, & \text{for } 0<\hat{\theta}<\theta  \\
+	|\theta-\hat{\theta}|*(a*r), & \text{for } 0<\theta<\hat{\theta} \\
+	|\theta-\hat{\theta}|*(b*r), & \text{for } \theta\leq0<\hat{\theta} \\
+	|\theta-\hat{\theta}|*(c*(r^{-0.5})), & \text{for } \hat{\theta}\leq0<\theta 
+	\end{cases},
+	\text{ with } a,b,c,r \in \mathbb{Q}
+	\end{equation}
 
-	L(θ,θ̂ )=|(θ−θ̂ )|∗(a∗r)
+	According to this, for \textit{r} = 1 the risk-neutral loss function is returned, since \textit{a, b} and \textit{c} are not altered.
+For \textit{r} $<$ 1, the weight on overestimating is reduced and increased for fatal underestimation.
+This represents a more risk-friendly actor that is willing to bid on a higher estimate to attain a greater gain.
+For \textit{r} $>$ 1, the overestimation weight is increased, the underestimation weight decreased in the loss function and respectively more risk-averse actors are prompted to bid on lower estimates.
 
-	L(θ,θ̂ )=|(θ−θ̂ )|∗(b∗r)
+	The factor \textit{r} can take basically any positive values.
+However, since risk-neutrality is expressed by\textit{r} = 1, values 0 $<$ \textit{r} $<$ 2 are considered to be the most appropriate choices to represent both sides of risk-affinity equally.
 
-	Thus, the risk-neutral loss function is returned for r = 1.
-For r < 1, the weight of overestimating is reduced.
-This represents a more risk-friendly actor that is willing to bid on a higher estimate to attain a higher return in value.
-For r > 1, the overestimation weight is increased and respectively more risk-averse actors are prompted to bid on lower estimates.
-They prefer a lower but safer return value, due to the higher loss of wrong overestimation.
-The implementation of a risk-affinity factor r leads to different steepnesses of the plotted curves of the loss functions on the positive side of estimates (see Figures 9 and 10).
-The same three given deterministic values are used as true scores in the three plots in Figure 9.
-For certain determined values, the steepnesses of the functions vary, but their loss minima are found at the same location.
-However, in Figure 10 it can be seen that the minima for expected loss, i.e.
+	Implementing it as a factor \textit{r} leads to different steepnesses of the plotted curves depicting loss and expected loss.
+The effect on the latter is visualized in Figure \ref{fig:LFR}, where 0.5, 0.75, 1, 1.25 and 1.5 were chosen as values for \textit{r}.
+	It can be observed that the minima for expected loss, i.e.
 the Bayesian estimators for differently risk-affine actors, are located at different estimates.
-Mean and median are clearly surpassed by the best estimate of the most risk-fiendly actor, while for the most risk-averse actor the Bayes action equals a zero estimate (and thus the decision to take on action).
+Mean and median are clearly surpassed by the best estimate of the most risk-friendly actor (\textit{r} = 0.5), while for the most risk-averse actors (\textit{r} 0.5 and \textit{r} = 0.75), the Bayes action equals a zero score estimate (and thus the decision to take on action).
+In Figure \ref{fig:LFR} it can also be recognized that the expected loss is generally lower for risk-friendlier actors (on the side of positive estimates).
+
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=1\textwidth]{Figures/LFR.png}
+		\caption{Plotting of expected loss realizations after including the risk factor r in the loss function (IV) for actors with risk-affinities ranging from risk-averse (r = 0.5 and 0.75), over risk-neutral (r = 1), to risk-friendly (r = 1.25 and r = 1.5).}\label{fig:LFR}
+	\end{figure}
 
@@
 
-	De la Varga (2016) made use of Bayesian inference to reduce the uncertainty in this type of one-dimensional model.
-The same is done here.
+	\citet{delaVarga2016} made use of Bayesian inference to reduce the uncertainty in this type of one-dimensional model. The same is conducted here in the following.
+
 	The probability distributions for the location of the layer boundaries are treated as priors.
-Likelihood functions for the thicknesses of the layers are introduced as new information in the model.
-For the seal, a normal distribution with a mean of 25 m and a standard devation of 20 m is chosen to reflect the likelihood.
-For the reservoir, a normal distribution with a mean of 180 m and a standard deviation of 60 m is chosen.
-Implementing these likelihoods in the model, the uncertainty in the posterior probability distributions for the layer boundary depths is reduced (see Figure 11).
+Now it is assumed that new observations have been made, providing additional information on the likelihoods of the thicknesses of the two layers.
+Likelihood functions for reservoir and seal thicknesses are introduced in the form of normal distributions, defined by means and standard deviations.
+These parameters vary according to nature of the observations made.
+Using the principle of Bayesian inference as explained in \ref{cha:met}, the model is updated and new posterior distributions for our true reservoir score are attained.
 
-	Modelling of a reservoir score is conducted as above, now based on these new distributions.
-In Figure 12 it can be recognized that the bulk of the distribution was shifted to the positive side and the peak at -2000 was raised.
-The probability of scores between -2000 and 0 decreased significantly.
-An interpretation of this would be that the true score is most likely either positive or -2000 if it is negative.
+	- VOI Quantification ???
 
-	The customized loss function is now applied on this updated distribution of the reservoir score.
-This is visualized in Figure 13 in which the expected losses are compared before and after uncertainty reduction.
-The risk-neutral loss function is depicted in plots A1 and A2, while in plots B1 and B2 different risk-affinities are included.
-It is observable, that after adding information about layer thickness likelihoods, Bayesian estimators shift relative to the nature of the information.
-In this case, the added data generally reinforces the likelihoods of the reservoir to be significantly thick and for the seal to be appropriately reliable.
-Thus, the minima of expected loss are found at higher estimates for all actors.
+	Different examples with different sets of likelihoods are conducted in the following, so that various possible results can be compared.
+Each take the layer boundaries defined in  \ref{sec:1D_construction} as priors and use Bayesian inference with data defined below for each example.
 
-	\section{Quantifying the value of adding information and reducing uncertainty}
+	\subsection{Updating example I (moderate reinforcement)- Seal: mean = 25~m, std = 20~m - Reservoir: mean = 180~m, std = 60~m}
 
-	TEXT NOT UP TO DATE!!!
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=1\textwidth]{Figures/update_moderate1.png}
+		\caption{...}\label{fig:update_moderate1} 
+	\end{figure}
 
-	This shift in Bayesian estimator can be quantified and shows in this case, that the risk-neutral actor experienced the greatest shift and actors most different in risk-affinity were were influenced the least, but still significantly.
+	In this first case, a normal distribution with a mean of 25~m and a standard deviation of 20~m is chosen to reflect the likelihood of the seal thickness.
+For the reservoir, a normal distribution with a mean of 180~m and a standard deviation of 60~m is chosen.
+Using these likelihoods to update the one-dimensional model, the uncertainty in the posterior probability distributions for the positions of layer boundaries in depth is reduced (see Figure \ref{fig:update_moderate1}).
 
-	\section{Case results}
+	Modeling the reservoir score is conducted as above, now based on these new distributions.
+In Figure \ref{fig:update_moderate2} it can be recognized that the bulk of the distribution was shifted to the positive side of scores, while the peak at -~2000 was raised.
+The probability of scores between -~2000 and 0 decreased significantly, i.e.
+the true score is most likely either positive or -~2000 if it is negative.
 
-	Bla Bla
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=1\textwidth]{Figures/update_moderate2.png}
+		\caption{...}\label{fig:update_moderate2} 
+	\end{figure}
+
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=1\textwidth]{Figures/update_moderate3.png}
+		\caption{...}\label{fig:update_moderate3} 
+	\end{figure}
+
+	The customized loss function (IV) is now applied on this updated distribution of the reservoir score.
+This is visualized in Figure \ref{fig:update_moderate3} in which the expected losses are compared before (B1 )and after (B2) uncertainty reduction.
+It is observable, that by adding information about layer thickness likelihoods, Bayes actions are shifted relative to the nature of the information.
+In this case, the added data generally reinforces the likelihoods of the reservoir to be significantly thick.
+Information on the seal, however, based on a normal distribution around 25~m thickness, leaves uncertainty about the reliability of the seal, as the safety threshold is defined as 20~m.
+
+	\begin{table}
+		\centering
+		\begin{tabular}[c]{| l | l | l |}
+			\hline
+			Risk factor \textit{r} & Shift in Bayes action & Change in expected loss \\ \hline
+			0.50 & +~553.67 & -~105.49 \\ 
+			0.75 & +~673.40 & -~81.67  \\ 
+			1.00 & +~750.05 & -~33.56 \\ 
+			1.25 & +~713.19 & +~75.71 \\ 
+			1.50 & +~0.00 & +~282.64  \\ 
+			\hline
+		\end{tabular}
+		\caption{Bla}
+		\label{tab:update_moderate_tab}
+	\end{table}
+
+	Increased certainty about the reservoir thickness is sufficient to shift Bayes actions to higher estimates for all actors, but the most risk-averse one.
+These shifts are quantified in Table \ref{tab:update_moderate_tab}.
+According to these numbers, the risk-neutral actor's estimate is increased the most.
+Expected losses decreased for the risk-neutral and the two risk-friendlier individuals.
+It is clear that the expected loss was reduced the most for the risk-friendliest actor and it increased most significantly for the most risk-averse actor.
+
+	\subsection{Updating example II (reliable seal) - Seal: mean = 50 m, std = 20 m - Reservoir: mean = 180 m, std = 60 m}
+
+	In this second case, the reservoir thickness likelihood is defined in the same way as before.
+For the seal, a mean of 50~m with a standard deviation of 20~m is chosen, favoring the likelihood of a reliable seal relative to the threshold of 20~m thickness.
+This results in the score distribution depicted in \ref{fig:update_goodseal2}.
+The bulk of the distribution is narrowed on the positive side of estimates.
+Mean and median are clearly shifted to higher values as well.
+The "seal failure" peak at -~2000 is significantly.
+decreased.
+
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=1\textwidth]{Figures/update_goodseal2.png}
+		\caption{...}\label{fig:update_goodseal2} 
+	\end{figure}
+
+	Applying the loss function on this new score distribution results in the realizations of expected loss illustrated in \ref{fig:update_goodseal3}.
+Bayes actions are shifted clearly to higher estimates and expected losses of these minima are significantly reduced for all actors.
+This is quantified in \ref{tab:update_goodseal_tab}.
+According to these values, the risk-neutral to risk-averse individuals seem to profit the most.
+
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=1\textwidth]{Figures/update_goodseal3.png}
+		\caption{...}\label{fig:update_goodseal3} 
+	\end{figure}
+
+	\begin{table}
+		\centering
+		\begin{tabular}[c]{| l | l | l |}
+			\hline
+			Risk factor \textit{r} & Shift in Bayes action & Change in expected loss \\ \hline
+			0.50 & +~522.95 & -~436.26 \\ 
+			0.75 & +~697.73 & -~551.92  \\ 
+			1.00 & +~855.57 & -~631.92 \\ 
+			1.25 & +~912.43 & -~647.10 \\ 
+			1.50 & +~819.26 & -~545.96  \\ 
+			\hline
+		\end{tabular}
+		\caption{Bla}
+		\label{tab:update_goodseal_tab}
+	\end{table}
+
+	\subsection{Updating example III (safe seal, but reservoir under average) - Seal: mean = 70 m, std = 10 m - Reservoir: mean = 100 m, std = 40 m}
+
+	In this third example, seal safety is ensured by using a mean of 70~m with a standard deviation of 10~m for the seal thickness likelihood.
+However, the new observations are assumed to provide information about the reservoir that makes it likely to be thinner as expected.
+Reservoir thickness likelihood is assigned a mean of 100~m and a standard deviation of 10~m.
+
+	The subsequent score distribution is depicted in \ref{fig:update_smallres2}.
+It can be seen that while the whole distribution is narrowed, it also is shifted to the left, to lower and negative estimates.
+Mean and median are almost the same.
+As seal reliability is practically guaranteed, the peak at -~2000~m vanishes.
+
+	Based on this probability distribution of the reservoir score, Bayes actions are shifted to lower estimates for all actors (\ref{fig:update_smallres3}).
+In fact, also but the most risk-friendly individual (\textit{r} = 0.5) find their Bayesian estimator to be zero now, i.e.
+project development is deemed to be too risky to them.
+Furthermore, the spread of expected loss values around the minima for all actors is diminished (the expected loss values of the Bayes actions are now much closer to each other).
+
+	Quantified changes in the position and the values of the Bayes actions are comparable to the results in example II (there is no shift for risk-averse actors, as they already found their best estimates to be zero).
+Most notable is the large reduction of expected loss in general (see \ref{tab:update_smallres_tab}).
+
+	\begin{figure}[h]
+			\centering
+			\includegraphics[width=1\textwidth]{Figures/update_smallres2.png}
+			\caption{...}\label{fig:update_smallres2} 
+	\end{figure}
+
+	\begin{figure}[h]
+				\centering
+				\includegraphics[width=1\textwidth]{Figures/update_smallres3.png}
+				\caption{...}\label{fig:update_smallres3} 
+	\end{figure}
+
+	\begin{table}
+		\centering
+		\begin{tabular}[c]{| l | l | l |}
+			\hline
+			Risk factor \textit{r} & Shift in Bayes action & Change in expected loss \\ \hline
+			0.50 & -~498.98 & -~538.08 \\ 
+			0.75 & -~346.16 & -~717.83  \\ 
+			1.00 & -~176.13 & -~872.02 \\ 
+			1.25 & -~0.00 & -~946.04 \\ 
+			1.50 & -~0.00 & -~951.57  \\ 
+			\hline
+		\end{tabular}
+		\caption{Bla}
+		\label{tab:update_smallres_tab}
+	\end{table}
+
+	\section{General case results}

### 2017-04-24 Risk analysis/Text/Layout/Layout/chapter3_1Dcase.tex
updated

@@
 These parameters vary according to nature of the observations made.
-Using the principle of Bayesian inference as explained in \ref{cha:met}, the model is updated and new posterior distributions for our true reservoir score are attained.
+Using the principle of Bayesian inference as explained in chapter \ref{cha:met}, the model is updated and new posterior distributions for our true reservoir score are attained.
 
@@
 	Different examples with different sets of likelihoods are conducted in the following, so that various possible results can be compared.
-Each take the layer boundaries defined in  \ref{sec:1D_construction} as priors and use Bayesian inference with data defined below for each example.
+Each take the layer boundaries defined in section \ref{sec:1D_construction} about model construction as priors and use Bayesian inference with data defined below for each example.
 
@@
 
-	\begin{table}
-		\centering
-		\begin{tabular}[c]{| l | l | l |}
-			\hline
-			Risk factor \textit{r} & Shift in Bayes action & Change in expected loss \\ \hline
-			0.50 & +~553.67 & -~105.49 \\ 
-			0.75 & +~673.40 & -~81.67  \\ 
-			1.00 & +~750.05 & -~33.56 \\ 
-			1.25 & +~713.19 & +~75.71 \\ 
-			1.50 & +~0.00 & +~282.64  \\ 
-			\hline
-		\end{tabular}
-		\caption{Bla}
-		\label{tab:update_moderate_tab}
-	\end{table}
+	%\begin{table}
+	%	\centering
+	%	\begin{tabular}[c]{| l | l | l |}
+	%		\hline
+	%		Risk factor \textit{r} & Shift in Bayes action & Change in expected loss \\ \hline
+	%		0.50 & +~553.67 & -~105.49 \\ 
+	%		0.75 & +~673.40 & -~81.67  \\ 
+	%		1.00 & +~750.05 & -~33.56 \\ 
+	%		1.25 & +~713.19 & +~75.71 \\ 
+	%		1.50 & +~0.00 & +~282.64  \\ 
+	%		\hline
+	%	\end{tabular}
+	%	\caption{Bla}
+	%	\label{tab:update_moderate_tab}
+	%\end{table}
 
 	Increased certainty about the reservoir thickness is sufficient to shift Bayes actions to higher estimates for all actors, but the most risk-averse one.
-These shifts are quantified in Table \ref{tab:update_moderate_tab}.
+These shifts are quantified in Table \ref{tab:update_examples_all}.
 According to these numbers, the risk-neutral actor's estimate is increased the most.
@@
 For the seal, a mean of 50~m with a standard deviation of 20~m is chosen, favoring the likelihood of a reliable seal relative to the threshold of 20~m thickness.
-This results in the score distribution depicted in \ref{fig:update_goodseal2}.
+This results in the score distribution depicted in Figure \ref{fig:update_goodseal2}.
 The bulk of the distribution is narrowed on the positive side of estimates.
@@
 
-	Applying the loss function on this new score distribution results in the realizations of expected loss illustrated in \ref{fig:update_goodseal3}.
+	Applying the loss function on this new score distribution results in the realizations of expected loss illustrated in Figure \ref{fig:update_goodseal3}.
 Bayes actions are shifted clearly to higher estimates and expected losses of these minima are significantly reduced for all actors.
-This is quantified in \ref{tab:update_goodseal_tab}.
+This is quantified in Table \ref{tab:update_examples_all}.
 According to these values, the risk-neutral to risk-averse individuals seem to profit the most.
@@
 
-	\begin{table}
-		\centering
-		\begin{tabular}[c]{| l | l | l |}
-			\hline
-			Risk factor \textit{r} & Shift in Bayes action & Change in expected loss \\ \hline
-			0.50 & +~522.95 & -~436.26 \\ 
-			0.75 & +~697.73 & -~551.92  \\ 
-			1.00 & +~855.57 & -~631.92 \\ 
-			1.25 & +~912.43 & -~647.10 \\ 
-			1.50 & +~819.26 & -~545.96  \\ 
-			\hline
-		\end{tabular}
-		\caption{Bla}
-		\label{tab:update_goodseal_tab}
-	\end{table}
+	%\begin{table}
+	%	\centering
+	%	\begin{tabular}[c]{| l | l | l |}
+	%		\hline
+	%		Risk factor \textit{r} & Shift in Bayes action & Change in expected loss \\ \hline
+	%		0.50 & +~522.95 & -~436.26 \\ 
+	%		0.75 & +~697.73 & -~551.92  \\ 
+	%		1.00 & +~855.57 & -~631.92 \\ 
+	%		1.25 & +~912.43 & -~647.10 \\ 
+	%		1.50 & +~819.26 & -~545.96  \\ 
+	%		\hline
+	%	\end{tabular}
+	%	\caption{Bla}
+	%	\label{tab:update_goodseal_tab}
+	%\end{table}
 
@@
 
-	The subsequent score distribution is depicted in \ref{fig:update_smallres2}.
+	The subsequent score distribution is depicted in Figure \ref{fig:update_smallres2}.
 It can be seen that while the whole distribution is narrowed, it also is shifted to the left, to lower and negative estimates.
@@
 
-	Based on this probability distribution of the reservoir score, Bayes actions are shifted to lower estimates for all actors (\ref{fig:update_smallres3}).
+	Based on this probability distribution of the reservoir score, Bayes actions are shifted to lower estimates for all actors (see Figure \ref{fig:update_smallres3}).
 In fact, also but the most risk-friendly individual (\textit{r} = 0.5) find their Bayesian estimator to be zero now, i.e.
@@
 	Quantified changes in the position and the values of the Bayes actions are comparable to the results in example II (there is no shift for risk-averse actors, as they already found their best estimates to be zero).
-Most notable is the large reduction of expected loss in general (see \ref{tab:update_smallres_tab}).
+Most notable is the large reduction of expected loss in general (see Table \ref{tab:update_examples_all}).
 
@@
 
+	%\begin{table}
+	%	\centering
+	%	\begin{tabular}[c]{| l | l | l |}
+	%		\hline
+	%		Risk factor \textit{r} & Shift in Bayes action & Change in expected loss \\ \hline
+	%		0.50 & -~498.98 & -~538.08 \\ 
+	%		0.75 & -~346.16 & -~717.83  \\ 
+	%		1.00 & -~176.13 & -~872.02 \\ 
+	%		1.25 & -~0.00 & -~946.04 \\ 
+	%		1.50 & -~0.00 & -~951.57  \\ 
+	%		\hline
+	%	\end{tabular}
+	%	\caption{Bla}
+	%	\label{tab:update_smallres_tab}
+	%\end{table}
+
 	\begin{table}
 		\centering
-		\begin{tabular}[c]{| l | l | l |}
+		\begin{tabular}[c]{| c || l | l || l | l || l | l |}
 			\hline
-			Risk factor \textit{r} & Shift in Bayes action & Change in expected loss \\ \hline
-			0.50 & -~498.98 & -~538.08 \\ 
-			0.75 & -~346.16 & -~717.83  \\ 
-			1.00 & -~176.13 & -~872.02 \\ 
-			1.25 & -~0.00 & -~946.04 \\ 
-			1.50 & -~0.00 & -~951.57  \\ 
+			& Example I & & Example II & & Example III & \\
+			\hline
+			\hline
+			\textit{r} & $\delta$ BA & $\delta$ EL & $\delta$ BA & $\delta$ EL & $\delta$ BA & $\delta$ EL\\ 
+			\hline
+			0.50 & +~553.67 & -~105.49 & +~522.95 & -~436.26 & -~498.98 & -~538.08 \\ 
+			0.75 & +~673.40 & -~81.67 & +~697.73 & -~551.92 & -~346.16 & -~717.83  \\ 
+			1.00 & +~750.05 & -~33.56 & +~855.57 & -~631.92 & -~176.13 & -~872.02 \\ 
+			1.25 & +~713.19 & +~75.71 & +~912.43 & -~647.10 & -~0.00 & -~946.04 \\ 
+			1.50 & +~0.00 & +~282.64 & +~819.26 & -~545.96 & -~0.00 & -~951.57  \\ 
 			\hline
@@
 		\caption{Bla}
-		\label{tab:update_smallres_tab}
+		\label{tab:update_examples_all}
 	\end{table}

### 2017-05-13 Risk analysis/Text/Layout/Layout/chapter3_1Dcase.tex
Block data from Miguel
Text updates

@@
 
-	\subsection{Updating example I (moderate reinforcement)- Seal: mean = 25~m, std = 20~m - Reservoir: mean = 180~m, std = 60~m}
+	\subsection{Updating example I: Moderately reinforcing information}
 
@@
 		\includegraphics[width=1\textwidth]{Figures/update_moderate1.png}
-		\caption{...}\label{fig:update_moderate1} 
+		\caption{Prior (A1) and posterior distributions (A2) of the layer boundary positions in depth and respective representations (A2, B2).}\label{fig:update_moderate1} 
 	\end{figure}
@@
 	Modeling the reservoir score is conducted as above, now based on these new distributions.
-In Figure \ref{fig:update_moderate2} it can be recognized that the bulk of the distribution was shifted to the positive side of scores, while the peak at -~2000 was raised.
+In Figure \ref{fig:update_moderate2_3} A it can be recognized that the bulk of the distribution was shifted to the positive side of scores, while the peak at -~2000 was raised.
 The probability of scores between -~2000 and 0 decreased significantly, i.e.
@@
 	\begin{figure}[h]
+		\begin{subfigure}{1\textwidth}
 			\centering
-		\includegraphics[width=1\textwidth]{Figures/update_moderate2.png}
-		\caption{...}\label{fig:update_moderate2} 
-	\end{figure}
-
-	\begin{figure}[h]
+			\includegraphics[width=1\linewidth]{Figures/update_moderate2.png}
+			%\caption{1a}
+			%\label{fig:sfig1}
+		\end{subfigure}%
+		\\
+		\begin{subfigure}{1\textwidth}
 			\centering
-		\includegraphics[width=1\textwidth]{Figures/update_moderate3.png}
-		\caption{...}\label{fig:update_moderate3} 
+			\includegraphics[width=1\linewidth]{Figures/update_moderate3.png}
+			%\caption{1b}
+			%\label{fig:sfig2}
+		\end{subfigure}
+		\caption{Reservoir score distributions (A) and change in the realizations of expected loss for several risk parameters (B1, B2) before and after Bayesian updating based on likelihoods defined as follows: Seal thickness mean = 25~m, std = 20~m.
+Reservoir thickness mean = 180~m, std = 60~m.}
+		\label{fig:update_moderate2_3}
 	\end{figure}
@@
 	The customized loss function (IV) is now applied on this updated distribution of the reservoir score.
-This is visualized in Figure \ref{fig:update_moderate3} in which the expected losses are compared before (B1 )and after (B2) uncertainty reduction.
+This is visualized in Figure \ref{fig:update_moderate2_3} in which the expected losses are compared before (B1) and after (B2) uncertainty reduction.
 It is observable, that by adding information about layer thickness likelihoods, Bayes actions are shifted relative to the nature of the information.
@@
 Information on the seal, however, based on a normal distribution around 25~m thickness, leaves uncertainty about the reliability of the seal, as the safety threshold is defined as 20~m.
-
 	%\begin{table}
@@
 
-	\subsection{Updating example II (reliable seal) - Seal: mean = 50 m, std = 20 m - Reservoir: mean = 180 m, std = 60 m}
+	\subsection{Updating example II: Likely reliable seal}
 
@@
 For the seal, a mean of 50~m with a standard deviation of 20~m is chosen, favoring the likelihood of a reliable seal relative to the threshold of 20~m thickness.
-This results in the score distribution depicted in Figure \ref{fig:update_goodseal2}.
+This results in the score distribution depicted in Figure \ref{fig:update_goodseal2_3}.
 The bulk of the distribution is narrowed on the positive side of estimates.
@@
 
-	\begin{figure}[h]
-		\centering
-		\includegraphics[width=1\textwidth]{Figures/update_goodseal2.png}
-		\caption{...}\label{fig:update_goodseal2} 
-	\end{figure}
-
-	Applying the loss function on this new score distribution results in the realizations of expected loss illustrated in Figure \ref{fig:update_goodseal3}.
+	Applying the loss function on this new score distribution results in the realizations of expected loss illustrated in Figure \ref{fig:update_goodseal2_3}.
 Bayes actions are shifted clearly to higher estimates and expected losses of these minima are significantly reduced for all actors.
@@
 	\begin{figure}[h]
+		\begin{subfigure}{1\textwidth}
 			\centering
-		\includegraphics[width=1\textwidth]{Figures/update_goodseal3.png}
-		\caption{...}\label{fig:update_goodseal3} 
+			\includegraphics[width=1\linewidth]{Figures/update_goodseal2.png}
+			%\caption{1a}
+			%\label{fig:sfig1}
+		\end{subfigure}%
+		\\
+		\begin{subfigure}{1\textwidth}
+			\centering
+			\includegraphics[width=1\linewidth]{Figures/update_goodseal3.png}
+			%\caption{1b}
+			%\label{fig:sfig2}
+		\end{subfigure}
+		\caption{Reservoir score distributions (A) and change in the realizations of expected loss for several risk parameters (B1, B2) before and after Bayesian updating based on likelihoods defined as follows: Seal thickness mean = 50~m, std = 20~m.
+Reservoir thickness mean = 180~m, std = 60~m.}
+		\label{fig:update_goodseal2_3}
 	\end{figure}
@@
 
-	\subsection{Updating example III (safe seal, but reservoir under average) - Seal: mean = 70 m, std = 10 m - Reservoir: mean = 100 m, std = 40 m}
+	\subsection{Updating example III: Safe seal but likely subpar reservoir thickness}
 
@@
 However, the new observations are assumed to provide information about the reservoir that makes it likely to be thinner as expected.
-Reservoir thickness likelihood is assigned a mean of 100~m and a standard deviation of 10~m.
+Reservoir thickness likelihood is assigned a mean of 100~m and a standard deviation of 40~m.
 
-	The subsequent score distribution is depicted in Figure \ref{fig:update_smallres2}.
+	The subsequent score distribution is depicted in Figure \ref{fig:update_smallres2_3}.
 It can be seen that while the whole distribution is narrowed, it also is shifted to the left, to lower and negative estimates.
@@
 
-	Based on this probability distribution of the reservoir score, Bayes actions are shifted to lower estimates for all actors (see Figure \ref{fig:update_smallres3}).
+	Based on this probability distribution of the reservoir score, Bayes actions are shifted to lower estimates for all actors (see Figure \ref{fig:update_smallres2_3}).
 In fact, also but the most risk-friendly individual (\textit{r} = 0.5) find their Bayesian estimator to be zero now, i.e.
@@
 	\begin{figure}[h]
+		\begin{subfigure}{1\textwidth}
 			\centering
-			\includegraphics[width=1\textwidth]{Figures/update_smallres2.png}
-			\caption{...}\label{fig:update_smallres2} 
-	\end{figure}
-
-	\begin{figure}[h]
+			\includegraphics[width=1\linewidth]{Figures/update_smallres2.png}
+			%\caption{1a}
+			%\label{fig:sfig1}
+		\end{subfigure}%
+		\\
+		\begin{subfigure}{1\textwidth}
 			\centering
-				\includegraphics[width=1\textwidth]{Figures/update_smallres3.png}
-				\caption{...}\label{fig:update_smallres3} 
+			\includegraphics[width=1\linewidth]{Figures/update_smallres3.png}
+			%\caption{1b}
+			%\label{fig:sfig2}
+		\end{subfigure}
+		\caption{Reservoir score distributions (A) and change in the realizations of expected loss for several risk parameters (B1, B2) before and after Bayesian updating based on likelihoods defined as follows: Seal thickness mean = 70~m, std = 10~m.
+Reservoir thickness mean = 100~m, std = 40~m.}
+		\label{fig:update_smallres2_3}
 	\end{figure}
@@
 			\hline
-			\textit{r} & $\delta$ BA & $\delta$ EL & $\delta$ BA & $\delta$ EL & $\delta$ BA & $\delta$ EL\\ 
+			\textit{r} & $\Delta$ BA & $\Delta$ EL & $\Delta$ BA & $\Delta$ EL & $\Delta$ BA & $\Delta$ EL\\ 
 			\hline
@@
 		\end{tabular}
-		\caption{Bla}
+		\caption{Changes in Bayes action (BA) and minimal expected loss (EL) for Bayesian updating examples I, II and III and respective actors with risk parameters \textit{r}.}
 		\label{tab:update_examples_all}
@@
 	\section{General case results}
+
+	-abstract case: easy model construction and straightforward design of a loss function making basic assumptions and taking relative values that can simply be exchanged
+
+	-so representative in a "relative" way, mostly appropriate to illustrate principles and benefits of the methodology
+
+	-decision making/ estimation is defined by the design of the loss function which includes framework parameters which depend on the problem environment (e.g. market, technical constraints, etc.)
+
+	-similar actors but different behaviors concerning risk: different decisions but same general loss function path (just different minima of EL)
+
+	-additional information changes decision (BA) and EL
+
+	-magnitude of change varies not only with the nature of the information (magnitude of uncertainty reduction) but also with the risk parameter
+
+	-some actors might benefit more from some specific additional information than others
+
+	-quantifiable value of information??

### 2017-08-09 Risk analysis/Text/Layout/Layout/chapter4_3Dcase.tex
Chaos - nothing is working

\chapter{Three-dimensional reservoir case}

	\section{Constructing the three-dimensional structural geological model}
	- place layer positional points in three lines, resembling three hypothetical seismic lines as base information
	-

	\section{Identifying trap structures in the geological model}

### 2017-08-29 Risk analysis/Text/Layout/Layout/chapter4_3Dcase.tex
Progress model and text

@@
 \chapter{Three-dimensional reservoir case}
+- concept of 1D case in chapter 3 transferred to a 3D case of a structural geological model
+- same again here: relatively simple model including some important typical geological and structural features and giving it an economic meaning
+- alternating layers of sandstones and shales, so typical reservoir and seal formations
+- structurally defined by a tilted dome-formation (folding along in the direction of two perpendicular axes) which is displaced by a standard normal fault
+- these features are arranged in a way, that they result in the formation of a trap, defined by the anticlinal features and the fault, as well as enabled by the presence of a reservoir rock
+- considering a petroleum production industry perspective, actors/ decision makers would be interested in the evaluation of value this potential reservoir
+- "traditionally", this is represented by volumetric calculations or estimations of the recoverable reserves
+- here: maximum volume of entrapment and subsequently, relative or hypothetical volumes of recoverable reserve can be calculated (for each model realization)
+- (later: numbers from example case are taken, to achieve estimations of the NPV for a development project realization)
+- looking at this structural model, not only the calculated volume can be taken into account, but also risk factors, such as fault permeabity seal safety (and capacity?), possible juxtapositions and leaks
+
+- as in the 1D case, uncertainties are introduced for a number of parameters in the model, in a way that they affect the resulting maximum volume of the trap (and following also the recoverable reserves)
+- again, a specific loss function is designed and applied on the results from numerous iterations of the constructed uncertain model, thus resembling the different value estimations/predictions or decisions different actors might make
+- (to bring this Bayesian method of using loss functions closer to the traditional application of deterministic decision-making in decision-trees, the continuous estimation space of the reservoir value (volume) is combined with the assumption, that the estimated value is linked to a consequent investment into a determined project development size.
+thus, the continuous approach is used to make the decision in a deterministic decision-making space.
+))
+
+- later on,  the effect of adding information (in the form of likelihoods) is examined, particularly by looking at the changes in decision-making after applying this Bayesian updating and uncertainty has been reduced
+
+- in the following, the construction and design of this model is described
 
 	\section{Constructing the three-dimensional structural geological model}
+	- base of the structural geological model is a 3Dcubic  space (block)  with 2000 m size in X, Y and Z x3
 	- place layer positional points in three lines, resembling three hypothetical seismic lines as base information

### 2017-08-31 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Text

@@
     The methods utilized in this work are presented in the following chapter.
-This includes introductions into Bayesian analysis and decision theory.
-In particular Bayesian inference, estimation of uncertain values and how loss function can be used in this context.
-These methods are incorporated in the context of numerical modeling of structural geological settings.
-This is done by programming in a python environment.
-Central tools for model construction and conduction of the statistical evaluation are Geomodeler3D, GemPy and PyMC3 in particular and are also presented in this chapter.
+Bayesian analysis and decision theory are introduced.
+Focus is laid on Bayesian inference, estimation of uncertain values and the use of loss functions in this context.
+These methods are incorporated in computational modeling of structural geological settings by programming in a python 3 environment.
+Central tools for model construction and conduction of the statistical evaluation are GeMpy and PyMC 2 in particular.
+These are also presented in this chapter.
 
         \section{Bayesian analysis and decision theory}
-	    Problems and reasoning behind decision making are examined in the field of decision theory, as implied by the name \citep{berger2013stat}.
+	    As implied by the name, the problems and reasoning behind decision-making are examined in the field of decision theory \citep{berger2013stat}.
 Such decision problems are commonly influenced by parameters that are uncertain.
 In statistical decision theory, available statistical knowledge is used to gain information on the nature of these uncertainties.
-(Such uncertain parameters can be considered as numerical quantities.) In order to find the best decision to a problem, it is possible to combine sample information with other aspects such as the possible consequences of decision making and the availability of prior information on our uncertainties.
+(Such uncertain parameters can be considered as numerical quantities.) In order to find the best decision to a problem, it is possible to combine sample information with other aspects such as the possible consequences of decision-making and the availability of prior information on our uncertainties.
 Decision consequences are expressed as gains in economic decision theory and as losses, which equal negative gains, in statistics.
-Prior information might be given for example due to experience from previous similar problems or from expert knowledge (see Batvold and Begg).
-The approach of utilizing priors is known as Bayesian analysis and is explained in the following \citep{berger2013stat}.
+Prior information might be given for example based on experience from previous similar problems or from expert knowledge (see Batvold and Begg).
+The approach of utilizing priors is known as Bayesian analysis, which is explained in the following \citep{berger2013stat}.
 (It goes well with decision theory.)
 
-	    \section{Basic elements}
-	    Some basic elements are to be defined.
-The unknown (uncertain) quantity influencing decision making is usually denominated as $\theta$, the state of nature \citep{berger2013stat}.
+	    \subsection{Basic elements}
+	    First, some basic elements are to be defined.
+The unknown (uncertain) quantity influencing decision-making is usually denominated as the state of nature $\theta$ \citep{berger2013stat}.
 Given statistical information on $\theta$ in the form of probability distributions, $\theta$ is called the parameter.
@@
 	    The outcome of statistical tests in form of information or statistical evidence is denoted as $X$.	    
-	    Loss is defined as $L(\theta,a)$, so $L(\theta_1,a-1)$ is the actual loss incurred when action $a_1$ is taken while the true state of nature is $\theta_1$ \cite{pberger2013stat}.
-Loss, expected loss and loss functions are further explained below.
+	    Loss is defined as $L(\theta,a)$, so $L(\theta_1,a_1)$ is the actual loss incurred when action $a_1$ is taken while the true state of nature is $\theta_1$ \citep{berger2013stat}.
+Loss, expected loss and loss functions are explained in detail further below.
 
@@
 Probability is seen as a measure of belief for an event to occur.
-It has been argued by \citep{davidson2015}  that this Bayesian approach is intuitive and inherent in the natural human perspective.
+It has been argued by \cite{davidson2015}  that this Bayesian approach is intuitive and inherent in the natural human perspective.
 These beliefs can be assigned to individuals \citep{davidson2015}.
@@
 
+        - Bayes Theorem in geological modeling (see Miguel)
+
         -Law of Large Numbers!
 
-        -MCMC sampling methods!
+        \subsection{Markov Chain Monte Carlo sampling (MCMC)}
+        Despite the apparent simplicity of the Bayes Theorem, a direct analytical calculation and exact inference of the posterior distribution P(A$|$X) is rarely possible in non-idealized cases, due to intractability in high-dimensional spaces \citep{homan2014no, delaVarga2016}.
+
+        - how to obtain posterior parameter distribution
+        - intractability of multi-dimensional space makes analytical calculation impossible (directly) --> need for MCMC
+        - MCMC: exploration of target distribution by sampling, then accepting or rejecting
+        - here used: adaptive Metropolis
 
@@
 
-        So, "how bad" a current estimate is, depends on the way a loss function weights accuracy errors and returns respective losses.
+        So, "how bad" a current estimate is, depends on the way a loss function weighs accuracy errors and returns respective losses.
 Two standard loss functions are the absolute-error and the squared-error loss function.
@@
 
-        As implied by its name, the absolute-error loss function returns loss as the absolute error, i.e. difference between the estimate and the true parameter \citep{davidson2015}:
+        As implied by its name, the absolute-error loss function returns loss as the absolute error, i.e. the difference between the estimate and the true parameter \citep{davidson2015}:
 
@@
 
-        \citet{davidson2015} and (LF DESIGN PAPER?) propose that it might be useful to move away from these type of objective loss functions to the design of new loss functions that are specifically customized to an reflect an individual's, i.e.
-the decision maker's objectives, preferences and outcomes.
+        \citet{davidson2015} and \citet{hennig2007} propose that it might be useful to move away from these type of objective loss functions to the design of customized loss functions that specifically reflect an individual's (i.e.
+the decision maker's) objectives, preferences and outcomes.
 

### 2017-08-31 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
trying to write MCMC part

@@
 
-        \subsection{Markov Chain Monte Carlo sampling (MCMC)}
-        Despite the apparent simplicity of the Bayes Theorem, a direct analytical calculation and exact inference of the posterior distribution P(A$|$X) is rarely possible in non-idealized cases, due to intractability in high-dimensional spaces \citep{homan2014no, delaVarga2016}.
+        \subsection{Markov chain Monte Carlo sampling (MCMC)}
+        Despite the apparent simplicity of the Bayes Theorem, a direct analytical calculation and exact inference of the posterior distribution P(A$|$X) is rarely possible in non-idealized cases, due to intractability in multi-dimensional spaces \citep{hoffman2014no, delaVarga2016}.
+Thereby arises the necessity to resort to methods of statistical inference approximation.
+Markov chain Monte Carlo (MCMC) sampling has proven to be a generally applicable and reliable method for exploring multi-dimensional parameter spaces in an intelligent way \citep{hoffman2014no, davidson2015}.
+\citet{gilks2005mcmc} has emphasized the significance of MCMC for the application of Bayesian statistics.
+        %In Monte Carlo integration, random samples are drawn from a target distribution, such as our posterior distribution P(A$|$X), to approximate the integral or probability density function in terms of an expectation:
 
-        - how to obtain posterior parameter distribution
-        - intractability of multi-dimensional space makes analytical calculation impossible (directly) --> need for MCMC
-        - MCMC: exploration of target distribution by sampling, then accepting or rejecting
-        - here used: adaptive Metropolis
+        %\begin{equation}\label{eq:MonteCarlo}
+        %        E[P(A|X)] \approx \frac{1}{N}\sum_{n=1}^{N}P(A|X)
+        %\end{equation}
+        In the ordinary Monte Carlo approach, random independent samples are drawn from a target distribution in order to approximate its shape \citep{gilks2005mcmc}.
+High-dimensional spaces as found in Bayesian applications lead to complex shapes and often make independent sampling infeasible \citep{gilks2005mcmc}.
+This can be solved by extending the Monte Carlo principle with a Markov chain, in which every sample iteration of the parameter $\theta^i$ is dependent uniquely on the previous value $\theta^{(i-1)}$ \citep{gilks2005mcmc, delaVarge2016}.
+        Various algorithms for random MCMC walks have been developed over more than six decades and are still an active field of research.
+Common examples are the Metropolis-Hastings (devised by \citet{metropolis1953equation} and continued by \citet{hastings1970}) and the Gibbs sampler \citep{geman1984stochastic}.
+        In this work, an adaptive Metropolis-Hastings sampling method is used.
 

### 2017-09-04 Risk analysis/Text/Layout/Layout/MSc_layout_Fabian.tex
Big update: Chapter 3! Some figures in notebooks. New model exports.
Text updates
Chaos - nothing is working
Text
Text Methods Progress

@@
 %\usepackage{pifont}
+\usepackage{graphicx}
+\usepackage{natbib}
+\usepackage{subcaption}
+\graphicspath{Figures/}
+%\usepackage{mathpazo}
 \mscName{Fabian Antonio Stamm}
@@
 
-%
-% First Part
-    \part{First Part} % you can divide your thesis into different parts, for example Theory&Modeling in part 1 and real data examples in part 2.
 		\include{chapter2_methods}
 		\include{chapter3_1Dcase}
+		\include{chapter4_3Dcase}
 
@@
     \bibliographystyle{apalike}
-    \printbib{complete_biblio.bib} % in MyBib.bib you add all your reference information, following the correct format.
+    \printbib{complete_biblio} % in MyBib.bib you add all your reference information, following the correct format.
     % The *.bin file needs to be compiled using the Bib-compile option of your LaTeX text editor.

### 2017-09-04 Risk analysis/Text/Layout/Layout/chapter1.tex
Text Methods Progress

@@
 
-    Welcome to the standard layout for your IDEA LEAGUE MSc thesis written in \LaTeX.
-\LaTeX\  has a variety of advantages over conventional/ standard text editing programs, which you will soon enough discover yourself.
-\LaTeX\  almost forms a standard in the Scientific Community, especially due to its effective and straightforward mathematical capabilities.\\
-    This is Chapter\ \ref{chap::intro}. If you want to know more about \LaTeX\ you better read
-    \cite{texbook} or use the extensive help available on the internet.
-\index{LaTeX}.
-This 'hidden' index command helps you making an index at the end of your thesis.
-You can add this flag anywhere you want to make an index hit.
-You can see here also how to use acronyms, like \ac{DUT}.
-The acronyms are automatically listed in the corresponding section.
-Also, hyperlinks are created automatically with the developed class file, such that your digital PDF version of your thesis can be read dynamically.
- Have fun with \LaTeX\ and your M.Sc. research project and good luck! \\ \\
+    - mention developments in geological modeling
 
-The purpose of the introduction is to tell readers why they should want to read what follows the introduction.
-This chapter should provide sufficient background information to allow readers to understand the context and significance of the problem.
-This does not mean, however, that authors should use the introduction to rederive established results or to indulge in other needless repetition.
-The introduction should (1) present the nature and scope of the problem; (2) review the pertinent literature, within reason; (3) state the objectives; (4) describe the method of investigation; and (5) describe the principal results of the investigation.
+    - economic significance of geological modeling: examples of application in petroleum, mining, aquifers, CCS, etc.
+
+    - Bayesian analysis developments
+
+    - decision-making in for such economic scenarios and projects which involve geosciences
+
+    - classical approach (decision trees) vs. Bayesian approach, expression of decision-maker's interests in continuous loss functions
+
+    - influence of uncertainty and work done by Wellmann & DeLaVarga

### 2017-09-04 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Text Methods Progress

--- 2017-08-31 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
@@
 
-        - Bayes Theorem in geological modeling (see Miguel)
-
         -Law of Large Numbers!
-
-        \subsection{Markov Chain Monte Carlo sampling (MCMC)}
-        Despite the apparent simplicity of the Bayes Theorem, a direct analytical calculation and exact inference of the posterior distribution P(A$|$X) is rarely possible in non-idealized cases, due to intractability in high-dimensional spaces \citep{homan2014no, delaVarga2016}.
-
-        - how to obtain posterior parameter distribution
-        - intractability of multi-dimensional space makes analytical calculation impossible (directly) --> need for MCMC
-        - MCMC: exploration of target distribution by sampling, then accepting or rejecting
-        - here used: adaptive Metropolis
 
@@
         Accordingly, losses increasing linearly with the distance to the true value are returned for respective estimates.
+This means that all differences between relative errors are weighed equally, no matter whether they are found in the realm of relatively small or relatively large errors \citep{hennig2007}.
 
@@
 
-        This exponential growth of loss also means that large errors are weighted much stronger than small errors.
+        This exponential growth of loss also means that large errors are weighed much stronger than small errors.
 This might come with over-valuation of distant outliers and misrepresentation of magnitudes in distance.
-Regarding this, the absolute-error loss function can be seen as more robust \citep{davidson2015}.
-
-        - each give mean or median --> estimator with accuracy metric? objective and symmetric! aim at high precision.
-
-        -shift in perspective. look at outcome instead of estimation precision!
-
+Regarding this, the absolute-error loss function can be seen as more robust \citep{davidson2015}.\\
+        Both of these standard loss functions are symmetric and can be described as objectively aiming at a high precision in estimating the true parameter value.
         \citet{davidson2015} and \citet{hennig2007} propose that it might be useful to move away from these type of objective loss functions to the design of customized loss functions that specifically reflect an individual's (i.e.
 the decision maker's) objectives, preferences and outcomes.
-
-        The standard loss function defined above are symmetric, but can easily be adapted to be asymmetric, for example by weighting on the negative side stronger than those on the positive side.
-Preference over estimates larger than the true value is thus incorporated in an uncomplicated way \citep{davidson2015}.
+\citet{hennig2007} argue that choosing and designing a loss function involves the translation of informal aims and interests into mathematical terms.
+This process naturally implies the integration of subjective decisions and subjective elements.
+According to \citet{hennig2007}, this is not necessarily unfavorable or "less objective", as it may better reflect an expert's perspective on the situation and contribute to a productive scientific discussion.\\
+        The standard loss functions defined above are symmetric, but can easily be adapted to be asymmetric, for example by weighing errors on the negative side stronger than those on the positive side.
+Preference over estimates larger than the true value (i.e.
+overestimation) is thus incorporated in an uncomplicated way \citep{davidson2015, hennig2007}.
 Much more complicated designs of loss functions are possible, depending on purpose, objective and application \citep{davidson2015}.
 Some case-specific loss functions are designed in the following chapters of this work.
-
-        - Cite from paper about the design of loss functions! Subjectivity of customized loss functions!
-
+\\
         The presence of uncertainty during decision making implies that the true parameter is unknown and thus the truly incurred loss $L(\theta,a)$ cannot be known at the time of making the decision \citep{berger2013stat, davidson2015}.
@@
 A suitable alternative to the actual loss is to consider a decision's expected loss and to make a decision that is optimal in relation to this expected loss \citep{berger2013stat}.
-
+\\
         Given a posterior distribution $P(\theta|X)$, the expected loss of choosing an estimate $\hat{\theta}$ over the true parameter $\theta$ (after evidence X has been observed) is defined by the function below \citep{davidson2015}:
@@
         The expectation symbol $E$ is subscripted with $\theta$, by which it is indicated that $\theta$ is the respective unknown variable.
-This expected loss as defined above, is also referred to the risk of estimate $\hat{\theta}$ \citep{davidson2015}.
+This expected loss as defined above, is also referred to as the risk of estimate $\hat{\theta}$ \citep{davidson2015}.
 
-        By the Law of Large Numbers, the expected loss of $\hat{\theta}$ can be approximated drawing $N$ samples from the posterior distribution, respectively applying a loss function $L$ and averaging of the number of samples \citep{davidson2015}:
+        By the Law of Large Numbers, the expected loss of $\hat{\theta}$ can be approximated drawing a large sample size $N$ from the posterior distribution, respectively applying a loss function $L$ and averaging of the number of samples \citep{davidson2015}:
 
@@
 
-        Minimization of \ref{eq:ExpectedLoss} or \ref{eq:ExpectedLoss2} returns a Bayesian point estimate known as Bayes action $\delta^P(X)$ (???), which is the estimate, action or decision with the least expected loss according to the loss function \citep{berger2013stat}.
+        Minimization of a loss function returns a Bayesian point estimate known as Bayes action $\delta^P(X)$, which is the estimate, action or decision with the least expected loss according to the loss function \citep{berger2013stat}.
 For a unimodal and symmetric absolute-error loss function, the Bayes action is simply the median of the posterior distribution, while squared-error loss it is the mean \citep{davidson2015, berger2013stat}.
-The MAP estimate is returned when using zero-one loss \citep{davidson2015}  (WRITE THIS?).
-The possibility of more than on minimum also implies that several Bayes actions can exist for one problem \citep{berger2013stat}.
-
-        - MAP?
-
+The MAP (maximum a posteriori) estimate is the minimizing solution for the posterior using zero-one loss \citep{davidson2015}.
+The possibility of more than on minimum also implies that several Bayes actions can exist for one problem \citep{berger2013stat}.\\
         \citet{davidson2015} implemented different risk affinities by simply  introducing a risk parameter into the loss function.
 By using different values for this parameter, it can be represented how comfortable an individual is with being wrong and furthermore which "side of wrong" is preferred by this decision maker \citep{davidson2015}.
-In the chapter below, different risk parameters are introduced to alter the weighting in the loss functions (???).
+This approach to expressing risk-affinities is used the design of loss functions in this work.
 
-        \subsection{Value of information}
-        - considering the change in gain (or loss) after regarding new information or evidence, the value of this information can be calculated
-        - in the presence of uncertainty we look at expected value of information
-        - this can be used as a measure, to assess if the effort to attain new evidence or information is worth it
-        - also as a measure to see how much additional information is worth to different actors with different risk affinities
-        - here we use the comparison between Bayes actions before and after Bayesian updating
+        %\subsection{Value of information}
+        %- considering the change in gain (or loss) after regarding new information or evidence, the value of this information can be calculated
+        %- in the presence of uncertainty we look at expected value of information
+        %- this can be used as a measure, to assess if the effort to attain new evidence or information is worth it
+        %- also as a measure to see how much additional information is worth to different actors with different risk affinities
+        %- here we use the comparison between Bayes actions before and after Bayesian updating
 
         \section{Application in structural geological modeling}
+        In this work, these methods of Bayesian analysis and decision theory are applied in the field of geological structural modeling.
+The fundamental approach follows closely the research conducted by \citet{delaVarga2016} and builds upon their findings.\\
+        According to them, structural geological modeling can be regarded as a forward problem and the elements of Bayesian inference can be specified in this context as follows:
+        \begin{enumerate}
+        	\item \textbf{Mathematical forward model ($M$)}: The connections between parameters $\theta$ and observed data $y$ are defined in this mathematical model. ...
+        	\item \textbf{Model parameters ($\theta$)}: These model-defining parameters, such as the position or dip of layers and faults in geological settings, can be deterministic or stochastic.
+In the latter case, they are uncertain parameters to which a probability distribution is assigned.
+        	\item \textbf{Observed data ($y$)}: This is any type of additional information that can be related to the forward modeling results and might possibly be used to reduce uncertainty.
+Such data can be gained by measurements and observations, for example by core sampling, well-log analysis or seismic acquisition.
+        	\item \textbf{Likelihood functions ($p(y|\theta)$}: Links between the previous parameters $\theta$ and the additional data $y$ are established by these functions in way that they reflect the likelihood of the parameter states given the observations \citep{delaVarga2016}.
+        \end{enumerate}
 
-        - applying these statistical theories in the context of geological modeling 
-        - as done by De la Varga (2016)
+        A fundamental sequence of the inference process was proposed by \citet{gelman2014bayesian}, adapted by \citet{delaVarga2016} and is subsequently adjusted for the application in this work as follows:
+        \begin{enumerate}
+        	\item \textbf{Setting up a full probability model}: A multi-dimensional joint probability space is to be generated, taking into account the probability distributions of every model parameter $\theta$.
+(- here, this not only results in the creation of geological models, but is subsequently also coupled with algorithms which deduct economic realizations from these models.)
+        	\item \textbf{Conditioning on observed data}: Subsequently, an appropriate posterior distribution $p(\theta|y)$ is to be calculated by conditioning the parameters $\theta$ on the observed data $y$ given the likelihood $p(\theta|y)$.
+This is the step of Bayesian updating of the belief about the parameter uncertainty given new information.
+        	In a chosen model ($M$), this is achieved by linking parameters and data through deterministic operations which are additionally compared to the likelihood functions.
+It is pointed out by \citet{delaVarga2016} that any combination of parameter-observation connections is allowed, i.e.
+not all parameters need to be necessarily connected to all observed data.
+After all conditional probabilities have been set up, the Bayes Theorem (Equation \ref{eq:BayesTheorem}) is applied to attain the posterior.
+However, due to the multi-dimensionality given in geological problems, the use of Markov chain Monte Carlo methods is advised to achieve this as described in Section \ref{sec:mcmc} below.
+        	\item \textbf{Evaluation of the posterior model}: Depending on the aim of the study, a post-processing analysis can be conducted accordingly.
+\citet{delaVarga2016} focused on the examination of the posterior distributions of the parameters $\theta$ and the generated models, particularly regarding information entropy within the model space.
+In this work, the geological models are assigned an economic meaning by declaring them potential petroleum reservoirs and introducing customized loss functions to reflect the economic interest of decision makers in developing respective resource extraction projects (see Section \ref{sec:loss}).
+Shifts in Bayes actions are considered a measure for the influence of Bayesian inference on decision-making and the significance of additional observations for different decision makers.
+        \end{enumerate}
 
-        - equivalence of basic elements in geologic models, some examples
+        \subsection{Markov chain Monte Carlo sampling (MCMC)}\label{sec:mcmc}
+        Despite the apparent simplicity of the Bayes Theorem, a direct analytical calculation and exact inference of the posterior distribution P(A$|$X) is rarely possible in non-idealized cases, due to intractability in multi-dimensional spaces \citep{hoffman2014no, delaVarga2016}.
+Thereby arises the necessity to resort to methods of statistical inference approximation.
+Markov chain Monte Carlo (MCMC) sampling has proven to be a generally applicable and reliable method for exploring multi-dimensional parameter spaces in an intelligent way \citep{hoffman2014no, davidson2015}.
+\citet{gilks2005markov} has emphasized the significance of MCMC for the application of Bayesian statistics.
+        %In Monte Carlo integration, random samples are drawn from a target distribution, such as our posterior distribution P(A$|$X), to approximate the integral or probability density function in terms of an expectation:
 
-        - sampling methods
+        %\begin{equation}\label{eq:MonteCarlo}
+        %        E[P(A|X)] \approx \frac{1}{N}\sum_{n=1}^{N}P(A|X)
+        %\end{equation}
+        In the ordinary Monte Carlo approach, random independent samples are drawn from a target distribution in order to approximate its shape \citep{gilks2005markov, delaVarga2016}.
+High-dimensional parameter spaces as found in Bayesian applications lead to complex shapes and often make independent sampling infeasible \citep{gilks2005markov}.
+This can be solved by extending the Monte Carlo principle with a Markov chain, in which every sample iteration of the parameter $\theta^{(t+1)}$ is dependent uniquely on the previous value $\theta^{(t)}$ \citep{gilks2005markov, delaVarga2016}.\\
+        The general principle of MCMC can be described as follows: 
+        Drawing representative samples from an target distribution of unknown shape is based on the conduction of a so-called random walk on the parameter distribution space.
+T sampling steps are to be performed.
+The first sampling location is chosen at random.
+With each subsequent step, a new sampling location is proposed.
+The new sample value is then related to the previous step.
+According to a weight defined by the scaled up candidate density of the value, the proposed step is then accepted or rejected.
+In the case of acceptance, the value is added to the sample trace and the process is continued from the current location.
+In the case of rejection, sampling is reverted to the previous accepted step \citep{schaaf2017, delaVarga2016}.
+The intention behind this concept it to achieve convergence of the sampling algorithm towards areas of high probability \citep{davidson2015}.\\ Variations in the way of how new  sample steps are proposed and in the acceptance-rejection condition result in different single MCMC sampling methods \citep{schaaf2017, delaVarga2016}.
+Various algorithms for random MCMC walks have been developed for over more than six decades and advancements have still been made in recent years.
+Common examples for such algorithms are the Metropolis-Hastings samplers as devised by \citet{metropolis1953equation} and generalized by \citet{hastings1970}.
+The Gibbs sampler \citep{geman1984stochastic} is another well-known method.
+        For the purspose of this work, an adaptive Metropolis-Hastings sampler is used.\\
+        In Metropolis-Hastings methods, each sampling step at iteration t is determined by a candidate probability distribution $q(\theta,\theta^{'})$, from which a proposed sample $\theta^{'}$ is drawn \citep{delaVarga2016}.
+The acceptance-rejection condition is defined by the acceptance ratio $a(\theta^{'},\theta)$:
 
-        - numerical implementation in python using pyMC
+        \begin{equation}\label{eq:acceptance_ratio}
+        a(\theta^{'},\theta)=\frac{p(\theta^{'})p(y|\theta^{'})}{p(\theta)p(y|\theta)}
+        \end{equation}
 
-        - construction of 1D model in python and pyMC
+        To ensure a thorough exploration of the probability space, the transition to higher probability densities should not be enforced in every case but selectively.
+This is assured by relating the acceptance ratio from Equation \ref{eq:acceptance_ratio} to a random value $u$ from a Uniform distribution $U(0,1)$ as follows:
 
-        - construction of 3D model using GeoModeller3D and GeMPy additionally
+        \begin{equation}\label{eq:transition_selectivity}
+        \theta^{(t+1)}=\begin{cases}
+        \theta^{'} & $if $ a(\theta^{'},\theta)>U(0,1)\\
+        \theta^{t} & $otherwise$\\
+        \end{cases}
+        \end{equation}
 
-        - what about theory of petroleum reservoir and volumetric calculation? explain this in later chapter of 3D model???
+        Thereby, the algorithm assigns high probabilities to high-density points and low probabilities to low-density points, so that the chain state is moved accordingly \citep{delaVarga2016}.\\
+        Metropolis methods are furthermore defined by the step size scale factor that is chosen.
+While large steps are good for exploration of the space and mixture in the chain, acceptance rates are low.
+Small steps have better acceptance rates, but lead to slower exploration and convergence of the algorithm \citep{delaVarga2016}.\\
+        The Adaptive Metropolis (AM) by \citet{haario2001adaptive} is used in this work.
+It adapts the traditional Metropolis-Hastings by incorporating the ability of continuous step size tuning during convergence, by taking into account the full information saved along the process.
+This is achieved by generating a covariance matrix that is updated every iteration.
+The adaptive nature of the process enables fast convergence for non-linear distributions while maintaining ergodicity \citet{haario2001adaptive, delaVarga2016}.
+Its suitability for multi-dimensional distribution spaces make it an excellent method for dealing with complex models such as the structural geological models in this work \citep{schaaf2017}
 
-		\subsection{Implementing Bayesian analysis numerically with Python and PyMC3}
+        \subsection{Structural geological forward modeling}\label{sec:struc_geo_modeling}       
+        Performing forward modeling in the context of structural geology requires the use of a suitable modeling step.
+Additionally, for the application in a probabilistic setting, the method should enable fully automatic reconstruction of the model, when parameters are changed.
+The application in this work follows the example of \citet{delaVarga2016} and relies on the use of implicit interpolation for geological modeling, a method developed by \citet{Lajaunie1997} and \citet{calcagno2008geological}.\\
+        This implicit method relies on the interpolation of a potential-field scalar function $T(p)$ that reflects the geometry of geological structures \citep{calcagno2008geological}.
+Modeling $T(p)$ is achieved by cokriging (i) contact points on geological interfaces which are regarded through increments of the potential field $T(\vec{x})-T(\vec{x}^{\,'})$ and (ii) orientation data as gradients, i.e.
+partial derivatives of the potential field $\delta T(\vec{x})/\delta u_\beta$ in each directon $u$ \citep{calcagno2008geological}.
+
+		\subsection{Implementing Bayesian analysis numerically with Python and PyMC}
 		Bayesian analysis can be conducted using probabilistic programming \citep{salvatier2016pymc3}.
@@
 		\citet{salvatier2016pymc3} point out that the development of PyMC3 continuing, as the inclusion of further tools is planned for future updates.
+
+		\subsection{Reservoir Volumetric Calculation}

### 2017-09-04 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Continued methods text
Interpolation method done

@@
 Additionally, for the application in a probabilistic setting, the method should enable fully automatic reconstruction of the model, when parameters are changed.
-The application in this work follows the example of \citet{delaVarga2016} and relies on the use of implicit interpolation for geological modeling, a method developed by \citet{Lajaunie1997} and \citet{calcagno2008geological}.\\
-        This implicit method relies on the interpolation of a potential-field scalar function $T(p)$ that reflects the geometry of geological structures \citep{calcagno2008geological}.
-Modeling $T(p)$ is achieved by cokriging (i) contact points on geological interfaces which are regarded through increments of the potential field $T(\vec{x})-T(\vec{x}^{\,'})$ and (ii) orientation data as gradients, i.e.
+The application in this work follows the example of \citet{delaVarga2016} and relies on the use of implicit interpolation for geological modeling, a method developed and elaborated by \citet{Lajaunie1997} and \citet{calcagno2008geological}.\\
+        This implicit method relies on the interpolation of a potential field scalar function $T(\vec{x})$ of any point $\vec{(x)}$ in a 3D space, and thus reflects the geometry of geological structures \citep{calcagno2008geological}.
+Modeling $T(\vec{x}$ is achieved by cokriging that regards two forms of data: (i) contact points on geological interfaces through increments of the potential field $T(\vec{x})-T(\vec{x}^{\,'})$ and (ii) orientation data as gradients, i.e.
 partial derivatives of the potential field $\delta T(\vec{x})/\delta u_\beta$ in each directon $u$ \citep{calcagno2008geological}.
+The respective estimator is defined as follows:
+        \begin{equation}\label{eq:Cokriging_Estimator}
+                T(\vec{x})-T(\vec{x_0})=\sum_{\alpha=1}^{M}\mu_\alpha(T\vec{x_\alpha}-T(\vec{x}^{\,'}_\alpha))+\sum_{\beta=1}^{N}\nu_\beta\frac{\delta T}{\delta u_\beta}(\vec{x}_\beta),
+        \end{equation}
+
+        where $\vec{x_0}$ is an arbitrary origin, $M$ and $N$ are the total number of data points and partial derivatives respectively, and their relative contributions are weighed by the factors $\mu_\alpha$ and $\nu_\beta$.
+Furthermore, $T$ is assumed to be a random function defined by polynomial drift and a stationary covariance $K(h)$ \citep{calcagno2008geological}.
+The use of a cubic covariance model is suggested by \citet{calcagno2008geological}, based on the results from studies conducted by \citet{aug2004modelisation} and \citet{chiles2004modelling}.\\
+		\begin{figure}[h]
+			\centering
+			\includegraphics[width=1\textwidth]{Figures/calcagno_pot_field.png}
+			\caption{woo a potential field! from \citet{calcagno2008geological}}\label{fig:pot_field}
+		\end{figure}        
+        Resulting potential fields can be used to describe geological interfaces as iso-surfaces in any kind of 3D geometry \citep{calcagno2008geological}.
+Fault geometries can be interpolated analogously.
+These can be infinite in the 3D space, interrelated in a fault network or finite.
+To account for the effect of faults on geological layers, discontinuous potential fields are created by applying discontinuous drift functions in the cokriging system.
+Additionally, geological rules allow for the representation of several types of interactions between sets of geological layers \citep{calcagno2008geological}.\\
+        It is pointed out by \citet{calcagno2008geological}, that this method is particularly appropriate for cases in which knowledge about the geology is only given a sparse locations and is thus applicable for a wide variety of typical problem in geological settings.
+Its suitability for this work is furthermore emphasized by the possibility to modify the topology-defining geological pile to achieve different geometric realizations without altering the data.
+The model can thus be updated in the face of new data or interpretations \citep{calcagno2008geological}.
 

### 2017-09-22 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Continued text, on reservoir part

@@
 
-		\subsection{Implementing Bayesian analysis numerically with Python and PyMC}
+        \subsection{Economic significance and reservoir valuation}\label{sec:Reservoir_values}
+        In this work, the generated structural models are to be assigned an economic significance in order to approach a real setting, in which hypothetical actors could have an interest in taking a decision, and thus, to evaluate the influence of the Bayesian inference of the decision-making process.
+Changes in Bayes actions are considered as a measure for this effect.\\
+        For this purpose, it is assumed that a hydrocarbon reservoir system is represented by each geological model realization.
+These are typically comprised stratigraphically of at least one potential reservoir formation, succeeded by at least one sealing layer above.
+Additionally, a trapping mechanism is normally required (source?).\\
+        The resulting reservoirs are to be valuated in a way that respective estimations on the basis of loss functions can be conducted, i.e.
+Bayesian decision theory is applicable.
+In the typical petroleum industrial setting, the main question is one of how much of the resource can be produced and how high the return on investment will be \citep{dean2007volumetric}.\\
+        Before pressure and production tests have been conducted (i.e.
+before production has started), volumetric estimation is the only approach to assess the amount of hydrocarbons in place in a reservoir.
+From this value, recoverable reserves can be estimated based on an estimated recovery factor \citep{dean2007volumetric}.\\
+        Oil-in-place and gas-in-place volumes are calculated based on:
+        \begin{enumerate}
+        \item Subsurface rock volume containing hydrocarbons. This is mainly defined by thickness and areal extend of the accumulation.
+        \item Weighted average effective porosity of the reservoir rock.
+        \item Water saturation in the reservoir rock.
+        \item Hydrocarbon fluid properties \citep{dean2007volumetric}.
+        \end{enumerate}
+        As the cases used in this work are purely artificial, a pure oil accumulation is assumed.
+Using the factors listed above, the respective equation for original oil-in-place (OOIP) is formulated as follows:
+        \begin{equation}\label{eq:OOIP}
+        OOIP = A * h * \phi * (1 - S_W) * 1/FvF
+        \end{equation}
+        Where OOIP is returned in $m³$.
+The hydrocarbon-filled rock volume is defined by the drainage area $A$ in $m²$ and the net pay thickness $h$ in $m$.
+Porosity $\phi$ and water saturation $S_W$  (interstitial water) are given in fraction of the rock volume.
+A dimensionless factor for the change in oil volume between reservoir conditions and standard conditions at surface is represented by the formation volume factor $FvF$.
+Thus, shrinkage of the oil volume brought to the surface is determined by $1/FvF$ \citep{dean2007volumetric}.\\
+        Subsequently, the effectively recoverable oil reserves can be calculated by multiplying the OOIP with a recovery factor $RF$.
+It is influenced by a number of fluid properties, such as viscosity, density, solution oil/gas ratio and the formation volume factor.
+Thus, it is difficult to estimate \citep{dean2007volumetric}.
+According to
+
+        - in the end: NPV
+
+        - here: 1D abstract case: arbitrarily chosen "scoring-system" as a reservoir value measure
+
+		\subsection{Numerical computational implementation via Python, GemPy and PyMC}
 		Bayesian analysis can be conducted using probabilistic programming \citep{salvatier2016pymc3}.
-For doings this, the programming language of choice in this work is Python.
+The implicit method of forward geological modeling, described above, is to be embedded in such a framework.
+For doing so, the programming language of choice in this work is Python.
 The merits of Python have been pointed out by \citet{behnel2010, salvatier2016pymc3, Langtangen2008}.
@@
 Python is dynamic, compatible with multiple platforms and offers good support for numerical computing.
-Integration of other scientific libraries and extension via C, C++, Fortran or Cython is easily possible \citep{behnel2010, salvatier2016pymc3, Langtangen2008}.
-Python is thus a straightforward tool for the implementation of central components of Bayesian analysis, such as custom statistical distributions and samplers \citep{salvatier2016pymc3}.
-
-		In probabilistic programming, programming variables are used as components to build probabilistic models \citep{davidson2015}.
-
-		The numerical implementation of Bayesian analysis is further aided by the use of the Python library PyMC3, which was developed for conducting Bayesian inference and prediction problems in an open-source probabilistic programming framework \citep{davidson2015, salvatier2016pymc3}.
-Different model fitting techniques are provided in PyMC3, such as the \textit{maximum a posteriori} (MAP) method and several \textit{Monte Carlo Markov Chain} (MCMC) sampling methods \citep{salvatier2016pymc3}.
-		Name examples for state-of-the-art samplers included.
-		\citet{salvatier2016pymc3} point out that the development of PyMC3 continuing, as the inclusion of further tools is planned for future updates.
-
-		\subsection{Reservoir Volumetric Calculation}
+Integration of other scientific libraries and extension via C, C++, Fortran or Cython are easily possible \citep{behnel2010, salvatier2016pymc3, Langtangen2008}.
+Python is thus a straightforward tool for the implementation of central components of Bayesian analysis, such as custom statistical distributions and samplers \citep{salvatier2016pymc3}.\\
+		The 3D geological modeling step in this work is implemented using GeMpy, an open-source, Python-based software that is able to generate and visualize complex 3D structural geological models based on the potential field interpolation method elaborated in Section \ref{sec:struc_geo_modeling} \citet{delaVarga2017gempy}.
+Its design allows for its application in a probabilistic setting, particularly by coupling it with PyMC (see below).
+At the time of writing this, GemPy is still under development (version 0.995), but is already functioning for the purpose of this work.\\
+		For conducting the Bayesian analysis, GemPy is combined with the Python library PyMC, which was developed for conducting Bayesian inference and prediction problems in an open-source probabilistic programming framework \citep{davidson2015, salvatier2016pymc3}.
+Different model fitting techniques are provided in PyMC, such as the \textit{maximum a posteriori} (MAP) method and several (MCMC) sampling methods, including the Adaptive Metropolis explained in Section \ref{sec:mcmc} \citep{salvatier2016pymc3}.
+The components which are used to construct a statistical model, are represented by \textit{Deterministic} or \textit{Stochastic} variables in PyMC.
+the values of \textit{Deterministic} variables are completely dependent on its parents' values, as defined by a respective mathematical function \citep{salvatier2016pymc3}.
+\textit{Stochastic} variables are used to represent uncertain parameters $\theta$ or observed stochastic variables as likelihood functions $p(y|\theta)$ \citep{salvatier2016pymc3, delaVarga2016}.
+Complex mathematical relations between \textit{Stochastic} variables can be described through \textit{Deterministic} variables \citep{delaVarga2016}.
+Futhermore, PyMC allows for the creation of own object definitions inheriting from the class descriptions of these two variable types.\\
+		As of writing this, the newer PyMC3 is not yet supported by GemPy.
+Thus, it is resorted to PyMC 2 for 3D geological settings in this work.
+However, the newer PyMC 3 is used in a simple 1D scenario.
+\citet{salvatier2016pymc3} point out that the development of PyMC 3 is continuing, as the inclusion of further tools is planned for future updates.

### 2017-09-22 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Next: Time to transform 1D part into methods and results

@@
         Minimization of a loss function returns a Bayesian point estimate known as Bayes action $\delta^P(X)$, which is the estimate, action or decision with the least expected loss according to the loss function \citep{berger2013stat}.
-For a unimodal and symmetric absolute-error loss function, the Bayes action is simply the median of the posterior distribution, while squared-error loss it is the mean \citep{davidson2015, berger2013stat}.
+For a unimodal and symmetric absolute-error loss function, the Bayes action is simply the median of the posterior distribution, while using squared-error loss it is the mean \citep{davidson2015, berger2013stat}.
 The MAP (maximum a posteriori) estimate is the minimizing solution for the posterior using zero-one loss \citep{davidson2015}.
@@
 Thus, shrinkage of the oil volume brought to the surface is determined by $1/FvF$ \citep{dean2007volumetric}.\\
-        Subsequently, the effectively recoverable oil reserves can be calculated by multiplying the OOIP with a recovery factor $RF$.
-It is influenced by a number of fluid properties, such as viscosity, density, solution oil/gas ratio and the formation volume factor.
+        Subsequently, the effectively recoverable oil volumes (ROV) can be calculated by multiplying the OOIP with a recovery factor $RF$:
+        \begin{equation}\label{eq:ROV}
+                ROV = OOIP * RF = A * h * \phi * (1 - S_W) * 1/FvF * RF
+        \end{equation}
+        the recovery factor is influenced by a number of fluid properties, such as viscosity, density, solution oil/gas ratio and the formation volume factor.
 Thus, it is difficult to estimate \citep{dean2007volumetric}.
-According to
+Globally, the ultimate average recovery factor for oil fields is about 35\% \citep{labastie2011increasingRF}.
+For gas accumulations, however, recovery factors range typically between 70 and 90\% \citep{dean2007volumetric}.\\
+        The models in this work are completely hypothetical and not based on real data.
+The inputs to these equations can thus be chosen arbitrarily to test the applied methods.
+However, to come to significant conclusion, it makes sense to utilize values that approximately represent real possible scenarios.
+Furthermore, as 3D geological structures are modeled here, the hydrocarbon-filled rock volume ($A * h$ in Equations \ref{eq:OOIP} and \ref{eq:ROV}) is defined by every realization of the uncertain geological model.
+The rock volume is thus an uncertain factor.
+Remaining factors can be implemented deterministically or as uncertain values, too.
+For most factors, it seems most suitable to define a normal distribution around a known typical value or global average, as it will return a high probability for common scenarios and occasionally produce cases of rare, more extreme values.
+(INSERT WHAT IS CHOSEN IN THE END - MAYBE MORE SENSE TO TAKE FIXED VALUES FOR THE FACTORS THEN RV, SO THAT ITS INFLUENCE IS CLEAR; THEN MAYBE INCLUDE POROSITY?
+OR COMPARE BOTH POSSIBILITIES?).\\
+        Consequently, OOIP and ROV calculations represent the step following and using the results from computing the geological model, so that the realizations are assigned values that can be interpreted from a perspective of economic interest.
+The resulting values from numerous model runs are used as a base probability distribution for the true value, on which loss function estimation can be applied as described in Section \ref{sec:loss}.\\
 
+        \subsubsection{Abstract valuation using scores in the one-dimensional case}
         - in the end: NPV

### 2017-09-23 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Bit of progress

@@
 In the typical petroleum industrial setting, the main question is one of how much of the resource can be produced and how high the return on investment will be \citep{dean2007volumetric}.\\
+
+        \subsubsection{Original oil-in-place and recoverable volumes as value measures}
         Before pressure and production tests have been conducted (i.e.
@@
         Consequently, OOIP and ROV calculations represent the step following and using the results from computing the geological model, so that the realizations are assigned values that can be interpreted from a perspective of economic interest.
-The resulting values from numerous model runs are used as a base probability distribution for the true value, on which loss function estimation can be applied as described in Section \ref{sec:loss}.\\
+The resulting values from numerous model runs are used as a base probability distribution for the true value, on which loss function estimation can be applied as described in Section \ref{sec:loss}.
+        \subsubsection{Net present value}
+        hmmm...
 
-        \subsubsection{Abstract valuation using scores in the one-dimensional case}
-        - in the end: NPV
-
-        - here: 1D abstract case: arbitrarily chosen "scoring-system" as a reservoir value measure
+        \subsubsection{Abstract valuation using scores in the one-dimensional model}
+        For the simplified 1D model that is described further below, OOIP and recoverable volume calculations are not applicable, as these require a 3D setting.
+Given only layer positions in one vertical dimension, it is resorted to an abstract way of reservoir valuation by defining a scoring system.
+The reservoir score is made dependent on three uncertain parameters which can be deduced from the 1D layer positions: (1) reservoir thickness, (2) reservoir top depth and (3) seal thickness.
 
@@
 \citet{salvatier2016pymc3} point out that the development of PyMC 3 is continuing, as the inclusion of further tools is planned for future updates.
+
+		\section{Designing a case-specific loss function}
+
+		\section{Constructing a 1D geological reservoir model}
+
+		\section{Constructing a full 3D geological reservoir model}

### 2017-09-28 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Restructuring cjhapters

@@
 
-		\section{Designing a case-specific loss function}
+		\section{Designing a case-specific loss function}\label{sec:LF_design}	
+		As explained before in Section \ref{sec:loss}, the standard symmetric loss functions provide objectively good estimators minimizing expected loss by returning the median or mean respectively.
+However, assigning an economic notion to our model and assuming the case of an actor or decision maker in any field, naturally necessitates the consideration of preferences, interests and the overall subjective perspective such an individual or for example a company might have.
+Further constraints, properties and factors can also be specific to the field, industry or generally to the problem at hand.
+Consequently, the design of a more specific non-standard and possibly asymmetric loss function might be required, so that an adapted Bayesian estimator can be found.
+One that includes subjective aspects and difference in weighting of particular gains or losses, arising from an actor's inherent preferences and the environment in which the actor has to estimate or make a decision.
+In the face of several uncertain parameters, a perfectly true estimate is virtually unattainable.
+However, an attempt can be made to design a customized loss function that returns a Bayesian estimator involving the least bad consequences for an individual in a specific environment.
+Assuming the reservoir setting case and valuation methods introduced in Section \ref{sec:Reservoir_values}), such a customization attempt is made and explained step by step in the following.
+(SOURCES?)\\
+		For the purpose of estimation, it makes sense that one of the standard loss functions is chosen as a basis and a customized loss function is developed from there.
+The absolute-error loss seems most appropriate for this case of hydrocarbon reservoir value estimation.
+Ideally, an actor would like to know the exact true value of interest, say the OOIP, so that investments or resources can be allocated appropriately in order to acquire economic gains.
+This allocation is the decision to be made or action to be taken.
+Deviations from the unknown true value in the form of over- and underestimation bring about an error and loss accordingly.
+In this case, it is assumed that investments increase linearly with linear growth in the value of the resource.
+For this reason, the absolute-error loss function is favored here over the squared-error loss function.
+It is chosen as the base function to which further development steps refer, based on mostly logical case-specific assumptions:
+
+		\begin{enumerate}
+			\item Step I: The standard symmetrical absolute-error loss function is chosen as a starting point for further customization steps:
+
+			\begin{equation}
+			L(\theta,\hat{\theta}) = |\theta - \hat{\theta}|
+			\end{equation}
+
+			\item Step II: Considering the development of a hydrocarbon reservoir, it can be assumed that over-investing is worse than under-investing.
+Overestimating the size of an accumulation might for example lead to the installation of equipment or facilites that are actually redundant or unnecessary.
+This would come with additional unrecoverable expenditures.
+Consequences from underestimating, however, may presumably be easier to resolve.
+Additional equipment can often be installed later on.
+Hence, overestimation is weighted stronger in this loss function by multiplying the error with an overestimation factor \textit{a} (= 1.25):
+
+			\begin{equation}\label{eq:LF_II}
+			L(\theta,\hat{\theta}) = |(\theta-\hat{\theta})|*a
+			\end{equation}
+
+			\item Step III: The worst case for any project would be that its development is set into motion, expecting a gain, only to discover later that the value in the reservoir does not cover the costs of realizing the project, resulting in an overall loss.
+A petroleum system might also turn out to be a complete failure, containing no value at all, although the actor's estimate indicated the opposite.
+Here, this is referred to as worst case or fatal overestimation.
+A positive value is estimated, but the true value is zero or negative.
+This is worse than the "normal" non-fatal overestimation, where both values are positive and a net gain is still achieved, which is only smaller then the best possible gain of expecting the true value.
+Fatal overestimation is included in the loss function by using another weighting factor \textit{b} (= 2) that replaces \textit{a}:
+
+			\begin{equation}\label{eq:LF_III}
+			L(\theta,\hat{\theta}) = |(\theta-\hat{\theta})|*b
+			\end{equation}
+
+			(In other words: Fatal overestimation is twice as bad as simple underestimation.)
+
+			\item Step IV: A worst case or fatal underestimation can also be derived from the idea of estimating a zero or negative value, when the true value is actually positive.
+This is assumed to be worse than non-fatal overestimation, but clearly better than fatal overestimation.
+No already owned resources are wasted, it is only the potential value that is lost, i.e.
+opportunity costs that arise from completely discarding a reservoir with a potential gain equal to the positive true value.
+Fatal underestimation is weighted using a third factor \textit{c}:
+
+			\begin{equation}\label{eq:LF_IV}
+			L(\theta,\hat{\theta}) = |(\theta-\hat{\theta})|*c
+			\end{equation}
+
+		\end{enumerate}
+
+		Combining these adaption steps and the conditions defined in them, results in the following customized loss function:
+
+		\begin{equation}\label{eq:LF_final}
+		L(\theta,\hat{\theta}) =
+		\begin{cases}
+		|\theta - \hat{\theta}|, & \text{for } 0<\hat{\theta}<\theta  \\
+		|\theta-\hat{\theta}|*a, & \text{for } 0<\theta<\hat{\theta} \\
+		|\theta-\hat{\theta}|*b, & \text{for } \theta\leq0<\hat{\theta} \\
+		|\theta-\hat{\theta}|*c, & \text{for } \hat{\theta}\leq0<\theta 
+		\end{cases},
+		\text{ with } a,b,c \in \mathbb{Q}
+		\end{equation}
+
+		Realizations of the four adaption steps are depicted in the plots in Figure \ref{fig:LF_4steps}.
+It is important to note that the weighting factors defined above can take basically any numerical values but should be chosen in a way that they appropriately represent the framework conditions of the problem.
+Here, for example, it is assumed that normal overestimation is 25\% (\textit{a} = 1.25), fatal overestimation 100\% (\textit{b} = 2) and fatal underestimation 50\% (\textit{c} = 1.5) worse than normal underestimation.
+
+		\begin{figure}[h]
+			\centering
+			\includegraphics[width=1\textwidth]{Figures/LF_4steps.png}
+			\caption{The single steps of customizing the loss function are depicted in plots I - IV.}\label{fig:LF_4steps}
+		\end{figure}
+
+		In Figure \ref{fig:LF_4steps} it can be observed, that assigning a stronger weight on overestimation (plot II) steepens the curve on the right hand side and shifts the minimum to the left, i.e.
+to a lower estimate.
+Using \textit{a} = 1.25, the Bayes action changes from the median, to a value close to the mean of the distribution.
+The shift and steepening are significantly reinforced by the introduction of fatal overestimation (plot III).
+With \textit{b} = 2, the Bayes action drops to the zero score estimate.
+It can also be noted, that by defining a condition dependent on the algebraic sign of the values, according to which only losses for positive estimates are multiplied by \textit{b}, a distinct jump appears at the zero score boundary.
+Due to a similar condition, the same effect is observed on the negative side of estimate values, where the curve has also been steepened, after including fatal underestimation (plot IV).
+This comes with a shift of the minimum towards positive values.
+It is also to be noted that with every customization step, the overall expected loss is increased.
+
+		\begin{figure}[h]
+			\centering
+			\includegraphics[width=1\textwidth]{Figures/LF4_det_values.png}
+			\caption{Loss based on the customized loss function (IV) for determined true scores of -750, 0 and 750.}\label{fig:LF4_det_values}
+		\end{figure}
+
+		The implementation of this customized loss function (IV) using single determined values for the true score is plotted in Figure \ref{fig:LF4_det_values}.
+This helps to clarify the way real losses result for each guess, relative to given true score values.
+The expected loss is acquired by arithmetically averaging such loss realizations based on the true score probability distribution by using formula \ref{eq:ExpectedLoss2}.
+
+		It has to be emphasized that this is just one possible proposal for loss function customization.
+There exists not one perfect design for such a case.
+Slight to strong changes can already be implemented by simply varying the values of the weighting factors \textit{a, b} and \textit{c}.
+Fundamentally different loss functions can also be based on a significantly different mathematical structure.
+Loss functions are customized regarding the problem environment and according to the to the subjective needs and objectives of an actor.
+Thus, they are mostly defined by the actor expressing his perspective.
+Changes in the individual's perception and viewpoint might lead to further customization needs even later on.
+(Especially considering individual persons as actors, psychological aspects may play a significant role.)
+
+		(Estimate or prediction? True value only known if a "yes"-action (development) is taken.)
+
+		\subsection{Including different risk-affinities in the loss function}
+
+		One can assume that several actors in one sector or decision environment may have the same general loss function, but different affinities concerning risks.
+This might be based for example on the different psychological factors or economic philosophies followed by companies.
+It might also be based on budgets and options such actors have available.
+An intuitive example is the comparison of a small and a large company.
+A certain false estimate or error might have a significantly stronger impact on a company which has a a generally lower market share and only few projects, than on a larger company which might possess a higher financial flexibility and for which one project is only one of many development options in a portfolio.
+
+		In the following, the loss function is further adapted to consider different risk-affinities of several actors.
+Representing risk behavior in a loss function can also be done in different ways and regarding different types of risks.
+Here, bidding lower is considered the cautious, risk-averse option, as smaller losses can be expected from underestimating.
+Guessing higher is deemed riskier.
+Losses from overestimation are greater.
+However, bidding correctly on a higher value, will also return a greater gain.
+It is assumed that risk-friendly actors care less about fatal underestimation, i.e.
+they will rather develop a project than discard it.
+In the loss function, risk is simply included using a risk factor \textit{r} which alters the weighting factors \textit{a, b} and \textit{c} respectively:
+
+		\begin{equation}\label{eq:LFR_final}
+		L(\theta,\hat{\theta}) =
+		\begin{cases}
+		|\theta - \hat{\theta}|, & \text{for } 0<\hat{\theta}<\theta  \\
+		|\theta-\hat{\theta}|*(a*r), & \text{for } 0<\theta<\hat{\theta} \\
+		|\theta-\hat{\theta}|*(b*r), & \text{for } \theta\leq0<\hat{\theta} \\
+		|\theta-\hat{\theta}|*(c*(r^{-0.5})), & \text{for } \hat{\theta}\leq0<\theta 
+		\end{cases},
+		\text{ with } a,b,c,r \in \mathbb{Q}
+		\end{equation}
+
+		According to this, for \textit{r} = 1 the risk-neutral loss function is returned, since \textit{a, b} and \textit{c} are not altered.
+For \textit{r} $<$ 1, the weight on overestimating is reduced and increased for fatal underestimation.
+This represents a more risk-friendly actor that is willing to bid on a higher estimate to attain a greater gain.
+For \textit{r} $>$ 1, the overestimation weight is increased, the underestimation weight decreased in the loss function and respectively more risk-averse actors are prompted to bid on lower estimates.
+
+		The factor \textit{r} can take basically any positive values.
+However, since risk-neutrality is expressed by\textit{r} = 1, values 0 $<$ \textit{r} $<$ 2 are considered to be the most appropriate choices to represent both sides of risk-affinity equally.
+
+		Implementing it as a factor \textit{r} leads to different steepnesses of the plotted curves depicting loss and expected loss.
+The effect on the latter is visualized in Figure \ref{fig:LFR}, where 0.5, 0.75, 1, 1.25 and 1.5 were chosen as values for \textit{r}.
+		It can be observed that the minima for expected loss, i.e.
+the Bayesian estimators for differently risk-affine actors, are located at different estimates.
+Mean and median are clearly surpassed by the best estimate of the most risk-friendly actor (\textit{r} = 0.5), while for the most risk-averse actors (\textit{r} 0.5 and \textit{r} = 0.75), the Bayes action equals a zero score estimate (and thus the decision to take on action).
+In Figure \ref{fig:LFR} it can also be recognized that the expected loss is generally lower for risk-friendlier actors (on the side of positive estimates).
+
+		\begin{figure}[h]
+			\centering
+			\includegraphics[width=1\textwidth]{Figures/LFR.png}
+			\caption{Plotting of expected loss realizations after including the risk factor r in the loss function (IV) for actors with risk-affinities ranging from risk-averse (r = 0.5 and 0.75), over risk-neutral (r = 1), to risk-friendly (r = 1.25 and r = 1.5).}\label{fig:LFR}
+		\end{figure}
 

### 2017-09-28 Risk analysis/Text/Layout/Layout/chapter3_results.tex
Restructuring cjhapters

	\chapter{Results}\label{cha:results}
	Text

		\section{Results using the 1D geological model}

			\subsection{Score system}

			\subsection{Applying the loss function}

			\subsection{Bayesian updating using thickness likelihoods}

				\subsubsection{Updating case I: Moderately reinforcing information}

				\subsubsection{Updating case II: Likely reliable seal}

				\subsubsection{Updating case III: Safe seal but likely subpar reservoir thickness}

		\section{Results using the 3D geological model}

### 2017-09-28 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Progress with methods and results

@@
 Bayesian decision theory is applicable.
-In the typical petroleum industrial setting, the main question is one of how much of the resource can be produced and how high the return on investment will be \citep{dean2007volumetric}.\\
+In the typical petroleum industrial setting, the main question is one of how much of the resource can be produced and how high the return on investment will be \citep{dean2007volumetric}.
 
@@
         As the cases used in this work are purely artificial, a pure oil accumulation is assumed.
-Using the factors listed above, the respective equation for original oil-in-place (OOIP) is formulated as follows:
+Using the factors listed above, the respective equation for original oil-in-place ($OOIP$) is formulated as follows:
         \begin{equation}\label{eq:OOIP}
@@
         \end{equation}
-        Where OOIP is returned in $m³$.
+        Where $OOIP$ is returned in $m³$.
 The hydrocarbon-filled rock volume is defined by the drainage area $A$ in $m²$ and the net pay thickness $h$ in $m$.
@@
 Thus, shrinkage of the oil volume brought to the surface is determined by $1/FvF$ \citep{dean2007volumetric}.\\
-        Subsequently, the effectively recoverable oil volumes (ROV) can be calculated by multiplying the OOIP with a recovery factor $RF$:
+        Subsequently, the effectively recoverable oil volumes ($ROV$) can be calculated by multiplying the $OOIP$ with a recovery factor $RF$:
         \begin{equation}\label{eq:ROV}
@@
 OR COMPARE BOTH POSSIBILITIES?).\\
-        Consequently, OOIP and ROV calculations represent the step following and using the results from computing the geological model, so that the realizations are assigned values that can be interpreted from a perspective of economic interest.
+        Consequently, $OOIP$ and $ROV$ calculations represent the step following and using the results from computing the geological model, so that the realizations are assigned values that can be interpreted from a perspective of economic interest.
 The resulting values from numerous model runs are used as a base probability distribution for the true value, on which loss function estimation can be applied as described in Section \ref{sec:loss}.
+$OOIP$ and $ROV$ are applicable in a 3D setting, the 1D geological model introduced below in Section \ref{sec:1D_model}, however, requires the use of an abstract valuation system that is explained in Section \ref{sec:1D_score_system}.
+
         \subsubsection{Net present value}
         hmmm...
-
-        \subsubsection{Abstract valuation using scores in the one-dimensional model}
-        For the simplified 1D model that is described further below, OOIP and recoverable volume calculations are not applicable, as these require a 3D setting.
-Given only layer positions in one vertical dimension, it is resorted to an abstract way of reservoir valuation by defining a scoring system.
-The reservoir score is made dependent on three uncertain parameters which can be deduced from the 1D layer positions: (1) reservoir thickness, (2) reservoir top depth and (3) seal thickness.
 
@@
 		\end{equation}  
-
-		Realizations of the four adaption steps are depicted in the plots in Figure \ref{fig:LF_4steps}.
 		It is important to note that the weighting factors defined above can take basically any numerical values but should be chosen in a way that they appropriately represent the framework conditions of the problem.
-Here, for example, it is assumed that normal overestimation is 25\% (\textit{a} = 1.25), fatal overestimation 100\% (\textit{b} = 2) and fatal underestimation 50\% (\textit{c} = 1.5) worse than normal underestimation.
-
-		\begin{figure}[h]
-			\centering
-			\includegraphics[width=1\textwidth]{Figures/LF_4steps.png}
-			\caption{The single steps of customizing the loss function are depicted in plots I - IV.}\label{fig:LF_4steps}
-		\end{figure}
-
-		In Figure \ref{fig:LF_4steps} it can be observed, that assigning a stronger weight on overestimation (plot II) steepens the curve on the right hand side and shifts the minimum to the left, i.e.
-to a lower estimate.
-Using \textit{a} = 1.25, the Bayes action changes from the median, to a value close to the mean of the distribution.
-The shift and steepening are significantly reinforced by the introduction of fatal overestimation (plot III).
-With \textit{b} = 2, the Bayes action drops to the zero score estimate.
-It can also be noted, that by defining a condition dependent on the algebraic sign of the values, according to which only losses for positive estimates are multiplied by \textit{b}, a distinct jump appears at the zero score boundary.
-Due to a similar condition, the same effect is observed on the negative side of estimate values, where the curve has also been steepened, after including fatal underestimation (plot IV).
-This comes with a shift of the minimum towards positive values.
-It is also to be noted that with every customization step, the overall expected loss is increased.
-
-		\begin{figure}[h]
-			\centering
-			\includegraphics[width=1\textwidth]{Figures/LF4_det_values.png}
-			\caption{Loss based on the customized loss function (IV) for determined true scores of -750, 0 and 750.}\label{fig:LF4_det_values}
-		\end{figure}
-
-		The implementation of this customized loss function (IV) using single determined values for the true score is plotted in Figure \ref{fig:LF4_det_values}.
-This helps to clarify the way real losses result for each guess, relative to given true score values.
-The expected loss is acquired by arithmetically averaging such loss realizations based on the true score probability distribution by using formula \ref{eq:ExpectedLoss2}.
-
+Here, based on the considerations named above, it is assumed that normal overestimation is 25\% (\textit{a} = 1.25), fatal overestimation 100\% (\textit{b} = 2) and fatal underestimation 50\% (\textit{c} = 1.5) worse than normal underestimation.
+\\
 		It has to be emphasized that this is just one possible proposal for loss function customization.
@@
 		(Estimate or prediction? True value only known if a "yes"-action (development) is taken.)
+		INSERT PLOTS BASED ON EXAMPLE NORMAL FUNCTIONS?
 
@@
 Here, bidding lower is considered the cautious, risk-averse option, as smaller losses can be expected from underestimating.
-Guessing higher is deemed riskier.
-Losses from overestimation are greater.
+Guessing higher is deemed riskier, as losses from overestimation are greater.
 However, bidding correctly on a higher value, will also return a greater gain.
@@
 This represents a more risk-friendly actor that is willing to bid on a higher estimate to attain a greater gain.
-For \textit{r} $>$ 1, the overestimation weight is increased, the underestimation weight decreased in the loss function and respectively more risk-averse actors are prompted to bid on lower estimates.
+For \textit{r} $>$ 1, the overestimation weight is increased in the loss function, the underestimation weight is decreased and respectively more risk-averse actors are prompted to bid on lower estimates.\\
+		The factor \textit{r} can take basically any positive values.
+However, since risk-neutrality is expressed by \textit{r} = 1, values 0 $<$ \textit{r} $<$ 2 are considered to be the most appropriate choices to represent both sides of risk-affinity equally here.
 
-		The factor \textit{r} can take basically any positive values.
-However, since risk-neutrality is expressed by\textit{r} = 1, values 0 $<$ \textit{r} $<$ 2 are considered to be the most appropriate choices to represent both sides of risk-affinity equally.
+		\section{1D geological reservoir modeling}\label{sec:1D_model}
+		For simple understanding and a preliminary assessment of the Bayesian statistical methods and the loss function described above, they are first to be applied in the context of an abstract one-dimensional reservoir case.
+The underlying model and basic approach are inherited from \citet{delaVarga2016}.
+Parameters are adapted to more appropriately represent a reasonable geological petroleum system, consisting of a reservoir formation with overlying seal in the subsurface.
+In this 1D setting, only the interface depths and thicknesses of layers such as the reservoir or seal unit can be observed.
+Other defining aspects, such as structural entrapment, are assumed to be given.
+Limiting the model to only one dimension and a small number of uncertain stratigraphical parameters allows for a relatively straightforward and simplified approach to assessing an abstract type of value for a reservoir and using the custom loss function (Equation \ref{eq:LF_final}) for value estimation.
+The construction of the 1D geological model is described in the following.
 
-		Implementing it as a factor \textit{r} leads to different steepnesses of the plotted curves depicting loss and expected loss.
-The effect on the latter is visualized in Figure \ref{fig:LFR}, where 0.5, 0.75, 1, 1.25 and 1.5 were chosen as values for \textit{r}.
-		It can be observed that the minima for expected loss, i.e.
-the Bayesian estimators for differently risk-affine actors, are located at different estimates.
-Mean and median are clearly surpassed by the best estimate of the most risk-friendly actor (\textit{r} = 0.5), while for the most risk-averse actors (\textit{r} 0.5 and \textit{r} = 0.75), the Bayes action equals a zero score estimate (and thus the decision to take on action).
-In Figure \ref{fig:LFR} it can also be recognized that the expected loss is generally lower for risk-friendlier actors (on the side of positive estimates).
+			\subsection{Construction of the 1D geological model}\label{sec:1D_construction}	
+			\citet{delaVarga2016} constructed a simple geological model using three uncertain positions in vertical one-dimensional space, marking hypothetical boundaries of layers in a subsurface column.
+The location probabilities for these points are defined by sampling from normal distributions.
+Standard deviations of these distributions increase with depth, representing an increase in uncertainty.
+For an approximate representation of a hydrocarbon reservoir system, the distribution means were set to depths of 2000 m (seal top), 2050 m (reservoir top) and 2200 m (reservoir bottom).
+STD-DEVIATIONS?
+These points confine two layers in the middle, from which the upper one can be labeled as seal and the lower one as reservoir.
+The resulting model with its possible layer boundary locations is illustrated in Figure \ref{fig:1D_model}.
 
@@
 				\centering
-			\includegraphics[width=1\textwidth]{Figures/LFR.png}
-			\caption{Plotting of expected loss realizations after including the risk factor r in the loss function (IV) for actors with risk-affinities ranging from risk-averse (r = 0.5 and 0.75), over risk-neutral (r = 1), to risk-friendly (r = 1.25 and r = 1.5).}\label{fig:LFR}
+				\includegraphics[width=1\textwidth]{Figures/1D_model.png}
+				\caption{Probability distributions for positions of layer boundaries in the subsurface and a respective representation using lines.}\label{fig:1D_model}
 			\end{figure}
 
-		\section{Constructing a 1D geological reservoir model}
+	        \subsection{Abstract valuation (scoring) of the 1D geological model}\label{sec:1D_score_system}
+	        For valuating a simplified 1D model, $OOIP$ and recoverable volume calculations are not applicable, as these require a 3D setting.
+Given only layer boundary positions in one vertical dimension, it is resorted to an abstract way of reservoir valuation by defining a scoring system.
+A dimensionless reservoir score is made dependent on three uncertain parameters which can be deduced from the 1D interface positions: (1) reservoir thickness, (2) reservoir top depth and (3) seal thickness (see Figure \ref{fig:3_parameters}).
+
+	        \begin{figure}[h]
+	        	\centering
+	        	\includegraphics[width=1\textwidth]{Figures/3_parameters.png}
+	        	\caption{Probability distributions for the three parameters deduced from the 1D model depicted in \ref{fig:1D_model}.
+Values from these parameter distributions are combined in \ref{eq:1D_score_system} attain a score to valuate reservoir model realizations.}\label{fig:3_parameters}
+	        \end{figure}
+
+	        Assuming that reservoir thickness is a simplified indicator for the amount of extractable oil or gas and thus value in place, a gain in score can be correlated with increase in thickness.
+Here, two score points are assigned to one meter of thickness.
+Increasing costs of drilling are indicated by increasing depth of the reservoir top.
+Consequently, one negative score point is ascribed to every meter in depth.
+Samples from the probability distributions of these two parameters are drawn to model the true score of the reservoir (depth scores are subtracted from reservoir thickness scores).
+	        A third parameter is defined by the seal thickness.
+Score points are not added or subtracted by this parameter directly.
+Instead, a threshold for seal reliability is defined beforehand.
+Here it is set at 20 m thickness.
+If the seal thickness falls below this threshold, it is assumed that the seal fails completely and thus all the potential value (positive score) of the reservoir is lost, while costs of depth (negative score) remain.
+Thus, a condition to check whether the seal is reliable is included in the model.
+A respective equation for the reservoir score $S_{res}$ is defined as follows:
+
+			\begin{equation}\label{eq:1D_score_system}
+			S_{res} = 
+			\begin{cases}
+			2h_{res} + d_{res}, & \text{for } h_{seal} >= 20  \\
+			d_{res}, & \text{for } h_{seal} < 20
+			\end{cases},
+			\end{equation}
+
+			Where reservoir thickness is given by $h_{res}$, seal thickness by $h_{seal}$ and reservoir depth, which is always a negative value, by $d_{res}$.
 
 		\section{Constructing a full 3D geological reservoir model}
+		- concept of 1D case in chapter 3 transferred to a 3D case of a structural geological model
+		- same again here: relatively simple model including some important typical geological and structural features and giving it an economic meaning
+		- alternating layers of sandstones and shales, so typical reservoir and seal formations
+		- structurally defined by a tilted dome-formation (folding along in the direction of two perpendicular axes) which is displaced by a standard normal fault
+		- these features are arranged in a way, that they result in the formation of a trap, defined by the anticlinal features and the fault, as well as enabled by the presence of a reservoir rock
+		- considering a petroleum production industry perspective, actors/ decision makers would be interested in the evaluation of value this potential reservoir
+		- "traditionally", this is represented by volumetric calculations or estimations of the recoverable reserves
+		- here: maximum volume of entrapment and subsequently, relative or hypothetical volumes of recoverable reserve can be calculated (for each model realization)
+		- (later: numbers from example case are taken, to achieve estimations of the NPV for a development project realization)
+		- looking at this structural model, not only the calculated volume can be taken into account, but also risk factors, such as fault permeabity seal safety (and capacity?), possible juxtapositions and leaks
+
+		- as in the 1D case, uncertainties are introduced for a number of parameters in the model, in a way that they affect the resulting maximum volume of the trap (and following also the recoverable reserves)
+		- again, a specific loss function is designed and applied on the results from numerous iterations of the constructed uncertain model, thus resembling the different value estimations/predictions or decisions different actors might make
+		- (to bring this Bayesian method of using loss functions closer to the traditional application of deterministic decision-making in decision-trees, the continuous estimation space of the reservoir value (volume) is combined with the assumption, that the estimated value is linked to a consequent investment into a determined project development size.
+thus, the continuous approach is used to make the decision in a deterministic decision-making space.
+))
+
+		- later on,  the effect of adding information (in the form of likelihoods) is examined, particularly by looking at the changes in decision-making after applying this Bayesian updating and uncertainty has been reduced
+
+		- in the following, the construction and design of this model is described
+
+		- base of the structural geological model is a 3Dcubic  space (block)  with 2000 m size in X, Y and Z x3
+		- place layer positional points in three lines, resembling three hypothetical seismic lines as base information
+		-

### 2017-09-28 Risk analysis/Text/Layout/Layout/chapter3_results.tex
Progress with methods and results

@@
 
-		\section{Results using the 1D geological model}
+		\section{1D geological reservoir model results}
 
-			\subsection{Score system}
+			\subsection{Scoring}
+			\begin{figure}[h]
+				\centering
+				\includegraphics[width=1\textwidth]{Figures/score_results1.png}
+				\caption{Posterior probability distributions from modeling scores using two (A) and three parameters (B).}\label{fig:score_results1}
+			\end{figure}
+			Results from scoring based on a 1D model constructed as defined in Section \ref{sec:1D_construction} are plotted in Figure \ref{fig:score_results1}.
+A test run of scoring with only the two parameters reservoir thickness and depth, is shown in \ref{fig:score_results1}-A.
+These results represented by an approximately normal distribution.
+The score is negative in about 17\% of the cases.
+Mean and median are about the same.\\
+			Full scoring results, including also seal reliability as a parameter, are visualized in Figure \ref{fig:score_results1}-B.
+The main distribution is not changed significantly, except for a striking peak of probability at the possibility for a score of -2000.
+This presumably represents the bulk of cases, in which the seal was assumed to have failed.
+It is to be noted, that the mean of the seal top distribution is found at -2000 m.
+It can be seen in Figure \ref{fig:1D_model}, that around that depth in the model column, reservoir top and seal top probability distributions significantly overlap.
+Thus, there is a possibility for a higher score, due to a shallower reservoir top position, but also a high probability for a seal that is to thin to be reliable.
+The negative score peak at -2000 is thus presumably caused by a high number of seal failures, due to both layer tops found in this area.
+Furthermore, mean and median of the score distribution have been shifted to lower values and are now found further apart.
 
-			\subsection{Applying the loss function}
+			\subsection{Applying the custom loss function on the 1D reservoir scores}
+
+			\begin{figure}[p!]
+				\centering
+				\includegraphics[width=1\textwidth]{Figures/LF_4steps.png}
+				\caption{The single steps of customizing the loss function are depicted in plots I to IV.}\label{fig:LF_4steps}
+				\centering
+				\includegraphics[width=1\textwidth]{Figures/LF4_det_values.png}
+				\caption{Loss based on the customized loss function (Equation \ref{eq:LF_final}) for determined true scores of -750, 0 and 750.}\label{fig:LF4_det_values}
+			\end{figure}			
+			\begin{figure}[h]
+				\centering
+				\includegraphics[width=1\textwidth]{Figures/LFR.png}
+				\caption{Plotting of expected loss realizations after including the risk factor $r$ in the loss function (Equation \ref{eq:LFR_final}) for actors with risk-affinities ranging from risk-averse ($r$ = 0.5 and 0.75), over risk-neutral ($r$ = 1), to risk-friendly ($r$ = 1.25 and $r$ = 1.5).}\label{fig:1D_LFR}
+			\end{figure}
+
+			Realizations of the four loss function adaption steps as elaborated in Section \ref{sec:LF_design} and based on the scoring results shown in Figure \ref{fig:score_results1}-B are depicted in the plots in Figure \ref{fig:LF_4steps}.
+			As expected, the median is returned for the Bayes action, when using the standard symmetric absolute-error loss function (Figure \ref{fig:LF_4steps}-I).
+Compared to this, it can be observed that assigning a stronger weight on overestimation (Figure \ref{fig:LF_4steps}-II) steepens the curve on the right hand side and shifts the minimum to the left, i.e.
+to a lower estimate.
+Using \textit{a} = 1.25, the Bayes action changes from the median, to a value close to the mean of the distribution.
+The shift and steepening are significantly reinforced by the introduction of fatal overestimation (Figure \ref{fig:LF_4steps}-III).
+With \textit{b} = 2, the Bayes action drops to the zero score estimate.
+It can also be noted, that by defining a condition dependent on the algebraic sign of the values, according to which only losses for positive estimates are multiplied by \textit{b}, a distinct jump appears at the zero score boundary.
+Due to a similar condition, the same effect is observed on the negative side of estimate values, where the curve has also been steepened, after including fatal underestimation (Figure \ref{fig:LF_4steps}-IV).
+This comes with a shift of the minimum towards positive values.
+It is also to be noted that with every customization step, the overall expected loss is increased.\\
+			The implementation of the final custom loss function (Figure \ref{fig:LF_4steps}-IV) using single determined values for the true score is plotted in Figure \ref{fig:LF4_det_values}.
+This helps to clarify the way real losses result for each guess, relative to given true score values.
+The expected loss is acquired by arithmetically averaging such loss realizations based on the true score probability distribution by using Equation \ref{eq:ExpectedLoss2}.\\
+			Implementing risk-affinity as a factor \textit{r} leads to different steepnesses of the plotted curves depicting loss and expected loss.
+The effect on the latter is visualized in Figure \ref{fig:1D_LFR}, where 0.5, 0.75, 1, 1.25 and 1.5 were chosen as values for \textit{r}.
+			It can be observed that the minima for expected loss, i.e.
+the Bayesian estimators for differently risk-affine actors, are located at different estimates.
+Mean and median are clearly surpassed by the best estimate of the most risk-friendly actor (\textit{r} = 0.5), while for the most risk-averse actors (\textit{r} = 1.25 and \textit{r} = 1.5), the Bayes action equals a zero score estimate and thus the decision to take no action.
+In Figure \ref{fig:1D_LFR} it can also be recognized that the expected loss is generally lower for risk-friendlier actors on the side of positive estimates, which is the relevant side for decision-making.
 
 			\subsection{Bayesian updating using thickness likelihoods}
+			\citet{delaVarga2016} made use of Bayesian inference to reduce the uncertainty in this type of one-dimensional model. The same is conducted here in the following.\\				
+			The probability distributions for the location of the layer boundaries are treated as priors.
+Now it is assumed that new observations have been made, providing additional information on the likelihoods of the thicknesses of the two layers.
+Likelihood functions for reservoir and seal thicknesses are introduced in the form of normal distributions, defined by means and standard deviations.
+These parameters vary according to nature of the observations made.
+Using the principle of Bayesian inference as explained in chapter \ref{cha:met}, the model is updated and new posterior distributions for our true reservoir score are attained\\
+			Different Bayesian updating cases with different sets of likelihoods are presented in the following, so that various possible results can be compared.
+Each take the layer boundaries defined in Section \ref{sec:1D_construction} as priors and use Bayesian inference with respective likelihood data defined below for each case.
 
 				\subsubsection{Updating case I: Moderately reinforcing information}
+				\begin{figure}[h]
+					\centering
+					\includegraphics[width=1\textwidth]{Figures/update_moderate1.png}
+					\caption{Prior (A1) and posterior distributions (A2) of the layer boundary positions in depth and respective representations (A2, B2).}\label{fig:update_moderate1} 
+				\end{figure}
+
+				In this first case, a normal distribution with a mean of 25~m and a standard deviation of 20~m is chosen to reflect the likelihood of the seal thickness.
+For the reservoir, a normal distribution with a mean of 180~m and a standard deviation of 60~m is chosen.
+Using these likelihoods to update the one-dimensional model, the uncertainty in the posterior probability distributions for the positions of layer boundaries in depth is reduced (see Figure \ref{fig:update_moderate1}).
+
+				Modeling the reservoir score is conducted as above, now based on these new distributions.
+In Figure \ref{fig:update_moderate2_3} A it can be recognized that the bulk of the distribution was shifted to the positive side of scores, while the peak at -~2000 was raised.
+The probability of scores between -~2000 and 0 decreased significantly, i.e.
+the true score is most likely either positive or -~2000 if it is negative.
+
+				\begin{figure}[h]
+					\begin{subfigure}{1\textwidth}
+						\centering
+						\includegraphics[width=1\linewidth]{Figures/update_moderate2.png}
+						%\caption{1a}
+						%\label{fig:sfig1}
+					\end{subfigure}%
+					\\
+					\begin{subfigure}{1\textwidth}
+						\centering
+						\includegraphics[width=1\linewidth]{Figures/update_moderate3.png}
+						%\caption{1b}
+						%\label{fig:sfig2}
+					\end{subfigure}
+					\caption{Reservoir score distributions (A) and change in the realizations of expected loss for several risk parameters (B1, B2) before and after Bayesian updating based on likelihoods defined as follows: Seal thickness mean = 25~m, std = 20~m.
+Reservoir thickness mean = 180~m, std = 60~m.}
+					\label{fig:update_moderate2_3}
+				\end{figure}
+
+				The customized loss function (IV) is now applied on this updated distribution of the reservoir score.
+This is visualized in Figure \ref{fig:update_moderate2_3} in which the expected losses are compared before (B1) and after (B2) uncertainty reduction.
+It is observable, that by adding information about layer thickness likelihoods, Bayes actions are shifted relative to the nature of the information.
+In this case, the added data generally reinforces the likelihoods of the reservoir to be significantly thick.
+Information on the seal, however, based on a normal distribution around 25~m thickness, leaves uncertainty about the reliability of the seal, as the safety threshold is defined as 20~m.
+				%\begin{table}
+				%	\centering
+				%	\begin{tabular}[c]{| l | l | l |}
+				%		\hline
+				%		Risk factor \textit{r} & Shift in Bayes action & Change in expected loss \\ \hline
+				%		0.50 & +~553.67 & -~105.49 \\ 
+				%		0.75 & +~673.40 & -~81.67  \\ 
+				%		1.00 & +~750.05 & -~33.56 \\ 
+				%		1.25 & +~713.19 & +~75.71 \\ 
+				%		1.50 & +~0.00 & +~282.64  \\ 
+				%		\hline
+				%	\end{tabular}
+				%	\caption{Bla}
+				%	\label{tab:update_moderate_tab}
+				%\end{table}
+
+				Increased certainty about the reservoir thickness is sufficient to shift Bayes actions to higher estimates for all actors, but the most risk-averse one.
+These shifts are quantified in Table \ref{tab:update_examples_all}.
+According to these numbers, the risk-neutral actor's estimate is increased the most.
+Expected losses decreased for the risk-neutral and the two risk-friendlier individuals.
+It is clear that the expected loss was reduced the most for the risk-friendliest actor and it increased most significantly for the most risk-averse actor.
 
 				\subsubsection{Updating case II: Likely reliable seal}
+				In this second case, the reservoir thickness likelihood is defined in the same way as before.
+For the seal, a mean of 50~m with a standard deviation of 20~m is chosen, favoring the likelihood of a reliable seal relative to the threshold of 20~m thickness.
+This results in the score distribution depicted in Figure \ref{fig:update_goodseal2_3}.
+The bulk of the distribution is narrowed on the positive side of estimates.
+Mean and median are clearly shifted to higher values as well.
+The "seal failure" peak at -~2000 is significantly.
+decreased.
+
+				Applying the loss function on this new score distribution results in the realizations of expected loss illustrated in Figure \ref{fig:update_goodseal2_3}.
+Bayes actions are shifted clearly to higher estimates and expected losses of these minima are significantly reduced for all actors.
+This is quantified in Table \ref{tab:update_examples_all}.
+According to these values, the risk-neutral to risk-averse individuals seem to profit the most.
+
+				\begin{figure}[h]
+					\begin{subfigure}{1\textwidth}
+						\centering
+						\includegraphics[width=1\linewidth]{Figures/update_goodseal2.png}
+						%\caption{1a}
+						%\label{fig:sfig1}
+					\end{subfigure}%
+					\\
+					\begin{subfigure}{1\textwidth}
+						\centering
+						\includegraphics[width=1\linewidth]{Figures/update_goodseal3.png}
+						%\caption{1b}
+						%\label{fig:sfig2}
+					\end{subfigure}
+					\caption{Reservoir score distributions (A) and change in the realizations of expected loss for several risk parameters (B1, B2) before and after Bayesian updating based on likelihoods defined as follows: Seal thickness mean = 50~m, std = 20~m.
+Reservoir thickness mean = 180~m, std = 60~m.}
+					\label{fig:update_goodseal2_3}
+				\end{figure}
+
+				%\begin{table}
+				%	\centering
+				%	\begin{tabular}[c]{| l | l | l |}
+				%		\hline
+				%		Risk factor \textit{r} & Shift in Bayes action & Change in expected loss \\ \hline
+				%		0.50 & +~522.95 & -~436.26 \\ 
+				%		0.75 & +~697.73 & -~551.92  \\ 
+				%		1.00 & +~855.57 & -~631.92 \\ 
+				%		1.25 & +~912.43 & -~647.10 \\ 
+				%		1.50 & +~819.26 & -~545.96  \\ 
+				%		\hline
+				%	\end{tabular}
+				%	\caption{Bla}
+				%	\label{tab:update_goodseal_tab}
+				%\end{table}
 
 				\subsubsection{Updating case III: Safe seal but likely subpar reservoir thickness}
+				In this third example, seal safety is ensured by using a mean of 70~m with a standard deviation of 10~m for the seal thickness likelihood.
+However, the new observations are assumed to provide information about the reservoir that makes it likely to be thinner as expected.
+Reservoir thickness likelihood is assigned a mean of 100~m and a standard deviation of 40~m.
+
+				The subsequent score distribution is depicted in Figure \ref{fig:update_smallres2_3}.
+It can be seen that while the whole distribution is narrowed, it also is shifted to the left, to lower and negative estimates.
+Mean and median are almost the same.
+As seal reliability is practically guaranteed, the peak at -~2000~m vanishes.
+
+				Based on this probability distribution of the reservoir score, Bayes actions are shifted to lower estimates for all actors (see Figure \ref{fig:update_smallres2_3}).
+In fact, also but the most risk-friendly individual (\textit{r} = 0.5) find their Bayesian estimator to be zero now, i.e.
+project development is deemed to be too risky to them.
+Furthermore, the spread of expected loss values around the minima for all actors is diminished (the expected loss values of the Bayes actions are now much closer to each other).
+
+				Quantified changes in the position and the values of the Bayes actions are comparable to the results in example II (there is no shift for risk-averse actors, as they already found their best estimates to be zero).
+Most notable is the large reduction of expected loss in general (see Table \ref{tab:update_examples_all}).
+
+				\begin{figure}[h]
+					\begin{subfigure}{1\textwidth}
+						\centering
+						\includegraphics[width=1\linewidth]{Figures/update_smallres2.png}
+						%\caption{1a}
+						%\label{fig:sfig1}
+					\end{subfigure}%
+					\\
+					\begin{subfigure}{1\textwidth}
+						\centering
+						\includegraphics[width=1\linewidth]{Figures/update_smallres3.png}
+						%\caption{1b}
+						%\label{fig:sfig2}
+					\end{subfigure}
+					\caption{Reservoir score distributions (A) and change in the realizations of expected loss for several risk parameters (B1, B2) before and after Bayesian updating based on likelihoods defined as follows: Seal thickness mean = 70~m, std = 10~m.
+Reservoir thickness mean = 100~m, std = 40~m.}
+					\label{fig:update_smallres2_3}
+				\end{figure}
+
+				%\begin{table}
+				%	\centering
+				%	\begin{tabular}[c]{| l | l | l |}
+				%		\hline
+				%		Risk factor \textit{r} & Shift in Bayes action & Change in expected loss \\ \hline
+				%		0.50 & -~498.98 & -~538.08 \\ 
+				%		0.75 & -~346.16 & -~717.83  \\ 
+				%		1.00 & -~176.13 & -~872.02 \\ 
+				%		1.25 & -~0.00 & -~946.04 \\ 
+				%		1.50 & -~0.00 & -~951.57  \\ 
+				%		\hline
+				%	\end{tabular}
+				%	\caption{Bla}
+				%	\label{tab:update_smallres_tab}
+				%\end{table}
+
+				\begin{table}
+					\centering
+					\begin{tabular}[c]{| c || l | l || l | l || l | l |}
+						\hline
+						& Example I & & Example II & & Example III & \\
+						\hline
+						\hline
+						\textit{r} & $\Delta$ BA & $\Delta$ EL & $\Delta$ BA & $\Delta$ EL & $\Delta$ BA & $\Delta$ EL\\ 
+						\hline
+						0.50 & +~553.67 & -~105.49 & +~522.95 & -~436.26 & -~498.98 & -~538.08 \\ 
+						0.75 & +~673.40 & -~81.67 & +~697.73 & -~551.92 & -~346.16 & -~717.83  \\ 
+						1.00 & +~750.05 & -~33.56 & +~855.57 & -~631.92 & -~176.13 & -~872.02 \\ 
+						1.25 & +~713.19 & +~75.71 & +~912.43 & -~647.10 & -~0.00 & -~946.04 \\ 
+						1.50 & +~0.00 & +~282.64 & +~819.26 & -~545.96 & -~0.00 & -~951.57  \\ 
+						\hline
+					\end{tabular}
+					\caption{Changes in Bayes action (BA) and minimal expected loss (EL) for Bayesian updating examples I, II and III and respective actors with risk parameters \textit{r}.}
+					\label{tab:update_examples_all}
+				\end{table}
+
+				\begin{table}
+						\centering
+						\begin{tabular}[c]{| l | l | l |}
+							\hline
+							Risk factor \textit{r} & Shift in Bayes action & Change in expected loss \\ \hline
+							0.50 & -~498.98 & -~538.08 \\ 
+							0.75 & -~346.16 & -~717.83  \\ 
+							1.00 & -~176.13 & -~872.02 \\ 
+							1.25 & -~0.00 & -~946.04 \\ 
+							1.50 & -~0.00 & -~951.57  \\ 
+							\hline
+						\end{tabular}
+						\caption{Bla}
+						\label{tab:update_examples_all}
+				\end{table}
+			\subsection{General case results}
+
+			-abstract case: easy model construction and straightforward design of a loss function making basic assumptions and taking relative values that can simply be exchanged
+
+			-so representative in a "relative" way, mostly appropriate to illustrate principles and benefits of the methodology
+
+			-decision making/ estimation is defined by the design of the loss function which includes framework parameters which depend on the problem environment (e.g. market, technical constraints, etc.)
+
+			-similar actors but different behaviors concerning risk: different decisions but same general loss function path (just different minima of EL)
+
+			-additional information changes decision (BA) and EL
+
+			-magnitude of change varies not only with the nature of the information (magnitude of uncertainty reduction) but also with the risk parameter
+
+			-some actors might benefit more from some specific additional information than others
+
+			-quantifiable value of information??
 

### 2017-09-29 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Now progress with 3D model construction text

@@
 
-        \section{Bayesian analysis and decision theory}
+        \section{Bayesian analysis and decision theory}\label{sec:bayes}
 	    As implied by the name, the problems and reasoning behind decision-making are examined in the field of decision theory \citep{berger2013stat}.
@@
 
-		\section{Constructing a full 3D geological reservoir model}
+		\section{3D geological reservoir model}
+		The concept of the 1D geological model is to be transferred and extended to a 3D setting that incorporates a more complete geological reservoir system.
+A three-dimensional space allows not only for better consideration of stratigraphical aspects, but also for the inclusion of structural formations, in particular reservoir trap-defining features.\\
+		For the purpose of exploring the effect of the methods of Bayesian analysis and decision theory in this work, the model is nevertheless to be kept relatively simple.
+
 		- concept of 1D case in chapter 3 transferred to a 3D case of a structural geological model

### 2017-09-29 Risk analysis/Text/Layout/Layout/chapter3_results.tex
Now progress with 3D model construction text

@@
 			\end{figure}			
-			\begin{figure}[h]
+			\begin{figure}[p!]
 				\centering
@@
 				\caption{Plotting of expected loss realizations after including the risk factor $r$ in the loss function (Equation \ref{eq:LFR_final}) for actors with risk-affinities ranging from risk-averse ($r$ = 0.5 and 0.75), over risk-neutral ($r$ = 1), to risk-friendly ($r$ = 1.25 and $r$ = 1.5).}\label{fig:1D_LFR}
+				\centering
+				\includegraphics[width=1\textwidth]{Figures/update_moderate1.png}
+				\caption{Prior (A1) and posterior distributions (A2) of the layer boundary positions in depth and respective representations (A2, B2).}\label{fig:update_moderate1} 
 			\end{figure}
@@
 These parameters vary according to nature of the observations made.
-Using the principle of Bayesian inference as explained in chapter \ref{cha:met}, the model is updated and new posterior distributions for our true reservoir score are attained\\
+Using the principle of Bayesian inference as explained in Chapter \ref{sec:bayes}, the model is updated and new posterior distributions for our true reservoir score are attained\\
 			Different Bayesian updating cases with different sets of likelihoods are presented in the following, so that various possible results can be compared.
@@
 				\subsubsection{Updating case I: Moderately reinforcing information}
-				\begin{figure}[h]
-					\centering
-					\includegraphics[width=1\textwidth]{Figures/update_moderate1.png}
-					\caption{Prior (A1) and posterior distributions (A2) of the layer boundary positions in depth and respective representations (A2, B2).}\label{fig:update_moderate1} 
-				\end{figure}
-
-				In this first case, a normal distribution with a mean of 25~m and a standard deviation of 20~m is chosen to reflect the likelihood of the seal thickness.
-For the reservoir, a normal distribution with a mean of 180~m and a standard deviation of 60~m is chosen.
-Using these likelihoods to update the one-dimensional model, the uncertainty in the posterior probability distributions for the positions of layer boundaries in depth is reduced (see Figure \ref{fig:update_moderate1}).
-
-				Modeling the reservoir score is conducted as above, now based on these new distributions.
-In Figure \ref{fig:update_moderate2_3} A it can be recognized that the bulk of the distribution was shifted to the positive side of scores, while the peak at -~2000 was raised.
-The probability of scores between -~2000 and 0 decreased significantly, i.e.
-the true score is most likely either positive or -~2000 if it is negative.
-
 				\begin{figure}[h]
@@
 
-				The customized loss function (IV) is now applied on this updated distribution of the reservoir score.
-This is visualized in Figure \ref{fig:update_moderate2_3} in which the expected losses are compared before (B1) and after (B2) uncertainty reduction.
+				In this first case, a normal distribution with a mean of 25~m and a standard deviation of 20~m is chosen to reflect the likelihood of the seal thickness.
+For the reservoir, a normal distribution with a mean of 180~m and a standard deviation of 60~m is chosen.
+Using these likelihoods to update the 1D geological model, the uncertainty in the posterior probability distributions for the positions of layer boundaries in depth is reduced (see Figure \ref{fig:update_moderate1}).\\
+				Scoring is applied based on these new distributions.
+In Figure \ref{fig:update_moderate2_3}-A, it can be recognized that the bulk of the score distribution was shifted to the positive side of values, while the peak at -2000 was raised.
+The probability of scores between -2000 and 0 decreased significantly, i.e.
+the true score is most likely either positive or -2000 if it is negative.\\
+				Application of the custom loss function (Equation \ref{eq:LFR_final}) is visualized in Figure \ref{fig:update_moderate2_3}, in which the expected losses are compared before (B1) and after (B2) uncertainty reduction.
 It is observable, that by adding information about layer thickness likelihoods, Bayes actions are shifted relative to the nature of the information.
-In this case, the added data generally reinforces the likelihoods of the reservoir to be significantly thick.
-Information on the seal, however, based on a normal distribution around 25~m thickness, leaves uncertainty about the reliability of the seal, as the safety threshold is defined as 20~m.
+In this case, the added data generally reinforces the probability of the reservoir to be significantly thick.
+Information on the seal, however, based on a normal distribution around 25~m thickness, leaves uncertainty about the reliability of the seal, as the safety threshold is defined as 20~m.\\
+				Increased certainty about the reservoir thickness is sufficient to shift Bayes actions to higher estimates for all actors, but the most risk-averse one ($r$ = 1.5).
+These shifts are quantified in Table \ref{tab:update_examples_all}.
+According to these numbers, the risk-neutral actor's estimate is increased the most, the risk-friendliest actor's estimate the least.
+Expected losses are decreased for the risk-neutral and the two risk-friendlier individuals.
+It is clear that the expected loss was reduced the most for the risk-friendliest actor and it increased most significantly for the most risk-averse actor.\\
 				%\begin{table}
@@
 
-				Increased certainty about the reservoir thickness is sufficient to shift Bayes actions to higher estimates for all actors, but the most risk-averse one.
-These shifts are quantified in Table \ref{tab:update_examples_all}.
-According to these numbers, the risk-neutral actor's estimate is increased the most.
-Expected losses decreased for the risk-neutral and the two risk-friendlier individuals.
-It is clear that the expected loss was reduced the most for the risk-friendliest actor and it increased most significantly for the most risk-averse actor.
-
 				\subsubsection{Updating case II: Likely reliable seal}
-				In this second case, the reservoir thickness likelihood is defined in the same way as before.
-For the seal, a mean of 50~m with a standard deviation of 20~m is chosen, favoring the likelihood of a reliable seal relative to the threshold of 20~m thickness.
-This results in the score distribution depicted in Figure \ref{fig:update_goodseal2_3}.
-The bulk of the distribution is narrowed on the positive side of estimates.
-Mean and median are clearly shifted to higher values as well.
-The "seal failure" peak at -~2000 is significantly.
-decreased.
-
-				Applying the loss function on this new score distribution results in the realizations of expected loss illustrated in Figure \ref{fig:update_goodseal2_3}.
-Bayes actions are shifted clearly to higher estimates and expected losses of these minima are significantly reduced for all actors.
-This is quantified in Table \ref{tab:update_examples_all}.
-According to these values, the risk-neutral to risk-averse individuals seem to profit the most.
-
 				\begin{figure}[h]
@@
 				\end{figure}
-
+				In this second case, the reservoir thickness likelihood is defined in the same way as in case I (mean = 180~m, standard deviation = 60~m).
+For the seal, a mean of 50~m with a standard deviation of 20~m is chosen, favoring the likelihood of a reliable seal relative to the threshold of 20~m thickness.
+Respective score results are depicted in Figure \ref{fig:update_goodseal2_3}-A.
+The bulk of the distribution is narrowed on the positive side of estimates.
+Mean and median are clearly shifted to higher values.
+The "seal failure peak" at -2000 is significantly decreased.\\
+				Applying the custom loss function on this new score distribution results in the realizations of expected loss illustrated in Figure \ref{fig:update_goodseal2_3}.
+Bayes actions are shifted clearly to higher estimates and expected losses of these minima are significantly reduced for all actors.
+This is quantified in Table \ref{tab:update_examples_all}.
+According to these values, the risk-neutral to risk-averse individuals seem to profit the most, due to a large change in estimate value and decreased expected loss.
+Compared to case I, the estimate shifts for both risk-friendlier actors remain approximately the same, expected losses, however, are lowered much more significantly.\\
 				%\begin{table}
@@
 				\subsubsection{Updating case III: Safe seal but likely subpar reservoir thickness}
-				In this third example, seal safety is ensured by using a mean of 70~m with a standard deviation of 10~m for the seal thickness likelihood.
-However, the new observations are assumed to provide information about the reservoir that makes it likely to be thinner as expected.
-Reservoir thickness likelihood is assigned a mean of 100~m and a standard deviation of 40~m.
-
-				The subsequent score distribution is depicted in Figure \ref{fig:update_smallres2_3}.
-It can be seen that while the whole distribution is narrowed, it also is shifted to the left, to lower and negative estimates.
-Mean and median are almost the same.
-As seal reliability is practically guaranteed, the peak at -~2000~m vanishes.
-
-				Based on this probability distribution of the reservoir score, Bayes actions are shifted to lower estimates for all actors (see Figure \ref{fig:update_smallres2_3}).
-In fact, also but the most risk-friendly individual (\textit{r} = 0.5) find their Bayesian estimator to be zero now, i.e.
-project development is deemed to be too risky to them.
-Furthermore, the spread of expected loss values around the minima for all actors is diminished (the expected loss values of the Bayes actions are now much closer to each other).
-
-				Quantified changes in the position and the values of the Bayes actions are comparable to the results in example II (there is no shift for risk-averse actors, as they already found their best estimates to be zero).
-Most notable is the large reduction of expected loss in general (see Table \ref{tab:update_examples_all}).
-
 				\begin{figure}[h]
@@
 				\end{figure}
-
+				In this third case, seal safety is ensured by using a mean of 70~m with a standard deviation of 10~m for the seal thickness likelihood.
+However, the new observations are assumed to provide information about the reservoir unit that makes it likely to be thinner than expected.
+Reservoir thickness likelihood is assigned a mean of 100~m and a standard deviation of 40~m.\\
+				The subsequent score distribution is depicted in Figure \ref{fig:update_smallres2_3}.
+It can be seen that while the whole distribution is narrowed, it also is shifted to the left, to lower and negative estimates.
+Mean and median are almost the same.
+As seal reliability is practically guaranteed, the peak at -2000~m vanishes.\\
+				Based on this reservoir score probability distribution, Bayes actions are shifted to lower estimates for all actors using the custom loss function (see Figure \ref{fig:update_smallres2_3}).
+In fact, all but the most risk-friendly individual (\textit{r}~=~0.5) find their Bayesian estimator to be zero now, i.e.
+project development is deemed to be too risky to them.
+Furthermore, the spread of expected loss values around the minima for all actors is diminished.
+In other words, the expected loss values of the Bayes actions are now much closer to each other.
+Notable is the large reduction of expected loss in general.
+Quantified changes in the positions and values of the Bayes actions are listed in Table \ref{tab:update_examples_all}.
+There is no shift in estimate for risk-averse actors, as they already found their best estimates to be zero before updating.\\
 				%\begin{table}
@@
 					\end{tabular}
-					\caption{Changes in Bayes action (BA) and minimal expected loss (EL) for Bayesian updating examples I, II and III and respective actors with risk parameters \textit{r}.}
+					\caption{Changes in Bayes action (BA) and minimal expected loss (EL) for Bayesian updating cases I, II and III and respective actors with risk parameters $r$.}
 					\label{tab:update_examples_all}
 				\end{table}
+				%\begin{table}
+				%		\begin{tabular}[c]{| l | l | l |}
+				%			\hline
+				%			Risk factor \textit{r} & Shift in Bayes action & Change in expected loss \\ \hline
+				%			0.50 & -~498.98 & -~538.08 \\ 
+				%			0.75 & -~346.16 & -~717.83  \\ 
+				%			1.00 & -~176.13 & -~872.02 \\ 
+				%			1.25 & -~0.00 & -~946.04 \\ 
+				%			1.50 & -~0.00 & -~951.57  \\ 
+				%			\hline
+				%		\end{tabular}
+				%		\caption{Bla}
+				%		\label{tab:update_examples_all}
+				%\end{table}
 
-				\begin{table}
-						\centering
-						\begin{tabular}[c]{| l | l | l |}
-							\hline
-							Risk factor \textit{r} & Shift in Bayes action & Change in expected loss \\ \hline
-							0.50 & -~498.98 & -~538.08 \\ 
-							0.75 & -~346.16 & -~717.83  \\ 
-							1.00 & -~176.13 & -~872.02 \\ 
-							1.25 & -~0.00 & -~946.04 \\ 
-							1.50 & -~0.00 & -~951.57  \\ 
-							\hline
-						\end{tabular}
-						\caption{Bla}
-						\label{tab:update_examples_all}
-				\end{table}
-			\subsection{General case results}
+				NEED TO INCLUDE CLEAR BA VALUES BEFORE AND AFTER!!! NOT VISIBLE IN PLOTS.
 
-			-abstract case: easy model construction and straightforward design of a loss function making basic assumptions and taking relative values that can simply be exchanged
-
-			-so representative in a "relative" way, mostly appropriate to illustrate principles and benefits of the methodology
-
-			-decision making/ estimation is defined by the design of the loss function which includes framework parameters which depend on the problem environment (e.g. market, technical constraints, etc.)
-
-			-similar actors but different behaviors concerning risk: different decisions but same general loss function path (just different minima of EL)
-
-			-additional information changes decision (BA) and EL
-
-			-magnitude of change varies not only with the nature of the information (magnitude of uncertainty reduction) but also with the risk parameter
-
-			-some actors might benefit more from some specific additional information than others
-
-			-quantifiable value of information??
+			\subsection{General 1D model updating results}
+			It is already shown by these three abstract 1D modeling cases, how Bayesian updating and change in uncertainties affects the way differently risk-affine decision-makers choose best estimates based on the custom loss function (Equation \ref{eq:LFR_final}).
+As can be expected intuitively, reassurance of positive scores, for example by narrowing of the distribution of scores on the side of positive estimates or by reducing the occurrence of negative extrema, leads to shifts towards higher Bayes action estimates for all actors.
+Conversely, actors are discouraged to expect gains when facing a higher probability of negative scores.
+Unsurprising is also the inverse behavior of risk-averse and risk-friendly individuals.
+However, it is most notable that a reduction of uncertainty in any case leads to a diminished spread of Bayes actions between all actors.
+In other words, given better information about the true value of the reservoir, different actors come to more similar decisions.
+Given perfect information, all actors would presumably choose the same estimate, which would equal the true score.
 

### 2017-09-29 Risk analysis/Text/Layout/Layout/chapter4_discussion.tex
Now progress with 3D model construction text

--- 2017-04-18 Risk analysis/Text/Layout/Layout/chapter3_1Dcase.tex
@@
-\chapter{One-dimensional reservoir case}
+	\chapter{Discussion}\label{cha:discussion}
 
-For simple understanding and working out the application of the methods, first, an abstract one-dimensional reservoir case is presented in this chapter.
-The underlying model and basic approach are inherited from \citet{delaVarga2016}.
-Parameters were adapted to more appropriately represent a reasonable geological petroleum system, consisting of a reservoir with overlying seal in the subsurface.
-In doing this, a certain economic significance is ascribed to the model and further relevant questions can be derived.
-Regarding the petroleum sector, the problem of interest is most commonly one of estimating the volume of resources or monetary value contained in an oil or gas reservoir.
-Limiting the model to only one dimension and a small number of uncertain parameters allows for a relatively straightforward and simplified approach to assessing an abstract type of value for a reservoir and designing a respective loss function for value estimation.
-A step by step derivation of such as case follows below.
+	1D CASE:
 
-	\section{Constructing the one-dimensional model}\label{sec:1D_construction}
+				-abstract case: easy model construction and straightforward design of a loss function making basic assumptions and taking relative values that can simply be exchanged
 
-	\citet{delaVarga2016} constructed a simple geological model using three uncertain positions in vertical one-dimensional space, marking hypothetical boundaries of layers in a subsurface column.
-The location probabilities for these points are defined by sampling from normal distributions.
-Standard deviations of these distributions increase with depth, representing an increase in uncertainty.
-For an approximate representation of a petroleum system, the distribution means were set to depths of 2000 m (seal top), 2050 m (reservoir top) and 2200 m (reservoir bottom).
-These points confine two layers in the middle, from which the upper one can be labeled as seal and the lower one as reservoir.
-The resulting model with its possible layer boundary locations is illustrated in Figure \ref{fig:1D_model}.
+				-so representative in a "relative" way, mostly appropriate to illustrate principles and benefits of the methodology
 
-	\begin{figure}[h]
-		\centering
-		\includegraphics[width=1\textwidth]{Figures/1D_model.png}
-		\caption{Probability distributions for positions of layer boundaries in the subsurface and a respective representation using lines.}\label{fig:1D_model}
-	\end{figure}
+				-decision making/ estimation is defined by the design of the loss function which includes framework parameters which depend on the problem environment (e.g. market, technical constraints, etc.)
 
-	\section{Assessing reservoir quality using scores}\label{sec:scoring}
+				-similar actors but different behaviors concerning risk: different decisions but same general loss function path (just different minima of EL)
 
-	The next step is to find a way to assess the quality of the reservoir from a petroleum industry perspective in such a simplified model.
-What can be deduced from the distributions of layer boundary locations, is the thickness of the seal and the reservoir, as well as the depth of both, again as probability distributions.
-To assess the reservoir quality in an abstract way, it can be assigned a score.
-This score is made dependent on the three uncertain parameters (1) reservoir thickness, (2) reservoir top depth and (3) seal thickness (see Figure \ref{fig:3_parameters}).
+				-additional information changes decision (BA) and EL
 
-	\begin{figure}[h]
-		\centering
-		\includegraphics[width=1\textwidth]{Figures/3_parameters.png}
-		\caption{Probability distributions for the three parameters used for modeling reservoir score.}\label{fig:3_parameters}
-	\end{figure}
+				-magnitude of change varies not only with the nature of the information (magnitude of uncertainty reduction) but also with the risk parameter
 
-	\subsection{Two parameters scoring}
+				-some actors might benefit more from some specific additional information than others
 
-	Assuming that reservoir thickness is a simplified indicator for the extractable oil or gas and thus value in place, a gain in score can be correlated with increase in thickness.
-Here, two score points are assigned to one meter of thickness.
-Increasing costs of drilling are indicated by increasing depth of the reservoir top.
-Consequently, one negative score point is ascribed to every meter in depth.
-Samples from the probability distributions of these two parameters are drawn to model the true score of the reservoir (depth scores are subtracted from reservoir thickness scores).
-For the data used here, it can be seen in Figure \ref{fig:score_results1}-A, that the results of modeling with only these two parameters, are represented by an approximately normal distribution.
-The score is negative in about 17\% of the cases.
-Mean and median are about the same.
-
-	\subsection{Three parameters scoring}
-
-	Next, influence by the third parameter seal thickness is included.
-Score points are not added or subtracted by this parameter directly.
-Instead, a threshold for seal reliability is defined beforehand.
-Here it is set at 20 m thickness.
-If the seal thickness falls below this threshold, it is assumed that the seal fails completely and thus all the potential value (positive score) of the reservoir is lost, while costs of depth (negative score) remain.
-Thus, a condition to check whether the seal is reliable is now included in the model.
-Results are visualized in Figure \ref{fig:score_results1}-B.
-The main distribution is not changed significantly, except for a striking peak of probability at the possibility of a score of -2000.
-Furthermore, mean and median have been shifted to lower values and are now found further apart.
-
-	\begin{figure}[h]
-		\centering
-		\includegraphics[width=1\textwidth]{Figures/score_results1.png}
-		\caption{Posterior probability distributions from modeling scores using two (A) and three parameters (B).}\label{fig:score_results1}
-	\end{figure}
-
-	\section{Designing a case-specific loss function}
-
-	Now that a distribution of reservoir score probabilities has been modeled, a loss function for estimation of the true score value can be developed.
-
-	\subsection{Testing standard loss functions}
-
-	Some standard loss function were presented in chapter \ref{sec:loss}.
-For this case, the absolute-error loss and the squared-error loss function were considered as starting points, to design a more case-specific loss function.
-	Calculating the expected loss for estimates ranging from -3000 to 6000 based on these two standard loss function results in the graphs depicted in Figure \ref{fig:abs_sqr_LF}.
-It can be observed that the median of the distribution coincides with the minimum of the absolute-error and the mean with the minimum of the squared-error expected loss.
-
-	\begin{figure}[h]
-		\centering
-		\includegraphics[width=1\textwidth]{Figures/abs_sqr_LF.png}
-		\caption{Expected loss based on the standard absolute-error (A) and squared-error (B) loss functions applied on the probability distribution resulting from \ref{sec:scoring} and for estimates ranging from -3000 to 6000.}\label{fig:abs_sqr_LF}
-	\end{figure}
-
-	\subsection{Adaptions for loss function customization}
-
-	These standard loss functions provide objectively good estimators minimizing expected loss.
-Due to their symmetric properties, both will always give the median or mean of the underlying distribution as minimizing estimator respectively.
-However, assigning an economic notion to our model and assuming the case of an actor or decision maker in any field, naturally necessitates the consideration of preferences, interests and the overall subjective perspective such an individual or for example a company might have.
-Further constraints, properties and factors can also be specific to the field, industry or generally to the problem at hand.
-Consequently, the design of a more specific non-standard and possibly asymmetric loss function might be required, so that an adapted Bayesian estimator can be found.
-One that includes subjective aspects and difference in weighting of particular gains or losses, arising from an actor's inherent preferences and the environment in which the actor has to estimate or make a decision.
-In the face of several uncertain parameters, a perfectly true estimate is virtually unattainable.
-However, an attempt can be made, to design a customized loss function that returns a Bayesian estimator involving the least bad consequences for an individual in a specific environment.
-Regarding the petroleum system case modeled above, such an attempt is made and explained step by step in the following.
-
-	For the purpose of estimation, it makes sense that one of the standard loss functions is chosen as a basis and a customized loss function is developed from there.
-The absolute-error loss seems most appropriate for this case of petroleum reservoir quality estimation.
-As stated above, the reservoir score is an abstract and simplified way to reflect a value contained in the reservoir.
-Ideally, an actor would like to know the exact true score, so that investments or resources can be allocated appropriately, in order to acquire economic gains.
-This allocation is the decision to be made or action to be taken.
-Deviations from the unknown true score in the form of over- and underestimation bring about an error and loss accordingly.
-In this case, there is no reason for loss to increase exponentially with distance from the true value.
-It is assumed that investments increase linearly with linear growth in the value of the resource.
-For this reason, the absolute-error loss function is favored over the squared-error loss function in this case.
-Some customization steps are taken below, based on mostly logical case-specific assumptions.
-
-	\begin{enumerate}
-		\item Step I: The standard symmetrical absolute-error loss function is chosen as a starting point for further customization steps:
-
-		\begin{equation}
-		L(\theta,\hat{\theta}) = |\theta - \hat{\theta}|
-		\end{equation}
-
-		\item Step II: Considering the development of a petroleum reservoir, it can be assumed that over-investing is worse than under-investing.
-Overestimating the size of a reservoir might for example lead to the installation of equipment or facilites that are actually redundant or unnecessary.
-This would come with additional unrecoverable expenditures.
-Consequences from underestimating, however, may presumably be easier to resolve.
-Additional equipment can often be installed later on.
-Hence, overestimation is weighted stronger in our loss function by multiplying the error with an overestimation factor \textit{a} (= 1.25):
-
-		\begin{equation}\label{eq:LF_II}
-		L(\theta,\hat{\theta}) = |(\theta-\hat{\theta})|*a
-		\end{equation}
-
-		\item Step III: The worst case for any project would be that its development is set into motion, expecting a gain, only to discover later that the value in the reservoir does not cover the costs of realizing the project, resulting in an overall loss.
-A petroleum system might also turn out to be a complete failure, containing no value at all, although the actor's estimate indicated the opposite.
-Here, this is referred to as a worst case or fatal overestimation.
-A positive score is estimated, but the true score is zero or negative.
-This is worse than the "normal" non-fatal overestimation, where both values are positive and a net gain is still achieved, which is only smaller then the best possible gain of expecting the true score.
-Fatal overestimation is included in the loss function by using another weighting factor \textit{b} (= 2) that replaces \textit{a}:
-
-		\begin{equation}\label{eq:LF_III}
-		L(\theta,\hat{\theta}) = |(\theta-\hat{\theta})|*b
-		\end{equation}
-
-		(In other words: Worst case or fatal overestimation is twice as bad as simple underestimation.)
-
-		\item Step IV: A worst case or fatal underestimation can also be derived from the idea of estimating a zero or negative score, when the true score is actually positive.
-This is assumed to be worse than non-fatal overestimation, but clearly better than fatal overestimation.
-No already owned resources are wasted, it is only the potential value that is lost, i.e.
-opportunity costs that arise from completely discarding a reservoir with a potential gain equal to the positive true score.
-Fatal underestimation is weighted using a third factor \textit{c}:
-
-		\begin{equation}\label{eq:LF_IV}
-		L(\theta,\hat{\theta}) = |(\theta-\hat{\theta})|*c
-		\end{equation}
-
-	\end{enumerate}
-
-	Combining these adaption steps and the conditions defined in them, results in the following customized loss function:
-
-	\begin{equation}\label{eq:LF_final}
-	L(\theta,\hat{\theta}) =
-	\begin{cases}
-	|\theta - \hat{\theta}|, & \text{for } 0<\hat{\theta}<\theta  \\
-	|\theta-\hat{\theta}|*a, & \text{for } 0<\theta<\hat{\theta} \\
-	|\theta-\hat{\theta}|*b, & \text{for } \theta\leq0<\hat{\theta} \\
-	|\theta-\hat{\theta}|*c, & \text{for } \hat{\theta}\leq0<\theta 
-	\end{cases},
-	\text{ with } a,b,c \in \mathbb{Q}
-	\end{equation}
-
-	Realizations of the four adaption steps are depicted in the plots in Figure \ref{fig:LF_4steps}.
-It is important to note that the weighting factors defined above can take basically any numerical values but should be chosen in a way that they appropriately represent the framework conditions of the problem.
-Here, for example, it is assumed that normal overestimation is 25\% (\textit{a} = 1.25), fatal overestimation 100\% (\textit{b} = 2) and fatal underestimation 50\% (\textit{c} = 1.5) worse than normal underestimation.
-
-	\begin{figure}[h]
-		\centering
-		\includegraphics[width=1\textwidth]{Figures/LF_4steps.png}
-		\caption{The single steps of customizing the loss function are depicted in plots I - IV.}\label{fig:LF_4steps}
-	\end{figure}
-
-	In Figure \ref{fig:LF_4steps} it can be observed, that assigning a stronger weight on overestimation (plot II) steepens the curve on the right hand side and shifts the minimum to the left, i.e.
-to a lower estimate.
-Using \textit{a} = 1.25, the Bayes action changes from the median, to a value close to the mean of the distribution.
-The shift and steepening are significantly reinforced by the introduction of fatal overestimation (plot III).
-With \textit{b} = 2, the Bayes action drops to the zero score estimate.
-It can also be noted, that by defining a condition dependent on the algebraic sign of the values, according to which only losses for positive estimates are multiplied by \textit{b}, a distinct jump appears at the zero score boundary.
-Due to a similar condition, the same effect is observed on the negative side of estimate values, where the curve has also been steepened, after including fatal underestimation (plot IV).
-This comes with a shift of the minimum towards positive values.
-It is also to be noted that with every customization step, the overall expected loss is increased.
-
-	\begin{figure}[h]
-		\centering
-		\includegraphics[width=1\textwidth]{Figures/LF4_det_values.png}
-		\caption{Loss based on the customized loss function (IV) for determined true scores of -750, 0 and 750.}\label{fig:LF4_det_values}
-	\end{figure}
-
-	The implementation of this customized loss function (IV) using single determined values for the true score is plotted in Figure \ref{fig:LF4_det_values}.
-This helps to clarify the way real losses result for each guess, relative to given true score values.
-The expected loss is acquired by arithmetically averaging such loss realizations based on the true score probability distribution by using formula \ref{eq:ExpectedLoss2}.
-
-	It has to be emphasized that this is just one possible proposal for loss function customization.
-There exists not one perfect design for such a case.
-Slight to strong changes can already be implemented by simply varying the values of the weighting factors \textit{a, b} and \textit{c}.
-Fundamentally different loss functions can also be based on a significantly different mathematical structure.
-Loss functions are customized regarding the problem environment and according to the to the subjective needs and objectives of an actor.
-Thus, they are mostly defined by the actor expressing his perspective.
-Changes in the individual's perception and viewpoint might lead to further customization needs even later on.
-(Especially considering individual persons as actors, psychological aspects may play a significant role.)
-
-	(Estimate or prediction? True value only known if a "yes"-action (development) is taken.)
-
-	\subsection{Including different risk-affinities in the loss function}
-
-	One can assume that several actors in one sector or decision environment may have the same general loss function, but different affinities concerning risks.
-This might be based for example on the different psychological factors or economic philosophies followed by companies.
-It might also be based on budgets and options such actors have available.
-An intuitive example is the comparison of a small and a large company.
-A certain false estimate or error might have a significantly stronger impact on a company which has a a generally lower market share and only few projects, than on a larger company which might possess a higher financial flexibility and for which one project is only one of many development options in a portfolio.
-
-	In the following, the loss function is further adapted to consider different risk-affinities of several actors.
-Representing risk behavior in a loss function can also be done in different ways and regarding different types of risks.
-Here, bidding lower is considered the cautious, risk-averse option, as smaller losses can be expected from underestimating.
-Guessing higher is deemed riskier.
-Losses from overestimation are greater.
-However, bidding correctly on a higher value, will also return a greater gain.
-It is assumed that risk-friendly actors care less about fatal underestimation, i.e.
-they will rather develop a project than discard it.
-In the loss function, risk is simply included using a risk factor \textit{r} which alters the weighting factors \textit{a, b} and \textit{c} respectively:
-
-	\begin{equation}\label{eq:LFR_final}
-	L(\theta,\hat{\theta}) =
-	\begin{cases}
-	|\theta - \hat{\theta}|, & \text{for } 0<\hat{\theta}<\theta  \\
-	|\theta-\hat{\theta}|*(a*r), & \text{for } 0<\theta<\hat{\theta} \\
-	|\theta-\hat{\theta}|*(b*r), & \text{for } \theta\leq0<\hat{\theta} \\
-	|\theta-\hat{\theta}|*(c*(r^{-0.5})), & \text{for } \hat{\theta}\leq0<\theta 
-	\end{cases},
-	\text{ with } a,b,c,r \in \mathbb{Q}
-	\end{equation}
-
-	According to this, for \textit{r} = 1 the risk-neutral loss function is returned, since \textit{a, b} and \textit{c} are not altered.
-For \textit{r} $<$ 1, the weight on overestimating is reduced and increased for fatal underestimation.
-This represents a more risk-friendly actor that is willing to bid on a higher estimate to attain a greater gain.
-For \textit{r} $>$ 1, the overestimation weight is increased, the underestimation weight decreased in the loss function and respectively more risk-averse actors are prompted to bid on lower estimates.
-
-	The factor \textit{r} can take basically any positive values.
-However, since risk-neutrality is expressed by\textit{r} = 1, values 0 $<$ \textit{r} $<$ 2 are considered to be the most appropriate choices to represent both sides of risk-affinity equally.
-
-	Implementing it as a factor \textit{r} leads to different steepnesses of the plotted curves depicting loss and expected loss.
-The effect on the latter is visualized in Figure \ref{fig:LFR}, where 0.5, 0.75, 1, 1.25 and 1.5 were chosen as values for \textit{r}.
-	It can be observed that the minima for expected loss, i.e.
-the Bayesian estimators for differently risk-affine actors, are located at different estimates.
-Mean and median are clearly surpassed by the best estimate of the most risk-friendly actor (\textit{r} = 0.5), while for the most risk-averse actors (\textit{r} 0.5 and \textit{r} = 0.75), the Bayes action equals a zero score estimate (and thus the decision to take on action).
-In Figure \ref{fig:LFR} it can also be recognized that the expected loss is generally lower for risk-friendlier actors (on the side of positive estimates).
-
-	\begin{figure}[h]
-		\centering
-		\includegraphics[width=1\textwidth]{Figures/LFR.png}
-		\caption{Plotting of expected loss realizations after including the risk factor r in the loss function (IV) for actors with risk-affinities ranging from risk-averse (r = 0.5 and 0.75), over risk-neutral (r = 1), to risk-friendly (r = 1.25 and r = 1.5).}\label{fig:LFR}
-	\end{figure}
-
-	\section{Updating the model with additional information in the form of thickness likelihoods}
-
-	\citet{delaVarga2016} made use of Bayesian inference to reduce the uncertainty in this type of one-dimensional model. The same is conducted here in the following.
-
-	The probability distributions for the location of the layer boundaries are treated as priors.
-Now it is assumed that new observations have been made, providing additional information on the likelihoods of the thicknesses of the two layers.
-Likelihood functions for reservoir and seal thicknesses are introduced in the form of normal distributions, defined by means and standard deviations.
-These parameters vary according to nature of the observations made.
-Using the principle of Bayesian inference as explained in \ref{cha:met}, the model is updated and new posterior distributions for our true reservoir score are attained.
-
-	- VOI Quantification ???
-
-	Different examples with different sets of likelihoods are conducted in the following, so that various possible results can be compared.
-Each take the layer boundaries defined in  \ref{sec:1D_construction} as priors and use Bayesian inference with data defined below for each example.
-
-	\subsection{Updating example I (moderate reinforcement)- Seal: mean = 25~m, std = 20~m - Reservoir: mean = 180~m, std = 60~m}
-
-	\begin{figure}[h]
-		\centering
-		\includegraphics[width=1\textwidth]{Figures/update_moderate1.png}
-		\caption{...}\label{fig:update_moderate1} 
-	\end{figure}
-
-	In this first case, a normal distribution with a mean of 25~m and a standard deviation of 20~m is chosen to reflect the likelihood of the seal thickness.
-For the reservoir, a normal distribution with a mean of 180~m and a standard deviation of 60~m is chosen.
-Using these likelihoods to update the one-dimensional model, the uncertainty in the posterior probability distributions for the positions of layer boundaries in depth is reduced (see Figure \ref{fig:update_moderate1}).
-
-	Modeling the reservoir score is conducted as above, now based on these new distributions.
-In Figure \ref{fig:update_moderate2} it can be recognized that the bulk of the distribution was shifted to the positive side of scores, while the peak at -~2000 was raised.
-The probability of scores between -~2000 and 0 decreased significantly, i.e.
-the true score is most likely either positive or -~2000 if it is negative.
-
-	\begin{figure}[h]
-		\centering
-		\includegraphics[width=1\textwidth]{Figures/update_moderate2.png}
-		\caption{...}\label{fig:update_moderate2} 
-	\end{figure}
-
-	\begin{figure}[h]
-		\centering
-		\includegraphics[width=1\textwidth]{Figures/update_moderate3.png}
-		\caption{...}\label{fig:update_moderate3} 
-	\end{figure}
-
-	The customized loss function (IV) is now applied on this updated distribution of the reservoir score.
-This is visualized in Figure \ref{fig:update_moderate3} in which the expected losses are compared before (B1 )and after (B2) uncertainty reduction.
-It is observable, that by adding information about layer thickness likelihoods, Bayes actions are shifted relative to the nature of the information.
-In this case, the added data generally reinforces the likelihoods of the reservoir to be significantly thick.
-Information on the seal, however, based on a normal distribution around 25~m thickness, leaves uncertainty about the reliability of the seal, as the safety threshold is defined as 20~m.
-
-	\begin{table}
-		\centering
-		\begin{tabular}[c]{| l | l | l |}
-			\hline
-			Risk factor \textit{r} & Shift in Bayes action & Change in expected loss \\ \hline
-			0.50 & +~553.67 & -~105.49 \\ 
-			0.75 & +~673.40 & -~81.67  \\ 
-			1.00 & +~750.05 & -~33.56 \\ 
-			1.25 & +~713.19 & +~75.71 \\ 
-			1.50 & +~0.00 & +~282.64  \\ 
-			\hline
-		\end{tabular}
-		\caption{Bla}
-		\label{tab:update_moderate_tab}
-	\end{table}
-
-	Increased certainty about the reservoir thickness is sufficient to shift Bayes actions to higher estimates for all actors, but the most risk-averse one.
-These shifts are quantified in Table \ref{tab:update_moderate_tab}.
-According to these numbers, the risk-neutral actor's estimate is increased the most.
-Expected losses decreased for the risk-neutral and the two risk-friendlier individuals.
-It is clear that the expected loss was reduced the most for the risk-friendliest actor and it increased most significantly for the most risk-averse actor.
-
-	\subsection{Updating example II (reliable seal) - Seal: mean = 50 m, std = 20 m - Reservoir: mean = 180 m, std = 60 m}
-
-	In this second case, the reservoir thickness likelihood is defined in the same way as before.
-For the seal, a mean of 50~m with a standard deviation of 20~m is chosen, favoring the likelihood of a reliable seal relative to the threshold of 20~m thickness.
-This results in the score distribution depicted in \ref{fig:update_goodseal2}.
-The bulk of the distribution is narrowed on the positive side of estimates.
-Mean and median are clearly shifted to higher values as well.
-The "seal failure" peak at -~2000 is significantly.
-decreased.
-
-	\begin{figure}[h]
-		\centering
-		\includegraphics[width=1\textwidth]{Figures/update_goodseal2.png}
-		\caption{...}\label{fig:update_goodseal2} 
-	\end{figure}
-
-	Applying the loss function on this new score distribution results in the realizations of expected loss illustrated in \ref{fig:update_goodseal3}.
-Bayes actions are shifted clearly to higher estimates and expected losses of these minima are significantly reduced for all actors.
-This is quantified in \ref{tab:update_goodseal_tab}.
-According to these values, the risk-neutral to risk-averse individuals seem to profit the most.
-
-	\begin{figure}[h]
-		\centering
-		\includegraphics[width=1\textwidth]{Figures/update_goodseal3.png}
-		\caption{...}\label{fig:update_goodseal3} 
-	\end{figure}
-
-	\begin{table}
-		\centering
-		\begin{tabular}[c]{| l | l | l |}
-			\hline
-			Risk factor \textit{r} & Shift in Bayes action & Change in expected loss \\ \hline
-			0.50 & +~522.95 & -~436.26 \\ 
-			0.75 & +~697.73 & -~551.92  \\ 
-			1.00 & +~855.57 & -~631.92 \\ 
-			1.25 & +~912.43 & -~647.10 \\ 
-			1.50 & +~819.26 & -~545.96  \\ 
-			\hline
-		\end{tabular}
-		\caption{Bla}
-		\label{tab:update_goodseal_tab}
-	\end{table}
-
-	\subsection{Updating example III (safe seal, but reservoir under average) - Seal: mean = 70 m, std = 10 m - Reservoir: mean = 100 m, std = 40 m}
-
-	In this third example, seal safety is ensured by using a mean of 70~m with a standard deviation of 10~m for the seal thickness likelihood.
-However, the new observations are assumed to provide information about the reservoir that makes it likely to be thinner as expected.
-Reservoir thickness likelihood is assigned a mean of 100~m and a standard deviation of 10~m.
-
-	The subsequent score distribution is depicted in \ref{fig:update_smallres2}.
-It can be seen that while the whole distribution is narrowed, it also is shifted to the left, to lower and negative estimates.
-Mean and median are almost the same.
-As seal reliability is practically guaranteed, the peak at -~2000~m vanishes.
-
-	Based on this probability distribution of the reservoir score, Bayes actions are shifted to lower estimates for all actors (\ref{fig:update_smallres3}).
-In fact, also but the most risk-friendly individual (\textit{r} = 0.5) find their Bayesian estimator to be zero now, i.e.
-project development is deemed to be too risky to them.
-Furthermore, the spread of expected loss values around the minima for all actors is diminished (the expected loss values of the Bayes actions are now much closer to each other).
-
-	Quantified changes in the position and the values of the Bayes actions are comparable to the results in example II (there is no shift for risk-averse actors, as they already found their best estimates to be zero).
-Most notable is the large reduction of expected loss in general (see \ref{tab:update_smallres_tab}).
-
-	\begin{figure}[h]
-			\centering
-			\includegraphics[width=1\textwidth]{Figures/update_smallres2.png}
-			\caption{...}\label{fig:update_smallres2} 
-	\end{figure}
-
-	\begin{figure}[h]
-				\centering
-				\includegraphics[width=1\textwidth]{Figures/update_smallres3.png}
-				\caption{...}\label{fig:update_smallres3} 
-	\end{figure}
-
-	\begin{table}
-		\centering
-		\begin{tabular}[c]{| l | l | l |}
-			\hline
-			Risk factor \textit{r} & Shift in Bayes action & Change in expected loss \\ \hline
-			0.50 & -~498.98 & -~538.08 \\ 
-			0.75 & -~346.16 & -~717.83  \\ 
-			1.00 & -~176.13 & -~872.02 \\ 
-			1.25 & -~0.00 & -~946.04 \\ 
-			1.50 & -~0.00 & -~951.57  \\ 
-			\hline
-		\end{tabular}
-		\caption{Bla}
-		\label{tab:update_smallres_tab}
-	\end{table}
-
-	\section{General case results}
+				-quantifiable value of information??

### 2017-10-02 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Gradient function to find min/max y X and Y directions

@@
 		The concept of the 1D geological model is to be transferred and extended to a 3D setting that incorporates a more complete geological reservoir system.
-A three-dimensional space allows not only for better consideration of stratigraphical aspects, but also for the inclusion of structural formations, in particular reservoir trap-defining features.\\
+A three-dimensional space allows not only for better consideration of stratigraphical aspects, but also for the inclusion of structural formations, in particular reservoir trap-defining features.
+
+		\subsection{Design of the 3D geological reservoir model}\label{sec:3D_design}
 		For the purpose of exploring the effect of the methods of Bayesian analysis and decision theory in this work, the model is nevertheless to be kept relatively simple.
+Stratigraphically, it is designed to include one main reservoir unit (sandstone), one main seal unit (shale), an underlying basement through which hydrocarbon fluids could have flown upwards and an overlying formation that is assumed to be permeable, so that hydrocarbons can leak upwards.
+Structurally, it is constructed to feature an anticlinal dome fold that is displaced by a normal fault.
+All layers are tilted so that they dip in the opposite direction of the fault plane dip.
+The original concept of this model is designed in a way, that a potential hydrocarbon trap is formed in the reservoir rock enclosed by the deformed and tilted seal and normal fault.\\
+		This trap is defined as the central feature of economic interest.
 
-		- concept of 1D case in chapter 3 transferred to a 3D case of a structural geological model
-		- same again here: relatively simple model including some important typical geological and structural features and giving it an economic meaning
-		- alternating layers of sandstones and shales, so typical reservoir and seal formations
-		- structurally defined by a tilted dome-formation (folding along in the direction of two perpendicular axes) which is displaced by a standard normal fault
-		- these features are arranged in a way, that they result in the formation of a trap, defined by the anticlinal features and the fault, as well as enabled by the presence of a reservoir rock
 		- considering a petroleum production industry perspective, actors/ decision makers would be interested in the evaluation of value this potential reservoir
@@
 
+		\subsection{Construction of the 3D geological reservoir model}\label{sec:3D_construction}
 		- in the following, the construction and design of this model is described
@@
 		- place layer positional points in three lines, resembling three hypothetical seismic lines as base information
-		-
+
+		\subsection{Determination of the trap volume}
+
+			\subsubsection{Finding spill and leak points}
+
+			\subsubsection{Calculating the maximum trap volume}

### 2017-10-16 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Model construction text. Need to merge with stuff at home!

@@
 The original concept of this model is designed in a way, that a potential hydrocarbon trap is formed in the reservoir rock enclosed by the deformed and tilted seal and normal fault.\\
-		This trap is defined as the central feature of economic interest.
+		This trap is defined as the central feature of economic interest. [...]
 
-		- considering a petroleum production industry perspective, actors/ decision makers would be interested in the evaluation of value this potential reservoir
-		- "traditionally", this is represented by volumetric calculations or estimations of the recoverable reserves
-		- here: maximum volume of entrapment and subsequently, relative or hypothetical volumes of recoverable reserve can be calculated (for each model realization)
-		- (later: numbers from example case are taken, to achieve estimations of the NPV for a development project realization)
-		- looking at this structural model, not only the calculated volume can be taken into account, but also risk factors, such as fault permeabity seal safety (and capacity?), possible juxtapositions and leaks
-
-		- as in the 1D case, uncertainties are introduced for a number of parameters in the model, in a way that they affect the resulting maximum volume of the trap (and following also the recoverable reserves)
-		- again, a specific loss function is designed and applied on the results from numerous iterations of the constructed uncertain model, thus resembling the different value estimations/predictions or decisions different actors might make
+		The maximum trap volume is assumed to equate the hydrocarbon-filled rock volume $A * h$ in Equation \ref{eq:OOIP}.
+$OOIP$ and recoverable oil volumes are thus to be used to valuate each realization of the geological model.
+Single results are expected to vary, depending on the uncertain input parameters defined in Section \ref{sec:3D_construction} below.
+The custom loss function \ref{eq:LFR_final} is to be applied on the resulting value distributions before and after performing Bayesian inference.
+This way, the effect of incorporating additional information in the form of likelihoods is to be assessed from a perspective of Bayesian decision theory.
 				- (to bring this Bayesian method of using loss functions closer to the traditional application of deterministic decision-making in decision-trees, the continuous estimation space of the reservoir value (volume) is combined with the assumption, that the estimated value is linked to a consequent investment into a determined project development size.
 thus, the continuous approach is used to make the decision in a deterministic decision-making space.
-))
+)?
 
-		- later on,  the effect of adding information (in the form of likelihoods) is examined, particularly by looking at the changes in decision-making after applying this Bayesian updating and uncertainty has been reduced
+		\subsection{Construction of the uncertain 3D geological reservoir model}\label{sec:3D_construction}
+		The 3D structural geological model is constructed as follows: In principle, it is defined as a cubic block with an extent of 2000~m in X, Y and Z directions.
+The basic input data for the interpolation of the geological features is comprised of 3D point coordinates for layer interfaces and fault surfaces, as well as measurements that indicate respective dip directions and angles.
+In a Python environment, the data is imported as comma-separated values (CSV) using gemPy.
+From the generated data frame, gemPy is able to interpolate surfaces following the potential-field method explained in Section \ref{sec:struc_geo_modeling}) and compute voxel-based 3D geological models (see Figure \ref{fig:base3Dmodel_section}).
 
-		\subsection{Construction of the 3D geological reservoir model}\label{sec:3D_construction}
-		- in the following, the construction and design of this model is described
+		\begin{figure}[h]
+			\centering
+			\includegraphics[width=1\textwidth]{Figures/base3Dmodel_plot.png}
+			\caption{Mid section of the base 3D geological model as defined by the original data frame (see Appendix).
+Layer interfaces are defined by points and foliation dipping indicated by perpendicularly oriented arrows.}\label{fig:base3Dmodel_section}
+		\end{figure}
 
-		- base of the structural geological model is a 3Dcubic  space (block)  with 2000 m size in X, Y and Z x3
-		- place layer positional points in three lines, resembling three hypothetical seismic lines as base information
+		Uncertainties are implemented in several ways:
+		\begin{enumerate}
+		        \item \textbf{Z-values for layer interfaces:} Analogously to the 1D model, uncertainties for the vertical position of layer interfaces are implemented by using normal probability distributions.
+A deviation from the original input data Z-value is hereby added and allows for the consideration of uncertainties regarding layer surface positions in depth, layer thicknesses, topological shape and degree of fault offset.
+Points belonging to the same layer interface are assigned the same base uncertainty by applying one shared probability distribution.
+Assuming an increase of uncertainty with depth, standard deviations for the distributions are increased for lower formations.
+Furthermore, uncertainty regarding the magnitude of fault offset is incorporated by adding a further normal probability distribution that is shared by all layer interface points on the hanging wall.
+The specific ditributions are listed in the Appendix.
+		        \item \textbf{Cross-fault sealing:} Uncertainty about the sealing capacity across the fault is implemented using a Bernoulli distribution.
+In doing so, a probability of success is defined beforehand.
+In the case of such a "success", a value of one is returned by the distribution and lateral sealing of the fault is assumed.
+Should a value of zero be returned, cross-fault leakage is determined to be generally possible, depending on the occurrence of juxtapositions.
+		        \item \textbf{Other non-structural parameters:} Other factors that determine the $OOIP$ and $ROV$ calculations, such as porosity and recovery factor, can be implemented stochastically or deterministically.
+WHAT DO I USE?
+		\end{enumerate}
 

### 2017-10-16 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Update 20 iter

--- 2017-10-02 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
@@
 		For the purpose of exploring the effect of the methods of Bayesian analysis and decision theory in this work, the model is nevertheless to be kept relatively simple.
-Stratigraphically, it is designed to include one main reservoir unit (sandstone), one main seal unit (shale), an underlying basement through which hydrocarbon fluids could have flown upwards and an overlying formation that is assumed to be permeable, so that hydrocarbons can leak upwards.
+Stratigraphically, it is designed to include one main reservoir unit (sandstone), one main seal unit (shale), an underlying basement through which hydrocarbon fluids could have flown upwards and overlying formations that are assumed to be permeable, so that hydrocarbons can leak upwards.
 Structurally, it is constructed to feature an anticlinal dome fold that is displaced by a normal fault.
 All layers are tilted so that they dip in the opposite direction of the fault plane dip.
-The original concept of this model is designed in a way, that a potential hydrocarbon trap is formed in the reservoir rock enclosed by the deformed and tilted seal and normal fault.\\
-		This trap is defined as the central feature of economic interest.
+The original concept of this model is designed in a way, that a potential hydrocarbon trap is formed in the reservoir rock enclosed by the deformed seal and the normal fault.\\
+		This trap is defined as the central feature of economic interest, more specifically, the trap volume.
+For conducting simple and straightforward volumetric calculations, it is assumed that defined safe traps are always "full-to-spill", i.e.
+the complete trap volume is hydrocarbon filled and the $OOIP$ is attained over this volume.
 

### 2017-10-16 Risk analysis/Text/Layout/Layout/chapter2_methods.tex

--- 2017-10-16 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
@@
 		\subsection{Design of the 3D geological reservoir model}\label{sec:3D_design}
-		For the purpose of exploring the effect of the methods of Bayesian analysis and decision theory in this work, the model is nevertheless to be kept relatively simple.
-Stratigraphically, it is designed to include one main reservoir unit (sandstone), one main seal unit (shale), an underlying basement through which hydrocarbon fluids could have flown upwards and an overlying formation that is assumed to be permeable, so that hydrocarbons can leak upwards.
-Structurally, it is constructed to feature an anticlinal dome fold that is displaced by a normal fault.
-All layers are tilted so that they dip in the opposite direction of the fault plane dip.
-The original concept of this model is designed in a way, that a potential hydrocarbon trap is formed in the reservoir rock enclosed by the deformed and tilted seal and normal fault.\\
-		This trap is defined as the central feature of economic interest. [...]
-
 		The maximum trap volume is assumed to equate the hydrocarbon-filled rock volume $A * h$ in Equation \ref{eq:OOIP}.
@@
 From the generated data frame, gemPy is able to interpolate surfaces following the potential-field method explained in Section \ref{sec:struc_geo_modeling}) and compute voxel-based 3D geological models (see Figure \ref{fig:base3Dmodel_section}).
+		For the purpose of exploring the effect of the methods of Bayesian analysis and decision theory in this work, the model is nevertheless to be kept relatively simple.
+Stratigraphically, it is designed to include one main reservoir unit (sandstone), one main seal unit (shale), an underlying basement through which hydrocarbon fluids could have flown upwards and overlying formations that are assumed to be permeable, so that hydrocarbons can leak upwards.
+Structurally, it is constructed to feature an anticlinal dome fold that is displaced by a normal fault.
+All layers are tilted so that they dip in the opposite direction of the fault plane dip.
+The original concept of this model is designed in a way, that a potential hydrocarbon trap is formed in the reservoir rock enclosed by the deformed seal and the normal fault.\\
+		This trap is defined as the central feature of economic interest, more specifically, the trap volume.
+For conducting simple and straightforward volumetric calculations, it is assumed that defined safe traps are always "full-to-spill", i.e.
+the complete trap volume is hydrocarbon filled and the $OOIP$ is attained over this volume.
 

### 2017-10-17 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Bit of text changes

@@
 		\subsection{Design of the 3D geological reservoir model}\label{sec:3D_design}
+		For the purpose of exploring the effect of the methods of Bayesian analysis and decision theory in this work, the model is nevertheless to be kept relatively simple.
+Stratigraphically, it is designed to include one main reservoir unit (sandstone), one main seal unit (shale), an underlying basement through which hydrocarbon fluids could have flown upwards and overlying formations that are assumed to be permeable, so that hydrocarbons can leak upwards.
+Structurally, it is constructed to feature an anticlinal dome fold that is displaced by a normal fault.
+All layers are tilted so that they dip in the opposite direction of the fault plane dip.
+The original concept of this model is designed in a way, that a potential hydrocarbon trap is formed in the reservoir rock enclosed by the deformed seal and the normal fault.\\
+		This trap is defined as the central feature of economic interest, more specifically, the trap volume.
+For conducting simple and straightforward volumetric calculations, it is assumed that defined safe traps are always "full-to-spill", i.e.
+the complete trap volume is hydrocarbon filled and the $OOIP$ is attained over this volume.
 		The maximum trap volume is assumed to equate the hydrocarbon-filled rock volume $A * h$ in Equation \ref{eq:OOIP}.
@@
 From the generated data frame, gemPy is able to interpolate surfaces following the potential-field method explained in Section \ref{sec:struc_geo_modeling}) and compute voxel-based 3D geological models (see Figure \ref{fig:base3Dmodel_section}).
-		For the purpose of exploring the effect of the methods of Bayesian analysis and decision theory in this work, the model is nevertheless to be kept relatively simple.
-Stratigraphically, it is designed to include one main reservoir unit (sandstone), one main seal unit (shale), an underlying basement through which hydrocarbon fluids could have flown upwards and overlying formations that are assumed to be permeable, so that hydrocarbons can leak upwards.
-Structurally, it is constructed to feature an anticlinal dome fold that is displaced by a normal fault.
-All layers are tilted so that they dip in the opposite direction of the fault plane dip.
-The original concept of this model is designed in a way, that a potential hydrocarbon trap is formed in the reservoir rock enclosed by the deformed seal and the normal fault.\\
-		This trap is defined as the central feature of economic interest, more specifically, the trap volume.
-For conducting simple and straightforward volumetric calculations, it is assumed that defined safe traps are always "full-to-spill", i.e.
-the complete trap volume is hydrocarbon filled and the $OOIP$ is attained over this volume.
 

### 2017-10-17 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Continued text. Now on spill point finding and saddle points

@@
 		\subsection{Determination of the trap volume}
+		In the course of this work, several algorithms are developed within the python environment, that in combination enable the automatic recognition and calculation of trap volumes in geological models computed by gemPy.
+To assign voxels of the model to the trap volume, it is checked whether the following conditions are satisfied by each particular voxel:
+		\begin{enumerate}
+				        \item \textbf{Labeled as reservoir formation:} The voxel has been assigned to the target reservoir formation in the lithology block model.
+This is determined by respective labeling of the input data and the computation conducted by gemPy.
+				        \item \textbf{Location in footwall:} The voxel is located on the footwall side of the fault.
+This condition is applicable to this specific model design, in which entrapment is assumed to occur between the footwall anticlinal enclosure and the normal fault.
+Due to respective dip of the reservoir formation, full leakage is assumed for the reservoir formation in the hanging wall.
+A distinction between foot- and hanging wall is easily achievable by using the fault model block that is generated by gemPy.
+In this fault block, voxels on both sides are assigned respective values for distinction.
+				        \item \textbf{Location above spill point horizon:} The voxel is located vertically above the final spill point of the trap.
+In the algorithm to find this final spill point, it is distinguished between a spill point defined by the folding structure, referred to as anticlinal spill point, and a cross-fault leak point, that depends on the magnitude of displacement and the resulting nature of juxtapositions.
+Once both of these points have been determined, the higher one is defined to be the final spill point used to determine the maximum fill capacity of the trap.
+In the case of fault sealing, however, the cross-fault leak point is discarded as irrelevant.
+The process of finding the decisive spill point is explained below in further detail.
+				        \item \textbf{Location inside of closed system:} The voxel is part of a model section inside of the main anticlinal feature.
+All of the voxels in this section are separated from the borders of the model by voxels that don't meet the conditions above, which primarily means that they are encapsulated by seal voxel laterally and upwards.
+This condition is applicable  under the assumption, that connection to the borders of the model lead to leakage.
+A trap is thus defined as a closed system in this model.
+		\end{enumerate} 
+		REGARD THE RELATIVITY OF THE CONDITIONS AND THE APPROACH TO THIS SPECIFIC MODEL!
 
-			\subsubsection{Finding spill and leak points}
+			\subsubsection{Finding the final spill point of a trap} %\label{sec:spill_leak_finding}
+			Regarding dome-shaped anticlinal structures, it can be observed that, geometrically and mathematically, a spill point is a saddle point of the reservoir top surface in 3D.
+Two examples for such saddle points are visualized in Figures \ref{fig:saddle_point} and \ref{fig:saddle_point_between_maxima}.
+The latter one in particular, illustrates well
+
+			\begin{figure}[h]
+				%\centering
+				\begin{subfigure}{.5\textwidth}
+					\centering
+					\includegraphics[width=1\linewidth]{Figures/Saddle_point.png}
+					\caption{Graph and saddle point of the function $z=x2−y2$.}\label{fig:saddle_point} 
+				\end{subfigure}
+				\begin{subfigure}{.5\textwidth}
+					\centering
+					\includegraphics[width=1\linewidth]{Figures/Saddle_Point_between_maxima.png}
+					\caption{Saddle point between two maxima.}\label{fig:saddle_point_between_maxima} 
+				\end{subfigure}
+				\caption{Two examples for saddle points on surfaces of 3D functions (from Wikimedia Commons)}
+			\end{figure}
 
 			\subsubsection{Calculating the maximum trap volume}
+
+		\subsection{Determination of layer thicknesses}

### 2017-10-18 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Spill point text continued

@@
 			\subsubsection{Finding the final spill point of a trap} %\label{sec:spill_leak_finding}
-			Regarding dome-shaped anticlinal structures, it can be observed that, geometrically and mathematically, a spill point is a saddle point of the reservoir top surface in 3D.
-Two examples for such saddle points are visualized in Figures \ref{fig:saddle_point} and \ref{fig:saddle_point_between_maxima}.
-The latter one in particular, illustrates well
+			Regarding dome-shaped anticlinal structures, it can be observed that, geometrically and mathematically, a spill point is a saddle point of the reservoir top surface in 3D. SADDLE POINT DEFINITION
+			Two examples for such saddle points are visualized in Figure \ref{fig:saddle_points_general}.
+Particularly Figure \ref{fig:saddle_point_between_maxima} can be seen as a representation of a point of spill between two dome structures (i.e.
+3D surface maxima) is defined by the surface's saddle point between both.
 
@@
 				\end{subfigure}
-				\caption{Two examples for saddle points on surfaces of 3D functions (from Wikimedia Commons)}
+				\caption{Two examples for saddle points on surfaces of 3D functions (from Wikimedia Commons)}\label{fig:saddle_points_general}
 			\end{figure}

### 2017-10-18 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Euclidian distance to be used? Text on Trap Volume calc written

@@
 		\end{enumerate} 
-		REGARD THE RELATIVITY OF THE CONDITIONS AND THE APPROACH TO THIS SPECIFIC MODEL!
+		It has to be emphasized that these conditions have been fitted to the geological model constructed as described in Section \ref{sec:3D_construction}.
+For other models featuring different geological properties, features and levels of complexities, these conditions might not apply at all or might have to be adapted.
+Models of higher complexities will surely require the introduction of further conditions.
 
 			\subsubsection{Finding the final spill point of a trap} %\label{sec:spill_leak_finding}
-			Regarding dome-shaped anticlinal structures, it can be observed that, geometrically and mathematically, a spill point is a saddle point of the reservoir top surface in 3D. SADDLE POINT DEFINITION
-			Two examples for such saddle points are visualized in Figure \ref{fig:saddle_points_general}.
-Particularly Figure \ref{fig:saddle_point_between_maxima} can be seen as a representation of a point of spill between two dome structures (i.e.
-3D surface maxima) is defined by the surface's saddle point between both.
-
+			Regarding dome-shaped anticlinal structures, it can be observed that, geometrically and mathematically, a spill point is a saddle point of the reservoir top surface in 3D.
+Regarding a surface defined by $f(x,y)$, a local maximum at $(x_0,y_0,z_0)$ would resemble a hill top \citep{guichard2013calculus}.
+Local maxima will be found looking at the cross-sections in the planes $y = y_0$ and $x = x_0$.
+Furthermore, the respective partial derivatives (i.e.
+gradients) $\frac{\delta z}{\delta x}$ and $\frac{\delta z}{\delta y}$ will equal zero at $x_0$ and $y_0$, i.e.
+that the extremum is a stationary point \citep{guichard2013calculus, weisstein2017saddlepoint}.
+In the context of a geological reservoir system, such a hill can be considered to represent an anticlinal structural trap.
+Local minima are defined analogously, presenting local minima in both planes at a stationary point \citep{guichard2013calculus}.
+A saddle point, however, is a stationary point, while not being an extremum \citep{weisstein2017saddlepoint}.
+In general, saddle points can be distinguished from extrema by applying the second derivative test \citep{guichard2013calculus, weisstein2017saddlepoint}: Considering a 2D function $f(x,y)$ with continuous partial derivatives at a  point $(x_0,y_0)$, so that $f_x(x_0,y_0) = 0$ and $f_x(x_0,y_0) = 0$, the following discriminant $D$ can be introduced:
+			\begin{equation}\label{eq:discriminant_D}
+			D(x_0,y_0) = f_{xx}(x_0,y_0)f_{yy}(x_0,y_0) - f_{xy}(x_0,y_0)^2
+			\end{equation}
+			Using this, the following holds for a point $(x_0,y_0)$:
+			\begin{enumerate}
+				\item If $D > 0$ and $f_{xx}(x_0,y_0) < 0$, there is a local maximum.
+				\item If $D > 0$ and $f_{xx}(x_0,y_0) > 0$, there is a local minimum.
+				\item If $D < 0$, there is a saddle point at the point $(x_0,y_0)$.
+				\item If $D = 0$, the test fails \citep{guichard2013calculus}.
+			\end{enumerate}
 			\begin{figure}[h]
@@
 				\includegraphics[width=1\linewidth]{Figures/Saddle_point.png}
-					\caption{Graph and saddle point of the function $z=x2−y2$.}\label{fig:saddle_point} 
+				\caption{Graph and saddle point of the function $z=x2~-~y2$.}\label{fig:saddle_point} 
 			\end{subfigure}
@@
 			\end{figure}		
+			Two examples for such saddle points are visualized in Figure \ref{fig:saddle_points_general}.
+Particularly Figure \ref{fig:saddle_point_between_maxima} can be seen as a representation of a point of spill between two dome structures (i.e.
+surface maxima or "hills") which is defined by the surface's saddle point between both.\\
+			In gemPy, layer boundary surfaces are returned in the form of discretized arrays of simplices and vertices.
+The latter can be transformed into a 2D rectangular array that contains Z-positional values for the surface in X and Y directions.
+According to \citet{verschelde2017programmingtools}, a saddle point in a matrix is maximal in its row and minimal in its column.
+This corresponds to the logical geometrical deduction, that a saddle point for a surface defined by $f(x,y)$ is marked by a local maximum in one plane, but a local minimum in the perpendicular plane.
+This is observable in Figure \ref{fig:saddle_points_general}.\\
+			Regarding leakage across the normal fault in this geological model, a similar approach is taken.
+If the fault offset remains small enough, that the seal layer is not broken, then the reservoir top surface can be used to find the cross-fault leak point.
+In this case, the leak point can be defined as the maximal point of hanging wall contact of the reservoir top surface with the fault.
+At the same time, this point will be a local minimum on the plane perpendicular to the normal fault and thus equal a saddle point.
+However, higher magnitudes of fault displacement can lead to breaking of the seal, i.e.
+the juxtaposition of the trap structure with overlying formations which are assumed to be complete permeable.
+In this case, the maximum footwall contact of the surface with the fault is to be defined as cross-fault leak point.\\
+			Following these observations, the algorithm developed in this work uses differentiation techniques to find local extrema in the 2D reservoir top surface arrays attained over gemPy.
+Points where minima and maxima of perpendicular axes X and Y coincide are then defined as saddle points.
+The anticlinal spill point is deduced as the maximal saddle point in the footwall.
+The cross-fault leak point is determined analogously in the hanging wall.
+In the end, whichever point of these two is maximal, is defined to be the final spill point of the trap.
+If no cross-fault leak point is found, it is presumed that full leakage is allowed to take place due to little displacement of the reservoir layer.
+However, if fault sealing is given, any fault leakage and respective leak points become irrelevant and only the anticlinal spill point is considered.
+The Z-value of the final spill point is used to determine the maximum down-to-horizon to which the trap feature can be filled.\\
+			\begin{figure}[h]
+				\centering
+				\includegraphics[width=1\textwidth]{Figures/restop_surface_spill_leak.png}
+				\caption{Plot of reservoir top surface points and determined anticlinal spill point (marked by 'X') and cross-fault leak point (marked by '+').}\label{fig:restop_surface_spill_leak}
+			\end{figure}\\	
+			\citet{kuijper2004detecting} has pointed out that using finding spatial critical points, saddle points especially, can be problematic using discretized methods and rectangular grids in particular.
+Finding all extrema and saddle points can be difficult \citep{kuijper2004detecting}.
+Some failures to detect saddle points in this work's model are resolved by adding a buffer around extrema that define the anticlinal structure.
+IMPLEMENT BETTER METHOD?
+However, it has to be pointed out that saddle points might still be missed in some rare cases, especially if they are located in an orientation which is "rotated" and not aligned with the X and Y axes.
+Furthermore, in this work, the geological model was designed in a way that saddle points are presumed to only occur within or near the structures of interest.
+FUTURE: MORE UNIVERSALLY APPLICAPLE METHODS
 
 			\subsubsection{Calculating the maximum trap volume}
+			Have all voxels that fulfill the conditions defined above been determined, these voxels are seen as part of the maximum trap volume $V_{trap}$.
+This volume can then be calculated by simply counting the number of these voxels and rescaling their cumulative volume depending on the resolution in which the model was computed:
+			\begin{equation}
+			V_{trap} = n_v*(\frac{S_{orig}}{R_{m}})^3
+			\end{equation}
+			Where $n_v$ is the number of trap voxels, $S_{orig}$ gives the original scale and $R_m$ the used resolution for the model.
+As mentioned before, $V_{trap}$ is assumed to equal the hydrocarbon-filled rock volume $A * h$ used in Equation \ref{eq:OOIP}.
+			For the example of a geological model with an original extent of 2000~m in three directions, computed using a resolution of 50, the scale factor is 40~m.
+Every voxels thus accounts for $40^3~m=64,000~m^3$ in volume.
+It has to be noted, that this direct approach to rescaling and calculating the volume requires the model to be designed as a cubic block.\\
 
-		\subsection{Determination of layer thicknesses}
+		\subsection{Implementation of reservoir-related likelihoods in the 3D model}

### 2017-10-18 Risk analysis/Text/Layout/Layout/chapter3_results.tex
inserted things to not forget in results section

@@
 Information on the seal, however, based on a normal distribution around 25~m thickness, leaves uncertainty about the reliability of the seal, as the safety threshold is defined as 20~m.\\
+				NOT INCREASED CERTAINTY; ONLY OVERALL VALUES!
 				Increased certainty about the reservoir thickness is sufficient to shift Bayes actions to higher estimates for all actors, but the most risk-averse one ($r$ = 1.5).
@@
 According to these values, the risk-neutral to risk-averse individuals seem to profit the most, due to a large change in estimate value and decreased expected loss.
-Compared to case I, the estimate shifts for both risk-friendlier actors remain approximately the same, expected losses, however, are lowered much more significantly.\\
+Compared to case I, the estimate shifts for both risk-friendlier actors remain approximately the same, expected losses, however, are lowered much more significantly.
+				POINT OUT: DECREASED UNCERTAINTY! AND HIGHER VALUES OVERALL!\\		
 				%\begin{table}
@@
 Quantified changes in the positions and values of the Bayes actions are listed in Table \ref{tab:update_examples_all}.
-There is no shift in estimate for risk-averse actors, as they already found their best estimates to be zero before updating.\\
+There is no shift in estimate for risk-averse actors, as they already found their best estimates to be zero before updating.
+				DECREASED UNCERTAINTY BUT AT THE SAME TIME LOWER VALUES OVERALL!\\
 				%\begin{table}

### 2017-10-19 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Tried to implement juxtaposition test, not working, yet!

@@
 		\subsection{Implementation of reservoir-related likelihoods in the 3D model}
+		As for the 1D case before, for the 3D case it is assumed that additional information is provided in the form of layers thickness likelihoods.
+In this geological scenario, the information might have been gained for example from additional seismic obersavtions or from core sample data from a well nearby.
+Assuming a possible correlation with the layers in our model section, layer thickness likelihoods can be deduced respectively.\\
+		First of all, a method to determine layer thicknesses in our model realizations has to be introduced. Here,

### 2017-10-23 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Some text update. Need to check merge with code changes at home

--- 2017-10-18 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
@@
 
-        \section{Application in structural geological modeling}
+        \section{Application in structural geological modeling}\label{sec:bayes_in_struc_modeling}
         In this work, these methods of Bayesian analysis and decision theory are applied in the field of geological structural modeling.
@@
 			Where reservoir thickness is given by $h_{res}$, seal thickness by $h_{seal}$ and reservoir depth, which is always a negative value, by $d_{res}$.
+			This 1D model is then placed inside the Bayesian probabilistic modeling framework
 
@@
 		\subsection{Implementation of reservoir-related likelihoods in the 3D model}
+		As for the 1D case before, for the 3D case it is assumed that additional information is provided in the form of layers thickness likelihoods.
+In this geological scenario, the information might have been gained for example from additional seismic observations or from core sample data from a well nearby.
+Likelihoods could also be derived from knowledge of the geological history of the formation, such as the depositional environment of sedimentary layers.
+Assuming a possible correlation of nearby core samples with the layers in our model section, layer thickness likelihoods can be deduced respectively.\\
+		First of all, a method to determine layer thicknesses in our model realizations has to be introduced.
+Here, as single X- and Y-point (1D column in Z) in the footwall of the model is chosen as a measure for the thickness of the layers.
+In this column, point coordinates for the layer interfaces had been defined in the original data frame and can be observed as they are altered in the interpolation data frame for each iteration.
+The column is located centrally in the overall model block, as well as the original potential trap structure (X~=~1100~m,Y~=~1000~m) and is thus presumed to be representative and of significance for the reservoir analysis.
+Thickness variations are traced and vary relative to the changes in the layer interface positions, given by their uncertainties in Z-values.\\
+		After defining these thickness to point-positional dependencies, thickness likelihoods are implemented as probability distributions, as done for the 1D model.
+Various types of geological information, conclusions and expectations can be expressed by adapting mean and standard deviation of these normal distributions.
+For core sampling, for example, information might be evenly provided for all layers, irrespective of depth, so that similar standard distributions might be assigned to all layer thicknesses.
+Alternatively, taking into account certain depositional environments for which high thickness variablity is characteristic, specific standard deviations can be raised accordingly.
+Some representative cases are presented in Chapter \ref{cha:results}.
+
+		\subsection{Bayesian analysis and valuation of the 3D model}
+		According to the methods introduced in Sections \ref{sec:bayes} and \ref{sec:bayes_in_struc_modeling}, the 3D geological model is to be incorporated in a probabilistic modeling framework and analyzed from a Bayesian perspective.

### 2017-10-24 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Bit of text progress in methods part

@@
 These beliefs can be assigned to individuals \citep{davidson2015}.
-Thus, different and even contradicting beliefs about the probability of an event might be held by different individuals, based on variations and disparities in the information available to each one individual \citep{davidson2015}.
-
+Thus, different and even contradicting beliefs about the probability of an event might be held by different individuals, based on variations and disparities in the information available to each one individual \citep{davidson2015}.\\
         The initial belief or guess about an event A can be denoted as P(A) \citep{davidson2015}.
@@
         \begin{enumerate}
-        	\item \textbf{Mathematical forward model ($M$)}: The connections between parameters $\theta$ and observed data $y$ are defined in this mathematical model. ...
-        	\item \textbf{Model parameters ($\theta$)}: These model-defining parameters, such as the position or dip of layers and faults in geological settings, can be deterministic or stochastic.
+        	\item \textbf{Mathematical forward model ($M$)}: The connections between parameters $\theta$ and observed data $y$ are defined in such mathematical models.
+According to \citet{delaVarga2016} (and GREENSTONE PAPER, CITE?), the realization of a geological model $M$ can be regarded as a direct function of a set of input parameters:
+        	\begin{equation}
+        	M = f(\vec{x}, \phi_i, k_j, \alpha_k, \beta_l)\label{eq:forward_model},
+        	\end{equation}
+        	where $\phi_i$ is a function of position $\vec{x}$, more precisely an interpolation function, to which respective additional parameters are given by $\alpha_k$.
+Available primary geological information, such as positions of layer interfaces and dips in the subsurface, is represented by $k_j$.
+A topological description, such as the relationships between faults and layer surfaces, is given by $\beta_l$.
+An essential aspect of this function is that it allows for a full automation of the modeling step, so that the consequences of a change in an input parameter are realized directly without the need for any further manual inspection or interaction (cite GREENSTONE paper).
+The modeling step and respective interpolation used in this work is presented in Section \ref{sec:struc_geo_modeling} below.
+        	\item \textbf{Model parameters ($\theta$)}: These model-defining parameters can be deterministic or stochastic.
 In the latter case, they are uncertain parameters to which a probability distribution is assigned.
+Regarding Equation \ref{eq:forward_model} above, these can be any of the factors $\vec{x}$, $k_j$, $\alpha_k$ and $\beta_l$, which define the realization of the forward model over $\phi_i$.
         	\item \textbf{Observed data ($y$)}: This is any type of additional information that can be related to the forward modeling results and might possibly be used to reduce uncertainty.
 Such data can be gained by measurements and observations, for example by core sampling, well-log analysis or seismic acquisition.
-        	\item \textbf{Likelihood functions ($p(y|\theta)$}: Links between the previous parameters $\theta$ and the additional data $y$ are established by these functions in way that they reflect the likelihood of the parameter states given the observations \citep{delaVarga2016}.
+        	\item \textbf{Likelihood functions ($p(y|\theta)$}: Links between the previous parameters $\theta$ and the additional data $y$ are established by these functions in way that they reflect the likelihood of the parameter states given the observations.
+They mathematically defined in the same way as probability functions, but are a function of the data $y$, instead of the parameters $\theta$ \citep{delaVarga2016}.
         \end{enumerate}
@@
         \subsection{Structural geological forward modeling}\label{sec:struc_geo_modeling}       
-        Performing forward modeling in the context of structural geology requires the use of a suitable modeling step.
-Additionally, for the application in a probabilistic setting, the method should enable fully automatic reconstruction of the model, when parameters are changed.
+        Performing forward modeling in the context of structural geology requires the use of a suitable modeling step $M$ (see Equation \ref{eq:forward_model}).
+For the application in a probabilistic setting, the method should enable fully automatic reconstruction of the model, when parameters are changed.
 The application in this work follows the example of \citet{delaVarga2016} and relies on the use of implicit interpolation for geological modeling, a method developed and elaborated by \citet{Lajaunie1997} and \citet{calcagno2008geological}.\\
-        This implicit method relies on the interpolation of a potential field scalar function $T(\vec{x})$ of any point $\vec{(x)}$ in a 3D space, and thus reflects the geometry of geological structures \citep{calcagno2008geological}.
+        This implicit method relies on the interpolation of a potential field scalar function $T(\vec{x})$ of any point $\vec{x}$ in a 3D space, and thus reflects the geometry of geological structures \citep{calcagno2008geological}.
 Modeling $T(\vec{x}$ is achieved by cokriging that regards two forms of data: (i) contact points on geological interfaces through increments of the potential field $T(\vec{x})-T(\vec{x}^{\,'})$ and (ii) orientation data as gradients, i.e.
@@
         \end{equation}        
-
         where $\vec{x_0}$ is an arbitrary origin, $M$ and $N$ are the total number of data points and partial derivatives respectively, and their relative contributions are weighed by the factors $\mu_\alpha$ and $\nu_\beta$.

### 2017-10-25 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Some text progress (methods)

@@
         Drawing representative samples from an target distribution of unknown shape is based on the conduction of a so-called random walk on the parameter distribution space.
-T sampling steps are to be performed.
+$T$ sampling steps are to be performed.
 The first sampling location is chosen at random.
-With each subsequent step, a new sampling location is proposed.
+With each subsequent step, a new position is proposed.
 The new sample value is then related to the previous step.
@@
 In the case of rejection, sampling is reverted to the previous accepted step \citep{schaaf2017, delaVarga2016}.
-The intention behind this concept it to achieve convergence of the sampling algorithm towards areas of high probability \citep{davidson2015}.\\ Variations in the way of how new  sample steps are proposed and in the acceptance-rejection condition result in different single MCMC sampling methods \citep{schaaf2017, delaVarga2016}.
+After perform the complete number of $T$ iterations, all accepted sampling locations (i.e.
+the trace) are returned.
+The intention behind this concept it to achieve convergence of the sampling algorithm towards areas of high probability \citep{davidson2015}.\\
+        Variations in the way of how new  sample steps are proposed and in the acceptance-rejection condition result in different single MCMC sampling methods \citep{schaaf2017, delaVarga2016}.
 Various algorithms for random MCMC walks have been developed for over more than six decades and advancements have still been made in recent years.
@@
 The Gibbs sampler \citep{geman1984stochastic} is another well-known method.
-        For the purspose of this work, an adaptive Metropolis-Hastings sampler is used.\\
-        In Metropolis-Hastings methods, each sampling step at iteration t is determined by a candidate probability distribution $q(\theta,\theta^{'})$, from which a proposed sample $\theta^{'}$ is drawn \citep{delaVarga2016}.
+        For the purpose of this work, an adaptive Metropolis-Hastings sampler is used.\\
+        In Metropolis-Hastings methods, each sampling step at iteration $t$ is determined by a candidate probability distribution $q(\theta,\theta^{'})$, from which a proposed sample $\theta^{'}$ is drawn \citep{delaVarga2016}.
 The acceptance-rejection condition is defined by the acceptance ratio $a(\theta^{'},\theta)$:
@@
 Small steps have better acceptance rates, but lead to slower exploration and convergence of the algorithm \citep{delaVarga2016}.\\
-        The Adaptive Metropolis (AM) by \citet{haario2001adaptive} is used in this work.
+        For the iterative sampling, Adaptive Metropolis (AM) by \citet{haario2001adaptive} is used in this work.
 It adapts the traditional Metropolis-Hastings by incorporating the ability of continuous step size tuning during convergence, by taking into account the full information saved along the process.
@@
 The adaptive nature of the process enables fast convergence for non-linear distributions while maintaining ergodicity \citet{haario2001adaptive, delaVarga2016}.
-Its suitability for multi-dimensional distribution spaces make it an excellent method for dealing with complex models such as the structural geological models in this work \citep{schaaf2017}
+Its suitability for multi-dimensional distribution spaces make it an excellent method for dealing with complex models such as the structural geological models in this work \citep{schaaf2017}.\\
+        A large enough number of iterations $T$ has to be chosen, so that a reliable and statistically significant exploration of the parameter space is assured.
+This is primarily dependent on the rate of convergence towards the true distribution.
+Considering empirical Bayes methods, in which either priors or likelihoods stem from empirical data, as is assumed for the models in this work, convergence can be expected to be reached almost immediately.
+However, for high dimensional problems more iterations are required, in order to ensure an accurate representation of posterior distributions (cite Miguel GREENSTONE!).
+Different realizations of the full 3D geological model are then constructed on the basis of the resulting posterior distributions, using an implicit modeling step described in the following.
 
@@
 $OOIP$ and $ROV$ are applicable in a 3D setting, the 1D geological model introduced below in Section \ref{sec:1D_model}, however, requires the use of an abstract valuation system that is explained in Section \ref{sec:1D_score_system}.
-
-        \subsubsection{Net present value}
-        hmmm...
+        %\subsubsection{Net present value}
+        %hmmm...
 
@@
 Python is thus a straightforward tool for the implementation of central components of Bayesian analysis, such as custom statistical distributions and samplers \citep{salvatier2016pymc3}.\\
-		The 3D geological modeling step in this work is implemented using GeMpy, an open-source, Python-based software that is able to generate and visualize complex 3D structural geological models based on the potential field interpolation method elaborated in Section \ref{sec:struc_geo_modeling} \citet{delaVarga2017gempy}.
-Its design allows for its application in a probabilistic setting, particularly by coupling it with PyMC (see below).
+		The 3D geological modeling step in this work is implemented using GeMpy, an open-source, Python-based software that is able to generate and visualize complex 3D structural geological models based on the potential field interpolation method elaborated in Section \ref{sec:struc_geo_modeling} \citep{delaVarga2017gempy}.
+Its design allows for its application in a probabilistic setting, particularly by coupling it with PyMC as elaborated below.
 At the time of writing this, GemPy is still under development (version 0.995), but is already functioning for the purpose of this work.\\
-		For conducting the Bayesian analysis, GemPy is combined with the Python library PyMC, which was developed for conducting Bayesian inference and prediction problems in an open-source probabilistic programming framework \citep{davidson2015, salvatier2016pymc3}.
+		For conducting the Bayesian analysis and embedding the geological modeling step in a probabilistic modeling framework, GemPy is combined with PyMC.
+This Python library was developed for conducting Bayesian inference and prediction problems in an open-source probabilistic programming environment \citep{davidson2015, salvatier2016pymc3}.
 Different model fitting techniques are provided in PyMC, such as the \textit{maximum a posteriori} (MAP) method and several (MCMC) sampling methods, including the Adaptive Metropolis explained in Section \ref{sec:mcmc} \citep{salvatier2016pymc3}.
 The components which are used to construct a statistical model, are represented by \textit{Deterministic} or \textit{Stochastic} variables in PyMC.
-the values of \textit{Deterministic} variables are completely dependent on its parents' values, as defined by a respective mathematical function \citep{salvatier2016pymc3}.
+The values of \textit{Deterministic} variables are completely dependent on its parents' values, as defined by a respective mathematical function \citep{salvatier2016pymc3}.
 \textit{Stochastic} variables are used to represent uncertain parameters $\theta$ or observed stochastic variables as likelihood functions $p(y|\theta)$ \citep{salvatier2016pymc3, delaVarga2016}.
@@
 Futhermore, PyMC allows for the creation of own object definitions inheriting from the class descriptions of these two variable types.\\
-		As of writing this, the newer PyMC3 is not yet supported by GemPy.
-Thus, it is resorted to PyMC 2 for 3D geological settings in this work.
-However, the newer PyMC 3 is used in a simple 1D scenario.
-\citet{salvatier2016pymc3} point out that the development of PyMC 3 is continuing, as the inclusion of further tools is planned for future updates.
+		%While conducting MCMC sampling, a burn-function can be implemented via PyMC, through which it can be defined that the $t_b$ first trace samples are discarded. This can be useful to d\\		
+		\citet{salvatier2016pymc3} point out that the development of PyMC is continuing, as the inclusion of further tools is planned for future updates.
 

### 2017-10-25 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Switched to fast_run, still broken. :( Progressed in methods text

@@
 
+		\subsection{Numerical computational implementation via Python, GemPy and PyMC}\label{sec:numerical_implementation}
+		Bayesian analysis can be conducted using probabilistic programming \citep{salvatier2016pymc3}.
+The implicit method of forward geological modeling, described above, is to be embedded in such a framework.
+For doing so, the programming language of choice in this work is Python.
+The merits of Python have been pointed out by \citet{behnel2010, salvatier2016pymc3, Langtangen2008}.
+Development is facilitated by an expressive but concise and clean syntax that is easy to learn.
+Python is dynamic, compatible with multiple platforms and offers good support for numerical computing.
+Integration of other scientific libraries and extension via C, C++, Fortran or Cython are easily possible \citep{behnel2010, salvatier2016pymc3, Langtangen2008}.
+Python is thus a straightforward tool for the implementation of central components of Bayesian analysis, such as custom statistical distributions and samplers \citep{salvatier2016pymc3}.\\
+		The 3D geological modeling step in this work is implemented using GeMpy, an open-source, Python-based software that is able to generate and visualize complex 3D structural geological models based on the potential field interpolation method elaborated in Section \ref{sec:struc_geo_modeling} \citep{delaVarga2017gempy}.
+Its design allows for its application in a probabilistic setting, particularly by coupling it with PyMC as elaborated below.
+At the time of writing this, GemPy is still under development (version 0.995), but is already functioning for the purpose of this work.\\
+		For conducting the Bayesian analysis and embedding the geological modeling step in a probabilistic modeling framework, GemPy is combined with PyMC.
+This Python library was developed for conducting Bayesian inference and prediction problems in an open-source probabilistic programming environment \citep{davidson2015, salvatier2016pymc3}.
+Different model fitting techniques are provided in PyMC, such as the \textit{maximum a posteriori} (MAP) method and several (MCMC) sampling methods, including the Adaptive Metropolis explained in Section \ref{sec:mcmc} \citep{salvatier2016pymc3}.
+The components which are used to construct a statistical model, are represented by \textit{Deterministic} or \textit{Stochastic} variables in PyMC.
+The values of \textit{Deterministic} variables are completely dependent on its parents' values, as defined by a respective mathematical function \citep{salvatier2016pymc3}.
+\textit{Stochastic} variables are used to represent uncertain parameters $\theta$ or observed stochastic variables as likelihood functions $p(y|\theta)$ \citep{salvatier2016pymc3, delaVarga2016}.
+Complex mathematical relations between \textit{Stochastic} variables can be described through \textit{Deterministic} variables \citep{delaVarga2016}.
+Futhermore, PyMC allows for the creation of own object definitions inheriting from the class descriptions of these two variable types.\\
+		%While conducting MCMC sampling, a burn-function can be implemented via PyMC, through which it can be defined that the $t_b$ first trace samples are discarded. This can be useful to d\\		
+		\citet{salvatier2016pymc3} point out that the development of PyMC is continuing, as the inclusion of further tools is planned for future updates.
+
         \subsection{Economic significance and reservoir valuation}\label{sec:Reservoir_values}
@@
         %hmmm...
-
-		\subsection{Numerical computational implementation via Python, GemPy and PyMC}
-		Bayesian analysis can be conducted using probabilistic programming \citep{salvatier2016pymc3}.
-The implicit method of forward geological modeling, described above, is to be embedded in such a framework.
-For doing so, the programming language of choice in this work is Python.
-The merits of Python have been pointed out by \citet{behnel2010, salvatier2016pymc3, Langtangen2008}.
-Development is facilitated by an expressive but concise and clean syntax that is easy to learn.
-Python is dynamic, compatible with multiple platforms and offers good support for numerical computing.
-Integration of other scientific libraries and extension via C, C++, Fortran or Cython are easily possible \citep{behnel2010, salvatier2016pymc3, Langtangen2008}.
-Python is thus a straightforward tool for the implementation of central components of Bayesian analysis, such as custom statistical distributions and samplers \citep{salvatier2016pymc3}.\\
-		The 3D geological modeling step in this work is implemented using GeMpy, an open-source, Python-based software that is able to generate and visualize complex 3D structural geological models based on the potential field interpolation method elaborated in Section \ref{sec:struc_geo_modeling} \citep{delaVarga2017gempy}.
-Its design allows for its application in a probabilistic setting, particularly by coupling it with PyMC as elaborated below.
-At the time of writing this, GemPy is still under development (version 0.995), but is already functioning for the purpose of this work.\\
-		For conducting the Bayesian analysis and embedding the geological modeling step in a probabilistic modeling framework, GemPy is combined with PyMC.
-This Python library was developed for conducting Bayesian inference and prediction problems in an open-source probabilistic programming environment \citep{davidson2015, salvatier2016pymc3}.
-Different model fitting techniques are provided in PyMC, such as the \textit{maximum a posteriori} (MAP) method and several (MCMC) sampling methods, including the Adaptive Metropolis explained in Section \ref{sec:mcmc} \citep{salvatier2016pymc3}.
-The components which are used to construct a statistical model, are represented by \textit{Deterministic} or \textit{Stochastic} variables in PyMC.
-The values of \textit{Deterministic} variables are completely dependent on its parents' values, as defined by a respective mathematical function \citep{salvatier2016pymc3}.
-\textit{Stochastic} variables are used to represent uncertain parameters $\theta$ or observed stochastic variables as likelihood functions $p(y|\theta)$ \citep{salvatier2016pymc3, delaVarga2016}.
-Complex mathematical relations between \textit{Stochastic} variables can be described through \textit{Deterministic} variables \citep{delaVarga2016}.
-Futhermore, PyMC allows for the creation of own object definitions inheriting from the class descriptions of these two variable types.\\
-		%While conducting MCMC sampling, a burn-function can be implemented via PyMC, through which it can be defined that the $t_b$ first trace samples are discarded. This can be useful to d\\		
-		\citet{salvatier2016pymc3} point out that the development of PyMC is continuing, as the inclusion of further tools is planned for future updates.
 
@@
 An intuitive example is the comparison of a small and a large company.
-A certain false estimate or error might have a significantly stronger impact on a company which has a a generally lower market share and only few projects, than on a larger company which might possess a higher financial flexibility and for which one project is only one of many development options in a portfolio.
-
+A certain false estimate or error might have a significantly stronger impact on a company which has a a generally lower market share and only few projects, than on a larger company which might possess a higher financial flexibility and for which one project is only one of many development options in a portfolio.\\
 		In the following, the loss function is further adapted to consider different risk-affinities of several actors.
@@
 In the loss function, risk is simply included using a risk factor \textit{r} which alters the weighting factors \textit{a, b} and \textit{c} respectively:
-
 		\begin{equation}\label{eq:LFR_final}
@@
 		\end{equation}	
-
-		According to this, for \textit{r} = 1 the risk-neutral loss function is returned, since \textit{a, b} and \textit{c} are not altered.
-For \textit{r} $<$ 1, the weight on overestimating is reduced and increased for fatal underestimation.
-This represents a more risk-friendly actor that is willing to bid on a higher estimate to attain a greater gain.
-For \textit{r} $>$ 1, the overestimation weight is increased in the loss function, the underestimation weight is decreased and respectively more risk-averse actors are prompted to bid on lower estimates.\\
+		According to this, for \textit{r} = 1 the risk-neutral loss function is returned, since \textit{a, b} and \textit{c} are not changed.
+For \textit{r} $<$ 1, the weight on overestimating (\textit{a, b}) is reduced and increased for fatal underestimation (\textit{c}).
+This represents a risk-friendlier actor that is willing to bid on a higher estimate to attain a greater gain.
+For \textit{r} $>$ 1, the overestimation weight (\textit{a, b}) is increased in the loss function, the fatal underestimation weight (\textit{c}) is decreased and respectively more risk-averse actors are prompted to bid on lower estimates.\\
 		The factor \textit{r} can take basically any positive values.
@@
 		\section{1D geological reservoir modeling}\label{sec:1D_model}
-		For simple understanding and a preliminary assessment of the Bayesian statistical methods and the loss function described above, they are first to be applied in the context of an abstract one-dimensional reservoir case.
+		For simple understanding and a preliminary assessment of the Bayesian statistical methods and the loss function described above, they are first to be applied in the context of an conceptual one-dimensional reservoir case.
 The underlying model and basic approach are inherited from \citet{delaVarga2016}.
@@
 Other defining aspects, such as structural entrapment, are assumed to be given.
-Limiting the model to only one dimension and a small number of uncertain stratigraphical parameters allows for a relatively straightforward and simplified approach to assessing an abstract type of value for a reservoir and using the custom loss function (Equation \ref{eq:LF_final}) for value estimation.
+Limiting the model to only one dimension and a small number of uncertain stratigraphical parameters allows for a relatively straightforward and simplified approach to assessing an abstract type of value for a reservoir, applying the custom loss function (Equation \ref{eq:LF_final}) for value estimation and observing respective effects of Bayesian inference.
 The construction of the 1D geological model is described in the following.
@@
 These points confine two layers in the middle, from which the upper one can be labeled as seal and the lower one as reservoir.
-The resulting model with its possible layer boundary locations is illustrated in Figure \ref{fig:1D_model}.
+The resulting model with its prior distributions of possible layer boundary locations is illustrated in Figure \ref{fig:1D_model}.
 
@@
 
-	        \subsection{Abstract valuation (scoring) of the 1D geological model}\label{sec:1D_score_system}
+			\subsection{Bayesian inference and analysis of the 1D case}\label{sec:1D_bayes}
+			For conducting Bayesian inference, layer thickness likelihood functions are introduced based on another set of normal probability distributions.
+These are directly related to the layer boundary positions, i.e.
+the prior distributions.
+Means and standard deviations of the likelihood distributions are variable, so that several scenarios based on the assumption of different observations can be tested.
+Parameter priors and likelihood functions are embedded into a probabilistic modeling framework as described in Section \ref{sec:numerical_implementation} and MCMC sampling is executed accordingly.
+Posterior distributions can then be evaluated, based on the valuation method described in the following.
+
+	        \subsection{Abstract valuation of the 1D geological model}\label{sec:1D_score_system}
 	        For valuating a simplified 1D model, $OOIP$ and recoverable volume calculations are not applicable, as these require a 3D setting.
@@
 Thus, a condition to check whether the seal is reliable is included in the model.
-A respective equation for the reservoir score $S_{res}$ is defined as follows:
+A respective equation for the reservoir score $S_{res}$ is defined as:
 
@@
 
-			Where reservoir thickness is given by $h_{res}$, seal thickness by $h_{seal}$ and reservoir depth, which is always a negative value, by $d_{res}$.
-			This 1D model is then placed inside the Bayesian probabilistic modeling framework
+			where reservoir thickness is given by $h_{res}$, seal thickness by $h_{seal}$ and reservoir depth, which is always a negative value, by $d_{res}$.\\
+			This valuation method is applied on the original prior-only 1D model as reference and then on each posterior distribution set after conducting Bayesian inference.
+Resulting score probability distributions can be compared directly and additionally under use of the case-specific custom loss function (Equation \ref{eq:LFR_final}) defined before.
 
@@
 		According to the methods introduced in Sections \ref{sec:bayes} and \ref{sec:bayes_in_struc_modeling}, the 3D geological model is to be incorporated in a probabilistic modeling framework and analyzed from a Bayesian perspective.
+
+		First, a simple Monte Carlo error propagation run, using only priors, is conducted so that reference distributions

### 2017-10-26 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Start hydrocarbon traps text (methods)

@@
         	\item \textbf{Mathematical forward model ($M$)}: The connections between parameters $\theta$ and observed data $y$ are defined in such mathematical models.
-According to \citet{delaVarga2016} (and GREENSTONE PAPER, CITE?), the realization of a geological model $M$ can be regarded as a direct function of a set of input parameters:
+According to \citet{delaVarga2016}, the realization of a geological model $M$ can be regarded as a direct function of a set of input parameters:
         	\begin{equation}
@@
 A topological description, such as the relationships between faults and layer surfaces, is given by $\beta_l$.
-An essential aspect of this function is that it allows for a full automation of the modeling step, so that the consequences of a change in an input parameter are realized directly without the need for any further manual inspection or interaction (cite GREENSTONE paper).
+An essential aspect of this function is that it allows for a full automation of the modeling step, so that the consequences of a change in an input parameter are realized directly without the need for any further manual inspection or interaction \citep{wellmann2017sandstone}.
 The modeling step and respective interpolation used in this work is presented in Section \ref{sec:struc_geo_modeling} below.
@@
 Considering empirical Bayes methods, in which either priors or likelihoods stem from empirical data, as is assumed for the models in this work, convergence can be expected to be reached almost immediately.
-However, for high dimensional problems more iterations are required, in order to ensure an accurate representation of posterior distributions (cite Miguel GREENSTONE!).
+However, for high dimensional problems more iterations are required, in order to ensure an accurate representation of posterior distributions \citep{wellmann2017sandstone}.
 Different realizations of the full 3D geological model are then constructed on the basis of the resulting posterior distributions, using an implicit modeling step described in the following.
@@
         In this work, the generated structural models are to be assigned an economic significance in order to approach a real setting, in which hypothetical actors could have an interest in taking a decision, and thus, to evaluate the influence of the Bayesian inference of the decision-making process.
-Changes in Bayes actions are considered as a measure for this effect.\\
+Changes in Bayes actions are considered as a measure for this effect.
 For this purpose, it is assumed that a hydrocarbon reservoir system is represented by each geological model realization.
+
+        \subsubsection{Reservoirs and structural hydrocarbon traps}
+        Such hydrocarbon are typically comprised of:
+        \begin{enumerate}
+        	\item
+        \end{enumerate} \citep{dolson2016basics}
         These are typically comprised stratigraphically of at least one potential reservoir formation, succeeded by at least one sealing layer above.

### 2017-10-26 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Writing hydrocarbon system part, progress!

@@
 
-        \subsection{Economic significance and reservoir valuation}\label{sec:Reservoir_values}
+        \section{Model evaluation: Economic significance and reservoir valuation}\label{sec:Reservoir_values}
         In this work, the generated structural models are to be assigned an economic significance in order to approach a real setting, in which hypothetical actors could have an interest in taking a decision, and thus, to evaluate the influence of the Bayesian inference of the decision-making process.
@@
 
-        \subsubsection{Reservoirs and structural hydrocarbon traps}
-        Such hydrocarbon are typically comprised of:
+        \subsection{Hydrocarbon reservoir system and structural traps}
+        Such hydrocarbon systems are typically comprised of:
         \begin{enumerate}
-        	\item
-        \end{enumerate} \citep{dolson2016basics}
-        These are typically comprised stratigraphically of at least one potential reservoir formation, succeeded by at least one sealing layer above.
-Additionally, a trapping mechanism is normally required (source?).\\
-        The resulting reservoirs are to be valuated in a way that respective estimations on the basis of loss functions can be conducted, i.e.
+        	\item A kerogen-rich \textbf{source rock} which produced and emitted hydrocarbons during its burial history, given sufficient heating \citep{dolson2016basics} These are typically organic-rich sedimentary rocks, shales in particular.
+For the purpose of this work, it is assumed that a  suitable source rock is present and hydrocarbons expulsion took place after the formation of any trap-defining structures in our geological models.
+        	\item A \textbf{pathway} that allows expelled oil and gas to migrate upwards through the subsurface from the source rock into potential reservoir formations \citep{dolson2016basics}.
+This condition is also assumed to be given in our model.
+Basement formations in the geological model are simply defined as permeable.
+        	\item \textbf{Seal} formations that are able to halt hydrocarbons on their migration pathway.
+Such sealing is generally defined by a reduction of pore space geometry relative to migration energy.
+Lithologically, seals are most commonly comprised of shales, siltstones, tight carbonates, evaporites and salts.
+Sealing is thus defined primarily by changes in stratigraphical composition to low porosities and micro-sized pore throats, which are characterized by a high capillary entry pressure \citep{dolson2016basics, dolson2016quantifying}.
+An additional factor to sealing is posed by the presence of faults.
+Fault zones may possess significantly higher or lower permeabilities then surrounding rock formations which means that they may serve as fluid flow conduits or migration barriers or respectively \citep{van2003lateral, sorkhabi2005place}.
+The possible influence of faults on structural hydrocarbon traps is elaborated in further detail below.
+For the model in this work, the reservoir cap rock is assumed to be a shale.
+        	\item Suitable \textbf{reservoir formations}, in which these hydrocarbons volumes can be stored.
+For this, a suitable percentage of void space in a rock formation, i.e.
+a sufficiently high porosity, is required.
+A second significant factor is the connectivity of the porosity through pore throats.
+The larger the pore throats, the easier it is for fluids to migrate in a rock.
+Typical reservoir rocks are sandstones and carbonates, but even normally tight lithologies might serve as reservoirs, if they are fractured \citep{dolson2016basics}.
+In the models of this work, reservoir formations are assumed to be conventional sandstones.
+        \end{enumerate}
+        Given the migration and a combination of appropriate reservoir-seal sequence geometries, many different types of traps, in which hydrocarbons can accumulate, are possible.
+A common trap type is a structural four-way closure, i.e.
+as reservoir-seal system that was folded into closed anticlines, often resembling a dome and is thus predominantly defined by the structural closure relief and the cap rock sealing capacity.
+Such a trap type can be considered to be closed, as the reservoir surface contours are closed in a circle or oval around a relief high.
+Closure can also defined by any other intersection of a contour with a seal.
+The point at which a a contour fails to close is referred to as spill point \citep{dolson2016basics}.
+
+        In this work, reservoir systems and traps resulting from modeling are to be valuated in a way that respective estimations on the basis of loss functions can be conducted, i.e.
 Bayesian decision theory is applicable.
@@
 
-        \subsubsection{Original oil-in-place and recoverable volumes as value measures}
+        \subsection{Original oil-in-place and recoverable volumes as value measures}
         Before pressure and production tests have been conducted (i.e.

### 2017-10-27 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Trap figures inserted

@@
 Basement formations in the geological model are simply defined as permeable.
-        	\item \textbf{Seal} formations that are able to halt hydrocarbons on their migration pathway.
+        	\item \textbf{Seal} formations that are able to halt hydrocarbons on their migration pathway as top, lateral and possibly bottom seals \citep{dolson2016basics, sorkhabi2005place}.
 Such sealing is generally defined by a reduction of pore space geometry relative to migration energy.
-Lithologically, seals are most commonly comprised of shales, siltstones, tight carbonates, evaporites and salts.
-Sealing is thus defined primarily by changes in stratigraphical composition to low porosities and micro-sized pore throats, which are characterized by a high capillary entry pressure \citep{dolson2016basics, dolson2016quantifying}.
+Sealing is thus defined primarily by changes in stratigraphical composition to low porosities and micro-sized pore throats.These are characterized by a high capillary entry pressure able to resist the buoyancy pressure of hydrocarbons.
+Lithologically, seals are most commonly comprised of shales, siltstones, tight carbonates, evaporites and salts \citep{dolson2016basics, sorkhabi2005place, dolson2016quantifying}.
 An additional factor to sealing is posed by the presence of faults.
-Fault zones may possess significantly higher or lower permeabilities then surrounding rock formations which means that they may serve as fluid flow conduits or migration barriers or respectively \citep{van2003lateral, sorkhabi2005place}.
+Fault zones may possess significantly higher or lower permeabilities then surrounding rock formations which means that they may serve as fluid flow conduits or migration barriers respectively \citep{van2003lateral, sorkhabi2005place}.
 The possible influence of faults on structural hydrocarbon traps is elaborated in further detail below.
@@
 Such a trap type can be considered to be closed, as the reservoir surface contours are closed in a circle or oval around a relief high.
-Closure can also defined by any other intersection of a contour with a seal.
-The point at which a a contour fails to close is referred to as spill point \citep{dolson2016basics}.
-
+Such closure can also be given through any other crossing of a structural contour with a seal, including faults.
+A point at which a contour fails to close is referred to as "spill point" \citep{dolson2016basics}.
+        \begin{figure}[h]
+        	\centering
+        	\includegraphics[width=0.95\textwidth]{Figures/trap_closure.png}
+        	\caption{Conceptual examples of a dome-shaped four-way-closure and a simple fault trap from top view (relief countour) and cross section perspective (modified from \citet{dolson2016basics}.)}
+        \end{figure}        
+        The influence of faults and fault sealing potential on structural traps has been outlined by several researches such as \citet{sorkhabi2005place} and \citet{van2003lateral}.
+		\begin{figure}[h]
+			%\centering
+			\begin{subfigure}{.8\textwidth}
+				\centering
+				\includegraphics[width=1\linewidth]{Figures/fault_trap_mechA.png}
+				%\caption{Graph and saddle point of the function $z=x2~-~y2$.}\label{fig:saddle_point} 
+			\end{subfigure}
+			\begin{subfigure}{0.18\textwidth}
+				\centering
+				\includegraphics[width=1\linewidth]{Figures/fault_trap_mechB.png}
+				%\caption{Saddle point between two maxima.}\label{fig:saddle_point_between_maxima} 
+			\end{subfigure}
+			\caption{Bla}\label{fig:fault_trap_spills}
+		\end{figure}        
         In this work, reservoir systems and traps resulting from modeling are to be valuated in a way that respective estimations on the basis of loss functions can be conducted, i.e.

### 2017-10-28 Risk analysis/Text/Layout/Layout/chapter2_methods.tex

@@
 
+        \subsection{Bayesian hierarchical networks}\label{sec:hierarchical_bayes}
+
         \subsection{Markov chain Monte Carlo sampling (MCMC)}\label{sec:mcmc}
@@
 Such sealing is generally defined by a reduction of pore space geometry relative to migration energy.
-Sealing is thus defined primarily by changes in stratigraphical composition to low porosities and micro-sized pore throats.These are characterized by a high capillary entry pressure able to resist the buoyancy pressure of hydrocarbons.
+Sealing is thus defined primarily by changes in stratigraphical composition to low porosities and micro-sized pore throats.
+These are characterized by a high capillary entry pressure able to resist the buoyancy pressure of hydrocarbons.
+This capillary pressure is also the main controlling factor for seal capacity.
 Lithologically, seals are most commonly comprised of shales, siltstones, tight carbonates, evaporites and salts \citep{dolson2016basics, sorkhabi2005place, dolson2016quantifying}.
@@
 a sufficiently high porosity, is required.
-A second significant factor is the connectivity of the porosity through pore throats.
+Further significant factors are the connectivity of the porosity through pore throats, as well as pore geometry and resulting permeability \citep{dolson2016basics, sorkhabi2005place}.
 The larger the pore throats, the easier it is for fluids to migrate in a rock.
@@
 In the models of this work, reservoir formations are assumed to be conventional sandstones.
-        \end{enumerate}
-        Given the migration and a combination of appropriate reservoir-seal sequence geometries, many different types of traps, in which hydrocarbons can accumulate, are possible.
+        	\item A \textbf{trap} resulting from an appropriate reservoir-seal sequence and a three-dimensional geometric closure.
+Regarding a top view on a structural relief, closure is generally defined by any intersection of a structural contour and the seal, that is closed on both sides \citep{dolson2016basics, sorkhabi2005place}.
+A point at which a contour fails to close is referred to as "spill point".
+The maximum column depth to which a trap can be filled, and ultimately its maximum volume, is defined by this spill point.
+A trap that contains hydrocarbons down to the spill point is thus referred to as "filled to spill".
+traps can also be "filled to seal capacity" when the seal capacity is exceeded by the buoyant pressure of a fill column of greater height.
+Lower accumulations can also result from insufficient migration, i.e.
+low charge volumes of oil or gas in the first place.
+Such a trap is referred to as "charge limited" \citep{dolson2016basics}.
+Many different types of traps, in which hydrocarbons can accumulate, are possible.
 A common trap type is a structural four-way closure, i.e.
@@
 Such a trap type can be considered to be closed, as the reservoir surface contours are closed in a circle or oval around a relief high.
-Such closure can also be given through any other crossing of a structural contour with a seal, including faults.
-A point at which a contour fails to close is referred to as "spill point" \citep{dolson2016basics}.
+Closure can also be given through the intersection of a fault with the seal \citep{dolson2016basics}.
+Two common examples are illustrated in Figure \ref{fig:trap_types}.
+        \end{enumerate}
         \begin{figure}[h]
         	\centering
-        	\includegraphics[width=0.95\textwidth]{Figures/trap_closure.png}
-        	\caption{Conceptual examples of a dome-shaped four-way-closure and a simple fault trap from top view (relief countour) and cross section perspective (modified from \citet{dolson2016basics}.)}
+        	\includegraphics[width=0.95\textwidth]{Figures/trap_closure.PNG}
+        	\caption{Conceptual examples of a dome-shaped four-way-closure and a simple fault trap from top view (relief contour) and cross section perspective (modified from \citet{dolson2016basics}.)}\label{fig:trap_types}
         \end{figure}        
-        The influence of faults and fault sealing potential on structural traps has been outlined by several researches such as \citet{sorkhabi2005place} and \citet{van2003lateral}.
+        %The influence of faults and fault sealing potential on structural traps has been outlined by several researches such as \citet{sorkhabi2005place} and \citet{van2003lateral}.
+        The presence of a fault in the reservoir-seal sequence can have different impacts on trap closure and the volume of a hydrocarbon accumulation.
+This is mainly defined by the transport properties of a fault, i.e.
+in which way it behaves as a seal or flow conduit respectively.
+Cases might range from complete fault sealing to complete leakage along a fault zone.
+More complex in-between cases are possible, such as faults that leak across, but are sealed along their plane.
+This ultimately defines the depth to which a trap may be filled and whether this point is controlled by a point of leakage on the fault plane or a spill point related to the remainder of the 3D structural closure.
+Another decisive aspects to consider, it the nature of juxtapositions across the fault, especially regarding the possibility of cross-fault leakage \citep{van2003lateral}.
+This is summarized in Figure \ref{fig:fault_trap_spills}.\\
 		\begin{figure}[h]
@@
 				\centering
-				\includegraphics[width=1\linewidth]{Figures/fault_trap_mechA.png}
+				\includegraphics[width=1\linewidth]{Figures/fault_trap_mechA.PNG}
 				%\caption{Graph and saddle point of the function $z=x2~-~y2$.}\label{fig:saddle_point} 
@@
 				\centering
-				\includegraphics[width=1\linewidth]{Figures/fault_trap_mechB.png}
+				\includegraphics[width=1\linewidth]{Figures/fault_trap_mechB.PNG}
 				%\caption{Saddle point between two maxima.}\label{fig:saddle_point_between_maxima} 
 			\end{subfigure}
-			\caption{Bla}\label{fig:fault_trap_spills}
+			\caption{In (a), a structural trap resulting from the combination of 4-way anticlinal closure and a simple normal fault is illustrated.
+The two reservoir layers in both sides are juxtaposed, so that the hydrocarbon accumulation potential is controlled by three points of possible leak pathways: (1) the anticlinal spill point, (2) leakage upwards along the fault and (3) leakage across the fault enabled by the juxtaposition.
+Possibilities (2) and (3) depend on the transport properties of the fault zone, i.e.
+potential fault sealing.
+Given leakage along the fault (2), that size volume would be reduced to a small relatively small volume solely defined by the 4-way closure down to maximum contact of the reservoir-seal boundary with the fault plane.
+In case (b), the reservoirs are not juxtaposed and thus laterally sealed.
+Fault-related leakage can thus only occur along the plane.
+Leakage along and across the fault is enabled by the juxtaposition in (c) (modified from \citet{van2003lateral}).}\label{fig:fault_trap_spills}
 		\end{figure}        
-        In this work, reservoir systems and traps resulting from modeling are to be valuated in a way that respective estimations on the basis of loss functions can be conducted, i.e.
+		The transport properties of a fault zone are primarily defined by the internal structure and composition of its fault gouge.
+Fault gouge properties and ultimately the variation in permeability in the fault zone is primarily dependent on clay content.
+Clay smearing has thus been a central topic of research surrounding fault sealing (see \citep{yielding1997quantitative, van2003lateral, van2005processes, schmatz2010clay}.
+It generally is defined as all processes that lead to the incorporation of clay from the wall rock (e.g.
+a shale as seal) into the fault zone \citet{van2003lateral, vrolijk2016clay}.
+Clay smearing can be described using different equations, such as the Shale Smear Factor and the Shale Gouge Ratio \citep{yielding1997quantitative, vrolijk2016clay}.
+These are primarily based on the relation of displaced shale to fault throw (IF I STILL GET TO USE THESE, ELABORATE!).
+A critical review of the respective research published over the last four decades was recently presented by \citet{vrolijk2016clay}.\\
+        The 3D structural model in this work (see Section \ref{sec:3D_design}) is designed in a way that a potential conventional hydrocarbon trap is included in the form of anticlinal folding combined with normal faulting of a reservoir-seal sequence.
+Results from modeling are to be evaluated in a way that respective estimations on the basis of loss functions can be conducted, i.e.
 Bayesian decision theory is applicable.
 In the typical petroleum industrial setting, the main question is one of how much of the resource can be produced and how high the return on investment will be \citep{dean2007volumetric}.
+Thus, several of the factors named above are to be considered to attain a calculable volume of the structural trap which can further be economically interpreted using the approaches presented below.
 

### 2017-10-30 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Things

@@
         \end{equation}        
+        \begin{figure}[h]
+        	\centering
+        	\includegraphics[width=1\textwidth]{Figures/loss_functions.png}
+        	\caption{Expected loss based on the standard absolute-error loss (A) and squared-loss function (B).}\label{fig:standard_LF} 
+        \end{figure}
 
@@
         \end{enumerate}
-
-        \subsection{Bayesian hierarchical networks}\label{sec:hierarchical_bayes}
 
@@
 		\citet{salvatier2016pymc3} point out that the development of PyMC is continuing, as the inclusion of further tools is planned for future updates.
+
+		\subsection{Bayesian hierarchical networks}\label{sec:hierarchical_bayes}
 
@@
 		\begin{enumerate}
-			\item Step I: The standard symmetrical absolute-error loss function is chosen as a starting point for further customization steps:
+			\item \textbf{Step I}: The standard symmetrical absolute-error loss function is chosen as a starting point for further customization steps:
 
@@
 
-			\item Step II: Considering the development of a hydrocarbon reservoir, it can be assumed that over-investing is worse than under-investing.
+			\item \textbf{Step II}: Considering the development of a hydrocarbon reservoir, it can be assumed that over-investing is worse than under-investing.
 Overestimating the size of an accumulation might for example lead to the installation of equipment or facilites that are actually redundant or unnecessary.
@@
 
-			\item Step III: The worst case for any project would be that its development is set into motion, expecting a gain, only to discover later that the value in the reservoir does not cover the costs of realizing the project, resulting in an overall loss.
+			\item \textbf{Step III}: The worst case for any project would be that its development is set into motion, expecting a gain, only to discover later that the value in the reservoir does not cover the costs of realizing the project, resulting in an overall loss.
 A petroleum system might also turn out to be a complete failure, containing no value at all, although the actor's estimate indicated the opposite.
@@
 
-			\item Step IV: A worst case or fatal underestimation can also be derived from the idea of estimating a zero or negative value, when the true value is actually positive.
+			\item \textbf{Step IV}: A worst case or fatal underestimation can also be derived from the idea of estimating a zero or negative value, when the true value is actually positive.
 This is assumed to be worse than non-fatal overestimation, but clearly better than fatal overestimation.
@@
 		\end{enumerate}
+		\begin{figure}[p!]
+			\centering
+			\includegraphics[width=1\textwidth]{Figures/LF_4steps_normal.png}
+			\caption{The single steps of customizing the loss function are depicted in plots I to IV.}\label{fig:LF_4steps_normal}
+			\centering
+			\includegraphics[width=1\textwidth]{Figures/LF4_det_values.png}
+			\caption{Loss based on the customized loss function (Equation \ref{eq:LF_final}) for determined true scores of -750, 0 and 750.}\label{fig:LF4_det_values}
+		\end{figure}
 
@@
 		\end{equation}	
+		\begin{figure}[h]
+			\centering
+			\includegraphics[width=1\textwidth]{Figures/LFR_normal.png}
+			\caption{Plotting of expected loss realizations after including the risk factor $r$ in the loss function (Equation \ref{eq:LFR_final}) for actors with risk-affinities ranging from risk-averse ($r$ = 0.5 and 0.75), over risk-neutral ($r$ = 1), to risk-friendly ($r$ = 1.25 and $r$ = 1.5).}\label{fig:LFR_normal}
+		\end{figure}
 		According to this, for \textit{r} = 1 the risk-neutral loss function is returned, since \textit{a, b} and \textit{c} are not changed.

### 2017-10-30 Risk analysis/Text/Layout/Layout/chapter3_results.tex
Things

@@
 
-			\begin{figure}[p!]
-				\centering
-				\includegraphics[width=1\textwidth]{Figures/LF_4steps.png}
-				\caption{The single steps of customizing the loss function are depicted in plots I to IV.}\label{fig:LF_4steps}
-				\centering
-				\includegraphics[width=1\textwidth]{Figures/LF4_det_values.png}
-				\caption{Loss based on the customized loss function (Equation \ref{eq:LF_final}) for determined true scores of -750, 0 and 750.}\label{fig:LF4_det_values}
-			\end{figure}			
+			%\begin{figure}[p!]
+			%	\centering
+			%	\includegraphics[width=1\textwidth]{Figures/LF_4steps.png}
+			%	\caption{The single steps of customizing the loss function are depicted in plots I to IV.}\label{fig:LF_4steps}
+			%\end{figure}			
 			\begin{figure}[p!]

### 2017-10-31 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Worked on some Figures

@@
 Focus is laid on Bayesian inference, estimation of uncertain values and the use of loss functions in this context.
-These methods are incorporated in computational modeling of structural geological settings by programming in a python 3 environment.
-Central tools for model construction and conduction of the statistical evaluation are GeMpy and PyMC 2 in particular.
+These methods are incorporated in computational modeling of structural geological settings by programming in a Python 3 environment.
+Central tools for model construction and conduction of the statistical evaluation are GeMpy and PyMC in particular.
 These are also presented in this chapter.
@@
 	    Decisions are also referred to as actions $a$.
-	    The outcome of statistical tests in form of information or statistical evidence is denoted as $X$.	    
+	    The outcome of statistical tests in form of information or statistical evidence is denoted as $y$.	    
 	    Loss is defined as $L(\theta,a)$, so $L(\theta_1,a_1)$ is the actual loss incurred when action $a_1$ is taken while the true state of nature is $\theta_1$ \citep{berger2013stat}.
@@
 Thus, different and even contradicting beliefs about the probability of an event might be held by different individuals, based on variations and disparities in the information available to each one individual \citep{davidson2015}.\\
-        The initial belief or guess about an event A can be denoted as P(A) \citep{davidson2015}.
+        The initial belief or guess about an event $\theta$ can be denoted as $p(\theta)$ \citep{davidson2015, delaVarga2016}.
 This is used as the so-called prior probability on which Bayesian updating is based.
 The beliefs about the occurrence of an event are revalued in the presence of additional information, i.e.
-the observation of new evidence X.
-These observations are included as likelihoods P(X$|$A).
-This process of updating results in a posterior probability P(A$|$X) \citep{davidson2015}.
+the observation of new evidence $y$.
+These observations are included as likelihoods $p(y|\theta)$.
+This process of updating results in a posterior probability $p(\theta|y)$ \citep{davidson2015, delaVarga2016}.
 It is important to note that the prior is not simply discarded but re-weighted by Bayesian updating.
@@
 This means that Bayesian updating is about reducing uncertainty in a belief and reaching a guess that is less wrong \citep{davidson2015}.
-        Bayesian updating is defined by and conducted via the following equation, called the Bayes' Theorem:
-
+        Bayesian updating is defined by and conducted via the following equation, called the Bayes' Theorem \citep{davidson2015}:
         \begin{equation}\label{eq:BayesTheorem}
-        P(A|X) = \frac{P(X|A)P(A)}{P(X)}
-        \propto P(X|A)P(A)
+        p(\theta|y) = \frac{p(y|\theta)p(\theta)}{p(y)}
+        \propto P(y|\theta)p(\theta)
         \end{equation}
-
-        -Law of Large Numbers!
 
@@
         The magnitude of an estimate's loss is defined by a loss function, which is a function of the estimate of the parameter and the true value of the parameter \citep{davidson2015}:        
-
         \begin{equation}\label{eq:LossFunction}
@@
         \end{equation}        
-
         So, "how bad" a current estimate is, depends on the way a loss function weighs accuracy errors and returns respective losses.
 Two standard loss functions are the absolute-error and the squared-error loss function.
-Both are simple to understand and commonly used \citep{davidson2015}.
-
+Both are simple to understand and commonly used \citep{davidson2015}.\\
         As implied by its name, the absolute-error loss function returns loss as the absolute error, i.e. the difference between the estimate and the true parameter \citep{davidson2015}:        
-
         \begin{equation}\label{eq:AbsLossFunction}
@@
         \end{equation}                
-
         Accordingly, losses increasing linearly with the distance to the true value are returned for respective estimates.
-This means that all differences between relative errors are weighed equally, no matter whether they are found in the realm of relatively small or relatively large errors \citep{hennig2007}.
-
+This means that all differences between relative errors are weighed equally, no matter whether they are found in the realm of relatively small or relatively large errors \citep{hennig2007}.\\
         Using the squared-error loss function returns losses that increase quadratically with distance of the estimator to the true parameter value \citep{davidson2015}:        
-
         \begin{equation}\label{eq:SqrLossFunction}
@@
         \end{equation}         
+        This exponential growth of loss also means that large errors are weighed much stronger than small errors.
+This might come with over-valuation of distant outliers and misrepresentation of magnitudes in distance.
+Regarding this, the absolute-error loss function can be seen as more robust \citep{davidson2015}.
         \begin{figure}[h]
         	\centering
-        	\includegraphics[width=1\textwidth]{Figures/loss_functions.png}
-        	\caption{Expected loss based on the standard absolute-error loss (A) and squared-loss function (B).}\label{fig:standard_LF} 
+        	\includegraphics[width=0.5\textwidth]{Figures/abs_sqr_det.png}
+        	\caption{Realizations of loss based on the absolute-error loss function (blue) and the squared-error loss function (red) for a determined true value $\theta$ = 0.}\label{fig:abs_sqr_det} 
         \end{figure}    
-
-        This exponential growth of loss also means that large errors are weighed much stronger than small errors.
-This might come with over-valuation of distant outliers and misrepresentation of magnitudes in distance.
-Regarding this, the absolute-error loss function can be seen as more robust \citep{davidson2015}.\\
-        Both of these standard loss functions are symmetric and can be described as objectively aiming at a high precision in estimating the true parameter value.
+        Both of these standard loss functions are symmetric and can be described as objectively aiming at a high precision in estimating the true parameter value (see Figure \ref{fig:abs_sqr_det}).
         \citet{davidson2015} and \citet{hennig2007} propose that it might be useful to move away from these type of objective loss functions to the design of customized loss functions that specifically reflect an individual's (i.e.
@@
 Much more complicated designs of loss functions are possible, depending on purpose, objective and application \citep{davidson2015}.
-Some case-specific loss functions are designed in the following chapters of this work.
-\\
+Some case-specific loss functions are designed in the following chapters of this work.\\
         The presence of uncertainty during decision making implies that the true parameter is unknown and thus the truly incurred loss $L(\theta,a)$ cannot be known at the time of making the decision \citep{berger2013stat, davidson2015}.
@@
 \\
-        Given a posterior distribution $P(\theta|X)$, the expected loss of choosing an estimate $\hat{\theta}$ over the true parameter $\theta$ (after evidence X has been observed) is defined by the function below \citep{davidson2015}:
+        Given a posterior distribution $p(\theta|y)$, the expected loss of choosing an estimate $\hat{\theta}$ over the true parameter $\theta$ (after evidence $y$ has been observed) is defined by the function below \citep{davidson2015}:
 
@@
         \end{equation}
+
+        \begin{figure}[h]
+        	\centering
+        	\includegraphics[width=1\textwidth]{Figures/loss_functions.png}
+        	\caption{Expected loss based on the standard absolute-error loss (A) and squared-loss function (B).}\label{fig:standard_LF} 
+        \end{figure}
 
@@
 Such data can be gained by measurements and observations, for example by core sampling, well-log analysis or seismic acquisition.
-        	\item \textbf{Likelihood functions ($p(y|\theta)$}: Links between the previous parameters $\theta$ and the additional data $y$ are established by these functions in way that they reflect the likelihood of the parameter states given the observations.
+        	\item \textbf{Likelihood functions $p(y|\theta)$}: Links between the previous parameters $\theta$ and the additional data $y$ are established by these functions in way that they reflect the likelihood of the parameter states given the observations.
 They mathematically defined in the same way as probability functions, but are a function of the data $y$, instead of the parameters $\theta$ \citep{delaVarga2016}.
@@
         	\item \textbf{Setting up a full probability model}: A multi-dimensional joint probability space is to be generated, taking into account the probability distributions of every model parameter $\theta$.
-(- here, this not only results in the creation of geological models, but is subsequently also coupled with algorithms which deduct economic realizations from these models.)
+A 2D example is illustrated in Figure \ref{fig:joint_prob}.
+	        \begin{figure}[h]
+				\centering
+				\includegraphics[width=0.8\textwidth]{Figures/joint_prob.png}
+				\caption{Example visualization of a 2D joint probability space generated by two random parameters $\theta$ (from \citet{delaVarga2016}).}\label{fig:joint_prob} 
+			\end{figure}
         	\item \textbf{Conditioning on observed data}: Subsequently, an appropriate posterior distribution $p(\theta|y)$ is to be calculated by conditioning the parameters $\theta$ on the observed data $y$ given the likelihood $p(\theta|y)$.
@@
         \subsection{Markov chain Monte Carlo sampling (MCMC)}\label{sec:mcmc}
-        Despite the apparent simplicity of the Bayes Theorem, a direct analytical calculation and exact inference of the posterior distribution P(A$|$X) is rarely possible in non-idealized cases, due to intractability in multi-dimensional spaces \citep{hoffman2014no, delaVarga2016}.
+        Despite the apparent simplicity of the Bayes Theorem, a direct analytical calculation and exact inference of the posterior distribution $p(\theta|y)$ is rarely possible in non-idealized cases, due to intractability in multi-dimensional spaces \citep{hoffman2014no, delaVarga2016}.
 Thereby arises the necessity to resort to methods of statistical inference approximation.
@@
 Furthermore, $T$ is assumed to be a random function defined by polynomial drift and a stationary covariance $K(h)$ \citep{calcagno2008geological}.
-The use of a cubic covariance model is suggested by \citet{calcagno2008geological}, based on the results from studies conducted by \citet{aug2004modelisation} and \citet{chiles2004modelling}.\\
+The use of a cubic covariance model is suggested by \citet{calcagno2008geological}, based on the results from studies conducted by \citet{aug2004modelisation} and \citet{chiles2004modelling}.
 		\begin{figure}[h]
@@
 			\includegraphics[width=1\textwidth]{Figures/calcagno_pot_field.png}
-			\caption{woo a potential field! from \citet{calcagno2008geological}}\label{fig:pot_field}
+			\caption{Illustration of the concept of interpolating to attain a poentail field.
+The original data is depicted in (a), with contact points as dots and orientation measurements indicated by arrows.
+Colors represent respective assignments to different formations.
+An accordingly calculated potential field is shown in (b) (from \citet{calcagno2008geological}).}\label{fig:pot_field}
 		\end{figure}        
@@
         	\centering
-        	\includegraphics[width=0.95\textwidth]{Figures/trap_closure.PNG}
+        	\includegraphics[width=0.5\textwidth]{Figures/trap_closure1.PNG}
         	\caption{Conceptual examples of a dome-shaped four-way-closure and a simple fault trap from top view (relief contour) and cross section perspective (modified from \citet{dolson2016basics}.)}\label{fig:trap_types}
@@
 				\centering
-				\includegraphics[width=1\linewidth]{Figures/fault_trap_mechA.PNG}
+				\includegraphics[width=1\linewidth]{Figures/fault_trap_mechA1.PNG}
 				%\caption{Graph and saddle point of the function $z=x2~-~y2$.}\label{fig:saddle_point} 
@@
 				\centering
-				\includegraphics[width=1\linewidth]{Figures/fault_trap_mechB.PNG}
+				\includegraphics[width=1\linewidth]{Figures/fault_trap_mechB1.PNG}
 				%\caption{Saddle point between two maxima.}\label{fig:saddle_point_between_maxima} 
@@
 
+			%\begin{figure}[h]
+			%	\centering
+			%	\includegraphics[width=1\textwidth]{Figures/1D_model.png}
+%				\caption{Probability distributions for positions of layer boundaries in the subsurface and a respective representation using lines.}\label{fig:1D_model}
+%			\end{figure}
 			\begin{figure}[h]
 			\centering
-				\includegraphics[width=1\textwidth]{Figures/1D_model.png}
-				\caption{Probability distributions for positions of layer boundaries in the subsurface and a respective representation using lines.}\label{fig:1D_model}
+			\includegraphics[width=1\textwidth]{Figures/1D_principle.png}
+			\caption{ (from \citet{delaVarga2016}).}\label{fig:1D_principle}
 			\end{figure}
-
 			\subsection{Bayesian inference and analysis of the 1D case}\label{sec:1D_bayes}

### 2017-11-01 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Several updates in model, text and figures

@@
 Another decisive aspects to consider, it the nature of juxtapositions across the fault, especially regarding the possibility of cross-fault leakage \citep{van2003lateral}.
-This is summarized in Figure \ref{fig:fault_trap_spills}.\\
+This is summarized in Figure \ref{fig:fault_trap_spills}.
 		\begin{figure}[h]
@@
 Leakage along and across the fault is enabled by the juxtaposition in (c) (modified from \citet{van2003lateral}).}\label{fig:fault_trap_spills}
-		\end{figure}        
+		\end{figure}\\
 		The transport properties of a fault zone are primarily defined by the internal structure and composition of its fault gouge.
 Fault gouge properties and ultimately the variation in permeability in the fault zone is primarily dependent on clay content.
-Clay smearing has thus been a central topic of research surrounding fault sealing (see \citep{yielding1997quantitative, van2003lateral, van2005processes, schmatz2010clay}.
+Clay smearing has thus been a central topic of research surrounding fault sealing (see \citet{lindsay1993outcrop, yielding1997quantitative, van2003lateral, van2005processes, schmatz2010clay}.
 It generally is defined as all processes that lead to the incorporation of clay from the wall rock (e.g.
-a shale as seal) into the fault zone \citet{van2003lateral, vrolijk2016clay}.
-Clay smearing can be described using different equations, such as the Shale Smear Factor and the Shale Gouge Ratio \citep{yielding1997quantitative, vrolijk2016clay}.
-These are primarily based on the relation of displaced shale to fault throw (IF I STILL GET TO USE THESE, ELABORATE!).
+a shale as seal) into the fault zone \citep{van2003lateral, vrolijk2016clay}.
+Clay smearing can be described using different equations, such as the Shale Smear Factor ($SSF$) and the Shale Gouge Ratio ($SGR$) \citep{lindsay1993outcrop, yielding1997quantitative, vrolijk2016clay}.
+The $SSF$ provides a simple conceptual approach to assessing the clay smearing on a fault plane, dependent on ratio of fault throw magnitude $D$ to displaced shale thickness $T$ \citep{lindsay1993outcrop, yielding1997quantitative, yielding2012using}:
+		\begin{equation}
+			SSF = \frac{D}{T}
+		\end{equation}
+		It has been observed, that as the displacement and thus the ratio increases, a critical value $SSF_c$ will be reached, at which the clay smear is breached and no longer continuous.
+Initial observations by \citet{lindsay1993outcrop} indicated this threshold to be $SSF_c$ = 7.
+However, the critical value was later suggested to be scale-dependent and smaller for thick clays dispalced by large faults \citep{yielding2012using}.
+\citet{faerseth2006shale} reported that for such scales, including fault throws greater than 100~m, clay smear conitinuity can be expected for $SSF\le4$.
+Any larger values come with a decrease in confidence \citep{faerseth2006shale}.\\
 		A critical review of the respective research published over the last four decades was recently presented by \citet{vrolijk2016clay}.\\		
@@
 		This trap is defined as the central feature of economic interest, more specifically, the trap volume.
-For conducting simple and straightforward volumetric calculations, it is assumed that defined safe traps are always "full-to-spill", i.e.
+For conducting simple and straightforward volumetric calculations, it is assumed that defined safe traps are always "filled to spill", i.e.
 the complete trap volume is hydrocarbon filled and the $OOIP$ is attained over this volume.
@@
 In a Python environment, the data is imported as comma-separated values (CSV) using gemPy.
-From the generated data frame, gemPy is able to interpolate surfaces following the potential-field method explained in Section \ref{sec:struc_geo_modeling}) and compute voxel-based 3D geological models (see Figure \ref{fig:base3Dmodel_section}).
+From the generated data frame, gemPy is able to interpolate surfaces following the potential-field method explained in Section \ref{sec:struc_geo_modeling}) and compute voxel-based 3D geological models (see Figure \ref{fig:3Dmodel_section}).
 
@@
 			\centering
-			\includegraphics[width=1\textwidth]{Figures/base3Dmodel_plot.png}
+			\includegraphics[width=1\textwidth]{Figures/3Dmodel_section.png}
 			\caption{Mid section of the base 3D geological model as defined by the original data frame (see Appendix).
-Layer interfaces are defined by points and foliation dipping indicated by perpendicularly oriented arrows.}\label{fig:base3Dmodel_section}
+Layer interfaces are defined by points and foliation dipping indicated by perpendicularly oriented arrows.}\label{fig:3Dmodel_section}
 		\end{figure}
@@
 
-			\subsubsection{Finding the final spill point of a trap} %\label{sec:spill_leak_finding}
+			\subsection{Finding the final spill point of a trap} %\label{sec:spill_leak_finding}
 			Regarding dome-shaped anticlinal structures, it can be observed that, geometrically and mathematically, a spill point is a saddle point of the reservoir top surface in 3D.
+This was observed by \citet{collignon2015fold_linkage}, who pointed out that the linkage of folds is given by saddle points.
+These are thus a controling factor for spill-related migration from respective structural traps.
+For anticlinal traps, closure can consequently be defined as the distance between the saddle point (i.e.
+spill point) and maximal point of the trap \citep{collignon2015fold_linkage}.\\
 			Regarding a surface defined by $f(x,y)$, a local maximum at $(x_0,y_0,z_0)$ would resemble a hill top \citep{guichard2013calculus}.
@@
 			\begin{figure}[h]
-				%\centering
-				\begin{subfigure}{.5\textwidth}
 				\centering
-				\includegraphics[width=1\linewidth]{Figures/Saddle_point.png}
-				\caption{Graph and saddle point of the function $z=x2~-~y2$.}\label{fig:saddle_point} 
-			\end{subfigure}
-				\begin{subfigure}{.5\textwidth}
-				\centering
-				\includegraphics[width=1\linewidth]{Figures/Saddle_Point_between_maxima.png}
-				\caption{Saddle point between two maxima.}\label{fig:saddle_point_between_maxima} 
-				\end{subfigure}
-				\caption{Two examples for saddle points on surfaces of 3D functions (from Wikimedia Commons)}\label{fig:saddle_points_general}
+				\includegraphics[width=1\textwidth]{Figures/Saddle_Point_between_maxima.png}
+				\caption{Saddle point (red dot) between two maxima. In a geological setting, these surface maxima may be interpreted as two dome-shaped four-way closure traps.}\label{fig:saddle_point_maxima}
 			\end{figure}
-			Two examples for such saddle points are visualized in Figure \ref{fig:saddle_points_general}.
-Particularly Figure \ref{fig:saddle_point_between_maxima} can be seen as a representation of a point of spill between two dome structures (i.e.
+			%\begin{figure}[h]
+			%	%\centering
+			%	\begin{subfigure}{.5\textwidth}
+			%	\centering
+			%	\includegraphics[width=1\linewidth]{Figures/Saddle_point.png}
+			%	\caption{Graph and saddle point of the function $z=x2~-~y2$.}\label{fig:saddle_point} 
+			%\end{subfigure}
+			%	\begin{subfigure}{.5\textwidth}
+			%	\centering
+			%	\includegraphics[width=1\linewidth]{Figures/Saddle_Point_between_maxima.png}
+			%	\caption{Saddle point between two maxima.}\label{fig:saddle_point_between_maxima} 
+			%	\end{subfigure}
+			%	\caption{Two examples for saddle points on surfaces of 3D functions (from Wikimedia Commons)}\label{fig:saddle_points_general}
+			%\end{figure}		
+			Figure \ref{fig:saddle_point_maxima} can be seen as a representation of a point of spill between two dome structures (i.e.
 surface maxima or "hills") which is defined by the surface's saddle point between both.\\
@@
 This corresponds to the logical geometrical deduction, that a saddle point for a surface defined by $f(x,y)$ is marked by a local maximum in one plane, but a local minimum in the perpendicular plane.
-This is observable in Figure \ref{fig:saddle_points_general}.\\
+This is observable in Figure \ref{fig:saddle_point_flow}.\\
+			\begin{figure}[h]
+				\centering
+				\includegraphics[width=1\textwidth]{Figures/saddle_point_flow.png}
+				\caption{Surface and saddle point (red dot) generated by the function $z=x2~-~y2$.
+Maxima in direction $y$ is marked by a red, minima in direction $x$ by a white dashed line.
+Considering the surface as a seal, consequent buoyant flow directions are indicated by arrows.
+Such flow would be directed away from minima and towards maxima, with the saddle point indicating a junction and divide, i.e.
+a possible spill point.}\label{fig:saddle_point_flow}
+			\end{figure}
 			Regarding leakage across the normal fault in this geological model, a similar approach is taken.
@@
 However, if fault sealing is given, any fault leakage and respective leak points become irrelevant and only the anticlinal spill point is considered.
-The Z-value of the final spill point is used to determine the maximum down-to-horizon to which the trap feature can be filled.\\
-			\begin{figure}[h]
-				\centering
-				\includegraphics[width=1\textwidth]{Figures/restop_surface_spill_leak.png}
-				\caption{Plot of reservoir top surface points and determined anticlinal spill point (marked by 'X') and cross-fault leak point (marked by '+').}\label{fig:restop_surface_spill_leak}
-			\end{figure}\\	
+The Z-value of the final spill point is used to determine the maximum down-to-horizon to which the trap feature can be filled.
+			%\begin{figure}[h]
+			%	\centering
+			%	\includegraphics[width=1\textwidth]{Figures/restop_surface_spill_leak.png}
+			%	\caption{Plot of reservoir top surface points and determined anticlinal spill point (marked by 'X') and cross-fault leak point (marked by '+').}\label{fig:restop_surface_spill_leak}
+			%\end{figure}\\	
 			\citet{kuijper2004detecting} has pointed out that using finding spatial critical points, saddle points especially, can be problematic using discretized methods and rectangular grids in particular.
@@
 
-			\subsubsection{Calculating the maximum trap volume}
+			\subsection{Checking for juxtapositions and possible fault sealing}
+
+			\subsection{Calculating the maximum trap volume}
 			Have all voxels that fulfill the conditions defined above been determined, these voxels are seen as part of the maximum trap volume $V_{trap}$.

### 2017-11-01 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
bit more

@@
 		The concept of the 1D geological model is to be transferred and extended to a 3D setting that incorporates a more complete geological reservoir system.
-A three-dimensional space allows not only for better consideration of stratigraphical aspects, but also for the inclusion of structural formations, in particular reservoir trap-defining features.
+A three-dimensional space allows not only for better consideration of stratigraphical aspects, but also for the inclusion of structural formations, in particular hydrocarbon trap-defining features.
 
 		\subsection{Design of the 3D geological reservoir model}\label{sec:3D_design}
-		For the purpose of exploring the effect of the methods of Bayesian analysis and decision theory in this work, the model is nevertheless to be kept relatively simple.
+		For the purpose of exploring the effect of the methods of Bayesian analysis and decision theory in this work, the model is nevertheless to be kept conceptual and relatively simple.
 Stratigraphically, it is designed to include one main reservoir unit (sandstone), one main seal unit (shale), an underlying basement through which hydrocarbon fluids could have flown upwards and overlying formations that are assumed to be permeable, so that hydrocarbons can leak upwards.
-Structurally, it is constructed to feature an anticlinal dome fold that is displaced by a normal fault.
+Structurally, it is constructed to feature an anticlinal fold that is displaced by a normal fault.
 All layers are tilted so that they dip in the opposite direction of the fault plane dip.
 The original concept of this model is designed in a way, that a potential hydrocarbon trap is formed in the reservoir rock enclosed by the deformed seal and the normal fault.\\
-		This trap is defined as the central feature of economic interest, more specifically, the trap volume.
-For conducting simple and straightforward volumetric calculations, it is assumed that defined safe traps are always "filled to spill", i.e.
-the complete trap volume is hydrocarbon filled and the $OOIP$ is attained over this volume.
+		This trap, more specifically the trap volume, is defined as the central feature of economic interest.
+For conducting simple and straightforward volumetric calculations, it is assumed that found closed traps are always filled to spill, i.e.
+the complete trap volume is hydrocarbon filled and the $OOIP$ is attained over this volume.\\
 		The maximum trap volume is assumed to equate the hydrocarbon-filled rock volume $A * h$ in Equation \ref{eq:OOIP}.
@@
 The basic input data for the interpolation of the geological features is comprised of 3D point coordinates for layer interfaces and fault surfaces, as well as measurements that indicate respective dip directions and angles.
-In a Python environment, the data is imported as comma-separated values (CSV) using gemPy.
-From the generated data frame, gemPy is able to interpolate surfaces following the potential-field method explained in Section \ref{sec:struc_geo_modeling}) and compute voxel-based 3D geological models (see Figure \ref{fig:3Dmodel_section}).
+In a Python environment, the data is imported as comma-separated values (CSV) using GemPy.
+From the generated data frame, GemPy is able to interpolate surfaces following the potential-field method explained in Section \ref{sec:struc_geo_modeling} and compute voxel-based 3D geological models (see Figure \ref{fig:3Dmodel_section}).
 
@@
 		\subsection{Determination of the trap volume}
-		In the course of this work, several algorithms are developed within the python environment, that in combination enable the automatic recognition and calculation of trap volumes in geological models computed by gemPy.
+		In the course of this work, several algorithms are developed within the Python environment, that in combination enable the automatic recognition and calculation of trap volumes in geological models computed by GemPy.
 To assign voxels of the model to the trap volume, it is checked whether the following conditions are satisfied by each particular voxel:
@@
 In the case of fault sealing, however, the cross-fault leak point is discarded as irrelevant.
-The process of finding the decisive spill point is explained below in further detail.
+Given a juxtapostion with layers overlying the seal, due to fault displacement, the respective section is checked for fault sealing by taking into account the Shale Smear Factor value.
+The processes of finding the decisive spill point, as well as the determination of juxtapostions and $SSF$ is explained below in further detail.
 				        \item \textbf{Location inside of closed system:} The voxel is part of a model section inside of the main anticlinal feature.
@@
 A trap is thus defined as a closed system in this model.
-		\end{enumerate} 
 		It has to be emphasized that these conditions have been fitted to the geological model constructed as described in Section \ref{sec:3D_construction}.

### 2017-11-01 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
One more figure

@@
 
-        \section{Model evaluation: Economic significance and reservoir valuation}\label{sec:Reservoir_values}
+        \section{Model evaluation}
+
+        \subsection{Economic significance and reservoir system valuation}\label{sec:Reservoir_values}
         In this work, the generated structural models are to be assigned an economic significance in order to approach a real setting, in which hypothetical actors could have an interest in taking a decision, and thus, to evaluate the influence of the Bayesian inference of the decision-making process.
@@
 For this purpose, it is assumed that a hydrocarbon reservoir system is represented by each geological model realization.
-
-        \subsection{Hydrocarbon reservoir system and structural traps}
         Such hydrocarbon systems are typically comprised of:
@@
 Clay smearing can be described using different equations, such as the Shale Smear Factor ($SSF$) and the Shale Gouge Ratio ($SGR$) \citep{lindsay1993outcrop, yielding1997quantitative, vrolijk2016clay}.
-The $SSF$ provides a simple conceptual approach to assessing the clay smearing on a fault plane, dependent on ratio of fault throw magnitude $D$ to displaced shale thickness $T$ \citep{lindsay1993outcrop, yielding1997quantitative, yielding2012using}:
+The $SSF$ provides a simple conceptual approach to assessing the clay smearing on a fault plane, dependent on ratio of fault throw magnitude $D$ to displaced shale thickness $T$ \citep{lindsay1993outcrop, yielding1997quantitative, yielding2012using} (see Figure \ref{fig:ssf}):
 		\begin{equation}
@@
 		\end{equation}
+		\begin{figure}[h]
+			\centering
+			\includegraphics[width=0.7\textwidth]{Figures/SSF.PNG}
+			\caption{Conceptual fault section illustrating the SSF approach to assess shale smear potential (modified from \citet{vrolijk2016clay}.)}\label{fig:ssf}
+		\end{figure}
 		It has been observed, that as the displacement and thus the ratio increases, a critical value $SSF_c$ will be reached, at which the clay smear is breached and no longer continuous.

### 2017-11-02 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Wrote part on information entropy

@@
 				\centering
-				\includegraphics[width=0.8\textwidth]{Figures/joint_prob.png}
+				\includegraphics[width=0.8\textwidth]{Figures/joint_prob}
 				\caption{Example visualization of a 2D joint probability space generated by two random parameters $\theta$ (from \citet{delaVarga2016}).}\label{fig:joint_prob} 
@@
         \section{Model evaluation}
+        After conducting Bayesian inference, posterior distributions and resulting model realizations are to be evaluated.
+For this, the sampling traces and posterior distributions of single parameters can be observed.
+The entirety of realized 3D geological models is to be assessed in terms of uncertainty quantification.
+Furthermore, a central part of this work is the assignment of an economic significance to the modeling results.
+To achieve this, realized models are evaluated by regarding them as potential hydrocarbon systems.
+This process is further explained in Section \ref{sec:Reservoir_values}.
+Information entropy is introduced as a measure for uncertainty in the following.
+
+        \subsection{Visualizing uncertainty}
+        Information entropy was first defined by \citet{shannon1948mathematical} and adopted as a method to visualize uncertainties in 3D geological models by \citet{wellmann2012uncertainties}.
+This approach was later also used by \citet{delaVarga2016} and \citet{schaaf2017}.
+Information entropy is applied to predict the model accuracy at every location in the model space by visualizing its respective uncertainty and thus returning a measure for the quality of the model.
+This is achieved by subdiving the model space into a regular raster of equally sized cells (voxels) and measuring the accuracy for every such cell.
+In the context of geological modeling, the central question is based on the knowledge of how frequent a single geological feature occurs in a voxel \citep{wellmann2012uncertainties}.
+\citet{de1972definition} established that information entropy can serve as a measure of fuzziness, a concept described by \citet{zadeh1965fuzzy}.
+Given a fuzzy set in which fuzziness of each part is indicated by $f\epsilon[0,1]$, the following conditions hold:
+        \begin{enumerate}
+	        \item The measure should only be 0 if $f$ is 0 or 1 in every cell.
+	        \item The measure is maximal given $f = 0.5$ in every cell \citep{wellmann2012uncertainties}.
+        \end{enumerate}
+        While the first condition represents the complete absence of uncertainty everywhere in the model, the second condition is met in the opposite case of full uncertainty in the whole model space, i.e.
+all outcomes are equally likely in each cell \citep{schaaf2017}.
+From this, \citet{wellmann2012uncertainties} derived the following equation for application in geological modeling:
+        \begin{equation}
+	        H_m = -\frac{1}{N}*\sum_{x=1}^{N}[p_m(x)*\log(p_m(x))+(1-p_m(x))*\log(1-p_m(x))]
+        \end{equation}
+        In which $f$ is denoted as the probability $p_m$ of an outcome $m\epsilon M$ in a cell with a position $x$.
+The fuzziness is quantified as the entropy $H_m$, normalized by the total number of cells $N$.
+Subsequently, a total information entropy can be calculated for a model as:
+        \begin{equation}
+	        H_T = -\frac{1}{N}\sum_{x=1}^{N}H(x)
+        \end{equation}
+        This total entropy $H_T$ is 0 when no uncertainty is found in any cell, oppositely it is maximal in the case of equal likelihood of outcoms in every cell \citep{wellmann2012uncertainties, schaaf2017}.
+The application of these entropy measures allows for the calculation and visualization of uncertainties in each voxel, the assessment of uncertainties of entire geological units, as well as the quantification of total model uncertainty represented by a single number \citep{wellmann2012uncertainties}.
+        \begin{figure}
+	        \centering
+	        \includegraphics[width=1\textwidth]{Figures/information_entropy.png}
+	        \caption{Illustrations depicting the concept of information entropy to visualize uncertainties in spatial applications.
+In (a), a map composed of uncertain units is subdivided into a regular grid of equally sized cells.
+The probabilities of possible outcomes regarding each unit is estimated for each cell in (b).
+From this results a map of cell information entropies (c), with the highest values where all units can occur equally frequent, and lowest for cells where no uncertainty is given and entropy is zero (from \citet{wellmann2012uncertainties}).}\label{fig:information_entropy}
+        \end{figure}
 
@@
         	\centering
-        	\includegraphics[width=0.5\textwidth]{Figures/trap_closure1.PNG}
+        	\includegraphics[width=0.5\textwidth]{Figures/trap_closure1}
         	\caption{Conceptual examples of a dome-shaped four-way-closure and a simple fault trap from top view (relief contour) and cross section perspective (modified from \citet{dolson2016basics}.)}\label{fig:trap_types}
@@
 				\centering
-				\includegraphics[width=1\linewidth]{Figures/fault_trap_mechA1.PNG}
+				\includegraphics[width=1\linewidth]{Figures/fault_trap_mechA1}
 				%\caption{Graph and saddle point of the function $z=x2~-~y2$.}\label{fig:saddle_point} 
@@
 				\centering
-				\includegraphics[width=1\linewidth]{Figures/fault_trap_mechB1.PNG}
+				\includegraphics[width=1\linewidth]{Figures/fault_trap_mechB1}
 				%\caption{Saddle point between two maxima.}\label{fig:saddle_point_between_maxima} 
@@
 			\centering
-			\includegraphics[width=1\textwidth]{Figures/1D_principle.png}
+			\includegraphics[width=1\textwidth]{Figures/1D_principle.PNG}
 			\caption{ (from \citet{delaVarga2016}).}\label{fig:1D_principle}
@@
 A trap is thus defined as a closed system in this model.
+		\end{enumerate}
 		It has to be emphasized that these conditions have been fitted to the geological model constructed as described in Section \ref{sec:3D_construction}.
@@
 				\centering
-				\includegraphics[width=1\textwidth]{Figures/saddle_point_flow.png}
+				\includegraphics[width=1\textwidth]{Figures/Saddle_point_flow}
 				\caption{Surface and saddle point (red dot) generated by the function $z=x2~-~y2$.

### 2017-11-03 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Little progress in text

@@
 This Python library was developed for conducting Bayesian inference and prediction problems in an open-source probabilistic programming environment \citep{davidson2015, salvatier2016pymc3}.
-Different model fitting techniques are provided in PyMC, such as the \textit{maximum a posteriori} (MAP) method and several (MCMC) sampling methods, including the Adaptive Metropolis explained in Section \ref{sec:mcmc} \citep{salvatier2016pymc3}.
-The components which are used to construct a statistical model, are represented by \textit{Deterministic} or \textit{Stochastic} variables in PyMC.
+Different model fitting techniques are provided in PyMC, such as the \textit{maximum a posteriori} (MAP) method and several (MCMC) sampling methods, including the Adaptive Metropolis explained in Section \ref{sec:mcmc}.
+The components which are used to construct a statistical model, are represented by \textit{Deterministic} or \textit{Stochastic} variables in PyMC \citep{salvatier2016pymc3}.
+These variables are described as hierarchically related nodes, specifically parent and child nodes, in a Bayesian network:
+		\begin{enumerate}
+			\item \textbf{Parent} nodes include variables that affect other variables.
+			\item \textbf{Child} nodes contain variables the are influenced by other variables (they depend on parent variables) \citep{delaVarga2016}.
+		\end{enumerate}
+		Bayesian networks as a concept has been presented more in detail by \citet{koller2009probabilistic}.\\
 		The values of \textit{Deterministic} variables are completely dependent on its parents' values, as defined by a respective mathematical function \citep{salvatier2016pymc3}.

### 2017-11-06 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Reworked 1D results text.

@@
 The absolute-error loss seems most appropriate for this case of hydrocarbon reservoir value estimation.
-Ideally, an actor would like to know the exact true value of interest, say the OOIP, so that investments or resources can be allocated appropriately in order to acquire economic gains.
+Ideally, an actor would like to know the exact true value of interest, say the $OOIP$, so that investments or resources can be allocated appropriately in order to acquire economic gains.
 This allocation is the decision to be made or action to be taken.
@@
 For this reason, the absolute-error loss function is favored here over the squared-error loss function.
-It is chosen as the base function to which further development steps refer, based on mostly logical case-specific assumptions:
-
+It is chosen as the base function to which further development steps refer, based on mostly logical case-specific assumptions.
+The steps are illustrated in Figure \ref{fig:LF_4steps_normal} and listed in the following:
+		\begin{figure}[h]
+			\centering
+			\includegraphics[width=1\textwidth]{Figures/LF_4steps_normal.png}
+			\caption{Single steps (I-IV) of customizing the loss function applied on an exemplary normal distribution ($\mu=0; \sigma=500$).
+The median (in this case also mean and average) is returned for the Bayes action, when using the standard symmetric absolute-error loss function (I).
+Assigning a stronger weight on overestimation steepens the curve on the right hand side and shifts the minimum to the left, i.e.
+to a lower estimate (II).
+Steepening is significantly reinforced by the introduction of fatal overestimation in (III).
+This affects only the positive side of estimates and leads to a jump at zero, where signs change.
+Due to a similar condition, the same effect is observed on the negative side of estimate values, where the curve is also steepened, after including fatal underestimation (IV).
+Due to this, the Bayes action is shifted back to the zero estimate.
+The final custom loss function is represented by the curve in (IV).}\label{fig:LF_4steps_normal}
+			\centering
+		\end{figure}		
 		\begin{enumerate}
@@
 
-			(In other words: Fatal overestimation is twice as bad as simple underestimation.)
+			In other words: With \textit{b} (= 2), fatal overestimation is twice as bad as simple underestimation.
 
@@
 			\end{equation}			
-
 		\end{enumerate}
-		\begin{figure}[p!]
-			\centering
-			\includegraphics[width=1\textwidth]{Figures/LF_4steps_normal.png}
-			\caption{The single steps of customizing the loss function are depicted in plots I to IV.}\label{fig:LF_4steps_normal}
-			\centering
+		\begin{figure}[h]
 			\includegraphics[width=1\textwidth]{Figures/LF4_det_values.png}
-			\caption{Loss based on the customized loss function (Equation \ref{eq:LF_final}) for determined true scores of -750, 0 and 750.}\label{fig:LF4_det_values}
+			\caption{Loss based on the customized loss function (Equation \ref{eq:LF_final}) for determined true scores of -750, 0 and 750.
+This plot is meant to clarify the way real losses are incurred for each guess, relative to a given true score value.
+The expected loss, as seen in Figure \ref{fig:LF_4steps_normal}, is acquired by arithmetically averaging such deterministic loss realizations based on the true score probability distribution by using Equation \ref{eq:ExpectedLoss2}.}\label{fig:LF4_det_values}
 		\end{figure}
@@
 (Especially considering individual persons as actors, psychological aspects may play a significant role.)
-
-		(Estimate or prediction? True value only known if a "yes"-action (development) is taken.)
-		INSERT PLOTS BASED ON EXAMPLE NORMAL FUNCTIONS?
 
@@
 The construction of the 1D geological model is described in the following.
-
+			\begin{figure}[h]
+				\centering
+				\includegraphics[width=1\textwidth]{Figures/1D_model.png}
+				\caption{Probability distributions for positions of layer boundaries in the subsurface and a respective representation using lines.}\label{fig:1D_model}
+			\end{figure}
 			\subsection{Construction of the 1D geological model}\label{sec:1D_construction}	
@@
 The location probabilities for these points are defined by sampling from normal distributions.
-Standard deviations of these distributions increase with depth, representing an increase in uncertainty.
-For an approximate representation of a hydrocarbon reservoir system, the distribution means were set to depths of 2000 m (seal top), 2050 m (reservoir top) and 2200 m (reservoir bottom).
-STD-DEVIATIONS?
 These points confine two layers in the middle, from which the upper one can be labeled as seal and the lower one as reservoir.
+The standard deviations ($sigma$) of the distributions increase with depth, representing an increase in uncertainty.
+For an approximate representation of a hydrocarbon reservoir system, the distribution means ($\mu$) are set to depths of 2000 m (seal top, $\sigma=7$), 2050 m (reservoir top, $\sigma=25$) and 2200 m (reservoir bottom, $\sigma=45$).
 The resulting model with its prior distributions of possible layer boundary locations is illustrated in Figure \ref{fig:1D_model}.
-
-			%\begin{figure}[h]
-			%	\centering
-			%	\includegraphics[width=1\textwidth]{Figures/1D_model.png}
-%				\caption{Probability distributions for positions of layer boundaries in the subsurface and a respective representation using lines.}\label{fig:1D_model}
-%			\end{figure}
 			\begin{figure}[h]
@@
 			\subsection{Bayesian inference and analysis of the 1D case}\label{sec:1D_bayes}
-			For conducting Bayesian inference, layer thickness likelihood functions are introduced based on another set of normal probability distributions.
+			For conducting Bayesian inference, it is assumed that new observations have been made, providing additional information on the likelihoods of the thicknesses of the two layers of interest.
+Likelihood functions for reservoir and seal thicknesses are introduced based on another set of normal probability distributions.
 These are directly related to the layer boundary positions, i.e.
-the prior distributions.
+the prior parameter distributions.
 Means and standard deviations of the likelihood distributions are variable, so that several scenarios based on the assumption of different observations can be tested.
 Parameter priors and likelihood functions are embedded into a probabilistic modeling framework as described in Section \ref{sec:numerical_implementation} and MCMC sampling is executed accordingly.
-Posterior distributions can then be evaluated, based on the valuation method described in the following.
+Posterior distributions are then to be evaluated, based on the valuation method described in the following and subsequent application of the custom loss function.
 

### 2017-11-06 Risk analysis/Text/Layout/Layout/chapter3_results.tex
Reworked 1D results text.

@@
 	\chapter{Results}\label{cha:results}
-	Text
 
@@
 				\includegraphics[width=1\textwidth]{Figures/score_results1.png}
-				\caption{Posterior probability distributions from modeling scores using two (A) and three parameters (B).}\label{fig:score_results1}
+				\caption{Posterior probability distributions from modeling scores using two (A) and three parameters (B) (reservoir thickness, reservoir top depth and seal thickness with a safety threshold of 20~m).}\label{fig:score_results1}
 			\end{figure}
-			Results from scoring based on a 1D model constructed as defined in Section \ref{sec:1D_construction} are plotted in Figure \ref{fig:score_results1}.
-A test run of scoring with only the two parameters reservoir thickness and depth, is shown in \ref{fig:score_results1}-A.
-These results represented by an approximately normal distribution.
+			Results from scoring based on simple Monte Carlo error propagation using only the priors of the 1D model (as defined in Section \ref{sec:1D_construction}) are plotted in Figure \ref{fig:score_results1}.
+A test run of scoring with only the two parameters reservoir thickness and depth, is shown in Figure \ref{fig:score_results1} (A).
+These results are represented by an approximately normal distribution.
 The score is negative in about 17\% of the cases.
 Mean and median are about the same.\\
-			Full scoring results, including also seal reliability as a parameter, are visualized in Figure \ref{fig:score_results1}-B.
+			Full scoring results, including also seal reliability (with a threshold of 20~m) as a parameter, are visualized in Figure \ref{fig:score_results1} (B).
 The main distribution is not changed significantly, except for a striking peak of probability at the possibility for a score of -2000.
-This presumably represents the bulk of cases, in which the seal was assumed to have failed.
-It is to be noted, that the mean of the seal top distribution is found at -2000 m.
-It can be seen in Figure \ref{fig:1D_model}, that around that depth in the model column, reservoir top and seal top probability distributions significantly overlap.
-Thus, there is a possibility for a higher score, due to a shallower reservoir top position, but also a high probability for a seal that is to thin to be reliable.
-The negative score peak at -2000 is thus presumably caused by a high number of seal failures, due to both layer tops found in this area.
-Furthermore, mean and median of the score distribution have been shifted to lower values and are now found further apart.
-
-			\subsection{Applying the custom loss function on the 1D reservoir scores}
-
+This presumably represents the bulk of cases, in which the seal is assumed to have failed.
+Regarding this, it is to be noted, that the mean of the seal top distribution is found at -2000 m.
+It can be seen in Figure \ref{fig:update_moderate1} (A1), that around that depth in the model column, reservoir top and seal top probability distributions significantly overlap.
+Thus, there is a possibility for a higher score, due to a shallower reservoir top position, but also a high probability for a seal thickness below the safety cut-off threshold of 20~m.
+The negative score peak at -2000 is thus presumably caused by a high number of seal failures, due to both layer interfaces located within this range.
+Furthermore, as a consequence of this negative peak, mean and median of the score distribution have been shifted to lower values and are now found further apart (see Figure \ref{fig:score_results1} (B)).\\
+			Using this prior score distribution as a base, the custom loss function including risk was applied (see Figure \ref{fig:1D_LFR}).
+It can be observed that the minima for expected loss, i.e.
+the Bayesian estimators for differently risk-affine actors, are located at different estimates.
+Mean and median of the score distribution are clearly surpassed by the best estimate of the most risk-friendly actor (\textit{r} = 0.5), while for the most risk-averse actors (\textit{r} = 1.25 and \textit{r} = 1.5), the Bayes action equals a zero score estimate and thus the decision to take no action.
+It can also be recognized that the expected loss is generally lower for risk-friendlier actors on the side of positive estimates, which in this case is the relevant side for decision-making.
+			%\subsection{Applying the custom loss function on the 1D reservoir scores}
 			%\begin{figure}[p!]
@@
 			%\end{figure}			
-			\begin{figure}[p!]
+			\begin{figure}[h]
 				\centering
 				\includegraphics[width=1\textwidth]{Figures/LFR.png}
-				\caption{Plotting of expected loss realizations after including the risk factor $r$ in the loss function (Equation \ref{eq:LFR_final}) for actors with risk-affinities ranging from risk-averse ($r$ = 0.5 and 0.75), over risk-neutral ($r$ = 1), to risk-friendly ($r$ = 1.25 and $r$ = 1.5).}\label{fig:1D_LFR}
-				\centering
-				\includegraphics[width=1\textwidth]{Figures/update_moderate1.png}
-				\caption{Prior (A1) and posterior distributions (A2) of the layer boundary positions in depth and respective representations (A2, B2).}\label{fig:update_moderate1} 
+				\caption{Plotting of expected loss realizations after including the risk factor $r$ in the loss function  for actors with risk-affinities ranging from risk-averse ($r$ = 0.5 and 0.75), over risk-neutral ($r$ = 1), to risk-friendly ($r$ = 1.25 and $r$ = 1.5), based on the application of the custom loss function (Equation \ref{eq:LFR_final}) on a prior score distribution (see Figure \ref{fig:score_results1} (B)) from simple Monte Carlo error propagation.}\label{fig:1D_LFR}
 			\end{figure}			
-
-			Realizations of the four loss function adaption steps as elaborated in Section \ref{sec:LF_design} and based on the scoring results shown in Figure \ref{fig:score_results1}-B are depicted in the plots in Figure \ref{fig:LF_4steps}.
-			As expected, the median is returned for the Bayes action, when using the standard symmetric absolute-error loss function (Figure \ref{fig:LF_4steps}-I).
+			%Realizations of the four loss function adaption steps as elaborated in Section \ref{sec:LF_design} and based on the scoring results shown in Figure \ref{fig:score_results1}-B are depicted in the plots in Figure \ref{fig:LF_4steps}.
+			%As expected, the median is returned for the Bayes action, when using the standard symmetric absolute-error loss function (Figure \ref{fig:LF_4steps}-I).
 Compared to this, it can be observed that assigning a stronger weight on overestimation (Figure \ref{fig:LF_4steps}-II) steepens the curve on the right hand side and shifts the minimum to the left, i.e.
@@
 It is also to be noted that with every customization step, the overall expected loss is increased.\\
-			The implementation of the final custom loss function (Figure \ref{fig:LF_4steps}-IV) using single determined values for the true score is plotted in Figure \ref{fig:LF4_det_values}.
+			%The implementation of the final custom loss function (Figure \ref{fig:LF_4steps}-IV) using single determined values for the true score is plotted in Figure \ref{fig:LF4_det_values}.
 This helps to clarify the way real losses result for each guess, relative to given true score values.
 The expected loss is acquired by arithmetically averaging such loss realizations based on the true score probability distribution by using Equation \ref{eq:ExpectedLoss2}.\\
-			Implementing risk-affinity as a factor \textit{r} leads to different steepnesses of the plotted curves depicting loss and expected loss.
+			%Implementing risk-affinity as a factor \textit{r} leads to different steepnesses of the plotted curves depicting loss and expected loss.
 The effect on the latter is visualized in Figure \ref{fig:1D_LFR}, where 0.5, 0.75, 1, 1.25 and 1.5 were chosen as values for \textit{r}.
-			It can be observed that the minima for expected loss, i.e.
-the Bayesian estimators for differently risk-affine actors, are located at different estimates.
-Mean and median are clearly surpassed by the best estimate of the most risk-friendly actor (\textit{r} = 0.5), while for the most risk-averse actors (\textit{r} = 1.25 and \textit{r} = 1.5), the Bayes action equals a zero score estimate and thus the decision to take no action.
-In Figure \ref{fig:1D_LFR} it can also be recognized that the expected loss is generally lower for risk-friendlier actors on the side of positive estimates, which is the relevant side for decision-making.
 
-			\subsection{Bayesian updating using thickness likelihoods}
-			\citet{delaVarga2016} made use of Bayesian inference to reduce the uncertainty in this type of one-dimensional model. The same is conducted here in the following.\\				
-			The probability distributions for the location of the layer boundaries are treated as priors.
+			\subsection{Bayesian inference using thickness likelihoods}	
+			\begin{figure}[h]
+				\centering
+				\includegraphics[width=1\textwidth]{Figures/update_moderate1.png}
+				\caption{Prior (A1) and posterior distributions (A2) of the layer boundary positions in depth and respective representations (A2, B2).
+Bayesian inference was conducted using likelihoods defined as follows: Seal thickness: $\mu = 25~m; \sigma = 20~m$; reservoir thickness: $\mu = 180~m; \sigma = 60~m$.
+From (A1) to (B1), the distributions are slightly narrowed.
+Seal top and reservoir top distributions are still overlapping in (B1).
+A moderate reduction in uncertainty is also indicated by the representation in (B2), compared to (A2).}\label{fig:update_moderate1}
+			\end{figure}
+			%\citet{delaVarga2016} made use of Bayesian inference to reduce the uncertainty in this type of one-dimensional model.
+The same was conducted here, as shown in the following.The locations of the layer boundaries are treated as priors with respective probability distributions.
 Now it is assumed that new observations have been made, providing additional information on the likelihoods of the thicknesses of the two layers.
@@
 These parameters vary according to nature of the observations made.
-Using the principle of Bayesian inference as explained in Chapter \ref{sec:bayes}, the model is updated and new posterior distributions for our true reservoir score are attained\\
-			Different Bayesian updating cases with different sets of likelihoods are presented in the following, so that various possible results can be compared.
-Each take the layer boundaries defined in Section \ref{sec:1D_construction} as priors and use Bayesian inference with respective likelihood data defined below for each case.
+Using the principle of Bayesian inference as explained in Chapter \ref{sec:bayes}, the model is updated and new posterior distributions for our true reservoir score are attained.\\
+			Three representative Bayesian updating cases based on different sets of likelihoods in the 1D model are presented in the following.
+In each case, the layer boundaries defined in Section \ref{sec:1D_construction} were adopted as prior parameters.
+The various results from modeling with likelihoods, i.e.
+applying Bayesian inference, are compared to the original results and evaluations based on simple Monte Carlo error propagation using only priors.
 
-				\subsubsection{Updating case I: Moderately reinforcing information}
+				\subsubsection{Inference case I: Moderately reinforcing information}
 				\begin{figure}[h]
@@
 					\end{subfigure}
-					\caption{Reservoir score distributions (A) and change in the realizations of expected loss for several risk parameters (B1, B2) before and after Bayesian updating based on likelihoods defined as follows: Seal thickness mean = 25~m, std = 20~m.
-Reservoir thickness mean = 180~m, std = 60~m.}
+					\caption{Reservoir score distributions (A) and change in the realizations of expected loss for several risk parameters (B1, B2) before and after Bayesian updating based on likelihoods defined as follows: Seal thickness: $\mu = 25~m; \sigma = 20~m$; reservoir thickness: $\mu = 180~m; \sigma = 60~m$.}
 					\label{fig:update_moderate2_3}
@@
 
-				In this first case, a normal distribution with a mean of 25~m and a standard deviation of 20~m is chosen to reflect the likelihood of the seal thickness.
-For the reservoir, a normal distribution with a mean of 180~m and a standard deviation of 60~m is chosen.
-Using these likelihoods to update the 1D geological model, the uncertainty in the posterior probability distributions for the positions of layer boundaries in depth is reduced (see Figure \ref{fig:update_moderate1}).\\
-				Scoring is applied based on these new distributions.
-In Figure \ref{fig:update_moderate2_3}-A, it can be recognized that the bulk of the score distribution was shifted to the positive side of values, while the peak at -2000 was raised.
-The probability of scores between -2000 and 0 decreased significantly, i.e.
-the true score is most likely either positive or -2000 if it is negative.\\
-				Application of the custom loss function (Equation \ref{eq:LFR_final}) is visualized in Figure \ref{fig:update_moderate2_3}, in which the expected losses are compared before (B1) and after (B2) uncertainty reduction.
-It is observable, that by adding information about layer thickness likelihoods, Bayes actions are shifted relative to the nature of the information.
-In this case, the added data generally reinforces the probability of the reservoir to be significantly thick.
+				In this first case, a normal distribution with a mean ($\mu$) of 25~m and a standard deviation ($\sigma$) of 20~m was chosen to reflect the likelihood of the seal thickness.
+For the reservoir, a normal distribution with a mean of 180~m and a standard deviation of 60~m was used respectively.
+As can be observed in Figure \ref{fig:update_moderate1}), the uncertainty in the probability distributions for the positions of layer boundaries in depth was reduced only moderately by conducting Bayesian inference with these values.\\
+				Scoring was applied based on these new distributions.
+In Figure \ref{fig:update_moderate2_3} (A), it can be recognized that the bulk of the score distribution was shifted to the positive side of values, while the peak at -2000 was raised.
+The probability of scores between -2000 and 0 was decreased significantly, i.e.
+the true score is most likely either positive or -2000, if it is negative.\\
+				Application of the custom loss function (Equation \ref{eq:LFR_final}) is visualized in Figure \ref{fig:update_moderate2_3}, in which the expected losses are compared before (B1) and after (B2) Bayesian inference.
+It is observable, that by adding information about layer thickness likelihoods, Bayes actions were shifted relative to the nature of the information.
+In this case, the added data generally reinforces the probability of the reservoir to be sufficiently thick.
 Information on the seal, however, based on a normal distribution around 25~m thickness, leaves uncertainty about the reliability of the seal, as the safety threshold is defined as 20~m.\\
-				NOT INCREASED CERTAINTY; ONLY OVERALL VALUES!
-				Increased certainty about the reservoir thickness is sufficient to shift Bayes actions to higher estimates for all actors, but the most risk-averse one ($r$ = 1.5).
-These shifts are quantified in Table \ref{tab:update_examples_all}.
-According to these numbers, the risk-neutral actor's estimate is increased the most, the risk-friendliest actor's estimate the least.
-Expected losses are decreased for the risk-neutral and the two risk-friendlier individuals.
-It is clear that the expected loss was reduced the most for the risk-friendliest actor and it increased most significantly for the most risk-averse actor.\\
+				The decrease in uncertainty can overall be described as moderate.
+Increased certainty about the reservoir thickness was sufficient to shift Bayes actions to higher estimates for all actors, but the most risk-averse one ($r$ = 1.5).
+The risk-neutral actor's estimate was increased the most, the risk-friendliest actor's estimate the least.
+This has led to a slight convergence of the Bayes action in the realm of positive estimates.
+Expected losses remained at approximately the same magnitude, as before the likelihoods were considered.\\
+				%These shifts are quantified in Table \ref{tab:update_examples_all}.Expected losses are decreased for the risk-neutral and the two risk-friendlier individuals.
+It is clear that the expected loss was reduced the most for the risk-friendliest actor and it increased most significantly for the most risk-averse actor.
 				%\begin{table}
@@
 
-				\subsubsection{Updating case II: Likely reliable seal}
+				\subsubsection{Inference case II: Likely reliable seal}
 				\begin{figure}[h]
@@
 					\end{subfigure}
-					\caption{Reservoir score distributions (A) and change in the realizations of expected loss for several risk parameters (B1, B2) before and after Bayesian updating based on likelihoods defined as follows: Seal thickness mean = 50~m, std = 20~m.
-Reservoir thickness mean = 180~m, std = 60~m.}
+					\caption{Reservoir score distributions (A) and change in the realizations of expected loss for several risk parameters (B1, B2) before and after Bayesian updating based on likelihoods defined as follows: Seal thickness: $\mu = 50~m; \sigma = 20~m$; reservoir thickness: $\mu = 180~m; \sigma = 60~m$.}
 					\label{fig:update_goodseal2_3}
 				\end{figure}
-				In this second case, the reservoir thickness likelihood is defined in the same way as in case I (mean = 180~m, standard deviation = 60~m).
-For the seal, a mean of 50~m with a standard deviation of 20~m is chosen, favoring the likelihood of a reliable seal relative to the threshold of 20~m thickness.
-Respective score results are depicted in Figure \ref{fig:update_goodseal2_3}-A.
+				In this second case, the reservoir thickness likelihood was defined in the same way as in case I ($\mu = 180~m; \sigma = 60~m$).
+For the seal, a higher mean of 50~m with a standard deviation of 20~m was chosen, favoring the likelihood of a reliable seal relative to the threshold of 20~m thickness.
+Respective score results are depicted in Figure \ref{fig:update_goodseal2_3} (A).
 The bulk of the distribution is narrowed on the positive side of estimates.
-Mean and median are clearly shifted to higher values.
-The "seal failure peak" at -2000 is significantly decreased.\\
-				Applying the custom loss function on this new score distribution results in the realizations of expected loss illustrated in Figure \ref{fig:update_goodseal2_3}.
-Bayes actions are shifted clearly to higher estimates and expected losses of these minima are significantly reduced for all actors.
-This is quantified in Table \ref{tab:update_examples_all}.
-According to these values, the risk-neutral to risk-averse individuals seem to profit the most, due to a large change in estimate value and decreased expected loss.
-Compared to case I, the estimate shifts for both risk-friendlier actors remain approximately the same, expected losses, however, are lowered much more significantly.
-				POINT OUT: DECREASED UNCERTAINTY! AND HIGHER VALUES OVERALL!\\		
+Very apparent is the significant diminishment of the "seal failure peak" at -2000.
+Due to this in combination with the overall distribution narrowing, mean and median were clearly shifted to higher values.\\
+				Applying the custom loss function on this new score distribution resulted in the realizations of expected loss illustrated in Figure \ref{fig:update_goodseal2_3}.
+Bayes actions were shifted clearly to higher estimates and expected losses of these minima were significantly reduced for all actors.
+The risk-neutral to risk-averse individuals seem to have profited the most, due to a large change in estimate value and decreased expected loss.
+Even the most risk-averse decision-maker is now willing expect a positive true score.
+Compared to the foregone case, a higher degree of uncertainty reduction was achieved in this case.
+Consequently, Bayes actions were not only shifted to greater values, but also narrowed significantly in their range, i.e.
+the optimal decisions of the different actors have converged to a greater extent than in case I.\\
+				%This is quantified in Table \ref{tab:update_examples_all}.
+According to these values,  Compared to case I, the estimate shifts for both risk-friendlier actors remain approximately the same, expected losses, however, are lowered much more significantly.
 				%\begin{table}
@@
 
-				\subsubsection{Updating case III: Safe seal but likely subpar reservoir thickness}
+				\subsubsection{Inference case III: Safe seal but likely subpar reservoir thickness}
 				\begin{figure}[h]
@@
 					\end{subfigure}
-					\caption{Reservoir score distributions (A) and change in the realizations of expected loss for several risk parameters (B1, B2) before and after Bayesian updating based on likelihoods defined as follows: Seal thickness mean = 70~m, std = 10~m.
-Reservoir thickness mean = 100~m, std = 40~m.}
+					\caption{Reservoir score distributions (A) and change in the realizations of expected loss for several risk parameters (B1, B2) before and after Bayesian updating based on likelihoods defined as follows: Seal thickness: $\mu = 70~m; \sigma = 10~m$; reservoir thickness: $\mu = 100~m; \sigma = 40~m$.}
 					\label{fig:update_smallres2_3}
 				\end{figure}
-				In this third case, seal safety is ensured by using a mean of 70~m with a standard deviation of 10~m for the seal thickness likelihood.
-However, the new observations are assumed to provide information about the reservoir unit that makes it likely to be thinner than expected.
-Reservoir thickness likelihood is assigned a mean of 100~m and a standard deviation of 40~m.\\
-				The subsequent score distribution is depicted in Figure \ref{fig:update_smallres2_3}.
-It can be seen that while the whole distribution is narrowed, it also is shifted to the left, to lower and negative estimates.
-Mean and median are almost the same.
-As seal reliability is practically guaranteed, the peak at -2000~m vanishes.\\
-				Based on this reservoir score probability distribution, Bayes actions are shifted to lower estimates for all actors using the custom loss function (see Figure \ref{fig:update_smallres2_3}).
-In fact, all but the most risk-friendly individual (\textit{r}~=~0.5) find their Bayesian estimator to be zero now, i.e.
-project development is deemed to be too risky to them.
-Furthermore, the spread of expected loss values around the minima for all actors is diminished.
+				In this third case, seal safety was ensured by using a mean of 70~m with a standard deviation of 10~m for the seal thickness likelihood.
+However, the new observations were assumed to provide information about the reservoir unit that makes it likely to be thinner than expected.
+Reservoir thickness likelihood was assigned a mean of 100~m and a standard deviation of 40~m.\\
+				The subsequent score distribution is depicted in Figure \ref{fig:update_smallres2_3} (A).
+It can be seen that while the whole distribution was narrowed, it also was shifted to the left, to lower and negative estimates.
+Mean and median are almost equal, indicating narrowing towards symmetry in the shape of a normal distribution.
+As seal reliability is practically guaranteed, the peak at -2000 has vanished.\\
+				Based on this posterior score probability distribution, Bayes actions were shifted to lower estimates for all actors using the custom loss function (see Figure \ref{fig:update_smallres2_3} (B2)).
+The risk-neutral and both risk-averse individuals find their Bayesian estimators to be zero after updating, i.e.
+bidding on a positive score is deemed to be too risky to them.
+There is no shift in estimate for risk-averse actors, as they already found their best estimates to be zero before updating.
+Notable is the large reduction of expected loss in general.
+The range of Bayes actions was significantly narrowed.
+Overall, it is apparent that uncertainty was significantly reduced (see the narrowed distribution).
+However, although the seal can expected to be safe, the reservoir thickness is likely subpar.
+Due to this, the distribution was shifted to lower and negative values, strongly increasing the probability of an unfavorable true score.
+				%Quantified changes in the positions and values of the Bayes actions are listed in Table \ref{tab:update_examples_all}.
+Furthermore, the spread of expected loss values around the minima for all actors was diminished.
 In other words, the expected loss values of the Bayes actions are now much closer to each other.
-Notable is the large reduction of expected loss in general.
-Quantified changes in the positions and values of the Bayes actions are listed in Table \ref{tab:update_examples_all}.
-There is no shift in estimate for risk-averse actors, as they already found their best estimates to be zero before updating.
-				DECREASED UNCERTAINTY BUT AT THE SAME TIME LOWER VALUES OVERALL!\\
 				%\begin{table}
@@
 				%\end{table}
-
-				\begin{table}
-					\centering
-					\begin{tabular}[c]{| c || l | l || l | l || l | l |}
-						\hline
-						& Example I & & Example II & & Example III & \\
-						\hline
-						\hline
-						\textit{r} & $\Delta$ BA & $\Delta$ EL & $\Delta$ BA & $\Delta$ EL & $\Delta$ BA & $\Delta$ EL\\ 
-						\hline
-						0.50 & +~553.67 & -~105.49 & +~522.95 & -~436.26 & -~498.98 & -~538.08 \\ 
-						0.75 & +~673.40 & -~81.67 & +~697.73 & -~551.92 & -~346.16 & -~717.83  \\ 
-						1.00 & +~750.05 & -~33.56 & +~855.57 & -~631.92 & -~176.13 & -~872.02 \\ 
-						1.25 & +~713.19 & +~75.71 & +~912.43 & -~647.10 & -~0.00 & -~946.04 \\ 
-						1.50 & +~0.00 & +~282.64 & +~819.26 & -~545.96 & -~0.00 & -~951.57  \\ 
-						\hline
-					\end{tabular}
-					\caption{Changes in Bayes action (BA) and minimal expected loss (EL) for Bayesian updating cases I, II and III and respective actors with risk parameters $r$.}
-					\label{tab:update_examples_all}
-				\end{table}
+				% % % %
+				% % % %
+				%\begin{table}
+				%	\centering
+				%	\begin{tabular}[c]{| c || l | l || l | l || l | l |}
+				%		\hline
+				%		& Example I & & Example II & & Example III & \\
+				%		\hline
+				%		\hline
+				%		\textit{r} & $\Delta$ BA & $\Delta$ EL & $\Delta$ BA & $\Delta$ EL & $\Delta$ BA & $\Delta$ EL\\ 
+				%		\hline
+				%		0.50 & +~553.67 & -~105.49 & +~522.95 & -~436.26 & -~498.98 & -~538.08 \\ 
+				%		0.75 & +~673.40 & -~81.67 & +~697.73 & -~551.92 & -~346.16 & -~717.83  \\ 
+				%		1.00 & +~750.05 & -~33.56 & +~855.57 & -~631.92 & -~176.13 & -~872.02 \\ 
+				%		1.25 & +~713.19 & +~75.71 & +~912.43 & -~647.10 & -~0.00 & -~946.04 \\ 
+				%		1.50 & +~0.00 & +~282.64 & +~819.26 & -~545.96 & -~0.00 & -~951.57  \\ 
+				%		\hline
+				%	\end{tabular}
+				%	\caption{Changes in Bayes action (BA) and minimal expected loss (EL) for Bayesian updating cases I, II and III and respective actors with risk parameters $r$.}
+				%	\label{tab:update_examples_all}
+				%\end{table}
+				% % % %
+				% % % %
 				%\begin{table}
@@
 
-				NEED TO INCLUDE CLEAR BA VALUES BEFORE AND AFTER!!! NOT VISIBLE IN PLOTS.
-
-			\subsection{General 1D model updating results}
-			It is already shown by these three abstract 1D modeling cases, how Bayesian updating and change in uncertainties affects the way differently risk-affine decision-makers choose best estimates based on the custom loss function (Equation \ref{eq:LFR_final}).
-As can be expected intuitively, reassurance of positive scores, for example by narrowing of the distribution of scores on the side of positive estimates or by reducing the occurrence of negative extrema, leads to shifts towards higher Bayes action estimates for all actors.
+			\subsection{General 1D model inference results}
+			This conceptual application in an abstract 1D geological model already serves to illuminate some of the main effects of Bayesian inference on decision-making.
+Two main mechanisms were observable in these three cases:
+			\begin{enumerate}
+				\item \textbf{Lateral shifting} of the score distribution and consequent shift in Bayes actions along the axis of possible score estimates, independent from change in uncertainty.
+According to this observation, uncertainty reduction is not a necessary condition to significantly alter the decisions for all individuals.
+An incentive to bid on higher estimates is given solely by the shift of the posterior distribution to higher values, while the inherited degree of uncertainty may remain the same.
+				\item \textbf{Narrowing} of the score probability distribution, due to uncertainty reduction and consequent convergence of the Bayes actions of different actor's towards a similar estimate.
+Additionally, overall expected loss seems to decrease with uncertainty.
+If perfect information about our parameters was available, all individuals would bid on the same estimate and with an expected loss of zero.
+			\end{enumerate}
+			The second mechanism, i.e.
+uncertainty reduction, can thus presumed to be most significant regarding the process of decision-making, as actor's wish to minimize their expected loss and aim to bid on the most accurate estimate possible.
+In an optimal case, this would come in combination with a shift of probabilities (first mechanism) to more positive values, so that a higher economic value may be returned from the acquisition of information.
+			%As can be expected intuitively, reassurance of positive scores, for example by narrowing of the distribution of scores on the side of positive estimates or by reducing the occurrence of negative extrema, leads to shifts towards higher Bayes action estimates for all actors.
 Conversely, actors are discouraged to expect gains when facing a higher probability of negative scores.

### 2017-11-06 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Small text fixes

@@
 		\section{3D geological reservoir model}
-		The concept of the 1D geological model is to be transferred and extended to a 3D setting that incorporates a more complete geological reservoir system.
+		The concept of the 1D geological model is to be transferred and extended to a 3D setting that incorporates a more realistic and complete geological reservoir system.
 A three-dimensional space allows not only for better consideration of stratigraphical aspects, but also for the inclusion of structural formations, in particular hydrocarbon trap-defining features.
@@
 		\subsection{Design of the 3D geological reservoir model}\label{sec:3D_design}
-		For the purpose of exploring the effect of the methods of Bayesian analysis and decision theory in this work, the model is nevertheless to be kept conceptual and relatively simple.
-Stratigraphically, it is designed to include one main reservoir unit (sandstone), one main seal unit (shale), an underlying basement through which hydrocarbon fluids could have flown upwards and overlying formations that are assumed to be permeable, so that hydrocarbons can leak upwards.
+		For the purpose of exploring the application and effects of Bayesian analysis and decision theory in this work, the model is nevertheless to be kept conceptual and relatively simple.
+Stratigraphically, it is designed to include one main reservoir unit (sandstone), one main seal unit (shale), an underlying basement through which hydrocarbon fluids could have flown upwards and overlying formations that are assumed to be permeable, so that hydrocarbons can escape upwards.
 Structurally, it is constructed to feature an anticlinal fold that is displaced by a normal fault.
@@
 		This trap, more specifically the trap volume, is defined as the central feature of economic interest.
-For conducting simple and straightforward volumetric calculations, it is assumed that found closed traps are always filled to spill, i.e.
-the complete trap volume is hydrocarbon filled and the $OOIP$ is attained over this volume.\\
-		The maximum trap volume is assumed to equate the hydrocarbon-filled rock volume $A * h$ in Equation \ref{eq:OOIP}.
-$OOIP$ and recoverable oil volumes are thus to be used to valuate each realization of the geological model.
+For conducting simple and straightforward volumetric calculations, it is assumed that found closed traps are always filled to spill with oil, i.e.
+the complete trap volume is hydrocarbon filled and the $OOIP$ can be attained over this volume.\\
+		The maximum trap volume is assumed to equate the hydrocarbon-filled rock volume $A * h$ in Equation \ref{eq:OOIP} and is thus to be used to valuate each realization of the geological model.
 Single results are expected to vary, depending on the uncertain input parameters defined in Section \ref{sec:3D_construction} below.
@@
 This way, the effect of incorporating additional information in the form of likelihoods is to be assessed from a perspective of Bayesian decision theory.
-				- (to bring this Bayesian method of using loss functions closer to the traditional application of deterministic decision-making in decision-trees, the continuous estimation space of the reservoir value (volume) is combined with the assumption, that the estimated value is linked to a consequent investment into a determined project development size.
-thus, the continuous approach is used to make the decision in a deterministic decision-making space.
-)?
+It is furthermore assumed that the hydrocarbon trap volume is directly linked to a project development decision, i.e.
+investment and allocation of resources is represented by bidding on a volume estimate.
 

### 2017-11-07 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Working on 3D model text

@@
 
-		Uncertainties are implemented in several ways:
-		\begin{enumerate}
-		        \item \textbf{Z-values for layer interfaces:} Analogously to the 1D model, uncertainties for the vertical position of layer interfaces are implemented by using normal probability distributions.
+		This model allows for the consideration of various kinds of uncertainties and the implementation in several different ways.
+Here, they are primarily included by assigning uncertainties to the Z-positions of points which mark layer tops and interfaces in the 3D space.
+This is achieved by drawing from normal distributions that are
+		%\begin{enumerate}
+		        %\item \textbf{Z-values for layer interfaces:} Analogously to the 1D model, uncertainties for the vertical position of layer interfaces are implemented by using normal probability distributions.
 A deviation from the original input data Z-value is hereby added and allows for the consideration of uncertainties regarding layer surface positions in depth, layer thicknesses, topological shape and degree of fault offset.
@@
 The specific ditributions are listed in the Appendix.
-		        \item \textbf{Cross-fault sealing:} Uncertainty about the sealing capacity across the fault is implemented using a Bernoulli distribution.
+		        %\item \textbf{Cross-fault sealing:} Uncertainty about the sealing capacity across the fault is implemented using a Bernoulli distribution.
 In doing so, a probability of success is defined beforehand.
@@
 Should a value of zero be returned, cross-fault leakage is determined to be generally possible, depending on the occurrence of juxtapositions.
-		        \item \textbf{Other non-structural parameters:} Other factors that determine the $OOIP$ and $ROV$ calculations, such as porosity and recovery factor, can be implemented stochastically or deterministically.
+		        %\item \textbf{Other non-structural parameters:} Other factors that determine the $OOIP$ and $ROV$ calculations, such as porosity and recovery factor, can be implemented stochastically or deterministically.
 WHAT DO I USE?
-		\end{enumerate}
+		%\end{enumerate}
 

### 2017-11-07 Risk analysis/Text/Layout/Layout/chapter2_methods.tex

@@
     \chapter{Methods}\label{cha:met}
-
     The methods utilized in this work are presented in the following chapter.
-Bayesian analysis and decision theory are introduced.
+Bayesian analysis and decision theory are introduced in Section \ref{sec:bayes}.
 Focus is laid on Bayesian inference, estimation of uncertain values and the use of loss functions in this context.
-These methods are incorporated in computational modeling of structural geological settings by programming in a Python 3 environment.
-Central tools for model construction and conduction of the statistical evaluation are GeMpy and PyMC in particular.
-These are also presented in this chapter.
+    These methods are to be applied in the context of structural geological modeling.
+Respective approaches to computational modeling, specifically in a probabilistic framework, are described in Section \ref{sec:bayes_in_struc_modeling}.
+Regarding numerical implementation, GemPy and PyMC are presented as central tools used in a Python environment.
+In Section \ref{sec:model_evaluation}, methods to evaluate consequent modeling results are outlined.
+Particularly the potential for valuation from an economic perspective is elaborated in this regard and basic principles of subsurface hydrocarbon systems are introduced.
+Considering the respective setting in the economic sector of petroleum production, the development of an according case-specific custom loss function is described in Section \ref{sec:LF_design}.
+These methodologies are to be applied on a conceptual 1D geological model first (see Section \ref{sec:1D_model}).
+Design, construction and methods to identify hydrocarbon trap features in a full 3D structural model are explained in Section \ref{sec:3D_model}.
 
@@
 In statistical decision theory, available statistical knowledge is used to gain information on the nature of these uncertainties.
-(Such uncertain parameters can be considered as numerical quantities.) In order to find the best decision to a problem, it is possible to combine sample information with other aspects such as the possible consequences of decision-making and the availability of prior information on our uncertainties.
+Such uncertain parameters can be considered as numerical quantities.
+In order to find the best decision to a problem, it is possible to combine sample information with other aspects such as the possible consequences of decision-making and the availability of prior information on our uncertainties.
 Decision consequences are expressed as gains in economic decision theory and as losses, which equal negative gains, in statistics.
-Prior information might be given for example based on experience from previous similar problems or from expert knowledge (see Batvold and Begg).
+Prior information might be given for example based on experience from previous similar problems or from expert knowledge \citep{bratvold2010making}.
 The approach of utilizing priors is known as Bayesian analysis, which is explained in the following \citep{berger2013stat}.
-(It goes well with decision theory.)
+%(It goes well with decision theory.)
 
@@
 It has been argued by \cite{davidson2015}  that this Bayesian approach is intuitive and inherent in the natural human perspective.
-These beliefs can be assigned to individuals \citep{davidson2015}.
+These beliefs can be assigned to individuals.
 Thus, different and even contradicting beliefs about the probability of an event might be held by different individuals, based on variations and disparities in the information available to each one individual \citep{davidson2015}.\\
@@
 This means that Bayesian updating is about reducing uncertainty in a belief and reaching a guess that is less wrong \citep{davidson2015}.
-        Bayesian updating is defined by and conducted via the following equation, called the Bayes' Theorem \citep{davidson2015}:
+        Bayesian inference is defined by and conducted via the following equation, called the Bayes' Theorem \citep{davidson2015}:
         \begin{equation}\label{eq:BayesTheorem}
         p(\theta|y) = \frac{p(y|\theta)p(\theta)}{p(y)}
-        \propto P(y|\theta)p(\theta)
+        \propto p(y|\theta)p(\theta).
         \end{equation}
@@
         \subsection{Expected loss and loss functions}\label{sec:loss} 
-        Loss is a statistical measure of how bad an estimate is, i.e. how much is lost by making a certain decision. Gains are considered by statisticians as negative losses \citep{davidson2015}.
+        Loss is a statistical measure of how "bad" an estimate is, i.e. how much is lost by making a certain decision. Gains are considered by statisticians as negative losses.
         The magnitude of an estimate's loss is defined by a loss function, which is a function of the estimate of the parameter and the true value of the parameter \citep{davidson2015}:        
         \begin{equation}\label{eq:LossFunction}
-        L(\theta,\hat{\theta}) = f(\theta,\hat{\theta})
+        L(\theta,\hat{\theta}) = f(\theta,\hat{\theta}).
         \end{equation}        
-        So, "how bad" a current estimate is, depends on the way a loss function weighs accuracy errors and returns respective losses.
+        So, how "bad" a current estimate is, depends on the way a loss function weighs accuracy errors and returns respective losses.
 Two standard loss functions are the absolute-error and the squared-error loss function.
@@
         \begin{equation}\label{eq:AbsLossFunction}
-        L(\theta,\hat{\theta}) = |\theta - \hat{\theta}|
+        L(\theta,\hat{\theta}) = |\theta - \hat{\theta}|.
         \end{equation}                
@@
         \begin{equation}\label{eq:SqrLossFunction}
-        L(\theta,\hat{\theta}) = |\theta - \hat{\theta}|^2
+        L(\theta,\hat{\theta}) = |\theta - \hat{\theta}|^2.
         \end{equation}         
@@
         	\caption{Realizations of loss based on the absolute-error loss function (blue) and the squared-error loss function (red) for a determined true value $\theta$ = 0.}\label{fig:abs_sqr_det} 
-        \end{figure}    
+        \end{figure}\\
         Both of these standard loss functions are symmetric and can be described as objectively aiming at a high precision in estimating the true parameter value (see Figure \ref{fig:abs_sqr_det}).
@@
 Much more complicated designs of loss functions are possible, depending on purpose, objective and application \citep{davidson2015}.
-Some case-specific loss functions are designed in the following chapters of this work.\\
+A case-specific loss functions is designed in Section \ref{sec:LF_design} of this work.\\
         The presence of uncertainty during decision making implies that the true parameter is unknown and thus the truly incurred loss $L(\theta,a)$ cannot be known at the time of making the decision \citep{berger2013stat, davidson2015}.
@@
 all possible true values are represented by this distribution \citep{davidson2015}.
-A suitable alternative to the actual loss is to consider a decision's expected loss and to make a decision that is optimal in relation to this expected loss \citep{berger2013stat}.
+A suitable alternative to the actual loss is to consider each decision's expected loss and to make a decision that is optimal in relation to this expected loss \citep{berger2013stat}.
 \\
@@
         \begin{equation}\label{eq:ExpectedLoss}
-        l(\hat{\theta}) = E_{\theta}[L(\theta,\hat{\theta})]
+        l(\hat{\theta}) = E_{\theta}[L(\theta,\hat{\theta})].
         \end{equation}
@@
         The expectation symbol $E$ is subscripted with $\theta$, by which it is indicated that $\theta$ is the respective unknown variable.
-This expected loss as defined above, is also referred to as the risk of estimate $\hat{\theta}$ \citep{davidson2015}.
-
-        By the Law of Large Numbers, the expected loss of $\hat{\theta}$ can be approximated drawing a large sample size $N$ from the posterior distribution, respectively applying a loss function $L$ and averaging of the number of samples \citep{davidson2015}:
+This expected loss is also referred to as the risk of estimate $\hat{\theta}$ \citep{davidson2015}.\\
+        By the Law of Large Numbers, the expected loss of $\hat{\theta}$ can be approximated drawing a large sample size $N$ from the posterior distribution, respectively applying a loss function $L$ and averaging over the number of samples \citep{davidson2015}:
 
         \begin{equation}\label{eq:ExpectedLoss2}
-        \frac{1}{N}\sum_{i=1}^{N} L(\theta_i,\hat{\theta}) \approx E_{\theta}[L(\theta,\hat{\theta})] = l(\hat{\theta})
+        \frac{1}{N}\sum_{i=1}^{N} L(\theta_i,\hat{\theta}) \approx E_{\theta}[L(\theta,\hat{\theta})] = l(\hat{\theta}).
         \end{equation}
 
-        Minimization of a loss function returns a Bayesian point estimate known as Bayes action $\delta^P(X)$, which is the estimate, action or decision with the least expected loss according to the loss function \citep{berger2013stat}.
+        Minimization of a loss function returns a Bayesian point estimate known as Bayes action $\delta^p(y)$, which is the estimate, action or decision with the least expected loss according to the loss function \citep{berger2013stat}.
 For a unimodal and symmetric absolute-error loss function, the Bayes action is simply the median of the posterior distribution, while using squared-error loss it is the mean \citep{davidson2015, berger2013stat}.
@@
 By using different values for this parameter, it can be represented how comfortable an individual is with being wrong and furthermore which "side of wrong" is preferred by this decision maker \citep{davidson2015}.
-This approach to expressing risk-affinities is used the design of loss functions in this work.
+This approach to expressing risk-affinities is used for the design of the custom loss functions in Section \ref{sec:LF_design}.
 
@@
         	where $\phi_i$ is a function of position $\vec{x}$, more precisely an interpolation function, to which respective additional parameters are given by $\alpha_k$.
-Available primary geological information, such as positions of layer interfaces and dips in the subsurface, is represented by $k_j$.
+Available primary geological information, such as positions and dips of layer interfaces and in the subsurface, is represented by $k_j$.
 A topological description, such as the relationships between faults and layer surfaces, is given by $\beta_l$.
-An essential aspect of this function is that it allows for a full automation of the modeling step, so that the consequences of a change in an input parameter are realized directly without the need for any further manual inspection or interaction \citep{wellmann2017sandstone}.
-The modeling step and respective interpolation used in this work is presented in Section \ref{sec:struc_geo_modeling} below.
+An essential aspect of this function is that it allows for a full automation of the modeling step, so that the consequences of a change in an input parameter are realized directly without the need for any further manual inspection or interaction \citep{wellmann2017sandstone, delaVarga2016}.
+The modeling step and the respective interpolation method used in this work are presented in Section \ref{sec:struc_geo_modeling} below.
         	\item \textbf{Model parameters ($\theta$)}: These model-defining parameters can be deterministic or stochastic.
@@
 			\end{figure}
-        	\item \textbf{Conditioning on observed data}: Subsequently, an appropriate posterior distribution $p(\theta|y)$ is to be calculated by conditioning the parameters $\theta$ on the observed data $y$ given the likelihood $p(\theta|y)$.
+        	\item \textbf{Conditioning on observed data}: Subsequently, an appropriate posterior distribution $p(\theta|y)$ is to be calculated by conditioning the parameters $\theta$ on the observed data $y$ given the likelihood $p(y|\theta)$.
 This is the step of Bayesian updating of the belief about the parameter uncertainty given new information.
@@
 not all parameters need to be necessarily connected to all observed data.
-After all conditional probabilities have been set up, the Bayes Theorem (Equation \ref{eq:BayesTheorem}) is applied to attain the posterior.
+After all conditional probabilities have been set up, the Bayes Theorem (Equation \ref{eq:BayesTheorem}) is applied to attain the posterior citep{delaVarga2016}.
 However, due to the multi-dimensionality given in geological problems, the use of Markov chain Monte Carlo methods is advised to achieve this as described in Section \ref{sec:mcmc} below.
@@
 \citet{delaVarga2016} focused on the examination of the posterior distributions of the parameters $\theta$ and the generated models, particularly regarding information entropy within the model space.
-In this work, the geological models are assigned an economic meaning by declaring them potential petroleum reservoirs and introducing customized loss functions to reflect the economic interest of decision makers in developing respective resource extraction projects (see Section \ref{sec:loss}).
-Shifts in Bayes actions are considered a measure for the influence of Bayesian inference on decision-making and the significance of additional observations for different decision makers.
+In this work, the geological models are additionally assigned an economic meaning by declaring them potential petroleum reservoirs and introducing customized loss functions to reflect the economic interest of decision makers in developing respective resource extraction projects (see Section \ref{sec:loss}).
+Changes in Bayes actions are considered a measure for the influence of Bayesian inference on decision-making and the significance of additional observations for different decision makers.
         \end{enumerate}
@@
 Markov chain Monte Carlo (MCMC) sampling has proven to be a generally applicable and reliable method for exploring multi-dimensional parameter spaces in an intelligent way \citep{hoffman2014no, davidson2015}.
-\citet{gilks2005markov} has emphasized the significance of MCMC for the application of Bayesian statistics.
+\citet{gilks2005markov} has emphasized the significance of MCMC for the application in Bayesian statistics in particular.\\
         %In Monte Carlo integration, random samples are drawn from a target distribution, such as our posterior distribution P(A$|$X), to approximate the integral or probability density function in terms of an expectation:
-
         %\begin{equation}\label{eq:MonteCarlo}
@@
         The general principle of MCMC can be described as follows: 
-        Drawing representative samples from an target distribution of unknown shape is based on the conduction of a so-called random walk on the parameter distribution space.
+        Drawing representative samples from a target distribution of unknown shape is based on the conduction of a so-called random walk on the parameter distribution space.
 $T$ sampling steps are to be performed.
@@
 In the case of rejection, sampling is reverted to the previous accepted step \citep{schaaf2017, delaVarga2016}.
-After perform the complete number of $T$ iterations, all accepted sampling locations (i.e.
+After performing the complete number of $T$ iterations, all accepted sampling locations (i.e.
 the trace) are returned.
@@
         In Metropolis-Hastings methods, each sampling step at iteration $t$ is determined by a candidate probability distribution $q(\theta,\theta^{'})$, from which a proposed sample $\theta^{'}$ is drawn \citep{delaVarga2016}.
-The acceptance-rejection condition is defined by the acceptance ratio $a(\theta^{'},\theta)$:
+The acceptance-rejection condition is defined by the acceptance ratio $a(\theta^{'},\theta) \citep{delaVarga2016}$:
 
         \begin{equation}\label{eq:acceptance_ratio}
-        a(\theta^{'},\theta)=\frac{p(\theta^{'})p(y|\theta^{'})}{p(\theta)p(y|\theta)}
+        a(\theta^{'},\theta)=\frac{p(\theta^{'})p(y|\theta^{'})}{p(\theta)p(y|\theta)}.
         \end{equation}
@@
         To ensure a thorough exploration of the probability space, the transition to higher probability densities should not be enforced in every case but selectively.
-This is assured by relating the acceptance ratio from Equation \ref{eq:acceptance_ratio} to a random value $u$ from a Uniform distribution $U(0,1)$ as follows:
+This is assured by relating the acceptance ratio from Equation \ref{eq:acceptance_ratio} to a random value $u$ from a Uniform distribution $U(0,1)$ as follows \citep{delaVarga2016}:
 
@@
         \theta^{t} & $otherwise$\\
-        \end{cases}
+        \end{cases}.
         \end{equation}
@@
 Small steps have better acceptance rates, but lead to slower exploration and convergence of the algorithm \citep{delaVarga2016}.\\
-        For the iterative sampling, Adaptive Metropolis (AM) by \citet{haario2001adaptive} is used in this work.
+        For the iterative sampling in this work, Adaptive Metropolis (AM) by \citet{haario2001adaptive} is used.
 It adapts the traditional Metropolis-Hastings by incorporating the ability of continuous step size tuning during convergence, by taking into account the full information saved along the process.
 This is achieved by generating a covariance matrix that is updated every iteration.
-The adaptive nature of the process enables fast convergence for non-linear distributions while maintaining ergodicity \citet{haario2001adaptive, delaVarga2016}.
-Its suitability for multi-dimensional distribution spaces make it an excellent method for dealing with complex models such as the structural geological models in this work \citep{schaaf2017}.\\
-        A large enough number of iterations $T$ has to be chosen, so that a reliable and statistically significant exploration of the parameter space is assured.
+The adaptive nature of the process enables fast convergence for non-linear distributions while maintaining ergodicity \citep{haario2001adaptive, delaVarga2016}.
+Its suitability for multi-dimensional distribution spaces make it an excellent method for dealing with complex models such as structural geological models \citep{schaaf2017}.\\
+        For sampling, a large enough number of iterations $T$ has to be chosen, so that a reliable and statistically significant exploration of the parameter space is assured.
 This is primarily dependent on the rate of convergence towards the true distribution.
@@
 However, for high dimensional problems more iterations are required, in order to ensure an accurate representation of posterior distributions \citep{wellmann2017sandstone}.
-Different realizations of the full 3D geological model are then constructed on the basis of the resulting posterior distributions, using an implicit modeling step described in the following.
+Different realizations of the full 3D geological model are then constructed on the basis of the approximated posterior distributions, using an implicit modeling step described in the following.
 
@@
         This implicit method relies on the interpolation of a potential field scalar function $T(\vec{x})$ of any point $\vec{x}$ in a 3D space, and thus reflects the geometry of geological structures \citep{calcagno2008geological}.
-Modeling $T(\vec{x}$ is achieved by cokriging that regards two forms of data: (i) contact points on geological interfaces through increments of the potential field $T(\vec{x})-T(\vec{x}^{\,'})$ and (ii) orientation data as gradients, i.e.
+Modeling $T(\vec{x})$ is achieved by cokriging that regards two forms of data: (i) contact points on geological interfaces through increments of the potential field $T(\vec{x})-T(\vec{x}^{\,'})$ and (ii) orientation data as gradients, i.e.
 partial derivatives of the potential field $\delta T(\vec{x})/\delta u_\beta$ in each directon $u$ \citep{calcagno2008geological}.
@@
 Additionally, geological rules allow for the representation of several types of interactions between sets of geological layers \citep{calcagno2008geological}.\\
-        It is pointed out by \citet{calcagno2008geological}, that this method is particularly appropriate for cases in which knowledge about the geology is only given a sparse locations and is thus applicable for a wide variety of typical problem in geological settings.
+        It is pointed out by \citet{calcagno2008geological}, that this method is particularly appropriate for cases in which knowledge about the geology is only given for sparse locations and is thus applicable for a wide variety of typical problems in geological settings.
 Its suitability for this work is furthermore emphasized by the possibility to modify the topology-defining geological pile to achieve different geometric realizations without altering the data.
@@
 For doing so, the programming language of choice in this work is Python.
-The merits of Python have been pointed out by \citet{behnel2010, salvatier2016pymc3, Langtangen2008}.
+The merits of Python have been pointed out by \citet{behnel2010, Langtangen2008} and \citet{salvatier2016pymc3}.
 Development is facilitated by an expressive but concise and clean syntax that is easy to learn.
@@
 		The 3D geological modeling step in this work is implemented using GeMpy, an open-source, Python-based software that is able to generate and visualize complex 3D structural geological models based on the potential field interpolation method elaborated in Section \ref{sec:struc_geo_modeling} \citep{delaVarga2017gempy}.
-Its design allows for its application in a probabilistic setting, particularly by coupling it with PyMC as elaborated below.
+Its design allows for its application in a probabilistic setting as elaborated below.
 At the time of writing this, GemPy is still under development (version 0.995), but is already functioning for the purpose of this work.\\
@@
 This Python library was developed for conducting Bayesian inference and prediction problems in an open-source probabilistic programming environment \citep{davidson2015, salvatier2016pymc3}.
-Different model fitting techniques are provided in PyMC, such as the \textit{maximum a posteriori} (MAP) method and several (MCMC) sampling methods, including the Adaptive Metropolis explained in Section \ref{sec:mcmc}.
+Different model fitting techniques are provided in PyMC, such as the \textit{maximum a posteriori} (MAP) method and several (MCMC) sampling methods, including the Adaptive Metropolis introduced in Section \ref{sec:mcmc}.
 The components which are used to construct a statistical model, are represented by \textit{Deterministic} or \textit{Stochastic} variables in PyMC \citep{salvatier2016pymc3}.
@@
 			\item \textbf{Parent} nodes include variables that affect other variables.
-			\item \textbf{Child} nodes contain variables the are influenced by other variables (they depend on parent variables) \citep{delaVarga2016}.
+			\item \textbf{Child} nodes contain variables that are influenced by other variables (they depend on parent variables) \citep{delaVarga2016}.
 		\end{enumerate}
-		Bayesian networks as a concept has been presented more in detail by \citet{koller2009probabilistic}.\\
 		The values of \textit{Deterministic} variables are completely dependent on its parents' values, as defined by a respective mathematical function \citep{salvatier2016pymc3}.
@@
 Futhermore, PyMC allows for the creation of own object definitions inheriting from the class descriptions of these two variable types.\\
+		Bayesian networks as a concept have been presented in more detail by \citet{koller2009probabilistic}.
+\citet{salvatier2016pymc3} have pointed out that the development of PyMC is continuing, as the inclusion of further tools is planned for future updates.
 		%While conducting MCMC sampling, a burn-function can be implemented via PyMC, through which it can be defined that the $t_b$ first trace samples are discarded. This can be useful to d\\
-		\citet{salvatier2016pymc3} point out that the development of PyMC is continuing, as the inclusion of further tools is planned for future updates.
 
-		\subsection{Bayesian hierarchical networks}\label{sec:hierarchical_bayes}
-
-        \section{Model evaluation}
+		%\subsection{Bayesian hierarchical networks}\label{sec:hierarchical_bayes}
+		\section{Model evaluation}\label{sec:model_evaluation}
         After conducting Bayesian inference, posterior distributions and resulting model realizations are to be evaluated.
@@
 To achieve this, realized models are evaluated by regarding them as potential hydrocarbon systems.
-This process is further explained in Section \ref{sec:Reservoir_values}.
+This process is further in Section \ref{sec:Reservoir_values}.
 Information entropy is introduced as a measure for uncertainty in the following.
@@
         \subsection{Visualizing uncertainty}
+        \begin{figure}
+	        \centering
+	        \includegraphics[width=1\textwidth]{Figures/information_entropy.png}
+	        \caption{Illustrations depicting the concept of information entropy to visualize uncertainties in spatial applications.
+In (a), a map composed of uncertain units is subdivided into a regular grid of equally sized cells.
+The probabilities of possible outcomes regarding each unit is estimated for each cell in (b).
+From this results a map of cell information entropies (c), with the highest values where all units can occur equally frequent, and lowest for cells where no uncertainty is given and entropy is zero (from \citet{wellmann2012uncertainties}).}\label{fig:information_entropy}
+        \end{figure}
         Information entropy was first defined by \citet{shannon1948mathematical} and adopted as a method to visualize uncertainties in 3D geological models by \citet{wellmann2012uncertainties}.
@@
 \citet{de1972definition} established that information entropy can serve as a measure of fuzziness, a concept described by \citet{zadeh1965fuzzy}.
-Given a fuzzy set in which fuzziness of each part is indicated by $f\epsilon[0,1]$, the following conditions hold:
+Given a fuzzy set in which fuzziness of each part is indicated by $f\in [0,1]$, the following conditions hold:
         \begin{enumerate}
@@
         \begin{equation}
-	        H_m = -\frac{1}{N}*\sum_{x=1}^{N}[p_m(x)*\log(p_m(x))+(1-p_m(x))*\log(1-p_m(x))]
+	        H_m = -\frac{1}{N}*\sum_{x=1}^{N}[p_m(x)*\log(p_m(x))+(1-p_m(x))*\log(1-p_m(x))],
         \end{equation}
-        In which $f$ is denoted as the probability $p_m$ of an outcome $m\epsilon M$ in a cell with a position $x$.
+        in which $f$ is denoted as the probability $p_m$ of an outcome $m\in M$ in a cell with a position $x$.
 The fuzziness is quantified as the entropy $H_m$, normalized by the total number of cells $N$.
-Subsequently, a total information entropy can be calculated for a model as:
+Subsequently, a total information entropy can be calculated for a model as \citep{wellmann2012uncertainties}:
         \begin{equation}
-	        H_T = -\frac{1}{N}\sum_{x=1}^{N}H(x)
+	        H_T = -\frac{1}{N}\sum_{x=1}^{N}H(x).
         \end{equation}
@@
 The application of these entropy measures allows for the calculation and visualization of uncertainties in each voxel, the assessment of uncertainties of entire geological units, as well as the quantification of total model uncertainty represented by a single number \citep{wellmann2012uncertainties}.
-        \begin{figure}
-	        \centering
-	        \includegraphics[width=1\textwidth]{Figures/information_entropy.png}
-	        \caption{Illustrations depicting the concept of information entropy to visualize uncertainties in spatial applications.
-In (a), a map composed of uncertain units is subdivided into a regular grid of equally sized cells.
-The probabilities of possible outcomes regarding each unit is estimated for each cell in (b).
-From this results a map of cell information entropies (c), with the highest values where all units can occur equally frequent, and lowest for cells where no uncertainty is given and entropy is zero (from \citet{wellmann2012uncertainties}).}\label{fig:information_entropy}
-        \end{figure}
 
-        \subsection{Economic significance and reservoir system valuation}\label{sec:Reservoir_values}
-        In this work, the generated structural models are to be assigned an economic significance in order to approach a real setting, in which hypothetical actors could have an interest in taking a decision, and thus, to evaluate the influence of the Bayesian inference of the decision-making process.
+        \subsection{Economic significance of structural geological models as hydrocarbon systems}\label{sec:Reservoir_values}
+        In this work, the generated structural geological models are to be assigned an economic significance in order to approach a real setting, in which hypothetical actors could have an interest in taking a decision, and thus, to evaluate the influence of the Bayesian inference of the decision-making process.
 Changes in Bayes actions are considered as a measure for this effect.
@@
         \begin{enumerate}
-        	\item A kerogen-rich \textbf{source rock} which produced and emitted hydrocarbons during its burial history, given sufficient heating \citep{dolson2016basics} These are typically organic-rich sedimentary rocks, shales in particular.
+        	\item \textbf{Source rock:} A kerogen-rich source rock which produced and emitted hydrocarbons during its burial history, given sufficient heating, is needed to provide oil or gas in the first place \citep{dolson2016basics}.
+These are typically organic-rich sedimentary rocks, shales in particular.
 For the purpose of this work, it is assumed that a  suitable source rock is present and hydrocarbons expulsion took place after the formation of any trap-defining structures in our geological models.
-        	\item A \textbf{pathway} that allows expelled oil and gas to migrate upwards through the subsurface from the source rock into potential reservoir formations \citep{dolson2016basics}.
+        	\item \textbf{Migration pathway:} A pathway that allows expelled oil and gas to migrate upwards through the subsurface from the source rock into potential reservoir formations is necessary \citep{dolson2016basics}.
 This condition is also assumed to be given in our model.
-Basement formations in the geological model are simply defined as permeable.
-        	\item \textbf{Seal} formations that are able to halt hydrocarbons on their migration pathway as top, lateral and possibly bottom seals \citep{dolson2016basics, sorkhabi2005place}.
+Basement formations in the geological model are simply defined to be permeable.
+        	\item \textbf{Seal}: These are formations that are able to halt hydrocarbons on their migration pathway as top, lateral and possibly bottom seals \citep{dolson2016basics, sorkhabi2005place}.
 Such sealing is generally defined by a reduction of pore space geometry relative to migration energy.
@@
 This capillary pressure is also the main controlling factor for seal capacity.
-Lithologically, seals are most commonly comprised of shales, siltstones, tight carbonates, evaporites and salts \citep{dolson2016basics, sorkhabi2005place, dolson2016quantifying}.
+Lithologically, seals are most commonly comprised of shales, siltstones, tight carbonates, evaporites and salts \citep{dolson2016basics, dolson2016quantifying, sorkhabi2005place}.
 An additional factor to sealing is posed by the presence of faults.
-Fault zones may possess significantly higher or lower permeabilities then surrounding rock formations which means that they may serve as fluid flow conduits or migration barriers respectively \citep{van2003lateral, sorkhabi2005place}.
+Fault zones may possess significantly higher or lower permeabilities then surrounding rock formations, which means that they may serve as fluid flow conduits or migration barriers respectively \citep{van2003lateral, sorkhabi2005place}.
 The possible influence of faults on structural hydrocarbon traps is elaborated in further detail below.
-For the model in this work, the reservoir cap rock is assumed to be a shale.
-        	\item Suitable \textbf{reservoir formations}, in which these hydrocarbons volumes can be stored.
-For this, a suitable percentage of void space in a rock formation, i.e.
-a sufficiently high porosity, is required.
+For the model in this work, the cap rock is assumed to be a shale.
+        	\item \textbf{Reservoir formations:} Suitable lithologies, in which hydrocarbon volumes can be stored, have to be present.
+For this, a suitable percentage of void space, i.e.
+a sufficiently high porosity, is required in a rock formation.
 Further significant factors are the connectivity of the porosity through pore throats, as well as pore geometry and resulting permeability \citep{dolson2016basics, sorkhabi2005place}.
@@
 In the models of this work, reservoir formations are assumed to be conventional sandstones.
-        	\item A \textbf{trap} resulting from an appropriate reservoir-seal sequence and a three-dimensional geometric closure.
+        	\item \textbf{Trap:} Entrapment results from an appropriate reservoir-seal sequence and a three-dimensional geometric closure.
 Regarding a top view on a structural relief, closure is generally defined by any intersection of a structural contour and the seal, that is closed on both sides \citep{dolson2016basics, sorkhabi2005place}.
@@
 A trap that contains hydrocarbons down to the spill point is thus referred to as "filled to spill".
-traps can also be "filled to seal capacity" when the seal capacity is exceeded by the buoyant pressure of a fill column of greater height.
+Traps can also be "filled to seal capacity" when the seal capacity is exceeded by the buoyant pressure of a fill column of greater height.
 Lower accumulations can also result from insufficient migration, i.e.
@@
 Closure can also be given through the intersection of a fault with the seal \citep{dolson2016basics}.
-Two common examples are illustrated in Figure \ref{fig:trap_types}.
+Two common examples for traps are illustrated in Figure \ref{fig:trap_types}.
         \end{enumerate}
@@
         	\includegraphics[width=0.5\textwidth]{Figures/trap_closure1}
-        	\caption{Conceptual examples of a dome-shaped four-way-closure and a simple fault trap from top view (relief contour) and cross section perspective (modified from \citet{dolson2016basics}.)}\label{fig:trap_types}
+        	\caption{Conceptual examples of a dome-shaped four-way-closure and a simple fault trap from top view (relief contour) and cross section perspective (modified from \citet{dolson2016basics}.}\label{fig:trap_types}
         \end{figure}        
@@
 This ultimately defines the depth to which a trap may be filled and whether this point is controlled by a point of leakage on the fault plane or a spill point related to the remainder of the 3D structural closure.
-Another decisive aspects to consider, it the nature of juxtapositions across the fault, especially regarding the possibility of cross-fault leakage \citep{van2003lateral}.
+Another decisive aspects to consider, is the nature of juxtapositions across the fault, especially regarding the possibility of cross-fault leakage \citep{van2003lateral}.
 This is summarized in Figure \ref{fig:fault_trap_spills}.
@@
 		\begin{equation}
-			SSF = \frac{D}{T}
+			SSF = \frac{D}{T}.
 		\end{equation}
@@
 Initial observations by \citet{lindsay1993outcrop} indicated this threshold to be $SSF_c$ = 7.
-However, the critical value was later suggested to be scale-dependent and smaller for thick clays dispalced by large faults \citep{yielding2012using}.
-\citet{faerseth2006shale} reported that for such scales, including fault throws greater than 100~m, clay smear conitinuity can be expected for $SSF\le4$.
-Any larger values come with a decrease in confidence \citep{faerseth2006shale}.\\
-		A critical review of the respective research published over the last four decades was recently presented by \citet{vrolijk2016clay}.\\		
+However, the critical value was later suggested to be scale-dependent and smaller for thick clays displaced by large faults \citep{yielding2012using}.
+\citet{faerseth2006shale} reported that for such large scales, including fault throws greater than 100~m, clay smear continuity can be expected for $SSF\le4$.
+Any larger values come with a decrease in confidence \citep{faerseth2006shale}.
+A critical review of the research on clay smearing published over the last four decades was recently presented by \citet{vrolijk2016clay}.
+
         The 3D structural model in this work (see Section \ref{sec:3D_design}) is designed in a way that a potential conventional hydrocarbon trap is included in the form of anticlinal folding combined with normal faulting of a reservoir-seal sequence.
@@
         \begin{equation}\label{eq:OOIP}
-        OOIP = A * h * \phi * (1 - S_W) * 1/FvF
+        OOIP = A * h * \phi * (1 - S_W) * 1/FvF,
         \end{equation}
-        Where $OOIP$ is returned in $m³$.
-The hydrocarbon-filled rock volume is defined by the drainage area $A$ in $m²$ and the net pay thickness $h$ in $m$.
-Porosity $\phi$ and water saturation $S_W$ (interstitial water) are given in fraction of the rock volume.
+        where $OOIP$ is returned in $m^3$.
+The hydrocarbon-filled rock volume is defined by the drainage area $A$ in $m^2$ and the net pay thickness $h$ in $m$.
+Porosity $\phi$ and water saturation $S_W$ (interstitial water) are given in fractions of the rock volume.
 A dimensionless factor for the change in oil volume between reservoir conditions and standard conditions at surface is represented by the formation volume factor $FvF$.
@@
         \begin{equation}\label{eq:ROV}
-                ROV = OOIP * RF = A * h * \phi * (1 - S_W) * 1/FvF * RF
+                ROV = OOIP * RF = A * h * \phi * (1 - S_W) * 1/FvF * RF.
         \end{equation}
-        the recovery factor is influenced by a number of fluid properties, such as viscosity, density, solution oil/gas ratio and the formation volume factor.
+        The recovery factor is influenced by a number of fluid properties, such as viscosity, density, solution oil/gas ratio and the formation volume factor.
 Thus, it is difficult to estimate \citep{dean2007volumetric}.
@@
 For gas accumulations, however, recovery factors range typically between 70 and 90\% \citep{dean2007volumetric}.\\
-        The models in this work are completely hypothetical and not based on real data.
+        The models in this work are artificial and not based on real data.
 The inputs to these equations can thus be chosen arbitrarily to test the applied methods.
@@
 Furthermore, as 3D geological structures are modeled here, the hydrocarbon-filled rock volume ($A * h$ in Equations \ref{eq:OOIP} and \ref{eq:ROV}) is defined by every realization of the uncertain geological model.
-The rock volume is thus an uncertain factor.
-Remaining factors can be implemented deterministically or as uncertain values, too.
-For most factors, it seems most suitable to define a normal distribution around a known typical value or global average, as it will return a high probability for common scenarios and occasionally produce cases of rare, more extreme values.
-(INSERT WHAT IS CHOSEN IN THE END - MAYBE MORE SENSE TO TAKE FIXED VALUES FOR THE FACTORS THEN RV, SO THAT ITS INFLUENCE IS CLEAR; THEN MAYBE INCLUDE POROSITY?
-OR COMPARE BOTH POSSIBILITIES?).\\
+The hydrocarbon-filled rock volume is thus an uncertain factor.
+Remaining factors can be implemented deterministically or as uncertain values as well.
+WHAT DO?\\%For most factors, it seems most suitable to define a normal distribution around a known typical value or global average, as it will return a high probability for common scenarios and occasionally produce cases of rare, more extreme values.\\
         Consequently, $OOIP$ and $ROV$ calculations represent the step following and using the results from computing the geological model, so that the realizations are assigned values that can be interpreted from a perspective of economic interest.
 The resulting values from numerous model runs are used as a base probability distribution for the true value, on which loss function estimation can be applied as described in Section \ref{sec:loss}.
-$OOIP$ and $ROV$ are applicable in a 3D setting, the 1D geological model introduced below in Section \ref{sec:1D_model}, however, requires the use of an abstract valuation system that is explained in Section \ref{sec:1D_score_system}.
+$OOIP$ and $ROV$ are applicable in a 3D setting, the 1D geological model introduced below in Section \ref{sec:1D_model}, however, requires the use of an abstract valuation system.
         %\subsubsection{Net present value}
@@
 
-		\section{3D geological reservoir model}
+		\section{3D geological reservoir model}\label{sec:3D_model}
 		The concept of the 1D geological model is to be transferred and extended to a 3D setting that incorporates a more realistic and complete geological reservoir system.
@@
 
-		This model allows for the consideration of various kinds of uncertainties and the implementation in several different ways.
-Here, they are primarily included by assigning uncertainties to the Z-positions of points which mark layer tops and interfaces in the 3D space.
-This is achieved by drawing from normal distributions that are
+		This model allows for the consideration of various kinds of uncertainties and their implementation in several different ways.
+Here, they are included by assigning uncertainties to the Z-positions of points which mark layer tops and interfaces in the 3D space.
+This is achieved based on respective probability distributions from which deviation values are drawn, which are then added to the original input data Z-value.
+Uncertainties regarding layer surface positions in depth, layer thicknesses, topological shape and degree of fault offset can be incorporated.
+These can be assigned as homogeneous sets to groups which are to share a common degree of uncertainty.
+Here, points belonging to the same layer interface are assigned the same base uncertainty by applying one shared probability distribution.
+Assuming an increase of uncertainty with depth, standard deviations for the distributions are increased for lower formations.
+Furthermore, uncertainty regarding the magnitude of fault offset is incorporated by adding a skew normal probability distribution that is shared by all layer interface points on the hanging wall.
+A left skewed ($\alpha=-4$) normal distribution is chosen to reflect the nature of throw on a normal fault, particularly the slip motion of the hanging wall block.
+Mainly negative values are returned by this distribution.
+This way, the offset nature of the normal fault is maintained and the inversion to a reverse fault is avoided.
+Specific ditributions used to represent uncertainties are listed in the Appendix.
 		%\begin{enumerate}
@@
 		\begin{enumerate}
-			\item \textbf{Labeled as reservoir formation:} The voxel has been assigned to the target reservoir formation in the lithology block model.
-This is determined by respective labeling of the input data and the computation conducted by gemPy.
+			\item \textbf{Labeled as reservoir formation:} The voxel has been assigned to the target reservoir formation (here: Sandstone 1 in Figure \ref{fig:3Dmodel_section}) in the lithology block model.
+This is determined by respective labeling of the input data and the computation conducted by GemPy.
 			\item \textbf{Location in footwall:} The voxel is located on the footwall side of the fault.
@@
 Due to respective dip of the reservoir formation, full leakage is assumed for the reservoir formation in the hanging wall.
-A distinction between foot- and hanging wall is easily achievable by using the fault model block that is generated by gemPy.
+A distinction between foot- and hanging wall is easily achievable by using the fault model block that is generated by GemPy.
 In this fault block, voxels on both sides are assigned respective values for distinction.
@@
 Once both of these points have been determined, the higher one is defined to be the final spill point used to determine the maximum fill capacity of the trap.
-In the case of fault sealing, however, the cross-fault leak point is discarded as irrelevant.
 Given a juxtapostion with layers overlying the seal, due to fault displacement, the respective section is checked for fault sealing by taking into account the Shale Smear Factor value.
+In the case of a suitable $SSF$ and consequent fault sealing, leakage across this type of juxtaposition is discarded as irrelevant.
 The processes of finding the decisive spill point, as well as the determination of juxtapostions and $SSF$ is explained below in further detail.
 			\item \textbf{Location inside of closed system:} The voxel is part of a model section inside of the main anticlinal feature.
-All of the voxels in this section are separated from the borders of the model by voxels that don't meet the conditions above, which primarily means that they are encapsulated by seal voxel laterally and upwards.
+All of the voxels inside this particular section are separated from the borders of the model by voxels that do not meet the conditions above, which primarily means that they are encapsulated by seal voxels laterally and upwards.
 This condition is applicable  under the assumption, that connection to the borders of the model lead to leakage.
-A trap is thus defined as a closed system in this model.
+A trap is thus defined as a closed system in this model and trap closure is assumed to be void outside of the space of information, i.e.
+the model space.
+
 		\end{enumerate}
@@
 			\subsection{Finding the final spill point of a trap} %\label{sec:spill_leak_finding}
-			Regarding dome-shaped anticlinal structures, it can be observed that, geometrically and mathematically, a spill point is a saddle point of the reservoir top surface in 3D.
+			Regarding anticlinal structures and traps, it can be observed that, geometrically and mathematically, a spill point is a saddle point of the reservoir top surface in 3D.
 This was observed by \citet{collignon2015fold_linkage}, who pointed out that the linkage of folds is given by saddle points.
@@
 surface maxima or "hills") which is defined by the surface's saddle point between both.\\
-			In gemPy, layer boundary surfaces are returned in the form of discretized arrays of simplices and vertices.
-The latter can be transformed into a 2D rectangular array that contains Z-positional values for the surface in X and Y directions.
+			In GemPy, layer boundary surfaces are returned in the form of discretized arrays of simplices and vertices.
+The latter can be interpolated to a 2D rectangular grid that contains Z-positional values for the surface in X and Y directions.
 According to \citet{verschelde2017programmingtools}, a saddle point in a matrix is maximal in its row and minimal in its column.
 This corresponds to the logical geometrical deduction, that a saddle point for a surface defined by $f(x,y)$ is marked by a local maximum in one plane, but a local minimum in the perpendicular plane.
-This is observable in Figure \ref{fig:saddle_point_flow}.\\
+This is observable in Figure \ref{fig:saddle_point_flow} and made use of in the respective algorithm in this work.
 			\begin{figure}[h]
@@
 			\end{figure}
-			Regarding leakage across the normal fault in this geological model, a similar approach is taken.
+			%Regarding leakage across the normal fault in this geological model, a similar approach is taken.
 If the fault offset remains small enough, that the seal layer is not broken, then the reservoir top surface can be used to find the cross-fault leak point.
@@
 In this case, the maximum footwall contact of the surface with the fault is to be defined as cross-fault leak point.\\
-			Following these observations, the algorithm developed in this work uses differentiation techniques to find local extrema in the 2D reservoir top surface arrays attained over gemPy.
-Points where minima and maxima of perpendicular axes X and Y coincide are then defined as saddle points.
+			Regarding leakage across the normal fault in this geological model, a similar approach is taken.
+The reservoir top surface can be used to find the cross-fault leak point regarding juxtapositions with layers underlying the seal.
+In this case, the leak point can be defined as the maximal point of the hanging wall contact of the reservoir top surface with the fault.
+At the same time, this point will be a local minimum on the plane perpendicular to the normal fault and thus equal a saddle point.
+Higher magnitudes of fault displacement can lead to detachment of the seal in the hanging wall and consequently the juxtaposition of the trap section with seal-overlying formations which are assumed to be complete permeable.
+This case is dealt with independently, by checking this area of juxtaposition with regard to its $SSF$ and potential fault sealing, as is further elaborated in Section \ref{sec:check_juxta} below.\\
+			Following these observations, the algorithm developed in this work uses differentiation techniques to find local extrema in the 2D reservoir top surface arrays attained over GemPy.
+Points where local minima and maxima of perpendicular axes X and Y coincide are then defined as saddle points.
 The anticlinal spill point is deduced as the maximal saddle point in the footwall.
@@
 In the end, whichever point of these two is maximal, is defined to be the final spill point of the trap.
-If no cross-fault leak point is found, it is presumed that full leakage is allowed to take place due to little displacement of the reservoir layer.
-However, if fault sealing is given, any fault leakage and respective leak points become irrelevant and only the anticlinal spill point is considered.
+If no cross-fault leak point is found, it is presumed that full leakage is allowed to take place due to little or no displacement of the reservoir layer.
 The Z-value of the final spill point is used to determine the maximum down-to-horizon to which the trap feature can be filled.
+			%However, if fault sealing is given, any fault leakage and respective leak points become irrelevant and only the anticlinal spill point is considered.
 			%\begin{figure}[h]
@@
 			%\end{figure}\\	
-			\citet{kuijper2004detecting} has pointed out that using finding spatial critical points, saddle points especially, can be problematic using discretized methods and rectangular grids in particular.
+			\citet{kuijper2004detecting} has pointed out that finding spatial critical points, saddle points especially, can be problematic using discretized methods and rectangular grids in particular.
 Finding all extrema and saddle points can be difficult \citep{kuijper2004detecting}.
-Some failures to detect saddle points in this work's model are resolved by adding a buffer around extrema that define the anticlinal structure.
-IMPLEMENT BETTER METHOD?
+Failures to detect saddle points in this work's model are resolved by adding a buffer around extrema that define the anticlinal structure.
 However, it has to be pointed out that saddle points might still be missed in some rare cases, especially if they are located in an orientation which is "rotated" and not aligned with the X and Y axes.
-Furthermore, in this work, the geological model was designed in a way that saddle points are presumed to only occur within or near the structures of interest.
-FUTURE: MORE UNIVERSALLY APPLICAPLE METHODS
+Furthermore, in this work, the geological model was designed in a way that saddle points are presumed to only occur within or near the structures of interest and are thus easier to differentiate.
 
-			\subsection{Checking for juxtapositions and possible fault sealing}
+			\subsection{Checking for juxtapositions and possible fault sealing}\label{sec:check_juxta}
+			Given a juxtaposition of a trap section with layers overlying the seal across the fault, the respective area is to be examined with regard to its clay smearing potential.
+Here, this is achieved by calculating the $SSF$ as introduced in Section \ref{sec:Reservoir_values}.
+The area of this juxtaposition can be attained using a topology function that is part of GemPy.
+On the base of this, a new function is created, with which the Z-extent, i.e.
+the height of the juxtaposition can be deduced.
+Consequently, the fault throw can be attained by adding the seal thickness to this.
+The resulting displacement value can then be related to the seal thickness to obtain the Shale Smear Factor.\\
+			As mentioned before, it has been reported that the critical value $SSF_c$ is dependent on scale and can expected to be lower with larger shale thicknesses and greater fault throws.
+So, as the scale in the present geological model is relatively large, but also for the purpose of shedding light on different conceptual scenarios by allowing the threshold value to be exceeded, the critical Shale Smear Factor is set to a respectively low value of $SSF_c = 2.5$.
+Consequently, in model realizations where $SSF < 2.5$, the fault surface on which the seal slipped, is assumed to be sealing due to clay smearing.
+Otherwise, for $SSF \le 2.5$, the sealing is assumed to be breached.
+Due to the topological shape of the trap, full leakage is expected and the maximum trap volume is set to zero by not assigning any voxels to the trap structure.
 
 			\subsection{Calculating the maximum trap volume}
-			Have all voxels that fulfill the conditions defined above been determined, these voxels are seen as part of the maximum trap volume $V_{trap}$.
+			Have all trap voxels (fulfilling the four conditions defined above), been determined, these cells are seen as part of the maximum trap volume $V_{trap}$.
 This volume can then be calculated by simply counting the number of these voxels and rescaling their cumulative volume depending on the resolution in which the model was computed:
 			\begin{equation}
-			V_{trap} = n_v*(\frac{S_{orig}}{R_{m}})^3
+			V_{trap} = n_v*(\frac{S_{orig}}{R_{m}})^3,
 			\end{equation}
@@
 As mentioned before, $V_{trap}$ is assumed to equal the hydrocarbon-filled rock volume $A * h$ used in Equation \ref{eq:OOIP}.
-			For the example of a geological model with an original extent of 2000~m in three directions, computed using a resolution of 50, the scale factor is 40~m.
+			For the example of a cubic geological model with an original extent of 2000~m in three directions, computed using a resolution of 50, the scale factor is 40~m.
 Every voxels thus accounts for $40^3~m=64,000~m^3$ in volume.
@@
 		\subsection{Implementation of reservoir-related likelihoods in the 3D model}
-		As for the 1D case before, for the 3D case it is assumed that additional information is provided in the form of layers thickness likelihoods.
+		As for the 1D case before, for the 3D case it is also assumed that additional information is provided in the form of layer thickness likelihoods.
 In this geological scenario, the information might have been gained for example from additional seismic observations or from core sample data from a well nearby.
 Likelihoods could also be derived from knowledge of the geological history of the formation, such as the depositional environment of sedimentary layers.
-Assuming a possible correlation of nearby core samples with the layers in our model section, layer thickness likelihoods can be deduced respectively.\\
-		First of all, a method to determine layer thicknesses in our model realizations has to be introduced.
-Here, as single X- and Y-point (1D column in Z) in the footwall of the model is chosen as a measure for the thickness of the layers.
-In this column, point coordinates for the layer interfaces had been defined in the original data frame and can be observed as they are altered in the interpolation data frame for each iteration.
-The column is located centrally in the overall model block, as well as the original potential trap structure (X~=~1100~m,Y~=~1000~m) and is thus presumed to be representative and of significance for the reservoir analysis.
-Thickness variations are traced and vary relative to the changes in the layer interface positions, given by their uncertainties in Z-values.\\
+Assuming a possible correlation of nearby core samples with the layers in our model section, layer thickness likelihoods can be defined respectively as normal distributions.\\
+		First of all, a method to determine layer thicknesses in the geological model realizations has to be introduced.
+Here, as single X- and Y-point, i.e.
+a 1D column in Z in the footwall of the model is chosen as a measure for the thickness of the layers.
+In this column, point coordinates for the layer interfaces had been defined in the original data frame and can be observed as they are altered in the interpolation data frame with each probabilistic modeling iteration.
+Thicknesses are simply calculated as the distance between two layer boundary points in this column.
+Variations in thicknesses can be traced and vary relative to the changes in the layer interface positions, defined by the uncertainties in their Z-values.
+At X~=~1100~m and Y~=~1000~m, the point of measurement is located centrally in the overall model block, as well as the original potential trap structure and is thus presumed to be representative and of significance for the reservoir analysis.\\
 		After defining these thickness to point-positional dependencies, thickness likelihoods are implemented as probability distributions, as done for the 1D model.
 Various types of geological information, conclusions and expectations can be expressed by adapting mean and standard deviation of these normal distributions.
-For core sampling, for example, information might be evenly provided for all layers, irrespective of depth, so that similar standard distributions might be assigned to all layer thicknesses.
-Alternatively, taking into account certain depositional environments for which high thickness variablity is characteristic, specific standard deviations can be raised accordingly.
+For core sampling, for example, information might be evenly provided for all layers, irrespective of depth, so that similar degrees of standard deviation might be assigned to all layer thicknesses.
+Alternatively, taking into account certain depositional environments, for which high thickness variablity is characteristic, specific standard deviations can be raised accordingly.
 Some representative cases are presented in Chapter \ref{cha:results}.
 
-		\subsection{Bayesian analysis and valuation of the 3D model}
+		\subsection{Bayesian analysis and evaluation for the 3D geological model}
 		According to the methods introduced in Sections \ref{sec:bayes} and \ref{sec:bayes_in_struc_modeling}, the 3D geological model is to be incorporated in a probabilistic modeling framework and analyzed from a Bayesian perspective.
-
-		First, a simple Monte Carlo error propagation run, using only priors, is conducted so that reference distributions
+		First, a simple Monte Carlo error propagation run, using only priors, is conducted so that a base reference distribution for the original model is generated.
+Subsequently, Bayesian inference is conducted based on different assumptions concerning the additional observations made and by incorporating likelihoods which respectively are designed to reflect the nature of this auxiliary data.
+Posterior distributions, in particular resulting volume distributions, can then be examined and evaluated by applying the methods described in Section \ref{sec:model_evaluation}.

### 2017-11-08 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Corrected the juxtaposition check to attain SSF!

@@
 Deviations from the unknown true value in the form of over- and underestimation bring about an error and loss accordingly.
-In this case, it is assumed that investments increase linearly with linear growth in the value of the resource.
+In this case, it is assumed that investments increase linearly with linear growth in the value of the reserve.
 For this reason, the absolute-error loss function is favored here over the squared-error loss function.
-It is chosen as the base function to which further development steps refer, based on mostly logical case-specific assumptions.
+It is chosen as the base function to which further development steps refer, based on logical case-specific assumptions.
 The steps are illustrated in Figure \ref{fig:LF_4steps_normal} and listed in the following:
@@
 Additional equipment can often be installed later on.
-Hence, overestimation is weighted stronger in this loss function by multiplying the error with an overestimation factor \textit{a} (= 1.25):
+Hence, overestimation is weighted stronger in this loss function by multiplying the error with an overestimation factor $a = 1.25$:
 

### 2017-11-08 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
NOW juxta check seems to be finally working! Does it?

@@
 			\begin{equation}
-			L(\theta,\hat{\theta}) = |\theta - \hat{\theta}|
+			L(\theta,\hat{\theta}) = |\theta - \hat{\theta}|.
 			\end{equation}
@@
 Additional equipment can often be installed later on.
-Hence, overestimation is weighted stronger in this loss function by multiplying the error with an overestimation factor $a = 1.25$:
+Hence, overestimation is weighted stronger in this loss function by multiplying the error with an overestimation factor $a$:
 
 			\begin{equation}\label{eq:LF_II}
-			L(\theta,\hat{\theta}) = |(\theta-\hat{\theta})|*a
+			L(\theta,\hat{\theta}) = |(\theta-\hat{\theta})|*a.
 			\end{equation}
@@
 This is worse than the "normal" non-fatal overestimation, where both values are positive and a net gain is still achieved, which is only smaller then the best possible gain of expecting the true value.
-Fatal overestimation is included in the loss function by using another weighting factor \textit{b} (= 2) that replaces \textit{a}:
+Fatal overestimation is included in the loss function by using another weighting factor $b$ that replaces \textit{a}:
 
 			\begin{equation}\label{eq:LF_III}
-			L(\theta,\hat{\theta}) = |(\theta-\hat{\theta})|*b
+			L(\theta,\hat{\theta}) = |(\theta-\hat{\theta})|*b.
 			\end{equation}
 
-			In other words: With \textit{b} (= 2), fatal overestimation is twice as bad as simple underestimation.
+			In other words: With $b = 2$, fatal overestimation is twice as bad as simple underestimation.
 
@@
 opportunity costs that arise from completely discarding a reservoir with a potential gain equal to the positive true value.
-Fatal underestimation is weighted using a third factor \textit{c}:
+Fatal underestimation is weighted using a third factor:
 
 			\begin{equation}\label{eq:LF_IV}
-			L(\theta,\hat{\theta}) = |(\theta-\hat{\theta})|*c
+			L(\theta,\hat{\theta}) = |(\theta-\hat{\theta})|*c.
 			\end{equation}			
@@
 		\end{cases},
-		\text{ with } a,b,c \in \mathbb{Q}
+		\text{ with } a,b,c \in \mathbb{Q}.
 		\end{equation}  
@@
 Fundamentally different loss functions can also be based on a significantly different mathematical structure.
-Loss functions are customized regarding the problem environment and according to the to the subjective needs and objectives of an actor.
+Loss functions are customized regarding the problem environment and according to the to subjective needs and objectives of decision-maker.
 Thus, they are mostly defined by the actor expressing his perspective.
-Changes in the individual's perception and viewpoint might lead to further customization needs even later on.
-(Especially considering individual persons as actors, psychological aspects may play a significant role.)
+Changes in the individual's perception and attitude might lead to further customization needs at a future point in time.
+Especially considering individual persons as actors, psychological aspects may play a significant role.
 
@@
 
-		One can assume that several actors in one sector or decision environment may have the same general loss function, but different affinities concerning risks.
-This might be based for example on the different psychological factors or economic philosophies followed by companies.
+		It can be assumed that several actors in one sector or decision environment may have the same general loss function but different affinities concerning risks.
+This might be based for example on different psychological factors or economic philosophies followed by companies.
 It might also be based on budgets and options such actors have available.
@@
 		\end{cases},
-		\text{ with } a,b,c,r \in \mathbb{Q}
+		\text{ with } a,b,c,r \in \mathbb{Q}.
 		\end{equation}	
@@
 			\includegraphics[width=1\textwidth]{Figures/LFR_normal.png}
-			\caption{Plotting of expected loss realizations after including the risk factor $r$ in the loss function (Equation \ref{eq:LFR_final}) for actors with risk-affinities ranging from risk-averse ($r$ = 0.5 and 0.75), over risk-neutral ($r$ = 1), to risk-friendly ($r$ = 1.25 and $r$ = 1.5).}\label{fig:LFR_normal}
+			\caption{Plotting of expected loss realizations after including the risk factor $r$ in the loss function (Equation \ref{eq:LFR_final}) for actors with risk-affinities ranging from risk-averse ($r$ = 0.5 and 0.75), over risk-neutral ($r$ = 1), to risk-friendly ($r$ = 1.25 and $r$ = 1.5).
+The loss function was applied on a normal distribution around zero ($\mu=0; \sigma=500$).}\label{fig:LFR_normal}
 		\end{figure}
@@
 		\section{1D geological reservoir modeling}\label{sec:1D_model}
-		For simple understanding and a preliminary assessment of the Bayesian statistical methods and the loss function described above, they are first to be applied in the context of an conceptual one-dimensional reservoir case.
-The underlying model and basic approach are inherited from \citet{delaVarga2016}.
-Parameters are adapted to more appropriately represent a reasonable geological petroleum system, consisting of a reservoir formation with overlying seal in the subsurface.
+		For simple understanding and a preliminary assessment of the Bayesian statistical methods and the loss function described above, they are first to be applied in the context of a conceptual one-dimensional hydrocarbon system case.
+The underlying model and basic approach are inherited from \citet{delaVarga2016}, but the parameters are adapted to more appropriately represent a reasonable geological petroleum system, consisting of a reservoir formation with overlying seal in the subsurface.
 In this 1D setting, only the interface depths and thicknesses of layers such as the reservoir or seal unit can be observed.
@@
 				\includegraphics[width=1\textwidth]{Figures/1D_model.png}
-				\caption{Probability distributions for positions of layer boundaries in the subsurface and a respective representation using lines.}\label{fig:1D_model}
+				\caption{Probability distributions for positions of layer boundaries in the subsurface and a respective representation using lines.
+These normal distributions are determinedby the following values: Seal top: $\mu = 2000 m, \sigma = 7$; reservoir top: $\mu = 2050 m, \sigma = 25$; reservoir bottom: $\mu = 2200 m, \sigma = 45$.
+}\label{fig:1D_model}
 			\end{figure}
@@
 These points confine two layers in the middle, from which the upper one can be labeled as seal and the lower one as reservoir.
-The standard deviations ($sigma$) of the distributions increase with depth, representing an increase in uncertainty.
+The standard deviations ($\sigma$) of the distributions increase with depth, representing an increase in uncertainty.
 For an approximate representation of a hydrocarbon reservoir system, the distribution means ($\mu$) are set to depths of 2000 m (seal top, $\sigma=7$), 2050 m (reservoir top, $\sigma=25$) and 2200 m (reservoir bottom, $\sigma=45$).
@@
 			\subsection{Bayesian inference and analysis of the 1D case}\label{sec:1D_bayes}
-			For conducting Bayesian inference, it is assumed that new observations have been made, providing additional information on the likelihoods of the thicknesses of the two layers of interest.
+			For conducting Bayesian inference, it is assumed that new observations are made, providing additional information on the likelihoods of the thicknesses of the two layers of interest.
 Likelihood functions for reservoir and seal thicknesses are introduced based on another set of normal probability distributions.
@@
 Means and standard deviations of the likelihood distributions are variable, so that several scenarios based on the assumption of different observations can be tested.
-Parameter priors and likelihood functions are embedded into a probabilistic modeling framework as described in Section \ref{sec:numerical_implementation} and MCMC sampling is executed accordingly.
-Posterior distributions are then to be evaluated, based on the valuation method described in the following and subsequent application of the custom loss function.
+Parameter priors and likelihood functions are embedded into a probabilistic modeling framework as described in Section \ref{sec:numerical_implementation} and MCMC sampling is conducted accordingly.
+Posterior distributions are then to be evaluated based on the valuation method described in the following.
 
@@
 	        	\includegraphics[width=1\textwidth]{Figures/3_parameters.png}
-	        	\caption{Probability distributions for the three parameters deduced from the 1D model depicted in \ref{fig:1D_model}.
-Values from these parameter distributions are combined in \ref{eq:1D_score_system} attain a score to valuate reservoir model realizations.}\label{fig:3_parameters}
+	        	\caption{Probability distributions for the three parameters deduced from the 1D model depicted in Figure \ref{fig:1D_model}.
+Values from these parameter distributions are combined in Equation \ref{eq:1D_score_system} to attain a score for valuation of reservoir model realizations.}\label{fig:3_parameters}
 	        \end{figure}
@@
 Instead, a threshold for seal reliability is defined beforehand.
-Here it is set at 20 m thickness.
+Here, it is set to 20 m thickness.
 If the seal thickness falls below this threshold, it is assumed that the seal fails completely and thus all the potential value (positive score) of the reservoir is lost, while costs of depth (negative score) remain.
@@
 			where reservoir thickness is given by $h_{res}$, seal thickness by $h_{seal}$ and reservoir depth, which is always a negative value, by $d_{res}$.\\
-			This valuation method is applied on the original prior-only 1D model as reference and then on each posterior distribution set after conducting Bayesian inference.
+			This valuation method is applied on the original prior-only Monte Carlo sampling of the 1D model as reference and then on each posterior distribution set after conducting Bayesian inference.
 Resulting score probability distributions can be compared directly and additionally under use of the case-specific custom loss function (Equation \ref{eq:LFR_final}) defined before.

### 2017-11-10 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Uncertainty parameters seem good!
To Do: Implement return of trap control mechanism:
Leak - Spill - Juxta... maybe SFF/Throw Likelihood?

@@
 In a Python environment, the data is imported as comma-separated values (CSV) using GemPy.
-From the generated data frame, GemPy is able to interpolate surfaces following the potential-field method explained in Section \ref{sec:struc_geo_modeling} and compute voxel-based 3D geological models (see Figure \ref{fig:3Dmodel_section}).
-
+From the generated data frame, GemPy is able to interpolate surfaces following the potential-field method explained in Section \ref{sec:struc_geo_modeling} and compute voxel-based 3D geological models (see Figure \ref{fig:3Dmodel_construction}).
 		\begin{figure}[h]
 			\centering
-			\includegraphics[width=1\textwidth]{Figures/3Dmodel_section.png}
-			\caption{Mid section of the base 3D geological model as defined by the original data frame (see Appendix).
-Layer interfaces are defined by points and foliation dipping indicated by perpendicularly oriented arrows.}\label{fig:3Dmodel_section}
+			\includegraphics[width=1\textwidth]{Figures/3Dmodel_construction.png}
+			\caption{FIGURE TO BE UPDATED!
+(A): 2D section through the moddle of the model and perpendicular to the fault.
+(B) 3D visualization of model (rather Blender or Paraview?).
+(C): Implementation of uncertainties for the z-positions of layer interfaces.
+Boundaries indicated by the same color, share a common respective uncertainty in form of a normal distribution.
+(D): The offset uncertainty is expressed by a skew normal distribution that only affects layer interface points in the hanging wall.}\label{fig:3Dmodel_construction}
 		\end{figure}
-
 		This model allows for the consideration of various kinds of uncertainties and their implementation in several different ways.
@@
 		\begin{enumerate}
-			\item \textbf{Labeled as reservoir formation:} The voxel has been assigned to the target reservoir formation (here: Sandstone 1 in Figure \ref{fig:3Dmodel_section}) in the lithology block model.
+			\item \textbf{Labeled as reservoir formation:} The voxel has been assigned to the target reservoir formation (here: Sandstone 1 in Figure \ref{fig:3Dmodel_construction}) in the lithology block model.
 This is determined by respective labeling of the input data and the computation conducted by GemPy.

### 2017-11-10 Risk analysis/Text/Layout/Layout/chapter1.tex
Wrote introduction. Needs revision and a bit more

@@
-\chapter{Introduction} \label{chap::intro}
-
-    - mention developments in geological modeling
-
-    - economic significance of geological modeling: examples of application in petroleum, mining, aquifers, CCS, etc.
-
-    - Bayesian analysis developments
-
-    - decision-making in for such economic scenarios and projects which involve geosciences
-
-    - classical approach (decision trees) vs. Bayesian approach, expression of decision-maker's interests in continuous loss functions
-
-    - influence of uncertainty and work done by Wellmann & DeLaVarga
+\chapter{Introduction} \label{chap:intro}
+Bayesian methods are an intuitive approach to inference, naturally inherent in human thinking patterns and closely tied to processes of decision-making \citep{berger2013stat, davidson2015, jaynes1986bayesian}.
+Individuals are constantly faced with situations in which a decision has to be made, but only incomplete information is available.
+Such a problem necessitates an approach based on plausible reasoning, one which is intuitively structured in five stages \cite{jaynes1986bayesian}:
+\begin{enumerate}
+	\item Identify uncertainties and attempt to consider all possibilities that might arise.
+	\item Based on all the information and past experience available, evaluate how likely every possibility is.
+	\item Assess the probable consequences of single possible actions.
+	\item Based on the foregone steps, make a decision \citep{jaynes1986bayesian}.
+\end{enumerate}
+This concept is relatable to a vast variety of problems, ranging from casual every-day situations to complex scenarios in large-scale economic decision-making: As a private person, should I take an umbrella with me today?
+As a company, should we invest in the development and realization of a certain project?
+Following this process of plausible reasoning, the quality of a decision is to be measured based on the preceding state of knowledge and reasonable expectations, not on the subsequent actual consequences \citep{jaynes1986bayesian}.
+In other words: A decision is optimal, as long as it is the best action given the information available to the decision-maker before making the decision, no matter if actual loss was incurred afterwards.\\
+Bayesian decision theory and the related concepts of expected loss and loss functions have found increasingly common application in several economic sectors and fields of research, such as medicine \citep{ashby2000evidence, ashby2006bayesian, moye2006statistical} and machine learning \citep{barber2012bayesian, theodoridis2015machine}.
+Probabilistic approaches to decision-making have also become prevalent in the sector of hydrocarbon exploration and production (SOURCE).
+However, the methods here a mainly limited on (p10-p90, decision trees).......(see BRATVOLD)???\\
+In geosciences, Bayesian inference has prominently found use in the context of geophysical inversion problems (see \citet{tarantola1982inverse, mosegaard2002probabilistic} and \citet{sambridge2002monte}).
+Recently, it was transferred by \citet{delaVarga2016} to the field of structural geological modeling.
+This has been enabled by progressing developments regarding implicit geological modeling functions based on interpolation \citep{hillier2014three, mallet1992discrete, lajaunie1997foliation} and the possibility of fully automated model reconstruction in particular.
+\citet{delaVarga2016} regarded geological modeling as a Bayesian inference problem by relating additional geological information to prior model parameters in the form of likelihood functions, linking them in a non-parametric Bayesian network.
+Using Markov chain Monte Carlo sampling to explore resulting probability spaces, they attained posterior model suites with reduced uncertainties \citep{delaVarga2016}\\
+This work builds upon their concept, exploring the potential significance their findings might have in the context of decision-making.
+Bayesian decision-theory is to be included in the step of model evaluation.
+This is achieved by assigning an economic meaning to the structural model and designing a case-specific custom loss functions to find decisions which are optimal related to the state of knowledge and the preferences of actor's with different risk-affinities.
+More specifically, the models are designed to represent potential petroleum systems.
+Consequently, the development of algorithms for automatic hydrocarbon trap recognition and volume calculation represent a central part of this work.\\
+The main hypothesis of this work is that Bayesian inference and resulting changes in uncertainties in a geological setting have a significant effect on related value estimation and decision-making.
+It is furthermore postulated that loss functions can be customized to appropriately represent preferences of actors in the hydrocarbon sector and moreover illustrate the nature of decisions such actor's might make depending on their individual attitudes towards risk and in the face of different types of uncertainties.
+Changes in their respective Bayes actions are treated as a suitable measure to assess the effect of updating prior model parameters with new geological information.

### 2017-11-10 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Wrote introduction. Needs revision and a bit more

@@
 This means that all differences between relative errors are weighed equally, no matter whether they are found in the realm of relatively small or relatively large errors \citep{hennig2007}.\\
-        Using the squared-error loss function returns losses that increase quadratically with distance of the estimator to the true parameter value \citep{davidson2015}:        
+        Using the squared-error loss function returns losses that increase quadratically with distance of the estimator to the true parameter value \citep{davidson2015, moye2006statistical}:        
         \begin{equation}\label{eq:SqrLossFunction}
@@
 
-        Minimization of a loss function returns a Bayesian point estimate known as Bayes action $\delta^p(y)$, which is the estimate, action or decision with the least expected loss according to the loss function \citep{berger2013stat}.
+        Minimization of a loss function returns a Bayesian point estimate known as Bayes action or Bayesian estimator $\delta^p(y)$, which is the estimate, action or decision with the least expected loss according to the loss function \citep{berger2013stat, moye2006statistical}.
 For a unimodal and symmetric absolute-error loss function, the Bayes action is simply the median of the posterior distribution, while using squared-error loss it is the mean \citep{davidson2015, berger2013stat}.
@@
 For the application in a probabilistic setting, the method should enable fully automatic reconstruction of the model, when parameters are changed.
-The application in this work follows the example of \citet{delaVarga2016} and relies on the use of implicit interpolation for geological modeling, a method developed and elaborated by \citet{Lajaunie1997} and \citet{calcagno2008geological}.\\
+The application in this work follows the example of \citet{delaVarga2016} and relies on the use of implicit interpolation for geological modeling, a method developed and elaborated by \citet{lajaunie1997foliation} and \citet{calcagno2008geological}.\\
         This implicit method relies on the interpolation of a potential field scalar function $T(\vec{x})$ of any point $\vec{x}$ in a 3D space, and thus reflects the geometry of geological structures \citep{calcagno2008geological}.

### 2017-11-12 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Saved some runs, 1000iter. New Figures for methods

@@
 			\centering
-			\includegraphics[width=1\textwidth]{Figures/3Dmodel_construction.png}
-			\caption{FIGURE TO BE UPDATED!
-(A): 2D section through the moddle of the model and perpendicular to the fault.
-(B) 3D visualization of model (rather Blender or Paraview?).
-(C): Implementation of uncertainties for the z-positions of layer interfaces.
-Boundaries indicated by the same color, share a common respective uncertainty in form of a normal distribution.
-(D): The offset uncertainty is expressed by a skew normal distribution that only affects layer interface points in the hanging wall.}\label{fig:3Dmodel_construction}
+			\includegraphics[width=1\textwidth]{Figures/Uncertainties_Likelihoods.png}
+			\caption{...}\label{fig:3Dmodel_construction}
 		\end{figure}
@@
 		In the course of this work, several algorithms are developed within the Python environment, that in combination enable the automatic recognition and calculation of trap volumes in geological models computed by GemPy.
-To assign voxels of the model to the trap volume, it is checked whether the following conditions are satisfied by each particular voxel:
+To assign voxels of the model to the trap volume, it is checked whether the following conditions (illustrated in Figure \ref{fig:ig:trap_recognition})are satisfied by each particular voxel:
+		\begin{figure}[h]
+			\centering
+			\includegraphics[width=1\textwidth]{Figures/Trap_Cond_H.png}
+			\caption{...}\label{fig:trap_recognition}
+		\end{figure}
 		\begin{enumerate}
-			\item \textbf{Labeled as reservoir formation:} The voxel has been assigned to the target reservoir formation (here: Sandstone 1 in Figure \ref{fig:3Dmodel_construction}) in the lithology block model.
+			\item \textbf{Labeled as reservoir formation:} The voxel has been assigned to the target reservoir formation (see Sandstone 1 in Figure \ref{fig:trap_recognitionn} (1)) in the lithology block model.
 This is determined by respective labeling of the input data and the computation conducted by GemPy.
@@
 the model space.
-
 		\end{enumerate}

### 2017-11-13 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
More models running. Updated text and figures

@@
 A critical review of the research on clay smearing published over the last four decades was recently presented by \citet{vrolijk2016clay}.
-
         The 3D structural model in this work (see Section \ref{sec:3D_design}) is designed in a way that a potential conventional hydrocarbon trap is included in the form of anticlinal folding combined with normal faulting of a reservoir-seal sequence.
@@
 			\includegraphics[width=1\textwidth]{Figures/Uncertainties_Likelihoods.png}
-			\caption{...}\label{fig:3Dmodel_construction}
+			\caption{Illustration of the design and construction of the 3D structural geological model.
+A 2D cross section through the middle of the model, perpendicular to the normal fault, is shown in (A).
+A 3D representation of the model, highlighting the reservoir and seal formations as voxels, is visualized in (B).
+In (C) and (D), the inclusion of parameter uncertainties is presented.
+Colors indicate certain layer bottoms (i.e.
+boundaries) which are assigned a shared normal distribution on which their z-positional uncertainties are based (C).
+All points in the hanging wall are additionally assigned a fault offset uncertainty based on a skew normal distribution (D).
+Thicknesses of the three middle layers are defined by the distances of boundary points (E) and are thus directly dependent on (C).
+These thicknesses are to be defined as likelihoods.}\label{fig:3Dmodel_construction}
 		\end{figure}
@@
 Furthermore, uncertainty regarding the magnitude of fault offset is incorporated by adding a skew normal probability distribution that is shared by all layer interface points on the hanging wall.
-A left skewed ($\alpha=-4$) normal distribution is chosen to reflect the nature of throw on a normal fault, particularly the slip motion of the hanging wall block.
+A left skewed ($\alpha=-2$) normal distribution is chosen to reflect the nature of throw on a normal fault, particularly the slip motion of the hanging wall block.
 Mainly negative values are returned by this distribution.
@@
 		In the course of this work, several algorithms are developed within the Python environment, that in combination enable the automatic recognition and calculation of trap volumes in geological models computed by GemPy.
-To assign voxels of the model to the trap volume, it is checked whether the following conditions (illustrated in Figure \ref{fig:ig:trap_recognition})are satisfied by each particular voxel:
-		\begin{figure}[h]
+To assign voxels of the model to the trap volume, it is checked whether the following conditions (illustrated in Figure \ref{fig:trap_recognition}) are satisfied by each particular voxel:
+		\begin{figure}[p!]
 			\centering
 			\includegraphics[width=1\textwidth]{Figures/Trap_Cond_H.png}
-			\caption{...}\label{fig:trap_recognition}
+			\caption{Illustration of the process of trap recognition, i.e.
+the conditions that have to be met by a model voxel, to be accepted as belonging to a valid trap.
+A voxel has to be labeled as part of the target reservoir formation (1) and positioned in the footwall (2).
+Trap closure is defined by the seal shape and the normal fault (3).
+Consequently, the maximum trap fill is defined by either the anticlinal spill point (3;a) or a point of leakage across the fault, depending on juxtapositions with layers underlying (3;b) or overlying the seal (3;c).
+The latter is only relevant if the critical Shale Smear Factor is exceeded, as determined over $D$ and $T$ in (4).
+In this example, assuming sealing of the fault due to clay smearing, the fill horizon is determined by the spill point in (4).
+Subsequently, only trap section 1 is isolated from the model borders in (4) and can thus be considered a closed trap.}\label{fig:trap_recognition}
 		\end{figure}
 		\begin{enumerate}
-			\item \textbf{Labeled as reservoir formation:} The voxel has been assigned to the target reservoir formation (see Sandstone 1 in Figure \ref{fig:trap_recognitionn} (1)) in the lithology block model.
+			\item \textbf{Labeled as reservoir formation:} The voxel has been assigned to the target reservoir formation (see Sandstone 1 in Figure \ref{fig:trap_recognition} (1)) in the lithology block model.
 This is determined by respective labeling of the input data and the computation conducted by GemPy.
@@
 			As mentioned before, it has been reported that the critical value $SSF_c$ is dependent on scale and can expected to be lower with larger shale thicknesses and greater fault throws.
-So, as the scale in the present geological model is relatively large, but also for the purpose of shedding light on different conceptual scenarios by allowing the threshold value to be exceeded, the critical Shale Smear Factor is set to a respectively low value of $SSF_c = 2.5$.
+So, as the scale in the present geological model is relatively large, but also for the purpose of shedding light on different conceptual scenarios by allowing the threshold value to be exceeded, the critical Shale Smear Factor is set to a respectively low value of $SSF_c = 3$.
 Consequently, in model realizations where $SSF < 2.5$, the fault surface on which the seal slipped, is assumed to be sealing due to clay smearing.

### 2017-11-13 Risk analysis/Text/Layout/Layout/chapter1.tex
Began with discussion text. Corrected loss function: r for underest.

@@
 Individuals are constantly faced with situations in which a decision has to be made, but only incomplete information is available.
-Such a problem necessitates an approach based on plausible reasoning, one which is intuitively structured in five stages \cite{jaynes1986bayesian}:
+Such a problem necessitates an approach based on plausible reasoning, one which is intuitively structured in five stages \citep{jaynes1986bayesian}:
 \begin{enumerate}
@@
 It is furthermore postulated that loss functions can be customized to appropriately represent preferences of actors in the hydrocarbon sector and moreover illustrate the nature of decisions such actor's might make depending on their individual attitudes towards risk and in the face of different types of uncertainties.
-Changes in their respective Bayes actions are treated as a suitable measure to assess the effect of updating prior model parameters with new geological information.
+Changes in their respective decisions are treated as a suitable measure to assess the effect of updating model parameters with new geological information.

### 2017-11-13 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Began with discussion text. Corrected loss function: r for underest.

@@
 Remaining factors can be implemented deterministically or as uncertain values as well.
-WHAT DO?\\%For most factors, it seems most suitable to define a normal distribution around a known typical value or global average, as it will return a high probability for common scenarios and occasionally produce cases of rare, more extreme values.\\
+WHICH ONE TO USE?\\%For most factors, it seems most suitable to define a normal distribution around a known typical value or global average, as it will return a high probability for common scenarios and occasionally produce cases of rare, more extreme values.\\
         Consequently, $OOIP$ and $ROV$ calculations represent the step following and using the results from computing the geological model, so that the realizations are assigned values that can be interpreted from a perspective of economic interest.
@@
 they will rather develop a project than discard it.
-In the loss function, risk is simply included using a risk factor \textit{r} which alters the weighting factors \textit{a, b} and \textit{c} respectively:
+In the loss function, risk is simply included using a risk factor \textit{r} which alters the incurred losses respectively:
 		\begin{equation}\label{eq:LFR_final}
@@
 		\begin{cases}
-		|\theta - \hat{\theta}|, & \text{for } 0<\hat{\theta}<\theta  \\
-		|\theta-\hat{\theta}|*(a*r), & \text{for } 0<\theta<\hat{\theta} \\
-		|\theta-\hat{\theta}|*(b*r), & \text{for } \theta\leq0<\hat{\theta} \\
-		|\theta-\hat{\theta}|*(c*(r^{-0.5})), & \text{for } \hat{\theta}\leq0<\theta 
+		|\theta - \hat{\theta}|*r^{-0.5}, & \text{for } 0<\hat{\theta}<\theta  \\
+		|\theta-\hat{\theta}|*a*r, & \text{for } 0<\theta<\hat{\theta} \\
+		|\theta-\hat{\theta}|*b*r, & \text{for } \theta\leq0<\hat{\theta} \\
+		|\theta-\hat{\theta}|*c*r^{-0.5}, & \text{for } \hat{\theta}\leq0<\theta 
 		\end{cases},
@@
 		\end{figure}
-		According to this, for \textit{r} = 1 the risk-neutral loss function is returned, since \textit{a, b} and \textit{c} are not changed.
-For \textit{r} $<$ 1, the weight on overestimating (\textit{a, b}) is reduced and increased for fatal underestimation (\textit{c}).
+		According to this, for \textit{r} = 1 the risk-neutral loss function is returned, as no re-weighting takes place.
+For \textit{r} $<$ 1, the weight on overestimating (\textit{a, b}) is reduced and increased for fatal underestimation (\textit{c}), as well as normal underestimation.
 This represents a risk-friendlier actor that is willing to bid on a higher estimate to attain a greater gain.
-For \textit{r} $>$ 1, the overestimation weight (\textit{a, b}) is increased in the loss function, the fatal underestimation weight (\textit{c}) is decreased and respectively more risk-averse actors are prompted to bid on lower estimates.\\
+For \textit{r} $>$ 1, the overestimation weight (\textit{a, b}) is increased in the loss function, underestimation and fatal underestimation weight (\textit{c}) are decreased and respectively more risk-averse actors are prompted to bid on lower estimates.\\
 		The factor \textit{r} can take basically any positive values.
@@
 			\includegraphics[width=1\textwidth]{Figures/1D_principle.PNG}
-			\caption{ (from \citet{delaVarga2016}).}\label{fig:1D_principle}
+			\caption{Conceptual illustration of applying Bayesian inference in an abstract 1D geological model as conducted by \citet{delaVarga2016}.
+Prior distributions of layer boundaries in depth (a) are updated using thickness likelihoods (b) to attain posterior distributions (c).
+Model parameters which are uncorrelated in their joint probability distribution before the inference (d), are correlated after performing Bayesian inference (e) (from \citet{delaVarga2016}).}\label{fig:1D_principle}
 			\end{figure}
@@
 Parameter priors and likelihood functions are embedded into a probabilistic modeling framework as described in Section \ref{sec:numerical_implementation} and MCMC sampling is conducted accordingly.
-Posterior distributions are then to be evaluated based on the valuation method described in the following.
+Performing Bayesian inference on such a 1D geological model follows the conceptual example devised by \citet{delaVarga2016} (see Figure \ref{fig:1D_principle}).
+Posterior distributions are to be evaluated based on the valuation method described in the following.
 
@@
 
-			\subsection{Finding the final spill point of a trap} %\label{sec:spill_leak_finding}
+			\subsection{Finding the decisive spill or leak point of a trap} %\label{sec:spill_leak_finding}
 			Regarding anticlinal structures and traps, it can be observed that, geometrically and mathematically, a spill point is a saddle point of the reservoir top surface in 3D.
-This was observed by \citet{collignon2015fold_linkage}, who pointed out that the linkage of folds is given by saddle points.
+This was described by \citet{collignon2015fold_linkage}, who pointed out that the linkage of folds is given by saddle points.
 These are thus a controling factor for spill-related migration from respective structural traps.
@@
 that the extremum is a stationary point \citep{guichard2013calculus, weisstein2017saddlepoint}.
-In the context of a geological reservoir system, such a hill can be considered to represent an anticlinal structural trap.
+In the context of a geological reservoir system, such a hill can be regarded as a representation of an anticlinal structural trap.
 Local minima are defined analogously, presenting local minima in both planes at a stationary point \citep{guichard2013calculus}.
@@
 			\begin{equation}\label{eq:discriminant_D}
-			D(x_0,y_0) = f_{xx}(x_0,y_0)f_{yy}(x_0,y_0) - f_{xy}(x_0,y_0)^2
+			D(x_0,y_0) = f_{xx}(x_0,y_0)f_{yy}(x_0,y_0) - f_{xy}(x_0,y_0)^2.
 			\end{equation}
@@
 				\includegraphics[width=1\textwidth]{Figures/Saddle_Point_between_maxima.png}
-				\caption{Saddle point (red dot) between two maxima. In a geological setting, these surface maxima may be interpreted as two dome-shaped four-way closure traps.}\label{fig:saddle_point_maxima}
+				\caption{Saddle point (red dot) between two maxima.
+In a geological setting, these surface maxima may be interpreted as two dome-shaped four-way closure traps.
+Spill-related migration between the two structures would take place over the saddle point, i.e.
+the spill point.}\label{fig:saddle_point_maxima}
 			\end{figure}
@@
 At the same time, this point will be a local minimum on the plane perpendicular to the normal fault and thus equal a saddle point.
-Higher magnitudes of fault displacement can lead to detachment of the seal in the hanging wall and consequently the juxtaposition of the trap section with seal-overlying formations which are assumed to be complete permeable.
+Higher magnitudes of fault displacement can lead to detachment of the seal in the hanging wall and consequently the juxtaposition of the trap section with seal-overlying formations which are assumed to be completely permeable.
 This case is dealt with independently, by checking this area of juxtaposition with regard to its $SSF$ and potential fault sealing, as is further elaborated in Section \ref{sec:check_juxta} below.\\
@@
 The cross-fault leak point is determined analogously in the hanging wall.
-In the end, whichever point of these two is maximal, is defined to be the final spill point of the trap.
+In the end, whichever point of these two is maximal, is defined to be the decisive spill point of the trap.
 If no cross-fault leak point is found, it is presumed that full leakage is allowed to take place due to little or no displacement of the reservoir layer.
-The Z-value of the final spill point is used to determine the maximum down-to-horizon to which the trap feature can be filled.
+Given only minor fault offset, no local maximum is found in the plane perpendicular to the fault.
+The Z-value of the decisive spill point is used to determine the maximum down-to-horizon to which the trap feature can be filled.
 			%However, if fault sealing is given, any fault leakage and respective leak points become irrelevant and only the anticlinal spill point is considered.
@@
 However, it has to be pointed out that saddle points might still be missed in some rare cases, especially if they are located in an orientation which is "rotated" and not aligned with the X and Y axes.
-Furthermore, in this work, the geological model was designed in a way that saddle points are presumed to only occur within or near the structures of interest and are thus easier to differentiate.
+Furthermore, in this work, the geological model is designed in a way that saddle points are presumed to only occur within or near the structures of interest and are thus easier to differentiate.
 
@@
 Here, this is achieved by calculating the $SSF$ as introduced in Section \ref{sec:Reservoir_values}.
-The area of this juxtaposition can be attained using a topology function that is part of GemPy.
+Juxtaposition areas can be attained using a topology function that is part of GemPy.
 On the base of this, a new function is created, with which the Z-extent, i.e.
-the height of the juxtaposition can be deduced.
+the height of relevant juxtapositions can be deduced.
+These are juxtapositions of all layers on which the seal slipped and seal-overlying formations.
+In the case of high displacement, for example, the footwall reservoir, but also the basement might be in contact with units above the seal in the hanging wall.
+The  heights of each of these single juxtapositions, below the point of maximum displacement at the trap, are added up to attain a total juxtaposition height.
 Consequently, the fault throw can be attained by adding the seal thickness to this.
@@
 So, as the scale in the present geological model is relatively large, but also for the purpose of shedding light on different conceptual scenarios by allowing the threshold value to be exceeded, the critical Shale Smear Factor is set to a respectively low value of $SSF_c = 3$.
-Consequently, in model realizations where $SSF < 2.5$, the fault surface on which the seal slipped, is assumed to be sealing due to clay smearing.
-Otherwise, for $SSF \le 2.5$, the sealing is assumed to be breached.
+Consequently, in model realizations where $SSF < 3$, the fault surface on which the seal slipped, is assumed to be sealing due to clay smearing.
+Otherwise, for $SSF \le 3$, the sealing is assumed to be breached.
 Due to the topological shape of the trap, full leakage is expected and the maximum trap volume is set to zero by not assigning any voxels to the trap structure.

### 2017-11-13 Risk analysis/Text/Layout/Layout/chapter4_discussion.tex
Reworked 1D results text.
Began with discussion text. Corrected loss function: r for underest.

@@
 	\chapter{Discussion}\label{cha:discussion}
+	Comparing the 1D and the 3D geological model, there is a large difference in complexity behind the model construction, as well as the way in which the decisive posterior value distributions (scores and recoverable volumes, respectively) are attained.
+Nevertheless, results of evaluating these models through application of the custom loss function are widely similar in principle.
+The reference distributions from the prior-only Monte Carlo uncertainty propagation sampling are comparable between the two cases.
+They are generally characterized by broad (approximately normal) distributions with striking peaks at zero or negative values.
+Thus, the variety of possible value realizations depending on the prior parameter distributions, but also the cut-off threshold dependent cases of seal failures are represented.
+It was observed that Bayesian inference has to main effects on the characteristics of these value distributions, and consequently also on the realization of decisions from application of the custom loss function: (1) lateral shifting of the probability distributions and the Bayes actions; (2) narrowing of the distributions and convergence of the decisions of different risk-affine actors.
+	As these two mechanism seem to act independently from one another, at least to a certain degree, the following conclusions can be derived:
+	\begin{enumerate}
+	\item \textbf{More information does not necessarily lead to better decisions.} In cases, in which no considerable reduction of uncertainty is achieved by Bayesian inference, but posterior distributions are primarily shifted, the range of different actor's decisions will shift accordingly, but not converge, while expected losses remain approximately constant.
+This was observed in particular for cases, in which the standard deviations provided in the likelihoods were relatively large, or uncertainty was reduced very proximal to a seal failure-defining threshold value, so that a relatively narrow distribution still resulted in an approximately 50-50 binary probability of success versus failure (see 1D case: reduction of uncertainty for a seal formation close to the threshold of 20~m).
+	\item \text{The higher the uncertainty reduction, the better the decisions,} i.e. it is the nature of the information
+	\end{enumerate}
 
@@
 				-quantifiable value of information??
+
+				- The GREATER the REDUCTION IN UNCERTAINTY, the LOWER the relevance of the risk-factor and difference in actor's preferences
+
+				- accurate decision-making vs. gaining of information that yields encouraging results (vs "disappointing" results)

### 2017-11-15 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Figures, Text, Notebooks for Figures...

@@
 before production has started), volumetric estimation is the only approach to assess the amount of hydrocarbons in place in a reservoir.
-From this value, recoverable reserves can be estimated based on an estimated recovery factor \citep{dean2007volumetric}.\\
+From this value, recoverable reserves can be deduced based on an estimated recovery factor \citep{dean2007volumetric}.\\
         Oil-in-place and gas-in-place volumes are calculated based on:
@@
         \end{enumerate}
-        As the cases used in this work are purely artificial, a pure oil accumulation is assumed.
-Using the factors listed above, the respective equation for original oil-in-place ($OOIP$) is formulated as follows:
+        As the hydrocarbon systems in this work are purely artificial, a pure oil accumulation is assumed for simplicity.
+Furthermore, defined formations are regarded as homogeneous in their lithological composition and petrological properties It is assumed, for example, that no shale is interbedded in the main sandstone reservoir unit.
+Using the factors listed above, the equation for original oil-in-place ($OOIP$) is formulated as follows:
         \begin{equation}\label{eq:OOIP}
@@
 Thus, shrinkage of the oil volume brought to the surface is determined by $1/FvF$ \citep{dean2007volumetric}.\\
-        Subsequently, the effectively recoverable oil volumes ($ROV$) can be calculated by multiplying the $OOIP$ with a recovery factor $RF$:
+        Subsequently, the effectively recoverable oil volumes ($ROV$) can be calculated by multiplying the $OOIP$ with a recovery factor $RF$ \citep{dean2007volumetric}:
         \begin{equation}\label{eq:ROV}
@@
 For gas accumulations, however, recovery factors range typically between 70 and 90\% \citep{dean2007volumetric}.\\
-        The models in this work are artificial and not based on real data.
-The inputs to these equations can thus be chosen arbitrarily to test the applied methods.
-However, to come to significant conclusion, it makes sense to utilize values that approximately represent real possible scenarios.
-Furthermore, as 3D geological structures are modeled here, the hydrocarbon-filled rock volume ($A * h$ in Equations \ref{eq:OOIP} and \ref{eq:ROV}) is defined by every realization of the uncertain geological model.
-The hydrocarbon-filled rock volume is thus an uncertain factor.
-Remaining factors can be implemented deterministically or as uncertain values as well.
-WHICH ONE TO USE?\\%For most factors, it seems most suitable to define a normal distribution around a known typical value or global average, as it will return a high probability for common scenarios and occasionally produce cases of rare, more extreme values.\\
-        Consequently, $OOIP$ and $ROV$ calculations represent the step following and using the results from computing the geological model, so that the realizations are assigned values that can be interpreted from a perspective of economic interest.
-The resulting values from numerous model runs are used as a base probability distribution for the true value, on which loss function estimation can be applied as described in Section \ref{sec:loss}.
-$OOIP$ and $ROV$ are applicable in a 3D setting, the 1D geological model introduced below in Section \ref{sec:1D_model}, however, requires the use of an abstract valuation system.
+        In this work, the constructed geological models focus on the representation of the structural control in this equation.
+This refers in particular to potential structural traps and ultimately the hydrocarbon-filled rock volume represented by $A * h$ in Equations \ref{eq:OOIP} and \ref{eq:ROV}.
+        Including the remaining factors as simple uncertain parameters would be easily possible and is usually done in the probabilistic estimation procedures commonly conducted in the petroleum industry \citep{wim2001guidelines}.
+However, the focus of this work lies on the structural aspects of a hydrocarbon system.
+Respective results and observations might possibly be overprinted and distorted by the incorporation of other parameters of petrology and engineering as uncertain.
+Effects of interest might become unrecognizable.
+Thus, fixed values are used for the non-structural factors ($\phi=30\%$, $S_W=20\%$, $FvF=1.05$, $RF = 80\%$).
+These are chosen arbitrarily but within a realistic range to considering a conventional oil reservoir.
+Consequently, calculating the $ROV$ from the maximum trap volume simply equates a rescaling of values, as proportions are conserved.
+However, translation of trap volumes to recoverable oil volumes is deemed appropriate for a formally better consideration of the results in the respective context of decision-making in the hydrocarbon sector.\\
+        %For most factors, it seems most suitable to define a normal distribution around a known typical value or global average, as it will return a high probability for common scenarios and occasionally produce cases of rare, more extreme values.\\
+        Attaining the hydrocarbon-filled rock volume (i.e.
+the maximum trap volume) and calculating the $ROV$ is a central step of the model evaluation.
+Model realizations are hereby assigned the economic value on which decisions are to be based.
+Realized volumetric values from numerous iterations and several simulations are combined to sets of probability distributions of the $ROV$.
+Considering the $ROV$ as the true value to be estimated, loss functions can then be applied as described in Section \ref{sec:loss}.
+$OOIP$ and $ROV$ calculations are applicable in a 3D setting, the 1D geological model introduced below in Section \ref{sec:1D_model}, however, requires the use of an abstract valuation system.
         %\subsubsection{Net present value}
@@
 However, an attempt can be made to design a customized loss function that returns a Bayesian estimator involving the least bad consequences for an individual in a specific environment.
-Assuming the reservoir setting case and valuation methods introduced in Section \ref{sec:Reservoir_values}), such a customization attempt is made and explained step by step in the following.
-(SOURCES?)\\
+Assuming the reservoir setting case and valuation methods introduced in Section \ref{sec:Reservoir_values}), such a customization attempt is made and explained step by step in the following.\\
+		%%%(SOURCES?)
 		For the purpose of estimation, it makes sense that one of the standard loss functions is chosen as a basis and a customized loss function is developed from there.
@@
 The steps are illustrated in Figure \ref{fig:LF_4steps_normal} and listed in the following:
-		\begin{figure}[h]
+		\begin{figure}[p!]
 			\centering
@@
 			\centering
-			\includegraphics[width=1\textwidth]{Figures/LFR_normal.png}
+			\includegraphics[width=1\textwidth]{Figures/LFR_normal2.png}
 			\caption{Plotting of expected loss realizations after including the risk factor $r$ in the loss function (Equation \ref{eq:LFR_final}) for actors with risk-affinities ranging from risk-averse ($r$ = 0.5 and 0.75), over risk-neutral ($r$ = 1), to risk-friendly ($r$ = 1.25 and $r$ = 1.5).

### 2017-11-15 Risk analysis/Text/Layout/Layout/chapter3_results.tex
Figures, Text, Notebooks for Figures...

@@
 			\subsection{Bayesian inference using thickness likelihoods}	
-			\begin{figure}[h]
+			\begin{figure}[p!]
 				\centering
@@
 				\caption{Prior (A1) and posterior distributions (A2) of the layer boundary positions in depth and respective representations (A2, B2).
-Bayesian inference was conducted using likelihoods defined as follows: Seal thickness: $\mu = 25~m; \sigma = 20~m$; reservoir thickness: $\mu = 180~m; \sigma = 60~m$.
+Bayesian inference was conducted using likelihoods defined as follows: Seal thickness: $\mu = 25~m; \sigma = 20~m$; reservoir thickness: $\mu = 180~m; \sigma = 40~m$.
 From (A1) to (B1), the distributions are slightly narrowed.
@@
 
-				\subsubsection{Inference case I: Moderately reinforcing information}
+				\subsubsection{Inference case I: Moderately reinforcing or ambivalent information}
 				\begin{figure}[h]
@@
 					\end{subfigure}
-					\caption{Reservoir score distributions (A) and change in the realizations of expected loss for several risk parameters (B1, B2) before and after Bayesian updating based on likelihoods defined as follows: Seal thickness: $\mu = 25~m; \sigma = 20~m$; reservoir thickness: $\mu = 180~m; \sigma = 60~m$.}
+					\caption{Reservoir score distributions (A) and change in the realizations of expected loss for several risk parameters (B1, B2) before and after Bayesian updating based on likelihoods defined as follows: Seal thickness: $\mu = 25~m; \sigma = 20~m$; reservoir thickness: $\mu = 180~m; \sigma = 40~m$.}
 					\label{fig:update_moderate2_3}
 				\end{figure}				
-
 				In this first case, a normal distribution with a mean ($\mu$) of 25~m and a standard deviation ($\sigma$) of 20~m was chosen to reflect the likelihood of the seal thickness.
-For the reservoir, a normal distribution with a mean of 180~m and a standard deviation of 60~m was used respectively.
-As can be observed in Figure \ref{fig:update_moderate1}), the uncertainty in the probability distributions for the positions of layer boundaries in depth was reduced only moderately by conducting Bayesian inference with these values.\\
+For the reservoir, a normal distribution with a mean of 180~m and a standard deviation of 40~m was used respectively.
+As can be observed in Figure \ref{fig:update_moderate1}), the uncertainty in the probability distributions for the positions of layer boundaries in depth was reduced moderately, mostly for the reservoir bottom, by conducting Bayesian inference with these values.\\
 				Scoring was applied based on these new distributions.
-In Figure \ref{fig:update_moderate2_3} (A), it can be recognized that the bulk of the score distribution was shifted to the positive side of values, while the peak at -2000 was raised.
-The probability of scores between -2000 and 0 was decreased significantly, i.e.
-the true score is most likely either positive or -2000, if it is negative.\\
+In Figure \ref{fig:update_moderate2_3} (A), it can be recognized that the bulk of the score distribution was shifted to the positive side of values and narrowed, approximately resembling a normal distribution there.
+However, the peak at -2000 was raised, while the probability of scores between -2000 and 0 was decreased to be negligible, i.e.
+the true score is most likely either positive or -2000, if it is negative.
+So while uncertainties were reduced in areas of opposite sign, the divide between these was significantly increased.
+The overall uncertainty seems thus to have been widely maintained and barely transformed into a problem of duality.
+The change in uncertainty might thus me described as ambivalent.\\
 				Application of the custom loss function (Equation \ref{eq:LFR_final}) is visualized in Figure \ref{fig:update_moderate2_3}, in which the expected losses are compared before (B1) and after (B2) Bayesian inference.
@@
 In this case, the added data generally reinforces the probability of the reservoir to be sufficiently thick.
-Information on the seal, however, based on a normal distribution around 25~m thickness, leaves uncertainty about the reliability of the seal, as the safety threshold is defined as 20~m.\\
-				The decrease in uncertainty can overall be described as moderate.
-Increased certainty about the reservoir thickness was sufficient to shift Bayes actions to higher estimates for all actors, but the most risk-averse one ($r$ = 1.5).
-The risk-neutral actor's estimate was increased the most, the risk-friendliest actor's estimate the least.
-This has led to a slight convergence of the Bayes action in the realm of positive estimates.
-Expected losses remained at approximately the same magnitude, as before the likelihoods were considered.\\
+Information on the seal, however, based on a normal distribution around 25~m thickness, leaves high uncertainty about the reliability of the seal, as the safety threshold is defined as 20~m.
+Consequently, the risk of complete seal failure remains a major concern.\\
+				Increased certainty about the reservoir thickness was sufficient to shift Bayes actions to higher estimates for all actors.
+Estimates of the more risk neutral actors ($r = 0.5$ to $r=1.25$) were increased the most.
+This has led to a slight convergence of the Bayes actions.
+Expected losses were altered only slighty.
+For the risk neutral actor, this change is negligible.
+For risk-friendlier actors, the expected loss was decreased, but for risk-averse actors, as they shifted from zero ("take no action") to positive estimates, increased.\\
 				%These shifts are quantified in Table \ref{tab:update_examples_all}.Expected losses are decreased for the risk-neutral and the two risk-friendlier individuals.
@@
 					\end{subfigure}
-					\caption{Reservoir score distributions (A) and change in the realizations of expected loss for several risk parameters (B1, B2) before and after Bayesian updating based on likelihoods defined as follows: Seal thickness: $\mu = 50~m; \sigma = 20~m$; reservoir thickness: $\mu = 180~m; \sigma = 60~m$.}
+					\caption{Reservoir score distributions (A) and change in the realizations of expected loss for several risk parameters (B1, B2) before and after Bayesian updating based on likelihoods defined as follows: Seal thickness: $\mu = 50~m; \sigma = 20~m$; reservoir thickness: $\mu = 180~m; \sigma = 40~m$.}
 					\label{fig:update_goodseal2_3}
 				\end{figure}
-				In this second case, the reservoir thickness likelihood was defined in the same way as in case I ($\mu = 180~m; \sigma = 60~m$).
+				In this second case, the reservoir thickness likelihood was defined in the same way as in case I ($\mu = 180~m; \sigma = 40~m$).
 For the seal, a higher mean of 50~m with a standard deviation of 20~m was chosen, favoring the likelihood of a reliable seal relative to the threshold of 20~m thickness.
@@
 Very apparent is the significant diminishment of the "seal failure peak" at -2000.
-Due to this in combination with the overall distribution narrowing, mean and median were clearly shifted to higher values.\\
+Due to this, in combination with the overall distribution narrowing, mean and median were clearly shifted to higher values and are now much closer together.\\
 				Applying the custom loss function on this new score distribution resulted in the realizations of expected loss illustrated in Figure \ref{fig:update_goodseal2_3}.
@@
 The risk-neutral to risk-averse individuals seem to have profited the most, due to a large change in estimate value and decreased expected loss.
-Even the most risk-averse decision-maker is now willing expect a positive true score.
-Compared to the foregone case, a higher degree of uncertainty reduction was achieved in this case.
+Compared to the foregone case, a higher degree of overall uncertainty reduction was achieved.
+The risk of seal failure is now much lower.
 Consequently, Bayes actions were not only shifted to greater values, but also narrowed significantly in their range, i.e.
-the optimal decisions of the different actors have converged to a greater extent than in case I.\\
+the optimal decisions of the different actors have converged to a much greater extent than in case I.\\
 				%This is quantified in Table \ref{tab:update_examples_all}.
@@
 					\end{subfigure}
-					\caption{Reservoir score distributions (A) and change in the realizations of expected loss for several risk parameters (B1, B2) before and after Bayesian updating based on likelihoods defined as follows: Seal thickness: $\mu = 70~m; \sigma = 10~m$; reservoir thickness: $\mu = 100~m; \sigma = 40~m$.}
+					\caption{Reservoir score distributions (A) and change in the realizations of expected loss for several risk parameters (B1, B2) before and after Bayesian updating based on likelihoods defined as follows: Seal thickness: $\mu = 70~m; \sigma = 10~m$; reservoir thickness: $\mu = 100~m; \sigma = 30~m$.}
 					\label{fig:update_smallres2_3}
@@
 However, the new observations were assumed to provide information about the reservoir unit that makes it likely to be thinner than expected.
-Reservoir thickness likelihood was assigned a mean of 100~m and a standard deviation of 40~m.\\
+Reservoir thickness likelihood was assigned a mean of 100~m and a standard deviation of 30~m.\\
 				The subsequent score distribution is depicted in Figure \ref{fig:update_smallres2_3} (A).
-It can be seen that while the whole distribution was narrowed, it also was shifted to the left, to lower and negative estimates.
+It can be seen that while the whole distribution was strongly narrowed, it also was shifted to the left, to lower and negative estimates.
 Mean and median are almost equal, indicating narrowing towards symmetry in the shape of a normal distribution.
@@
 Due to this, the distribution was shifted to lower and negative values, strongly increasing the probability of an unfavorable true score.
+Consequently, only risk-friendly actors are willing to take action.
 				%Quantified changes in the positions and values of the Bayes actions are listed in Table \ref{tab:update_examples_all}.
@@
 
-			\subsection{General 1D model inference results}
-			This conceptual application in an abstract 1D geological model already serves to illuminate some of the main effects of Bayesian inference on decision-making.
-Two main mechanisms were observable in these three cases:
-			\begin{enumerate}
-				\item \textbf{Lateral shifting} of the score distribution and consequent shift in Bayes actions along the axis of possible score estimates, independent from change in uncertainty.
+			%\subsection{General 1D model inference results}
+			%This conceptual application in an abstract 1D geological model already serves to illuminate some of the main effects of Bayesian inference on decision-making.
+Three main mechanisms were observable in these cases:
+			%\begin{enumerate}
+			%	\item \textbf{Lateral shifting} of the score distribution and consequent shift in Bayes actions along the axis of possible score estimates, while their expected losses remain approximately constant.
+To a certain degree, this seems to be independent from the overall uncertainty reduction and is primarily related to the shift of probability to a different area.
 According to this observation, uncertainty reduction is not a necessary condition to significantly alter the decisions for all individuals.
-An incentive to bid on higher estimates is given solely by the shift of the posterior distribution to higher values, while the inherited degree of uncertainty may remain the same.
-				\item \textbf{Narrowing} of the score probability distribution, due to uncertainty reduction and consequent convergence of the Bayes actions of different actor's towards a similar estimate.
-Additionally, overall expected loss seems to decrease with uncertainty.
-If perfect information about our parameters was available, all individuals would bid on the same estimate and with an expected loss of zero.
-			\end{enumerate}
-			The second mechanism, i.e.
-uncertainty reduction, can thus presumed to be most significant regarding the process of decision-making, as actor's wish to minimize their expected loss and aim to bid on the most accurate estimate possible.
+				%An incentive to bid on higher estimates is given solely by the shift of the posterior distribution to higher values, while the inherited degree of uncertainty may remain the same.
+			%	\item \textbf{Narrowing} of the score probability distribution, due to uncertainty reduction and consequent convergence of the Bayes actions of different actor's towards a similar estimate.
+Additionally, expected loss seem to decrease with uncertainty.
+If perfect information about our parameters was available, all individuals would bid on the same estimate and expected a loss of zero.
+			%	\item \textbf{Dualization} of uncertainty, apparently due to reduction of uncertainty that narrows probability to areas of two possible outcomes, such as success and failure.
+The area at which uncertainty is reduced seems to be of significance.
+A high reduction of uncertainty in the form of standard deviation near a cut-off threshold value seems to lead to a transformation of the nature of uncertainty to a problem of duality.
+CONNECTION TO SHIFTING???
+			%\end{enumerate}
+			%The second mechanism, i.e.
+uncertainty reduction, can thus presumed to be most significant regarding the process of "good" decision-making, as actor's wish to minimize their expected loss from bidding on an estimate.
 In an optimal case, this would come in combination with a shift of probabilities (first mechanism) to more positive values, so that a higher economic value may be returned from the acquisition of information.
@@
 		\section{Results using the 3D geological model}
+
+		- results of ML1 generally similar as for ML3:
+			i.e. thicker reservoir does not mean higher volume, as we didn't implement a direct uncertainty to the shape, the spill point will stay more or less the same
+			-> spill and leak point as limiting factors to max volume
+			-> max volume already defined in prior (ROV max 5,500,000)
+		- see ML2: seal thickness as decisive factor
+			-> if seal thin and reservoir thick see the ratio of both!), higher likelihood of surpassing SSF

### 2017-11-15 Risk analysis/Text/Layout/Layout/chapter4_discussion.tex
Figures, Text, Notebooks for Figures...

@@
 				- accurate decision-making vs. gaining of information that yields encouraging results (vs "disappointing" results)
+
+	WHERE in the MODEL uncertainty is reduced is SIGNIFICANT for the DECISION MAKING PERSPECTIVE:
+	uncertainty might be reduced greatly in a model, e.g. looking at IE,
+	but this might happen to lead to high probabilities at areas which are critical to the decision-making problem, such as areas at which a threshold has been defined, leading to a stark cut-off limit between YES and NO (two extrema of decisions)
+	==> so maybe, while the objective model-related uncertainty has been reduced, the DECISION UNCERTAINTY remains constant or might even rise??

### 2017-11-15 Risk analysis/Text/Layout/Layout/chapter3_results.tex
Including Figures

@@
 
-		\section{Results using the 3D geological model}
+		\section{3D geological model results}
+		- first: MC simulation of only prior uncertainty error propagation to generate a prior reference model
+		- second: Bayesian inference using only thickness likelihoods
+		- third: Bayesian inference including SSF likelihood
+		- last test: BI using only SSF likelihood
 
+		\subsection{Prior model}
+		\begin{figure}[h]
+			\centering
+			\includegraphics[width=1\textwidth]{Figures/M_prior}
+			\caption{....}\label{fig:M_prior}
+		\end{figure}
+
+		\subsection{Posterior model I}
+		\begin{figure}[p!]
+			\centering
+			\includegraphics[width=1\textwidth]{Figures/ML1}
+			\caption{....}\label{fig:ML1}
+		\end{figure}
+
+		\subsection{Posterior model II}%ML3
+		-results similar to I, basically same
 		- results of ML1 generally similar as for ML3:
@@
 		-> max volume already defined in prior (ROV max 5,500,000)
+
+		\subsection{Posterior model III}%ML2
+		\begin{figure}[p!]
+			\centering
+			\includegraphics[width=1\textwidth]{Figures/ML2}
+			\caption{....}\label{fig:ML2}
+		\end{figure}
 		- see ML2: seal thickness as decisive factor

### 2017-11-15 Risk analysis/Text/Layout/Layout/chapter3_results.tex
Working on 3D results text

@@
 		\section{3D geological model results}
-		- first: MC simulation of only prior uncertainty error propagation to generate a prior reference model
-		- second: Bayesian inference using only thickness likelihoods
-		- third: Bayesian inference including SSF likelihood
-		- last test: BI using only SSF likelihood
+		In the following, results from applying the methods on the artificial 3D structural model are presented.
+Several representative examples of posterior models based on the use of different likelihoods are referenced to the original prior model.
+The following measures are central to the comparison of simulation results: (1) information entropy, (2) occurrence of trap control mechanisms, (3) resulting recoverable oil volumes and (3) consequent realization of expected losses and related Bayes actions (i.e.
+decisions).
 
 		\subsection{Prior model}
+		To generate a set of reference results, a prior model was first generated by running a simple simulation of Monte Carlo sampling and error propagation regarding only the uncertainty of prior parameters.
+No likelihood functions were included in this prior model.
+Prior parameter uncertainties were chosen to be identical for all models, prior and posteriors.
+		\begin{table}[h]
+		\centering
+		\begin{tabular}{lrrr} 
+			\toprule
+			(a) Uncertainty: layer boundaries $z$-position (normal distributions)\\  
+			\midrule 
+			Formation bottom & $\mu$ [m] & $\sigma$ [m]\\ 
+			\midrule 
+			Overlying & 0 & 40 \\
+			Sandstone 2 & 0 & 60 \\
+			Shale (Seal) & 0 & 80\\ 
+			Sandstone 1 (Reservoir) & 0 & 100 \\
+			\bottomrule
+			\toprule
+			(b) Uncertainty: normal fault offset (skew normal distribution)\\
+			\midrule
+			Fault side & $\mu$ [m] & $\sigma$ [m] & $\alpha$\\
+			\midrule
+			Hanging wall offset & 0 & -~150 & -~2\\
+			\bottomrule 
+		\end{tabular}
+		\caption{Prior parameter uncertainties defined by distributions with respective means $\mu$, standard deviations $\sigma$ and shape factor $\alpha$, as used in all models.
+Uncertainties about the position of layer interfaces (a) are implemented by drawing values from normal distributions.
+All points belonging to one layer are assigned one shared distribution.
+Drawn values are added or subtracted respectively from the originally defined $z$-positional values.
+The offset uncertainty (b) is based on a skew normal distribution that affects all positional points in the hanging wall.}
+		\label{tab:3D_prior_parameters}
+		\end{table}
 		\begin{figure}[h]
@@
 			\includegraphics[width=1\textwidth]{Figures/M_prior}
-			\caption{....}\label{fig:M_prior}
+			\caption{Recoverable oil volumes probability distribution for the prior model (a) and respective realization of the custom loss function (b).}\label{fig:M_prior}
 		\end{figure}
+		The nature of the prior uncertainties is reflected by the visualization using information entropy (see Figure \ref{fig:ML1} (a)).
+High entropies are found at and near the possible locations of layer boundaries.
+This effect is amplified by the offset uncertainty in the footwall, where the highest information entropies are found.\\
+		Regarding the different mechanisms of maximum trap volume control, it can be observed (see Figure \ref{fig:ML1} (c)) that all four relevant mechanisms occurred.
+The dominant factor is the anticlinal spill point (labeled "SPILL POINT"), accounting for more than 60\% of model realizations.
+It is followed by cross-fault leakage due to juxtaposition with formations underlying ("LEAK UNDER", ~20\%) and formations overlying ("LEAK OVER", ~13\%) the seal in the footwall.
+Leakage due to stratigraphical connections to seal-overlying units ("STRAT"), i.e.
+stratigraphical breaches of the seal, were registered to be decisive in only about 3\% of realizations.
+A satisfactory reliability of model algorithms and computation is indicated by the lack of unrecognized mechanisms (labeled "UNCLEAR", see Figure \ref{fig:ML1} (c)).\\
+		Recoverable oil volumes were calculated for each model iteration and plotted as a probability distribution in Figure \ref{fig:M_prior} (a).
+It can clearly be seen that while a large portion of the distribution is found in highly positive volumes, forming a normal distribution shape around 3.8 million $m³$, there also a high probability of complete failure, i.e.
+a maximum trap volume of zero.
+This results in a bimodal characterization, with low probability of the $ROV$ to be found between zero and three million $m³$.\\
+		Consequently, applying the custom loss function on this distribution results in divergent Bayes actions for the different risk-affine actors (see Figure \ref{fig:M_prior} (b)).
+Although all actors find their estimates to be positive, only the risk-neutral to risk-friendliest actors come close to the positive mode of the distribution.
+Risk-averse individuals bid on significantly lower estimates.
+It can also be observed, that expected losses seem to be generally lower for risk-affine actors.
+The different Bayes actions have apparently been torn apart by the bimodality of the $ROV$ probability distribution.
 
 		\subsection{Posterior model I}
+		\begin{table}[h]
+			\centering
+			\begin{tabular}{lrr} 
+				\toprule
+				Model I - Likelihoods: layer thicknesses (normal distributions)\\  
+				\midrule 
+				Formation & $\mu$ [m] & $\sigma$ [m]\\ 
+				\midrule 
+				Sandstone 2 & 150 & 30 \\
+				Shale (Seal) & 350 & 30\\ 
+				Sandstone 1 (Reservoir) & 250 & 30 \\
+				\bottomrule
+			\end{tabular}
+			\caption{Layer thickness likelihoods determined by normal distributions as used for posterior model I.}
+			\label{tab:ML1_likelihoods}
+		\end{table}
+		For the simulation of this first posterior model, thickness likelihoods for the three central layers were defined (see Table \ref{tab:ML1_likelihoods}).
 		\begin{figure}[p!]

### 2017-11-15 Risk analysis/Text/Layout/Layout/chapter3_results.tex
Auto stash before merge of "master" and "origin/master"

@@
 The dominant factor is the anticlinal spill point (labeled "SPILL POINT"), accounting for more than 60\% of model realizations.
-It is followed by cross-fault leakage due to juxtaposition with formations underlying ("LEAK UNDER", ~20\%) and formations overlying ("LEAK OVER", ~13\%) the seal in the footwall.
+It is followed by cross-fault leakage due to juxtaposition with formations underlying ("LEAK UNDER", ~20\%) and formations overlying ("LEAK OVER", ~13\%) the seal in the hanging wall.
 Leakage due to stratigraphical connections to seal-overlying units ("STRAT"), i.e.
@@
 
-		\subsection{Posterior model I}
+		\subsection{Posterior model I: Seal and reservoir likely thick}
 		\begin{table}[h]
@@
 		\end{table}
-		For the simulation of this first posterior model, thickness likelihoods for the three central layers were defined (see Table \ref{tab:ML1_likelihoods}).
+		For the simulation of this first posterior model, thickness likelihoods for the three central layers were included into the probabilistic model.
+These were defined in a way that reinforced the probabilities of the relevant layers to be significantly thick (see Table \ref{tab:ML1_likelihoods}).
 		\begin{figure}[p!]
@@
 		\end{figure}		
+		Respective results are summarized in Figure \ref{fig:ML1}.
+Considering the changes in information entropy, it can be seen that the uncertainty about layer interface positions in $z$ was clearly reduced in the footwall.
+Information entropy decreased also in the middle section of the hanging wall, but remained high in upper and lower parts, presumably due to the remaining influence of the offset uncertainty.\\
+		Regarding trap volume control (Figure \ref{fig:ML1} (c)), only spill point-related migration and cross-fault leakage due to juxtaposition with permeable layers below the seal in the hanging wall remain relevant in the posterior model.
+As no occurrences of leakage through overlying layers was observed, the two other mechanisms increased in frequency, while staying proportionally the same to one another.
+It can be presumed, that by introducing the respective likelihoods, a high thickness of the seal was ensured, as can also be recognized in Figure \ref{fig:ML1} (b).
+This not only made stratigraphic leakage in this model impossible, but also enforced a favorable $SSF$.
+A minor emergence of unrecognized control factors (label: "UNCLEAR") hints at a singular cases in which the model was possibly not correctly computable or the structural analysis failed for a different reason.\\
+		The bulk that accounted for the positive mode in the prior $ROV$ distribution was raised and shifted laterally to higher values, while the probability of failure ($ROV = 0$) was diminished significantly.
+This can presumably be explained by the disappearing of leakage to overlying formations as a trap control mechanism, which was programmed to always lead to complete failure of the trap volume.\\
+		As the probability of failure and thus the bimodality of the probability distribution disappeared, all Bayes actions were shifted to higher estimates, most significantly those of the two risk-averse individuals (see Figure \ref{fig:ML1} (e).
+There is thus clear convergence of the different points of minimum expected loss, which was also lowered overall.
+It is clear that the posterior state of knowledge, provided by the updated model, encourages even the most risk-averse decision-maker to expect recoverable oil volumes from 3.5 million m$^3$ upwards.
 
-		\subsection{Posterior model II}%ML3
-		-results similar to I, basically same
-		- results of ML1 generally similar as for ML3:
-		i.e. thicker reservoir does not mean higher volume, as we didn't implement a direct uncertainty to the shape, the spill point will stay more or less the same
-		-> spill and leak point as limiting factors to max volume
-		-> max volume already defined in prior (ROV max 5,500,000)
+		\subsection{Posterior model II: Likely thin reservoir formation}%ML3
+		For the following model, thickness likelihoods were defined in a way, that reinforced the probability of a very thick seal, but at the same time made restricted the reservoir formation to a likely significantly thinner reservoir (around 50 m) (see Table in Appendix).
+The results (see Appendix) are very similar to those of model I above.
+This can be explained by the influence of leak and spill point.
+As the shape of the general structure itself is not defined to be uncertain, the position of the anticlinal spill point relative to the fault top remains mostly unchanged, irrespective of layer thickness variations.
+Thus, reservoir thickness is a secondary factor to trap volume control.
+Given all other parameters are fixed, the maximum fill horizon is most likely the same for a 300 m as for a 50 m thick reservoir unit.
+The thickness of the reservoir only becomes respectively significant, if it thin enough to set the layer bottom above the leak and spill point depth in the trap section.
+Ultimately, the spill point appears to be the primariy limiting factor to maximum trap fill.
+It can furthermore be derived from this that the maximum trap volume capacity, as defined by the given structural features, is already represented in the $ROV$ probability distribution of the prior model.
+Examining this in Figure \ref{fig:ML1} (d), it seems that this overall maximum volume is found at around 5.7 millon m$3$.
+Higher values were not observed in any other model realization and are presumed to be unattainable due to the restriction given by the spill point.
 
@@
 		\end{figure}
+		Results from this third model show that in contrast to the reservoir thickness, seal thickness is a primary and decisive factor for trap volume.
+Using likelihoods which favor a thick reservoir, might only a thin seal of about 50 m thickness, leakage due to
 		- see ML2: seal thickness as decisive factor

### 2017-11-16 Risk analysis/Text/Layout/Layout/chapter3_results.tex
Some text progress, new IE Figure

@@
 
-		\subsection{Posterior model III}%ML2
+		\subsection{Posterior model III: Likely very thin seal}%ML2
 		\begin{figure}[p!]
@@
 		Results from this third model show that in contrast to the reservoir thickness, seal thickness is a primary and decisive factor for trap volume.
-Using likelihoods which favor a thick reservoir, might only a thin seal of about 50 m thickness, leakage due to
-		- see ML2: seal thickness as decisive factor
-			-> if seal thin and reservoir thick see the ratio of both!), higher likelihood of surpassing SSF
+Using likelihoods which favor a thick reservoir, but only a thin seal of about 50 m thickness, leakage due to juxtaposition with seal-overlying formations across the fault becomes the primary mechanism for trap volume control (see Figure \ref{fig:ML2} (c).
+It can recognized in Figure \ref{fig:ML2} (b) that while uncertainty was reduced clearly in the footwall, it remains high in the hanging wall.
+The state of information entropy in furthermore suggests that the seal layer in the appeared often significantly deeper hanging wall than in the footwall.
+Combined with a generally very thin seal, this allowed for the occurrence of high displacement to seal thickness ratios, i.e.
+high Shale Smear Factors.
+Thus, a frequenting surpassing of the critical $SSF$ and subsequent fault seal breaching was enabled.
+It is additionally apparent in Figure \ref{fig:ML2} (b), that while fault-related leakage has become predominant, control by the spill point has become widely irrelevant in this posterior model.\\	%It is suggested by the footwall information entropy in Figure \ref{label} (b), that the general shape of the structure has been flattened.
+		A dramatically higher probability of complete trap failure is induced by the high frequency of cross-fault leakage to overlying formations.
+Consequently, the bulk of the posterior $ROV$ probability distribution is dropped to a volume of zero, as depicted in Figure \ref{fig:ML2} (d).
+Any positive trap volumes have become highly improbable.\\
+		Bayes actions changed accordingly.
+Given this posterior $ROV$ distribution, all actors find their minimum of expected loss at $ROV = 0$, which equals the decision to "take no action" (see Figure \ref{fig:ML2} (e)).
+It has to be pointed out that all Bayes actions are found at the same point, due to the lack of negative values, into which more risk-averse actors would presumably move, if given the possibility.
+Compared to the prior model, expected losses were decreased overall.

### 2017-11-16 Risk analysis/Text/Layout/Layout/chapter3_results.tex
Results text pre-finished!

--- 2017-11-15 Risk analysis/Text/Layout/Layout/chapter3_results.tex
@@
 No likelihood functions were included in this prior model.
-Prior parameter uncertainties were chosen to be identical for all models, prior and posteriors.
+Prior parameter uncertainties were chosen to be identical in all model simulations.
 		\begin{table}[h]
@@
 		\end{figure}
-		The nature of the prior uncertainties is reflected by the visualization using information entropy (see Figure \ref{fig:ML1} (a)).
+		The nature of the prior uncertainties was visualized using information entropy (see Figure \ref{fig:ML1} (a)).
 High entropies are found at and near the possible locations of layer boundaries.
 This effect is amplified by the offset uncertainty in the footwall, where the highest information entropies are found.\\
-		Regarding the different mechanisms of maximum trap volume control, it can be observed (see Figure \ref{fig:ML1} (c)) that all four relevant mechanisms occurred.
-The dominant factor is the anticlinal spill point (labeled "SPILL POINT"), accounting for more than 60\% of model realizations.
-It is followed by cross-fault leakage due to juxtaposition with formations underlying ("LEAK UNDER", ~20\%) and formations overlying ("LEAK OVER", ~13\%) the seal in the footwall.
-Leakage due to stratigraphical connections to seal-overlying units ("STRAT"), i.e.
-stratigraphical breaches of the seal, were registered to be decisive in only about 3\% of realizations.
-A satisfactory reliability of model algorithms and computation is indicated by the lack of unrecognized mechanisms (labeled "UNCLEAR", see Figure \ref{fig:ML1} (c)).\\
+		The occurrence of the different mechanisms of maximum trap control were traced for all simulations and labeled as follows:
+		\begin{enumerate}
+		\item \textbf{SPILL POINT}: Migration related to the anticlinal spill point of the trap.
+		\item \textbf{LEAK UNDER}: Leakage due to juxtaposition of the reservoir with formations underlying the seal in the hanging wall.
+		\item \textbf{LEAK OVER}: Leakage due to juxtaposition of the reservoir with formations overlying the seal in the hanging wall.
+This was defined to lead to full leakage and complete failure of the trap.
+		\item \textbf{STRAT}: Leakage due to stratigraphical connections to seal-overlying units, i.e.
+stratigraphical breaches of the seal.
+Full leakage and trap failure were assumed for this mechanism as well.
+		\item \textbf{UNCLEAR}: Label defined to track cases in which a mechanism was not recognized.
+This is deemed an indicator for failures of the algorithms or the computation, i.e.
+the reliability of the overall model and incorporated functions.
+		\end{enumerate}
+		It is shown in Figure \ref{fig:ML1} (c) that all four relevant mechanisms occurred for the prior model.
+The dominant factor is the anticlinal spill point accounting for more than 60\% of model realizations.
+It is followed by "LEAK UNDER" ($\sim20\%$) and "LEAK OVER" ($\sim13\%$).
+Stratigraphical breaches of the seal were registered to be decisive in only about 3\% of iterations.
+A satisfactory reliability of model algorithms and computation is indicated by the lack of unrecognized mechanisms.\\
 		Recoverable oil volumes were calculated for each model iteration and plotted as a probability distribution in Figure \ref{fig:M_prior} (a).
-It can clearly be seen that while a large portion of the distribution is found in highly positive volumes, forming a normal distribution shape around 3.8 million $m³$, there also a high probability of complete failure, i.e.
-a maximum trap volume of zero.
-This results in a bimodal characterization, with low probability of the $ROV$ to be found between zero and three million $m³$.\\
+It can clearly be seen that while a large portion of the distribution is found in a range of highly positive volumes, forming a normal distribution shape around 3.8 million m$^3$, there is also a high probability of complete failure, i.e.
+the trap volume to be zero.
+This results in a bimodal characterization, with low probability of the $ROV$ to be found between zero and 3 million m$^3$.\\
 		Consequently, applying the custom loss function on this distribution results in divergent Bayes actions for the different risk-affine actors (see Figure \ref{fig:M_prior} (b)).
-Although all actors find their estimates to be positive, only the risk-neutral to risk-friendliest actors come close to the positive mode of the distribution.
+Although all actors find their estimates to be positive, only the risk-neutral to risk-friendliest actors come close to the described positive mode of the distribution ($ROV > 3*10^6$ m$^3$).
 Risk-averse individuals bid on significantly lower estimates.
-It can also be observed, that expected losses seem to be generally lower for risk-affine actors.
-The different Bayes actions have apparently been torn apart by the bimodality of the $ROV$ probability distribution.
+It can also be observed that expected losses seem to be generally lower for risk-affine actors.
+The divergence of different Bayes actions presumably resulted from the bimodality of the $ROV$ probability distribution.
 
-		\subsection{Posterior model I}
+		\subsection{Posterior model I: Seal and reservoir likely thick}
 		\begin{table}[h]
@@
 		\end{table}
-		For the simulation of this first posterior model, thickness likelihoods for the three central layers were defined (see Table \ref{tab:ML1_likelihoods}).
 		\begin{figure}[p!]
@@
 		\end{figure}		
+		For the simulation of this first posterior model, thickness likelihoods for the three central layers were included into the probabilistic model.
+These were defined in a way that reinforced the probabilities of the relevant layers to be significantly thick (see Table \ref{tab:ML1_likelihoods}).
+Respective results are summarized in Figure \ref{fig:ML1}.\\
+		Considering the changes in information entropy, it can be seen that the uncertainty about layer interface positions in $z$ was clearly reduced in the footwall.
+Information entropy was also decreased in the middle section of the hanging wall, but remained high in upper and lower parts, presumably due to the remaining influence of the offset uncertainty.\\
+		Regarding trap volume control (Figure \ref{fig:ML1} (c)), only spill point-related migration ("SPILL POINT") and cross-fault leakage due to juxtaposition with permeable layers below the seal in the hanging wall ("LEAK UNDER") remain relevant in the posterior model.
+As no occurrences of leakage through overlying layers was observed, the two other mechanisms increased in frequency, while staying proportionally the same to one another.
+It can be presumed, that by introducing the respective likelihoods, a high thickness of the seal was ensured, as can also be recognized in Figure \ref{fig:ML1} (b).
+This not only made stratigraphic leakage in this model impossible, but also enforced a favorable $SSF$.
+A minor emergence of unrecognized control factors ("UNCLEAR") suggests that, for a few iterations, the model was possibly not correctly computable or the structural analysis failed for a different reason.\\
+		The bulk that accounted for the positive mode in the prior $ROV$ distribution was raised and shifted laterally to higher values, while the probability of failure ($ROV = 0$) was diminished significantly (see Figure Figure \ref{fig:ML1} (d)).
+This can presumably be explained by the disappearing of leakage to overlying formations as a trap control mechanism, which was programmed to always lead to complete failure of the trap volume.\\
+		As the probability of failure and thus the bimodality of the probability distribution disappeared, all Bayes actions were shifted to higher estimates, most significantly those of the two risk-averse individuals (see Figure \ref{fig:ML1} (e).
+These have thus converged into an area of similar expectations.
+Additionally, expected losses were lowered overall.
+The posterior state of knowledge, provided by the updated model, encouraged even the most risk-averse decision-maker to expect recoverable oil volumes from 3.5 million m$^3$ upwards.
 
-		\subsection{Posterior model II}%ML3
-		-results similar to I, basically same
-		- results of ML1 generally similar as for ML3:
-		i.e. thicker reservoir does not mean higher volume, as we didn't implement a direct uncertainty to the shape, the spill point will stay more or less the same
-		-> spill and leak point as limiting factors to max volume
-		-> max volume already defined in prior (ROV max 5,500,000)
+		\subsection{Posterior model II: Likely thin reservoir formation}%ML3
+		For the second model, thickness likelihoods were defined in a way that reinforced the probability of a very thick seal but at the same time restricted the reservoir formation to a likely significantly thinner thickness of around 50 m (see Table in Appendix).
+The results (see Appendix) are very similar, in fact practically identical, to those of model I above.
+This can be explained by the influence of leak and spill point.
+As the shape of the general structure itself was not defined to be uncertain, the position of the anticlinal spill point relative to the fault top remains mostly unchanged, irrespective of layer thickness variations.
+Thus, reservoir thickness is only a secondary factor to trap volume control.
+Assuming fixed values for all other parameters, the maximum fill horizon is most likely the same for a 300 m as for a 50 m thick reservoir unit.
+The thickness of the reservoir only becomes significant, if it is thin enough to locate the respective layer bottom above the leak and spill point depth in the trap section.
+Ultimately, the spill point appears to be the primariy limiting factor to maximum trap fill.
+It can furthermore be derived from this that the maximum trap volume potential, as defined by the given structural features, is already represented in the $ROV$ probability distribution of the prior model.
+Examining this in Figure \ref{fig:ML1} (d), it seems that this overall maximum volume is found at approximately 5.7 millon m$3$.
+Higher values were not observed in any other model realization and are presumed to be unattainable due to the restriction given by the spill point.
 
-		\subsection{Posterior model III}%ML2
+		\subsection{Posterior model III: Likely very thin seal}%ML2
 		\begin{figure}[p!]
@@
 		\end{figure}
-		- see ML2: seal thickness as decisive factor
-			-> if seal thin and reservoir thick see the ratio of both!), higher likelihood of surpassing SSF
+		Results from this third model show that in contrast to the reservoir thickness, seal thickness is a primary and decisive factor for the manifestation of a reliable trap structure.
+Using likelihoods which favor a thick reservoir, but only a thin seal of about 50 m thickness, leakage due to juxtaposition with seal-overlying formations across the fault ("LEAK OVER") becomes the primary mechanism for trap volume control (see Figure \ref{fig:ML2} (c).
+It can be recognized in Figure \ref{fig:ML2} (b) that, while uncertainty was reduced clearly in the footwall, it remains high in the hanging wall.
+The state of information entropy furthermore suggests that the seal layer frequently appeared significantly deeper in the hanging wall than in the footwall.
+Combined with a generally very thin seal, this allowed for the occurrence of high displacement to seal thickness ratios, i.e.
+a high $SSF$.
+Thus, frequent surpassing of the critical $SSF$ and subsequent fault seal breaching were enabled.
+It is additionally apparent in Figure \ref{fig:ML2} (b), that while fault-related leakage has become predominant, control by the spill point was suppressed to be widely irrelevant in this posterior model.\\	%It is suggested by the footwall information entropy in Figure \ref{label} (b), that the general shape of the structure has been flattened.
+		A dramatically higher probability of complete trap failure is induced by the high frequency of cross-fault leakage to overlying formations.
+Consequently, the bulk of the posterior $ROV$ probability distribution is shifted to a volume of zero, as depicted in Figure \ref{fig:ML2} (d).
+Any positive trap volumes have become highly improbable.\\
+		Bayes actions changed accordingly.
+Given this posterior $ROV$ distribution, all actors find their minimum of expected loss at $ROV = 0$, which equals the decision to "take no action" (see Figure \ref{fig:ML2} (e)).
+It has to be pointed out that all Bayes actions are found at the same point due to the lack of negative values into which more risk-averse actors would presumably move, given the possibility.
+This is a difference to the score system implemented for the 1D geological model evaluation.
+Volumes cannot be negative, thus, Bayes actions cannot be shifted past the limit of $ROV = 0$.
+Compared to the prior model, expected losses were decreased overall, but are higher for more risk-friendly individuals at the minimum.
+
+		\subsection{Posterior model IV: SSF likelihood reinforcing trap reliability}%ML4
+		For this fourth model, thickness likelihoods were chosen to be comparable to those of model I, i.e.
+supporting the probability of thick layers in general.
+Additionally, a $SSF$-related likelihood function was implemented.
+With a mean of 1.5 and a standard deviation of 0.5, it was chosen to mostly reinforce the likelihood of a $SSF$ below the critical value of $SSF_c = 3$.\\
+		In Figure \ref{ML4} (b), it can be seen that the uncertainty about the position of interfaces in the footwall was greatly reduced for all layers.
+In the hanging wall, information entropies remained high but were shifted upwards and mainly concentrated in an area that is juxtaposed to the footwall seal formation.\\
+		In the resulting $ROV$ distribution, the possibility of trap failure was completely extinguished, while the probability of high volumes in the range between 3.5  and 5 million m$^3$ was significantly increased.
+Mean and median are found at almost the same value, but the bulk of the distribution found in this range appears to be slightly bimodal in itself, resembling approximately the shape of two normal distributions overlapping around a value of 4 million m$m^3$.\\
+		Accordingly, Bayes actions converged within this range of positive values.
+Considering the two modes in this range, risk-averse to risk-neutral bids remained in the middle of the mode of lower volumes, while risk-friendly actors were encouraged to estimate higher volumes as found in the second mode.
+Overall, significant convergence was achieved, also in respect to expected losses.\\
+		The importance of a reliable seal is once again emphasized by these results.
+Reliable fault sealing was favored by the chosen $SSF$ likelihood.
+It should be pointed out that computation of the complete model is required to attain the $SSF$.
+This value is furthermore characterized by a complex interdependence with other parameters and is most directly related to layer thicknesses and fault offset.
+For this reason, information entropy and offset uncertainty in the hanging wall was not greatly reduced.
+However, the certainty about the ratio of displacement to layer thicknesses was improved by using the $SSF$ in a likelihood function.
+
+		\subsection{Posterior model V: Controversial information about the Shale Smear Factor}%ML5
+		In this last model simulation, no thickness likelihoods, only a $SSF$-related likelihood was used.
+In contrast to model IV, this was chosen to narrow the probability of the $SSF$ to be around the threshold of its critical value, by basing it on a normal distribution with a mean of 3 and a standard deviation of 0.3.\\
+		Results are relatively similar to those of the prior model.
+It is notable that the information entropy regarding the position of layer boundaries in the footwall was decreased, while the entropy seems to have been raised significantly in the hanging wall (Figure \ref{ML5} (b)).
+Furthermore, an increase in the relevance of cross-fault leakage to overlying formations ("LEAK OVER") as a trap volume mechanism (increase from $\sim13\5$ to $\sim18\%$) was observed (Figure \ref{ML5} (c)).
+Consequently, the bimodality between the bulk of positive values and the probability of complete failure in the $ROV$ distribution was apparently reinforced.
+This is also indicated by a shift of mean and median to lower volumes (Figure \ref{ML5} (d)).
+Accordingly, a higher degree of decision divergence, i.e.
+a greater separation of different Bayes actions, was induced (Figure \ref{ML5} (e)).
+Expected losses of the minima were raised as well.\\
+		It is notable that the implementation of a likelihood which can be expected to significantly decrease the $SSF$ related uncertainty, overall leads to an amplification of the duality in the $ROV$ probability distribution and to dispersion of different risk-affine decisions.
+This effect is presumably caused by conservation of decisive uncertainty between "success" and complete failure of the trap, by narrowing the probability of the $SSF$ to be around its critical value.

### 2017-11-17 Risk analysis/Text/Layout/Layout/chapter3_results.tex
Figures in Appendix. Next: Discussion!

@@
 		\subsection{Posterior model V: Controversial information about the Shale Smear Factor}%ML5
-		In this last model simulation, no thickness likelihoods, only a $SSF$-related likelihood was used.
+		\begin{figure}[p!]
+			\centering
+			\includegraphics[width=1\textwidth]{Figures/ML5}
+			\caption{....}\label{fig:ML5}
+		\end{figure}
+		In this last case example, no thickness likelihoods, only a $SSF$-related likelihood was used.
 In contrast to model IV, this was chosen to narrow the probability of the $SSF$ to be around the threshold of its critical value, by basing it on a normal distribution with a mean of 3 and a standard deviation of 0.3.\\
 		Results are relatively similar to those of the prior model.
-It is notable that the information entropy regarding the position of layer boundaries in the footwall was decreased, while the entropy seems to have been raised significantly in the hanging wall (Figure \ref{ML5} (b)).
-Furthermore, an increase in the relevance of cross-fault leakage to overlying formations ("LEAK OVER") as a trap volume mechanism (increase from $\sim13\5$ to $\sim18\%$) was observed (Figure \ref{ML5} (c)).
+It is notable that the information entropy regarding the position of layer boundaries in the footwall was decreased, while the entropy seems to have been raised significantly in the hanging wall (Figure \ref{fig:ML5} (b)).
+Furthermore, an increase in the relevance of cross-fault leakage to overlying formations ("LEAK OVER") as a trap volume mechanism (increase from $\sim13\%$ to $\sim18\%$) was observed (Figure \ref{fig:ML5} (c)).
 Consequently, the bimodality between the bulk of positive values and the probability of complete failure in the $ROV$ distribution was apparently reinforced.
-This is also indicated by a shift of mean and median to lower volumes (Figure \ref{ML5} (d)).
+This is also indicated by a shift of mean and median to lower volumes (Figure \ref{fig:ML5} (d)).
 Accordingly, a higher degree of decision divergence, i.e.
-a greater separation of different Bayes actions, was induced (Figure \ref{ML5} (e)).
+a greater separation of different Bayes actions, was induced (Figure \ref{fig:ML5} (e)).
 Expected losses of the minima were raised as well.\\

### 2017-11-17 Risk analysis/Text/Layout/Layout/chapter4_discussion.tex
Continued discussion text (1 page)

--- 2017-09-29 Risk analysis/Text/Layout/Layout/chapter4_discussion.tex
@@
 	\chapter{Discussion}\label{cha:discussion}
+	Based on the recent findings by \citet{delaVarga2016}, showing that structural modeling can be viewed as a problem of Bayesian inference, the aim of this work was to extend this approach by considering practical applications and the utility geological modeling might have in an economic context.
+A sector in which structural geological modeling is of central importance and commonly used is hydrocarbon exploration.
+This field is characterized by the necessity to regularly make decisions in the face of high risks and potentially high rewards.
+As these decisions are often closely linked to geological modeling and the estimation of reservoir-related values, this work aimed to use this context to extend the Bayesian inference step in geological modeling by evaluating its results in terms of influence on decision making.
+It was initially hypothesized that Bayesian updating by using likelihoods and the resulting change in uncertainty in a model, should have significant effects on subsequent decision-making.\\
+	To observe this, geological models were interpreted as potential hydrocarbon systems.
+Algorithms for structural trap recognition and calculation of values relevant to respective decision-making were developed.
+A custom loss function was designed to express the environment in which such values are to be estimated and to represent preferences of differently risk-affine decision makers.
+These methods were for a conceptual 1D geological model and subsequently for a 3D structural model.
+Despite the great leap in complexity, the results from both models are widely similar.
+According to these, conducting Bayesian inference in the context of structual geological modeling of hydrocarbon system induced several main effects on reservoir value probability distributions and the consequent estimation step, i.e.
+the decision-making.\\
+	Changes in the posterior score and $ROV$ distributions were characterized by shifting of probabilities, as well as raising and narrowing of modes.
+The latter effect can be interpreted as a translation of uncertainty reduction in the model, to uncertainty reduction in the resulting distribution.
+Lateral shifting can be related to modes of the distribution as well as the distribution as a whole.
+...\\
+	One major observation to be named is that it seem to be of central importance "where" in the model uncertainty is reduced, i.e.
+in which spatial area or regarding which model parameter.
+This appears to be of particular importance considering threshold values that lead to an abrupt cut-off and duality in the decision-making.
+In both types of model construction, 1D and 3D, this was directly related to sealing reliability.
+Thresholds regarding seal thickness (1D model) and Shale Smear Factor ($SSF_c$ in 3D model) had been defined in a way that introduced a significant possibility of complete failure of a trap.
+Consequently, it was observed that reducing the uncertainty in a way that narrowed the probability of a threshold-related parameter around its cut-off value, led to an amplification of large-scale bimodality of the score or $ROV$ probability distribution.
+Resulting Bayes action were characterized by an increased divergence, i.e.
+disagreement between the estimates of different decision-makers.
+Thus, it appears suitable to introduce a differentiation between model uncertainty and decision uncertainty.
+Reducing model uncertainty in the "wrong" areas seems to simply lead to a transformation of this uncertainty in the realm of decision-making.
+Overall uncertainty seems to be conserved, as the duality of the decision problem is amplified in the form of a more stretched-out bimodal distribution and diverging Bayes actions.
+According to this, it should be a foremost importance for each actor, to reduce the uncertainty regarding threshold-related factors which might decide between "success" and complete failure of a project.\\
+	It furthermore can be argued that the degree of convergence of Bayes actions is a measure for the state of knowledge during the decision-making process.
+
+	3: high risk related to seal-reliability --> main factor for decision making: important to look for right "area" to reduce uncertainty
+
+	%Comparing the 1D and the 3D geological model, there is a large difference in complexity behind the model construction, as well as the way in which the decisive posterior value distributions (scores and recoverable volumes, respectively) are attained.
+Nevertheless, results of evaluating these models through application of the custom loss function are widely similar in principle.
+The reference distributions from the prior-only Monte Carlo uncertainty propagation sampling are comparable between the two cases.
+They are generally characterized by broad (approximately normal) distributions with striking peaks at zero or negative values.
+Thus, the variety of possible value realizations depending on the prior parameter distributions, but also the cut-off threshold dependent cases of seal failures are represented.
+It was observed that Bayesian inference has to main effects on the characteristics of these value distributions, and consequently also on the realization of decisions from application of the custom loss function: (1) lateral shifting of the probability distributions and the Bayes actions; (2) narrowing of the distributions and convergence of the decisions of different risk-affine actors.
+	%As these two mechanism seem to act independently from one another, at least to a certain degree, the following conclusions can be derived:
+	%\begin{enumerate}
+	%\item \textbf{More information does not necessarily lead to better decisions.} In cases, in which no considerable reduction of uncertainty is achieved by Bayesian inference, but posterior distributions are primarily shifted, the range of different actor's decisions will shift accordingly, but not converge, while expected losses remain approximately constant.
+This was observed in particular for cases, in which the standard deviations provided in the likelihoods were relatively large, or uncertainty was reduced very proximal to a seal failure-defining threshold value, so that a relatively narrow distribution still resulted in an approximately 50-50 binary probability of success versus failure (see 1D case: reduction of uncertainty for a seal formation close to the threshold of 20~m).
+	%\item \text{The higher the uncertainty reduction, the better the decisions,} i.e. it is the nature of the information
+	%\end{enumerate}
 
@@
 				-quantifiable value of information??
+
+				- The GREATER the REDUCTION IN UNCERTAINTY, the LOWER the relevance of the risk-factor and difference in actor's preferences
+
+				- accurate decision-making vs. gaining of information that yields encouraging results (vs "disappointing" results)
+
+	WHERE in the MODEL uncertainty is reduced is SIGNIFICANT for the DECISION MAKING PERSPECTIVE:
+	uncertainty might be reduced greatly in a model, e.g. looking at IE,
+	but this might happen to lead to high probabilities at areas which are critical to the decision-making problem, such as areas at which a threshold has been defined, leading to a stark cut-off limit between YES and NO (two extrema of decisions)
+	==> so maybe, while the objective model-related uncertainty has been reduced, the DECISION UNCERTAINTY remains constant or might even rise??

### 2017-11-18 Risk analysis/Text/Layout/Layout/chapter4_discussion.tex
In discussion

@@
 	\chapter{Discussion}\label{cha:discussion}
-	Based on the recent findings by \citet{delaVarga2016}, showing that structural modeling can be viewed as a problem of Bayesian inference, the aim of this work was to extend this approach by considering practical applications and the utility geological modeling might have in an economic context.
+	Building upon the recent findings by \citet{delaVarga2016}, showing that structural modeling can be viewed as a problem of Bayesian inference, the aim of this work was to extend this approach by considering practical applications and the utility geological modeling might have in an economic context.
 A sector in which structural geological modeling is of central importance and commonly used is hydrocarbon exploration.
-This field is characterized by the necessity to regularly make decisions in the face of high risks and potentially high rewards.
-As these decisions are often closely linked to geological modeling and the estimation of reservoir-related values, this work aimed to use this context to extend the Bayesian inference step in geological modeling by evaluating its results in terms of influence on decision making.
-It was initially hypothesized that Bayesian updating by using likelihoods and the resulting change in uncertainty in a model, should have significant effects on subsequent decision-making.\\
-	To observe this, geological models were interpreted as potential hydrocarbon systems.
+This field is characterized by the necessity to make decisions in the face of high risks and potentially high rewards.
+As these decisions are often closely linked to geological modeling and the estimation of reservoir-related values, this is work aimed to use this context to extend the Bayesian inference step in geological modeling by evaluating its results in terms of influence on decision-making.
+It was initially hypothesized that Bayesian updating by using likelihoods and the resulting change of uncertainty in a model, should have significant effects on subsequent decision-making.\\
+	To analyze this, geological models were interpreted as potential hydrocarbon systems.
 Algorithms for structural trap recognition and calculation of values relevant to respective decision-making were developed.
-A custom loss function was designed to express the environment in which such values are to be estimated and to represent preferences of differently risk-affine decision makers.
-These methods were for a conceptual 1D geological model and subsequently for a 3D structural model.
+A custom loss function was designed to express the environment in which such values are to be estimated and to represent preferences of differently risk-affine decision-makers.
+These methods were applied for a conceptual 1D geological model and subsequently for a 3D structural model.
 Despite the great leap in complexity, the results from both models are widely similar.
-According to these, conducting Bayesian inference in the context of structual geological modeling of hydrocarbon system induced several main effects on reservoir value probability distributions and the consequent estimation step, i.e.
+According to these, conducting Bayesian inference in the context of structural geological modeling of hydrocarbon system induces several main effects on reservoir value probability distributions and the consequent estimation step, i.e.
 the decision-making.\\
-	Changes in the posterior score and $ROV$ distributions were characterized by shifting of probabilities, as well as raising and narrowing of modes.
-The latter effect can be interpreted as a translation of uncertainty reduction in the model, to uncertainty reduction in the resulting distribution.
-Lateral shifting can be related to modes of the distribution as well as the distribution as a whole.
-...\\
+	As score and $ROV$ probability distributions were defined to depend on model parameters directly, but also on their values related to other parameters and specific thresholds, changing the model uncertainty naturally resulted in a change of shape of related posterior distributions.
+For both model designs, the prior value probability distributions were characterized by a distinct bimodality formed by widely separated modes of opposing values (positive vs.
+negative or zero).
+By introducing likelihood functions, these modes were most often shifted or altered in their shape.
+The latter was observed as either an amplification of the mode, by raising its mean probability, while narrowing the range of probabilities (i.e.
+reducing the standard deviation), or the opposite, a diminishment of the mode up to total erasure of its probability.
+These changes were often accompanied by lateral shifting of a mode.
+It can be argued that these processes reflect a translation of the model uncertainty into uncertainty in the reservoir value distribution.
+Within this it can be differentiated between uncertainty related to single modes, and uncertainty regarding the overall distribution.
+In some inference cases, the degree of bimodality was increased.
+This was caused by amplificaton of one of the modes, while the other remained constant or was also raised.
+So while uncertainty was reduced for one or both modes, the overall spread of probabilities and thus the total uncertainty was conserved, indicated by the mean of the whole distribution located in between both modes.
+In other cases, one mode was diminished or even completely erased, while the other was amplified.
+Resulting in overall unimodal distributions with centralized means, characterized accordingly by a reduction in total uncertainty.
+As the process of decision-making is based on these value distributions by applying the custom loss function respectively, this overall uncertainty might well be referred to as "decision uncertainty", here.\\
+	The decision-making of differently risk-affine actors, i.e.
+the position and spread of Bayes actions, changed relative to the properties of value probability distributions and their inherent decision uncertainty.
+Distinct bimodality between two extremes, i.e.
+high decision uncertainty, resulted in a higher spread of Bayes estimators.
+Reduction of the distribution to one mode conversely led to the convergence of different Bayes actions.
+Reduction of the overall decision uncertainty furthermore was accompanied by a decrease in expected loss for each decision.
+Consequently, it can be argued that the degree of convergence of Bayes actions and their expected loss can be considered measures for the state of knowledge during the decision-making process.
+The better this is, the more similar the decisions of differently risk-affine actors are.
+It can be assumed that given perfect information, all actors would bid on the same estimate, the true value, and expect no loss.
+The relevance of the risk-factor consequently decreases with higher reduction of decision uncertainty.\\
 	One major observation to be named is that it seem to be of central importance "where" in the model uncertainty is reduced, i.e.
 in which spatial area or regarding which model parameter.
-This appears to be of particular importance considering threshold values that lead to an abrupt cut-off and duality in the decision-making.
-In both types of model construction, 1D and 3D, this was directly related to sealing reliability.
-Thresholds regarding seal thickness (1D model) and Shale Smear Factor ($SSF_c$ in 3D model) had been defined in a way that introduced a significant possibility of complete failure of a trap.
-Consequently, it was observed that reducing the uncertainty in a way that narrowed the probability of a threshold-related parameter around its cut-off value, led to an amplification of large-scale bimodality of the score or $ROV$ probability distribution.
-Resulting Bayes action were characterized by an increased divergence, i.e.
-disagreement between the estimates of different decision-makers.
-Thus, it appears suitable to introduce a differentiation between model uncertainty and decision uncertainty.
+This appears to be of particular significance considering threshold values that lead to an abrupt cut-off between two extrema of decision options.
+In both types of model construction, 1D and 3D, this is directly related to sealing reliability.
+Thresholds regarding seal thickness (1D model) and Shale Smear Factor ($SSF_c$ in 3D model) were defined in a way that introduced a significant possibility of complete failure of a trap.
+Consequently, it was observed that reducing the uncertainty in a way that narrowed the probability of a threshold-related parameter around its cut-off value, led to an amplification of the respective mode (negative or zero values) and thus emphasized the risk of complete failure.
+Resulting Bayes actions were characterized by an increase in divergence and expected losses.
 Reducing model uncertainty in the "wrong" areas seems to simply lead to a transformation of this uncertainty in the realm of decision-making.
-Overall uncertainty seems to be conserved, as the duality of the decision problem is amplified in the form of a more stretched-out bimodal distribution and diverging Bayes actions.
-According to this, it should be a foremost importance for each actor, to reduce the uncertainty regarding threshold-related factors which might decide between "success" and complete failure of a project.\\
-	It furthermore can be argued that the degree of convergence of Bayes actions is a measure for the state of knowledge during the decision-making process.
+Overall uncertainty seems to be conserved or even increased (see 3D posterior model V), as the duality of the decision problem is amplified in the form of a more stretched-out bimodal distribution and diverging Bayes actions.
+According to this, it should be of foremost importance for each actor, to reduce the uncertainty regarding threshold-related factors which might decide between "success" and complete failure of a project.
+Some type of information might improve the potential for a positive outcome, but maintain the risk of failure.
+For making better decisions, elimination of such high risks should be a priority.
+In the context of hydrocarbon traps, this is greatly influenced by seal reliability.\\
 
-	3: high risk related to seal-reliability --> main factor for decision making: important to look for right "area" to reduce uncertainty
+	- IE not adequate method for finding this "where", as such a threshold value might dependent on more on the relation of different parameters, than on the realization of one in the model
+
+	It has too be emphasized that the models constructed in this work were purely artificial and not based on real data.
+However, the 3D model in particular was designed to include some typical structural characteristics related to hydrocarbon systems and algorithms were developed to consider the most common conditions that define structural traps.
+The fact that similar observations were made for the 1D, as well as the 3D model, indicates that these might be appropriately representative to conceptually visualize the effects of the methods.\\
+	For the purpose of a conceptual application, the customization of the loss functions was kept relatively simple.
+In fact, as it incorporated only a few basic assumptions, it might thus be more generally representative than a complex function which takes into account too specific aspects.
+The consideration of more details would have furthermore required extensive speculation, which would have presumably impaired the generalization potential of respective observations.
 
@@
 	%\end{enumerate}
-
-	1D CASE:
-
-				-abstract case: easy model construction and straightforward design of a loss function making basic assumptions and taking relative values that can simply be exchanged
-
-				-so representative in a "relative" way, mostly appropriate to illustrate principles and benefits of the methodology
-
-				-decision making/ estimation is defined by the design of the loss function which includes framework parameters which depend on the problem environment (e.g. market, technical constraints, etc.)
-
-				-similar actors but different behaviors concerning risk: different decisions but same general loss function path (just different minima of EL)
-
-				-additional information changes decision (BA) and EL
-
-				-magnitude of change varies not only with the nature of the information (magnitude of uncertainty reduction) but also with the risk parameter
-
-				-some actors might benefit more from some specific additional information than others
-
-				-quantifiable value of information??
-
-				- The GREATER the REDUCTION IN UNCERTAINTY, the LOWER the relevance of the risk-factor and difference in actor's preferences
-
-				- accurate decision-making vs. gaining of information that yields encouraging results (vs "disappointing" results)
-
-	WHERE in the MODEL uncertainty is reduced is SIGNIFICANT for the DECISION MAKING PERSPECTIVE:
-	uncertainty might be reduced greatly in a model, e.g. looking at IE,
-	but this might happen to lead to high probabilities at areas which are critical to the decision-making problem, such as areas at which a threshold has been defined, leading to a stark cut-off limit between YES and NO (two extrema of decisions)
-	==> so maybe, while the objective model-related uncertainty has been reduced, the DECISION UNCERTAINTY remains constant or might even rise??

### 2017-11-20 Risk analysis/Text/Layout/Layout/chapter3_results.tex
Discussion needs rework!

@@
 		\subsection{Prior model}
-		To generate a set of reference results, a prior model was first generated by running a simple simulation of Monte Carlo sampling and error propagation regarding only the uncertainty of prior parameters.
-No likelihood functions were included in this prior model.
-Prior parameter uncertainties were chosen to be identical in all model simulations.
+		To generate a set of reference results, a prior model was first generated by running a simple simulation of Monte Carlo sampling considering only the uncertainty of prior parameters.
+No likelihood functions were included.
+The prior parameter uncertainties, defined in Table \ref{tab:3D_prior_parameters} were chosen to be identical for all posterior simulations.
 		\begin{table}[h]
@@
 		\end{tabular}
-		\caption{Prior parameter uncertainties defined by distributions with respective means $\mu$, standard deviations $\sigma$ and shape factor $\alpha$, as used in all models.
-Uncertainties about the position of layer interfaces (a) are implemented by drawing values from normal distributions.
-All points belonging to one layer are assigned one shared distribution.
-Drawn values are added or subtracted respectively from the originally defined $z$-positional values.
+		\caption{Prior parameter uncertainties defined by distributions with respective means $\mu$, standard deviations $\sigma$ and shape factor $\alpha$, as used in all model simulations.
+Points belonging to the same layer interface share the same normal distribution (a).
 The offset uncertainty (b) is based on a skew normal distribution that affects all positional points in the hanging wall.}
@@
 			\caption{Recoverable oil volumes probability distribution for the prior model (a) and respective realization of the custom loss function (b).}\label{fig:M_prior}
-		\end{figure}
+		\end{figure}\\
 		The nature of the prior uncertainties was visualized using information entropy (see Figure \ref{fig:ML1} (a)).
 High entropies are found at and near the possible locations of layer boundaries.
-This effect is amplified by the offset uncertainty in the footwall, where the highest information entropies are found.\\
+This effect was amplified by the offset uncertainty in the footwall, where the highest information entropies are found.\\
 		The occurrence of the different mechanisms of maximum trap control were traced for all simulations and labeled as follows:
@@
 This was defined to lead to full leakage and complete failure of the trap.
-		\item \textbf{STRAT}: Leakage due to stratigraphical connections to seal-overlying units, i.e.
+		\item \textbf{STRAT}: Leakage due to stratigraphical connections of the trap to seal-overlying units, i.e.
 stratigraphical breaches of the seal.
@@
 The dominant factor is the anticlinal spill point accounting for more than 60\% of model realizations.
-It is followed by "LEAK UNDER" ($\sim20\%$) and "LEAK OVER" ($\sim13\%$).
+It is followed by cross-fault leakage labeled "LEAK UNDER" ($\sim20\%$) and "LEAK OVER" ($\sim13\%$).
 Stratigraphical breaches of the seal were registered to be decisive in only about 3\% of iterations.

### 2017-11-20 Risk analysis/Text/Layout/Layout/chapter4_discussion.tex
Discussion needs rework!

--- 2017-11-15 Risk analysis/Text/Layout/Layout/chapter4_discussion.tex
@@
 	\chapter{Discussion}\label{cha:discussion}
-	Comparing the 1D and the 3D geological model, there is a large difference in complexity behind the model construction, as well as the way in which the decisive posterior value distributions (scores and recoverable volumes, respectively) are attained.
-Nevertheless, results of evaluating these models through application of the custom loss function are widely similar in principle.
-The reference distributions from the prior-only Monte Carlo uncertainty propagation sampling are comparable between the two cases.
-They are generally characterized by broad (approximately normal) distributions with striking peaks at zero or negative values.
-Thus, the variety of possible value realizations depending on the prior parameter distributions, but also the cut-off threshold dependent cases of seal failures are represented.
-It was observed that Bayesian inference has to main effects on the characteristics of these value distributions, and consequently also on the realization of decisions from application of the custom loss function: (1) lateral shifting of the probability distributions and the Bayes actions; (2) narrowing of the distributions and convergence of the decisions of different risk-affine actors.
-	As these two mechanism seem to act independently from one another, at least to a certain degree, the following conclusions can be derived:
-	\begin{enumerate}
-	\item \textbf{More information does not necessarily lead to better decisions.} In cases, in which no considerable reduction of uncertainty is achieved by Bayesian inference, but posterior distributions are primarily shifted, the range of different actor's decisions will shift accordingly, but not converge, while expected losses remain approximately constant.
-This was observed in particular for cases, in which the standard deviations provided in the likelihoods were relatively large, or uncertainty was reduced very proximal to a seal failure-defining threshold value, so that a relatively narrow distribution still resulted in an approximately 50-50 binary probability of success versus failure (see 1D case: reduction of uncertainty for a seal formation close to the threshold of 20~m).
-	\item \text{The higher the uncertainty reduction, the better the decisions,} i.e. it is the nature of the information
-	\end{enumerate}
+	Building upon the recent findings by \citet{delaVarga2016}, showing that structural modeling can be viewed as a problem of Bayesian inference, the aim of this work was to extend this approach by considering practical applications and the utility geological modeling might have in an economic context.
+A sector in which structural geological modeling is of central importance is hydrocarbon exploration.
+This field is characterized by the necessity to make decisions in the face of high risks and potentially high rewards.
+As these decisions are often closely linked to geological modeling and the estimation of reservoir-related values, this is work aimed to use this context to extend the Bayesian inference step in geological modeling by evaluating its results in terms of influence on respective decision-making.
+It was initially hypothesized that Bayesian updating via likelihoods and the resulting change of uncertainty in a model, should have significant effects on subsequent decision-making.\\
+	To analyze this, geological models were interpreted as potential hydrocarbon systems.
+Algorithms for structural trap recognition and calculation of values relevant to respective decision-making were developed.
+Based on many iterations, probability distributions for such economic values (score and $ROV$, respectively) were generated.
+A custom loss function was designed to express the environment in which such values are to be estimated and to represent preferences of differently risk-affine decision-makers.
+These methods were applied for a conceptual 1D geological model and subsequently for a 3D structural model.
+Despite the great leap in complexity, the results from both models are widely similar.\\
+	Scores in the 1D model and $ROV$ in the 3D model were chosen to represent a measure for the economic value realized in each model construction.
+As such, they were defined to depend on model parameters, either directly or indirectly by valuing their relations to other parameters or specific thresholds and conditions.
+Accordingly, their posterior probability distributions and respective uncertainties changed with the implementation of Bayesian inference.\\
+	These changes related mostly to main modes in the distributions.
+For both model designs (1D and 3D), the prior value probability distributions were characterized by a distinct bimodality formed by widely separated modes of opposing values, i.e.
+one in the range of highly postive values, another at zero or negative values.
+By introducing likelihood functions, these modes were most often shifted or altered in their shape.
+The latter was observed as either an amplification of the mode, by raising its mean probability, while narrowing the range of probabilities (i.e.
+reducing the standard deviation), or the opposite, a diminishment of the mode up to total erasure of its probability.
+	Considering that these mode changes correlated with increased or decreased relevance of specific trap control mechanisms in the 3D model, it can be argued that each mode represents the occurrence of a specific mechanism-related scenario.
+These are mostly distinctly separated with little probability in between.
+This presumably is based on the assumption that traps are "filled-to-spill" in respect to spill and leak point, and that full leakage is assumed for any type of seal failure, resulting in an approximate discretization into respective scenarios.\\
+	Bayesian inference primarily seems to lead to a change in uncertainty in respect to these scenarios or modes.
+The proportionality regarding these modes, their relative probabilities and distance to each other, in turn influence the overall uncertainty of the probability distribution as a whole.
+This was often indicated by the position of the mean and median relative to the modes.
+A mean located in the middle between two widely separated modes, which might be of low intrinsic uncertainty themselves, would indicate a much higher overall uncertainty, than a mean found in the middle of a narrow unimodal distribution.
+	As the process of decision-making is based on these value distributions by applying the custom loss function respectively, it is proposed here, that this overall uncertainty might be referred to as "decision uncertainty".\\
+	The decision-making of differently risk-affine actors, i.e.
+the position and spread of Bayes actions, changed relative to the properties of value probability distributions and their inherent decision uncertainty.
+Distinct bimodality between two extremes, i.e.
+high decision uncertainty, resulted in a higher spread of Bayes estimators.
+Reduction of the distribution to one mode conversely led to the convergence of different Bayes actions.
+A decrease decision uncertainty furthermore was accompanied by a reduction in expected loss for each decision.
+Risk-friendly actors were the most robust in their decision-making in the face of possible trap failure.
+Eliminating this risk proved to be far less significant to the most risk-friendly actor.
+The respective estimate was generally found close to the positive mode and experienced only minor lateral shifting, after diminishing the probability of failure.
+Greater shifts were caused for this actor, in cases in which the positive mode itself was shifted laterally or erased completely.
+Conversely, the Bayes action of the most-averse individual was influenced greatly by the probability of the "worst scenario".
+Only a stark decrease of the respective mode encouraged this actor to bid on an estimate close to the risk-friendlier decision-makers.
+These observations match the principles and conditions conceived in the design of the loss function and the applied loss function.
+A clear difference in the decision-making relative to risk-affinity was observed.\\
+	It can be argued that the degree of convergence of Bayes actions and their expected loss can be considered measures for the state of knowledge during the decision-making process.
+The better this is, the more similar the decisions of differently risk-affine actors are.
+It can be assumed that given perfect information, all actors would bid on the same estimate, the true value, and expect no loss.
+Thus, the relevance of the risk-factor decreases with higher reduction of decision uncertainty.\\
+	While in most cases the inclusion of additional information as likelihoods led to a reduction of decision uncertainty, this was not always the case.
+It appears that the change in decision uncertainty is not necessarily strictly aligned with the change in uncertainty regarding model parameters and their derivatives.
+In respect to this, it seems to be of central importance "where" in the model uncertainty is reduced, i.e.
+in which spatial area or regarding which model parameter.
+This appears to be of particular significance considering threshold values that lead to an abrupt cut-off between two extrema of decision options.
+In both types of model construction, 1D and 3D, this is directly related to sealing reliability.
+Thresholds regarding seal thickness (1D model) and Shale Smear Factor ($SSF_c$ in 3D model) were defined in a way that introduced a significant possibility of complete failure of a trap.
+Consequently, it was observed that reducing the uncertainty in a way that narrows the probability of a threshold-related parameter around its cut-off value, can lead to an amplification of the respective mode and thus emphasize the risk of complete failure.
+This can well be seen in the posterior $ROV$ distribution of model V (3D).
+Resulting Bayes actions were characterized by an increase in divergence and expected losses.\\
+	Reducing model uncertainty in the "wrong" areas seems to simply lead to a transformation of this uncertainty in the realm of decision-making.
+Overall uncertainty seems to be conserved to a certain extent or even increased, as the duality of the decision problem is amplified in the form of a more stretched-out bimodal distribution and diverging Bayes actions.
+According to this, it should be of foremost importance for each actor, to reduce the uncertainty regarding threshold-related factors which might decide between "success" and complete failure of a project.
+Some type of information might improve the potential magnitude of a positive outcome, but maintain the risk of failure.
+For making better decisions, elimination of such high risks should be a priority.
+In the context of hydrocarbon traps, seal reliability is decisive.\\
+	Such risks might be easily assessed, if they are dependent on only one or a few parameters, such as seal thickness in the 1D model.
+In other cases, they are derived from more complex parameter inter-relations, as is the case for the Shale Smear Factor, the ratio of displacement to seal thickness.
+Models IV and V (3D) showed that reduced uncertainty about the $SSF$ is not necessarily directly recognizable by information entropy visualization, especially if uncertainty regarding its parent parameters remains high.
+It follows that, to approach an effective mitigation of high risks, the complexities behind decisive factors need to be assessed and respective parent parameters and inter-dependencies need to be recognized.
+This might enable a better understanding of which type of information is missing and where in the model, additional data might be of use for improved decision-making.\\
 
-	1D CASE:
+	It has too be emphasized that the models constructed in this work were purely artificial and not based on real data.
+Nevertheless, the 3D model in particular was designed to include some typical structural characteristics related to hydrocarbon systems and algorithms were developed to consider the most common conditions that define structural traps.
+The fact that similar observations were made for the 1D, as well as the 3D model, indicates a certain degree of continuity with respect to representativity.
+However, it also has to be pointed out, that the uncertainties employed in the 3D model regarded positional values in the $z$-axis only, and were of one-dimensional nature.
+This might party be reason for the parallelism in the results.
+There was thereby also no real uncertainty concerning the overall structural shape, particularly anticlinal features and the position of the spill point in relation to the trap maximum.\\
 
-				-abstract case: easy model construction and straightforward design of a loss function making basic assumptions and taking relative values that can simply be exchanged
+	On algorithms
 
-				-so representative in a "relative" way, mostly appropriate to illustrate principles and benefits of the methodology
+	For the purpose of a conceptual application, the customization of the loss functions was kept relatively simple.
+As mentioned above, despite its simplicity, the design of the function and the incorporation of a risk factor appears to have been suitable to achieve a clear distinction between differently risk-affine actors and respective behavior.
+In contrast to this, the use of standard loss function would have only returned mean or median as estimates.
+Customization of such loss functions offers a flexible approach to express different objective and subjective weighting factors that are specified to certain decision-making environments and decision-makers' perspectives.
+It is thereby possible to extend this design to attain very complex loss functions.
+As the loss function in this work was designed based on only a few basic assumptions, it might be more generally representative than a complex function which takes into account numerous very specific aspects.
+The consideration of more details, without a base of real data, would have furthermore required extensive speculation, which would have presumably impaired the generalization potential of respective observations.\\
 
-				-decision making/ estimation is defined by the design of the loss function which includes framework parameters which depend on the problem environment (e.g. market, technical constraints, etc.)
-
-				-similar actors but different behaviors concerning risk: different decisions but same general loss function path (just different minima of EL)
-
-				-additional information changes decision (BA) and EL
-
-				-magnitude of change varies not only with the nature of the information (magnitude of uncertainty reduction) but also with the risk parameter
-
-				-some actors might benefit more from some specific additional information than others
-
-				-quantifiable value of information??
-
-				- The GREATER the REDUCTION IN UNCERTAINTY, the LOWER the relevance of the risk-factor and difference in actor's preferences
-
-				- accurate decision-making vs. gaining of information that yields encouraging results (vs "disappointing" results)
-
-	WHERE in the MODEL uncertainty is reduced is SIGNIFICANT for the DECISION MAKING PERSPECTIVE:
-	uncertainty might be reduced greatly in a model, e.g. looking at IE,
-	but this might happen to lead to high probabilities at areas which are critical to the decision-making problem, such as areas at which a threshold has been defined, leading to a stark cut-off limit between YES and NO (two extrema of decisions)
-	==> so maybe, while the objective model-related uncertainty has been reduced, the DECISION UNCERTAINTY remains constant or might even rise??
+	Considering the findings of this work, there is still many points that could be expanded on in future research.
+It would be of interest to apply the same overall concept and methods on an authentic case based on real datasets.
+Given a realistic economic scenario, including capital and operational expenditures of a project, possibly a full net-present-value ($NPV$) analysis could be conducted.
+Recoverable reserves could be replaced by the $NPV$ to evaluate modeling results and serve as a base for decision-making.
+A more elaborate respective loss function could be customized on the base of surveys, acquiring the specific preferences of one or more companies and thus attaining a better profile of the economic environment, as well as the individuals acting in it.
+Although the algorithms for automatic 3D hydrocarbon trap recognition were developed to fit the the model in this work, they presumably could be easily adapted to other structural cases or even engineered further to be universally applicable.
+	Furthermore, additional and different uncertainty parameters should be considered.
+A respective next step regarding this model would be the incorporation of uncertainties which to a wider extent affect structural shapes in all three dimension.
+Otherwise, non-structural reservoir parameters could be included as uncertain, in particular those which are part of the $ROV$ equation.
+	Hydrocarbon systems and petroleum exploration as a sector were chosen for an exemplary application in this work, but other settings can be found, in which geological modeling is of central significance for decision-making.
+Related to the context and results of this work would be subsurface storage of fluids in reservoir, such as Carbon Capture and Storage.
+Questions of storage capacity and safety deal with similar conditions and geological problems, as the ones presented in this here, most importantly seal reliability and the risk of leakage.
+In respect to this, the models and approaches in this work might provide a base to raise new cases that can be analyzed.

### 2017-11-20 Risk analysis/Text/Layout/Layout/MSc_layout_Fabian.tex
Restructuring cjhapters
Layout changes
Some text progress (methods)
Uncertainty parameters seem good!
To Do: Implement return of trap control mechanism:
Leak - Spill - Juxta... maybe SFF/Throw Likelihood?
More models running. Updated text and figures
Working on 3D results text
Figures in Appendix. Next: Discussion!
Continued discussion text (1 page)
Text progress

@@
 \usepackage{subcaption}
+\usepackage{wrapfig}
+\usepackage{booktabs}
 \graphicspath{Figures/}
@@
 \mscDate{\today}
-\mscTitle{Test}
-\mscSubTitle{Subtitle}
+\mscTitle{Bayesian Decision Theory in Structural Geological Modeling}
+\mscSubTitle{How Reducing Uncertainties Affects Reservoir Value Estimations}
 \mscKeyWords{thesis, msc, subject}
@@
 	\nonumchap{Abstract}
+Bayesian methods are an intuitive approach to inference and subsequent decision-making, relatable to a wide spectrum of problems.
+Hydrocarbon exploration and production is a high-risk, high-reward sector in which good decision-making is indispensable.
+However, actors in this field are faced by numerous uncertainties that have to be considered.
+Structural geological modeling is of central importance for the estimation of uncertain hydrocarbon accumulations in potential reservoirs.
+Recent developments have shown, that such modeling can be regarded as a Bayesian inference problem.
+This allows for the incorporation of additional geological information as likelihood functions in a probabilistic framework and a consequent reduction of model uncertainties.
 
+---
 Please pay particular attention to the preparation of your abstract; use this text as a guide.
@@
 % Introduction
-\include{chapter1} % notice how the reference to the introduction takes place.. It refers to the name.tex
+%\include{chapter1} % notice how the reference to the introduction takes place.. It refers to the name.tex
 
 \cleardoublepage
-
+		\include{chapter1}
 		\include{chapter2_methods}
-		\include{chapter3_1Dcase}
-		\include{chapter4_3Dcase}
+		\include{chapter3_results}
+		\include{chapter4_discussion}
+		%\include{chapter5_conclusion}
+		%\include{chapter3_1Dcase}
+		%\include{chapter4_3Dcase}
 
@@
     \section{An appendix section}
+
+    \begin{figure}[h]
+    	\centering
+    	\includegraphics[width=0.62\textwidth]{Figures/Appendix/CSV_input1}
+    	\caption{....}\label{fig:CSV_input1}
+    \end{figure}
+
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=0.62\textwidth]{Figures/Appendix/CSV_input2}
+		\caption{....}\label{fig:CSV_input2}
+	\end{figure}
+
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=1\textwidth]{Figures/Appendix/CSV_input3}
+		\caption{....}\label{fig:CSV_input3}
+	\end{figure}
+
+    \begin{figure}[p!]
+    	\centering
+    	\includegraphics[width=1\textwidth]{Figures/Appendix/ML4}
+    	\caption{....}\label{fig:ML4}
+    \end{figure}
 

### 2017-11-20 Risk analysis/Text/Layout/Layout/chapter3_results.tex
Text progress

@@
 Several representative examples of posterior models based on the use of different likelihoods are referenced to the original prior model.
-The following measures are central to the comparison of simulation results: (1) information entropy, (2) occurrence of trap control mechanisms, (3) resulting recoverable oil volumes and (3) consequent realization of expected losses and related Bayes actions (i.e.
+For all model simulations, 1000 sampling steps were conducted, with a burn-in phase of 50 iterations for MCMC sampling.
+The following measures are central to the comparison of results: (1) information entropy, (2) occurrence of trap control mechanisms, (3) resulting recoverable oil volumes and (3) consequent realization of expected losses and related Bayes actions (i.e.
 decisions).
@@
 This results in a bimodal characterization, with low probability of the $ROV$ to be found between zero and 3 million m$^3$.\\
-		Consequently, applying the custom loss function on this distribution results in divergent Bayes actions for the different risk-affine actors (see Figure \ref{fig:M_prior} (b)).
-Although all actors find their estimates to be positive, only the risk-neutral to risk-friendliest actors come close to the described positive mode of the distribution ($ROV > 3*10^6$ m$^3$).
+		Consequently, applying the custom loss function on this distribution resulted in divergent Bayes actions for the different risk-affine actors (see Figure \ref{fig:M_prior} (b)).
+Although all actors found their estimates to be positive, only the risk-neutral to risk-friendliest actors came close to the described positive mode of the distribution ($ROV > 3*10^6$ m$^3$).
 Risk-averse individuals bid on significantly lower estimates.
-It can also be observed that expected losses seem to be generally lower for risk-affine actors.
-The divergence of different Bayes actions presumably resulted from the bimodality of the $ROV$ probability distribution.
+It can also be observed that expected losses resulted to be generally lower for risk-affine actors.
+It can be presumed that the divergence of different Bayes actions was caused the bimodality of the $ROV$ probability distribution.
 
@@
 		\end{figure}		
-		For the simulation of this first posterior model, thickness likelihoods for the three central layers were included into the probabilistic model.
+		For the simulation of this first posterior model, thickness likelihoods for the three central layers were included in the probabilistic model.
 These were defined in a way that reinforced the probabilities of the relevant layers to be significantly thick (see Table \ref{tab:ML1_likelihoods}).
@@
 Information entropy was also decreased in the middle section of the hanging wall, but remained high in upper and lower parts, presumably due to the remaining influence of the offset uncertainty.\\
-		Regarding trap volume control (Figure \ref{fig:ML1} (c)), only spill point-related migration ("SPILL POINT") and cross-fault leakage due to juxtaposition with permeable layers below the seal in the hanging wall ("LEAK UNDER") remain relevant in the posterior model.
-As no occurrences of leakage through overlying layers was observed, the two other mechanisms increased in frequency, while staying proportionally the same to one another.
+		Regarding trap volume control (Figure \ref{fig:ML1} (c)), only spill point-related migration ("SPILL POINT") and cross-fault leakage due to juxtaposition with permeable layers below the seal in the hanging wall ("LEAK UNDER") remained relevant in the posterior model.
+As no occurrences of leakage through overlying layers were observed, the two other mechanisms increased in frequency, while staying proportionally the same to one another.
 It can be presumed, that by introducing the respective likelihoods, a high thickness of the seal was ensured, as can also be recognized in Figure \ref{fig:ML1} (b).
@@
 		For the second model, thickness likelihoods were defined in a way that reinforced the probability of a very thick seal but at the same time restricted the reservoir formation to a likely significantly thinner thickness of around 50 m (see Table in Appendix).
-The results (see Appendix) are very similar, in fact practically identical, to those of model I above.
+The results (see Figure \ref{fig:ML3}) are very similar, in fact almost identical, to those of model I above.
 This can be explained by the influence of leak and spill point.
@@
 It can furthermore be derived from this that the maximum trap volume potential, as defined by the given structural features, is already represented in the $ROV$ probability distribution of the prior model.
-Examining this in Figure \ref{fig:ML1} (d), it seems that this overall maximum volume is found at approximately 5.7 millon m$3$.
+After examining Figure \ref{fig:ML1} (d), it seems that this overall maximum volume is found at approximately 5.7 millon m$3$.
 Higher values were not observed in any other model realization and are presumed to be unattainable due to the restriction given by the spill point.
@@
 		\end{figure}
-		Results from this third model show that in contrast to the reservoir thickness, seal thickness is a primary and decisive factor for the manifestation of a reliable trap structure.
-Using likelihoods which favor a thick reservoir, but only a thin seal of about 50 m thickness, leakage due to juxtaposition with seal-overlying formations across the fault ("LEAK OVER") becomes the primary mechanism for trap volume control (see Figure \ref{fig:ML2} (c).
+		Using likelihoods which favor a thick reservoir, but only a thin seal of about 50 m thickness, leakage due to juxtaposition with seal-overlying formations across the fault ("LEAK OVER") became the primary mechanism for trap volume control (see Figure \ref{fig:ML2} (c).
 It can be recognized in Figure \ref{fig:ML2} (b) that, while uncertainty was reduced clearly in the footwall, it remains high in the hanging wall.
@@
 Thus, frequent surpassing of the critical $SSF$ and subsequent fault seal breaching were enabled.
-It is additionally apparent in Figure \ref{fig:ML2} (b), that while fault-related leakage has become predominant, control by the spill point was suppressed to be widely irrelevant in this posterior model.\\	%It is suggested by the footwall information entropy in Figure \ref{label} (b), that the general shape of the structure has been flattened.
-		A dramatically higher probability of complete trap failure is induced by the high frequency of cross-fault leakage to overlying formations.
-Consequently, the bulk of the posterior $ROV$ probability distribution is shifted to a volume of zero, as depicted in Figure \ref{fig:ML2} (d).
+It is additionally apparent in Figure \ref{fig:ML2} (b), that while fault-related leakage became predominant, control by the spill point was suppressed to be widely irrelevant in this posterior model.\\	%It is suggested by the footwall information entropy in Figure \ref{label} (b), that the general shape of the structure has been flattened.
+		A dramatically higher probability of complete trap failure was induced by the high frequency of cross-fault leakage to overlying formations.
+Consequently, the bulk of the posterior $ROV$ probability distribution was shifted to a volume of zero, as depicted in Figure \ref{fig:ML2} (d).
 Any positive trap volumes have become highly improbable.\\
 		Bayes actions changed accordingly.
-Given this posterior $ROV$ distribution, all actors find their minimum of expected loss at $ROV = 0$, which equals the decision to "take no action" (see Figure \ref{fig:ML2} (e)).
-It has to be pointed out that all Bayes actions are found at the same point due to the lack of negative values into which more risk-averse actors would presumably move, given the possibility.
-This is a difference to the score system implemented for the 1D geological model evaluation.
-Volumes cannot be negative, thus, Bayes actions cannot be shifted past the limit of $ROV = 0$.
-Compared to the prior model, expected losses were decreased overall, but are higher for more risk-friendly individuals at the minimum.
+Given this posterior $ROV$ distribution, all actors found their minimum of expected loss at $ROV = 0$, which equals the decision to "take no action" (see Figure \ref{fig:ML2} (e)).
+It has to be pointed out that all Bayes actions were found at the same point due to the lack of negative values into which more risk-averse actors would presumably have moved, given the possibility.
+This poses a difference to the score system implemented for the 1D geological model evaluation.
+As volumes cannot be negative, Bayes actions cannot be shifted past the limit of $ROV = 0$.
+Compared to the prior model, expected losses were decreased overall, but are higher for risk-friendlier individuals at the minimum.
+Results from this third model show that in contrast to the reservoir thickness, seal thickness is a primary and decisive factor for the manifestation of a reliable trap structure.
 
@@
 		In Figure \ref{ML4} (b), it can be seen that the uncertainty about the position of interfaces in the footwall was greatly reduced for all layers.
-In the hanging wall, information entropies remained high but were shifted upwards and mainly concentrated in an area that is juxtaposed to the footwall seal formation.\\
+In the hanging wall, information entropy remained high but was shifted upwards and mainly concentrated in an area that is juxtaposed to the footwall seal formation.\\
 		In the resulting $ROV$ distribution, the possibility of trap failure was completely extinguished, while the probability of high volumes in the range between 3.5  and 5 million m$^3$ was significantly increased.
-Mean and median are found at almost the same value, but the bulk of the distribution found in this range appears to be slightly bimodal in itself, resembling approximately the shape of two normal distributions overlapping around a value of 4 million m$m^3$.\\
+Mean and median are found at almost the same value, but the bulk of the distribution found in this range appears to be slightly bimodal in itself, resembling approximately the shape of two normal distributions overlapping around the value of 4 million m$^3$.\\
 		Accordingly, Bayes actions converged within this range of positive values.
@@
 This value is furthermore characterized by a complex interdependence with other parameters and is most directly related to layer thicknesses and fault offset.
-For this reason, information entropy and offset uncertainty in the hanging wall was not greatly reduced.
-However, the certainty about the ratio of displacement to layer thicknesses was improved by using the $SSF$ in a likelihood function.
+For this reason, information entropy and offset uncertainty in the hanging wall were not greatly reduced.
+However, the certainty about the ratio of displacement to layer thicknesses was presumably improved by using the $SSF$ in a likelihood function.
 
@@
 Furthermore, an increase in the relevance of cross-fault leakage to overlying formations ("LEAK OVER") as a trap volume mechanism (increase from $\sim13\%$ to $\sim18\%$) was observed (Figure \ref{fig:ML5} (c)).
-Consequently, the bimodality between the bulk of positive values and the probability of complete failure in the $ROV$ distribution was apparently reinforced.
+Consequently, the bimodal character between the bulk of positive values and the probability of complete failure in the $ROV$ distribution was reinforced.
 This is also indicated by a shift of mean and median to lower volumes (Figure \ref{fig:ML5} (d)).
@@
 a greater separation of different Bayes actions, was induced (Figure \ref{fig:ML5} (e)).
-Expected losses of the minima were raised as well.\\
-		It is notable that the implementation of a likelihood which can be expected to significantly decrease the $SSF$ related uncertainty, overall leads to an amplification of the duality in the $ROV$ probability distribution and to dispersion of different risk-affine decisions.
-This effect is presumably caused by conservation of decisive uncertainty between "success" and complete failure of the trap, by narrowing the probability of the $SSF$ to be around its critical value.
+Expected losses of these minima were raised as well.\\
+		It is notable that the implementation of a likelihood function, which can be expected to significantly decrease the $SSF$ related uncertainty, overall leads to an amplification of the duality in the $ROV$ probability distribution and to dispersion of differently risk-affine decisions.
+This effect is presumably caused by conservation of decisive uncertainty between "success" and complete failure of the trap, due to narrowing of the $SSF$ probability around its critical value.

### 2017-11-20 Risk analysis/Text/Layout/Layout/chapter4_discussion.tex
Text progress

@@
 	\chapter{Discussion}\label{cha:discussion}
-	Building upon the recent findings by \citet{delaVarga2016}, showing that structural modeling can be viewed as a problem of Bayesian inference, the aim of this work was to extend this approach by considering practical applications and the utility geological modeling might have in an economic context.
+	Building upon the recent findings by \citet{delaVarga2016}, who showed that structural modeling can be viewed as a problem of Bayesian inference, the aim of this work was to extend this approach by considering practical applications and the utility geological modeling combined with Bayesian updating might have in an economic context.
 A sector in which structural geological modeling is of central importance is hydrocarbon exploration.
 This field is characterized by the necessity to make decisions in the face of high risks and potentially high rewards.
-As these decisions are often closely linked to geological modeling and the estimation of reservoir-related values, this is work aimed to use this context to extend the Bayesian inference step in geological modeling by evaluating its results in terms of influence on respective decision-making.
-It was initially hypothesized that Bayesian updating via likelihoods and the resulting change of uncertainty in a model, should have significant effects on subsequent decision-making.\\
+As these decisions are often closely linked to geological modeling and the estimation of reservoir-related values, we aimed to use this context to extend the Bayesian inference step by considering its influence on respective decision-making.
+To observe this, a case-specific loss function was developed in this work.
+In doing so, we intend to furthermore show, that the loss function estimation approach is suitable and useful to describe varying degrees of complexities behind decision-making environments and risk-related behavior of the actors in it.
+	% It was initially hypothesized that Bayesian updating via likelihoods and the resulting change of uncertainty in a model should have significant effects on subsequent decision-making.\\
 	To analyze this, geological models were interpreted as potential hydrocarbon systems.
-Algorithms for structural trap recognition and calculation of values relevant to respective decision-making were developed.
-Based on many iterations, probability distributions for such economic values (score and $ROV$, respectively) were generated.
-A custom loss function was designed to express the environment in which such values are to be estimated and to represent preferences of differently risk-affine decision-makers.
+A central part of our work was comprised of the development of algorithms capable of automatic structural trap recognition and subsequent calculation of values relevant to decision-making.
+Different prior and posterior probability distributions for such economic values were generated to form a basis for application of the custom loss function for value estimation.
 These methods were applied for a conceptual 1D geological model and subsequently for a 3D structural model.
-Despite the great leap in complexity, the results from both models are widely similar.\\
-	Scores in the 1D model and $ROV$ in the 3D model were chosen to represent a measure for the economic value realized in each model construction.
-As such, they were defined to depend on model parameters, either directly or indirectly by valuing their relations to other parameters or specific thresholds and conditions.
-Accordingly, their posterior probability distributions and respective uncertainties changed with the implementation of Bayesian inference.\\
-	These changes related mostly to main modes in the distributions.
+	Despite a significant leap in complexity, the results from both models proved to be widely similar.\\
+
+	The economic value realized in each model construction (score and $ROV$, respectively), had been defined to depend on model parameters, either directly or indirectly by taking into account their relations to other parameters or to specific thresholds and conditions.
+This meant in general, that in this step of valuation, model uncertainties were in a way translated and reduced to the probability distribution of a single decisive value.
+As this was set to be the foundation for loss function estimation, we consider this value distribution to be an expression of the state of knowledge on which the decision-making is to be based.
+Thereby, it can furthermore be proposed, that the overall uncertainty inherent in this probability distribution can be referred to as "decision uncertainty" and that this entity is to be viewed separately from the model uncertainty.\\
+
+	This state of knowledge and the decision uncertainty accordingly experienced alterations related to the different Bayesian inference cases.
+These changes related foremost to modes in the value distribution.
 For both model designs (1D and 3D), the prior value probability distributions were characterized by a distinct bimodality formed by widely separated modes of opposing values, i.e.
-one in the range of highly postive values, another at zero or negative values.
+one in the range of highly positive values, another at zero or negative values.
 By introducing likelihood functions, these modes were most often shifted or altered in their shape.
 The latter was observed as either an amplification of the mode, by raising its mean probability, while narrowing the range of probabilities (i.e.
-reducing the standard deviation), or the opposite, a diminishment of the mode up to total erasure of its probability.
+reducing the standard deviation), or the opposite, a diminishment of the mode up to total erasure of its probability.\\
 	Considering that these mode changes correlated with increased or decreased relevance of specific trap control mechanisms in the 3D model, it can be argued that each mode represents the occurrence of a specific mechanism-related scenario.
 These are mostly distinctly separated with little probability in between.
-This presumably is based on the assumption that traps are "filled-to-spill" in respect to spill and leak point, and that full leakage is assumed for any type of seal failure, resulting in an approximate discretization into respective scenarios.\\
-	Bayesian inference primarily seems to lead to a change in uncertainty in respect to these scenarios or modes.
-The proportionality regarding these modes, their relative probabilities and distance to each other, in turn influence the overall uncertainty of the probability distribution as a whole.
+This presumably is based on the assumption that traps are always "filled-to-spill" in respect to spill and leak point, and that full leakage is assumed for any type of seal failure, resulting in an approximate discretization into respective scenarios.\\
+	From a decision-making perspective, these scenarios might be interpreted as different decision options.
+In a case of distinct bimodality as seen in the prior of the 3D model, two main options would be given: (1) to not take any action, or (2) to expect a positive value in the range of the positive mode and invest accordingly.
+However, most often, low probabilities for the values between these two opposite options remain.
+Especially considering the use of loss functions which lead to best estimates located in these areas of low probability, it is debatable, how these ranges should be treated.
+A possibility could be to subdivide the value probability distribution accordingly into regions.
+Regarding respective integral probabilities for each such section and by introducing a critical integral probability as a threshold, it could be determined whether such a subregion is to be considered an actual option for decision-making or not.
+If a Bayes action is located outside of such a decision option, it could be automatically assigned to the closest region of sufficient integral probability.
+This approach should be easily applicable as an extension to our custom loss function.
+Furthermore, it could serve to build a bridge between this continuous loss function method and commonly used discrete approaches to decision-making, such as decision trees.
+In real cases, normally only a limited number of options is given.
+In the context of hydrocarbon exploration and production, this would relate to fixed magnitudes of resource allocation, such as a certain number of required drilling wells or the size of a production platform.
+A discretization of decision subregions in a probability distribution could thus also be based on previously defined actual options.\\
+
+	The proportionality regarding modes, their relative probabilities and distances to each other, in turn seemed to influence the overall uncertainty of the probability distribution, i.e.
+the decision uncertainty.
 This was often indicated by the position of the mean and median relative to the modes.
-A mean located in the middle between two widely separated modes, which might be of low intrinsic uncertainty themselves, would indicate a much higher overall uncertainty, than a mean found in the middle of a narrow unimodal distribution.
-	As the process of decision-making is based on these value distributions by applying the custom loss function respectively, it is proposed here, that this overall uncertainty might be referred to as "decision uncertainty".\\
+A mean located in the middle between two widely separated modes, which might be of low intrinsic uncertainty themselves, would indicate a much higher overall uncertainty, than a mean found in the middle of a narrow unimodal distribution.\\
+
 	The decision-making of differently risk-affine actors, i.e.
-the position and spread of Bayes actions, changed relative to the properties of value probability distributions and their inherent decision uncertainty.
+the position and spread of Bayes actions, manifest according to the properties of value probability distributions and their inherent decision uncertainty.
 Distinct bimodality between two extremes, i.e.
@@
 Reduction of the distribution to one mode conversely led to the convergence of different Bayes actions.
-A decrease decision uncertainty furthermore was accompanied by a reduction in expected loss for each decision.
-Risk-friendly actors were the most robust in their decision-making in the face of possible trap failure.
-Eliminating this risk proved to be far less significant to the most risk-friendly actor.
-The respective estimate was generally found close to the positive mode and experienced only minor lateral shifting, after diminishing the probability of failure.
+A decrease in decision uncertainty furthermore was accompanied by a reduction in expected loss for each Bayes estimator.
+Risk-friendly actors were the most robust in their decision-making, particularly in the face of possible trap failure.
+Eliminating this risk proved to be far less significant to the most risk-friendly, than for risk-averse actors.
+The respective estimate resilient to risk was generally found close to the positive mode and experienced only minor lateral shifting, after diminishing the probability of failure.
 Greater shifts were caused for this actor, in cases in which the positive mode itself was shifted laterally or erased completely.
-Conversely, the Bayes action of the most-averse individual was influenced greatly by the probability of the "worst scenario".
+Conversely, the Bayes action of the most-averse individual was influenced greatly by the probability of the "worst case" scenario, complete trap failure.
 Only a stark decrease of the respective mode encouraged this actor to bid on an estimate close to the risk-friendlier decision-makers.
-These observations match the principles and conditions conceived in the design of the loss function and the applied loss function.
+These observations match the principles and conditions conceived in the design of the loss function and the chosen range of risk factors.
 A clear difference in the decision-making relative to risk-affinity was observed.\\
-	It can be argued that the degree of convergence of Bayes actions and their expected loss can be considered measures for the state of knowledge during the decision-making process.
-The better this is, the more similar the decisions of differently risk-affine actors are.
-It can be assumed that given perfect information, all actors would bid on the same estimate, the true value, and expect no loss.
-Thus, the relevance of the risk-factor decreases with higher reduction of decision uncertainty.\\
-	While in most cases the inclusion of additional information as likelihoods led to a reduction of decision uncertainty, this was not always the case.
+
+	Considering the nature of changes experienced by the different Bayes action, depending on variations in the underlying reservoir value distributions, we derive that the degree of convergence of Bayes actions and their expected losses can be considered measures for the state of knowledge during the decision-making process.
+The better this is, the more similar the decisions of differently risk-affine actors and the lower their loss expectations are.
+It can be assumed, that given perfect information, all actors would bid on the same estimate, the true value, and expect no loss.
+It follows from this that the relevance of risk factors decreases with higher reduction of decision uncertainty.\\
+
+	While in most cases Bayesian inference led to a reduction of decision uncertainty, this was not always the case.
 It appears that the change in decision uncertainty is not necessarily strictly aligned with the change in uncertainty regarding model parameters and their derivatives.
 In respect to this, it seems to be of central importance "where" in the model uncertainty is reduced, i.e.
-in which spatial area or regarding which model parameter.
+in which spatial area or regarding which model parameters.
 This appears to be of particular significance considering threshold values that lead to an abrupt cut-off between two extrema of decision options.
-In both types of model construction, 1D and 3D, this is directly related to sealing reliability.
-Thresholds regarding seal thickness (1D model) and Shale Smear Factor ($SSF_c$ in 3D model) were defined in a way that introduced a significant possibility of complete failure of a trap.
-Consequently, it was observed that reducing the uncertainty in a way that narrows the probability of a threshold-related parameter around its cut-off value, can lead to an amplification of the respective mode and thus emphasize the risk of complete failure.
-This can well be seen in the posterior $ROV$ distribution of model V (3D).
-Resulting Bayes actions were characterized by an increase in divergence and expected losses.\\
-	Reducing model uncertainty in the "wrong" areas seems to simply lead to a transformation of this uncertainty in the realm of decision-making.
-Overall uncertainty seems to be conserved to a certain extent or even increased, as the duality of the decision problem is amplified in the form of a more stretched-out bimodal distribution and diverging Bayes actions.
-According to this, it should be of foremost importance for each actor, to reduce the uncertainty regarding threshold-related factors which might decide between "success" and complete failure of a project.
-Some type of information might improve the potential magnitude of a positive outcome, but maintain the risk of failure.
+In both types of model construction, 1D and 3D, this was directly related to sealing reliability.
+Thresholds regarding seal thickness (1D model) and Shale Smear Factor ($SSF_c$ in 3D model) were defined in a way that introduced a significant possibility of complete trap failure.
+Consequently, it was observed that reducing the uncertainty in a way that narrowed the probability of a threshold-related parameter around its cut-off value, led to an amplification of the respective mode and thus emphasized the risk of complete failure.
+This can be seen in the posterior $ROV$ distribution of model V (3D).
+Resulting Bayes actions were characterized by an increase in divergence and expected losses, indicating a deterioration of the state of knowledge and an according increase in decision uncertainty.
+Thus, reducing model uncertainty in the "wrong" areas seems to simply lead to a transformation of this uncertainty in the realm of decision-making.\\
+	%The decision uncertainty seems to be conserved to a certain extent or even increased, as the duality of the decision problem is amplified in the form of a more stretched-out bimodal distribution and diverging Bayes actions.
+	According to this, it should be of foremost importance, especially for risk-averse actors, to reduce the uncertainty regarding threshold-related factors which might decide between the success and complete failure of a project.
+In the context of hydrocarbon traps, seal reliability is decisive.
 For making better decisions, elimination of such high risks should be a priority.
-In the context of hydrocarbon traps, seal reliability is decisive.\\
-	Such risks might be easily assessed, if they are dependent on only one or a few parameters, such as seal thickness in the 1D model.
-In other cases, they are derived from more complex parameter inter-relations, as is the case for the Shale Smear Factor, the ratio of displacement to seal thickness.
+As pointed out above, this might be less relevant for risk-friendly decisions makers, who respectively might acquire a comparable benefit from knowing more about the probability of positive outcomes, as they are less discouraged from overestimating and less afraid of failure.
+Nevertheless, it is suggested by the results from the different inference cases, that the best states of knowledge are attained by eliminating bimodality and narrowing the range of probabilities to one mode anywhere along the value axis.\\
+
+	Crucial risks might be easily assessed, if they are dependent on only one or a few parameters, such as seal thickness in the 1D model.
+In other cases, they are derived from more complex parameter inter-relations, as is the case for the Shale Smear Factor, which is the ratio of displacement to seal thickness.
 Models IV and V (3D) showed that reduced uncertainty about the $SSF$ is not necessarily directly recognizable by information entropy visualization, especially if uncertainty regarding its parent parameters remains high.
-It follows that, to approach an effective mitigation of high risks, the complexities behind decisive factors need to be assessed and respective parent parameters and inter-dependencies need to be recognized.
-This might enable a better understanding of which type of information is missing and where in the model, additional data might be of use for improved decision-making.\\
+It follows that, to approach an effective mitigation of high risks, the complexities behind decisive factors need to be assessed thoroughly and respective parent parameters, as well as their interdependencies need to be recognized.
+This might enable a better understanding of which type of information is missing and where in the model additional data might be of use for improved decision-making.\\
+
+	A central conclusion from these observations seems to be, that more of simply any type of information does not necessarily lead to better decisions.
+Instead, improvement of decision-making is achieved by attaining the right kind of information that is able to shed light on uncertainties which are relevant to an individuals own goals and preferences.
+These findings pair well with statements made by \citet{bratvold2010making}, who emphasized that the purpose of models in the hydrocarbon sector is to attain insights to improve decisions.
+In their book, they point out that value is not generated by uncertainty quantification or reduction in itself, but is created to the extent that these processes have potential to change a decision.
+This is reflected by our results which showed that one set of information might have a vastly different impact on the Bayes estimators of different actors.
+Assuming discretized options, new observations might lead to a change in decision for one actor, but not for the other.
 
@@
 The fact that similar observations were made for the 1D, as well as the 3D model, indicates a certain degree of continuity with respect to representativity.
-However, it also has to be pointed out, that the uncertainties employed in the 3D model regarded positional values in the $z$-axis only, and were of one-dimensional nature.
-This might party be reason for the parallelism in the results.
-There was thereby also no real uncertainty concerning the overall structural shape, particularly anticlinal features and the position of the spill point in relation to the trap maximum.\\
-
-	On algorithms
+However, it also has to be pointed out, that the uncertainties employed in the 3D model related to $z$-axis positional values only, and were thus of mostly  one-dimensional nature.
+This may in part explain the parallelism compared to the 1D model results.
+Furthermore, it follows that no effective uncertainty concerning the overall structural shape was implemented, particularly regarding anticlinal features and the position of the spill point in relation to the trap maximum.\\
 
@@
 As mentioned above, despite its simplicity, the design of the function and the incorporation of a risk factor appears to have been suitable to achieve a clear distinction between differently risk-affine actors and respective behavior.
-In contrast to this, the use of standard loss function would have only returned mean or median as estimates.
-Customization of such loss functions offers a flexible approach to express different objective and subjective weighting factors that are specified to certain decision-making environments and decision-makers' perspectives.
-It is thereby possible to extend this design to attain very complex loss functions.
+It was stated by \citet{bratvold2010making} that uncertainty has to possible consequences: risk and opportunity.
+According to them, risk is a consequence which is subjectively undesirable to an actor, while opportunity is the opposite.
+In our loss function, this seems to be well represented by varying affinities to rather over- or underestimate.
+For our risk-averse actors, a major risk is posed by seal failure.
+To our risk-friendliest individual, missing out on opportunity seems to be a greater risk.
+While standard loss functions would have only returned mean or median as estimates, loss function customization offers a flexible approach to express different objective and subjective weighting factors that are specified to certain decision-making environments and decision-makers' perspectives.\\
+	It would be possible to extend this design to attain very complex loss functions.
 As the loss function in this work was designed based on only a few basic assumptions, it might be more generally representative than a complex function which takes into account numerous very specific aspects.
@@
 It would be of interest to apply the same overall concept and methods on an authentic case based on real datasets.
-Given a realistic economic scenario, including capital and operational expenditures of a project, possibly a full net-present-value ($NPV$) analysis could be conducted.
-Recoverable reserves could be replaced by the $NPV$ to evaluate modeling results and serve as a base for decision-making.
+Given a realistic economic scenario, including capital and operational expenditures of a project, possibly a full net-present-value (NPV) analysis could be conducted.
+Recoverable reserves could be replaced by the NPV to evaluate modeling results and serve as a base for decision-making.
 A more elaborate respective loss function could be customized on the base of surveys, acquiring the specific preferences of one or more companies and thus attaining a better profile of the economic environment, as well as the individuals acting in it.
-Although the algorithms for automatic 3D hydrocarbon trap recognition were developed to fit the the model in this work, they presumably could be easily adapted to other structural cases or even engineered further to be universally applicable.
-	Furthermore, additional and different uncertainty parameters should be considered.
-A respective next step regarding this model would be the incorporation of uncertainties which to a wider extent affect structural shapes in all three dimension.
-Otherwise, non-structural reservoir parameters could be included as uncertain, in particular those which are part of the $ROV$ equation.
+Value of information could be considered as a parameter to better quantify the impact of Bayesian inference as observed in this work.
+	Although the algorithms for automatic 3D hydrocarbon trap recognition were developed to fit the primarily model in this work, we presume that they could easily be adapted to other structural cases or even engineered further to be universally applicable.\\
+	Furthermore, additional and different uncertainty parameters should be considered in the future.
+A respective next step regarding our model would be the incorporation of uncertainties which to a wider extent affect structural shapes in all three dimensions.
+Otherwise, non-structural reservoir parameters could be included as uncertain values, such as porosities and permeabilities of different layers.
+This might be of particular interest considering factors which are part of the $ROV$ equation or other aspects that related to high risks in decision-making.\\
 	Hydrocarbon systems and petroleum exploration as a sector were chosen for an exemplary application in this work, but other settings can be found, in which geological modeling is of central significance for decision-making.
-Related to the context and results of this work would be subsurface storage of fluids in reservoir, such as Carbon Capture and Storage.
-Questions of storage capacity and safety deal with similar conditions and geological problems, as the ones presented in this here, most importantly seal reliability and the risk of leakage.
-In respect to this, the models and approaches in this work might provide a base to raise new cases that can be analyzed.
+One example would be subsurface storage of fluids in a reservoir, such as carbon capture and storage (CCS).
+Questions about storage capacity and safety deal with similar conditions and geological problems, as the ones presented this work, most importantly seal reliability and the risk of leakage.
+In respect to this, the models and approaches in this work might provide a basis to raise new and advanced explorable cases.

### 2017-11-20 Risk analysis/Text/Layout/Layout/chapter1.tex
Enough for today

@@
 Individuals are constantly faced with situations in which a decision has to be made, but only incomplete information is available.
-Such a problem necessitates an approach based on plausible reasoning, one which is intuitively structured in five stages \citep{jaynes1986bayesian}:
+Such a problem necessitates an approach based on plausible reasoning, one which is intuitively structured in four stages \citep{jaynes1986bayesian}:
 \begin{enumerate}
@@
 Bayesian decision-theory is to be included in the step of model evaluation.
-This is achieved by assigning an economic meaning to the structural model and designing a case-specific custom loss functions to find decisions which are optimal related to the state of knowledge and the preferences of actor's with different risk-affinities.
+This is achieved by assigning an economic meaning to the structural model and designing a case-specific custom loss functions to find decisions which are optimal related to the state of knowledge and the preferences of actors with different risk-affinities.
 More specifically, the models are designed to represent potential petroleum systems.

### 2017-11-20 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Enough for today

@@
         \subsection{Bayesian inference}
-        Bayesian inference is most importantly characterized by its preservation of uncertainty, in contrast to standard statistical inference \citep{davidson2015}.
+        Bayesian inference is most importantly characterized by its preservation of uncertainty, in contrast to standard statistical inference \citep{box2011bayesian, jaynes2003probability, harney2013bayesian, davidson2015}.
+The Bayesian approach is widely seen as intuitive and inherent in the natural human perspective.
 Probability is seen as a measure of belief for an event to occur.
-It has been argued by \cite{davidson2015}  that this Bayesian approach is intuitive and inherent in the natural human perspective.
 These beliefs can be assigned to individuals.
-Thus, different and even contradicting beliefs about the probability of an event might be held by different individuals, based on variations and disparities in the information available to each one individual \citep{davidson2015}.\\
-        The initial belief or guess about an event $\theta$ can be denoted as $p(\theta)$ \citep{davidson2015, delaVarga2016}.
+Thus, different and even contradicting beliefs about the probability of an event might be held by different individuals, based on variations and disparities in the information available to each one individual.\\
+        The initial belief or guess about an event $\theta$ can be denoted as $p(\theta)$.
 This is used as the so-called prior probability on which Bayesian updating is based.
@@
 These observations are included as likelihoods $p(y|\theta)$.
-This process of updating results in a posterior probability $p(\theta|y)$ \citep{davidson2015, delaVarga2016}.
-It is important to note that the prior is not simply discarded but re-weighted by Bayesian updating.
-It was also pointed out by \citet{davidson2015} that by utilizing an uncertain prior, the potential for wrongfulness of the initial guess is already included.
-This means that Bayesian updating is about reducing uncertainty in a belief and reaching a guess that is less wrong \citep{davidson2015}.
-        Bayesian inference is defined by and conducted via the following equation, called the Bayes' Theorem \citep{davidson2015}:
+By conducting this process of inference, a posterior probability $p(\theta|y)$ is attained.
+It is important to note that the prior is not simply discarded but re-weighted by Bayesian inference.
+By utilizing an uncertain prior, the potential for wrongfulness of the initial guess is already included.
+This means that Bayesian updating is about reducing uncertainty in a belief and reaching a guess that is less wrong.
+Bayesian inference is defined by and conducted via the following equation, called the Bayes' Theorem \citep{box2011bayesian, jaynes2003probability, harney2013bayesian, davidson2015}:
         \begin{equation}\label{eq:BayesTheorem}
@@
         Loss is a statistical measure of how "bad" an estimate is, i.e. how much is lost by making a certain decision. Gains are considered by statisticians as negative losses.
-        The magnitude of an estimate's loss is defined by a loss function, which is a function of the estimate of the parameter and the true value of the parameter \citep{davidson2015}:        
+        The magnitude of incurred loss related to an estimate is defined by a loss function, which is a function of the estimate of the parameter and the true value of the parameter \citep{wald1950statistical, davidson2015}:
         \begin{equation}\label{eq:LossFunction}
@@
 all possible true values are represented by this distribution \citep{davidson2015}.
-A suitable alternative to the actual loss is to consider each decision's expected loss and to make a decision that is optimal in relation to this expected loss \citep{berger2013stat}.
+A probabilistic alternative to the actual loss is to consider each decision's expected loss and to make a decision that is optimal in relation to this expected loss \citep{berger2013stat}.
 \\
@@
         The expectation symbol $E$ is subscripted with $\theta$, by which it is indicated that $\theta$ is the respective unknown variable.
-This expected loss is also referred to as the risk of estimate $\hat{\theta}$ \citep{davidson2015}.\\
+This expected loss $l$ is also referred to as the risk of estimate $\hat{\theta}$ \citep{davidson2015}.\\
         By the Law of Large Numbers, the expected loss of $\hat{\theta}$ can be approximated drawing a large sample size $N$ from the posterior distribution, respectively applying a loss function $L$ and averaging over the number of samples \citep{davidson2015}:
@@
         	where $\phi_i$ is a function of position $\vec{x}$, more precisely an interpolation function, to which respective additional parameters are given by $\alpha_k$.
-Available primary geological information, such as positions and dips of layer interfaces and in the subsurface, is represented by $k_j$.
+Available primary geological information, such as positions and dips of layer interfaces, is represented by $k_j$.
 A topological description, such as the relationships between faults and layer surfaces, is given by $\beta_l$.
@@
         \end{equation}
-        This total entropy $H_T$ is 0 when no uncertainty is found in any cell, oppositely it is maximal in the case of equal likelihood of outcoms in every cell \citep{wellmann2012uncertainties, schaaf2017}.
+        This total entropy $H_T$ is 0 when no uncertainty is found in any cell, oppositely it is maximal in the case of equal likelihood of outcomes in every cell \citep{wellmann2012uncertainties, schaaf2017}.
 The application of these entropy measures allows for the calculation and visualization of uncertainties in each voxel, the assessment of uncertainties of entire geological units, as well as the quantification of total model uncertainty represented by a single number \citep{wellmann2012uncertainties}.
@@
 Fundamentally different loss functions can also be based on a significantly different mathematical structure.
-Loss functions are customized regarding the problem environment and according to the to subjective needs and objectives of decision-maker.
+Loss functions are customized regarding the problem environment and according to the subjective needs and objectives of the decision-maker.
 Thus, they are mostly defined by the actor expressing his perspective.
-Changes in the individual's perception and attitude might lead to further customization needs at a future point in time.
+Changes in the individual's perception and attitude might lead to further customization needs at a future point in time, as was reported by \citet{hennig2007}.
 Especially considering individual persons as actors, psychological aspects may play a significant role.
@@
 The resulting model with its prior distributions of possible layer boundary locations is illustrated in Figure \ref{fig:1D_model}.
-			\begin{figure}[h]
-			\centering
-			\includegraphics[width=1\textwidth]{Figures/1D_principle.PNG}
-			\caption{Conceptual illustration of applying Bayesian inference in an abstract 1D geological model as conducted by \citet{delaVarga2016}.
+			%\begin{figure}[h]
+			%\centering
+			%\includegraphics[width=1\textwidth]{Figures/1D_principle.PNG}
+			%\caption{Conceptual illustration of applying Bayesian inference in an abstract 1D geological model as conducted by \citet{delaVarga2016}.
 Prior distributions of layer boundaries in depth (a) are updated using thickness likelihoods (b) to attain posterior distributions (c).
 Model parameters which are uncorrelated in their joint probability distribution before the inference (d), are correlated after performing Bayesian inference (e) (from \citet{delaVarga2016}).}\label{fig:1D_principle}
-			\end{figure}
-			\subsection{Bayesian inference and analysis of the 1D case}\label{sec:1D_bayes}
-			For conducting Bayesian inference, it is assumed that new observations are made, providing additional information on the likelihoods of the thicknesses of the two layers of interest.
+			%\end{figure}
+			%\subsection{Bayesian inference and analysis of the 1D case}\label{sec:1D_bayes}
+			For conducting Bayesian inference, it is assumed that new observations are made, providing additional information considered here in the form of likelihood functions for the thicknesses of the two layers of interest.
 Likelihood functions for reservoir and seal thicknesses are introduced based on another set of normal probability distributions.
-These are directly related to the layer boundary positions, i.e.
-the prior parameter distributions.
+These are directly related to the layer boundary positions, i.e.\ the prior parameter distributions.
 Means and standard deviations of the likelihood distributions are variable, so that several scenarios based on the assumption of different observations can be tested.
 Parameter priors and likelihood functions are embedded into a probabilistic modeling framework as described in Section \ref{sec:numerical_implementation} and MCMC sampling is conducted accordingly.
-Performing Bayesian inference on such a 1D geological model follows the conceptual example devised by \citet{delaVarga2016} (see Figure \ref{fig:1D_principle}).
+Performing Bayesian inference on such a 1D geological model follows the conceptual example devised by \citet{delaVarga2016}.
 Posterior distributions are to be evaluated based on the valuation method described in the following.
@@
 All layers are tilted so that they dip in the opposite direction of the fault plane dip.
-The original concept of this model is designed in a way, that a potential hydrocarbon trap is formed in the reservoir rock enclosed by the deformed seal and the normal fault.\\
+The original concept of this model is designed in a way that a potential hydrocarbon trap is formed in the reservoir rock enclosed by the deformed seal and the normal fault.\\
 		This trap, more specifically the trap volume, is defined as the central feature of economic interest.

### 2017-11-21 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Improved discussion part

@@
 At X~=~1100~m and Y~=~1000~m, the point of measurement is located centrally in the overall model block, as well as the original potential trap structure and is thus presumed to be representative and of significance for the reservoir analysis.\\
-		After defining these thickness to point-positional dependencies, thickness likelihoods are implemented as probability distributions, as done for the 1D model.
+		After defining these thickness to point-positional dependencies, thickness likelihood functions are implemented based on probability distributions, as done for the 1D model.
 Various types of geological information, conclusions and expectations can be expressed by adapting mean and standard deviation of these normal distributions.
 For core sampling, for example, information might be evenly provided for all layers, irrespective of depth, so that similar degrees of standard deviation might be assigned to all layer thicknesses.
-Alternatively, taking into account certain depositional environments, for which high thickness variablity is characteristic, specific standard deviations can be raised accordingly.
-Some representative cases are presented in Chapter \ref{cha:results}.
+Alternatively, taking into account certain depositional environments for which high thickness variablity is characteristic, specific standard deviations can be raised accordingly.\\
+		We additionally consider a likelihood function related to the Shale Smear Factor.
+This function is based on a normal distribution as well.
+As the $SSF$ is a more complex derivate of parameter relations, it also more difficult to argue, how respective observations were sourced.
+Related knowledge about fault displacement and thicknesses in an area nearby, would come with respective likelihood functions for those parent parameters as well.
+Possibly, the additional information about the $SSF$ was acquired indirectly, for example by having tested the leakage potential of other faults in the same general region.
+In any case, the inclusion of our likelihood functions based on purely hypothetical assumptions and is intended to provide the opportunity, to explore the effects different types of information might have.
+While the thickness likelihood functions are dependent on input parameters directly, the implementation of the $SSF$ likelihood function requires a full computation of the model and extended algorithms of structural analysis.\\
+		Some representative cases of Bayesian inference applied on the 3D model are presented in Chapter \ref{cha:results}.
 
@@
 		According to the methods introduced in Sections \ref{sec:bayes} and \ref{sec:bayes_in_struc_modeling}, the 3D geological model is to be incorporated in a probabilistic modeling framework and analyzed from a Bayesian perspective.
-		First, a simple Monte Carlo error propagation run, using only priors, is conducted so that a base reference distribution for the original model is generated.
+		First, a simple Monte Carlo simulation is conducted using only priors, so that a base reference distribution for the original model is generated.
 Subsequently, Bayesian inference is conducted based on different assumptions concerning the additional observations made and by incorporating likelihoods which respectively are designed to reflect the nature of this auxiliary data.
 Posterior distributions, in particular resulting volume distributions, can then be examined and evaluated by applying the methods described in Section \ref{sec:model_evaluation}.
+Furthermore, for every simulation, the mechanisms which ultimately determined the trap closure in every iteration, such as the spill point and cross-fault leakage, are traced as trap volume controls factors.

### 2017-11-21 Risk analysis/Text/Layout/Layout/chapter4_discussion.tex
Improved discussion part

@@
 	\chapter{Discussion}\label{cha:discussion}
-	Building upon the recent findings by \citet{delaVarga2016}, who showed that structural modeling can be viewed as a problem of Bayesian inference, the aim of this work was to extend this approach by considering practical applications and the utility geological modeling combined with Bayesian updating might have in an economic context.
+	Building upon the recent findings by \citet{delaVarga2016}, the aim of this work was to extend their approach by considering practical applications and the utility structural geological modeling viewed as an Bayesian inference problem might have in an economic context.
 A sector in which structural geological modeling is of central importance is hydrocarbon exploration.
@@
 	Despite a significant leap in complexity, the results from both models proved to be widely similar.\\	
-
 	The economic value realized in each model construction (score and $ROV$, respectively), had been defined to depend on model parameters, either directly or indirectly by taking into account their relations to other parameters or to specific thresholds and conditions.
@@
 As this was set to be the foundation for loss function estimation, we consider this value distribution to be an expression of the state of knowledge on which the decision-making is to be based.
-Thereby, it can furthermore be proposed, that the overall uncertainty inherent in this probability distribution can be referred to as "decision uncertainty" and that this entity is to be viewed separately from the model uncertainty.\\
-
-	This state of knowledge and the decision uncertainty accordingly experienced alterations related to the different Bayesian inference cases.
+Thereby, we furthermore propose, that the overall uncertainty inherent in this probability distribution can be referred to as "decision uncertainty" and that this entity is to be viewed separately from the model uncertainty.\\
+	By viewing decision-making as a problem of optimizing a case-specific customized loss function based on such a state of knowledge and decision uncertainty, we were able to observe clear differences in the respective behavior of distinctly risk-affine actors.
+	The position and spread of their Bayes actions, i.e.
+their decision, manifested according to the properties of value probability distributions and their inherent decision uncertainty.
+These were in turn determined by the proportionality regarding distribution modes, their relative probabilities and distances to each other.
+This was often indicated by the position of mean and median relative to these modes.
+A mean located in the middle between two widely separated modes, which might be of low intrinsic uncertainty themselves, would indicate a much higher overall decision uncertainty, than a mean found in the middle of a narrow unimodal distribution.
+Pronounced bimodality between two extremes, i.e.
+high decision uncertainty, resulted in a higher spread of Bayes estimators.
+Reduction of the distribution to one mode conversely led to the convergence of different Bayes actions.
+A decrease in decision uncertainty furthermore was accompanied by a reduction in expected loss for each Bayes estimator.
+	Considering these observations, we derive that the degree of Bayes action convergence and respective expected losses can be considered measures for the state of knowledge and decision uncertainty at the moment of making a decision.
+The better this is, the more similar the decisions of differently risk-affine actors and the lower their loss expectations are.
+It can be assumed, that given perfect information, all actors would bid on the same estimate (the true value) and expect no loss, since no risk is present.
+It furthermore follows from this that the relevance of risk factors decreases with higher reduction of decision uncertainty.\\
+	Cases of various states of knowledge and decision uncertainties, i.e.
+different realizations of the value probability distribution, were attained by the implementation of Bayesian inference.
+While in most settings, inference led to a reduction of decision uncertainty, this was not always the case.
+We observed that the impact on decision uncertainty, induced by Bayesian inference, is not necessarily strictly aligned with the change in uncertainty regarding model parameters and their derivatives.
+In respect to this, two aspects seems to be of central importance: (1) "where" in the model uncertainty is reduced, i.e.
+in which spatial area or regarding which model parameters and (2) which possible outcome is enhanced in terms of probability.
+An increased probability of a thick or thin seal in our 3D model, for example, equally reduced decision uncertainty significantly, by raising the probability of a positive outcome or negative outcome, respectively.
+Improved certainty about our reservoir thickness, however, had little to no impact on decision-making.
+This shows us that some areas and parameters of the model have a much greater influence on the decision uncertainty than others.\\
+	Adding to this, we observed a particular effect related to threshold values that lead to an abrupt cut-off between two extrema of decision ranges.
+In both types of model construction, 1D and 3D, this was directly related to sealing reliability.
+Thresholds regarding seal thickness (1D model) and Shale Smear Factor ($SSF_c$ in 3D model) were defined in a way that introduced a significant possibility of complete trap failure.
+Consequently, it was observed that reducing the uncertainty in a way that narrowed the probability of a threshold-related parameter around its cut-off value, led to an amplification of the respective mode and thus emphasized the risk of complete failure.
+This can be seen in the posterior $ROV$ distribution of model V (3D).
+Resulting Bayes actions were characterized by an increase in divergence and expected losses, indicating a deterioration of the state of knowledge and an according increase in decision uncertainty.
+Thus, in some cases, implementing Bayesian inference might leave actors in greater disagreement than before.\\
+	However, we furthermore have to consider that actors weight possible outcomes of the probability distribution differently.
+They consequently are affected differently by the same type of additional information.
+Risk-friendly actors were the most robust in their decision-making in the face of possible trap failure.
+Eliminating this risk proved to be far less significant to the most risk-friendly, than for risk-averse actors.
+Accordingly, it should be of foremost importance, for risk-averse actors, to reduce the uncertainty regarding threshold-related factors which might decide between the success and complete failure of a project.
+This is less relevant for risk-friendly decisions makers, who respectively might acquire a comparable benefit from knowing more about the probability of positive outcomes, as they are less discouraged from overestimating and less afraid of failure.
+It was stated by \citet{bratvold2010making} that uncertainty has to possible consequences: risk and opportunity.
+According to them, risk is a consequence which is subjectively undesirable to an actor, while opportunity is the opposite.
+In our loss function, this seems to be well represented by varying affinities to rather over- or underestimate.
+For our risk-averse actors, a major risk is posed by seal failure.
+Our risk-friendliest individual sees a greater risk in missing out on opportunity.
+Nevertheless, it is suggested by the results from the different inference cases, that the best states of knowledge are attained by eliminating multimodality and narrowing the range of probabilities to one mode anywhere along the value axis.\\
+	%The respective estimate resilient to risk was generally found close to the positive mode and experienced only minor lateral shifting, after diminishing the probability of failure.
+Greater shifts were caused for this actor, in cases in which the positive mode itself was shifted laterally or erased completely.
+Conversely, the Bayes action of the most-averse individual was influenced greatly by the probability of the "worst case" scenario, complete trap failure.
+Only a stark decrease of the respective mode encouraged this actor to bid on an estimate close to the risk-friendlier decision-makers.
+These observations match the principles and conditions conceived in the design of the loss function and the chosen range of risk factors.
+	A central conclusion from these observations seems to be, that more of simply any type of information does not necessarily lead to better decisions.
+Instead, improvement of decision-making is achieved by attaining the right kind of information that is able to shed light on uncertainties which are relevant to an individuals own goals and preferences, as well as the general problem at hand.
+These findings pair well with statements made by \citet{bratvold2010making}, who emphasized that the purpose of models in the hydrocarbon sector is to attain insights to improve decisions.
+In their book, they point out that value is not generated by uncertainty quantification or reduction in itself, but is created to the extent that these processes have potential to change a decision.
+This is reflected by our results which showed that one set of information might have a vastly different impact on the Bayes estimators of different actors.
+New observations might lead to a great change in decision for one actor, but be negligible the other.\\
+		%This state of knowledge and the decision uncertainty accordingly experienced alterations related to the different Bayesian inference cases.
 These changes related foremost to modes in the value distribution.
@@
 reducing the standard deviation), or the opposite, a diminishment of the mode up to total erasure of its probability.\\
-	Considering that these mode changes correlated with increased or decreased relevance of specific trap control mechanisms in the 3D model, it can be argued that each mode represents the occurrence of a specific mechanism-related scenario.
-These are mostly distinctly separated with little probability in between.
-This presumably is based on the assumption that traps are always "filled-to-spill" in respect to spill and leak point, and that full leakage is assumed for any type of seal failure, resulting in an approximate discretization into respective scenarios.\\
+	Effects of Bayesian inference were reflected in changing appearances of modes in the reservoir value distributions.
+Considering that these changes correlated with increased or decreased relevance of specific trap control mechanisms in the 3D model, we derive that each mode represents the occurrence of a specific mechanism-related scenario.
+These were mostly distinctly separated with little probability in between.
+This presumably resulted from the assumption that traps are always "filled-to-spill" in respect to spill and leak point, and that full leakage is assumed for any type of seal failure, resulting in an approximate discretization into respective scenarios.\\
 	From a decision-making perspective, these scenarios might be interpreted as different decision options.
-In a case of distinct bimodality as seen in the prior of the 3D model, two main options would be given: (1) to not take any action, or (2) to expect a positive value in the range of the positive mode and invest accordingly.
+In a case of distinct bimodality as seen in the prior of the 3D model, two main options would be given: (1) to not take any action, or (2) to allocate resources according to an estimated value within the positive mode.
 However, most often, low probabilities for the values between these two opposite options remain.
@@
 A possibility could be to subdivide the value probability distribution accordingly into regions.
-Regarding respective integral probabilities for each such section and by introducing a critical integral probability as a threshold, it could be determined whether such a subregion is to be considered an actual option for decision-making or not.
+Considering respective integral probabilities for each such section and by introducing a critical integral probability as a threshold, it could be determined whether such a subregion is to be considered an actual option for decision-making or not.
 If a Bayes action is located outside of such a decision option, it could be automatically assigned to the closest region of sufficient integral probability.
 This approach should be easily applicable as an extension to our custom loss function.
-Furthermore, it could serve to build a bridge between this continuous loss function method and commonly used discrete approaches to decision-making, such as decision trees.
+Furthermore, it could serve to build a bridge between this continuous loss function method and discrete approaches to decision-making, such as decision trees.
 In real cases, normally only a limited number of options is given.
@@
 A discretization of decision subregions in a probability distribution could thus also be based on previously defined actual options.\\
-
-	The proportionality regarding modes, their relative probabilities and distances to each other, in turn seemed to influence the overall uncertainty of the probability distribution, i.e.
-the decision uncertainty.
-This was often indicated by the position of the mean and median relative to the modes.
-A mean located in the middle between two widely separated modes, which might be of low intrinsic uncertainty themselves, would indicate a much higher overall uncertainty, than a mean found in the middle of a narrow unimodal distribution.\\
-
-	The decision-making of differently risk-affine actors, i.e.
-the position and spread of Bayes actions, manifest according to the properties of value probability distributions and their inherent decision uncertainty.
-Distinct bimodality between two extremes, i.e.
-high decision uncertainty, resulted in a higher spread of Bayes estimators.
-Reduction of the distribution to one mode conversely led to the convergence of different Bayes actions.
-A decrease in decision uncertainty furthermore was accompanied by a reduction in expected loss for each Bayes estimator.
-Risk-friendly actors were the most robust in their decision-making, particularly in the face of possible trap failure.
-Eliminating this risk proved to be far less significant to the most risk-friendly, than for risk-averse actors.
-The respective estimate resilient to risk was generally found close to the positive mode and experienced only minor lateral shifting, after diminishing the probability of failure.
-Greater shifts were caused for this actor, in cases in which the positive mode itself was shifted laterally or erased completely.
-Conversely, the Bayes action of the most-averse individual was influenced greatly by the probability of the "worst case" scenario, complete trap failure.
-Only a stark decrease of the respective mode encouraged this actor to bid on an estimate close to the risk-friendlier decision-makers.
-These observations match the principles and conditions conceived in the design of the loss function and the chosen range of risk factors.
-A clear difference in the decision-making relative to risk-affinity was observed.\\
-
-	Considering the nature of changes experienced by the different Bayes action, depending on variations in the underlying reservoir value distributions, we derive that the degree of convergence of Bayes actions and their expected losses can be considered measures for the state of knowledge during the decision-making process.
-The better this is, the more similar the decisions of differently risk-affine actors and the lower their loss expectations are.
-It can be assumed, that given perfect information, all actors would bid on the same estimate, the true value, and expect no loss.
-It follows from this that the relevance of risk factors decreases with higher reduction of decision uncertainty.\\
-
-	While in most cases Bayesian inference led to a reduction of decision uncertainty, this was not always the case.
-It appears that the change in decision uncertainty is not necessarily strictly aligned with the change in uncertainty regarding model parameters and their derivatives.
-In respect to this, it seems to be of central importance "where" in the model uncertainty is reduced, i.e.
-in which spatial area or regarding which model parameters.
-This appears to be of particular significance considering threshold values that lead to an abrupt cut-off between two extrema of decision options.
-In both types of model construction, 1D and 3D, this was directly related to sealing reliability.
-Thresholds regarding seal thickness (1D model) and Shale Smear Factor ($SSF_c$ in 3D model) were defined in a way that introduced a significant possibility of complete trap failure.
-Consequently, it was observed that reducing the uncertainty in a way that narrowed the probability of a threshold-related parameter around its cut-off value, led to an amplification of the respective mode and thus emphasized the risk of complete failure.
-This can be seen in the posterior $ROV$ distribution of model V (3D).
-Resulting Bayes actions were characterized by an increase in divergence and expected losses, indicating a deterioration of the state of knowledge and an according increase in decision uncertainty.
-Thus, reducing model uncertainty in the "wrong" areas seems to simply lead to a transformation of this uncertainty in the realm of decision-making.\\
-	%The decision uncertainty seems to be conserved to a certain extent or even increased, as the duality of the decision problem is amplified in the form of a more stretched-out bimodal distribution and diverging Bayes actions.
-	According to this, it should be of foremost importance, especially for risk-averse actors, to reduce the uncertainty regarding threshold-related factors which might decide between the success and complete failure of a project.
-In the context of hydrocarbon traps, seal reliability is decisive.
-For making better decisions, elimination of such high risks should be a priority.
-As pointed out above, this might be less relevant for risk-friendly decisions makers, who respectively might acquire a comparable benefit from knowing more about the probability of positive outcomes, as they are less discouraged from overestimating and less afraid of failure.
-Nevertheless, it is suggested by the results from the different inference cases, that the best states of knowledge are attained by eliminating bimodality and narrowing the range of probabilities to one mode anywhere along the value axis.\\
-
-	Crucial risks might be easily assessed, if they are dependent on only one or a few parameters, such as seal thickness in the 1D model.
+	DRAW COMPARISON TO SIMPLE MCS AND DISCRETIZED DECISION THEORY IN HYDROCARBON E\&P HERE!\\	
+	As follows from our elaborations above, it is important to recognize crucial risks resulting from uncertainty.
+Such risks might be easily assessed, if they are dependent on only one or a few parameters, such as seal thickness in the 1D model.
 In other cases, they are derived from more complex parameter inter-relations, as is the case for the Shale Smear Factor, which is the ratio of displacement to seal thickness.
-Models IV and V (3D) showed that reduced uncertainty about the $SSF$ is not necessarily directly recognizable by information entropy visualization, especially if uncertainty regarding its parent parameters remains high.
-It follows that, to approach an effective mitigation of high risks, the complexities behind decisive factors need to be assessed thoroughly and respective parent parameters, as well as their interdependencies need to be recognized.
+Models IV and V (3D) showed that reduced uncertainty about the $SSF$ is not necessarily directly recognizable via Shannon entropy visualization, especially if uncertainty regarding its parent parameters remains high.
+It follows that, to approach an effective mitigation of high risks, the complexities behind decisive factors need to be assessed thoroughly and respective parent parameters, as well as their interdependencies, need to be identified.
 This might enable a better understanding of which type of information is missing and where in the model additional data might be of use for improved decision-making.\\
-
-	A central conclusion from these observations seems to be, that more of simply any type of information does not necessarily lead to better decisions.
-Instead, improvement of decision-making is achieved by attaining the right kind of information that is able to shed light on uncertainties which are relevant to an individuals own goals and preferences.
-These findings pair well with statements made by \citet{bratvold2010making}, who emphasized that the purpose of models in the hydrocarbon sector is to attain insights to improve decisions.
-In their book, they point out that value is not generated by uncertainty quantification or reduction in itself, but is created to the extent that these processes have potential to change a decision.
-This is reflected by our results which showed that one set of information might have a vastly different impact on the Bayes estimators of different actors.
-Assuming discretized options, new observations might lead to a change in decision for one actor, but not for the other.
-
-	It has too be emphasized that the models constructed in this work were purely artificial and not based on real data.
-Nevertheless, the 3D model in particular was designed to include some typical structural characteristics related to hydrocarbon systems and algorithms were developed to consider the most common conditions that define structural traps.
+	It has too be emphasized that the models constructed in this work were synthetic and not based on real data.
+Nevertheless, the 3D model in particular was designed to include some typical structural characteristics related to hydrocarbon systems.
+We developed algorithms aimed to consider the most common conditions that define structural traps.
 The fact that similar observations were made for the 1D, as well as the 3D model, indicates a certain degree of continuity with respect to representativity.
-However, it also has to be pointed out, that the uncertainties employed in the 3D model related to $z$-axis positional values only, and were thus of mostly  one-dimensional nature.
-This may in part explain the parallelism compared to the 1D model results.
-Furthermore, it follows that no effective uncertainty concerning the overall structural shape was implemented, particularly regarding anticlinal features and the position of the spill point in relation to the trap maximum.\\
-
+However, we also need to address that the uncertainties employed in the 3D model related to $z$-axis positional values only and were thus of mostly  one-dimensional nature.
+This may in part explain the parallelism to the 1D model results.
+Furthermore, it follows that no effective uncertainty concerning the overall structural shape was implemented, particularly regarding anticlinal features and the position of the spill point in relation to the trap top.
+Due to this, trap volumes tended to occur primarily in distinct "filled-to-spill" scenarios.
+While being unrealistic, this contributed to pronounced influences related to the various control mechanisms.\\
 	For the purpose of a conceptual application, the customization of the loss functions was kept relatively simple.
-As mentioned above, despite its simplicity, the design of the function and the incorporation of a risk factor appears to have been suitable to achieve a clear distinction between differently risk-affine actors and respective behavior.
-It was stated by \citet{bratvold2010making} that uncertainty has to possible consequences: risk and opportunity.
-According to them, risk is a consequence which is subjectively undesirable to an actor, while opportunity is the opposite.
-In our loss function, this seems to be well represented by varying affinities to rather over- or underestimate.
-For our risk-averse actors, a major risk is posed by seal failure.
-To our risk-friendliest individual, missing out on opportunity seems to be a greater risk.
-While standard loss functions would have only returned mean or median as estimates, loss function customization offers a flexible approach to express different objective and subjective weighting factors that are specified to certain decision-making environments and decision-makers' perspectives.\\
+We argue that despite its simplicity, the design of the function and the incorporation of a risk factor was suitable to achieve a clear distinction between differently risk-affine actors and respective behavior.
 It would be possible to extend this design to attain very complex loss functions.
-As the loss function in this work was designed based on only a few basic assumptions, it might be more generally representative than a complex function which takes into account numerous very specific aspects.
-The consideration of more details, without a base of real data, would have furthermore required extensive speculation, which would have presumably impaired the generalization potential of respective observations.\\
-
+As the loss function in this work was designed based on only a few basic assumptions, it might be more generally representative than a complex function which would takes into account numerous very specific aspects.
+The consideration of more details, without a basis on real data, would have furthermore required extensive speculation, which would have presumably impaired the generalization potential of respective observations.
+While standard loss functions only return mean or median as estimates, loss function customization offers a flexible approach to express different objective and subjective weighting factors that are specified to certain decision-making environments and decision-makers' perspectives.\\
 	Considering the findings of this work, there is still many points that could be expanded on in future research.

### 2017-11-22 Risk analysis/Text/Layout/Layout/chapter1.tex
Kind of finished all main parts

@@
 Bayesian decision theory and the related concepts of expected loss and loss functions have found increasingly common application in several economic sectors and fields of research, such as medicine \citep{ashby2000evidence, ashby2006bayesian, moye2006statistical} and machine learning \citep{barber2012bayesian, theodoridis2015machine}.
-Probabilistic approaches to decision-making have also become prevalent in the sector of hydrocarbon exploration and production (SOURCE).
-However, the methods here a mainly limited on (p10-p90, decision trees).......(see BRATVOLD)???\\
+Probabilistic approaches to decision-making have also become prevalent in hydrocarbon exploration and production \citep{murtha1997monte, mudford2000valuing,bratvold2010making}.
+However, in this sector, the methods are mainly limited to conducting simple Monte Carlo simulations and results are often interpreted viewing only percentiles and single values such as the mean.
+Bayesian inference and Markov chain Monte Carlo methods have only recently received more attention by researchers in this field (see \citet{wadsley2005markov}, \citet{ma2006multistage} and \citet{liu2010continuous}).\\
 In geosciences, Bayesian inference has prominently found use in the context of geophysical inversion problems (see \citet{tarantola1982inverse, mosegaard2002probabilistic} and \citet{sambridge2002monte}).
@@
 This work builds upon their concept, exploring the potential significance their findings might have in the context of decision-making.
-Bayesian decision-theory is to be included in the step of model evaluation.
-This is achieved by assigning an economic meaning to the structural model and designing a case-specific custom loss functions to find decisions which are optimal related to the state of knowledge and the preferences of actors with different risk-affinities.
+For this, we include Bayesian decision theory in the step of model evaluation.
+This is achieved by assigning an economic meaning to the structural model and designing a case-specific custom loss function to find decisions which are optimal related to the state of knowledge and the preferences of actors with different risk-affinities.
 More specifically, the models are designed to represent potential petroleum systems.
 Consequently, the development of algorithms for automatic hydrocarbon trap recognition and volume calculation represent a central part of this work.\\
-The main hypothesis of this work is that Bayesian inference and resulting changes in uncertainties in a geological setting have a significant effect on related value estimation and decision-making.
-It is furthermore postulated that loss functions can be customized to appropriately represent preferences of actors in the hydrocarbon sector and moreover illustrate the nature of decisions such actor's might make depending on their individual attitudes towards risk and in the face of different types of uncertainties.
+The main hypothesis of this work is that Bayesian inference and resulting changes in uncertainties in a geological setting have a significant effect on reservoir value estimation and decision-making.
+We furthermore postulate that loss functions can be customized to appropriately represent preferences of actors in the hydrocarbon sector and moreover illustrate the nature of decisions such actor's might make depending on their individual attitudes towards risk and in the face of different types of uncertainties.
 Changes in their respective decisions are treated as a suitable measure to assess the effect of updating model parameters with new geological information.

### 2017-11-22 Risk analysis/Text/Layout/Layout/chapter4_discussion.tex
Kind of finished all main parts

@@
 This approach should be easily applicable as an extension to our custom loss function.
-Furthermore, it could serve to build a bridge between this continuous loss function method and discrete approaches to decision-making, such as decision trees.
+Furthermore, it could serve to build a bridge between this continuous loss function method and discrete approaches to structuring decision problems, such as decision trees.
 In real cases, normally only a limited number of options is given.
@@
 A discretization of decision subregions in a probability distribution could thus also be based on previously defined actual options.\\
-	DRAW COMPARISON TO SIMPLE MCS AND DISCRETIZED DECISION THEORY IN HYDROCARBON E\&P HERE!\\	
+	In remains to be argued, to what extent our methodology and findings contribute regarding practical applications in hydrocarbon exploration and production.
+Monte Carlo simulation for reservoir estimation and risk assessment have become common in this sector and are often used in combination with decision trees to make decisions (see \citet{murtha1997monte}, \citet{mudford2000valuing} and \citet{bratvold2010making}).
+However, it seems to us that distributions resulting from probabilistic modeling are mostly only considered to attain best estimates in the form of means.
+Most likely and extreme outcomes are identified as percentiles, typically P50 (the median), P10 and P90.
+We believe that this practice does not harness the full potential of such a probabilistic distribution and that much of the inherent information potential is discarded.
+Contrary to that, customized loss functions as a Bayesian method, take into account the full probability distribution and enable the inclusion of various conditions in the process of finding an optimal estimate.
+While used in statistical decision theory and other scientific fields, loss functions have, to the best of our knowledge, found no significant application in the field of petroleum exploration and production.
+Thus, we intend to provide a new perspective with our methodology.
+\citet{murtha1997monte} emphasized that Monte Carlo simulation does not make decisions, it merely prepares for it.
+We believe that loss functions have the potential to go one step further.
+A hypothetical ideal loss function would consider all conditions in an economic environment, as well as perfectly represent preferences and goals of an actor and consequently be able to automatically find an optimal decision.
+While this is obviously unrealistic, we presume that an elaborate loss function might at least provide a very good preliminary decision recommendation.
+It might furthermore weight \\
+	More recently, Bayesian inference and MCMC methods were applied for OOIP estimation and forecasting of reservoir productivity by \citet{wadsley2005markov}, \citet{ma2006multistage} and \citet{liu2010continuous}.
+However, their research focused on simulations based on history matching for already producing fields.
+Our approach of applying Bayesian inference for structural geological modeling and volumetric reservoir calculations is intended to support decision-making in the earliest stages of a reservoir, when it has to be decided whether a project should be developed or not.
+Nevertheless, in was shown in the work by \citet{wadsley2005markov} that early volumetric OOIP estimates can be combined with later calculations from production data via MCMC methods.
+This is an area where our methods could possibly be integrated.\\
+	\citet{bratvold2010making} highlighted the potential to use Monte Carlo simulation to identify, via sensitivity analysis, which uncertainties have the greatest impact on  decisions.
+This was virtually achieved by performing several cases of Bayesian inference in our work, as we recognized primarily the importance of seal reliability.\\
 	As follows from our elaborations above, it is important to recognize crucial risks resulting from uncertainty.

### 2017-11-22 Risk analysis/Text/Layout/Layout/chapter1.tex
Working on corrections

@@
 \chapter{Introduction} \label{chap:intro}
-Bayesian methods are an intuitive approach to inference, naturally inherent in human thinking patterns and closely tied to processes of decision-making \citep{berger2013stat, davidson2015, jaynes1986bayesian}.
+Bayesian methods are an intuitive approach to inference, naturally inherent in human thinking patterns and closely tied to processes of decision making \citep{berger2013stat, davidson2015, jaynes1986bayesian}.
 Individuals are constantly faced with situations in which a decision has to be made, but only incomplete information is available.
@@
 \end{enumerate}
-This concept is relatable to a vast variety of problems, ranging from casual every-day situations to complex scenarios in large-scale economic decision-making: As a private person, should I take an umbrella with me today?
+This concept is relatable to a vast variety of problems, ranging from casual every-day situations to complex scenarios in large-scale economic decision making: As a private person, should I take an umbrella with me today?
 As a company, should we invest in the development and realization of a certain project?
 Following this process of plausible reasoning, the quality of a decision is to be measured based on the preceding state of knowledge and reasonable expectations, not on the subsequent actual consequences \citep{jaynes1986bayesian}.
-In other words: A decision is optimal, as long as it is the best action given the information available to the decision-maker before making the decision, no matter if actual loss was incurred afterwards.\\
+In other words: A decision is optimal, as long as it is the best action given the information available to the decision maker before making the decision, no matter if actual loss was incurred afterwards.\\
 Bayesian decision theory and the related concepts of expected loss and loss functions have found increasingly common application in several economic sectors and fields of research, such as medicine \citep{ashby2000evidence, ashby2006bayesian, moye2006statistical} and machine learning \citep{barber2012bayesian, theodoridis2015machine}.
-Probabilistic approaches to decision-making have also become prevalent in hydrocarbon exploration and production \citep{murtha1997monte, mudford2000valuing,bratvold2010making}.
+Probabilistic approaches to decision making have also become prevalent in hydrocarbon exploration and production \citep{murtha1997monte, mudford2000valuing,bratvold2010making}.
 However, in this sector, the methods are mainly limited to conducting simple Monte Carlo simulations and results are often interpreted viewing only percentiles and single values such as the mean.
@@
 Using Markov chain Monte Carlo sampling to explore resulting probability spaces, they attained posterior model suites with reduced uncertainties \citep{delaVarga2016}\\
-This work builds upon their concept, exploring the potential significance their findings might have in the context of decision-making.
+This work builds upon their concept, exploring the potential significance their findings might have in the context of decision making.
 For this, we include Bayesian decision theory in the step of model evaluation.
@@
 Consequently, the development of algorithms for automatic hydrocarbon trap recognition and volume calculation represent a central part of this work.\\
-The main hypothesis of this work is that Bayesian inference and resulting changes in uncertainties in a geological setting have a significant effect on reservoir value estimation and decision-making.
+The main hypothesis of this work is that Bayesian inference and resulting changes in uncertainties in a geological setting have a significant effect on reservoir value estimation and decision making.
 We furthermore postulate that loss functions can be customized to appropriately represent preferences of actors in the hydrocarbon sector and moreover illustrate the nature of decisions such actor's might make depending on their individual attitudes towards risk and in the face of different types of uncertainties.

### 2017-11-22 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Working on corrections

@@
     \chapter{Methods}\label{cha:met}
-    The methods utilized in this work are presented in the following chapter.
+    In the following chapter, we present the methodology utilized in this work.
 Bayesian analysis and decision theory are introduced in Section \ref{sec:bayes}.
 Focus is laid on Bayesian inference, estimation of uncertain values and the use of loss functions in this context.
-    These methods are to be applied in the context of structural geological modeling.
+    We apply these methods in the context of structural geological modeling.
 Respective approaches to computational modeling, specifically in a probabilistic framework, are described in Section \ref{sec:bayes_in_struc_modeling}.
 Regarding numerical implementation, GemPy and PyMC are presented as central tools used in a Python environment.
-In Section \ref{sec:model_evaluation}, methods to evaluate consequent modeling results are outlined.
+In Section \ref{sec:model_evaluation}, we outline our methods to evaluate consequent modeling results are.
 Particularly the potential for valuation from an economic perspective is elaborated in this regard and basic principles of subsurface hydrocarbon systems are introduced.
 Considering the respective setting in the economic sector of petroleum production, the development of an according case-specific custom loss function is described in Section \ref{sec:LF_design}.
-These methodologies are to be applied on a conceptual 1D geological model first (see Section \ref{sec:1D_model}).
-Design, construction and methods to identify hydrocarbon trap features in a full 3D structural model are explained in Section \ref{sec:3D_model}.
+We first apply these methodologies on a conceptual 1D geological model (see Section \ref{sec:1D_model}), subsequently on a full 3D structural geological model.
+Design, construction and methods to identify hydrocarbon trap features in such the 3D model are explained in Section \ref{sec:3D_model}.
 
         \section{Bayesian analysis and decision theory}\label{sec:bayes}
-	    As implied by the name, the problems and reasoning behind decision-making are examined in the field of decision theory \citep{berger2013stat}.
+	    As implied by the name, the problems and reasoning behind decision making are examined in the field of decision theory \citep{berger2013stat}.
 Such decision problems are commonly influenced by parameters that are uncertain.
@@
 Such uncertain parameters can be considered as numerical quantities.
-In order to find the best decision to a problem, it is possible to combine sample information with other aspects such as the possible consequences of decision-making and the availability of prior information on our uncertainties.
+In order to find the best decision to a problem, it is possible to combine sample information with other aspects such as the possible consequences of decision making and the availability of prior information on our uncertainties.
 Decision consequences are expressed as gains in economic decision theory and as losses, which equal negative gains, in statistics.
@@
 	    First, some basic elements are to be defined.
-The unknown (uncertain) quantity influencing decision-making is usually denominated as the state of nature $\theta$ \citep{berger2013stat}.
+The unknown (uncertain) quantity influencing decision making is usually denominated as the state of nature $\theta$ \citep{berger2013stat}.
 Given statistical information on $\theta$ in the form of probability distributions, $\theta$ is called the parameter.
@@
         \subsection{Bayesian inference}
-        Bayesian inference is most importantly characterized by its preservation of uncertainty, in contrast to standard statistical inference \citep{box2011bayesian, jaynes2003probability, harney2013bayesian, davidson2015}.
+        Bayesian inference is most importantly characterized by its preservation of uncertainty, in contrast to standard statistical inference \citep{jaynes2003probability, box2011bayesian, harney2013bayesian, davidson2015}.
 The Bayesian approach is widely seen as intuitive and inherent in the natural human perspective.
@@
 This means that Bayesian updating is about reducing uncertainty in a belief and reaching a guess that is less wrong.
-Bayesian inference is defined by and conducted via the following equation, called the Bayes' Theorem \citep{box2011bayesian, jaynes2003probability, harney2013bayesian, davidson2015}:
+Bayesian inference is defined by and conducted via the following equation, called the Bayes' Theorem \citep{jaynes2003probability, box2011bayesian, harney2013bayesian, davidson2015}:
         \begin{equation}\label{eq:BayesTheorem}
@@
 In this work, the geological models are additionally assigned an economic meaning by declaring them potential petroleum reservoirs and introducing customized loss functions to reflect the economic interest of decision makers in developing respective resource extraction projects (see Section \ref{sec:loss}).
-Changes in Bayes actions are considered a measure for the influence of Bayesian inference on decision-making and the significance of additional observations for different decision makers.
+Changes in Bayes actions are considered a measure for the influence of Bayesian inference on decision making and the significance of additional observations for different decision makers.
         \end{enumerate}
@@
         \subsection{Economic significance of structural geological models as hydrocarbon systems}\label{sec:Reservoir_values}
-        In this work, the generated structural geological models are to be assigned an economic significance in order to approach a real setting, in which hypothetical actors could have an interest in taking a decision, and thus, to evaluate the influence of the Bayesian inference of the decision-making process.
+        In this work, the generated structural geological models are to be assigned an economic significance in order to approach a real setting, in which hypothetical actors could have an interest in taking a decision, and thus, to evaluate the influence of the Bayesian inference of the decision making process.
 Changes in Bayes actions are considered as a measure for this effect.
@@
 Consequently, calculating the $ROV$ from the maximum trap volume simply equates a rescaling of values, as proportions are conserved.
-However, translation of trap volumes to recoverable oil volumes is deemed appropriate for a formally better consideration of the results in the respective context of decision-making in the hydrocarbon sector.\\
+However, translation of trap volumes to recoverable oil volumes is deemed appropriate for a formally better consideration of the results in the respective context of decision making in the hydrocarbon sector.\\
         %For most factors, it seems most suitable to define a normal distribution around a known typical value or global average, as it will return a high probability for common scenarios and occasionally produce cases of rare, more extreme values.\\
@@
 Fundamentally different loss functions can also be based on a significantly different mathematical structure.
-Loss functions are customized regarding the problem environment and according to the subjective needs and objectives of the decision-maker.
+Loss functions are customized regarding the problem environment and according to the subjective needs and objectives of the decision maker.
 Thus, they are mostly defined by the actor expressing his perspective.

### 2017-11-22 Risk analysis/Text/Layout/Layout/chapter4_discussion.tex
Working on corrections

@@
 One example would be subsurface storage of fluids in a reservoir, such as carbon capture and storage (CCS).
-Questions about storage capacity and safety deal with similar conditions and geological problems, as the ones presented this work, most importantly seal reliability and the risk of leakage.
+Questions regaring storage capacity and safety deal with similar conditions and geological problems, as the ones presented this work, most importantly seal reliability and the risk of leakage.
 In respect to this, the models and approaches in this work might provide a basis to raise new and advanced explorable cases.

### 2017-11-23 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
correction of important errors

@@
 Given only layer boundary positions in one vertical dimension, it is resorted to an abstract way of reservoir valuation by defining a scoring system.
-A dimensionless reservoir score is made dependent on three uncertain parameters which can be deduced from the 1D interface positions: (1) reservoir thickness, (2) reservoir top depth and (3) seal thickness (see Figure \ref{fig:3_parameters}).
-
+A dimensionless reservoir score is made dependent on three derivatives of the uncertain 1D interface positions: (1) reservoir thickness, (2) reservoir top depth and (3) seal thickness.
+From these, (2) equals the prior parameter that indicates the reservoir-seal boundary in depth.
+The other two are respectively deduced from the distance between parameter values.
+According to this dependence, these values, which we use to valuate the reservoir, inherit the uncertainty from their parent parameters.
+An exemplary comparison between the distribution of these values depending on prior and posterior parameters is depicted in Figure \ref{fig:3_parameters}.
+We hereby illustrate how model uncertainties are translated to uncertainties related to valuation measures.\\
 	        \begin{figure}[h]
@@
 	        	\includegraphics[width=1\textwidth]{Figures/3_parameters.png}
-	        	\caption{Probability distributions for the three parameters deduced from the 1D model depicted in Figure \ref{fig:1D_model}.
-Values from these parameter distributions are combined in Equation \ref{eq:1D_score_system} to attain a score for valuation of reservoir model realizations.}\label{fig:3_parameters}
+	        	\caption{Probability distributions for the three factors used in the abstract reservoir valuation as formulated in Equation \ref{eq:1D_score_system}.
+These are deduced from the 1D model input parameters and are thus uncertain.
+Here, we show factors deduced from prior parameters in Figure \ref{fig:1D_model} and factors resulting from exemplary posterior parameter distributions.}\label{fig:3_parameters}
 	        \end{figure}
-
 	        Assuming that reservoir thickness is a simplified indicator for the amount of extractable oil or gas and thus value in place, a gain in score can be correlated with increase in thickness.
@@
 			where reservoir thickness is given by $h_{res}$, seal thickness by $h_{seal}$ and reservoir depth, which is always a negative value, by $d_{res}$.\\
-			This valuation method is applied on the original prior-only Monte Carlo sampling of the 1D model as reference and then on each posterior distribution set after conducting Bayesian inference.
+			This valuation method is applied on the original prior-only Monte Carlo simulation of the 1D model as reference and then on each posterior distribution set after conducting Bayesian inference.
 Resulting score probability distributions can be compared directly and additionally under use of the case-specific custom loss function (Equation \ref{eq:LFR_final}) defined before.

### 2017-11-23 Risk analysis/Text/Layout/Layout/chapter3_results.tex
Enough for today
correction of important errors

@@
 The main distribution is not changed significantly, except for a striking peak of probability at the possibility for a score of -2000.
-This presumably represents the bulk of cases, in which the seal is assumed to have failed.
+This presumably represents the bulk of cases in which the seal is assumed to have failed.
 Regarding this, it is to be noted, that the mean of the seal top distribution is found at -2000 m.
@@
 		\subsection{Posterior model IV: SSF likelihood reinforcing trap reliability}%ML4
+		\begin{figure}[p!]
+			\centering
+			\includegraphics[width=1\textwidth]{Figures/Appendix/ML4}
+			\caption{....}\label{fig:ML4}
+		\end{figure}
 		For this fourth model, thickness likelihoods were chosen to be comparable to those of model I, i.e.
@@
 With a mean of 1.5 and a standard deviation of 0.5, it was chosen to mostly reinforce the likelihood of a $SSF$ below the critical value of $SSF_c = 3$.\\
-		In Figure \ref{ML4} (b), it can be seen that the uncertainty about the position of interfaces in the footwall was greatly reduced for all layers.
+		In Figure \ref{fig:ML4} (b), it can be seen that the uncertainty about the position of interfaces in the footwall was greatly reduced for all layers.
 In the hanging wall, information entropy remained high but was shifted upwards and mainly concentrated in an area that is juxtaposed to the footwall seal formation.\\

### 2017-11-23 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
gurke home

@@
         	\includegraphics[width=1\textwidth]{Figures/loss_functions.png}
-        	\caption{Expected loss based on the standard absolute-error loss (A) and squared-loss function (B).}\label{fig:standard_LF} 
+        	\caption{Expected loss and respective Bayes actions based on the standard absolute-error loss (A) and squared-loss function (B).
+In this example, we applied the functions on a normal distribution characterized by $\mu = 0$ and $\sigma = 500$.}\label{fig:standard_LF}
         \end{figure}
@@
         The expectation symbol $E$ is subscripted with $\theta$, by which it is indicated that $\theta$ is the respective unknown variable.
-This expected loss $l$ is also referred to as the risk of estimate $\hat{\theta}$ \citep{davidson2015}.\\
+This expected loss $l$ is also referred to as the Bayes risk of estimate $\hat{\theta}$ \citep{berger2013stat, davidson2015}.
+Exemplary realizations of the expected loss according to both standard loss functions and minimized by the Bayes action are plotted in Figure \ref{fig:standard_LF}.\\
         By the Law of Large Numbers, the expected loss of $\hat{\theta}$ can be approximated drawing a large sample size $N$ from the posterior distribution, respectively applying a loss function $L$ and averaging over the number of samples \citep{davidson2015}:
@@
 The MAP (maximum a posteriori) estimate is the minimizing solution for the posterior using zero-one loss \citep{davidson2015}.
-The possibility of more than on minimum also implies that several Bayes actions can exist for one problem \citep{berger2013stat}.\\
+The possibility of more than one minimum also implies that several Bayes actions can exist for one problem \citep{berger2013stat}.\\
         \citet{davidson2015} implemented different risk affinities by simply  introducing a risk parameter into the loss function.
@@
 The fundamental approach follows closely the research conducted by \citet{delaVarga2016} and builds upon their findings.\\
-        According to them, structural geological modeling can be regarded as a forward problem and the elements of Bayesian inference can be specified in this context as follows:
+        According to them, structural geological modeling can be regarded as a statistical problem and the elements of Bayesian inference can be specified in this context as follows:
         \begin{enumerate}
         	\item \textbf{Mathematical forward model ($M$)}: The connections between parameters $\theta$ and observed data $y$ are defined in such mathematical models.
-According to \citet{delaVarga2016}, the realization of a geological model $M$ can be regarded as a direct function of a set of input parameters:
+According to \citet{wellmann2010towards} and \citet{delaVarga2016}, the realization of a geological model $M$ can be regarded as a direct function of a set of input parameters:
         	\begin{equation}
@@
         	\end{equation}
-        	where $\phi_i$ is a function of position $\vec{x}$, more precisely an interpolation function, to which respective additional parameters are given by $\alpha_k$.
+        	where $\phi_i$ is a function of position $\vec{x}$, more precisely an interpolation function, to which respective additional interpolation parameters are given by $\alpha_k$.
 Available primary geological information, such as positions and dips of layer interfaces, is represented by $k_j$.
-A topological description, such as the relationships between faults and layer surfaces, is given by $\beta_l$.
+A topological description, such as the relationships between faults and layer surfaces, is given by $\beta_\ell$.
 An essential aspect of this function is that it allows for a full automation of the modeling step, so that the consequences of a change in an input parameter are realized directly without the need for any further manual inspection or interaction \citep{wellmann2017sandstone, delaVarga2016}.
@@
 Regarding Equation \ref{eq:forward_model} above, these can be any of the factors $\vec{x}$, $k_j$, $\alpha_k$ and $\beta_l$, which define the realization of the forward model over $\phi_i$.
-        	\item \textbf{Observed data ($y$)}: This is any type of additional information that can be related to the forward modeling results and might possibly be used to reduce uncertainty.
+        	\item \textbf{Observed data ($y$)}: This is any type of additional information that can be related to the forward modeling results or to the parameters and their combinations, and might possibly be used to reduce uncertainty.
 Such data can be gained by measurements and observations, for example by core sampling, well-log analysis or seismic acquisition.
-        	\item \textbf{Likelihood functions $p(y|\theta)$}: Links between the previous parameters $\theta$ and the additional data $y$ are established by these functions in way that they reflect the likelihood of the parameter states given the observations.
-They mathematically defined in the same way as probability functions, but are a function of the data $y$, instead of the parameters $\theta$ \citep{delaVarga2016}.
+        	\item \textbf{Likelihood functions $p(y|\theta)$}: Links between the previous parameters $\theta$ and the additional data $y$ are established by these functions in a way that they reflect the likelihood of the parameter states given the observations.
+They are mathematically defined in the same way as probability functions, but are a function of the data $y$, instead of the parameters $\theta$ \citep{mackay2003information, patil2010pymc, delaVarga2016}.
         \end{enumerate}
@@
 This is the step of Bayesian updating of the belief about the parameter uncertainty given new information.
-        	In a chosen model ($M$), this is achieved by linking parameters and data through deterministic operations which are additionally compared to the likelihood functions.
+        	In a chosen model ($M$), this is achieved by linking parameters and data through deterministic operations to the likelihood functions.
 It is pointed out by \citet{delaVarga2016} that any combination of parameter-observation connections is allowed, i.e.
 not all parameters need to be necessarily connected to all observed data.
-After all conditional probabilities have been set up, the Bayes Theorem (Equation \ref{eq:BayesTheorem}) is applied to attain the posterior citep{delaVarga2016}.
+After all conditional probabilities have been set up, the Bayes Theorem (Equation \ref{eq:BayesTheorem}) is applied to attain the posterior \citep{delaVarga2016}.
 However, due to the multi-dimensionality given in geological problems, the use of Markov chain Monte Carlo methods is advised to achieve this as described in Section \ref{sec:mcmc} below.
@@
         In the ordinary Monte Carlo approach, random independent samples are drawn from a target distribution in order to approximate its shape \citep{gilks2005markov, delaVarga2016}.
-High-dimensional parameter spaces as found in Bayesian applications lead to complex shapes and often make independent sampling infeasible \citep{gilks2005markov}.
+High-dimensional parameter spaces as found in Bayesian applications lead to vast spaces of likelihood and often make independent sampling infeasible \citep{gilks2005markov}.
 This can be solved by extending the Monte Carlo principle with a Markov chain, in which every sample iteration of the parameter $\theta^{(t+1)}$ is dependent uniquely on the previous value $\theta^{(t)}$ \citep{gilks2005markov, delaVarga2016}.\\
@@
 The new sample value is then related to the previous step.
-According to a weight defined by the scaled up candidate density of the value, the proposed step is then accepted or rejected.
+According to a weight defined by the scaled-up candidate density of the value, the proposed step is then accepted or rejected.
 In the case of acceptance, the value is added to the sample trace and the process is continued from the current location.
@@
         For the purpose of this work, an adaptive Metropolis-Hastings sampler is used.\\
-        In Metropolis-Hastings methods, each sampling step at iteration $t$ is determined by a candidate probability distribution $q(\theta,\theta^{'})$, from which a proposed sample $\theta^{'}$ is drawn \citep{delaVarga2016}.
-The acceptance-rejection condition is defined by the acceptance ratio $a(\theta^{'},\theta) \citep{delaVarga2016}$:
+        In Metropolis-Hastings methods, each sampling step at iteration $t$ is determined by a candidate probability distribution $q(\theta,\theta^{'})$, from which a proposed sample $\theta^{'}$ is drawn.
+The acceptance-rejection condition is defined by the acceptance ratio $a(\theta^{'},\theta)$ \citep{haario2001adaptive}:
 
@@
         For the iterative sampling in this work, Adaptive Metropolis (AM) by \citet{haario2001adaptive} is used.
-It adapts the traditional Metropolis-Hastings by incorporating the ability of continuous step size tuning during convergence, by taking into account the full information saved along the process.
+It adapts the traditional Metropolis-Hastings by incorporating the ability of continuous step-size tuning during convergence, by taking into account the full information saved along the process.
 This is achieved by generating a covariance matrix that is updated every iteration.
@@
 This is primarily dependent on the rate of convergence towards the true distribution.
-Considering empirical Bayes methods, in which either priors or likelihoods stem from empirical data, as is assumed for the models in this work, convergence can be expected to be reached almost immediately.
+Considering empirical Bayes methods, in which either priors or likelihoods stem from empirical data, as is assumed for the models in this work, convergence can be expected to be reached almost immediately \citep{delaVarga2016}.
 However, for high dimensional problems more iterations are required, in order to ensure an accurate representation of posterior distributions \citep{wellmann2017sandstone}.
@@
 For the application in a probabilistic setting, the method should enable fully automatic reconstruction of the model, when parameters are changed.
-The application in this work follows the example of \citet{delaVarga2016} and relies on the use of implicit interpolation for geological modeling, a method developed and elaborated by \citet{lajaunie1997foliation} and \citet{calcagno2008geological}.\\
+The application in this work follows the example of \citet{wellmann2010towards} and \citet{delaVarga2016}, and relies on the use of implicit interpolation for geological modeling, a method developed and elaborated by \citet{lajaunie1997foliation} and \citet{calcagno2008geological}.\\
         This implicit method relies on the interpolation of a potential field scalar function $T(\vec{x})$ of any point $\vec{x}$ in a 3D space, and thus reflects the geometry of geological structures \citep{calcagno2008geological}.
-Modeling $T(\vec{x})$ is achieved by cokriging that regards two forms of data: (i) contact points on geological interfaces through increments of the potential field $T(\vec{x})-T(\vec{x}^{\,'})$ and (ii) orientation data as gradients, i.e.
+Modeling $T(\vec{x})$ is achieved by cokriging two forms of data: (i) contact points on geological interfaces through increments of the potential field $T(\vec{x})-T(\vec{x}^{\,'})$ and (ii) orientation data as gradients, i.e.
 partial derivatives of the potential field $\delta T(\vec{x})/\delta u_\beta$ in each directon $u$ \citep{calcagno2008geological}.
@@
 		\subsection{Numerical computational implementation via Python, GemPy and PyMC}\label{sec:numerical_implementation}
-		Bayesian analysis can be conducted using probabilistic programming \citep{salvatier2016pymc3}.
+		Bayesian analysis can be conducted numerically using probabilistic programming methods \citep{salvatier2016pymc3}.
 The implicit method of forward geological modeling, described above, is to be embedded in such a framework.
 For doing so, the programming language of choice in this work is Python.
-The merits of Python have been pointed out by \citet{behnel2010, Langtangen2008} and \citet{salvatier2016pymc3}.
+The merits of Python have been pointed out by \citet{behnel2010}, \citet{Langtangen2008} and \citet{salvatier2016pymc3}.
 Development is facilitated by an expressive but concise and clean syntax that is easy to learn.
@@
 At the time of writing this, GemPy is still under development (version 0.995), but is already functioning for the purpose of this work.\\
-		For conducting the Bayesian analysis and embedding the geological modeling step in a probabilistic modeling framework, GemPy is combined with PyMC.
+		For conducting the Bayesian analysis and embedding the geological modeling step in a probabilistic context, GemPy is integrated into the probabilistic framework of PyMC.
 This Python library was developed for conducting Bayesian inference and prediction problems in an open-source probabilistic programming environment \citep{davidson2015, salvatier2016pymc3}.
 Different model fitting techniques are provided in PyMC, such as the \textit{maximum a posteriori} (MAP) method and several (MCMC) sampling methods, including the Adaptive Metropolis introduced in Section \ref{sec:mcmc}.
-The components which are used to construct a statistical model, are represented by \textit{Deterministic} or \textit{Stochastic} variables in PyMC \citep{salvatier2016pymc3}.
+The components which are used to construct a statistical model, are represented by \textit{Deterministic} or \textit{Stochastic} functions in PyMC \citep{salvatier2016pymc3}.
 These variables are described as hierarchically related nodes, specifically parent and child nodes, in a Bayesian network:
@@
 \textit{Stochastic} variables are used to represent uncertain parameters $\theta$ or observed stochastic variables as likelihood functions $p(y|\theta)$ \citep{salvatier2016pymc3, delaVarga2016}.
-Complex mathematical relations between \textit{Stochastic} variables can be described through \textit{Deterministic} variables \citep{delaVarga2016}.
+Complex mathematical relations between \textit{Stochastic} variables can be described through \textit{Deterministic} functions \citep{delaVarga2016}.
 Futhermore, PyMC allows for the creation of own object definitions inheriting from the class descriptions of these two variable types.\\
@@
 For this, the sampling traces and posterior distributions of single parameters can be observed.
-The entirety of realized 3D geological models is to be assessed in terms of uncertainty quantification.
+The entirety of accepted 3D geological models (sampled after burn-in phase) is to be assessed in terms of uncertainty quantification.
 Furthermore, a central part of this work is the assignment of an economic significance to the modeling results.
@@
 	        \includegraphics[width=1\textwidth]{Figures/information_entropy.png}
-	        \caption{Illustrations depicting the concept of information entropy to visualize uncertainties in spatial applications.
+	        \caption{Illustrations depicting the concept of Shannon entropy to visualize uncertainties in spatial applications.
 In (a), a map composed of uncertain units is subdivided into a regular grid of equally sized cells.
 The probabilities of possible outcomes regarding each unit is estimated for each cell in (b).
-From this results a map of cell information entropies (c), with the highest values where all units can occur equally frequent, and lowest for cells where no uncertainty is given and entropy is zero (from \citet{wellmann2012uncertainties}).}\label{fig:information_entropy}
+From this results a map of cell Shannon entropies (c), with the highest values where all units can occur equally frequent, and lowest for cells where no uncertainty is given and entropy is zero (from \citet{wellmann2012uncertainties}).}\label{fig:information_entropy}
         \end{figure}
-        Information entropy was first defined by \citet{shannon1948mathematical} and adopted as a method to visualize uncertainties in 3D geological models by \citet{wellmann2012uncertainties}.
+        Shannon entropy, commonly also referred to as information entropy, was first defined by \citet{shannon1948mathematical} and adopted as a method to visualize uncertainties in 3D geological models by \citet{wellmann2012uncertainties}.
 This approach was later also used by \citet{delaVarga2016} and \citet{schaaf2017}.
-Information entropy is applied to predict the model accuracy at every location in the model space by visualizing its respective uncertainty and thus returning a measure for the quality of the model.
+Shannon entropy is applied to predict the model accuracy at every location in the model space by visualizing its respective uncertainty and thus returning a measure for the quality of the model.
 This is achieved by subdiving the model space into a regular raster of equally sized cells (voxels) and measuring the accuracy for every such cell.
 In the context of geological modeling, the central question is based on the knowledge of how frequent a single geological feature occurs in a voxel \citep{wellmann2012uncertainties}.
-\citet{de1972definition} established that information entropy can serve as a measure of fuzziness, a concept described by \citet{zadeh1965fuzzy}.
+\citet{de1972definition} established that Shannon entropy can serve as a measure of fuzziness, a concept described by \citet{zadeh1965fuzzy}.
 Given a fuzzy set in which fuzziness of each part is indicated by $f\in [0,1]$, the following conditions hold:
@@
 	        \item The measure should only be 0 if $f$ is 0 or 1 in every cell.
-	        \item The measure is maximal given $f = 0.5$ in every cell \citep{wellmann2012uncertainties}.
+	        \item The measure is maximal given $f = 0.5$ in every cell \citep{de1972definition}.
         \end{enumerate}
@@
 all outcomes are equally likely in each cell \citep{schaaf2017}.
-From this, \citet{wellmann2012uncertainties} derived the following equation for application in geological modeling:
+From this, the following equation for application in geological modeling was derived \citet{Mann1993,wellmann2012uncertainties}:
         \begin{equation}
@@
 The fuzziness is quantified as the entropy $H_m$, normalized by the total number of cells $N$.
-Subsequently, a total information entropy can be calculated for a model as \citep{wellmann2012uncertainties}:
+Subsequently, an average Shannon entropy can be calculated for a model as \citep{wellmann2012uncertainties}:
         \begin{equation}
@@
         \end{equation}
-        This total entropy $H_T$ is 0 when no uncertainty is found in any cell, oppositely it is maximal in the case of equal likelihood of outcomes in every cell \citep{wellmann2012uncertainties, schaaf2017}.
+        This total entropy $H_T$ is 0 when no uncertainty is found in any cell, oppositely it is maximal in the case of equal probabilities of outcomes in every cell \citep{wellmann2012uncertainties, schaaf2017}.
 The application of these entropy measures allows for the calculation and visualization of uncertainties in each voxel, the assessment of uncertainties of entire geological units, as well as the quantification of total model uncertainty represented by a single number \citep{wellmann2012uncertainties}.

### 2017-11-24 Risk analysis/Text/Layout/Layout/MSc_layout_Fabian.tex
correction of important errors
Abstract (needs revision)

@@
 	\nonumchap{Abstract}
-Bayesian methods are an intuitive approach to inference and subsequent decision-making, relatable to a wide spectrum of problems.
-Hydrocarbon exploration and production is a high-risk, high-reward sector in which good decision-making is indispensable.
-However, actors in this field are faced by numerous uncertainties that have to be considered.
-Structural geological modeling is of central importance for the estimation of uncertain hydrocarbon accumulations in potential reservoirs.
-Recent developments have shown, that such modeling can be regarded as a Bayesian inference problem.
-This allows for the incorporation of additional geological information as likelihood functions in a probabilistic framework and a consequent reduction of model uncertainties.
+Bayesian decision theory is applicable to a wide spectrum of problems.
+Hydrocarbon exploration and production is a high-risk, high-reward sector in which good decision making is indispensable.
+Actors in this field are faced by numerous uncertainties that have to be considered.
+Structural geological modeling is of central importance for the assessment of uncertain hydrocarbon accumulations in potential reservoirs.
+By regarding such modeling as a Bayesian inference problem, additional geological information can be incorporated as likelihood functions linked to prior parameters in a probabilistic framework.
+Markov chain Monte Carlo sampling is used to approximate posterior models of reduced uncertainty.
+Here, synthetic geological models are constructed to represent potential hydrocarbon system.
+Algorithms for automatic trap recognition are developed to enable the reservoir valuation of prior and posterior models.
+These serve as a base for true value estimation by respective optimization of a case-specific loss function.
+This function is customized to reflect the decision making of differently risk-affine actors in this context.
+Results from conducting this for several inference cases show that the various Bayes estimators shift according to the characteristics of the underlying value distribution.
+While bimodality and overall uncertainty leads to separation, risk-averse and risk-friendly decisions converge and decrease in expected loss.
+These degree of these processes is considered a measure for the state of knowledge and decision uncertainty at the decision-making moment.
+This decision uncertainty does not change aligned with model uncertainty, but depends on alterations of critical parameter and interdependencies, relating to seal reliability in particular.
+Additionally, actors affected differently by one set of information, as defined by their risk-opportunity weighting.
+For good decision making, it appears to be most important to recognize in for which parameters reduction of uncertainty leads to changes in decisions.
 
----
-Please pay particular attention to the preparation of your abstract; use this text as a guide.
+%Please pay particular attention to the preparation of your abstract; use this text as a guide.
 Every master thesis report must be accompanied by an informative abstract of no more than one paragraph (max 300 words).
@@
     	\centering
-    	\includegraphics[width=1\textwidth]{Figures/Appendix/ML4}
-    	\caption{....}\label{fig:ML4}
+    	\includegraphics[width=1\textwidth]{Figures/Appendix/ML3}
+    	\caption{....}\label{fig:ML3}
     \end{figure}

### 2017-11-24 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Abstract (needs revision)

@@
 Fault gouge properties and ultimately the variation in permeability in the fault zone is primarily dependent on clay content.
-Clay smearing has thus been a central topic of research surrounding fault sealing (see \citet{lindsay1993outcrop, yielding1997quantitative, van2003lateral, van2005processes, schmatz2010clay}.
+Clay smearing has thus been a central topic of research surrounding fault sealing (see \citet{lindsay1993outcrop} \citet{yielding1997quantitative}, \citet{van2003lateral} , \citet{van2005processes} and \citet{schmatz2010clay}).
 It generally is defined as all processes that lead to the incorporation of clay from the wall rock (e.g.
@@
 Any larger values come with a decrease in confidence \citep{faerseth2006shale}.
-A critical review of the research on clay smearing published over the last four decades was recently presented by \citet{vrolijk2016clay}.
+A critical review of the research on clay smearing published over the last four decades was recently presented by \citet{vrolijk2016clay}.\\
         The 3D structural model in this work (see Section \ref{sec:3D_design}) is designed in a way that a potential conventional hydrocarbon trap is included in the form of anticlinal folding combined with normal faulting of a reservoir-seal sequence.
@@
 				\caption{Probability distributions for positions of layer boundaries in the subsurface and a respective representation using lines.
-These normal distributions are determinedby the following values: Seal top: $\mu = 2000 m, \sigma = 7$; reservoir top: $\mu = 2050 m, \sigma = 25$; reservoir bottom: $\mu = 2200 m, \sigma = 45$.
+These normal distributions are determined by the following values: Seal top: $\mu = 2000 m, \sigma = 7$; reservoir top: $\mu = 2050 m, \sigma = 25$; reservoir bottom: $\mu = 2200 m, \sigma = 45$.
 }\label{fig:1D_model}
@@
 	        	\includegraphics[width=1\textwidth]{Figures/3_parameters.png}
-	        	\caption{Probability distributions for the three factors used in the abstract reservoir valuation as formulated in Equation \ref{eq:1D_score_system}.
+	        	\caption{Probability distributions for the three factors used in the score-based reservoir valuation as formulated in Equation \ref{eq:1D_score_system}: (A) reservoir thickness $h_{res}$, (B) reservoir top depth $d_{res}$ and (C) seal thickness $h_{seal}$.
 These are deduced from the 1D model input parameters and are thus uncertain.
-Here, we show factors deduced from prior parameters in Figure \ref{fig:1D_model} and factors resulting from exemplary posterior parameter distributions.}\label{fig:3_parameters}
+Here, we show a comparison between the three factors as deduced from prior parameters in Figure \ref{fig:1D_model} and as determined from exemplary posterior parameter distributions.}\label{fig:3_parameters}
 	        \end{figure}
@@
 		\subsection{Determination of the trap volume}
-		In the course of this work, several algorithms are developed within the Python environment, that in combination enable the automatic recognition and calculation of trap volumes in geological models computed by GemPy.
+		In the course of this work, several algorithms are developed within the Python environment.
+This set of functions enables the automatic recognition and calculation of trap volumes in geological models computed by GemPy.
 To assign voxels of the model to the trap volume, it is checked whether the following conditions (illustrated in Figure \ref{fig:trap_recognition}) are satisfied by each particular voxel:
@@
 This case is dealt with independently, by checking this area of juxtaposition with regard to its $SSF$ and potential fault sealing, as is further elaborated in Section \ref{sec:check_juxta} below.\\
-			Following these observations, the algorithm developed in this work uses differentiation techniques to find local extrema in the 2D reservoir top surface arrays attained over GemPy.
-Points where local minima and maxima of perpendicular axes X and Y coincide are then defined as saddle points.
+			Following these observations, the algorithm developed in this work uses a function from the Python library Scipy to find local extrema in the 2D reservoir top surface arrays attained over GemPy.
+This function returns the position of relative extrema in an array, by comparing every grid value with a variable number of neighbors along the same row or column.
+Points where local minima and maxima of perpendicular axes X and Y coincide in an array, are then defined as saddle points.
 The anticlinal spill point is deduced as the maximal saddle point in the footwall.

### 2017-11-24 Risk analysis/Text/Layout/Layout/chapter3_results.tex
Abstract (needs revision)

@@
 		\subsection{Posterior model II: Likely thin reservoir formation}%ML3
-		For the second model, thickness likelihoods were defined in a way that reinforced the probability of a very thick seal but at the same time restricted the reservoir formation to a likely significantly thinner thickness of around 50 m (see Table in Appendix).
+		\begin{table}[h]
+			\centering
+			\begin{tabular}{lrr} 
+				\toprule
+				Model II - Likelihoods: layer thicknesses (normal distributions)\\  
+				\midrule 
+				Formation & $\mu$ [m] & $\sigma$ [m]\\ 
+				\midrule 
+				Sandstone 2 & 300 & 30 \\
+				Shale (Seal) & 400 & 40\\ 
+				Sandstone 1 (Reservoir) & 50 & 10 \\
+				\bottomrule
+			\end{tabular}
+			\caption{Layer thickness likelihoods determined by normal distributions as used for posterior model II.}
+			\label{tab:ML3_likelihoods}
+		\end{table}
+		For the second model, thickness likelihoods were defined in a way that reinforced the probability of a very thick seal but at the same time restricted the reservoir formation to a likely significantly thinner thickness of around 50 m (see Table \ref{tab:ML3_likelihoods}).
 The results (see Figure \ref{fig:ML3}) are very similar, in fact almost identical, to those of model I above.
@@
 		\end{figure}
-		Using likelihoods which favor a thick reservoir, but only a thin seal of about 50 m thickness, leakage due to juxtaposition with seal-overlying formations across the fault ("LEAK OVER") became the primary mechanism for trap volume control (see Figure \ref{fig:ML2} (c).
+		\begin{table}[h]
+			\centering
+			\begin{tabular}{lrr} 
+				\toprule
+				Model III - Likelihoods: layer thicknesses (normal distributions)\\  
+				\midrule 
+				Formation & $\mu$ [m] & $\sigma$ [m]\\ 
+				\midrule 
+				Sandstone 2 & 300 & 30 \\
+				Shale (Seal) & 50 & 10\\ 
+				Sandstone 1 (Reservoir) & 400 & 30 \\
+				\bottomrule
+			\end{tabular}
+			\caption{Layer thickness likelihoods determined by normal distributions as used for posterior model III.}
+			\label{tab:ML2_likelihoods}
+		\end{table}
+		Using likelihoods which favor a thick reservoir, but only a thin seal of about 50 m thickness (see Table \ref{tab:ML2_likelihoods}), leakage due to juxtaposition with seal-overlying formations across the fault ("LEAK OVER") became the primary mechanism for trap volume control (see Figure \ref{fig:ML2} (c).
 It can be recognized in Figure \ref{fig:ML2} (b) that, while uncertainty was reduced clearly in the footwall, it remains high in the hanging wall.
@@
 		\end{figure}
+		\begin{table}[h]
+			\centering
+			\begin{tabular}{lrr} 
+				\toprule
+				Model IV - Likelihood: layer thicknesses (normal distributions)\\  
+				\midrule 
+				Formation & $\mu$ [m] & $\sigma$ [m]\\ 
+				\midrule 
+				Sandstone 2 & 150 & 20 \\
+				Shale (Seal) & 300 & 30\\ 
+				Sandstone 1 (Reservoir) & 250 & 25 \\
+				\bottomrule
+				\toprule
+				Likelihood: Shale Smear Factor (normal distribution)\\
+				\midrule
+				- & $\mu$ & $\sigma$\\
+				\midrule
+				SSF & 1.5 & 0.5\\
+				\bottomrule
+			\end{tabular}
+			\caption{Layer thickness likelihoods determined by normal distributions as used for posterior model IV.}
+			\label{tab:ML4_likelihoods}
+		\end{table}
 		For this fourth model, thickness likelihoods were chosen to be comparable to those of model I, i.e.
@@
 Additionally, a $SSF$-related likelihood function was implemented.
-With a mean of 1.5 and a standard deviation of 0.5, it was chosen to mostly reinforce the likelihood of a $SSF$ below the critical value of $SSF_c = 3$.\\
+With a mean of 1.5 and a standard deviation of 0.5, it was chosen to mostly reinforce the likelihood of a $SSF$ below the critical value of $SSF_c = 3$ (see Table \ref{tab:ML4_likelihoods}).\\
 		In Figure \ref{fig:ML4} (b), it can be seen that the uncertainty about the position of interfaces in the footwall was greatly reduced for all layers.
@@
 		\end{figure}
+		\begin{table}[h]
+			\centering
+			\begin{tabular}{lrr} 
+				Model V - Likelihood: Shale Smear Factor (normal distribution)\\
+				\midrule
+				- & $\mu$ & $\sigma$\\
+				\midrule
+				SSF & 3 & 0.3\\
+				\bottomrule
+			\end{tabular}
+			\caption{SSF-likelihood determined by normal distributions as used for posterior model V.}
+			\label{tab:ML5_likelihoods}
+		\end{table}
 		In this last case example, no thickness likelihoods, only a $SSF$-related likelihood was used.
-In contrast to model IV, this was chosen to narrow the probability of the $SSF$ to be around the threshold of its critical value, by basing it on a normal distribution with a mean of 3 and a standard deviation of 0.3.\\
+In contrast to model IV, this was chosen to narrow the probability of the $SSF$ to be around the threshold of its critical value of $SSF_c = 3$ (see Table \ref{tab:ML5_likelihoods}).\\
 		Results are relatively similar to those of the prior model.

### 2017-11-24 Risk analysis/Text/Layout/Layout/chapter1.tex
Restructured discussion

@@
 \end{enumerate}
-This concept is relatable to a vast variety of problems, ranging from casual every-day situations to complex scenarios in large-scale economic decision making: As a private person, should I take an umbrella with me today?
+This concept can be related to a vast variety of problems, ranging from casual every-day situations to complex scenarios in large-scale economic decision making: As a private person, should I take an umbrella with me today?
 As a company, should we invest in the development and realization of a certain project?
@@
 Probabilistic approaches to decision making have also become prevalent in hydrocarbon exploration and production \citep{murtha1997monte, mudford2000valuing,bratvold2010making}.
-However, in this sector, the methods are mainly limited to conducting simple Monte Carlo simulations and results are often interpreted viewing only percentiles and single values such as the mean.
+However, in this sector, the methods are mainly limited to conducting simple Monte Carlo simulations and results are often interpreted viewing only percentiles and single values, such as the mean.
 Bayesian inference and Markov chain Monte Carlo methods have only recently received more attention by researchers in this field (see \citet{wadsley2005markov}, \citet{ma2006multistage} and \citet{liu2010continuous}).\\

### 2017-11-24 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Restructured discussion

@@
         	\item \textbf{Observed data ($y$)}: This is any type of additional information that can be related to the forward modeling results or to the parameters and their combinations, and might possibly be used to reduce uncertainty.
-Such data can be gained by measurements and observations, for example by core sampling, well-log analysis or seismic acquisition.
+Such data can be gained by measurements and observations, but is typically inadequate for utilization as model input parameters.
+Examples are gravity or magnetic measurements, but also information attained by core sampling, well-log analysis or seismic acquisition.
         	\item \textbf{Likelihood functions $p(y|\theta)$}: Links between the previous parameters $\theta$ and the additional data $y$ are established by these functions in a way that they reflect the likelihood of the parameter states given the observations.
@@
 In the face of several uncertain parameters, a perfectly true estimate is virtually unattainable.
-However, an attempt can be made to design a customized loss function that returns a Bayesian estimator involving the least bad consequences for an individual in a specific environment.
-Assuming the reservoir setting case and valuation methods introduced in Section \ref{sec:Reservoir_values}), such a customization attempt is made and explained step by step in the following.\\
+However, an attempt can be made to design a customized loss function that returns a Bayesian estimator involving the least bad consequences for an individual in a specific environment \citep{davidson2015, hennig2007}.
+Assuming the reservoir setting case and valuation methods introduced in Section \ref{sec:Reservoir_values}, such a customization attempt is made and explained step by step in the following.\\
 		%%%(SOURCES?)
@@
 This allocation is the decision to be made or action to be taken.
+We hereby follow the definition by \citet{bratvold2010making}, who, in the context of hydrocarbon exploration and production, referred to a decision as a "conscious, irrevocable allocation of resources to achieve desired objectives".
+Thus, estimating the true value is deciding.
 Deviations from the unknown true value in the form of over- and underestimation bring about an error and loss accordingly.
-In this case, it is assumed that investments increase linearly with linear growth in the value of the reserve.
+In this case, we assume that investments increase linearly with linear growth in the value of the reserve.
 For this reason, the absolute-error loss function is favored here over the squared-error loss function.
@@
 Fundamentally different loss functions can also be based on a significantly different mathematical structure.
-Loss functions are customized regarding the problem environment and according to the subjective needs and objectives of the decision maker.
+Loss functions are customized regarding the problem environment and according to the subjective needs and objectives of the decision maker \citep{davidson2015, hennig2007}.
 Thus, they are mostly defined by the actor expressing his perspective.
@@
 Especially considering individual persons as actors, psychological aspects may play a significant role.
+\citet{bratvold2010making} provided an overview on how biases can influence and impair the decision making of otherwise smart actors.
 

### 2017-11-24 Risk analysis/Text/Layout/Layout/chapter3_results.tex
Restructured discussion

@@
 Several representative examples of posterior models based on the use of different likelihoods are referenced to the original prior model.
-For all model simulations, 1000 sampling steps were conducted, with a burn-in phase of 50 iterations for MCMC sampling.
+For model simulations, 1000 sampling steps were conducted, with a burn-in phase of 50 iterations for MCMC sampling.
+However, for computational reasons, Shannon entropy visualization was based on 200 sampling steps and a burn-in phase of 50 iterations.
 The following measures are central to the comparison of results: (1) information entropy, (2) occurrence of trap control mechanisms, (3) resulting recoverable oil volumes and (3) consequent realization of expected losses and related Bayes actions (i.e.
@@
 
-		\subsection{Prior model}
+		\subsection{Prior model}\label{sec:prior_model3D}
 		To generate a set of reference results, a prior model was first generated by running a simple simulation of Monte Carlo sampling considering only the uncertainty of prior parameters.
@@
 
-		\subsection{Posterior model IV: SSF likelihood reinforcing trap reliability}%ML4
+		\subsection{Posterior model IV: SSF likelihood reinforcing trap reliability}\label{sec:model4}%ML4
 		\begin{figure}[p!]
@@
 
-		\subsection{Posterior model V: Controversial information about the Shale Smear Factor}%ML5
+		\subsection{Posterior model V: Controversial information about the Shale Smear Factor}\label{sec:model5}%ML5
 		\begin{figure}[p!]

### 2017-11-24 Risk analysis/Text/Layout/Layout/chapter4_discussion.tex
Restructured discussion

@@
 These methods were applied for a conceptual 1D geological model and subsequently for a 3D structural model.
-	Despite a significant leap in complexity, the results from both models proved to be widely similar.\\	
+	Despite a significant leap in complexity, the results from both models proved to be widely similar.
+	\subsubsection{State of knowledge, decision uncertainty and consequent decision making via loss function optimization}
 	The economic value realized in each model construction (score and $ROV$, respectively), had been defined to depend on model parameters, either directly or indirectly by taking into account their relations to other parameters or to specific thresholds and conditions.
@@
 Pronounced bimodality between two extremes, i.e.
-high decision uncertainty, resulted in a higher spread of Bayes estimators.
+high decision uncertainty, resulted in a wieder separation of Bayes estimators.
 Reduction of the distribution to one mode conversely led to the convergence of different Bayes actions.
@@
 It can be assumed, that given perfect information, all actors would bid on the same estimate (the true value) and expect no loss, since no risk is present.
-It furthermore follows from this that the relevance of risk factors decreases with higher reduction of decision uncertainty.\\
+It furthermore follows from this that the relevance of risk factors decreases with higher reduction of decision uncertainty.
+	\subsubsection{Impact of additional information on decision making}
 	Cases of various states of knowledge and decision uncertainties, i.e.
@@
 Consequently, it was observed that reducing the uncertainty in a way that narrowed the probability of a threshold-related parameter around its cut-off value, led to an amplification of the respective mode and thus emphasized the risk of complete failure.
-This can be seen in the posterior $ROV$ distribution of model V (3D).
+This can be seen in the posterior $ROV$ distribution of 3D model V (see Figure \ref{fig:ML5}).
 Resulting Bayes actions were characterized by an increase in divergence and expected losses, indicating a deterioration of the state of knowledge and an according increase in decision uncertainty.
-Thus, in some cases, implementing Bayesian inference might leave actors in greater disagreement than before.\\
+Thus, in some cases, adding information to the inference process might leave actors in greater disagreement than before.
 	However, we furthermore have to consider that actors weight possible outcomes of the probability distribution differently.
@@
 This is less relevant for risk-friendly decisions makers, who respectively might acquire a comparable benefit from knowing more about the probability of positive outcomes, as they are less discouraged from overestimating and less afraid of failure.
-It was stated by \citet{bratvold2010making} that uncertainty has to possible consequences: risk and opportunity.
+It was stated by \citet{bratvold2010making} that uncertainty has two possible consequences: risk and opportunity.
 According to them, risk is a consequence which is subjectively undesirable to an actor, while opportunity is the opposite.
@@
 Nevertheless, it is suggested by the results from the different inference cases, that the best states of knowledge are attained by eliminating multimodality and narrowing the range of probabilities to one mode anywhere along the value axis.\\
+	Crucial risks might be easily assessed, if they are dependent on only one or a few parameters, such as seal thickness in the 1D model.
+In other cases, they are derived from more complex parameter inter-relations, as is the case for the Shale Smear Factor, which is the ratio of displacement to seal thickness.
+3D models IV and V (see Sections \ref{sec:model4} and \ref{sec:model5}) showed that reduced uncertainty about the $SSF$ is not necessarily directly recognizable via Shannon entropy visualization, especially if uncertainty regarding its parent parameters remains high.
+It follows that, to approach an effective mitigation of high risks, the complexities behind decisive factors need to be assessed thoroughly and respective parent parameters, as well as their interdependencies, need to be identified.
+This might enable a better understanding of which type of information is missing and where in the model additional data might be of use for improved decision-making.\\
 	%The respective estimate resilient to risk was generally found close to the positive mode and experienced only minor lateral shifting, after diminishing the probability of failure.
@@
 This is reflected by our results which showed that one set of information might have a vastly different impact on the Bayes estimators of different actors.
-New observations might lead to a great change in decision for one actor, but be negligible the other.\\
+New observations might lead to a great change in decision for one actor, but be negligible the other.
 		%This state of knowledge and the decision uncertainty accordingly experienced alterations related to the different Bayesian inference cases.
@@
 reducing the standard deviation), or the opposite, a diminishment of the mode up to total erasure of its probability.\\
-	Effects of Bayesian inference were reflected in changing appearances of modes in the reservoir value distributions.
+	\subsubsection{Value distribution modes as possible model scenarios and potential decision options}	
+	Effects of adding information via Bayesian inference were reflected in changing appearances of modes in the reservoir value distributions.
 Considering that these changes correlated with increased or decreased relevance of specific trap control mechanisms in the 3D model, we derive that each mode represents the occurrence of a specific mechanism-related scenario.
@@
 In the context of hydrocarbon exploration and production, this would relate to fixed magnitudes of resource allocation, such as a certain number of required drilling wells or the size of a production platform.
-A discretization of decision subregions in a probability distribution could thus also be based on previously defined actual options.\\
+A discretization of decision subregions in a probability distribution could thus also be based on previously defined actual options.
+	\subsubsection{Comparison to common practices and other research in the hydrocarbon sector}		
 	In remains to be argued, to what extent our methodology and findings contribute regarding practical applications in hydrocarbon exploration and production.
@@
 While this is obviously unrealistic, we presume that an elaborate loss function might at least provide a very good preliminary decision recommendation.
-It might furthermore weight \\
+It might furthermore be able to weight risks that are not immediately apparent to an individual.
+Furthermore, the influence of human biases and psychological behavioral challenges, as described by \citet{bratvold2010making}, could be avoided.\\
 	More recently, Bayesian inference and MCMC methods were applied for OOIP estimation and forecasting of reservoir productivity by \citet{wadsley2005markov}, \citet{ma2006multistage} and \citet{liu2010continuous}.
@@
 	\citet{bratvold2010making} highlighted the potential to use Monte Carlo simulation to identify, via sensitivity analysis, which uncertainties have the greatest impact on decisions.
-This was virtually achieved by performing several cases of Bayesian inference in our work, as we recognized primarily the importance of seal reliability.\\
-	As follows from our elaborations above, it is important to recognize crucial risks resulting from uncertainty.
-Such risks might be easily assessed, if they are dependent on only one or a few parameters, such as seal thickness in the 1D model.
-In other cases, they are derived from more complex parameter inter-relations, as is the case for the Shale Smear Factor, which is the ratio of displacement to seal thickness.
-Models IV and V (3D) showed that reduced uncertainty about the $SSF$ is not necessarily directly recognizable via Shannon entropy visualization, especially if uncertainty regarding its parent parameters remains high.
-It follows that, to approach an effective mitigation of high risks, the complexities behind decisive factors need to be assessed thoroughly and respective parent parameters, as well as their interdependencies, need to be identified.
-This might enable a better understanding of which type of information is missing and where in the model additional data might be of use for improved decision-making.\\
+This was virtually achieved by performing several cases of Bayesian inference in our work, as we recognized primarily the importance of seal reliability.
+	\subsubsection{Representativity and limitations of this methodology}
 	It has too be emphasized that the models constructed in this work were synthetic and not based on real data.
@@
 The consideration of more details, without a basis on real data, would have furthermore required extensive speculation, which would have presumably impaired the generalization potential of respective observations.
-While standard loss functions only return mean or median as estimates, loss function customization offers a flexible approach to express different objective and subjective weighting factors that are specified to certain decision-making environments and decision-makers' perspectives.\\
+While standard loss functions only return mean or median as estimates, loss function customization offers a flexible approach to express different objective and subjective weighting factors that are specified to certain decision-making environments and decision-makers' perspectives.
+	\subsubsection{Possibilities for future research and extensions}	
 	Considering the findings of this work, there is still many points that could be expanded on in future research.

### 2017-11-25 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Cleaning Notebooks2

@@
         	\centering
-        	\includegraphics[width=1\textwidth]{Figures/loss_functions.png}
+        	\includegraphics[width=1\textwidth]{Figures/loss_functions2.png}
         	\caption{Expected loss and respective Bayes actions based on the standard absolute-error loss (A) and squared-loss function (B).
-In this example, we applied the functions on a normal distribution characterized by $\mu = 0$ and $\sigma = 500$.}\label{fig:standard_LF}
+In this example, we applied the functions on a normal distribution characterized by $\mu = 0$ and $\sigma = 500$ (shown in light blue).}\label{fig:standard_LF}
         \end{figure}
@@
 			\centering
-			\includegraphics[width=1\textwidth]{Figures/LF_4steps_normal.png}
-			\caption{Single steps (I-IV) of customizing the loss function applied on an exemplary normal distribution ($\mu=0; \sigma=500$).
+			\includegraphics[width=1\textwidth]{Figures/LF_4steps_normal2.png}
+			\caption{Single steps (I-IV) of customizing the loss function applied on an exemplary normal distribution ($\mu=0; \sigma=500$; illustrated n light blue).
 The median (in this case also mean and average) is returned for the Bayes action, when using the standard symmetric absolute-error loss function (I).
@@
 			\caption{Plotting of expected loss realizations after including the risk factor $r$ in the loss function (Equation \ref{eq:LFR_final}) for actors with risk-affinities ranging from risk-averse ($r$ = 0.5 and 0.75), over risk-neutral ($r$ = 1), to risk-friendly ($r$ = 1.25 and $r$ = 1.5).
-The loss function was applied on a normal distribution around zero ($\mu=0; \sigma=500$).}\label{fig:LFR_normal}
+The loss function was applied the same normal distribution ($\mu=0; \sigma=500$) as in Figure \ref{fig:LF_4steps_normal}.}\label{fig:LFR_normal}
 		\end{figure}

### 2017-11-25 Risk analysis/Text/Layout/Layout/chapter3_results.tex
Cleaning Notebooks2

@@
 				%\end{table}
-
 			%\subsection{General 1D model inference results}
@@
 			\includegraphics[width=1\textwidth]{Figures/ML1}
-			\caption{....}\label{fig:ML1}
+			\caption{Summary of results from evaluating posterior model I.
+Shannon entropies in prior (a) and posterior (b) are compared.
+So are the mechanisms of maximum trap volume control (c).
+Prior and posterior ROV probability distributions are shown in (d).
+Respective loss function and Bayes actions are plotted in (e).}\label{fig:ML1}
 		\end{figure}		
@@
 			\includegraphics[width=1\textwidth]{Figures/ML2}
-			\caption{....}\label{fig:ML2}
+			\caption{Summary of results from evaluating posterior model III.
+Shannon entropies in prior (a) and posterior (b) are compared.
+So are the mechanisms of maximum trap volume control (c).
+Prior and posterior ROV probability distributions are shown in (d).
+Respective loss function and Bayes actions are plotted in (e).}\label{fig:ML2}
 		\end{figure}
@@
 			\includegraphics[width=1\textwidth]{Figures/Appendix/ML4}
-			\caption{....}\label{fig:ML4}
+			\caption{Summary of results from evaluating posterior model IV.
+Shannon entropies in prior (a) and posterior (b) are compared.
+So are the mechanisms of maximum trap volume control (c).
+Prior and posterior ROV probability distributions are shown in (d).
+Respective loss function and Bayes actions are plotted in (e).}\label{fig:ML4}
 		\end{figure}
@@
 			\end{tabular}
-			\caption{Layer thickness likelihoods determined by normal distributions as used for posterior model IV.}
+			\caption{Layer thickness likelihoods and SSF-likelihood determined by normal distributions as used for posterior model IV.}
 			\label{tab:ML4_likelihoods}
@@
 			\includegraphics[width=1\textwidth]{Figures/ML5}
-			\caption{....}\label{fig:ML5}
+			\caption{Summary of results from evaluating posterior model V.
+Shannon entropies in prior (a) and posterior (b) are compared.
+So are the mechanisms of maximum trap volume control (c).
+Prior and posterior ROV probability distributions are shown in (d).
+Respective loss function and Bayes actions are plotted in (e).}\label{fig:ML5}
 		\end{figure}
@@
 			\begin{tabular}{lrr} 
+				\toprule
 				Model V - Likelihood: Shale Smear Factor (normal distribution)\\
@@
 			\end{tabular}
-			\caption{SSF-likelihood determined by normal distributions as used for posterior model V.}
+			\caption{SSF-likelihood determined by a normal distributions as used for posterior model V.}
 			\label{tab:ML5_likelihoods}

### 2017-11-25 Risk analysis/Text/Layout/Layout/chapter4_discussion.tex
Cleaning Notebooks2

@@
 	\subsubsection{Value distribution modes as possible model scenarios and potential decision options}	
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=1\textwidth]{Figures/scenarios_options}
+		\caption{ROV distribution of the prior model subdivided by modes into scenarios which could pose decision options (marked by greens lines).
+Red lines indicated a discretization of decision options according fixed scales of resource allocations for a project.
+For both cases, zero equals the option to "take no action", i.e.
+allocate no resources.}\label{fig:scenarios_options}
+	\end{figure}
 	Effects of adding information via Bayesian inference were reflected in changing appearances of modes in the reservoir value distributions.
@@
 A discretization of decision subregions in a probability distribution could thus also be based on previously defined actual options.
+A conceptual illustration of this is shown in Figure \ref{fig:scenarios_options}.
 	\subsubsection{Comparison to common practices and other research in the hydrocarbon sector}		

### 2017-11-25 Risk analysis/Text/Layout/Layout/chapter1.tex
Revision, in methods part

@@
 Probabilistic approaches to decision making have also become prevalent in hydrocarbon exploration and production \citep{murtha1997monte, mudford2000valuing,bratvold2010making}.
-However, in this sector, the methods are mainly limited to conducting simple Monte Carlo simulations and results are often interpreted viewing only percentiles and single values, such as the mean.
-Bayesian inference and Markov chain Monte Carlo methods have only recently received more attention by researchers in this field (see \citet{wadsley2005markov}, \citet{ma2006multistage} and \citet{liu2010continuous}).\\
-In geosciences, Bayesian inference has prominently found use in the context of geophysical inversion problems (see \citet{tarantola1982inverse, mosegaard2002probabilistic} and \citet{sambridge2002monte}).
-Recently, it was transferred by \citet{delaVarga2016} to the field of structural geological modeling.
-This has been enabled by progressing developments regarding implicit geological modeling functions based on interpolation \citep{hillier2014three, mallet1992discrete, lajaunie1997foliation} and the possibility of fully automated model reconstruction in particular.
+However, in this sector, the methods are mainly limited to conducting simple Monte Carlo simulations and results are often interpreted considering only percentiles and single values, such as the mean.
+Bayesian inference and Markov chain Monte Carlo methods (MCMC) have only recently received more attention by researchers in this field (see \citet{wadsley2005markov}, \citet{ma2006multistage} and \citet{liu2010continuous}).\\
+In general geosciences, Bayesian inference has prominently found use in the context of geophysical inversion problems (see \citet{tarantola1982inverse, mosegaard2002probabilistic} and \citet{sambridge2002monte}).
+Recent developments have transferred this probabilistic approach to address uncertainties the field of structural geological modeling.
+\citet{Caumon2010}, \citet{jessell2010towards} and \citet{Wellmann2010b} developed methods of geologically based inversion by considering secondary geological observations to constrain uncertain models.
+Advancements in this field have been enabled by progressing developments regarding implicit modeling functions based on interpolation and the possibility of fully automated model reconstruction in particular \citep{hillier2014three, mallet1992discrete, lajaunie1997foliation}.
 \citet{delaVarga2016} regarded geological modeling as a Bayesian inference problem by relating additional geological information to prior model parameters in the form of likelihood functions, linking them in a non-parametric Bayesian network.
-Using Markov chain Monte Carlo sampling to explore resulting probability spaces, they attained posterior model suites with reduced uncertainties \citep{delaVarga2016}\\
+Using Markov chain Monte Carlo sampling to explore resulting probability spaces, they attained posterior model suites with reduced uncertainties \citep{delaVarga2016}.\\
 This work builds upon their concept, exploring the potential significance their findings might have in the context of decision making.
@@
 This is achieved by assigning an economic meaning to the structural model and designing a case-specific custom loss function to find decisions which are optimal related to the state of knowledge and the preferences of actors with different risk-affinities.
-More specifically, the models are designed to represent potential petroleum systems.
-Consequently, the development of algorithms for automatic hydrocarbon trap recognition and volume calculation represent a central part of this work.\\
+More specifically, the models are designed to represent potential hydrocarbon reservoir systems.
+Consequently, the development of algorithms for automatic structural trap recognition and volumetric calculation represents a central part of this work.\\
 The main hypothesis of this work is that Bayesian inference and resulting changes in uncertainties in a geological setting have a significant effect on reservoir value estimation and decision making.
-We furthermore postulate that loss functions can be customized to appropriately represent preferences of actors in the hydrocarbon sector and moreover illustrate the nature of decisions such actor's might make depending on their individual attitudes towards risk and in the face of different types of uncertainties.
-Changes in their respective decisions are treated as a suitable measure to assess the effect of updating model parameters with new geological information.
+We furthermore postulate that loss functions can be customized to appropriately represent preferences of actors in the hydrocarbon sector and moreover illustrate the nature of decisions such actors might make depending on their individual attitudes towards risk and in the face of different types of uncertainties.
+Changes in their respective decisions are treated as a suitable measure to assess the effect of updating model parameters with additional geological information.

### 2017-11-25 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Revision, in methods part

@@
 Bayesian analysis and decision theory are introduced in Section \ref{sec:bayes}.
-Focus is laid on Bayesian inference, estimation of uncertain values and the use of loss functions in this context.
+Focus is laid on Bayesian inference, estimation of uncertain values and the use of loss functions.
 We apply these methods in the context of structural geological modeling.
@@
 Regarding numerical implementation, GemPy and PyMC are presented as central tools used in a Python environment.
-In Section \ref{sec:model_evaluation}, we outline our methods to evaluate consequent modeling results are.
-Particularly the potential for valuation from an economic perspective is elaborated in this regard and basic principles of subsurface hydrocarbon systems are introduced.
-Considering the respective setting in the economic sector of petroleum production, the development of an according case-specific custom loss function is described in Section \ref{sec:LF_design}.
-We first apply these methodologies on a conceptual 1D geological model (see Section \ref{sec:1D_model}), subsequently on a full 3D structural geological model.
-Design, construction and methods to identify hydrocarbon trap features in such the 3D model are explained in Section \ref{sec:3D_model}.
+In Section \ref{sec:model_evaluation}, we outline our methods to evaluate consequent modeling results.
+Particularly the potential for valuation from an economic perspective is elaborated and basic principles of subsurface hydrocarbon systems are introduced.
+Considering the respective setting in the economic sector of petroleum exploration and production, the development of an according case-specific custom loss function is described in Section \ref{sec:LF_design}.
+We first apply these methods to a conceptual 1D geological model (see Section \ref{sec:1D_model}), subsequently on a full 3D structural geological model.
+Design, construction and methods to identify hydrocarbon trap features in such a 3D model are explained in Section \ref{sec:3D_model}.
 
@@
 In order to find the best decision to a problem, it is possible to combine sample information with other aspects such as the possible consequences of decision making and the availability of prior information on our uncertainties.
-Decision consequences are expressed as gains in economic decision theory and as losses, which equal negative gains, in statistics.
+Decision consequences are expressed as gains in economic decision theory and as losses, which equal negative gains, in statistics \citep{berger2013stat}.
 Prior information might be given for example based on experience from previous similar problems or from expert knowledge \citep{bratvold2010making}.
-The approach of utilizing priors is known as Bayesian analysis, which is explained in the following \citep{berger2013stat}.
+The approach of utilizing priors is known as Bayesian analysis, which is explained in the following.
 %(It goes well with decision theory.)
@@
         \subsection{Bayesian inference}
-        Bayesian inference is most importantly characterized by its preservation of uncertainty, in contrast to standard statistical inference \citep{jaynes2003probability, box2011bayesian, harney2013bayesian, davidson2015}.
+        Bayesian inference is most importantly characterized by its preservation of uncertainty, in contrast to standard statistical inference \citep{jaynes2003probability, gelman2014bayesian, box2011bayesian, harney2013bayesian, davidson2015}.
 The Bayesian approach is widely seen as intuitive and inherent in the natural human perspective.
@@
 This means that Bayesian updating is about reducing uncertainty in a belief and reaching a guess that is less wrong.
-Bayesian inference is defined by and conducted via the following equation, called the Bayes' Theorem \citep{jaynes2003probability, box2011bayesian, harney2013bayesian, davidson2015}:
+Bayesian inference is defined by and conducted via the following equation, called the Bayes' Theorem \citep{jaynes2003probability, gelman2014bayesian, box2011bayesian, harney2013bayesian, davidson2015}:
         \begin{equation}\label{eq:BayesTheorem}
@@
 Two standard loss functions are the absolute-error and the squared-error loss function.
-Both are simple to understand and commonly used \citep{davidson2015}.\\
-        As implied by its name, the absolute-error loss function returns loss as the absolute error, i.e. the difference between the estimate and the true parameter \citep{davidson2015}:        
+Both are simple to understand and commonly used \citep{davidson2015, hennig2007}.\\
+        As implied by its name, the absolute-error loss function returns loss as the absolute error, i.e. the difference between the estimate and the true parameter:        
         \begin{equation}\label{eq:AbsLossFunction}
@@
 overestimation) is thus incorporated in an uncomplicated way \citep{davidson2015, hennig2007}.
-Much more complicated designs of loss functions are possible, depending on purpose, objective and application \citep{davidson2015}.
+Much more complicated designs of loss functions are possible, depending on purpose, objective and application.
 A case-specific loss functions is designed in Section \ref{sec:LF_design} of this work.\\
-        The presence of uncertainty during decision making implies that the true parameter is unknown and thus the truly incurred loss $L(\theta,a)$ cannot be known at the time of making the decision \citep{berger2013stat, davidson2015}.
+        The presence of uncertainty during decision making implies that the true parameter is unknown and thus the truly incurred loss $L(\theta,a)$ cannot be known at the time of making the decision.
 The Bayesian perspective considers unknown parameters as random variables and samples that are drawn from the posterior distribution as possible realizations of the unknown parameter, i.e.
-all possible true values are represented by this distribution \citep{davidson2015}.
-A probabilistic alternative to the actual loss is to consider each decision's expected loss and to make a decision that is optimal in relation to this expected loss \citep{berger2013stat}.
-\\
+all possible true values are represented by this distribution.
+A probabilistic alternative to the actual loss is to consider each decision's expected loss and to make a decision that is optimal in relation to this expected loss \citep{berger2013stat, davidson2015}.\\
         Given a posterior distribution $p(\theta|y)$, the expected loss of choosing an estimate $\hat{\theta}$ over the true parameter $\theta$ (after evidence $y$ has been observed) is defined by the function below \citep{davidson2015}:
@@
         	\includegraphics[width=1\textwidth]{Figures/loss_functions2.png}
-        	\caption{Expected loss and respective Bayes actions based on the standard absolute-error loss (A) and squared-loss function (B).
-In this example, we applied the functions on a normal distribution characterized by $\mu = 0$ and $\sigma = 500$ (shown in light blue).}\label{fig:standard_LF}
+        	\caption{Expected loss for estimates of the true value and Bayes actions based on the standard absolute-error loss (A) and squared-loss function (B).
+In this example, the probability distribution of the true value is normal around zero ($\mu = 0$; $\sigma = 500$; shown in light blue).
+It can be seen that expected loss increase exponentially with distance from zero for the squared-error loss function.}\label{fig:standard_LF}
         \end{figure}
@@
         The expectation symbol $E$ is subscripted with $\theta$, by which it is indicated that $\theta$ is the respective unknown variable.
-This expected loss $l$ is also referred to as the Bayes risk of estimate $\hat{\theta}$ \citep{berger2013stat, davidson2015}.
-Exemplary realizations of the expected loss according to both standard loss functions and minimized by the Bayes action are plotted in Figure \ref{fig:standard_LF}.\\
+This expected loss $l$ is also referred to as the Bayes risk of estimate $\hat{\theta}$ \citep{berger2013stat, davidson2015}.\\
         By the Law of Large Numbers, the expected loss of $\hat{\theta}$ can be approximated drawing a large sample size $N$ from the posterior distribution, respectively applying a loss function $L$ and averaging over the number of samples \citep{davidson2015}:
@@
         Minimization of a loss function returns a Bayesian point estimate known as Bayes action or Bayesian estimator $\delta^p(y)$, which is the estimate, action or decision with the least expected loss according to the loss function \citep{berger2013stat, moye2006statistical}.
-For a unimodal and symmetric absolute-error loss function, the Bayes action is simply the median of the posterior distribution, while using squared-error loss it is the mean \citep{davidson2015, berger2013stat}.
-The MAP (maximum a posteriori) estimate is the minimizing solution for the posterior using zero-one loss \citep{davidson2015}.
-The possibility of more than one minimum also implies that several Bayes actions can exist for one problem \citep{berger2013stat}.\\
+For a unimodal and symmetric absolute-error loss function, the Bayes action is simply the median of the posterior distribution, while using squared-error loss it is the mean.
+The MAP (maximum a posteriori) estimate is the minimizing solution for the posterior using zero-one loss.
+The possibility of more than one minimum also implies that several Bayes actions can exist for one problem \citep{davidson2015, berger2013stat}.
+Exemplary realizations of expected loss and Bayes actions according to both standard loss functions are plotted in Figure \ref{fig:standard_LF}.\\
         \citet{davidson2015} implemented different risk affinities by simply  introducing a risk parameter into the loss function.
@@
         \section{Application in structural geological modeling}\label{sec:bayes_in_struc_modeling}
-        In this work, these methods of Bayesian analysis and decision theory are applied in the field of geological structural modeling.
+        In this work, we apply these methods of Bayesian analysis and decision theory in the field of geological structural modeling.
 The fundamental approach follows closely the research conducted by \citet{delaVarga2016} and builds upon their findings.\\
@@
         	\item \textbf{Observed data ($y$)}: This is any type of additional information that can be related to the forward modeling results or to the parameters and their combinations, and might possibly be used to reduce uncertainty.
-Such data can be gained by measurements and observations, but is typically inadequate for utilization as model input parameters.
-Examples are gravity or magnetic measurements, but also information attained by core sampling, well-log analysis or seismic acquisition.
+This comprises mostly secondary data that can not be incorporated directly as input parameters, such as geophysical measurements (gravity and magnetics), information from well-log analysis or simply geological rules.
         	\item \textbf{Likelihood functions $p(y|\theta)$}: Links between the previous parameters $\theta$ and the additional data $y$ are established by these functions in a way that they reflect the likelihood of the parameter states given the observations.
@@
         	\item \textbf{Evaluation of the posterior model}: Depending on the aim of the study, a post-processing analysis can be conducted accordingly.
-\citet{delaVarga2016} focused on the examination of the posterior distributions of the parameters $\theta$ and the generated models, particularly regarding information entropy within the model space.
+\citet{delaVarga2016} focused on the examination of the posterior distributions of the parameters $\theta$ and the generated models, particularly regarding Shannon entropy which is explained here in Section \ref{sec:shannon_entropy}.
 In this work, the geological models are additionally assigned an economic meaning by declaring them potential petroleum reservoirs and introducing customized loss functions to reflect the economic interest of decision makers in developing respective resource extraction projects (see Section \ref{sec:loss}).
@@
 
-        \subsection{Visualizing uncertainty}
+        \subsection{Visualizing uncertainty}\label{sec:shannon_entropy}
         \begin{figure}
@@
 		\subsection{Including different risk-affinities in the loss function}		
-
 		It can be assumed that several actors in one sector or decision environment may have the same general loss function but different affinities concerning risks.

### 2017-11-27 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Revisions

@@
         	\caption{Expected loss for estimates of the true value and Bayes actions based on the standard absolute-error loss (A) and squared-loss function (B).
-In this example, the probability distribution of the true value is normal around zero ($\mu = 0$; $\sigma = 500$; shown in light blue).
-It can be seen that expected loss increase exponentially with distance from zero for the squared-error loss function.}\label{fig:standard_LF}
+In this example, the probability distribution of the true value is normal around zero ($\mu = 0$; $\sigma = 500$; shown in light blue).}\label{fig:standard_LF}
         \end{figure}
@@
 			\includegraphics[width=1\textwidth]{Figures/calcagno_pot_field.png}
-			\caption{Illustration of the concept of interpolating to attain a poentail field.
+			\caption{Illustration of the concept of interpolating to attain a potential field.
 The original data is depicted in (a), with contact points as dots and orientation measurements indicated by arrows.
@@
 An accordingly calculated potential field is shown in (b) (from \citet{calcagno2008geological}).}\label{fig:pot_field}
-		\end{figure}        
+		\end{figure}\\
         Resulting potential fields can be used to describe geological interfaces as iso-surfaces in any kind of 3D geometry \citep{calcagno2008geological}.
+A conceptual illustration is depicted in Figure \ref{fig:pot_field}.
 Fault geometries can be interpolated analogously.
@@
         It is pointed out by \citet{calcagno2008geological}, that this method is particularly appropriate for cases in which knowledge about the geology is only given for sparse locations and is thus applicable for a wide variety of typical problems in geological settings.
-Its suitability for this work is furthermore emphasized by the possibility to modify the topology-defining geological pile to achieve different geometric realizations without altering the data.
-The model can thus be updated in the face of new data or interpretations \citep{calcagno2008geological}.
+Furthermore, the model topology is defined by the knowledge contained in the geological pile (set of formations).
+Thus, the model can be updated in the face of new data or interpretations without the need to consider the impact on the geometry of formations, as this is automatically managed by the rules entailed in the geological pile \citep{calcagno2008geological}.
 
-		\subsection{Numerical computational implementation via Python, GemPy and PyMC}\label{sec:numerical_implementation}
+		\subsection{Numerical computational implementation}\label{sec:numerical_implementation}
 		Bayesian analysis can be conducted numerically using probabilistic programming methods \citep{salvatier2016pymc3}.
@@
 At the time of writing this, GemPy is still under development (version 0.995), but is already functioning for the purpose of this work.\\
-		For conducting the Bayesian analysis and embedding the geological modeling step in a probabilistic context, GemPy is integrated into the probabilistic framework of PyMC.
+		For conducting the Bayesian analysis and embedding the geological modeling step in a probabilistic context, we integrate GemPy into the probabilistic framework of PyMC.
 This Python library was developed for conducting Bayesian inference and prediction problems in an open-source probabilistic programming environment \citep{davidson2015, salvatier2016pymc3}.
 Different model fitting techniques are provided in PyMC, such as the \textit{maximum a posteriori} (MAP) method and several (MCMC) sampling methods, including the Adaptive Metropolis introduced in Section \ref{sec:mcmc}.
-The components which are used to construct a statistical model, are represented by \textit{Deterministic} or \textit{Stochastic} functions in PyMC \citep{salvatier2016pymc3}.
-These variables are described as hierarchically related nodes, specifically parent and child nodes, in a Bayesian network:
+The components which are used to construct a statistical model, are represented by \textit{Deterministic} or \textit{Stochastic} functions or variables in PyMC \citep{salvatier2016pymc3}.
+These are described as hierarchically related nodes, specifically parent and child nodes, in a Bayesian network:
 		\begin{enumerate}
 			\item \textbf{Parent} nodes include variables that affect other variables.
-			\item \textbf{Child} nodes contain variables that are influenced by other variables (they depend on parent variables) \citep{delaVarga2016}.
+			\item \textbf{Child} nodes contain variables that are influenced by other variables (they depend on parent variables) \citep{, ben2007bayesian, koller2009probabilistic}.
 		\end{enumerate}
@@
 Complex mathematical relations between \textit{Stochastic} variables can be described through \textit{Deterministic} functions \citep{delaVarga2016}.
-Futhermore, PyMC allows for the creation of own object definitions inheriting from the class descriptions of these two variable types.\\
+Futhermore, PyMC allows for the creation of own object definitions inheriting from the class descriptions of these two function and variable types.\\
 		Bayesian networks as a concept have been presented in more detail by \citet{koller2009probabilistic}.
@@
         After conducting Bayesian inference, posterior distributions and resulting model realizations are to be evaluated.
-For this, the sampling traces and posterior distributions of single parameters can be observed.
-The entirety of accepted 3D geological models (sampled after burn-in phase) is to be assessed in terms of uncertainty quantification.
+For this, we can examine sampling traces and posterior distributions of single parameters can be observed.
+The entirety of accepted 3D geological models (sampled after burn-in phase) is to be assessed in terms of uncertainty quantification and visualization.
 Furthermore, a central part of this work is the assignment of an economic significance to the modeling results.
-To achieve this, realized models are evaluated by regarding them as potential hydrocarbon systems.
-This process is further in Section \ref{sec:Reservoir_values}.
-Information entropy is introduced as a measure for uncertainty in the following.
+To achieve this, we valuate accepted models by regarding them as potential hydrocarbon systems.
+We explain this process in details in Section \ref{sec:Reservoir_values}.
+Shannon entropy is introduced as a measure for uncertainty in the following.
 
@@
 all outcomes are equally likely in each cell \citep{schaaf2017}.
-From this, the following equation for application in geological modeling was derived \citet{Mann1993,wellmann2012uncertainties}:
+From this, the following equation for application in geological modeling was derived \citep{Mann1993,wellmann2012uncertainties}:
         \begin{equation}
@@
         \begin{enumerate}
-        	\item \textbf{Source rock:} A kerogen-rich source rock which produced and emitted hydrocarbons during its burial history, given sufficient heating, is needed to provide oil or gas in the first place \citep{dolson2016basics}.
+        	\item \textbf{Source rock:} A kerogen-rich source rock which produced and emitted hydrocarbons during its burial history, given sufficient heating, is needed to provide oil or gas in the first place \citep{dolson2016basics, selley2014elements}.
 These are typically organic-rich sedimentary rocks, shales in particular.
 For the purpose of this work, it is assumed that a  suitable source rock is present and hydrocarbons expulsion took place after the formation of any trap-defining structures in our geological models.
-        	\item \textbf{Migration pathway:} A pathway that allows expelled oil and gas to migrate upwards through the subsurface from the source rock into potential reservoir formations is necessary \citep{dolson2016basics}.
+        	\item \textbf{Migration pathway:} A pathway that allows expelled oil and gas to migrate upwards through the subsurface from the source rock into potential reservoir formations is necessary \citep{dolson2016basics,north1985petroleum, selley2014elements}.
 This condition is also assumed to be given in our model.
@@
 a sufficiently high porosity, is required in a rock formation.
-Further significant factors are the connectivity of the porosity through pore throats, as well as pore geometry and resulting permeability \citep{dolson2016basics, sorkhabi2005place}.
+Further significant factors are the connectivity of the porosity through pore throats, as well as pore geometry and resulting permeability.
 The larger the pore throats, the easier it is for fluids to migrate in a rock.
-Typical reservoir rocks are sandstones and carbonates, but even normally tight lithologies might serve as reservoirs, if they are fractured \citep{dolson2016basics}.
+Typical reservoir rocks are sandstones and carbonates, but even normally tight lithologies might serve as reservoirs, if they are fractured \citep{dolson2016basics, sorkhabi2005place, north1985petroleum, selley2014elements}.
 In the models of this work, reservoir formations are assumed to be conventional sandstones.
         	\item \textbf{Trap:} Entrapment results from an appropriate reservoir-seal sequence and a three-dimensional geometric closure.
-Regarding a top view on a structural relief, closure is generally defined by any intersection of a structural contour and the seal, that is closed on both sides \citep{dolson2016basics, sorkhabi2005place}.
+Regarding a top view on a structural relief, closure is generally defined by any intersection of a structural contour and the seal, that is closed on both sides \citep{dolson2016basics, sorkhabi2005place, north1985petroleum}.
 A point at which a contour fails to close is referred to as "spill point".
@@
 Many different types of traps, in which hydrocarbons can accumulate, are possible.
-A common trap type is a structural four-way closure, i.e.
-as reservoir-seal system that was folded into closed anticlines, often resembling a dome and is thus predominantly defined by the structural closure relief and the cap rock sealing capacity.
+A common trap type is a structural four-way closure, i.e.\ as reservoir-seal system that was folded into closed anticlines, often resembling a dome and is thus predominantly defined by the structural closure relief and the cap rock sealing capacity.
 Such a trap type can be considered to be closed, as the reservoir surface contours are closed in a circle or oval around a relief high.
-Closure can also be given through the intersection of a fault with the seal \citep{dolson2016basics}.
+Closure can also be given through the intersection of a fault with the seal \citep{dolson2016basics, north1985petroleum}.
 Two common examples for traps are illustrated in Figure \ref{fig:trap_types}.
@@
         The presence of a fault in the reservoir-seal sequence can have different impacts on trap closure and the volume of a hydrocarbon accumulation.
-This is mainly defined by the transport properties of a fault, i.e.
-in which way it behaves as a seal or flow conduit respectively.
+This is mainly defined by the transport properties of a fault, i.e.\ in which way it behaves as a seal or flow conduit respectively.
 Cases might range from complete fault sealing to complete leakage along a fault zone.
@@
 a shale as seal) into the fault zone \citep{van2003lateral, vrolijk2016clay}.
-Clay smearing can be described using different equations, such as the Shale Smear Factor ($SSF$) and the Shale Gouge Ratio ($SGR$) \citep{lindsay1993outcrop, yielding1997quantitative, vrolijk2016clay}.
-The $SSF$ provides a simple conceptual approach to assessing the clay smearing on a fault plane, dependent on ratio of fault throw magnitude $D$ to displaced shale thickness $T$ \citep{lindsay1993outcrop, yielding1997quantitative, yielding2012using} (see Figure \ref{fig:ssf}):
+Clay smearing can be described using different equations, such as the Shale Smear Factor (SSF) and the Shale Gouge Ratio \citep{lindsay1993outcrop, yielding1997quantitative, vrolijk2016clay}.
+The SSF provides a simple conceptual approach to assessing the clay smearing on a fault plane, dependent on ratio of fault throw magnitude $D$ to displaced shale thickness $T$ \citep{lindsay1993outcrop, yielding1997quantitative, yielding2012using} (see Figure \ref{fig:ssf}):
 		\begin{equation}
-			SSF = \frac{D}{T}.
+			\text{SSF} = \frac{D}{T}.
 		\end{equation}
@@
 			\includegraphics[width=0.7\textwidth]{Figures/SSF.PNG}
-			\caption{Conceptual fault section illustrating the SSF approach to assess shale smear potential (modified from \citet{vrolijk2016clay}.)}\label{fig:ssf}
-		\end{figure}
-		It has been observed, that as the displacement and thus the ratio increases, a critical value $SSF_c$ will be reached, at which the clay smear is breached and no longer continuous.
-Initial observations by \citet{lindsay1993outcrop} indicated this threshold to be $SSF_c$ = 7.
+			\caption{Conceptual fault section illustrating the Shale Smear Factor approach to assess shale smear potential (modified from \citet{vrolijk2016clay}.)}\label{fig:ssf}
+		\end{figure}\\
+		It has been observed, that as the displacement and thus the ratio increases, a critical value SSF$_\text{c}$ will be reached, at which the clay smear is breached and no longer continuous.
+Initial observations by \citet{lindsay1993outcrop} indicated this threshold to be SSF$_\text{c}$ = 7.
 However, the critical value was later suggested to be scale-dependent and smaller for thick clays displaced by large faults \citep{yielding2012using}.
-\citet{faerseth2006shale} reported that for such large scales, including fault throws greater than 100~m, clay smear continuity can be expected for $SSF\le4$.
+\citet{faerseth2006shale} reported that for such large scales, including fault throws greater than 100~m, clay smear continuity can be expected for SSF $\le4$.
 Any larger values come with a decrease in confidence \citep{faerseth2006shale}.
@@
 before production has started), volumetric estimation is the only approach to assess the amount of hydrocarbons in place in a reservoir.
-From this value, recoverable reserves can be deduced based on an estimated recovery factor \citep{dean2007volumetric}.\\
+From this value, recoverable reserves can be deduced based on an estimated recovery factor \citep{dean2007volumetric, morton1993development}.\\
         Oil-in-place and gas-in-place volumes are calculated based on:
@@
         \item Water saturation in the reservoir rock.
-        \item Hydrocarbon fluid properties \citep{dean2007volumetric}.
+        \item Hydrocarbon fluid properties \citep{dean2007volumetric, morton1993development}.
         \end{enumerate}
-        As the hydrocarbon systems in this work are purely artificial, a pure oil accumulation is assumed for simplicity.
+        As the hydrocarbon systems in this work are purely synthetic, a pure oil accumulation is assumed for simplicity.
 Furthermore, defined formations are regarded as homogeneous in their lithological composition and petrological properties It is assumed, for example, that no shale is interbedded in the main sandstone reservoir unit.
-Using the factors listed above, the equation for original oil-in-place ($OOIP$) is formulated as follows:
+Using the factors listed above, the equation for original oil-in-place (OOIP) is formulated as follows:
         \begin{equation}\label{eq:OOIP}
-        OOIP = A * h * \phi * (1 - S_W) * 1/FvF,
+        \text{OOIP} = A * h * \phi * (1 - S_W) * 1/\text{FvF},
         \end{equation}
-        where $OOIP$ is returned in $m^3$.
-The hydrocarbon-filled rock volume is defined by the drainage area $A$ in $m^2$ and the net pay thickness $h$ in $m$.
+        where OOIP is returned in m$^3$.
+The hydrocarbon-filled rock volume is defined by the drainage area $A$ in m$^2$ and the net pay thickness $h$ in $m$.
 Porosity $\phi$ and water saturation $S_W$ (interstitial water) are given in fractions of the rock volume.
-A dimensionless factor for the change in oil volume between reservoir conditions and standard conditions at surface is represented by the formation volume factor $FvF$.
-Thus, shrinkage of the oil volume brought to the surface is determined by $1/FvF$ \citep{dean2007volumetric}.\\
-        Subsequently, the effectively recoverable oil volumes ($ROV$) can be calculated by multiplying the $OOIP$ with a recovery factor $RF$ \citep{dean2007volumetric}:
+A dimensionless factor for the change in oil volume between reservoir conditions and standard conditions at surface is represented by the formation volume factor FvF.
+Thus, shrinkage of the oil volume brought to the surface is determined by $1/FvF$ \citep{dean2007volumetric, morton1993development}.\\
+        Subsequently, the effectively recoverable oil volumes (ROV) can be calculated by multiplying the OOIP with a recovery factor RF \citep{dean2007volumetric, morton1993development}:
         \begin{equation}\label{eq:ROV}
-                ROV = OOIP * RF = A * h * \phi * (1 - S_W) * 1/FvF * RF.
+                \text{ROV}= \text{OOIP} * \text{RF} = A * h * \phi * (1 - S_W) * 1/\text{FvF} * \text{RF}.
         \end{equation}
@@
 This refers in particular to potential structural traps and ultimately the hydrocarbon-filled rock volume represented by $A * h$ in Equations \ref{eq:OOIP} and \ref{eq:ROV}.
-        Including the remaining factors as simple uncertain parameters would be easily possible and is usually done in the probabilistic estimation procedures commonly conducted in the petroleum industry \citep{wim2001guidelines}.
+        Including the remaining factors as simple uncertain parameters would be easily possible and is usually done in the probabilistic estimation procedures commonly conducted in the petroleum industry \citep{wim2001guidelines, murtha1997monte, mudford2000valuing}.
 However, the focus of this work lies on the structural aspects of a hydrocarbon system.
@@
 Effects of interest might become unrecognizable.
-Thus, fixed values are used for the non-structural factors ($\phi=30\%$, $S_W=20\%$, $FvF=1.05$, $RF = 80\%$).
-These are chosen arbitrarily but within a realistic range to considering a conventional oil reservoir.
-Consequently, calculating the $ROV$ from the maximum trap volume simply equates a rescaling of values, as proportions are conserved.
-However, translation of trap volumes to recoverable oil volumes is deemed appropriate for a formally better consideration of the results in the respective context of decision making in the hydrocarbon sector.\\
+Thus, fixed values are used for the non-structural factors ($\phi=30\%$, $S_W=20\%$, FvF $=1.05$, RF $= 80\%$).
+These are chosen arbitrarily and optimistic, but within a realistic range for a very good conventional oil reservoir (for example values see \citet{morton1993development}, \citet{ahmed2011advanced} and \citet{dandekar2013petroleum}).
+Consequently, calculating the ROV from the maximum trap volume simply equates a rescaling of values, as proportions are conserved.
+However, we deem the translation of trap volumes to recoverable oil volumes as appropriate for a formally better consideration of the results in the respective context of decision making in the hydrocarbon sector.\\
         %For most factors, it seems most suitable to define a normal distribution around a known typical value or global average, as it will return a high probability for common scenarios and occasionally produce cases of rare, more extreme values.\\
         Attaining the hydrocarbon-filled rock volume (i.e.
-the maximum trap volume) and calculating the $ROV$ is a central step of the model evaluation.
+the maximum trap volume) and calculating the ROV is a central step of the model evaluation.
 Model realizations are hereby assigned the economic value on which decisions are to be based.
-Realized volumetric values from numerous iterations and several simulations are combined to sets of probability distributions of the $ROV$.
-Considering the $ROV$ as the true value to be estimated, loss functions can then be applied as described in Section \ref{sec:loss}.
-$OOIP$ and $ROV$ calculations are applicable in a 3D setting, the 1D geological model introduced below in Section \ref{sec:1D_model}, however, requires the use of an abstract valuation system.
+Realized volumetric values from numerous iterations and several simulations are combined to sets of probability distributions of the ROV.
+Considering the ROV as the true value to be estimated, loss functions can then be applied as described in Section \ref{sec:loss}.
+OOIP and ROV calculations are applicable in a 3D setting, the 1D geological model introduced below in Section \ref{sec:1D_model}, however, requires the use of an abstract valuation system.
         %\subsubsection{Net present value}
@@
 Assuming the reservoir setting case and valuation methods introduced in Section \ref{sec:Reservoir_values}, such a customization attempt is made and explained step by step in the following.\\
-		%%%(SOURCES?)
 		For the purpose of estimation, it makes sense that one of the standard loss functions is chosen as a basis and a customized loss function is developed from there.
 The absolute-error loss seems most appropriate for this case of hydrocarbon reservoir value estimation.
-Ideally, an actor would like to know the exact true value of interest, say the $OOIP$, so that investments or resources can be allocated appropriately in order to acquire economic gains.
+Ideally, an actor would like to know the exact true value of interest, say the ROV, so that investments or resources can be allocated appropriately in order to acquire economic gains.
 This allocation is the decision to be made or action to be taken.
@@
 Here, based on the considerations named above, it is assumed that normal overestimation is 25\% (\textit{a} = 1.25), fatal overestimation 100\% (\textit{b} = 2) and fatal underestimation 50\% (\textit{c} = 1.5) worse than normal underestimation.
-\\
+For better understanding of how these determine influence the incurrence, actual losses for three fixed true values are plotted in Figure \ref{fig:LF4_det_values}.\\
 		It has to be emphasized that this is just one possible proposal for loss function customization.
-There exists not one perfect design for such a case.
+There exists not one perfect design for such a case \citep{hennig2007}.
 Slight to strong changes can already be implemented by simply varying the values of the weighting factors \textit{a, b} and \textit{c}.
@@
 This represents a risk-friendlier actor that is willing to bid on a higher estimate to attain a greater gain.
-For \textit{r} $>$ 1, the overestimation weight (\textit{a, b}) is increased in the loss function, underestimation and fatal underestimation weight (\textit{c}) are decreased and respectively more risk-averse actors are prompted to bid on lower estimates.\\
+For \textit{r} $>$ 1, the overestimation weight (\textit{a, b}) is increased in the loss function, underestimation and fatal underestimation weight (\textit{c}) are decreased and respectively more risk-averse actors are prompted to bid on lower estimates.
+Accordingly different loss function realizations are plotted in Figure \ref{fig:LFR_normal}.\\
 		The factor \textit{r} can take basically any positive values.
-However, since risk-neutrality is expressed by \textit{r} = 1, values 0 $<$ \textit{r} $<$ 2 are considered to be the most appropriate choices to represent both sides of risk-affinity equally here.
+However, since risk-neutrality is expressed by \textit{r} = 1, we consider values 0 $<$ \textit{r} $<$ 2 to be the most appropriate choices to represent both sides of risk-affinity equally here.
 
 		\section{1D geological reservoir modeling}\label{sec:1D_model}
-		For simple understanding and a preliminary assessment of the Bayesian statistical methods and the loss function described above, they are first to be applied in the context of a conceptual one-dimensional hydrocarbon system case.
+		For simple understanding and a preliminary assessment of the Bayesian statistical methods and the loss function described above, we first apply them to a conceptual one-dimensional hydrocarbon system case.
 The underlying model and basic approach are inherited from \citet{delaVarga2016}, but the parameters are adapted to more appropriately represent a reasonable geological petroleum system, consisting of a reservoir formation with overlying seal in the subsurface.
 In this 1D setting, only the interface depths and thicknesses of layers such as the reservoir or seal unit can be observed.
-Other defining aspects, such as structural entrapment, are assumed to be given.
+We assume other defining aspects, such as structural entrapment, to be given.
 Limiting the model to only one dimension and a small number of uncertain stratigraphical parameters allows for a relatively straightforward and simplified approach to assessing an abstract type of value for a reservoir, applying the custom loss function (Equation \ref{eq:LF_final}) for value estimation and observing respective effects of Bayesian inference.
-The construction of the 1D geological model is described in the following.
+The construction of this 1D geological model is described in the following.
 			\begin{figure}[h]
@@
 	        \subsection{Abstract valuation of the 1D geological model}\label{sec:1D_score_system}
-	        For valuating a simplified 1D model, $OOIP$ and recoverable volume calculations are not applicable, as these require a 3D setting.
-Given only layer boundary positions in one vertical dimension, it is resorted to an abstract way of reservoir valuation by defining a scoring system.
-A dimensionless reservoir score is made dependent on three derivatives of the uncertain 1D interface positions: (1) reservoir thickness, (2) reservoir top depth and (3) seal thickness.
+	        For valuating a simplified 1D model, OOIP and ROV calculations are not applicable, as these require a 3D setting.
+Given only layer boundary positions in one vertical dimension, we resort to an abstract way of reservoir valuation by defining a scoring system.
+A dimensionless reservoir score is made dependent on three factors derived from the uncertain 1D interface positions: (1) reservoir thickness, (2) reservoir top depth and (3) seal thickness.
 From these, (2) equals the prior parameter that indicates the reservoir-seal boundary in depth.
@@
 An exemplary comparison between the distribution of these values depending on prior and posterior parameters is depicted in Figure \ref{fig:3_parameters}.
-We hereby illustrate how model uncertainties are translated to uncertainties related to valuation measures.\\
+We hereby demonstrate how model uncertainties are translated to uncertainties related to valuation measures.
 	        \begin{figure}[h]
@@
 Here, we show a comparison between the three factors as deduced from prior parameters in Figure \ref{fig:1D_model} and as determined from exemplary posterior parameter distributions.}\label{fig:3_parameters}
-	        \end{figure}
+	        \end{figure}\\
 	        Assuming that reservoir thickness is a simplified indicator for the amount of extractable oil or gas and thus value in place, a gain in score can be correlated with increase in thickness.
-Here, two score points are assigned to one meter of thickness.
+Here, we assign two score points to one meter of thickness.
 Increasing costs of drilling are indicated by increasing depth of the reservoir top.
-Consequently, one negative score point is ascribed to every meter in depth.
-Samples from the probability distributions of these two parameters are drawn to model the true score of the reservoir (depth scores are subtracted from reservoir thickness scores).
+Consequently, we ascribe one negative score point to every meter in depth.
+Samples from the probability distributions of these two factors are drawn to model the true score of the reservoir (depth scores are subtracted from reservoir thickness scores).
 	        A third parameter is defined by the seal thickness.
 Score points are not added or subtracted by this parameter directly.
-Instead, a threshold for seal reliability is defined beforehand.
+Instead, we define a threshold for seal reliability beforehand.
 Here, it is set to 20 m thickness.
 If the seal thickness falls below this threshold, it is assumed that the seal fails completely and thus all the potential value (positive score) of the reservoir is lost, while costs of depth (negative score) remain.
-Thus, a condition to check whether the seal is reliable is included in the model.
-A respective equation for the reservoir score $S_{res}$ is defined as:
+Thus, we include a condition to check whether the seal is reliable.
+We define a respective equation for the reservoir score $S_{res}$ as:
 
@@
 			where reservoir thickness is given by $h_{res}$, seal thickness by $h_{seal}$ and reservoir depth, which is always a negative value, by $d_{res}$.\\
-			This valuation method is applied on the original prior-only Monte Carlo simulation of the 1D model as reference and then on each posterior distribution set after conducting Bayesian inference.
+			We apply this valuation method to the original prior-only Monte Carlo simulation of the 1D model as reference and then on each posterior distribution set after conducting Bayesian inference.
 Resulting score probability distributions can be compared directly and additionally under use of the case-specific custom loss function (Equation \ref{eq:LFR_final}) defined before.
@@
 For conducting simple and straightforward volumetric calculations, it is assumed that found closed traps are always filled to spill with oil, i.e.
-the complete trap volume is hydrocarbon filled and the $OOIP$ can be attained over this volume.\\
+the complete trap volume is hydrocarbon filled and the OOIP can be attained over this volume.\\
 		The maximum trap volume is assumed to equate the hydrocarbon-filled rock volume $A * h$ in Equation \ref{eq:OOIP} and is thus to be used to valuate each realization of the geological model.
@@
 		\subsection{Construction of the uncertain 3D geological reservoir model}\label{sec:3D_construction}
-		The 3D structural geological model is constructed as follows: In principle, it is defined as a cubic block with an extent of 2000~m in X, Y and Z directions.
+		We construct the 3D structural geological model as follows: In principle, it is defined as a cubic block with an extent of 2000~m in $x$-, $y$- and $z$-directions.
 The basic input data for the interpolation of the geological features is comprised of 3D point coordinates for layer interfaces and fault surfaces, as well as measurements that indicate respective dip directions and angles.
@@
 These thicknesses are to be defined as likelihoods.}\label{fig:3Dmodel_construction}
-		\end{figure}
+		\end{figure}\\
 		This model allows for the consideration of various kinds of uncertainties and their implementation in several different ways.
-Here, they are included by assigning uncertainties to the Z-positions of points which mark layer tops and interfaces in the 3D space.
-This is achieved based on respective probability distributions from which deviation values are drawn, which are then added to the original input data Z-value.
+Here, they are included by assigning uncertainties to the $z$-positions of points which mark layer interfaces in the 3D space.
+This is achieved based on respective probability distributions from which deviation values are drawn, which are then added to the original input data $z$-value.
 Uncertainties regarding layer surface positions in depth, layer thicknesses, topological shape and degree of fault offset can be incorporated.
@@
 This way, the offset nature of the normal fault is maintained and the inversion to a reverse fault is avoided.
-Specific ditributions used to represent uncertainties are listed in the Appendix.
+Specific distributions used to represent uncertainties are listed in Section \ref{sec:prior_model3D}.
 		%\begin{enumerate}
@@
 Given a juxtapostion with layers overlying the seal, due to fault displacement, the respective section is checked for fault sealing by taking into account the Shale Smear Factor value.
-In the case of a suitable $SSF$ and consequent fault sealing, leakage across this type of juxtaposition is discarded as irrelevant.
-The processes of finding the decisive spill point, as well as the determination of juxtapostions and $SSF$ is explained below in further detail.
+In the case of a suitable SSF and consequent fault sealing, leakage across this type of juxtaposition is discarded as irrelevant.
+The processes of finding the decisive spill point, as well as the determination of juxtapostions and SSF is explained below in further detail.
 			\item \textbf{Location inside of closed system:} The voxel is part of a model section inside of the main anticlinal feature.
@@
 			In GemPy, layer boundary surfaces are returned in the form of discretized arrays of simplices and vertices.
-The latter can be interpolated to a 2D rectangular grid that contains Z-positional values for the surface in X and Y directions.
+The latter can be interpolated to a 2D rectangular grid that contains $z$-positional values for the surface in $x$ and $y$ directions.
 According to \citet{verschelde2017programmingtools}, a saddle point in a matrix is maximal in its row and minimal in its column.
@@
 Higher magnitudes of fault displacement can lead to detachment of the seal in the hanging wall and consequently the juxtaposition of the trap section with seal-overlying formations which are assumed to be completely permeable.
-This case is dealt with independently, by checking this area of juxtaposition with regard to its $SSF$ and potential fault sealing, as is further elaborated in Section \ref{sec:check_juxta} below.\\
+This case is dealt with independently, by checking this area of juxtaposition with regard to its SSF and potential fault sealing, as is further elaborated in Section \ref{sec:check_juxta} below.\\
 			Following these observations, the algorithm developed in this work uses a function from the Python library Scipy to find local extrema in the 2D reservoir top surface arrays attained over GemPy.
 This function returns the position of relative extrema in an array, by comparing every grid value with a variable number of neighbors along the same row or column.
-Points where local minima and maxima of perpendicular axes X and Y coincide in an array, are then defined as saddle points.
+Points where local minima and maxima of perpendicular axes $x$ and $y$ coincide in an array, are then defined as saddle points.
 The anticlinal spill point is deduced as the maximal saddle point in the footwall.
@@
 Given only minor fault offset, no local maximum is found in the plane perpendicular to the fault.
-The Z-value of the decisive spill point is used to determine the maximum down-to-horizon to which the trap feature can be filled.
+The $z$-value of the decisive spill point is used to determine the maximum down-to-horizon to which the trap feature can be filled.\\
 			%However, if fault sealing is given, any fault leakage and respective leak points become irrelevant and only the anticlinal spill point is considered.
@@
 Failures to detect saddle points in this work's model are resolved by adding a buffer around extrema that define the anticlinal structure.
-However, it has to be pointed out that saddle points might still be missed in some rare cases, especially if they are located in an orientation which is "rotated" and not aligned with the X and Y axes.
+However, it has to be pointed out that saddle points might still be missed in some rare cases, especially if they are located in an orientation which is "rotated" and not aligned with the $x$- and $y$ axes.
 Furthermore, in this work, the geological model is designed in a way that saddle points are presumed to only occur within or near the structures of interest and are thus easier to differentiate.
@@
 			Given a juxtaposition of a trap section with layers overlying the seal across the fault, the respective area is to be examined with regard to its clay smearing potential.
-Here, this is achieved by calculating the $SSF$ as introduced in Section \ref{sec:Reservoir_values}.
+Here, this is achieved by calculating the SSF as introduced in Section \ref{sec:Reservoir_values}.
 Juxtaposition areas can be attained using a topology function that is part of GemPy.
-On the base of this, a new function is created, with which the Z-extent, i.e.
+On the base of this, a new function is created, with which the $z$-extent, i.e.
 the height of relevant juxtapositions can be deduced.
@@
 The resulting displacement value can then be related to the seal thickness to obtain the Shale Smear Factor.\\
-			As mentioned before, it has been reported that the critical value $SSF_c$ is dependent on scale and can expected to be lower with larger shale thicknesses and greater fault throws.
-So, as the scale in the present geological model is relatively large, but also for the purpose of shedding light on different conceptual scenarios by allowing the threshold value to be exceeded, the critical Shale Smear Factor is set to a respectively low value of $SSF_c = 3$.
-Consequently, in model realizations where $SSF < 3$, the fault surface on which the seal slipped, is assumed to be sealing due to clay smearing.
-Otherwise, for $SSF \le 3$, the sealing is assumed to be breached.
+			As mentioned before, it has been reported that the critical value SSF$_\text{c}$ is dependent on scale and can expected to be lower with larger shale thicknesses and greater fault throws.
+So, as the scale in the present geological model is relatively large, but also for the purpose of shedding light on different conceptual scenarios by allowing the threshold value to be exceeded, the critical Shale Smear Factor is set to a respectively low value of SSF$_\text{c} = 3$.
+Consequently, in model realizations where SSF $< 3$, the fault surface on which the seal slipped, is assumed to be sealing due to clay smearing.
+Otherwise, for SSF $\le 3$, the sealing is assumed to be breached.
 Due to the topological shape of the trap, full leakage is expected and the maximum trap volume is set to zero by not assigning any voxels to the trap structure.
@@
 			\subsection{Calculating the maximum trap volume}
-			Have all trap voxels (fulfilling the four conditions defined above), been determined, these cells are seen as part of the maximum trap volume $V_{trap}$.
-This volume can then be calculated by simply counting the number of these voxels and rescaling their cumulative volume depending on the resolution in which the model was computed:
+			Have all trap voxels (fulfilling the four conditions defined above) been determined, these cells are seen as part of the maximum trap volume $V_{trap}$.
+This volume can then be calculated by simply counting the number of trap voxels and rescaling their cumulative volume depending on the resolution in which the model was computed:
 			\begin{equation}
@@
 			Where $n_v$ is the number of trap voxels, $S_{orig}$ gives the original scale and $R_m$ the used resolution for the model.
-As mentioned before, $V_{trap}$ is assumed to equal the hydrocarbon-filled rock volume $A * h$ used in Equation \ref{eq:OOIP}.
-			For the example of a cubic geological model with an original extent of 2000~m in three directions, computed using a resolution of 50, the scale factor is 40~m.
-Every voxels thus accounts for $40^3~m=64,000~m^3$ in volume.
+As mentioned before, we assume $V_{trap}$ to equal the hydrocarbon-filled rock volume $A * h$ used in Equation \ref{eq:OOIP}.
+			For the example of a cubic geological model with an original extent of 2000~m in three directions, computed using a resolution of 100, the scale factor is 20~m.
+Every voxels thus accounts for $20~\text{m}*20~\text{m}*20~\text{m}=8,000~\text{m}^3$ in volume.
 It has to be noted, that this direct approach to rescaling and calculating the volume requires the model to be designed as a cubic block.\\
@@
 		\subsection{Implementation of reservoir-related likelihoods in the 3D model}
-		As for the 1D case before, for the 3D case it is also assumed that additional information is provided in the form of layer thickness likelihoods.
-In this geological scenario, the information might have been gained for example from additional seismic observations or from core sample data from a well nearby.
+		As for the 1D case before, for the 3D case we also assume that additional information is provided in the form of layer thickness likelihoods.
+In this geological scenario, the information might have been gained for example from additional seismic observations or from data from a drilling well nearby.
 Likelihoods could also be derived from knowledge of the geological history of the formation, such as the depositional environment of sedimentary layers.
@@
 Here, as single X- and Y-point, i.e.
-a 1D column in Z in the footwall of the model is chosen as a measure for the thickness of the layers.
+a 1D column in $z$ in the footwall of the model is chosen as a measure for the thickness of the layers.
 In this column, point coordinates for the layer interfaces had been defined in the original data frame and can be observed as they are altered in the interpolation data frame with each probabilistic modeling iteration.
 Thicknesses are simply calculated as the distance between two layer boundary points in this column.
-Variations in thicknesses can be traced and vary relative to the changes in the layer interface positions, defined by the uncertainties in their Z-values.
-At X~=~1100~m and Y~=~1000~m, the point of measurement is located centrally in the overall model block, as well as the original potential trap structure and is thus presumed to be representative and of significance for the reservoir analysis.\\
+Variations in thicknesses can be traced and vary relative to the changes in the layer interface positions, defined by the uncertainties in their $z$-values.
+At $x$~=~1100~m and $y$~=~1000~m, the point of measurement is located centrally in the overall model block, as well as the original potential trap structure, and is thus presumed to be of representative significance for the reservoir analysis.\\
 		After defining these thickness to point-positional dependencies, thickness likelihood functions are implemented based on probability distributions, as done for the 1D model.
@@
 This function is based on a normal distribution as well.
-As the $SSF$ is a more complex derivate of parameter relations, it also more difficult to argue, how respective observations were sourced.
+As the SSF is a more complex derivate of parameter relations, it also more difficult to argue, how respective observations were sourced.
 Related knowledge about fault displacement and thicknesses in an area nearby, would come with respective likelihood functions for those parent parameters as well.
-Possibly, the additional information about the $SSF$ was acquired indirectly, for example by having tested the leakage potential of other faults in the same general region.
-In any case, the inclusion of our likelihood functions based on purely hypothetical assumptions and is intended to provide the opportunity, to explore the effects different types of information might have.
-While the thickness likelihood functions are dependent on input parameters directly, the implementation of the $SSF$ likelihood function requires a full computation of the model and extended algorithms of structural analysis.\\
+Possibly, the additional information about the SSF was acquired indirectly, for example by having tested the leakage potential of other faults in the same general region.
+The inclusion of these likelihood functions is based on purely hypothetical assumptions and is intended to provide the opportunity, to explore the effects different types of information might have.
+While the thickness likelihood functions are dependent on input parameters directly, the implementation of the SSF likelihood function requires a full computation of the model and extended algorithms of structural analysis.\\
 		Some representative cases of Bayesian inference applied on the 3D model are presented in Chapter \ref{cha:results}.
@@
 		According to the methods introduced in Sections \ref{sec:bayes} and \ref{sec:bayes_in_struc_modeling}, the 3D geological model is to be incorporated in a probabilistic modeling framework and analyzed from a Bayesian perspective.
-		First, a simple Monte Carlo simulation is conducted using only priors, so that a base reference distribution for the original model is generated.
+		First, we conduct a simple Monte Carlo simulation considering only priors, so that a base reference distribution for the original model is generated.
 Subsequently, Bayesian inference is conducted based on different assumptions concerning the additional observations made and by incorporating likelihoods which respectively are designed to reflect the nature of this auxiliary data.

### 2017-11-27 Risk analysis/Text/Layout/Layout/chapter3_results.tex
Revisions

@@
 			\end{figure}
-			Results from scoring based on simple Monte Carlo error propagation using only the priors of the 1D model (as defined in Section \ref{sec:1D_construction}) are plotted in Figure \ref{fig:score_results1}.
+			Results from scoring based on simple Monte Carlo error propagation (10000 iterations) using only the priors of the 1D model (as defined in Section \ref{sec:1D_construction}) are plotted in Figure \ref{fig:score_results1}.
 A test run of scoring with only the two parameters reservoir thickness and depth, is shown in Figure \ref{fig:score_results1} (A).
@@
 
-			\subsection{Bayesian inference using thickness likelihoods}	
+			\subsection{Bayesian inference using thickness likelihood functions}	
 			\begin{figure}[p!]
@@
 Using the principle of Bayesian inference as explained in Chapter \ref{sec:bayes}, the model is updated and new posterior distributions for our true reservoir score are attained.\\
-			Three representative Bayesian updating cases based on different sets of likelihoods in the 1D model are presented in the following.
+			Three representative Bayesian updating cases (10000 iterations; burn-in phase of 1000 iterations) based on different sets of likelihoods (see Table \ref{tab:1D_likelihoods}) in the 1D model are presented in the following.
 In each case, the layer boundaries defined in Section \ref{sec:1D_construction} were adopted as prior parameters.
@@
 applying Bayesian inference, are compared to the original results and evaluations based on simple Monte Carlo error propagation using only priors.
+			\begin{table}[h]
+				\centering
+			\begin{tabular}{lrrr} 
+				\toprule
+				(a) Inference case I - Likelihoods: layer thicknesses (normal distributions)\\  
+				\midrule 
+				Formation bottom & $\mu$ [m] & $\sigma$ [m]\\ 
+				\midrule 
+				Seal & 25 & 20 \\
+				Reservoir & 180 & 40 \\
+				\bottomrule
+				\toprule
+				(b) Inference case II - Likelihoods: layer thicknesses (normal)\\
+				\midrule 
+				Formation bottom & $\mu$ [m] & $\sigma$ [m]\\ 
+				\midrule 
+				Seal & 50 & 20 \\
+				Reservoir & 180 & 40 \\
+				\bottomrule
+				\toprule
+				(c) Inference case III - Likelihoods: layer thicknesses (normal)\\
+				\midrule 
+				Formation bottom & $\mu$ [m] & $\sigma$ [m]\\ 
+				\midrule 
+				Seal & 70 & 10 \\
+				Reservoir & 100 & 30 \\
+				\bottomrule
+			\end{tabular}
+			\caption{Thickness likelihoods defined by normal distributions, as used for the three inference cases of the 1D model.}
+			\label{tab:1D_likelihoods}
+			\end{table}
 
-				\subsubsection{Inference case I: Moderately reinforcing or ambivalent information}
+				\subsubsection{1D - Inference case I: Moderately reinforcing information}
 				\begin{figure}[h]
@@
 					\end{subfigure}
-					\caption{Reservoir score distributions (A) and change in the realizations of expected loss for several risk parameters (B1, B2) before and after Bayesian updating based on likelihoods defined as follows: Seal thickness: $\mu = 25~m; \sigma = 20~m$; reservoir thickness: $\mu = 180~m; \sigma = 40~m$.}
+					\caption{1D - Inference case I.
+Reservoir score distributions (A) and change in the realizations of expected loss for several risk parameters (B1, B2) before and after Bayesian updating based on likelihoods defined as follows: Seal thickness: $\mu = 25~m; \sigma = 20~m$; reservoir thickness: $\mu = 180~m; \sigma = 40~m$.}
 					\label{fig:update_moderate2_3}
 				\end{figure}				
-				In this first case, a normal distribution with a mean ($\mu$) of 25~m and a standard deviation ($\sigma$) of 20~m was chosen to reflect the likelihood of the seal thickness.
-For the reservoir, a normal distribution with a mean of 180~m and a standard deviation of 40~m was used respectively.
-As can be observed in Figure \ref{fig:update_moderate1}), the uncertainty in the probability distributions for the positions of layer boundaries in depth was reduced moderately, mostly for the reservoir bottom, by conducting Bayesian inference with these values.\\
+				As can be observed in Figure \ref{fig:update_moderate1}, the uncertainty in the probability distributions for the positions of layer boundaries in depth was reduced moderately by implementing the likelihoods defined in Table \ref{tab:1D_likelihoods} (a).\\
 				Scoring was applied based on these new distributions.
@@
 So while uncertainties were reduced in areas of opposite sign, the divide between these was significantly increased.
-The overall uncertainty seems thus to have been widely maintained and barely transformed into a problem of duality.
-The change in uncertainty might thus me described as ambivalent.\\
-				Application of the custom loss function (Equation \ref{eq:LFR_final}) is visualized in Figure \ref{fig:update_moderate2_3}, in which the expected losses are compared before (B1) and after (B2) Bayesian inference.
+The overall uncertainty seems thus to have been widely maintained and barely transformed into a problem of duality.\\
+				Application of the custom loss function (Equation \ref{eq:LFR_final}) is visualized in Figure \ref{fig:update_moderate2_3}, in which the expected losses are compared before (B1) and after (B2) inference.
 It is observable, that by adding information about layer thickness likelihoods, Bayes actions were shifted relative to the nature of the information.
-In this case, the added data generally reinforces the probability of the reservoir to be sufficiently thick.
-Information on the seal, however, based on a normal distribution around 25~m thickness, leaves high uncertainty about the reliability of the seal, as the safety threshold is defined as 20~m.
-Consequently, the risk of complete seal failure remains a major concern.\\
+In this case, the added data generally reinforced the probability of the reservoir to be sufficiently thick.
+Information on the seal, however, based on a normal distribution around 25~m thickness, left high uncertainty about the reliability of the seal, as the safety threshold was defined to be 20~m.
+Consequently, the risk of complete seal failure remained a major concern.\\
 				Increased certainty about the reservoir thickness was sufficient to shift Bayes actions to higher estimates for all actors.
-Estimates of the more risk neutral actors ($r = 0.5$ to $r=1.25$) were increased the most.
+Estimators of the more risk-neutral actors ($r = 0.75$ to $r=1.25$) were increased the most.
 This has led to a slight convergence of the Bayes actions.
-Expected losses were altered only slighty.
-For the risk neutral actor, this change is negligible.
+For the risk-neutral actor, change in expected loss was negligible.
 For risk-friendlier actors, the expected loss was decreased, but for risk-averse actors, as they shifted from zero ("take no action") to positive estimates, increased.\\
@@
 
-				\subsubsection{Inference case II: Likely reliable seal}
+				\subsubsection{1D - Inference case II: Likely reliable seal}
 				\begin{figure}[h]
@@
 					\end{subfigure}
-					\caption{Reservoir score distributions (A) and change in the realizations of expected loss for several risk parameters (B1, B2) before and after Bayesian updating based on likelihoods defined as follows: Seal thickness: $\mu = 50~m; \sigma = 20~m$; reservoir thickness: $\mu = 180~m; \sigma = 40~m$.}
+					\caption{1D - Inference case II.
+Reservoir score distributions (A) and change in the realizations of expected loss for several risk parameters (B1, B2) before and after Bayesian updating based on likelihoods defined as follows: Seal thickness: $\mu = 50~m; \sigma = 20~m$; reservoir thickness: $\mu = 180~m; \sigma = 40~m$.}
 					\label{fig:update_goodseal2_3}
 				\end{figure}
-				In this second case, the reservoir thickness likelihood was defined in the same way as in case I ($\mu = 180~m; \sigma = 40~m$).
-For the seal, a higher mean of 50~m with a standard deviation of 20~m was chosen, favoring the likelihood of a reliable seal relative to the threshold of 20~m thickness.
+				In this second case, the reservoir thickness likelihood was defined in the same way as in case I (see Table \ref{tab:1D_likelihoods}).
+For the seal, a higher mean of 50~m chosen, favoring the likelihood of a safe thickness relative to the threshold of 20~m.
 Respective score results are depicted in Figure \ref{fig:update_goodseal2_3} (A).
@@
 Due to this, in combination with the overall distribution narrowing, mean and median were clearly shifted to higher values and are now much closer together.\\
-				Applying the custom loss function on this new score distribution resulted in the realizations of expected loss illustrated in Figure \ref{fig:update_goodseal2_3}.
+				Applying the custom loss function to this new score distribution resulted in the realizations of expected loss illustrated in Figure \ref{fig:update_goodseal2_3}.
 Bayes actions were shifted clearly to higher estimates and expected losses of these minima were significantly reduced for all actors.
-The risk-neutral to risk-averse individuals seem to have profited the most, due to a large change in estimate value and decreased expected loss.
+The risk-neutral to risk-averse individuals seem to have profited the most, due to a large change in estimator value and decreased expected loss (Bayes risk).
 Compared to the foregone case, a higher degree of overall uncertainty reduction was achieved.
@@
 
-				\subsubsection{Inference case III: Safe seal but likely subpar reservoir thickness}
+				\subsubsection{1D - Inference case III: Safe seal but likely subpar reservoir thickness}
 				\begin{figure}[h]
@@
 					\end{subfigure}
-					\caption{Reservoir score distributions (A) and change in the realizations of expected loss for several risk parameters (B1, B2) before and after Bayesian updating based on likelihoods defined as follows: Seal thickness: $\mu = 70~m; \sigma = 10~m$; reservoir thickness: $\mu = 100~m; \sigma = 30~m$.}
+					\caption{1D - Inference case III.
+Reservoir score distributions (A) and change in the realizations of expected loss for several risk parameters (B1, B2) before and after Bayesian updating based on likelihoods defined as follows: Seal thickness: $\mu = 70~m; \sigma = 10~m$; reservoir thickness: $\mu = 100~m; \sigma = 30~m$.}
 					\label{fig:update_smallres2_3}
@@
 				In this third case, seal safety was ensured by using a mean of 70~m with a standard deviation of 10~m for the seal thickness likelihood.
-However, the new observations were assumed to provide information about the reservoir unit that makes it likely to be thinner than expected.
-Reservoir thickness likelihood was assigned a mean of 100~m and a standard deviation of 30~m.\\
+However, the new observations were assumed to provide information about the reservoir unit that makes it likely to be thinner than priorly expected (see Table \ref{tab:1D_likelihoods} (c))\\
 				The subsequent score distribution is depicted in Figure \ref{fig:update_smallres2_3} (A).
-It can be seen that while the whole distribution was strongly narrowed, it also was shifted to the left, to lower and negative estimates.
+It can be seen that while the whole distribution was strongly narrowed, it also was shifted to lower and negative estimates.
 Mean and median are almost equal, indicating narrowing towards symmetry in the shape of a normal distribution.
-As seal reliability is practically guaranteed, the peak at -2000 has vanished.\\
-				Based on this posterior score probability distribution, Bayes actions were shifted to lower estimates for all actors using the custom loss function (see Figure \ref{fig:update_smallres2_3} (B2)).
-The risk-neutral and both risk-averse individuals find their Bayesian estimators to be zero after updating, i.e.
-bidding on a positive score is deemed to be too risky to them.
+As seal reliability was practically guaranteed, the peak at -2000 vanished.\\
+				Based on this posterior score probability distribution, Bayes actions were shifted to lower estimates for all actors (see Figure \ref{fig:update_smallres2_3} (B2)).
+The risk-neutral and both risk-averse individuals find their Bayesian estimators to be zero after updating, i.e.\ bidding on a positive score is deemed to be too risky to them.
 There is no shift in estimate for risk-averse actors, as they already found their best estimates to be zero before updating.
-Notable is the large reduction of expected loss in general.
+Notable is the large reduction of expected loss (Bayes risk) in general.
 The range of Bayes actions was significantly narrowed.
-Overall, it is apparent that uncertainty was significantly reduced (see the narrowed distribution).
-However, although the seal can expected to be safe, the reservoir thickness is likely subpar.
+Overall, it is apparent that uncertainty was greatly reduced (see the narrowed distribution).
+However, although the seal can now expected to be safe, the reservoir thickness is likely subpar.
 Due to this, the distribution was shifted to lower and negative values, strongly increasing the probability of an unfavorable true score.
@@
 		\section{3D geological model results}
-		In the following, results from applying the methods on the artificial 3D structural model are presented.
+		In the following, we present the results from applying the methods to the synthetic 3D structural geological model.
 Several representative examples of posterior models based on the use of different likelihoods are referenced to the original prior model.
-For model simulations, 1000 sampling steps were conducted, with a burn-in phase of 50 iterations for MCMC sampling.
+For each simulation, 1,000 sampling steps were conducted, with a burn-in phase of 50 iterations for MCMC sampling.
 However, for computational reasons, Shannon entropy visualization was based on 200 sampling steps and a burn-in phase of 50 iterations.
-The following measures are central to the comparison of results: (1) information entropy, (2) occurrence of trap control mechanisms, (3) resulting recoverable oil volumes and (3) consequent realization of expected losses and related Bayes actions (i.e.
+We consider in particular the following measures for the comparison of results: (1) Shannon entropy, (2) occurrence of trap control mechanisms, (3) resulting recoverable oil volumes and (4) consequent realization of expected losses and related Bayes actions (i.e.
 decisions).
 
-		\subsection{Prior model}\label{sec:prior_model3D}
+		\subsection{3D - Prior model}\label{sec:prior_model3D}
 		To generate a set of reference results, a prior model was first generated by running a simple simulation of Monte Carlo sampling considering only the uncertainty of prior parameters.
@@
 			\includegraphics[width=1\textwidth]{Figures/M_prior}
-			\caption{Recoverable oil volumes probability distribution for the prior model (a) and respective realization of the custom loss function (b).}\label{fig:M_prior}
+			\caption{Recoverable oil volumes probability distribution for the prior 3D model (a) and respective realization of the custom loss function (b).}\label{fig:M_prior}
 		\end{figure}\\
-		The nature of the prior uncertainties was visualized using information entropy (see Figure \ref{fig:ML1} (a)).
-High entropies are found at and near the possible locations of layer boundaries.
-This effect was amplified by the offset uncertainty in the footwall, where the highest information entropies are found.\\
+		The nature of the prior uncertainties was visualized using Shannon entropy (see Figure \ref{fig:ML1} (a)).
+High entropies can be seen at and near the possible locations of layer boundaries.
+This effect was strongly amplified by the offset uncertainty in the footwall, where the highest entropies are found.\\
 		The occurrence of the different mechanisms of maximum trap control were traced for all simulations and labeled as follows:
 		\begin{enumerate}
-		\item \textbf{SPILL POINT}: Migration related to the anticlinal spill point of the trap.
-		\item \textbf{LEAK UNDER}: Leakage due to juxtaposition of the reservoir with formations underlying the seal in the hanging wall.
-		\item \textbf{LEAK OVER}: Leakage due to juxtaposition of the reservoir with formations overlying the seal in the hanging wall.
+		\item \textbf{SPILL POINT:} Migration related to the anticlinal spill point of the trap.
+		\item \textbf{LEAK UNDER:} Leakage due to juxtaposition of the reservoir with formations underlying the seal in the hanging wall.
+		\item \textbf{LEAK OVER:} Leakage due to juxtaposition of the reservoir with formations overlying the seal in the hanging wall.
 This was defined to lead to full leakage and complete failure of the trap.
-		\item \textbf{STRAT}: Leakage due to stratigraphical connections of the trap to seal-overlying units, i.e.
+		\item \textbf{STRAT:} Leakage due to stratigraphical connections of the trap to seal-overlying units, i.e.
 stratigraphical breaches of the seal.
 Full leakage and trap failure were assumed for this mechanism as well.
-		\item \textbf{UNCLEAR}: Label defined to track cases in which a mechanism was not recognized.
+		\item \textbf{UNCLEAR:} Label defined to track cases in which a mechanism was not recognized.
 This is deemed an indicator for failures of the algorithms or the computation, i.e.
@@
 It can clearly be seen that while a large portion of the distribution is found in a range of highly positive volumes, forming a normal distribution shape around 3.8 million m$^3$, there is also a high probability of complete failure, i.e.
-the trap volume to be zero.
-This results in a bimodal characterization, with low probability of the $ROV$ to be found between zero and 3 million m$^3$.\\
-		Consequently, applying the custom loss function on this distribution resulted in divergent Bayes actions for the different risk-affine actors (see Figure \ref{fig:M_prior} (b)).
-Although all actors found their estimates to be positive, only the risk-neutral to risk-friendliest actors came close to the described positive mode of the distribution ($ROV > 3*10^6$ m$^3$).
+for the trap volume to be zero.
+This results in a bimodal characterization with low probability of the ROV to be found between zero and 3 million m$^3$.\\
+		Consequently, applying the custom loss function to this distribution resulted in divergent Bayes actions for the different risk-affine actors (see Figure \ref{fig:M_prior} (b)).
+Although all actors find their estimators to be positive, only the risk-neutral to risk-friendliest actors come close to the described positive mode of the distribution (ROV $> 3*10^6$ m$^3$).
 Risk-averse individuals bid on significantly lower estimates.
 It can also be observed that expected losses resulted to be generally lower for risk-affine actors.
-It can be presumed that the divergence of different Bayes actions was caused the bimodality of the $ROV$ probability distribution.
+It can be presumed that the divergence of different Bayes actions was caused by the bimodality of the ROV probability distribution.
 
-		\subsection{Posterior model I: Seal and reservoir likely thick}
+		\subsection{3D - Posterior model I: Seal and reservoir likely thick}
 		\begin{table}[h]
@@
 			\includegraphics[width=1\textwidth]{Figures/ML1}
-			\caption{Summary of results from evaluating posterior model I.
+			\caption{Summary of results from evaluating posterior model I (3D).
 Shannon entropies in prior (a) and posterior (b) are compared.
@@
 Respective results are summarized in Figure \ref{fig:ML1}.\\
-		Considering the changes in information entropy, it can be seen that the uncertainty about layer interface positions in $z$ was clearly reduced in the footwall.
+		Considering the changes in Shannon entropy, it can be seen that the uncertainty about layer interface positions in $z$ was clearly reduced in the footwall.
 Information entropy was also decreased in the middle section of the hanging wall, but remained high in upper and lower parts, presumably due to the remaining influence of the offset uncertainty.\\
 		Regarding trap volume control (Figure \ref{fig:ML1} (c)), only spill point-related migration ("SPILL POINT") and cross-fault leakage due to juxtaposition with permeable layers below the seal in the hanging wall ("LEAK UNDER") remained relevant in the posterior model.
-As no occurrences of leakage through overlying layers were observed, the two other mechanisms increased in frequency, while staying proportionally the same to one another.
+As no occurrences of leakage through overlying layers were observed, these two main mechanisms increased in frequency, while staying proportionally the same to one another.
 It can be presumed, that by introducing the respective likelihoods, a high thickness of the seal was ensured, as can also be recognized in Figure \ref{fig:ML1} (b).
-This not only made stratigraphic leakage in this model impossible, but also enforced a favorable $SSF$.
+This not only made stratigraphic leakage in this model impossible, but also enforced a favorable SSF.
 A minor emergence of unrecognized control factors ("UNCLEAR") suggests that, for a few iterations, the model was possibly not correctly computable or the structural analysis failed for a different reason.\\
-		The bulk that accounted for the positive mode in the prior $ROV$ distribution was raised and shifted laterally to higher values, while the probability of failure ($ROV = 0$) was diminished significantly (see Figure Figure \ref{fig:ML1} (d)).
-This can presumably be explained by the disappearing of leakage to overlying formations as a trap control mechanism, which was programmed to always lead to complete failure of the trap volume.\\
-		As the probability of failure and thus the bimodality of the probability distribution disappeared, all Bayes actions were shifted to higher estimates, most significantly those of the two risk-averse individuals (see Figure \ref{fig:ML1} (e).
+		The bulk that accounted for the positive mode in the prior ROV distribution was raised and shifted laterally to higher values, while the probability of failure (ROV = 0) was diminished significantly (see Figure Figure \ref{fig:ML1} (d)).
+This can presumably be explained by the disappearance of leakage to overlying formations as a trap control mechanism, which was programmed to always lead to complete failure of the trap volume.\\
+		As the probability of failure and thus the bimodality of the probability distribution were diminished, all Bayes actions were shifted to higher estimates, most significantly those of the two risk-averse individuals (see Figure \ref{fig:ML1} (e)).
 These have thus converged into an area of similar expectations.
-Additionally, expected losses were lowered overall.
+Additionally, expected losses (Bayes risks) were lowered overall.
 The posterior state of knowledge, provided by the updated model, encouraged even the most risk-averse decision-maker to expect recoverable oil volumes from 3.5 million m$^3$ upwards.
 
-		\subsection{Posterior model II: Likely thin reservoir formation}%ML3
+		\subsection{3D - Posterior model II: Likely thin reservoir formation}%ML3
 		\begin{table}[h]
@@
 		\end{table}
-		For the second model, thickness likelihoods were defined in a way that reinforced the probability of a very thick seal but at the same time restricted the reservoir formation to a likely significantly thinner thickness of around 50 m (see Table \ref{tab:ML3_likelihoods}).
+		For the second model, thickness likelihoods were defined in a way that reinforced the probability of a very thick seal, but at the same time restricted the reservoir formation to a likely significantly thinner thickness of around 50 m (see Table \ref{tab:ML3_likelihoods}).
 The results (see Figure \ref{fig:ML3}) are very similar, in fact almost identical, to those of model I above.
@@
 Ultimately, the spill point appears to be the primariy limiting factor to maximum trap fill.
-It can furthermore be derived from this that the maximum trap volume potential, as defined by the given structural features, is already represented in the $ROV$ probability distribution of the prior model.
-After examining Figure \ref{fig:ML1} (d), it seems that this overall maximum volume is found at approximately 5.7 millon m$3$.
+It can furthermore be derived from this that the maximum trap volume potential, as defined by the given structural features, is already represented in the ROV probability distribution of the prior model.
+After examining Figure \ref{fig:ML1} (d), it seems that this overall maximum volume is found at approximately 5.7 millon m$^3$.
 Higher values were not observed in any other model realization and are presumed to be unattainable due to the restriction given by the spill point.
 
-		\subsection{Posterior model III: Likely very thin seal}%ML2
+		\subsection{3D - Posterior model III: Likely thin seal}%ML2
 		\begin{figure}[p!]
@@
 			\includegraphics[width=1\textwidth]{Figures/ML2}
-			\caption{Summary of results from evaluating posterior model III.
+			\caption{Summary of results from evaluating posterior model III (3D).
 Shannon entropies in prior (a) and posterior (b) are compared.
@@
 		Using likelihoods which favor a thick reservoir, but only a thin seal of about 50 m thickness (see Table \ref{tab:ML2_likelihoods}), leakage due to juxtaposition with seal-overlying formations across the fault ("LEAK OVER") became the primary mechanism for trap volume control (see Figure \ref{fig:ML2} (c).
-It can be recognized in Figure \ref{fig:ML2} (b) that, while uncertainty was reduced clearly in the footwall, it remains high in the hanging wall.
-The state of information entropy furthermore suggests that the seal layer frequently appeared significantly deeper in the hanging wall than in the footwall.
+It can be recognized in Figure \ref{fig:ML2} (b) that, while uncertainty was reduced clearly in the footwall, it remained high in the hanging wall.
+The Shannon entropy visualization furthermore suggests that the seal layer frequently appeared significantly deeper in the hanging wall than in the footwall.
 Combined with a generally very thin seal, this allowed for the occurrence of high displacement to seal thickness ratios, i.e.
-a high $SSF$.
-Thus, frequent surpassing of the critical $SSF$ and subsequent fault seal breaching were enabled.
+a high SSF.
+Thus, frequent surpassing of the critical SSF and subsequent fault seal breaching were enabled.
 It is additionally apparent in Figure \ref{fig:ML2} (b), that while fault-related leakage became predominant, control by the spill point was suppressed to be widely irrelevant in this posterior model.\\	%It is suggested by the footwall information entropy in Figure \ref{label} (b), that the general shape of the structure has been flattened.
 		A dramatically higher probability of complete trap failure was induced by the high frequency of cross-fault leakage to overlying formations.
-Consequently, the bulk of the posterior $ROV$ probability distribution was shifted to a volume of zero, as depicted in Figure \ref{fig:ML2} (d).
+Consequently, the bulk of the posterior ROV probability distribution was shifted to a volume of zero, as depicted in Figure \ref{fig:ML2} (d).
 Any positive trap volumes have become highly improbable.\\
 		Bayes actions changed accordingly.
-Given this posterior $ROV$ distribution, all actors found their minimum of expected loss at $ROV = 0$, which equals the decision to "take no action" (see Figure \ref{fig:ML2} (e)).
-It has to be pointed out that all Bayes actions were found at the same point due to the lack of negative values into which more risk-averse actors would presumably have moved, given the possibility.
+Given this posterior ROV distribution, all actors find their minimum of expected loss at ROV = 0, which equals the decision to "take no action" (see Figure \ref{fig:ML2} (e)).
+It has to be pointed out that all Bayes actions are found at the same point due to the lack of negative values into which more risk-averse actors would presumably have moved, given the possibility.
 This poses a difference to the score system implemented for the 1D geological model evaluation.
-As volumes cannot be negative, Bayes actions cannot be shifted past the limit of $ROV = 0$.
-Compared to the prior model, expected losses were decreased overall, but are higher for risk-friendlier individuals at the minimum.
+As volumes cannot be negative, Bayes actions cannot be shifted past the limit of ROV = 0.
+Compared to the prior model, expected losses decreased overall, but are higher for risk-friendlier individuals at the minimum.
 Results from this third model show that in contrast to the reservoir thickness, seal thickness is a primary and decisive factor for the manifestation of a reliable trap structure.
 
-		\subsection{Posterior model IV: SSF likelihood reinforcing trap reliability}\label{sec:model4}%ML4
+		\subsection{3D - Posterior model IV: SSF likelihood reinforcing trap reliability}\label{sec:model4}%ML4
 		\begin{figure}[p!]
@@
 			\end{tabular}
-			\caption{Layer thickness likelihoods and SSF-likelihood determined by normal distributions as used for posterior model IV.}
+			\caption{Layer thickness likelihoods and SSF likelihood determined by normal distributions as used for posterior model IV (3D).}
 			\label{tab:ML4_likelihoods}
@@
 supporting the probability of thick layers in general.
-Additionally, a $SSF$-related likelihood function was implemented.
-With a mean of 1.5 and a standard deviation of 0.5, it was chosen to mostly reinforce the likelihood of a $SSF$ below the critical value of $SSF_c = 3$ (see Table \ref{tab:ML4_likelihoods}).\\
+Additionally, a SSF-related likelihood function was implemented.
+With a mean of 1.5 and a standard deviation of 0.5, it was chosen to mostly reinforce the likelihood of a SSF below the critical value of SSF$_\text{c} = 3$ (see Table \ref{tab:ML4_likelihoods}).\\
 		In Figure \ref{fig:ML4} (b), it can be seen that the uncertainty about the position of interfaces in the footwall was greatly reduced for all layers.
-In the hanging wall, information entropy remained high but was shifted upwards and mainly concentrated in an area that is juxtaposed to the footwall seal formation.\\
-		In the resulting $ROV$ distribution, the possibility of trap failure was completely extinguished, while the probability of high volumes in the range between 3.5  and 5 million m$^3$ was significantly increased.
-Mean and median are found at almost the same value, but the bulk of the distribution found in this range appears to be slightly bimodal in itself, resembling approximately the shape of two normal distributions overlapping around the value of 4 million m$^3$.\\
+In the hanging wall, Sannon entropy remained higher but was shifted upwards and mainly concentrated in an area that is juxtaposed to the footwall seal formation.\\
+		In the resulting ROV distribution, the possibility of trap failure was completely extinguished, while the probability of high volumes in the range between 3.5  and 5 million m$^3$ was significantly increased.
+Mean and median are found at almost the same value, but the bulk of the distribution found in this range appears to be slightly bimodal in itself, resembling approximately the shape of two normal distributions overlapping around the value of 4.3 million m$^3$.\\
 		Accordingly, Bayes actions converged within this range of positive values.
-Considering the two modes in this range, risk-averse to risk-neutral bids remained in the middle of the mode of lower volumes, while risk-friendly actors were encouraged to estimate higher volumes as found in the second mode.
+Considering the two modes in this range, risk-averse to risk-neutral bids remained in the middle of the mode of lower volumes, while the risk-friendliest actor was encouraged to bid on a higher volume found in the second mode.
 Overall, significant convergence was achieved, also in respect to expected losses.\\
-		The importance of a reliable seal is once again emphasized by these results.
-Reliable fault sealing was favored by the chosen $SSF$ likelihood.
-It should be pointed out that computation of the complete model is required to attain the $SSF$.
-This value is furthermore characterized by a complex interdependence with other parameters and is most directly related to layer thicknesses and fault offset.
+		The importance of a safe seal was once again emphasized by these results.
+Reliable fault sealing was favored by the chosen SSF likelihood.
+It should be pointed out that computation of the complete model is required to attain the SSF.
+This value is furthermore characterized by a complex interdependence with other parameters and is directly related to layer thicknesses and fault offset.
 For this reason, information entropy and offset uncertainty in the hanging wall were not greatly reduced.
-However, the certainty about the ratio of displacement to layer thicknesses was presumably improved by using the $SSF$ in a likelihood function.
+However, the certainty about the ratio of displacement to layer thicknesses was presumably improved by using the SSF in a likelihood function.
 
-		\subsection{Posterior model V: Controversial information about the Shale Smear Factor}\label{sec:model5}%ML5
+		\subsection{3D - Posterior model V: Controversial information about the Shale Smear Factor}\label{sec:model5}%ML5
 		\begin{figure}[p!]
@@
 			\end{tabular}
-			\caption{SSF-likelihood determined by a normal distributions as used for posterior model V.}
+			\caption{SSF-likelihood determined by a normal distributions as used for posterior model V (3D).}
 			\label{tab:ML5_likelihoods}
 		\end{table}
-		In this last case example, no thickness likelihoods, only a $SSF$-related likelihood was used.
-In contrast to model IV, this was chosen to narrow the probability of the $SSF$ to be around the threshold of its critical value of $SSF_c = 3$ (see Table \ref{tab:ML5_likelihoods}).\\
+		In this last case example, no thickness likelihoods, only a SSF-related likelihood was used.
+In contrast to model IV, this was chosen to narrow the probability of the SSF to be around the threshold of its critical value of SSF$_\text{c} = 3$ (see Table \ref{tab:ML5_likelihoods}).\\
 		Results are relatively similar to those of the prior model.
-It is notable that the information entropy regarding the position of layer boundaries in the footwall was decreased, while the entropy seems to have been raised significantly in the hanging wall (Figure \ref{fig:ML5} (b)).
-Furthermore, an increase in the relevance of cross-fault leakage to overlying formations ("LEAK OVER") as a trap volume mechanism (increase from $\sim13\%$ to $\sim18\%$) was observed (Figure \ref{fig:ML5} (c)).
-Consequently, the bimodal character between the bulk of positive values and the probability of complete failure in the $ROV$ distribution was reinforced.
+Shannon entropies were only slightly decreased overall (Figure \ref{fig:ML5} (b)).
+An increase in the relevance of cross-fault leakage to overlying formations ("LEAK OVER") as a trap volume mechanism (increase from $\sim13\%$ to $\sim18\%$) was observed (Figure \ref{fig:ML5} (c)).
+Consequently, the bimodal character between the bulk of positive values and the probability of complete failure in the ROV distribution was reinforced.
 This is also indicated by a shift of mean and median to lower volumes (Figure \ref{fig:ML5} (d)).
@@
 Expected losses of these minima were raised as well.\\
-		It is notable that the implementation of a likelihood function, which can be expected to significantly decrease the $SSF$ related uncertainty, overall leads to an amplification of the duality in the $ROV$ probability distribution and to dispersion of differently risk-affine decisions.
-This effect is presumably caused by conservation of decisive uncertainty between "success" and complete failure of the trap, due to narrowing of the $SSF$ probability around its critical value.
+		It is notable that the implementation of a likelihood function, which can be expected to significantly decrease the SSF related uncertainty, overall leads to an amplification of the duality in the ROV probability distribution and to dispersion of differently risk-affine decisions.
+This effect is presumably caused by conservation of decisive uncertainty between "success" and complete failure of the trap, due to narrowing of the SSF probability around its critical value.

### 2017-11-27 Risk analysis/Text/Layout/Layout/chapter4_discussion.tex
Revisions

@@
 To observe this, a case-specific loss function was developed in this work.
-In doing so, we intend to furthermore show, that the loss function estimation approach is suitable and useful to describe varying degrees of complexities behind decision-making environments and risk-related behavior of the actors in it.
-	% It was initially hypothesized that Bayesian updating via likelihoods and the resulting change of uncertainty in a model should have significant effects on subsequent decision-making.\\
+In doing so, we intended to furthermore show, that the loss function estimation approach is suitable and useful to describe varying degrees of complexities behind decision-making environments and the risk-related behavior of the actors in it.
 	To analyze this, geological models were interpreted as potential hydrocarbon systems.
@@
 Different prior and posterior probability distributions for such economic values were generated to form a basis for application of the custom loss function for value estimation.
-These methods were applied for a conceptual 1D geological model and subsequently for a 3D structural model.
+These methods were first applied to a conceptual 1D geological model and subsequently to a full 3D structural model.
 	Despite a significant leap in complexity, the results from both models proved to be widely similar.
 	\subsubsection{State of knowledge, decision uncertainty and consequent decision making via loss function optimization}
-	The economic value realized in each model construction (score and $ROV$, respectively), had been defined to depend on model parameters, either directly or indirectly by taking into account their relations to other parameters or to specific thresholds and conditions.
+	The economic value realized in each model construction (score and ROV, respectively), had been defined to depend on model parameters, either directly or indirectly by taking into account their relations to other parameters or to specific thresholds and conditions.
 This meant in general, that in this step of valuation, model uncertainties were in a way translated and reduced to the probability distribution of a single decisive value.
 As this was set to be the foundation for loss function estimation, we consider this value distribution to be an expression of the state of knowledge on which the decision-making is to be based.
-Thereby, we furthermore propose, that the overall uncertainty inherent in this probability distribution can be referred to as "decision uncertainty" and that this entity is to be viewed separately from the model uncertainty.\\
-	By viewing decision-making as a problem of optimizing a case-specific customized loss function based on such a state of knowledge and decision uncertainty, we were able to observe clear differences in the respective behavior of distinctly risk-affine actors.
-	The position and spread of their Bayes actions, i.e.
-their decision, manifested according to the properties of value probability distributions and their inherent decision uncertainty.
+Thereby, we furthermore propose that the overall uncertainty inherent in this probability distribution can be referred to as "decision uncertainty" and that this entity is to be viewed separately from the model uncertainty.\\
+	By viewing decision-making as a problem of optimizing a case-specific loss function based on such a state of knowledge and decision uncertainty, we were able to observe clear differences in the respective behavior of distinctly risk-affine actors.
+	The position and separation of their Bayes actions, i.e.\ their decisions, manifested according to the properties of the value distributions.
 These were in turn determined by the proportionality regarding distribution modes, their relative probabilities and distances to each other.
@@
 Pronounced bimodality between two extremes, i.e.
-high decision uncertainty, resulted in a wieder separation of Bayes estimators.
+high uncertainty, resulted in a wider separation of Bayes estimators.
 Reduction of the distribution to one mode conversely led to the convergence of different Bayes actions.
 A decrease in decision uncertainty furthermore was accompanied by a reduction in expected loss for each Bayes estimator.
-	Considering these observations, we derive that the degree of Bayes action convergence and respective expected losses can be considered measures for the state of knowledge and decision uncertainty at the moment of making a decision.
-The better this is, the more similar the decisions of differently risk-affine actors and the lower their loss expectations are.
+	Considering these observations, we derive that the degree of Bayes action convergence and respective expected losses, i.e.
+Bayes risks, can be considered measures for the state of knowledge and decision uncertainty at the moment of making a decision.
+The better these are, the more similar the decisions of differently risk-affine actors and the lower their loss expectations are.
 It can be assumed, that given perfect information, all actors would bid on the same estimate (the true value) and expect no loss, since no risk is present.
@@
 different realizations of the value probability distribution, were attained by the implementation of Bayesian inference.
-While in most settings, inference led to a reduction of decision uncertainty, this was not always the case.
-We observed that the impact on decision uncertainty, induced by Bayesian inference, is not necessarily strictly aligned with the change in uncertainty regarding model parameters and their derivatives.
+While in most settings, adding information via inference led to a reduction of decision uncertainty, this was not always the case.
+We observed that the impact on decision uncertainty, induced by Bayesian inference, is not necessarily strictly aligned with the change in uncertainty regarding model parameters and their combinations.
 In respect to this, two aspects seems to be of central importance: (1) "where" in the model uncertainty is reduced, i.e.
 in which spatial area or regarding which model parameters and (2) which possible outcome is enhanced in terms of probability.
-An increased probability of a thick or thin seal in our 3D model, for example, equally reduced decision uncertainty significantly, by raising the probability of a positive outcome or negative outcome, respectively.
+An increased probability of a thick or thin seal in our 3D model, for example, equally reduced decision uncertainty significantly, by raising the probability of a positive or negative outcome, respectively.
 Improved certainty about our reservoir thickness, however, had little to no impact on decision-making.
@@
 In both types of model construction, 1D and 3D, this was directly related to sealing reliability.
-Thresholds regarding seal thickness (1D model) and Shale Smear Factor ($SSF_c$ in 3D model) were defined in a way that introduced a significant possibility of complete trap failure.
+Thresholds regarding seal thickness (1D model) and Shale Smear Factor (SSF$_\text{c}$ in 3D model) were defined in a way that introduced a significant possibility of complete trap failure.
 Consequently, it was observed that reducing the uncertainty in a way that narrowed the probability of a threshold-related parameter around its cut-off value, led to an amplification of the respective mode and thus emphasized the risk of complete failure.
-This can be seen in the posterior $ROV$ distribution of 3D model V (see Figure \ref{fig:ML5}).
+This can be seen in the posterior ROV distribution of 3D model V (see Figure \ref{fig:ML5}).
 Resulting Bayes actions were characterized by an increase in divergence and expected losses, indicating a deterioration of the state of knowledge and an according increase in decision uncertainty.
@@
 Eliminating this risk proved to be far less significant to the most risk-friendly, than for risk-averse actors.
-Accordingly, it should be of foremost importance, for risk-averse actors, to reduce the uncertainty regarding threshold-related factors which might decide between the success and complete failure of a project.
+Accordingly, it should be of foremost importance for risk-averse actors, to reduce the uncertainty regarding threshold-related factors which might decide between the success and complete failure of a project.
 This is less relevant for risk-friendly decisions makers, who respectively might acquire a comparable benefit from knowing more about the probability of positive outcomes, as they are less discouraged from overestimating and less afraid of failure.
@@
 In other cases, they are derived from more complex parameter inter-relations, as is the case for the Shale Smear Factor, which is the ratio of displacement to seal thickness.
-3D models IV and V (see Sections \ref{sec:model4} and \ref{sec:model5}) showed that reduced uncertainty about the $SSF$ is not necessarily directly recognizable via Shannon entropy visualization, especially if uncertainty regarding its parent parameters remains high.
+Posterior models IV and V (see Sections \ref{sec:model4} and \ref{sec:model5}) showed that reduced uncertainty about the SSF is not necessarily directly recognizable via Shannon entropy visualization, especially if uncertainty regarding its parent parameters remains high.
 It follows that, to approach an effective mitigation of high risks, the complexities behind decisive factors need to be assessed thoroughly and respective parent parameters, as well as their interdependencies, need to be identified.
@@
 	A central conclusion from these observations seems to be, that more of simply any type of information does not necessarily lead to better decisions.
-Instead, improvement of decision-making is achieved by attaining the right kind of information that is able to shed light on uncertainties which are relevant to an individuals own goals and preferences, as well as the general problem at hand.
+Instead, improved of decision-making is achieved by attaining the right kind of information that is able to shed light on uncertainties which are relevant to an individuals own goals and preferences, as well as the general problem at hand.
 These findings pair well with statements made by \citet{bratvold2010making}, who emphasized that the purpose of models in the hydrocarbon sector is to attain insights to improve decisions.
 In their book, they point out that value is not generated by uncertainty quantification or reduction in itself, but is created to the extent that these processes have potential to change a decision.
-This is reflected by our results which showed that one set of information might have a vastly different impact on the Bayes estimators of different actors.
-New observations might lead to a great change in decision for one actor, but be negligible the other.
+This is reflected by our results which showed that one set of information might have a vastly different effects on a variety of Bayes estimators.
+New observations might lead to a great change in decision for one actor, but be negligible for the other.
 		%This state of knowledge and the decision uncertainty accordingly experienced alterations related to the different Bayesian inference cases.
@@
 	\subsubsection{Value distribution modes as possible model scenarios and potential decision options}	
+	Effects of adding information via Bayesian inference were reflected in changing appearances of modes in the reservoir value distributions.
+Considering that these changes correlated with increased or decreased relevance of specific trap control mechanisms in the 3D model, we derive that each mode represents the occurrence of a specific mechanism-related scenario.
+These were mostly distinctly separated with little probability in between.
+This presumably resulted from the assumption that traps are always "filled-to-spill" in respect to spill and leak point, and that full leakage is assumed for any type of seal failure, resulting in an approximate discretization into respective scenarios.
 	\begin{figure}[h]
@@
 For both cases, zero equals the option to "take no action", i.e.
-allocate no resources.}\label{fig:scenarios_options}
-	\end{figure}
-	Effects of adding information via Bayesian inference were reflected in changing appearances of modes in the reservoir value distributions.
-Considering that these changes correlated with increased or decreased relevance of specific trap control mechanisms in the 3D model, we derive that each mode represents the occurrence of a specific mechanism-related scenario.
-These were mostly distinctly separated with little probability in between.
-This presumably resulted from the assumption that traps are always "filled-to-spill" in respect to spill and leak point, and that full leakage is assumed for any type of seal failure, resulting in an approximate discretization into respective scenarios.\\
+to allocate no resources.}\label{fig:scenarios_options}
+	\end{figure}\\
 	From a decision-making perspective, these scenarios might be interpreted as different decision options.
-In a case of distinct bimodality as seen in the prior of the 3D model, two main options would be given: (1) to not take any action, or (2) to allocate resources according to an estimated value within the positive mode.
+In a case of distinct bimodality as seen in the prior of the 3D model, two main options would be given: (1) to not take any action or (2) to allocate resources according to an estimated value within the positive mode.
 However, most often, low probabilities for the values between these two opposite options remain.
@@
 A discretization of decision subregions in a probability distribution could thus also be based on previously defined actual options.
-A conceptual illustration of this is shown in Figure \ref{fig:scenarios_options}.
-	\subsubsection{Comparison to common practices and other research in the hydrocarbon sector}		
-	In remains to be argued, to what extent our methodology and findings contribute regarding practical applications in hydrocarbon exploration and production.
-Monte Carlo simulation for reservoir estimation and risk assessment have become common in this sector and are often used in combination with decision trees to make decisions (see \citet{murtha1997monte}, \citet{mudford2000valuing} and \citet{bratvold2010making}).
+A schematic illustration of these concepts is shown in Figure \ref{fig:scenarios_options}, where three predefined decision areas are labeled according to hypothetical project sizes options.
+	\subsubsection{Comparison to common practices and research in the hydrocarbon sector}		
+	In remains to be argued, to what extent our methodology and findings contribute to practical applications in hydrocarbon exploration and production.
+Monte Carlo simulation for reservoir estimation and risk assessment has become common in this sector and is often used in combination with decision trees to make decisions (see \citet{murtha1997monte}, \citet{mudford2000valuing}, \citet{wim2001guidelines} and \citet{bratvold2010making}).
 However, it seems to us that distributions resulting from probabilistic modeling are mostly only considered to attain best estimates in the form of means.
@@
 We believe that this practice does not harness the full potential of such a probabilistic distribution and that much of the inherent information potential is discarded.
-Contrary to that, customized loss functions as a Bayesian method, take into account the full probability distribution and enable the inclusion of various conditions in the process of finding an optimal estimate.
+Contrary to that, customized loss functions, as a Bayesian method, take into account the full probability distribution and enable the inclusion of various conditions in the process of finding an optimal estimate.
 While used in statistical decision theory and other scientific fields, loss functions have, to the best of our knowledge, found no significant application in the field of petroleum exploration and production.
@@
 It might furthermore be able to weight risks that are not immediately apparent to an individual.
-Furthermore, the influence of human biases and psychological behavioral challenges, as described by \citet{bratvold2010making}, could be avoided.\\
+Furthermore, the influence of human biases and psychological behavioral challenges, as described by \citet{bratvold2010making}, could be mitigated.\\
 	More recently, Bayesian inference and MCMC methods were applied for OOIP estimation and forecasting of reservoir productivity by \citet{wadsley2005markov}, \citet{ma2006multistage} and \citet{liu2010continuous}.
-However, their research focused on simulations based on history matching for already producing fields.
+However, their research focused on history-matching simulations for already producing fields.
 Our approach of applying Bayesian inference for structural geological modeling and volumetric reservoir calculations is intended to support decision-making in the earliest stages of a reservoir, when it has to be decided whether a project should be developed or not.
-Nevertheless, in was shown in the work by \citet{wadsley2005markov} that early volumetric OOIP estimates can be combined with later calculations from production data via MCMC methods.
+Nevertheless, it was shown in the research conducted by \citet{wadsley2005markov} that early volumetric OOIP estimates can be combined with later calculations from production data via MCMC methods.
 This is an area where our methods could possibly be integrated.\\
 	\citet{bratvold2010making} highlighted the potential to use Monte Carlo simulation to identify, via sensitivity analysis, which uncertainties have the greatest impact on decisions.
-This was virtually achieved by performing several cases of Bayesian inference in our work, as we recognized primarily the importance of seal reliability.
+This was virtually achieved by performing several cases of Bayesian inference in our work, as we recognized primarily the importance of seal safety.
 	\subsubsection{Representativity and limitations of this methodology}
@@
 The fact that similar observations were made for the 1D, as well as the 3D model, indicates a certain degree of continuity with respect to representativity.
-However, we also need to address that the uncertainties employed in the 3D model related to $z$-axis positional values only and were thus of mostly  one-dimensional nature.
-This may in part explain the parallelism to the 1D model results.
+However, we also need to address that the uncertainties employed in the 3D model related to $z$-axis positional values only and were thus of primarily one-dimensional nature.
+This may in part account for the parallelism to the 1D model results.
 Furthermore, it follows that no effective uncertainty concerning the overall structural shape was implemented, particularly regarding anticlinal features and the position of the spill point in relation to the trap top.
 Due to this, trap volumes tended to occur primarily in distinct "filled-to-spill" scenarios.
-While being unrealistic, this contributed to pronounced influences related to the various control mechanisms.\\
+While being unrealistic, this enabled the recognition of pronounced influences related to the various control mechanisms.\\
 	For the purpose of a conceptual application, the customization of the loss functions was kept relatively simple.
@@
 It would be possible to extend this design to attain very complex loss functions.
-As the loss function in this work was designed based on only a few basic assumptions, it might be more generally representative than a complex function which would takes into account numerous very specific aspects.
+As the loss function in this work was designed based on only a few basic assumptions, it might be more generally representative than a complex function that takes into account numerous very specific aspects.
 The consideration of more details, without a basis on real data, would have furthermore required extensive speculation, which would have presumably impaired the generalization potential of respective observations.
@@
 Recoverable reserves could be replaced by the NPV to evaluate modeling results and serve as a base for decision-making.
-A more elaborate respective loss function could be customized on the base of surveys, acquiring the specific preferences of one or more companies and thus attaining a better profile of the economic environment, as well as the individuals acting in it.
-Value of information could be considered as a parameter to better quantify the impact of Bayesian inference as observed in this work.
-	Although the algorithms for automatic 3D hydrocarbon trap recognition were developed to fit the primarily model in this work, we presume that they could easily be adapted to other structural cases or even engineered further to be universally applicable.\\
+A more elaborate respective loss function could be customized on the base of surveys, acquiring the specific preferences of one or several companies and thus attaining a better profile of the economic environment, as well as the individuals acting in it.
+Value of information could be considered as a parameter to better quantify the impact of Bayesian inference as observed in this work.\\
+	Although the algorithms for automatic 3D hydrocarbon trap recognition were developed to fit primarily the model in this work, we presume that they could easily be adapted to other structural cases or even engineered further to be universally applicable.\\
 	Furthermore, additional and different uncertainty parameters should be considered in the future.
@@
 Otherwise, non-structural reservoir parameters could be included as uncertain values, such as porosities and permeabilities of different layers.
-This might be of particular interest considering factors which are part of the $ROV$ equation or other aspects that related to high risks in decision-making.\\
-	Hydrocarbon systems and petroleum exploration as a sector were chosen for an exemplary application in this work, but other settings can be found, in which geological modeling is of central significance for decision-making.
+This might be of particular interest considering factors which are part of the ROV equation or other aspects that are related to high risks in decision-making.\\
+	We chose hydrocarbon systems and petroleum exploration as a sector for an exemplary application in this work, but other settings can be found, in which geological modeling is of central significance for decision-making.
 One example would be subsurface storage of fluids in a reservoir, such as carbon capture and storage (CCS).
-Questions regaring storage capacity and safety deal with similar conditions and geological problems, as the ones presented this work, most importantly seal reliability and the risk of leakage.
+Questions regaring storage capacity and safety deal with similar conditions and geological problems as the ones presented this work, most importantly seal reliability and the risk of leakage.
 In respect to this, the models and approaches in this work might provide a basis to raise new and advanced explorable cases.

### 2017-11-27 Risk analysis/Text/Layout/Layout/MSc_layout_Fabian.tex
Cleaning Notebooks2
Formatting

@@
 %
+Statutory Declaration in Lieu of an Oath
+
+\vspace*{0.1\textheight}
+
+\noindent\begin{minipage}[t]{0.5\textwidth}
+\begin{flushleft} \large
+
+%\authorname \\
+Stamm, Fabian \\
+\emph{\footnotesize Last Name, First Name}
+
+\end{flushleft}
+\end{minipage}%
+\begin{minipage}[t]{0.5\textwidth}
+\begin{flushright} \large
+
+%123456 \\
+%\emph{\footnotesize Matriculation No.}
+
+\end{flushright}
+\end{minipage}\\[1cm]
+
+\noindent I hereby declare in lieu of an oath that I have completed the present Master thesis entitled "\msctitle\ - \mscsubtitle" independently and without illegitimate assistance from third parties.
+I have used no other than the specified sources and aids.
+In case that the thesis is additionally submitted in an electronic format, I declare that the written and electronic versions are fully identical.
+The thesis has not been submitted to any examination body in this, or similar, form.\\[2cm]
+
+\noindent\begin{minipage}[t]{0.5\textwidth}
+\begin{flushleft} \large
+
+Aachen, \today \\
+\emph{\footnotesize City, Date}
+
+\end{flushleft}
+\end{minipage}%
+\begin{minipage}[t]{0.5\textwidth}
+\begin{flushright} \large
+
+\break
+\rule{10em}{0.5pt}\\ % This prints a line for the signature
+\emph{\footnotesize Signature}
+
+\end{flushright}
+\end{minipage}\\[3cm]
+
+\cleardoublepage
+
 % Abstract
@@
 Results from conducting this for several inference cases show that the various Bayes estimators shift according to the characteristics of the underlying value distribution.
-While bimodality and overall uncertainty leads to separation, risk-averse and risk-friendly decisions converge and decrease in expected loss.
-These degree of these processes is considered a measure for the state of knowledge and decision uncertainty at the decision-making moment.
-This decision uncertainty does not change aligned with model uncertainty, but depends on alterations of critical parameter and interdependencies, relating to seal reliability in particular.
-Additionally, actors affected differently by one set of information, as defined by their risk-opportunity weighting.
-For good decision making, it appears to be most important to recognize in for which parameters reduction of uncertainty leads to changes in decisions.
+While bimodality and overall uncertainty leads to separation, risk-averse and risk-friendly decisions converge and decrease in expected loss given more narrow unimodal distributions.
+The degree of decision convergence is considered a measure for the state of knowledge and decision uncertainty at the moment of decision making.
+This decision uncertainty does not change aligned with model uncertainty but depends on alterations of critical parameters and respective interdependencies, relating to seal reliability in particular.
+Additionally, actors are affected differently by one set of information, as defined by their risk weighting.
+For good decision making, it appears to be most important to recognize for which parameters reduction of uncertainty leads to changes in decisions.
 
@@
 % Nomenclature
-    \printnomencl %
+    %\printnomencl %
 %
@@
     \begin{acronym}%
-        \acro{RWTH}{Aachen University}%
+    	\acro{CSV}{Comma-separated values}%
+    	\acro{MAP}{Maximum a posteriori}%
+        \acro{MCMC}{Markov chain Monte Carlo}
+        \acro{NPV}{Net present value}%
+        \acro{OOIP}{Original oil-in-place}
+        \acro{ROV}{Recoverable oil volumes}%
+        \acro{SSF}{Shale Smear Factor}
     \end{acronym}%
@@
 
-    \chapter{The back of the thesis}
+    \chapter{Appendix}
+    \section{Code availability}
+    	GemPy and PyMC are both fully open source and to be used in a Python environment.
+At the time of writing this, GemPy is still under development (Version 0.997).
+It can be obtained from \url{https://github.com/cgre-aachen/gempy}.
+PyMC can be downloaded from \url{https://github.com/pymc-devs/pymc} (PyMC2) and \url{https://github.com/pymc-devs/pymc3} (PyMC3).\\
+    	The models, algorithms and functions developed and presented in this work can be found on \url{https://github.com/nre-aachen/Theses/tree/master/Risk analysis}.
 
-    \section{An appendix section}
+    \section{Additional tables and figures}
+
+    \begin{figure}[h]
+    	\centering
+    	\includegraphics[width=0.62\textwidth]{Figures/Appendix/update_goodseal1}
+    	\caption{Prior (A1) and posterior distributions (B1) of the layer boundary positions in depth and respective representations (A2, B2) for 1D inference case II.}\label{fig:update_goodseal1}
+    \end{figure}
+
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=0.62\textwidth]{Figures/Appendix/update_smallres1}
+		\caption{Prior (A1) and posterior distributions (B1) of the layer boundary positions in depth and respective representations (A2, B2) for 1D inference case III.}\label{fig:update_smallres1}
+	\end{figure}
 
@@
     	\includegraphics[width=0.62\textwidth]{Figures/Appendix/CSV_input1}
-    	\caption{....}\label{fig:CSV_input1}
+    	\caption{3D model construction input data 1: point-positional values for layer interfaces and the fault surface.}\label{fig:CSV_input1}
     \end{figure}
@@
 		\includegraphics[width=0.62\textwidth]{Figures/Appendix/CSV_input2}
-		\caption{....}\label{fig:CSV_input2}
+		\caption{3D model construction input data 2: point-positional values for layer interfaces and the fault surface.}\label{fig:CSV_input2}
 	\end{figure}
@@
 		\includegraphics[width=1\textwidth]{Figures/Appendix/CSV_input3}
-		\caption{....}\label{fig:CSV_input3}
+		\caption{3D model construction input data 3: dip -related values for layer interfaces and the fault surface.}\label{fig:CSV_input3}
 	\end{figure}
@@
     	\includegraphics[width=1\textwidth]{Figures/Appendix/ML3}
-    	\caption{....}\label{fig:ML3}
+    	\caption{Summary of results from evaluating posterior model II (3D).
+Shannon entropies in prior (a) and posterior (b) are compared.
+So are the mechanisms of maximum trap volume control (c).
+Prior and posterior ROV probability distributions are shown in (d).
+Respective loss function and Bayes actions are plotted in (e).}\label{fig:ML3}
     \end{figure}
 
-    \subsection{An appendix subsection with C++ Lisitng}
+    %\section{Another test section}
 
-    \lstset{language=C++}
-    \lstinputlisting{test.c}
-
-    \subsection{A \matlab $ $ Listing}
-
-    \lstset{language=matlab}
-    \lstinputlisting{test.m}
-
-    \chapter{Yet another appendix}
-
-    \section{Another test section}
-
-    Ok, all is well.
+    %Ok, all is well.
 

### 2017-11-27 Risk analysis/Text/Layout/Layout/MSc_layout_Fabian.tex
Some corrections

@@
 \usepackage{booktabs}
+\usepackage{pdfpages}
 \graphicspath{Figures/}
@@
 %
-Statutory Declaration in Lieu of an Oath
-
-\vspace*{0.1\textheight}
-
-\noindent\begin{minipage}[t]{0.5\textwidth}
-\begin{flushleft} \large
-
-%\authorname \\
-Stamm, Fabian \\
-\emph{\footnotesize Last Name, First Name}
-
-\end{flushleft}
-\end{minipage}%
-\begin{minipage}[t]{0.5\textwidth}
-\begin{flushright} \large
-
-%123456 \\
-%\emph{\footnotesize Matriculation No.}
-
-\end{flushright}
-\end{minipage}\\[1cm]
-
-\noindent I hereby declare in lieu of an oath that I have completed the present Master thesis entitled "\msctitle\ - \mscsubtitle" independently and without illegitimate assistance from third parties.
-I have used no other than the specified sources and aids.
-In case that the thesis is additionally submitted in an electronic format, I declare that the written and electronic versions are fully identical.
-The thesis has not been submitted to any examination body in this, or similar, form.\\[2cm]
-
-\noindent\begin{minipage}[t]{0.5\textwidth}
-\begin{flushleft} \large
-
-Aachen, \today \\
-\emph{\footnotesize City, Date}
-
-\end{flushleft}
-\end{minipage}%
-\begin{minipage}[t]{0.5\textwidth}
-\begin{flushright} \large
-
-\break
-\rule{10em}{0.5pt}\\ % This prints a line for the signature
-\emph{\footnotesize Signature}
-
-\end{flushright}
-\end{minipage}\\[3cm]
+\includepdf[pages={1}]{eidesstattliche_erklaerung.pdf}
 
@@
 While bimodality and overall uncertainty leads to separation, risk-averse and risk-friendly decisions converge and decrease in expected loss given more narrow unimodal distributions.
-The degree of decision convergence is considered a measure for the state of knowledge and decision uncertainty at the moment of decision making.
-This decision uncertainty does not change aligned with model uncertainty but depends on alterations of critical parameters and respective interdependencies, relating to seal reliability in particular.
+The degree of decision convergence is considered a measure for the state of knowledge and decisive uncertainty at the moment of decision making.
+This uncertainty does not change aligned with model uncertainty but depends on alterations of critical parameters and respective interdependencies, relating to seal reliability in particular.
 Additionally, actors are affected differently by one set of information, as defined by their risk weighting.
@@
     \nonumchap{Acknowledgements}%
-    First of all I want to thank all the people who have participated in this project ..
-    Remember, often more people have (in some way) contributed to your final thesis than you would initially think of....
+    First and foremost, I wish to express my greatest thanks to my supervisor Miguel de la Varga, for continually supporting and motivating me to pursue a better understanding of computational geosciences and probabilistic methods.
+
+    I also want to thank all of my friends and fellow researchers
     \vspace*{15mm}
@@
 PyMC can be downloaded from \url{https://github.com/pymc-devs/pymc} (PyMC2) and \url{https://github.com/pymc-devs/pymc3} (PyMC3).\\
-    	The models, algorithms and functions developed and presented in this work can be found on \url{https://github.com/nre-aachen/Theses/tree/master/Risk analysis}.
-
+    	The models, algorithms and functions developed and presented in this work are accessible as IPython Notebooks on github: \url{https://github.com/nre-aachen/Theses/tree/master/Risk analysis}.
+\newpage
     \section{Additional tables and figures}
@@
     	\centering
-    	\includegraphics[width=0.62\textwidth]{Figures/Appendix/update_goodseal1}
+    	\includegraphics[width=0.50\textwidth]{Figures/Appendix/update_goodseal1}
     	\caption{Prior (A1) and posterior distributions (B1) of the layer boundary positions in depth and respective representations (A2, B2) for 1D inference case II.}\label{fig:update_goodseal1}
@@
 		\centering
-		\includegraphics[width=0.62\textwidth]{Figures/Appendix/update_smallres1}
+		\includegraphics[width=0.50\textwidth]{Figures/Appendix/update_smallres1}
 		\caption{Prior (A1) and posterior distributions (B1) of the layer boundary positions in depth and respective representations (A2, B2) for 1D inference case III.}\label{fig:update_smallres1}
@@
     	\includegraphics[width=0.62\textwidth]{Figures/Appendix/CSV_input1}
-    	\caption{3D model construction input data 1: point-positional values for layer interfaces and the fault surface.}\label{fig:CSV_input1}
+    	\caption{3D model construction input data (1): point-positional values for layer interfaces and the fault surface.}\label{fig:CSV_input1}
     \end{figure}
@@
 		\includegraphics[width=0.62\textwidth]{Figures/Appendix/CSV_input2}
-		\caption{3D model construction input data 2: point-positional values for layer interfaces and the fault surface.}\label{fig:CSV_input2}
+		\caption{3D model construction input data (2): point-positional values for layer interfaces and the fault surface.}\label{fig:CSV_input2}
 	\end{figure}
@@
 		\includegraphics[width=1\textwidth]{Figures/Appendix/CSV_input3}
-		\caption{3D model construction input data 3: dip -related values for layer interfaces and the fault surface.}\label{fig:CSV_input3}
+		\caption{3D model construction input data (3): dip-related values for layer interfaces and the fault surface.}\label{fig:CSV_input3}
 	\end{figure}

### 2017-11-27 Risk analysis/Text/Layout/Layout/chapter1.tex
Some corrections

@@
 Recent developments have transferred this probabilistic approach to address uncertainties the field of structural geological modeling.
-\citet{Caumon2010}, \citet{jessell2010towards} and \citet{Wellmann2010b} developed methods of geologically based inversion by considering secondary geological observations to constrain uncertain models.
+\citet{Caumon2010}, \citet{jessell2010towards} and \citet{wellmann2010towards} developed methods of geologically based inversion by considering secondary geological observations to constrain uncertain models.
 Advancements in this field have been enabled by progressing developments regarding implicit modeling functions based on interpolation and the possibility of fully automated model reconstruction in particular \citep{hillier2014three, mallet1992discrete, lajaunie1997foliation}.

### 2017-11-27 Risk analysis/Text/Layout/Layout/chapter3_results.tex
Some corrections

@@
 
-			\subsection{Scoring}
+c			\subsection{Scoring}
 			\begin{figure}[h]
@@
 				In this second case, the reservoir thickness likelihood was defined in the same way as in case I (see Table \ref{tab:1D_likelihoods}).
-For the seal, a higher mean of 50~m chosen, favoring the likelihood of a safe thickness relative to the threshold of 20~m.
+For the seal, a higher mean of 50~m chosen, favoring the likelihood of a safe thickness relative to the threshold of 20~m (see Figure \ref{fig:update_goodseal1}).
 Respective score results are depicted in Figure \ref{fig:update_goodseal2_3} (A).
@@
 				In this third case, seal safety was ensured by using a mean of 70~m with a standard deviation of 10~m for the seal thickness likelihood.
-However, the new observations were assumed to provide information about the reservoir unit that makes it likely to be thinner than priorly expected (see Table \ref{tab:1D_likelihoods} (c))\\
+However, the new observations were assumed to provide information about the reservoir unit that makes it likely to be thinner than priorly expected (see Table \ref{tab:1D_likelihoods} (c) and Figure \ref{fig:update_smallres1})\\
 				The subsequent score distribution is depicted in Figure \ref{fig:update_smallres2_3} (A).

### 2017-11-27 Risk analysis/Text/Layout/Layout/MSc_layout_Fabian.tex
Acknowledgements

@@
     \nonumchap{Acknowledgements}%
-    First and foremost, I wish to express my greatest thanks to my supervisor Miguel de la Varga, for continually supporting and motivating me to pursue a better understanding of computational geosciences and probabilistic methods.
-
-    I also want to thank all of my friends and fellow researchers
+    First and foremost, I wish to express my greatest thanks to Miguel de la Varga, whose outstanding support I have received since he was my tutor in Numerical Reservoir Engineering.
+He encouraged me to challenge myself and to pursue a path into a research field that, not long ago, was very much unfamiliar to me.
+I am grateful for the continuous motivation, constant positive feedback and especially the immense patience he showed as my supervisor and mentor.\\
+    Furthermore, I want to voice my deep gratitude to Professor Florian Wellmann, for welcoming me as a part of his research group and thus showing me, how exciting and inspiring it can be, to participate in a free and innovative academic environment.
+With his interest, openness and positive attitude, he encouraged me to contribute to his group's work and to develop my own scientific ideas.\\
+    I also wish to thank Professor Janos Urai, for years of fascinating lectures on structural geology, hydrocarbon exploration and production, and related risk analysis.
+His courses provided me with many insights and served as a sources of inspiration and information that I was able to include in this project.\\
+    In general, I am grateful to my friends and fellow researchers from the CGRE group, especially for creating such a pleasant working atmosphere.
+Special thanks go to the modeling team.
+To Alexander Schaaf, for helping me when Miguel could not be found and reminding me to take a break now and then.
+To Marco van Veen, for endless inciting and exhilarating discussions, but also for the many shared adventures during our geoscientific studies.\\
+    And of course, I owe my deepest gratitude to my family, for a lifetime of unconditional support.
+To my parents, for raising me to be curious.
+For enabling me to attain such excellent education and for providing me with the freedom to follow my individual interests and aspirations.
     \vspace*{15mm}

### 2017-11-27 Risk analysis/Text/Layout/Layout/chapter1.tex
minor corrections

@@
 In general geosciences, Bayesian inference has prominently found use in the context of geophysical inversion problems (see \citet{tarantola1982inverse, mosegaard2002probabilistic} and \citet{sambridge2002monte}).
-Recent developments have transferred this probabilistic approach to address uncertainties the field of structural geological modeling.
+Recent developments have transferred this probabilistic approach to address uncertainties in the field of structural geological modeling.
 \citet{Caumon2010}, \citet{jessell2010towards} and \citet{wellmann2010towards} developed methods of geologically based inversion by considering secondary geological observations to constrain uncertain models.
@@
 \citet{delaVarga2016} regarded geological modeling as a Bayesian inference problem by relating additional geological information to prior model parameters in the form of likelihood functions, linking them in a non-parametric Bayesian network.
-Using Markov chain Monte Carlo sampling to explore resulting probability spaces, they attained posterior model suites with reduced uncertainties \citep{delaVarga2016}.\\
+Using Markov chain Monte Carlo sampling to explore resulting probability spaces, they attained posterior model suites with reduced uncertainties.\\
 This work builds upon their concept, exploring the potential significance their findings might have in the context of decision making.
@@
 More specifically, the models are designed to represent potential hydrocarbon reservoir systems.
-Consequently, the development of algorithms for automatic structural trap recognition and volumetric calculation represents a central part of this work.\\
+The development of algorithms for automatic structural trap recognition and volumetric calculation represents a central part of this work.\\
 The main hypothesis of this work is that Bayesian inference and resulting changes in uncertainties in a geological setting have a significant effect on reservoir value estimation and decision making.

### 2017-11-27 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Some corrections
minor corrections

@@
         \end{equation}        
-        So, how "bad" a current estimate is, depends on the way a loss function weighs accuracy errors and returns respective losses.
+        So, how "bad" a current estimate is, depends on the way a loss function weights accuracy errors and returns respective losses.
 Two standard loss functions are the absolute-error and the squared-error loss function.
@@
         Accordingly, losses increasing linearly with the distance to the true value are returned for respective estimates.
-This means that all differences between relative errors are weighed equally, no matter whether they are found in the realm of relatively small or relatively large errors \citep{hennig2007}.\\
+This means that all differences between relative errors are weighted equally, no matter whether they are found in the realm of relatively small or relatively large errors \citep{hennig2007}.\\
         Using the squared-error loss function returns losses that increase quadratically with distance of the estimator to the true parameter value \citep{davidson2015, moye2006statistical}:        
@@
         \end{equation}         
-        This exponential growth of loss also means that large errors are weighed much stronger than small errors.
+        This exponential growth of loss also means that large errors are weighted much stronger than small errors.
 This might come with over-valuation of distant outliers and misrepresentation of magnitudes in distance.
@@
 According to \citet{hennig2007}, this is not necessarily unfavorable or "less objective", as it may better reflect an expert's perspective on the situation and contribute to a productive scientific discussion.\\
-        The standard loss functions defined above are symmetric, but can easily be adapted to be asymmetric, for example by weighing errors on the negative side stronger than those on the positive side.
+        The standard loss functions defined above are symmetric, but can easily be adapted to be asymmetric, for example by weighting errors on the negative side stronger than those on the positive side.
 Preference over estimates larger than the true value (i.e.
@@
         To ensure a thorough exploration of the probability space, the transition to higher probability densities should not be enforced in every case but selectively.
-This is assured by relating the acceptance ratio from Equation \ref{eq:acceptance_ratio} to a random value $u$ from a Uniform distribution $U(0,1)$ as follows \citep{delaVarga2016}:
+This is assured by relating the acceptance ratio from Equation \ref{eq:acceptance_ratio} to a random value $u$ from a Uniform distribution $U(0,1)$ as follows:
 
@@
         \end{equation}        
-        where $\vec{x_0}$ is an arbitrary origin, $M$ and $N$ are the total number of data points and partial derivatives respectively, and their relative contributions are weighed by the factors $\mu_\alpha$ and $\nu_\beta$.
+        where $\vec{x_0}$ is an arbitrary origin, $M$ and $N$ are the total number of data points and partial derivatives respectively, and their relative contributions are weighted by the factors $\mu_\alpha$ and $\nu_\beta$.
 Furthermore, $T$ is assumed to be a random function defined by polynomial drift and a stationary covariance $K(h)$ \citep{calcagno2008geological}.
@@
 Its design allows for its application in a probabilistic setting as elaborated below.
-At the time of writing this, GemPy is still under development (version 0.995), but is already functioning for the purpose of this work.\\
+At the time of writing this, GemPy is still under development (version 0.997), but is already functioning for the purpose of this work.\\
 		For conducting the Bayesian analysis and embedding the geological modeling step in a probabilistic context, we integrate GemPy into the probabilistic framework of PyMC.
 This Python library was developed for conducting Bayesian inference and prediction problems in an open-source probabilistic programming environment \citep{davidson2015, salvatier2016pymc3}.
-Different model fitting techniques are provided in PyMC, such as the \textit{maximum a posteriori} (MAP) method and several (MCMC) sampling methods, including the Adaptive Metropolis introduced in Section \ref{sec:mcmc}.
+Different model fitting techniques are provided in PyMC, such as the \textit{maximum a posteriori} (MAP) method and several MCMC sampling methods, including the Adaptive Metropolis introduced in Section \ref{sec:mcmc}.
 The components which are used to construct a statistical model, are represented by \textit{Deterministic} or \textit{Stochastic} functions or variables in PyMC \citep{salvatier2016pymc3}.
@@
 		The values of \textit{Deterministic} variables are completely dependent on its parents' values, as defined by a respective mathematical function \citep{salvatier2016pymc3}.
-\textit{Stochastic} variables are used to represent uncertain parameters $\theta$ or observed stochastic variables as likelihood functions $p(y|\theta)$ \citep{salvatier2016pymc3, delaVarga2016}.
-Complex mathematical relations between \textit{Stochastic} variables can be described through \textit{Deterministic} functions \citep{delaVarga2016}.
-Futhermore, PyMC allows for the creation of own object definitions inheriting from the class descriptions of these two function and variable types.\\
-		Bayesian networks as a concept have been presented in more detail by \citet{koller2009probabilistic}.
+\textit{Stochastic} variables are used to represent uncertain parameters $\theta$ or observed stochastic variables as likelihood functions $p(y|\theta)$.
+Complex mathematical relations between \textit{Stochastic} variables can be described through \textit{Deterministic} functions.
+Futhermore, PyMC allows for the creation of own object definitions inheriting from the class descriptions of these two function and variable types \citep{salvatier2016pymc3, delaVarga2016}.\\
+		Bayesian networks as a concept have been presented in further detail by \citet{koller2009probabilistic}.
 \citet{salvatier2016pymc3} have pointed out that the development of PyMC is continuing, as the inclusion of further tools is planned for future updates.
@@
         After conducting Bayesian inference, posterior distributions and resulting model realizations are to be evaluated.
-For this, we can examine sampling traces and posterior distributions of single parameters can be observed.
-The entirety of accepted 3D geological models (sampled after burn-in phase) is to be assessed in terms of uncertainty quantification and visualization.
+For this, we can examine sampling traces and posterior distributions of single parameters or their combinations.
+The entirety of accepted 3D geological models (sampled after a burn-in phase) is to be assessed in terms of uncertainty quantification and visualization.
 Furthermore, a central part of this work is the assignment of an economic significance to the modeling results.
 To achieve this, we valuate accepted models by regarding them as potential hydrocarbon systems.
-We explain this process in details in Section \ref{sec:Reservoir_values}.
+We explain this process in detail in Section \ref{sec:Reservoir_values}.
 Shannon entropy is introduced as a measure for uncertainty in the following.
@@
         \end{equation}
-        This total entropy $H_T$ is 0 when no uncertainty is found in any cell, oppositely it is maximal in the case of equal probabilities of outcomes in every cell \citep{wellmann2012uncertainties, schaaf2017}.
+        This total entropy $H_T$ is 0 when no uncertainty is found in any cell, oppositely it is maximal in the case of equal probabilities of outcomes in every cell.
 The application of these entropy measures allows for the calculation and visualization of uncertainties in each voxel, the assessment of uncertainties of entire geological units, as well as the quantification of total model uncertainty represented by a single number \citep{wellmann2012uncertainties}.
@@
 
-		\section{1D geological reservoir modeling}\label{sec:1D_model}
+		\section{1D geological reservoir model}\label{sec:1D_model}
 		For simple understanding and a preliminary assessment of the Bayesian statistical methods and the loss function described above, we first apply them to a conceptual one-dimensional hydrocarbon system case.
@@
 			\end{figure}
-			\subsection{Construction of the 1D geological model}\label{sec:1D_construction}	
+			\subsection{Construction of the uncertain 1D geological model}\label{sec:1D_construction}	
 			\citet{delaVarga2016} constructed a simple geological model using three uncertain positions in vertical one-dimensional space, marking hypothetical boundaries of layers in a subsurface column.
@@
 The basic input data for the interpolation of the geological features is comprised of 3D point coordinates for layer interfaces and fault surfaces, as well as measurements that indicate respective dip directions and angles.
-In a Python environment, the data is imported as comma-separated values (CSV) using GemPy.
+In a Python environment, the data is imported as comma-separated values (see Figures \ref{fig:CSV_input1} to \ref{fig:CSV_input3}) using GemPy.
 From the generated data frame, GemPy is able to interpolate surfaces following the potential-field method explained in Section \ref{sec:struc_geo_modeling} and compute voxel-based 3D geological models (see Figure \ref{fig:3Dmodel_construction}).

### 2017-11-27 Risk analysis/Text/Layout/Layout/chapter3_results.tex
minor corrections

@@
 
-		\section{3D geological model results}
+		\section{3D geological reservoir model results}
 		In the following, we present the results from applying the methods to the synthetic 3D structural geological model.
@@
 			\end{tabular}
-			\caption{SSF-likelihood determined by a normal distributions as used for posterior model V (3D).}
+			\caption{SSF likelihood determined by a normal distributions as used for posterior model V (3D).}
 			\label{tab:ML5_likelihoods}

### 2017-11-28 Risk analysis/Text/Layout/Layout/MSc_layout_Fabian.tex
minor corrections
Added convergence diagnostics and Bayes network. Almost finished!

@@
 Special thanks go to the modeling team.
-To Alexander Schaaf, for helping me when Miguel could not be found and reminding me to take a break now and then.
+To Alexander Schaaf, for helping me with GemPy when Miguel could not be found and reminding me to take a break now and then.
 To Marco van Veen, for endless inciting and exhilarating discussions, but also for the many shared adventures during our geoscientific studies.\\
@@
     \begin{acronym}%
-    	\acro{CSV}{Comma-separated values}%
     	\acro{MAP}{Maximum a posteriori}%
@@
     	GemPy and PyMC are both fully open source and to be used in a Python environment.
-At the time of writing this, GemPy is still under development (Version 0.997).
+At the time of writing this, GemPy is still under development (version 0.997).
 It can be obtained from \url{https://github.com/cgre-aachen/gempy}.
@@
     	\centering
-    	\includegraphics[width=0.50\textwidth]{Figures/Appendix/update_goodseal1}
+    	\includegraphics[width=0.5\textwidth]{Figures/Appendix/update_goodseal1}
     	\caption{Prior (A1) and posterior distributions (B1) of the layer boundary positions in depth and respective representations (A2, B2) for 1D inference case II.}\label{fig:update_goodseal1}
@@
 		\centering
-		\includegraphics[width=0.50\textwidth]{Figures/Appendix/update_smallres1}
+		\includegraphics[width=0.5\textwidth]{Figures/Appendix/update_smallres1}
 		\caption{Prior (A1) and posterior distributions (B1) of the layer boundary positions in depth and respective representations (A2, B2) for 1D inference case III.}\label{fig:update_smallres1}
@@
 
+	\begin{figure}[h]
+		\begin{subfigure}{1\textwidth}
+			\centering
+			\includegraphics[width=1\linewidth]{Figures/Appendix/ML1/Geweke_ML1.eps}
+			\caption{Geweke plots for uncertain priors.}
+			%\label{fig:sfig1}
+		\end{subfigure}%
+		\\
+		\begin{subfigure}{1\textwidth}
+			\centering
+			\includegraphics[width=1\linewidth]{Figures/Appendix/ML1/Traces_ML1.eps}
+			\caption{Traces for uncertain priors.}
+			%\label{fig:sfig2}
+		\end{subfigure}
+		\caption{3D posterior model I: Geweke convergence analysis (a) and plotted traces (b) for uncertain priors.
+In (a), the standard deviation between the mean value of one trace interval and the following is represented by the Z-score.
+Scores in the range between -2 and 2 indicate convergence of the MCMC sampling process.}
+		\label{fig:gew_ML1}
+	\end{figure}
+
+	\begin{figure}[h]
+		\begin{subfigure}{1\textwidth}
+			\centering
+			\includegraphics[width=1\linewidth]{Figures/Appendix/ML3/Geweke_ML3.eps}
+			\caption{Geweke plots for uncertain priors.}
+			%\label{fig:sfig1}
+		\end{subfigure}%
+		\\
+		\begin{subfigure}{1\textwidth}
+			\centering
+			\includegraphics[width=1\linewidth]{Figures/Appendix/ML3/Traces_ML3.eps}
+			\caption{Traces for uncertain priors.}
+			%\label{fig:sfig2}
+		\end{subfigure}
+		\caption{3D posterior model II: Geweke convergence analysis (a) and plotted traces (b) for uncertain priors.
+In (a), the standard deviation between the mean value of one trace interval and the following is represented by the Z-score.
+Scores in the range between -2 and 2 indicate convergence of the MCMC sampling process.}
+		\label{fig:gew_ML2}
+	\end{figure}
+
+	\begin{figure}[h]
+		\begin{subfigure}{1\textwidth}
+			\centering
+			\includegraphics[width=1\linewidth]{Figures/Appendix/ML2/Geweke_ML2.eps}
+			\caption{Geweke plots for uncertain priors.}
+			%\label{fig:sfig1}
+		\end{subfigure}%
+		\\
+		\begin{subfigure}{1\textwidth}
+			\centering
+			\includegraphics[width=1\linewidth]{Figures/Appendix/ML2/Traces_ML2.eps}
+			\caption{Traces for uncertain priors.}
+			%\label{fig:sfig2}
+		\end{subfigure}
+		\caption{3D posterior model III: Geweke convergence analysis (a) and plotted traces (b) for uncertain priors.
+In (a), the standard deviation between the mean value of one trace interval and the following is represented by the Z-score.
+Scores in the range between -2 and 2 indicate convergence of the MCMC sampling process.}
+		\label{fig:gew_ML3}
+	\end{figure}
+
+	\begin{figure}[h]
+		\begin{subfigure}{1\textwidth}
+			\centering
+			\includegraphics[width=1\linewidth]{Figures/Appendix/ML4/Geweke_ML4.eps}
+			\caption{Geweke plots for uncertain priors.}
+			%\label{fig:sfig1}
+		\end{subfigure}%
+		\\
+		\begin{subfigure}{1\textwidth}
+			\centering
+			\includegraphics[width=1\linewidth]{Figures/Appendix/ML4/Traces_ML4.eps}
+			\caption{Traces for uncertain priors.}
+			%\label{fig:sfig2}
+		\end{subfigure}
+		\caption{3D posterior model IV: Geweke convergence analysis (a) and plotted traces (b) for uncertain priors.
+In (a), the standard deviation between the mean value of one trace interval and the following is represented by the Z-score.
+Scores in the range between -2 and 2 indicate convergence of the MCMC sampling process.}
+		\label{fig:gew_ML4}
+	\end{figure}
+
+	\begin{figure}[h]
+		\begin{subfigure}{1\textwidth}
+			\centering
+			\includegraphics[width=1\linewidth]{Figures/Appendix/ML5/Geweke_ML5.eps}
+			\caption{Geweke plots for uncertain priors.}
+			%\label{fig:sfig1}
+		\end{subfigure}%
+		\\
+		\begin{subfigure}{1\textwidth}
+			\centering
+			\includegraphics[width=1\linewidth]{Figures/Appendix/ML5/Traces_ML5.eps}
+			\caption{Traces for uncertain priors.}
+			%\label{fig:sfig2}
+		\end{subfigure}
+		\caption{3D posterior model V: Geweke convergence analysis (a) and plotted traces (b) for uncertain priors.
+In (a), the standard deviation between the mean value of one trace interval and the following is represented by the Z-score.
+Scores in the range between -2 and 2 indicate convergence of the MCMC sampling process.}
+		\label{fig:gew_ML5}
+	\end{figure}
     %\section{Another test section}

### 2017-11-28 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Added convergence diagnostics and Bayes network. Almost finished!

@@
 This is used as the so-called prior probability on which Bayesian updating is based.
-The beliefs about the occurrence of an event are revalued in the presence of additional information, i.e.
-the observation of new evidence $y$.
+The beliefs about the occurrence of an event are revalued in the presence of additional information, i.e.\ the observation of new evidence $y$.
 These observations are included as likelihoods $p(y|\theta)$.
@@
         The resulting posterior distribution can be used to acquire point estimates for the true state of nature $\theta$.
-Common and simple examples for such estimators are the mode (i.e.
-the generalized maximum likelihood estimate), the mean and the median of a distribution \citep{berger2013stat}.
+Common and simple examples for such estimators are the mode (i.e.\ the generalized maximum likelihood estimate), the mean and the median of a distribution \citep{berger2013stat}.
 The presentation of a point estimate should usually come with a measure for its estimation error.
@@
         \subsection{Expected loss and loss functions}\label{sec:loss} 
-        Loss is a statistical measure of how "bad" an estimate is, i.e. how much is lost by making a certain decision. Gains are considered by statisticians as negative losses.
+        Loss is a statistical measure of how "bad" an estimate is, i.e.\ how much is lost by making a certain decision. Gains are considered by statisticians as negative losses.
         The magnitude of incurred loss related to an estimate is defined by a loss function, which is a function of the estimate of the parameter and the true value of the parameter \citep{wald1950statistical, davidson2015}:
@@
 Both are simple to understand and commonly used \citep{davidson2015, hennig2007}.\\
-        As implied by its name, the absolute-error loss function returns loss as the absolute error, i.e. the difference between the estimate and the true parameter:        
+        As implied by its name, the absolute-error loss function returns loss as the absolute error, i.e.\ the difference between the estimate and the true parameter:        
         \begin{equation}\label{eq:AbsLossFunction}
@@
         Both of these standard loss functions are symmetric and can be described as objectively aiming at a high precision in estimating the true parameter value (see Figure \ref{fig:abs_sqr_det}).
-        \citet{davidson2015} and \citet{hennig2007} propose that it might be useful to move away from these type of objective loss functions to the design of customized loss functions that specifically reflect an individual's (i.e.
-the decision maker's) objectives, preferences and outcomes.
+        \citet{davidson2015} and \citet{hennig2007} propose that it might be useful to move away from these type of objective loss functions to the design of customized loss functions that specifically reflect an individual's (i.e.\ the decision maker's) objectives, preferences and outcomes.
 \citet{hennig2007} argue that choosing and designing a loss function involves the translation of informal aims and interests into mathematical terms.
@@
         The standard loss functions defined above are symmetric, but can easily be adapted to be asymmetric, for example by weighting errors on the negative side stronger than those on the positive side.
-Preference over estimates larger than the true value (i.e.
-overestimation) is thus incorporated in an uncomplicated way \citep{davidson2015, hennig2007}.
+Preference over estimates larger than the true value (i.e.\ overestimation) is thus incorporated in an uncomplicated way \citep{davidson2015, hennig2007}.
 Much more complicated designs of loss functions are possible, depending on purpose, objective and application.
@@
         The presence of uncertainty during decision making implies that the true parameter is unknown and thus the truly incurred loss $L(\theta,a)$ cannot be known at the time of making the decision.
-The Bayesian perspective considers unknown parameters as random variables and samples that are drawn from the posterior distribution as possible realizations of the unknown parameter, i.e.
-all possible true values are represented by this distribution.
+The Bayesian perspective considers unknown parameters as random variables and samples that are drawn from the posterior distribution as possible realizations of the unknown parameter, i.e.\ all possible true values are represented by this distribution.
 A probabilistic alternative to the actual loss is to consider each decision's expected loss and to make a decision that is optimal in relation to this expected loss \citep{berger2013stat, davidson2015}.\\
@@
         	In a chosen model ($M$), this is achieved by linking parameters and data through deterministic operations to the likelihood functions.
-It is pointed out by \citet{delaVarga2016} that any combination of parameter-observation connections is allowed, i.e.
-not all parameters need to be necessarily connected to all observed data.
+It is pointed out by \citet{delaVarga2016} that any combination of parameter-observation connections is allowed, i.e.\ not all parameters need to be necessarily connected to all observed data.
 After all conditional probabilities have been set up, the Bayes Theorem (Equation \ref{eq:BayesTheorem}) is applied to attain the posterior \citep{delaVarga2016}.
@@
 In the case of rejection, sampling is reverted to the previous accepted step \citep{schaaf2017, delaVarga2016}.
-After performing the complete number of $T$ iterations, all accepted sampling locations (i.e.
-the trace) are returned.
+After performing the complete number of $T$ iterations, all accepted sampling locations (i.e.\ the trace) are returned.
 The intention behind this concept it to achieve convergence of the sampling algorithm towards areas of high probability \citep{davidson2015}.\\
@@
         This implicit method relies on the interpolation of a potential field scalar function $T(\vec{x})$ of any point $\vec{x}$ in a 3D space, and thus reflects the geometry of geological structures \citep{calcagno2008geological}.
-Modeling $T(\vec{x})$ is achieved by cokriging two forms of data: (i) contact points on geological interfaces through increments of the potential field $T(\vec{x})-T(\vec{x}^{\,'})$ and (ii) orientation data as gradients, i.e.
-partial derivatives of the potential field $\delta T(\vec{x})/\delta u_\beta$ in each directon $u$ \citep{calcagno2008geological}.
+Modeling $T(\vec{x})$ is achieved by cokriging two forms of data: (i) contact points on geological interfaces through increments of the potential field $T(\vec{x})-T(\vec{x}^{\,'})$ and (ii) orientation data as gradients, i.e.\ partial derivatives of the potential field $\delta T(\vec{x})/\delta u_\beta$ in each directon $u$ \citep{calcagno2008geological}.
 The respective estimator is defined as follows:
@@
 Python is thus a straightforward tool for the implementation of central components of Bayesian analysis, such as custom statistical distributions and samplers \citep{salvatier2016pymc3}.\\
-		The 3D geological modeling step in this work is implemented using GeMpy, an open-source, Python-based software that is able to generate and visualize complex 3D structural geological models based on the potential field interpolation method elaborated in Section \ref{sec:struc_geo_modeling} \citep{delaVarga2017gempy}.
+		The 3D geological modeling step in this work is implemented using GemPy, an open-source, Python-based software that is able to generate and visualize complex 3D structural geological models based on the potential field interpolation method elaborated in Section \ref{sec:struc_geo_modeling} \citep{delaVarga2017gempy}.
 Its design allows for its application in a probabilistic setting as elaborated below.
@@
 Different model fitting techniques are provided in PyMC, such as the \textit{maximum a posteriori} (MAP) method and several MCMC sampling methods, including the Adaptive Metropolis introduced in Section \ref{sec:mcmc}.
+Furthermore, several approaches for checking MCMC convergence are included in PyMC.
+Here, we rely on a time-series method proposed by \citet{geweke1991evaluating}.
+Convergence is hereby assessed by comparing mean and variance of subsequent intervals in the sampling chain.
+Resulting scores (Z-scores) can be plotted using PyMC.
+If a trace has converged, most of the score points should be found within two standard deviations around zero.
+Also, the initial part of a sampling chain (before convergence) can be discarded by declaring a "burn-in" phase \citep{geweke1991evaluating, patil2010pymc, davidson2015}.\\
 		The components which are used to construct a statistical model, are represented by \textit{Deterministic} or \textit{Stochastic} functions or variables in PyMC \citep{salvatier2016pymc3}.
@@
         \end{enumerate}
-        While the first condition represents the complete absence of uncertainty everywhere in the model, the second condition is met in the opposite case of full uncertainty in the whole model space, i.e.
-all outcomes are equally likely in each cell \citep{schaaf2017}.
+        While the first condition represents the complete absence of uncertainty everywhere in the model, the second condition is met in the opposite case of full uncertainty in the whole model space, i.e.\ all outcomes are equally likely in each cell \citep{schaaf2017}.
 From this, the following equation for application in geological modeling was derived \citep{Mann1993,wellmann2012uncertainties}:
@@
         	\item \textbf{Reservoir formations:} Suitable lithologies, in which hydrocarbon volumes can be stored, have to be present.
-For this, a suitable percentage of void space, i.e.
-a sufficiently high porosity, is required in a rock formation.
+For this, a suitable percentage of void space, i.e.\ a sufficiently high porosity, is required in a rock formation.
 Further significant factors are the connectivity of the porosity through pore throats, as well as pore geometry and resulting permeability.
@@
 Traps can also be "filled to seal capacity" when the seal capacity is exceeded by the buoyant pressure of a fill column of greater height.
-Lower accumulations can also result from insufficient migration, i.e.
-low charge volumes of oil or gas in the first place.
+Lower accumulations can also result from insufficient migration, i.e.\ low charge volumes of oil or gas in the first place.
 Such a trap is referred to as "charge limited" \citep{dolson2016basics}.
@@
 The two reservoir layers in both sides are juxtaposed, so that the hydrocarbon accumulation potential is controlled by three points of possible leak pathways: (1) the anticlinal spill point, (2) leakage upwards along the fault and (3) leakage across the fault enabled by the juxtaposition.
-Possibilities (2) and (3) depend on the transport properties of the fault zone, i.e.
-potential fault sealing.
+Possibilities (2) and (3) depend on the transport properties of the fault zone, i.e.\ potential fault sealing.
 Given leakage along the fault (2), that size volume would be reduced to a small relatively small volume solely defined by the 4-way closure down to maximum contact of the reservoir-seal boundary with the fault plane.
@@
         The 3D structural model in this work (see Section \ref{sec:3D_design}) is designed in a way that a potential conventional hydrocarbon trap is included in the form of anticlinal folding combined with normal faulting of a reservoir-seal sequence.
-Results from modeling are to be evaluated in a way that respective estimations on the basis of loss functions can be conducted, i.e.
-Bayesian decision theory is applicable.
+Results from modeling are to be evaluated in a way that respective estimations on the basis of loss functions can be conducted, i.e.\ Bayesian decision theory is applicable.
 In the typical petroleum industrial setting, the main question is one of how much of the resource can be produced and how high the return on investment will be \citep{dean2007volumetric}.
@@
         \subsection{Original oil-in-place and recoverable volumes as value measures}
-        Before pressure and production tests have been conducted (i.e.
-before production has started), volumetric estimation is the only approach to assess the amount of hydrocarbons in place in a reservoir.
+        Before pressure and production tests have been conducted (i.e.\ before production has started), volumetric estimation is the only approach to assess the amount of hydrocarbons in place in a reservoir.
 From this value, recoverable reserves can be deduced based on an estimated recovery factor \citep{dean2007volumetric, morton1993development}.\\
@@
         %For most factors, it seems most suitable to define a normal distribution around a known typical value or global average, as it will return a high probability for common scenarios and occasionally produce cases of rare, more extreme values.\\
-        Attaining the hydrocarbon-filled rock volume (i.e.
-the maximum trap volume) and calculating the ROV is a central step of the model evaluation.
+        Attaining the hydrocarbon-filled rock volume (i.e.\ the maximum trap volume) and calculating the ROV is a central step of the model evaluation.
 Model realizations are hereby assigned the economic value on which decisions are to be based.
@@
 The median (in this case also mean and average) is returned for the Bayes action, when using the standard symmetric absolute-error loss function (I).
-Assigning a stronger weight on overestimation steepens the curve on the right hand side and shifts the minimum to the left, i.e.
-to a lower estimate (II).
+Assigning a stronger weight on overestimation steepens the curve on the right hand side and shifts the minimum to the left, i.e.\ to a lower estimate (II).
 Steepening is significantly reinforced by the introduction of fatal overestimation in (III).
@@
 This is assumed to be worse than non-fatal overestimation, but clearly better than fatal overestimation.
-No already owned resources are wasted, it is only the potential value that is lost, i.e.
-opportunity costs that arise from completely discarding a reservoir with a potential gain equal to the positive true value.
+No already owned resources are wasted, it is only the potential value that is lost, i.e.\ opportunity costs that arise from completely discarding a reservoir with a potential gain equal to the positive true value.
 Fatal underestimation is weighted using a third factor:
@@
 However, bidding correctly on a higher value, will also return a greater gain.
-It is assumed that risk-friendly actors care less about fatal underestimation, i.e.
-they will rather develop a project than discard it.
+It is assumed that risk-friendly actors care less about fatal underestimation, i.e.\ they will rather develop a project than discard it.
 In the loss function, risk is simply included using a risk factor \textit{r} which alters the incurred losses respectively:
@@
 		This trap, more specifically the trap volume, is defined as the central feature of economic interest.
-For conducting simple and straightforward volumetric calculations, it is assumed that found closed traps are always filled to spill with oil, i.e.
-the complete trap volume is hydrocarbon filled and the OOIP can be attained over this volume.\\
+For conducting simple and straightforward volumetric calculations, it is assumed that found closed traps are always filled to spill with oil, i.e.\ the complete trap volume is hydrocarbon filled and the OOIP can be attained over this volume.\\
 		The maximum trap volume is assumed to equate the hydrocarbon-filled rock volume $A * h$ in Equation \ref{eq:OOIP} and is thus to be used to valuate each realization of the geological model.
@@
 This way, the effect of incorporating additional information in the form of likelihoods is to be assessed from a perspective of Bayesian decision theory.
-It is furthermore assumed that the hydrocarbon trap volume is directly linked to a project development decision, i.e.
-investment and allocation of resources is represented by bidding on a volume estimate.
+It is furthermore assumed that the hydrocarbon trap volume is directly linked to a project development decision, i.e.\ investment and allocation of resources is represented by bidding on a volume estimate.
 
@@
 In (C) and (D), the inclusion of parameter uncertainties is presented.
-Colors indicate certain layer bottoms (i.e.
-boundaries) which are assigned a shared normal distribution on which their z-positional uncertainties are based (C).
+Colors indicate certain layer bottoms (i.e.\ boundaries) which are assigned a shared normal distribution on which their z-positional uncertainties are based (C).
 All points in the hanging wall are additionally assigned a fault offset uncertainty based on a skew normal distribution (D).
@@
 			\includegraphics[width=1\textwidth]{Figures/Trap_Cond_H.png}
-			\caption{Illustration of the process of trap recognition, i.e.
-the conditions that have to be met by a model voxel, to be accepted as belonging to a valid trap.
+			\caption{Illustration of the process of trap recognition, i.e.\ the conditions that have to be met by a model voxel, to be accepted as belonging to a valid trap.
 A voxel has to be labeled as part of the target reservoir formation (1) and positioned in the footwall (2).
@@
 This condition is applicable  under the assumption, that connection to the borders of the model lead to leakage.
-A trap is thus defined as a closed system in this model and trap closure is assumed to be void outside of the space of information, i.e.
-the model space.
+A trap is thus defined as a closed system in this model and trap closure is assumed to be void outside of the space of information, i.e.\ the model space.
 		\end{enumerate}
@@
 These are thus a controling factor for spill-related migration from respective structural traps.
-For anticlinal traps, closure can consequently be defined as the distance between the saddle point (i.e.
-spill point) and maximal point of the trap \citep{collignon2015fold_linkage}.\\
+For anticlinal traps, closure can consequently be defined as the distance between the saddle point (i.e.\ spill point) and maximal point of the trap \citep{collignon2015fold_linkage}.\\
 			Regarding a surface defined by $f(x,y)$, a local maximum at $(x_0,y_0,z_0)$ would resemble a hill top \citep{guichard2013calculus}.
 Local maxima will be found looking at the cross-sections in the planes $y = y_0$ and $x = x_0$.
-Furthermore, the respective partial derivatives (i.e.
-gradients) $\frac{\delta z}{\delta x}$ and $\frac{\delta z}{\delta y}$ will equal zero at $x_0$ and $y_0$, i.e.
-that the extremum is a stationary point \citep{guichard2013calculus, weisstein2017saddlepoint}.
+Furthermore, the respective partial derivatives (i.e.\ gradients) $\frac{\delta z}{\delta x}$ and $\frac{\delta z}{\delta y}$ will equal zero at $x_0$ and $y_0$, i.e.\ that the extremum is a stationary point \citep{guichard2013calculus, weisstein2017saddlepoint}.
 In the context of a geological reservoir system, such a hill can be regarded as a representation of an anticlinal structural trap.
@@
 In a geological setting, these surface maxima may be interpreted as two dome-shaped four-way closure traps.
-Spill-related migration between the two structures would take place over the saddle point, i.e.
-the spill point.}\label{fig:saddle_point_maxima}
+Spill-related migration between the two structures would take place over the saddle point, i.e.\ the spill point.}\label{fig:saddle_point_maxima}
 			\end{figure}
@@
 			%\end{figure}		
-			Figure \ref{fig:saddle_point_maxima} can be seen as a representation of a point of spill between two dome structures (i.e.
-surface maxima or "hills") which is defined by the surface's saddle point between both.\\
+			Figure \ref{fig:saddle_point_maxima} can be seen as a representation of a point of spill between two dome structures (i.e.\ surface maxima or "hills") which is defined by the surface's saddle point between both.\\
 			In GemPy, layer boundary surfaces are returned in the form of discretized arrays of simplices and vertices.
@@
 Considering the surface as a seal, consequent buoyant flow directions are indicated by arrows.
-Such flow would be directed away from minima and towards maxima, with the saddle point indicating a junction and divide, i.e.
-a possible spill point.}\label{fig:saddle_point_flow}
+Such flow would be directed away from minima and towards maxima, with the saddle point indicating a junction and divide, i.e.\ a possible spill point.}\label{fig:saddle_point_flow}
 			\end{figure}
@@
 At the same time, this point will be a local minimum on the plane perpendicular to the normal fault and thus equal a saddle point.
-However, higher magnitudes of fault displacement can lead to breaking of the seal, i.e.
-the juxtaposition of the trap structure with overlying formations which are assumed to be complete permeable.
+However, higher magnitudes of fault displacement can lead to breaking of the seal, i.e.\ the juxtaposition of the trap structure with overlying formations which are assumed to be complete permeable.
 In this case, the maximum footwall contact of the surface with the fault is to be defined as cross-fault leak point.\\
@@
 Juxtaposition areas can be attained using a topology function that is part of GemPy.
-On the base of this, a new function is created, with which the $z$-extent, i.e.
-the height of relevant juxtapositions can be deduced.
+On the base of this, a new function is created, with which the $z$-extent, i.e.\ the height of relevant juxtapositions can be deduced.
 These are juxtapositions of all layers on which the seal slipped and seal-overlying formations.
@@
 		First of all, a method to determine layer thicknesses in the geological model realizations has to be introduced.
-Here, as single X- and Y-point, i.e.
-a 1D column in $z$ in the footwall of the model is chosen as a measure for the thickness of the layers.
+Here, as single $x$- and $y$-point, i.e.\ a 1D column in $z$ in the footwall of the model is chosen as a measure for the thickness of the layers.
 In this column, point coordinates for the layer interfaces had been defined in the original data frame and can be observed as they are altered in the interpolation data frame with each probabilistic modeling iteration.
@@
 Furthermore, for every simulation, the mechanisms which ultimately determined the trap closure in every iteration, such as the spill point and cross-fault leakage, are traced as trap volume controls factors.
+
+		\begin{figure}[h]
+			\centering
+			\includegraphics[width=1\textwidth]{Figures/model_network}
+			\caption{Illustration of our probabilistic model (3D) as a hierarchical Bayesian network after \citet{koller2009probabilistic}.
+\textit{Stochastic} nodes are represented by ellipses (blue for priors, yellow for likelihoods).\ \textit{Deterministic} functions are depicted as triangles in red.
+Arrows indicate direct connections from parent to child nodes.
+We included the most important functions that contribute to the automatic trap recognition and volumetric calculation in our structural geological model.}\label{fig:model_network}
+		\end{figure}

### 2017-11-28 Risk analysis/Text/Layout/Layout/chapter3_results.tex
Added convergence diagnostics and Bayes network. Almost finished!

@@
 				\toprule
-				(a) Inference case I - Likelihoods: layer thicknesses (normal distributions)\\  
+				(a) Inference case I - Likelihoods: Layer thicknesses (normal distributions)\\  
 				\midrule 
@@
 				\toprule
-				(b) Inference case II - Likelihoods: layer thicknesses (normal)\\
+				(b) Inference case II - Likelihoods: Layer thicknesses (normal distributions)\\
 				\midrule 
@@
 				\toprule
-				(c) Inference case III - Likelihoods: layer thicknesses (normal)\\
+				(c) Inference case III - Likelihoods: Layer thicknesses (normal distributions)\\
 				\midrule 
@@
 For each simulation, 1,000 sampling steps were conducted, with a burn-in phase of 50 iterations for MCMC sampling.
-However, for computational reasons, Shannon entropy visualization was based on 200 sampling steps and a burn-in phase of 50 iterations.
-We consider in particular the following measures for the comparison of results: (1) Shannon entropy, (2) occurrence of trap control mechanisms, (3) resulting recoverable oil volumes and (4) consequent realization of expected losses and related Bayes actions (i.e.
-decisions).
+For computational reasons, Shannon entropy visualization was based on 200 sampling steps and a burn-in phase of 50 iterations.
+Convergence diagnostic results (Geweke and trace plots for the uncertain priors) can be found in the Appendix (Figures \ref{fig:gew_ML1} to \ref{fig:gew_ML5}).
+Considering that the majority of Z-score points in the Geweke plot are found within two standard deviations around zero, we deem that suitable convergence was achieved for the purpose of this work.
+For the comparison of results, we consider in particular the following measures : (1) Shannon entropy, (2) occurrence of trap control mechanisms, (3) resulting recoverable oil volumes and (4) consequent realization of expected losses and related Bayes actions (i.e.\ decisions).
 
@@
 			\toprule
-			(a) Uncertainty: layer boundaries $z$-position (normal distributions)\\  
+			(a) Uncertainty: Layer boundaries $z$-position (normal distributions)\\  
 			\midrule 
@@
 			\toprule
-			(b) Uncertainty: normal fault offset (skew normal distribution)\\
+			(b) Uncertainty: Normal fault offset (skew normal distribution)\\
 			\midrule
@@
 				\toprule
-				Model I - Likelihoods: layer thicknesses (normal distributions)\\  
+				Model I - Likelihoods: Layer thicknesses (normal distributions)\\  
 				\midrule 
@@
 				\toprule
-				Model II - Likelihoods: layer thicknesses (normal distributions)\\  
+				Model II - Likelihoods: Layer thicknesses (normal distributions)\\  
 				\midrule 
@@
 				\toprule
-				Model III - Likelihoods: layer thicknesses (normal distributions)\\  
+				Model III - Likelihoods: Layer thicknesses (normal distributions)\\  
 				\midrule 
@@
 				\toprule
-				Model IV - Likelihood: layer thicknesses (normal distributions)\\  
+				Model IV - Likelihood: Layer thicknesses (normal distributions)\\  
 				\midrule 
@@
 				\toprule
-				Likelihood: Shale Smear Factor (normal distribution)\\
+				Model IV - Likelihood: Shale Smear Factor (normal distribution)\\
 				\midrule

### 2017-11-28 Risk analysis/Text/Layout/Layout/chapter4_discussion.tex
Added convergence diagnostics and Bayes network. Almost finished!

@@
 A mean located in the middle between two widely separated modes, which might be of low intrinsic uncertainty themselves, would indicate a much higher overall decision uncertainty, than a mean found in the middle of a narrow unimodal distribution.
-Pronounced bimodality between two extremes, i.e.
-high uncertainty, resulted in a wider separation of Bayes estimators.
+Pronounced bimodality between two extremes, i.e.\ high uncertainty, resulted in a wider separation of Bayes estimators.
 Reduction of the distribution to one mode conversely led to the convergence of different Bayes actions.
 A decrease in decision uncertainty furthermore was accompanied by a reduction in expected loss for each Bayes estimator.
-	Considering these observations, we derive that the degree of Bayes action convergence and respective expected losses, i.e.
-Bayes risks, can be considered measures for the state of knowledge and decision uncertainty at the moment of making a decision.
+	Considering these observations, we derive that the degree of Bayes action convergence and respective expected losses, i.e.\ Bayes risks, can be considered measures for the state of knowledge and decision uncertainty at the moment of making a decision.
 The better these are, the more similar the decisions of differently risk-affine actors and the lower their loss expectations are.
@@
 	\subsubsection{Impact of additional information on decision making}
-	Cases of various states of knowledge and decision uncertainties, i.e.
-different realizations of the value probability distribution, were attained by the implementation of Bayesian inference.
+	Cases of various states of knowledge and decision uncertainties, i.e.\ different realizations of the value probability distribution, were attained by the implementation of Bayesian inference.
 While in most settings, adding information via inference led to a reduction of decision uncertainty, this was not always the case.
 We observed that the impact on decision uncertainty, induced by Bayesian inference, is not necessarily strictly aligned with the change in uncertainty regarding model parameters and their combinations.
-In respect to this, two aspects seems to be of central importance: (1) "where" in the model uncertainty is reduced, i.e.
-in which spatial area or regarding which model parameters and (2) which possible outcome is enhanced in terms of probability.
+In respect to this, two aspects seems to be of central importance: (1) "where" in the model uncertainty is reduced, i.e.\ in which spatial area or regarding which model parameters and (2) which possible outcome is enhanced in terms of probability.
 An increased probability of a thick or thin seal in our 3D model, for example, equally reduced decision uncertainty significantly, by raising the probability of a positive or negative outcome, respectively.
@@
 By introducing likelihood functions, these modes were most often shifted or altered in their shape.
-The latter was observed as either an amplification of the mode, by raising its mean probability, while narrowing the range of probabilities (i.e.
-reducing the standard deviation), or the opposite, a diminishment of the mode up to total erasure of its probability.\\
+The latter was observed as either an amplification of the mode, by raising its mean probability, while narrowing the range of probabilities (i.e.\ reducing the standard deviation), or the opposite, a diminishment of the mode up to total erasure of its probability.\\
 	\subsubsection{Value distribution modes as possible model scenarios and potential decision options}	
@@
 		\caption{ROV distribution of the prior model subdivided by modes into scenarios which could pose decision options (marked by greens lines).
-Red lines indicated a discretization of decision options according fixed scales of resource allocations for a project.
-For both cases, zero equals the option to "take no action", i.e.
-to allocate no resources.}\label{fig:scenarios_options}
+Red lines indicate a discretization of decision options according to fixed scales of resource allocations for a project.
+For both cases, zero equals the option to "take no action", i.e.\ to allocate no resources.}\label{fig:scenarios_options}
 	\end{figure}\\
@@
 Due to this, trap volumes tended to occur primarily in distinct "filled-to-spill" scenarios.
-While being unrealistic, this enabled the recognition of pronounced influences related to the various control mechanisms.\\
-	For the purpose of a conceptual application, the customization of the loss functions was kept relatively simple.
+While being unrealistic, this enabled the recognition of pronounced influences related to the various control mechanisms.
+For computational reasons, we had to keep the number of MCMC sampling steps relatively low.
+For improved convergence and a higher reliability of results, more iterations should be executed.
+Nevertheless, we believe that for the purpose of this study, sufficient convergence and meaningful posterior results were achieved.\\
+	For a conceptual application, the customization of the loss functions was kept relatively simple.
 We argue that despite its simplicity, the design of the function and the incorporation of a risk factor was suitable to achieve a clear distinction between differently risk-affine actors and respective behavior.
@@
 The consideration of more details, without a basis on real data, would have furthermore required extensive speculation, which would have presumably impaired the generalization potential of respective observations.
+We defined risk affinity to be dependent on arbitrarily chosen risk factors, which led to according re-weighting.
+\citet{davidson2015} used risk parameters determined by the maximal loss each actor could incur.
+Other approaches could be based on more tangible values, for example by making risk attitude dependent on a fixed budget.
 While standard loss functions only return mean or median as estimates, loss function customization offers a flexible approach to express different objective and subjective weighting factors that are specified to certain decision-making environments and decision-makers' perspectives.

### 2017-11-29 Risk analysis/Text/Layout/Layout/MSc_layout_Fabian.tex
Almost ready to print

@@
 %
-\includepdf[pages={1}]{eidesstattliche_erklaerung.pdf}
+\includepdf[pages={1}]{eidesstattliche_erklaerung3.pdf}
 
@@
 I am grateful for the continuous motivation, constant positive feedback and especially the immense patience he showed as my supervisor and mentor.\\
-    Furthermore, I want to voice my deep gratitude to Professor Florian Wellmann, for welcoming me as a part of his research group and thus showing me, how exciting and inspiring it can be, to participate in a free and innovative academic environment.
+    Furthermore, I want to voice my gratitude to Professor Florian Wellmann, for welcoming me as a part of his research group and thus showing me how exciting and inspiring it can be to participate in a free and forward-thinking academic environment.
 With his interest, openness and positive attitude, he encouraged me to contribute to his group's work and to develop my own scientific ideas.\\
-    I also wish to thank Professor Janos Urai, for years of fascinating lectures on structural geology, hydrocarbon exploration and production, and related risk analysis.
-His courses provided me with many insights and served as a sources of inspiration and information that I was able to include in this project.\\
+    I also wish to thank Professor Janos Urai, for years of fascinating lectures on structural geology, hydrocarbon exploration and production, as well as related risk analysis.
+His courses provided me with many insights and served as sources of inspiration and information that I was able to include in this project.\\
     In general, I am grateful to my friends and fellow researchers from the CGRE group, especially for creating such a pleasant working atmosphere.
@@
     And of course, I owe my deepest gratitude to my family, for a lifetime of unconditional support.
-To my parents, for raising me to be curious.
-For enabling me to attain such excellent education and for providing me with the freedom to follow my individual interests and aspirations.
+To my parents, for raising me to be curious, for enabling me to attain such excellent education and for providing me with the freedom to follow my individual interests and aspirations.
     \vspace*{15mm}

### 2017-11-29 Risk analysis/Text/Layout/Layout/chapter1.tex
Almost ready to print

@@
 For this, we include Bayesian decision theory in the step of model evaluation.
-This is achieved by assigning an economic meaning to the structural model and designing a case-specific custom loss function to find decisions which are optimal related to the state of knowledge and the preferences of actors with different risk-affinities.
+This is achieved by assigning an economic meaning to the structural model and designing a case-specific custom loss function to find decisions which are optimal related to the state of knowledge and the preferences of actors with different risk affinities.
 More specifically, the models are designed to represent potential hydrocarbon reservoir systems.

### 2017-11-29 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Almost ready to print

@@
         \subsection{Bayesian inference}
-        Bayesian inference is most importantly characterized by its preservation of uncertainty, in contrast to standard statistical inference \citep{jaynes2003probability, gelman2014bayesian, box2011bayesian, harney2013bayesian, davidson2015}.
+        The principles of Bayesian inference have been presented and discussed extensively in literature (see \citet{jaynes2003probability}, \citet{box2011bayesian}, \citet{harney2013bayesian}, \citet{gelman2014bayesian} and \citet{davidson2015}).\ It is most importantly characterized by its preservation of uncertainty, in contrast to standard statistical inference.
 The Bayesian approach is widely seen as intuitive and inherent in the natural human perspective.
@@
 By using different values for this parameter, it can be represented how comfortable an individual is with being wrong and furthermore which "side of wrong" is preferred by this decision maker \citep{davidson2015}.
-This approach to expressing risk-affinities is used for the design of the custom loss functions in Section \ref{sec:LF_design}.
+This approach to expressing risk affinities is used for the design of the custom loss functions in Section \ref{sec:LF_design}.
 
@@
         In this work, the generated structural geological models are to be assigned an economic significance in order to approach a real setting, in which hypothetical actors could have an interest in taking a decision, and thus, to evaluate the influence of the Bayesian inference of the decision making process.
-Changes in Bayes actions are considered as a measure for this effect.
+Changes in Bayes actions are considered a measure for this effect.
 For this purpose, it is assumed that a hydrocarbon reservoir system is represented by each geological model realization.
-        Such hydrocarbon systems are typically comprised of:
+        Such hydrocarbon systems are typically comprised of the following elements:
         \begin{enumerate}
@@
 Basement formations in the geological model are simply defined to be permeable.
-        	\item \textbf{Seal}: These are formations that are able to halt hydrocarbons on their migration pathway as top, lateral and possibly bottom seals \citep{dolson2016basics, sorkhabi2005place}.
-Such sealing is generally defined by a reduction of pore space geometry relative to migration energy.
-Sealing is thus defined primarily by changes in stratigraphical composition to low porosities and micro-sized pore throats.
+        	\item \textbf{Seal}: These are formations that able to halt hydrocarbons on their migration pathway as top, lateral and possibly bottom seals \citep{dolson2016basics, sorkhabi2005place}.
+Such sealing is generally defined by a reduction of pore space geometry relative to migration energy, primarily due to changes in stratigraphical composition to low porosities and micro-sized pore throats.
 These are characterized by a high capillary entry pressure able to resist the buoyancy pressure of hydrocarbons.
@@
 Many different types of traps, in which hydrocarbons can accumulate, are possible.
-A common trap type is a structural four-way closure, i.e.\ as reservoir-seal system that was folded into closed anticlines, often resembling a dome and is thus predominantly defined by the structural closure relief and the cap rock sealing capacity.
+A common trap type is a structural four-way closure, i.e.\ as reservoir-seal system that was folded into closed anticlines, often resembling a dome.
+It is thus predominantly defined by the structural closure relief and the cap rock sealing capacity.
 Such a trap type can be considered to be closed, as the reservoir surface contours are closed in a circle or oval around a relief high.
@@
         	\includegraphics[width=0.5\textwidth]{Figures/trap_closure1}
-        	\caption{Conceptual examples of a dome-shaped four-way-closure and a simple fault trap from top view (relief contour) and cross section perspective (modified from \citet{dolson2016basics}.}\label{fig:trap_types}
+        	\caption{Conceptual examples of a dome-shaped four-way-closure (a) and a simple fault trap (b) from top view (relief contour) and cross section perspective (modified from \citet{dolson2016basics}.}\label{fig:trap_types}
         \end{figure}        
@@
 Cases might range from complete fault sealing to complete leakage along a fault zone.
-More complex in-between cases are possible, such as faults that leak across, but are sealed along their plane.
+More complex in-between cases are possible, such as faults that leak across but are sealed along their plane.
 This ultimately defines the depth to which a trap may be filled and whether this point is controlled by a point of leakage on the fault plane or a spill point related to the remainder of the 3D structural closure.
-Another decisive aspects to consider, is the nature of juxtapositions across the fault, especially regarding the possibility of cross-fault leakage \citep{van2003lateral}.
+Another decisive aspect to consider is the nature of juxtapositions across the fault, especially regarding the possibility of cross-fault leakage \citep{van2003lateral}.
 This is summarized in Figure \ref{fig:fault_trap_spills}.
@@
         \end{enumerate}
-        As the hydrocarbon systems in this work are purely synthetic, a pure oil accumulation is assumed for simplicity.
-Furthermore, defined formations are regarded as homogeneous in their lithological composition and petrological properties It is assumed, for example, that no shale is interbedded in the main sandstone reservoir unit.
+        As the hydrocarbon systems in this work are purely synthetic, we assume a pure oil accumulation for simplicity.
+Furthermore, defined formations are regarded as homogeneous in their lithological composition and petrological properties.
 Using the factors listed above, the equation for original oil-in-place (OOIP) is formulated as follows:
@@
         where OOIP is returned in m$^3$.
-The hydrocarbon-filled rock volume is defined by the drainage area $A$ in m$^2$ and the net pay thickness $h$ in $m$.
+The hydrocarbon-filled rock volume is defined by the drainage area $A$ in m$^2$ and the net pay thickness $h$ in m.
 Porosity $\phi$ and water saturation $S_W$ (interstitial water) are given in fractions of the rock volume.
 A dimensionless factor for the change in oil volume between reservoir conditions and standard conditions at surface is represented by the formation volume factor FvF.
-Thus, shrinkage of the oil volume brought to the surface is determined by $1/FvF$ \citep{dean2007volumetric, morton1993development}.\\
+Thus, shrinkage of the oil volume brought to the surface is determined by 1/FvF \citep{dean2007volumetric, morton1993development}.\\
         Subsequently, the effectively recoverable oil volumes (ROV) can be calculated by multiplying the OOIP with a recovery factor RF \citep{dean2007volumetric, morton1993development}:
@@
         Including the remaining factors as simple uncertain parameters would be easily possible and is usually done in the probabilistic estimation procedures commonly conducted in the petroleum industry \citep{wim2001guidelines, murtha1997monte, mudford2000valuing}.
-However, the focus of this work lies on the structural aspects of a hydrocarbon system.
-Respective results and observations might possibly be overprinted and distorted by the incorporation of other parameters of petrology and engineering as uncertain.
+However, results and observations related to structural geology might possibly be overprinted and distorted by incorporating other uncertain parameters belonging to petrology and engineering.
 Effects of interest might become unrecognizable.
-Thus, fixed values are used for the non-structural factors ($\phi=30\%$, $S_W=20\%$, FvF $=1.05$, RF $= 80\%$).
-These are chosen arbitrarily and optimistic, but within a realistic range for a very good conventional oil reservoir (for example values see \citet{morton1993development}, \citet{ahmed2011advanced} and \citet{dandekar2013petroleum}).
-Consequently, calculating the ROV from the maximum trap volume simply equates a rescaling of values, as proportions are conserved.
-However, we deem the translation of trap volumes to recoverable oil volumes as appropriate for a formally better consideration of the results in the respective context of decision making in the hydrocarbon sector.\\
+Thus, fixed values are used for the non-structural factors.
+These are chosen arbitrarily and optimistic ($\phi=30\%$, $S_W=20\%$, FvF $=1.05$, RF $= 80\%$), but within a realistic range for a very good conventional oil reservoir (for example values see \citet{morton1993development}, \citet{ahmed2011advanced} and \citet{dandekar2013petroleum}).
+Consequently, calculating the ROV from the maximum trap volume merely equates a rescaling of values, as proportions are conserved.
+Nevertheless, we deem the translation of trap volumes to recoverable oil volumes as appropriate for a formally better consideration of the results in the respective context of decision making in the hydrocarbon sector.\\
         %For most factors, it seems most suitable to define a normal distribution around a known typical value or global average, as it will return a high probability for common scenarios and occasionally produce cases of rare, more extreme values.\\
@@
 			\includegraphics[width=1\textwidth]{Figures/LF_4steps_normal2.png}
-			\caption{Single steps (I-IV) of customizing the loss function applied on an exemplary normal distribution ($\mu=0; \sigma=500$; illustrated n light blue).
+			\caption{Single steps (I-IV) of customizing the loss function applied to an exemplary normal distribution ($\mu=0; \sigma=500$; illustrated n light blue).
 The median (in this case also mean and average) is returned for the Bayes action, when using the standard symmetric absolute-error loss function (I).
@@
 
-		Combining these adaption steps and the conditions defined in them, results in the following customized loss function:
+		Combining these adaption steps and the conditions defined in them results in the following customized loss function:
 
@@
 
-		\subsection{Including different risk-affinities in the loss function}		
+		\subsection{Including different risk affinities in the loss function}		
 		It can be assumed that several actors in one sector or decision environment may have the same general loss function but different affinities concerning risks.
@@
 A certain false estimate or error might have a significantly stronger impact on a company which has a a generally lower market share and only few projects, than on a larger company which might possess a higher financial flexibility and for which one project is only one of many development options in a portfolio.\\
-		In the following, the loss function is further adapted to consider different risk-affinities of several actors.
+		In the following, the loss function is further adapted to consider different risk affinities of several actors.
 Representing risk behavior in a loss function can also be done in different ways and regarding different types of risks.
@@
 			\includegraphics[width=1\textwidth]{Figures/LFR_normal2.png}
-			\caption{Plotting of expected loss realizations after including the risk factor $r$ in the loss function (Equation \ref{eq:LFR_final}) for actors with risk-affinities ranging from risk-averse ($r$ = 0.5 and 0.75), over risk-neutral ($r$ = 1), to risk-friendly ($r$ = 1.25 and $r$ = 1.5).
+			\caption{Plotting of expected loss realizations after including the risk factor $r$ in the loss function (Equation \ref{eq:LFR_final}) for actors with risk affinities ranging from risk-averse ($r$ = 0.5 and 0.75), over risk-neutral ($r$ = 1), to risk-friendly ($r$ = 1.25 and $r$ = 1.5).
 The loss function was applied the same normal distribution ($\mu=0; \sigma=500$) as in Figure \ref{fig:LF_4steps_normal}.}\label{fig:LFR_normal}
-		\end{figure}
+		\end{figure}\\
 		According to this, for \textit{r} = 1 the risk-neutral loss function is returned, as no re-weighting takes place.
@@
 		The factor \textit{r} can take basically any positive values.
-However, since risk-neutrality is expressed by \textit{r} = 1, we consider values 0 $<$ \textit{r} $<$ 2 to be the most appropriate choices to represent both sides of risk-affinity equally here.
+However, since risk neutrality is expressed by \textit{r} = 1, we consider values 0 $<$ \textit{r} $<$ 2 to be the most appropriate choices to represent both sides of risk affinity equally here.
 
 		\section{1D geological reservoir model}\label{sec:1D_model}
-		For simple understanding and a preliminary assessment of the Bayesian statistical methods and the loss function described above, we first apply them to a conceptual one-dimensional hydrocarbon system case.
+		For simple understanding and a preliminary assessment of the Bayesian statistical methods, as well as the custom loss function described above, we first apply them to a conceptual one-dimensional hydrocarbon system case.
 The underlying model and basic approach are inherited from \citet{delaVarga2016}, but the parameters are adapted to more appropriately represent a reasonable geological petroleum system, consisting of a reservoir formation with overlying seal in the subsurface.
@@
 				\caption{Probability distributions for positions of layer boundaries in the subsurface and a respective representation using lines.
-These normal distributions are determined by the following values: Seal top: $\mu = 2000 m, \sigma = 7$; reservoir top: $\mu = 2050 m, \sigma = 25$; reservoir bottom: $\mu = 2200 m, \sigma = 45$.
+These normal distributions are determined by the following values: Seal top: $\mu = 2000~\text{m}, \sigma = 7~\text{m}$; reservoir top: $\mu=2050~\text{m}, \sigma = 25~\text{m}$; reservoir bottom: $\mu = 2200~\text{m}, \sigma = 45~\text{m}$.
 }\label{fig:1D_model}
@@
 Single results are expected to vary, depending on the uncertain input parameters defined in Section \ref{sec:3D_construction} below.
-The custom loss function \ref{eq:LFR_final} is to be applied on the resulting value distributions before and after performing Bayesian inference.
+The custom loss function \ref{eq:LFR_final} is to be applied to the resulting value distributions before and after performing Bayesian inference.
 This way, the effect of incorporating additional information in the form of likelihoods is to be assessed from a perspective of Bayesian decision theory.
@@
 
-		\subsection{Determination of the trap volume}
+		\subsection{Determination of the maximum trap volume}
 		In the course of this work, several algorithms are developed within the Python environment.
@@
 In the case of a suitable SSF and consequent fault sealing, leakage across this type of juxtaposition is discarded as irrelevant.
+For simplicity, the fault is considered to be sealing along its plane.
 The processes of finding the decisive spill point, as well as the determination of juxtapostions and SSF is explained below in further detail.
@@
 While the thickness likelihood functions are dependent on input parameters directly, the implementation of the SSF likelihood function requires a full computation of the model and extended algorithms of structural analysis.\\
-		Some representative cases of Bayesian inference applied on the 3D model are presented in Chapter \ref{cha:results}.
+		Some representative cases of Bayesian inference applied to the 3D model are presented in Chapter \ref{cha:results}.
 

### 2017-11-29 Risk analysis/Text/Layout/Layout/chapter3_results.tex
Almost ready to print

@@
 This results in a bimodal characterization with low probability of the ROV to be found between zero and 3 million m$^3$.\\
-		Consequently, applying the custom loss function to this distribution resulted in divergent Bayes actions for the different risk-affine actors (see Figure \ref{fig:M_prior} (b)).
+		Consequently, applying the custom loss function to this distribution resulted in divergent Bayes actions for the differently risk-affine actors (see Figure \ref{fig:M_prior} (b)).
 Although all actors find their estimators to be positive, only the risk-neutral to risk-friendliest actors come close to the described positive mode of the distribution (ROV $> 3*10^6$ m$^3$).

### 2017-11-29 Risk analysis/Text/Layout/Layout/chapter4_discussion.tex
Almost ready to print

@@
 These were mostly distinctly separated with little probability in between.
-This presumably resulted from the assumption that traps are always "filled-to-spill" in respect to spill and leak point, and that full leakage is assumed for any type of seal failure, resulting in an approximate discretization into respective scenarios.
+This presumably resulted from the assumption that traps are always "filled to spill" in respect to spill and leak point, and that full leakage is assumed for any type of seal failure, resulting in an approximate discretization into respective scenarios.
 	\begin{figure}[h]
@@
 	Considering the findings of this work, there is still many points that could be expanded on in future research.
-It would be of interest to apply the same overall concept and methods on an authentic case based on real datasets.
+It would be of interest to apply the same overall concept and methods to an authentic case based on real datasets.
 Given a realistic economic scenario, including capital and operational expenditures of a project, possibly a full net-present-value (NPV) analysis could be conducted.

### 2017-12-11 Risk analysis/Text/Layout/Layout/MSc_layout_Fabian.tex
FINAL (printed and handed in)

@@
 Markov chain Monte Carlo sampling is used to approximate posterior models of reduced uncertainty.
-Here, synthetic geological models are constructed to represent potential hydrocarbon system.
+Here, synthetic geological models are constructed to represent potential hydrocarbon systems.
 Algorithms for automatic trap recognition are developed to enable the reservoir valuation of prior and posterior models.
 These serve as a base for true value estimation by respective optimization of a case-specific loss function.
-This function is customized to reflect the decision making of differently risk-affine actors in this context.
+This function is customized to reflect the decision making of differently risk-affine actors.
 Results from conducting this for several inference cases show that the various Bayes estimators shift according to the characteristics of the underlying value distribution.
-While bimodality and overall uncertainty leads to separation, risk-averse and risk-friendly decisions converge and decrease in expected loss given more narrow unimodal distributions.
-The degree of decision convergence is considered a measure for the state of knowledge and decisive uncertainty at the moment of decision making.
-This uncertainty does not change aligned with model uncertainty but depends on alterations of critical parameters and respective interdependencies, relating to seal reliability in particular.
-Additionally, actors are affected differently by one set of information, as defined by their risk weighting.
+While bimodality and overall uncertainty leads to separation, risk-averse and risk-friendly decisions converge and decrease in expected loss given narrower unimodal distributions.
+The degree of decision convergence is considered a measure for the state of knowledge and its inherent uncertainty at the moment of decision making.
+This decisive uncertainty does not change aligned with model uncertainty but depends on alterations of critical parameters and respective interdependencies, in particular relating to seal reliability.
+Additionally, actors are affected differently by one set of information, as defined by their risk affinity.
 For good decision making, it appears to be most important to recognize for which parameters reduction of uncertainty leads to changes in decisions.

### 2017-12-11 Risk analysis/Text/Layout/Layout/chapter3_results.tex
FINAL (printed and handed in)

@@
 
-c			\subsection{Scoring}
+			\subsection{Scoring}
 			\begin{figure}[h]
@@
 Mean and median of the score distribution are clearly surpassed by the best estimate of the most risk-friendly actor (\textit{r} = 0.5), while for the most risk-averse actors (\textit{r} = 1.25 and \textit{r} = 1.5), the Bayes action equals a zero score estimate and thus the decision to take no action.
-It can also be recognized that the expected loss is generally lower for risk-friendlier actors on the side of positive estimates, which in this case is the relevant side for decision-making.
+It can also be recognized that the expected loss is generally lower for risk-friendlier actors on the side of positive estimates, which in this case is the relevant side for decision making.
 			%\subsection{Applying the custom loss function on the 1D reservoir scores}
@@
 The overall uncertainty seems thus to have been widely maintained and barely transformed into a problem of duality.\\
-				Application of the custom loss function (Equation \ref{eq:LFR_final}) is visualized in Figure \ref{fig:update_moderate2_3}, in which the expected losses are compared before (B1) and after (B2) inference.
+				Application of the custom loss function (Equation \ref{eq:LFR_final}) is visualized in Figure \ref{fig:update_moderate2_3} in which the expected losses are compared before (B1) and after (B2) inference.
 It is observable, that by adding information about layer thickness likelihoods, Bayes actions were shifted relative to the nature of the information.
@@
 			%\subsection{General 1D model inference results}
-			%This conceptual application in an abstract 1D geological model already serves to illuminate some of the main effects of Bayesian inference on decision-making.
+			%This conceptual application in an abstract 1D geological model already serves to illuminate some of the main effects of Bayesian inference on decision making.
 Three main mechanisms were observable in these cases:
@@
 			%The second mechanism, i.e.
-uncertainty reduction, can thus presumed to be most significant regarding the process of "good" decision-making, as actor's wish to minimize their expected loss from bidding on an estimate.
+uncertainty reduction, can thus presumed to be most significant regarding the process of "good" decision making, as actor's wish to minimize their expected loss from bidding on an estimate.
 In an optimal case, this would come in combination with a shift of probabilities (first mechanism) to more positive values, so that a higher economic value may be returned from the acquisition of information.
@@
 Considering that the majority of Z-score points in the Geweke plot are found within two standard deviations around zero, we deem that suitable convergence was achieved for the purpose of this work.
-For the comparison of results, we consider in particular the following measures : (1) Shannon entropy, (2) occurrence of trap control mechanisms, (3) resulting recoverable oil volumes and (4) consequent realization of expected losses and related Bayes actions (i.e.\ decisions).
+For the comparison of results, we consider in particular the following measures: (1) Shannon entropy, (2) occurrence of trap control mechanisms, (3) resulting recoverable oil volumes and (4) consequent realization of expected losses and related Bayes actions (i.e.\ decisions).
 
 		\subsection{3D - Prior model}\label{sec:prior_model3D}
-		To generate a set of reference results, a prior model was first generated by running a simple simulation of Monte Carlo sampling considering only the uncertainty of prior parameters.
+		To generate a set of reference results, a prior model was first generated by running a simple Monte Carlo simulation considering only the uncertainty of prior parameters.
 No likelihood functions were included.
-The prior parameter uncertainties, defined in Table \ref{tab:3D_prior_parameters} were chosen to be identical for all posterior simulations.
+The prior parameter uncertainties (defined in Table \ref{tab:3D_prior_parameters}) were chosen to be identical for all posterior simulations.
 		\begin{table}[h]
@@
 		\end{tabular}
-		\caption{Prior parameter uncertainties defined by distributions with respective means $\mu$, standard deviations $\sigma$ and shape factor $\alpha$, as used in all model simulations.
+		\caption{Prior parameter uncertainties defined by distributions with respective means $\mu$, standard deviations $\sigma$ and shape factor $\alpha$, as used in all model simulations (3D).
 Points belonging to the same layer interface share the same normal distribution (a).
@@
 			\end{tabular}
-			\caption{Layer thickness likelihoods determined by normal distributions as used for posterior model I.}
+			\caption{Layer thickness likelihoods determined by normal distributions as used for posterior model I (3D).}
 			\label{tab:ML1_likelihoods}
@@
 		\end{figure}		
-		For the simulation of this first posterior model, thickness likelihoods for the three central layers were included in the probabilistic model.
+		For the simulation of this first posterior model, we included thickness likelihoods for the three central layers in the probabilistic model.
 These were defined in a way that reinforced the probabilities of the relevant layers to be significantly thick (see Table \ref{tab:ML1_likelihoods}).
@@
 		Regarding trap volume control (Figure \ref{fig:ML1} (c)), only spill point-related migration ("SPILL POINT") and cross-fault leakage due to juxtaposition with permeable layers below the seal in the hanging wall ("LEAK UNDER") remained relevant in the posterior model.
-As no occurrences of leakage through overlying layers were observed, these two main mechanisms increased in frequency, while staying proportionally the same to one another.
-It can be presumed, that by introducing the respective likelihoods, a high thickness of the seal was ensured, as can also be recognized in Figure \ref{fig:ML1} (b).
-This not only made stratigraphic leakage in this model impossible, but also enforced a favorable SSF.
+As no occurrences of leakage through overlying layers were observed, these two main mechanisms increased in frequency while staying proportionally the same to one another.
+It can be presumed that by introducing the respective likelihoods, a high thickness of the seal was ensured, as can also be recognized in Figure \ref{fig:ML1} (b).
+This not only made stratigraphic leakage in this model impossible but also enforced a favorable SSF.
 A minor emergence of unrecognized control factors ("UNCLEAR") suggests that, for a few iterations, the model was possibly not correctly computable or the structural analysis failed for a different reason.\\
@@
 Additionally, expected losses (Bayes risks) were lowered overall.
-The posterior state of knowledge, provided by the updated model, encouraged even the most risk-averse decision-maker to expect recoverable oil volumes from 3.5 million m$^3$ upwards.
+The posterior state of knowledge, provided by the updated model, encouraged even the most risk-averse decision maker to expect recoverable oil volumes from 3.5 million m$^3$ upwards.
 
@@
 			\end{tabular}
-			\caption{Layer thickness likelihoods determined by normal distributions as used for posterior model II.}
+			\caption{Layer thickness likelihoods determined by normal distributions as used for posterior model II (3D).}
 			\label{tab:ML3_likelihoods}
 		\end{table}
-		For the second model, thickness likelihoods were defined in a way that reinforced the probability of a very thick seal, but at the same time restricted the reservoir formation to a likely significantly thinner thickness of around 50 m (see Table \ref{tab:ML3_likelihoods}).
-The results (see Figure \ref{fig:ML3}) are very similar, in fact almost identical, to those of model I above.
+		For the second model, thickness likelihoods were defined in a way that reinforced the probability of a very thick seal but at the same time restricted the reservoir formation to a likely significantly thinner thickness of around 50 m (see Table \ref{tab:ML3_likelihoods}).
+The results (see Figure \ref{fig:ML3} in the Appendix) are very similar, in fact almost identical, to those of model I above.
 This can be explained by the influence of leak and spill point.
@@
 The thickness of the reservoir only becomes significant, if it is thin enough to locate the respective layer bottom above the leak and spill point depth in the trap section.
-Ultimately, the spill point appears to be the primariy limiting factor to maximum trap fill.
+Ultimately, the spill point appears to be the primary limiting factor to maximum trap fill.
 It can furthermore be derived from this that the maximum trap volume potential, as defined by the given structural features, is already represented in the ROV probability distribution of the prior model.
@@
 			\end{tabular}
-			\caption{Layer thickness likelihoods determined by normal distributions as used for posterior model III.}
+			\caption{Layer thickness likelihoods determined by normal distributions as used for posterior model III (3D).}
 			\label{tab:ML2_likelihoods}
 		\end{table}
-		Using likelihoods which favor a thick reservoir, but only a thin seal of about 50 m thickness (see Table \ref{tab:ML2_likelihoods}), leakage due to juxtaposition with seal-overlying formations across the fault ("LEAK OVER") became the primary mechanism for trap volume control (see Figure \ref{fig:ML2} (c).
+		Using likelihoods which favor a thick reservoir but only a thin seal of about 50 m thickness (see Table \ref{tab:ML2_likelihoods}), leakage due to juxtaposition with seal-overlying formations across the fault ("LEAK OVER") became the primary mechanism for trap volume control (see Figure \ref{fig:ML2} (c).
 It can be recognized in Figure \ref{fig:ML2} (b) that, while uncertainty was reduced clearly in the footwall, it remained high in the hanging wall.
@@
 supporting the probability of thick layers in general.
-Additionally, a SSF-related likelihood function was implemented.
+Additionally, we implemented a SSF-related likelihood function.
 With a mean of 1.5 and a standard deviation of 0.5, it was chosen to mostly reinforce the likelihood of a SSF below the critical value of SSF$_\text{c} = 3$ (see Table \ref{tab:ML4_likelihoods}).\\
@@
 This value is furthermore characterized by a complex interdependence with other parameters and is directly related to layer thicknesses and fault offset.
-For this reason, information entropy and offset uncertainty in the hanging wall were not greatly reduced.
-However, the certainty about the ratio of displacement to layer thicknesses was presumably improved by using the SSF in a likelihood function.
+For this reason, Shannon entropy and offset uncertainty in the hanging wall were not reduced significantly more than without considering a SSF likelihood in posterior model I.
+However, we presume that the certainty about the ratio of displacement to layer thicknesses was improved by using the SSF in a likelihood function.
 
@@
 This is also indicated by a shift of mean and median to lower volumes (Figure \ref{fig:ML5} (d)).
-Accordingly, a higher degree of decision divergence, i.e.
-a greater separation of different Bayes actions, was induced (Figure \ref{fig:ML5} (e)).
-Expected losses of these minima were raised as well.\\
-		It is notable that the implementation of a likelihood function, which can be expected to significantly decrease the SSF related uncertainty, overall leads to an amplification of the duality in the ROV probability distribution and to dispersion of differently risk-affine decisions.
+Accordingly, a higher degree of decision divergence, i.e.\ a greater separation of different Bayes actions, was induced (Figure \ref{fig:ML5} (e)).
+Expected losses for these estimators were raised as well.\\
+		It is notable that the implementation of a likelihood function, which can be expected to significantly decrease the SSF-related uncertainty, overall leads to an amplification of the duality in the ROV probability distribution and to dispersion of differently risk-affine decisions.
 This effect is presumably caused by conservation of decisive uncertainty between "success" and complete failure of the trap, due to narrowing of the SSF probability around its critical value.

### 2017-12-11 Risk analysis/Text/Layout/Layout/chapter4_discussion.tex
FINAL (printed and handed in)

@@
 This field is characterized by the necessity to make decisions in the face of high risks and potentially high rewards.
-As these decisions are often closely linked to geological modeling and the estimation of reservoir-related values, we aimed to use this context to extend the Bayesian inference step by considering its influence on respective decision-making.
+As these decisions are often closely linked to geological modeling and the estimation of reservoir-related values, we aimed to use this context to extend the Bayesian inference step by considering its influence on respective decision making.
 To observe this, a case-specific loss function was developed in this work.
-In doing so, we intended to furthermore show, that the loss function estimation approach is suitable and useful to describe varying degrees of complexities behind decision-making environments and the risk-related behavior of the actors in it.
+In doing so, we intended to furthermore show that the loss function estimation approach is suitable and useful to describe varying degrees of complexities behind decision-making environments and the risk-related behavior of actors in it.
 	To analyze this, geological models were interpreted as potential hydrocarbon systems.
-A central part of our work was comprised of the development of algorithms capable of automatic structural trap recognition and subsequent calculation of values relevant to decision-making.
-Different prior and posterior probability distributions for such economic values were generated to form a basis for application of the custom loss function for value estimation.
-These methods were first applied to a conceptual 1D geological model and subsequently to a full 3D structural model.
+A central part of our work was comprised of the development of algorithms capable of automatic structural trap recognition and subsequent calculation of volumetric values relevant to decision making.
+Different prior and posterior probability distributions for such values were generated to form a basis for estimation via application of the custom loss function.
+We applied these methods first to a conceptual 1D geological model and subsequently to a full 3D structural model.
 	Despite a significant leap in complexity, the results from both models proved to be widely similar.
@@
 	The economic value realized in each model construction (score and ROV, respectively), had been defined to depend on model parameters, either directly or indirectly by taking into account their relations to other parameters or to specific thresholds and conditions.
-This meant in general, that in this step of valuation, model uncertainties were in a way translated and reduced to the probability distribution of a single decisive value.
-As this was set to be the foundation for loss function estimation, we consider this value distribution to be an expression of the state of knowledge on which the decision-making is to be based.
+This meant, in general, that in this step of valuation, model uncertainties were in a way translated and reduced to the probability distribution of a single decisive value.
+As this was set to be the foundation for loss function estimation, we consider this value distribution to be an expression of the state of knowledge on which the decision making is to be based.
 Thereby, we furthermore propose that the overall uncertainty inherent in this probability distribution can be referred to as "decision uncertainty" and that this entity is to be viewed separately from the model uncertainty.\\
-	By viewing decision-making as a problem of optimizing a case-specific loss function based on such a state of knowledge and decision uncertainty, we were able to observe clear differences in the respective behavior of distinctly risk-affine actors.
+	By viewing decision making as a problem of optimizing a case-specific loss function based on such a state of knowledge and decision uncertainty, we were able to observe clear differences in the respective behavior of distinctly risk-affine actors.
 	The position and separation of their Bayes actions, i.e.\ their decisions, manifested according to the properties of the value distributions.
@@
 A mean located in the middle between two widely separated modes, which might be of low intrinsic uncertainty themselves, would indicate a much higher overall decision uncertainty, than a mean found in the middle of a narrow unimodal distribution.
-Pronounced bimodality between two extremes, i.e.\ high uncertainty, resulted in a wider separation of Bayes estimators.
+Pronounced bimodality between two extremes, i.e.\ high overall uncertainty, resulted in a wider separation of Bayes estimators.
 Reduction of the distribution to one mode conversely led to the convergence of different Bayes actions.
@@
 The better these are, the more similar the decisions of differently risk-affine actors and the lower their loss expectations are.
-It can be assumed, that given perfect information, all actors would bid on the same estimate (the true value) and expect no loss, since no risk is present.
-It furthermore follows from this that the relevance of risk factors decreases with higher reduction of decision uncertainty.
+It can be assumed that given perfect information all actors would bid on the same estimate (the true value) and expect no loss, since no risk is present.
+It furthermore follows from this that the relevance of risk factors decreases with greater reduction of decision uncertainty.
 	\subsubsection{Impact of additional information on decision making}
 	Cases of various states of knowledge and decision uncertainties, i.e.\ different realizations of the value probability distribution, were attained by the implementation of Bayesian inference.
-While in most settings, adding information via inference led to a reduction of decision uncertainty, this was not always the case.
+While, in most settings, adding information via inference led to a reduction of decision uncertainty, this was not always the case.
 We observed that the impact on decision uncertainty, induced by Bayesian inference, is not necessarily strictly aligned with the change in uncertainty regarding model parameters and their combinations.
-In respect to this, two aspects seems to be of central importance: (1) "where" in the model uncertainty is reduced, i.e.\ in which spatial area or regarding which model parameters and (2) which possible outcome is enhanced in terms of probability.
+In respect to this, two aspects seems to be of central importance: (1) "where" in the model uncertainty is reduced, i.e.\ in which spatial area or regarding which model parameters, and (2) which possible outcome is enhanced in terms of probability.
 An increased probability of a thick or thin seal in our 3D model, for example, equally reduced decision uncertainty significantly, by raising the probability of a positive or negative outcome, respectively.
-Improved certainty about our reservoir thickness, however, had little to no impact on decision-making.
-This shows us that some areas and parameters of the model have a much greater influence on the decision uncertainty than others.\\
-	Adding to this, we observed a particular effect related to threshold values that lead to an abrupt cut-off between two extrema of decision ranges.
+Improved certainty about our reservoir thickness, however, had little to no impact on decision making.
+This shows that some areas and parameters of the model have a much greater influence on the decision uncertainty than others.\\
+	Adding to this, we observed a particular effect related to threshold values that led to an abrupt cut-off between two extrema of decision ranges.
 In both types of model construction, 1D and 3D, this was directly related to sealing reliability.
@@
 Resulting Bayes actions were characterized by an increase in divergence and expected losses, indicating a deterioration of the state of knowledge and an according increase in decision uncertainty.
-Thus, in some cases, adding information to the inference process might leave actors in greater disagreement than before.
-	However, we furthermore have to consider that actors weight possible outcomes of the probability distribution differently.
+Thus, in some cases, adding information to the inference process might leave actors in greater disagreement than before.\\
+	However, we furthermore have to consider that actors weight possible outcomes of the value distribution differently.
 They consequently are affected differently by the same type of additional information.
-Risk-friendly actors were the most robust in their decision-making in the face of possible trap failure.
+Risk-friendly actors were the most robust in their decision making in the face of possible trap failure.
 Eliminating this risk proved to be far less significant to the most risk-friendly, than for risk-averse actors.
-Accordingly, it should be of foremost importance for risk-averse actors, to reduce the uncertainty regarding threshold-related factors which might decide between the success and complete failure of a project.
+Accordingly, it should be of foremost importance for risk-averse actors to reduce the uncertainty regarding threshold-related factors which might decide between the success and complete failure of a project.
 This is less relevant for risk-friendly decisions makers, who respectively might acquire a comparable benefit from knowing more about the probability of positive outcomes, as they are less discouraged from overestimating and less afraid of failure.
@@
 In our loss function, this seems to be well represented by varying affinities to rather over- or underestimate.
-For our risk-averse actors, a major risk is posed by seal failure.
-Our risk-friendliest individual sees a greater risk in missing out on opportunity.
+For risk-averse actors, a major risk is posed by seal failure.
+The risk-friendliest individual sees a greater risk in missing out on opportunity.
 Nevertheless, it is suggested by the results from the different inference cases, that the best states of knowledge are attained by eliminating multimodality and narrowing the range of probabilities to one mode anywhere along the value axis.\\
@@
 It follows that, to approach an effective mitigation of high risks, the complexities behind decisive factors need to be assessed thoroughly and respective parent parameters, as well as their interdependencies, need to be identified.
-This might enable a better understanding of which type of information is missing and where in the model additional data might be of use for improved decision-making.\\
+This might enable a better understanding of which type of information is missing and where in the model additional data might be of use for improved decision making.\\
 	%The respective estimate resilient to risk was generally found close to the positive mode and experienced only minor lateral shifting, after diminishing the probability of failure.
@@
 Conversely, the Bayes action of the most-averse individual was influenced greatly by the probability of the "worst case" scenario, complete trap failure.
-Only a stark decrease of the respective mode encouraged this actor to bid on an estimate close to the risk-friendlier decision-makers.
+Only a stark decrease of the respective mode encouraged this actor to bid on an estimate close to the risk-friendlier decision makers.
 These observations match the principles and conditions conceived in the design of the loss function and the chosen range of risk factors.
-	A central conclusion from these observations seems to be, that more of simply any type of information does not necessarily lead to better decisions.
-Instead, improved of decision-making is achieved by attaining the right kind of information that is able to shed light on uncertainties which are relevant to an individuals own goals and preferences, as well as the general problem at hand.
+	A central conclusion from these observations seems to be that more of simply any type of information does not necessarily lead to better decisions.
+Instead, improved decision making is achieved by attaining the right kind of information that is able to shed light on uncertainties which are relevant to an individuals own goals and preferences, as well as the general problem at hand.
 These findings pair well with statements made by \citet{bratvold2010making}, who emphasized that the purpose of models in the hydrocarbon sector is to attain insights to improve decisions.
@@
 		\includegraphics[width=1\textwidth]{Figures/scenarios_options}
-		\caption{ROV distribution of the prior model subdivided by modes into scenarios which could pose decision options (marked by greens lines).
+		\caption{ROV distribution of the prior model subdivided by modes into scenarios which could pose decision options (boundaries marked by greens lines).
 Red lines indicate a discretization of decision options according to fixed scales of resource allocations for a project.
@@
 However, most often, low probabilities for the values between these two opposite options remain.
-Especially considering the use of loss functions which lead to best estimates located in these areas of low probability, it is debatable, how these ranges should be treated.
+Especially considering the use of loss functions which lead to estimators located in these areas of low probability, it is debatable, how these ranges should be treated.
 A possibility could be to subdivide the value probability distribution accordingly into regions.
-Considering respective integral probabilities for each such section and by introducing a critical integral probability as a threshold, it could be determined whether such a subregion is to be considered an actual option for decision-making or not.
+Considering respective integral probabilities for each such section and by introducing a critical integral probability as a threshold, it could be determined whether such a subregion is to be considered an actual option for decision making or not.
 If a Bayes action is located outside of such a decision option, it could be automatically assigned to the closest region of sufficient integral probability.
@@
 	\subsubsection{Comparison to common practices and research in the hydrocarbon sector}		
-	In remains to be argued, to what extent our methodology and findings contribute to practical applications in hydrocarbon exploration and production.
-Monte Carlo simulation for reservoir estimation and risk assessment has become common in this sector and is often used in combination with decision trees to make decisions (see \citet{murtha1997monte}, \citet{mudford2000valuing}, \citet{wim2001guidelines} and \citet{bratvold2010making}).
+	It remains to be argued to what extent our methodology and findings contribute to practical applications in hydrocarbon exploration and production.
+Monte Carlo simulation for reservoir estimation and risk assessment has become common in this sector and is often used in combination with decision trees (see \citet{murtha1997monte}, \citet{mudford2000valuing}, \citet{wim2001guidelines} and \citet{bratvold2010making}).
 However, it seems to us that distributions resulting from probabilistic modeling are mostly only considered to attain best estimates in the form of means.
@@
 Thus, we intend to provide a new perspective with our methodology.
-\citet{murtha1997monte} emphasized that Monte Carlo simulation does not make decisions, it merely prepares for it.
+\citet{murtha1997monte} emphasized that Monte Carlo simulation does not make decisions.
+It merely prepares for it.
 We believe that loss functions have the potential to go one step further.
@@
 While this is obviously unrealistic, we presume that an elaborate loss function might at least provide a very good preliminary decision recommendation.
-It might furthermore be able to weight risks that are not immediately apparent to an individual.
+It might furthermore be able to weight risks that are not immediately apparent to an individual as a person.
 Furthermore, the influence of human biases and psychological behavioral challenges, as described by \citet{bratvold2010making}, could be mitigated.\\
@@
 However, their research focused on history-matching simulations for already producing fields.
-Our approach of applying Bayesian inference for structural geological modeling and volumetric reservoir calculations is intended to support decision-making in the earliest stages of a reservoir, when it has to be decided whether a project should be developed or not.
+Our approach of applying Bayesian inference for structural geological modeling and volumetric reservoir calculations is intended to support decision making in the earliest stages of a reservoir, when it has to be decided whether a project should be developed or not.
 Nevertheless, it was shown in the research conducted by \citet{wadsley2005markov} that early volumetric OOIP estimates can be combined with later calculations from production data via MCMC methods.
@@
 Furthermore, it follows that no effective uncertainty concerning the overall structural shape was implemented, particularly regarding anticlinal features and the position of the spill point in relation to the trap top.
-Due to this, trap volumes tended to occur primarily in distinct "filled-to-spill" scenarios.
+Due to this, trap volumes tended to occur primarily in distinct scenarios.
 While being unrealistic, this enabled the recognition of pronounced influences related to the various control mechanisms.
@@
 	For a conceptual application, the customization of the loss functions was kept relatively simple.
-We argue that despite its simplicity, the design of the function and the incorporation of a risk factor was suitable to achieve a clear distinction between differently risk-affine actors and respective behavior.
+We argue that, despite its simplicity, the design of the function and the incorporation of a risk factor was suitable to achieve a clear distinction between differently risk-affine actors and respective behaviors.
 It would be possible to extend this design to attain very complex loss functions.
@@
 The consideration of more details, without a basis on real data, would have furthermore required extensive speculation, which would have presumably impaired the generalization potential of respective observations.
-We defined risk affinity to be dependent on arbitrarily chosen risk factors, which led to according re-weighting.
+We defined risk affinity to be dependent on arbitrarily chosen risk factors which led to according re-weighting.
 \citet{davidson2015} used risk parameters determined by the maximal loss each actor could incur.
 Other approaches could be based on more tangible values, for example by making risk attitude dependent on a fixed budget.
-While standard loss functions only return mean or median as estimates, loss function customization offers a flexible approach to express different objective and subjective weighting factors that are specified to certain decision-making environments and decision-makers' perspectives.
+While standard loss functions only return mean or median as estimates, loss function customization offers a flexible approach to express different objective and subjective weighting factors that are specified to particular decision-making environments and decision makers' perspectives.
 	\subsubsection{Possibilities for future research and extensions}	
 	Considering the findings of this work, there is still many points that could be expanded on in future research.
-It would be of interest to apply the same overall concept and methods to an authentic case based on real datasets.
+It would be of interest to apply the same overall concept and methodology to an authentic case based on real datasets.
 Given a realistic economic scenario, including capital and operational expenditures of a project, possibly a full net-present-value (NPV) analysis could be conducted.
-Recoverable reserves could be replaced by the NPV to evaluate modeling results and serve as a base for decision-making.
-A more elaborate respective loss function could be customized on the base of surveys, acquiring the specific preferences of one or several companies and thus attaining a better profile of the economic environment, as well as the individuals acting in it.
-Value of information could be considered as a parameter to better quantify the impact of Bayesian inference as observed in this work.\\
+Recoverable volumes could be replaced by the NPV to evaluate modeling results and serve as a base for decision making.
+A more elaborate loss function could be customized on the base of surveys, acquiring the specific preferences of one or several companies and thus attaining a better profile of the economic environment, as well as the individuals acting in it.
+Value of information (see \citet{eidsvik2015voi}) could be considered as a parameter to better quantify the impact of Bayesian inference as observed in this work.\\
 	Although the algorithms for automatic 3D hydrocarbon trap recognition were developed to fit primarily the model in this work, we presume that they could easily be adapted to other structural cases or even engineered further to be universally applicable.\\
@@
 Otherwise, non-structural reservoir parameters could be included as uncertain values, such as porosities and permeabilities of different layers.
-This might be of particular interest considering factors which are part of the ROV equation or other aspects that are related to high risks in decision-making.\\
-	We chose hydrocarbon systems and petroleum exploration as a sector for an exemplary application in this work, but other settings can be found, in which geological modeling is of central significance for decision-making.
+This might be of particular interest considering factors which are part of the ROV equation or other aspects that are related to high risks in decision making.\\
+	We chose hydrocarbon systems and petroleum exploration as a sector for an exemplary application, but other settings can be found in which geological modeling is of central significance for decision making.
 One example would be subsurface storage of fluids in a reservoir, such as carbon capture and storage (CCS).
-Questions regaring storage capacity and safety deal with similar conditions and geological problems as the ones presented this work, most importantly seal reliability and the risk of leakage.
+Questions regarding storage capacity and safety deal with similar conditions and geological problems as the ones presented this work, most importantly seal reliability and the risk of leakage.
 In respect to this, the models and approaches in this work might provide a basis to raise new and advanced explorable cases.

### 2018-03-07 Risk analysis/Poster/Poster STC 2017/examples/rwth/poster.tex
Created 3D uncertainty model with structural analysis.
TODO:
- rework structural analysis algorithms so that they work with current version of GemPy (topology module seems to be a problem sitll)
- work on a potser presentation of the thesis

\documentclass[final,hyperref={pdfpagelabels=false}]{beamer}
\usepackage{grffile}
\mode<presentation>{\usetheme{RWTH}}
\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage{amsmath,amsthm, amssymb, latexsym}
%\usepackage{times}\usefonttheme{professionalfonts}  % obsolete
%\usefonttheme[onlymath]{serif}
\boldmath
\usepackage[orientation=portrait,size=a0,scale=1.4,debug]{beamerposter}
% change list indention level
% \setdefaultleftmargin{3em}{}{}{}{}{}

%\usepackage{snapshot} % will write a .dep file with all dependencies, allows for easy bundling

\usepackage{array,booktabs,tabularx}
\newcolumntype{Z}{>{\centering\arraybackslash}X} % centered tabularx columns
\newcommand{\pphantom}{\textcolor{ta3aluminium}} % phantom introduces a vertical space in p formatted table columns??!!

\listfiles

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\graphicspath{{figures/}}

\title{\huge CCS Summer School in Oslo}
\author{d \\ Marco van Veen}
\institute[RWTH Aachen University]{Computational Geoscience and Reservoir Engineering, RWTH Aachen University, Aachen, Germany}
\date[Sep. 8th, 2009]{Sep. 8th, 2009}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newlength{\columnheight}
\setlength{\columnheight}{105cm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\begin{frame}
  \begin{columns}
    % ---------------------------------------------------------%
    % Set up a column 
    \begin{column}{.49\textwidth}
      \begin{beamercolorbox}[center,wd=\textwidth]{postercolumn}
        \begin{minipage}[T]{.95\textwidth}  % tweaks the width, makes a new \textwidth
          \parbox[t][\columnheight]{\textwidth}{ % must be some better way to set the the height, width and textwidth simultaneously
            % Since all columns are the same length, it is all nice and tidy.  You have to get the height empirically
            % ---------------------------------------------------------%
            % fill each column with content            
            \begin{block}{Introduction}
              \begin{itemize}
              \item Most face recognition approaches are sensitive to registration errors
                \begin{itemize}
                \item rely on a very good initial alignment and illumination
                \end{itemize}
              \item We propose/analyze:
                \begin{itemize}
                \item grid-based and dense extraction of local features
                \item block-based matching accounting for different\\
	                  viewpoints and registration errors
                \end{itemize}
              \end{itemize}              
            \end{block}
            \vfill
            \begin{block}{Feature Extraction}
              \begin{columns}
                \begin{column}{.55\textwidth}
                  \begin{itemize}
                  \item Interest point based feature extraction
                    \begin{itemize}
                    \item SIFT or SURF interest point detector
                    \item leads to a \alert{very sparse} description
                    \end{itemize}
                  \item Grid-based feature extraction
                    \begin{itemize}
                    \item overlaid regular grid
                    \item leads to a \alert{dense} description
                    \end{itemize}
                  \end{itemize}
                \end{column}
                \begin{column}{.44\textwidth}
                  \centering
                  \begin{tabularx}{\linewidth}{ZZZ}
                     Orig.
                     &
                     IP
                     &
                     Grid
                     \\
                    \includegraphics[width=0.95\linewidth]{images/viola/m-012-1.png}
                    &
                    \includegraphics[width=0.95\linewidth]{images/viola/surf-64x64-ip/m-012-1.pngsurf.png}
                    &
                    \includegraphics[width=0.95\linewidth]{images/viola/surf-64x64-4/m-012-1.pngsurf.png}
                    \\
                    \includegraphics[width=0.95\linewidth]{images/viola/m-012-2.png}
                    &
                    \includegraphics[width=0.95\linewidth]{images/viola/surf-64x64-ip/m-012-2.pngsurf.png}
                    &
                    \includegraphics[width=0.95\linewidth]{images/viola/surf-64x64-4/m-012-2.pngsurf.png}
                    \\
                  \end{tabularx}
                \end{column}
              \end{columns}
              \vskip-1ex
            \end{block}
            \vfill
            \begin{block}{Feature Description}
              \begin{itemize}
              \item Scale Invariant Feature Transform (SIFT)
                \begin{itemize}
                \item 128-dimensional descriptor, histogram of gradients, scale invariant
                \end{itemize}
              \item Speeded Up Robust Features (SURF)
                \begin{itemize}
                \item 64-dimensional descriptor, histogram of gradients, scale invariant
                \end{itemize}
              \item face recognition: invariance w.r.t. rotation is often not necessary
                \begin{itemize}
                \item rotation dependent upright-versions U-SIFT, U-SURF-64, U-SURF-128
                \end{itemize}
              \end{itemize}
            \end{block}
            \vfill
            \begin{block}{Feature Matching}
              \begin{itemize}
              \item Recognition by Matching
                \begin{itemize}
                \item nearest neighbor matching strategy
                \item descriptor vectors extracted at keypoints in a test image $X$ are compared to all descriptor vectors extracted at keypoints from the reference images $Y_n, n=1,\cdots,N$ by the Euclidean distance
                \item decision rule:
                  \begin{align*}
                    X \rightarrow r(X) & = \arg \max_{c} \Big\{ \max_n \big\{ \sum_{x_i \in X} \delta(x_i,Y_{n,c})\big\} \Big\} 
                  \end{align*}
                \item additionally, a ratio constraint is applied in $\delta(x_i,Y_{n,c})$
                \end{itemize}
              \item Viewpoint Matching Constraints
                \begin{itemize}
                \item maximum matching: unconstrained
                \item grid-based matching: absolute box constraints
                \item grid-based best matching: absolute box constraints, overlapping
                \end{itemize}
              \item Postprocessing
                \begin{itemize}
                \item RANSAC-based outlier removal
                \item RANSAC-based system combination
                \end{itemize}
              \end{itemize}
            \end{block}
            \vfill
            \begin{block}{Matching Examples for the AR-Face and CMU-PIE Database}
              \begin{figure}
                \footnotesize
                \centering
                \begin{tabular}{p{.09\linewidth} | p{.12\linewidth} | p{.12\linewidth} | p{.12\linewidth} || p{.12\linewidth} | p{.12\linewidth} | p{.12\linewidth} | p{.09\linewidth} }
                  Feature
                  &
                  Maximum
                  & 
                  Grid
                  & 
                  Grid-Best
                  &
                  Maximum
                  & 
                  Grid
                  & 
                  Grid-Best
                  & 
                  Feature
                  \\
                  \hline
                  SIFT 
                  &
                  \includegraphics[width=1.0\linewidth]{paper/bmvc09-surf/figures/matchings/arface-sift/maximum_m-005-17.pgm--w-052-4}
                  &
                  \includegraphics[width=1.0\linewidth]{paper/bmvc09-surf/figures/matchings/arface-sift/grid_m-005-17.pgm--m-005-4}
                  &
                  \includegraphics[width=1.0\linewidth]{paper/bmvc09-surf/figures/matchings/arface-sift/grid-best_m-005-17.pgm--m-005-4}
                  &
                  \includegraphics[width=1.0\linewidth]{paper/bmvc09-surf/figures/matchings/cmupie-surf/maximum_07-27-22.pgm--04-27-08}
                  &
                  \includegraphics[width=1.0\linewidth]{paper/bmvc09-surf/figures/matchings/cmupie-surf/grid_07-27-22.pgm--07-27-08}
                  &
                  \includegraphics[width=1.0\linewidth]{paper/bmvc09-surf/figures/matchings/cmupie-surf/grid-best_07-27-22.pgm--07-27-08}
                  &
                  SURF
                  \\
                  \hline
                  U-SIFT
                  &
                  \includegraphics[width=1.0\linewidth]{paper/bmvc09-surf/figures/matchings/arface-usift/maximum_m-005-17.pgm--w-052-4}
                  &
                  \includegraphics[width=1.0\linewidth]{paper/bmvc09-surf/figures/matchings/arface-usift/grid_m-005-17.pgm--m-005-4}
                  &
                  \includegraphics[width=1.0\linewidth]{paper/bmvc09-surf/figures/matchings/arface-usift/grid-best_m-005-17.pgm--m-005-4}
                  &
                  \includegraphics[width=1.0\linewidth]{paper/bmvc09-surf/figures/matchings/cmupie-usurf/maximum_07-27-22.pgm--04-27-08}
                  &
                  \includegraphics[width=1.0\linewidth]{paper/bmvc09-surf/figures/matchings/cmupie-usurf/grid_07-27-22.pgm--07-27-08}
                  &
                  \includegraphics[width=1.0\linewidth]{paper/bmvc09-surf/figures/matchings/cmupie-usurf/grid-best_07-27-22.pgm--07-27-08}
                  &
                  U-SURF
                  \\
                  \hline
                \end{tabular}
              \end{figure}
              \begin{itemize}
              \item Matching results for the AR-Face (left) and the CMU-PIE database (right)
                \begin{itemize}
                \item maximum matching show false classification examples
                \item grid matchings show correct classification examples
                \item upright descriptor versions reduce the number of false matches
                \end{itemize}
              \end{itemize}
            \end{block}
          }
        \end{minipage}
      \end{beamercolorbox}
    \end{column}
    % ---------------------------------------------------------%
    % end the column

    % ---------------------------------------------------------%
    % Set up a column 
    \begin{column}{.49\textwidth}
      \begin{beamercolorbox}[center,wd=\textwidth]{postercolumn}
        \begin{minipage}[T]{.95\textwidth} % tweaks the width, makes a new \textwidth
          \parbox[t][\columnheight]{\textwidth}{ % must be some better way to set the the height, width and textwidth simultaneously
            % Since all columns are the same length, it is all nice and tidy.  You have to get the height empirically
            % ---------------------------------------------------------%
            % fill each column with content

            \begin{block}{Databases}
              \begin{columns}
                \begin{column}{.6\textwidth}
                  \begin{itemize}
                  \item AR-Face 
                    \begin{itemize}
                    \item variations in illumination
                    \item many different facial expressions
                    \end{itemize}
                  \item CMU-PIE
                    \begin{itemize}
                    \item variations in illumination (frontal images from the illumination subset)
                    \end{itemize}
                  \end{itemize}
                \end{column}
                \begin{column}{.39\textwidth}
                  \includegraphics[width=0.22\linewidth]{hanselmann-databases/arface/train/png/occlusions/arneutral/m-016-1}
                  \-
                  \includegraphics[width=0.22\linewidth]{hanselmann-databases/arface/train/png/cropped/m-016-4}
                  \-
                  \includegraphics[width=0.22\linewidth]{hanselmann-databases/arface/test/png/occlusions/ar1sun/m-016-8}
                  \-
                  \includegraphics[width=0.22\linewidth]{hanselmann-databases/arface/test/png/occlusions/ar1scarf/m-016-11}
                  \vskip3ex

                  \includegraphics[width=0.22\linewidth]{hanselmann-databases/cmupie/test/png/cropped/00-27-02}
                  \-
                  \includegraphics[width=0.22\linewidth]{hanselmann-databases/cmupie/test/png/cropped/00-27-17}
                  \-
                  \includegraphics[width=0.22\linewidth]{hanselmann-databases/cmupie/test/png/cropped/00-27-18}
                  \-
                  \includegraphics[width=0.22\linewidth]{hanselmann-databases/cmupie/test/png/cropped/00-27-20}
                \end{column}
              \end{columns}
            \end{block}
            \vfill
            \begin{block}{Results: Manually Aligned Faces}
              \begin{itemize}
              \item AR-Face: 110 classes, 770 train, 770 test
              \end{itemize}
              \vskip-0.5ex
              \begin{table}
                \centering
                \small
                \begin{tabular}{@{} p{.2\linewidth} p{.18\linewidth} p{.25\linewidth} r r r @{}}
                  \toprule 
                  Descriptor &  Extraction        & \multicolumn{1}{p{.2\linewidth}}{\# Features}           & \multicolumn{3}{c @{}}{Error Rates [\%]}         \\
                  \cmidrule(l){4-6}
                             &                    &                                                      & \multicolumn{1}{c}{Maximum} & \multicolumn{1}{c}{Grid} & \multicolumn{1}{c @{}}{Grid-Best} \\
                  \cmidrule(r){1-1}  \cmidrule(lr){2-2}   \cmidrule(lr){3-3}                               \cmidrule(lr){4-4}            \cmidrule(lr){5-5}         \cmidrule(l){6-6}  
                  SURF-64    & IPs                & \pphantom{1}64 $\times$ 5.6 (avg.)                    & 80.64     &  84.15  & 84.15     \\
                  SIFT       & IPs                & 128 $\times$ 633.78 (avg.)                           & 1.03      &  95.84  & 95.84      \\
                  \addlinespace
                  SURF-64    & 64x64-2 grid       & \pphantom{1}64 $\times$ 1024                          & 0.90      &  0.51   & 0.90      \\  
                  SURF-128   & 64x64-2 grid       & 128 $\times$ 1024                                    & 0.90      &  0.51   & 0.38       \\
                  SIFT       & 64x64-2 grid       & 128 $\times$ 1024                                    & 11.03     &  0.90   & 0.64       \\
                  \addlinespace
                  U-SURF-64  & 64x64-2 grid       & \pphantom{1}64 $\times$ 1024                          & 0.90      &  1.03   & 0.64      \\ 
                  U-SURF-128 & 64x64-2 grid       & 128 $\times$ 1024                                    & 1.55      &  1.29   & 1.03       \\ 
                  U-SIFT     & 64x64-2 grid       & 128 $\times$ 1024                                    & \textbf{0.25}  &  \textbf{0.25}   & \textbf{0.25}  \\
                  % \midrule
                  % \multicolumn{3}{l}{Modular PCA \cite{TanChen:subpatternPCA:Neurocomputing2005}}                         & 14.14 \\
                  % \multicolumn{3}{l}{Adaptively weighted Sub-Pattern PCA \cite{TanChen:subpatternPCA:Neurocomputing2005}} & 6.43  \\
                  % \multicolumn{3}{l}{DCT \cite{ekenel:LocalapFaceRecog:cvprbw2006}}                                       & 4.70  \\
                  \bottomrule
                \end{tabular}
              \end{table}

              \vskip1ex
              \begin{itemize}
              \item CMU-PIE: 68 classes, 68 train (``one-shot'' training), 1360 test
              \end{itemize}
              \vskip-0.5ex
              \begin{table}
                \centering
                \small
                \begin{tabular}{@{} p{.2\linewidth} p{.18\linewidth} p{.25\linewidth} r r r @{}}
                  \toprule 
                  Descriptor              &  Extraction    & \# Features                            & \multicolumn{3}{c @{}}{Error Rates [\%]}      \\
                  \cmidrule(l){4-6}
                                          &                    &                                    & \multicolumn{1}{c}{Maximum} & \multicolumn{1}{c}{Grid} & \multicolumn{1}{c @{}}{Grid-Best} \\
                  \cmidrule(r){1-1}  \cmidrule(lr){2-2}   \cmidrule(lr){3-3}                          \cmidrule(lr){4-4}            \cmidrule(lr){5-5}         \cmidrule(l){6-6}  
                  SURF-64    & IPs           & \pphantom{1}64 $\times$ 6.80 (avg.)                  & 93.95         & 95.21             & 95.21     \\  
                  SIFT       & IPs           & 128 $\times$ 723.17 (avg.)                           & 43.47         & 99.33             & 99.33     \\  
                  \addlinespace
                  SURF-64    & 64x64-2 grid  & \pphantom{1}64  $\times$ 1024                        &   13.41       &  4.12             &  7.82     \\
                  SURF-128   & 64x64-2 grid  & 128 $\times$ 1024                                    &  12.45        &  3.68             & 3.24      \\  
                  SIFT       &  64x64-2 grid & 128 $\times$ 1024                                    & 27.92         &  7.00             & 9.80      \\  
                  \addlinespace
                  U-SURF-64  & 64x64-2 grid  & \pphantom{1}64  $\times$ 1024                        & \textbf{3.83} &   \textbf{0.51}   & \textbf{0.66}    \\  
                  U-SURF-128 & 64x64-2 grid  & 128 $\times$ 1024                                    & 5.67          &  0.95             & 0.88      \\  
                  U-SIFT     & 64x64-2 grid  & 128 $\times$ 1024                                    & 16.28         &  1.40             & 6.41      \\
                  % \midrule
                  % \midrule
                  % \multicolumn{3}{l}{Spherical Harmonics \cite{ZhangSamaras:SphericalHarmonics:pami2006}}                              & 1.80 \\
                  % \multicolumn{3}{l}{Modeling Phase Spectra using GMM \cite{MitraSavvidesBrockwell:ModelingPhaseSpectra:lncs2005}}    & 12.83 \\
                  \bottomrule
                \end{tabular}
              \end{table}              
            \end{block}
            \vfill
            \begin{block}{Results: Unaligned Faces}
              \begin{columns}
                \begin{column}{.59\textwidth}
                  \begin{itemize}
                  \item Automatically aligned by Viola \& Jones
                  \end{itemize}
                  \vskip-0.5ex
                  \begin{table}
                    \centering
                    \small
                    \begin{tabular}{@{} p{.4\linewidth} r r @{}}
                      \toprule 
                      Descriptor  &      \multicolumn{2}{c @{}}{Error Rates [\%]}      \\
                      \cmidrule(l){2-3}   
                      &   AR-Face       & CMU-PIE  \\
                      \cmidrule(r){1-1}  \cmidrule(lr){2-2}  \cmidrule(l){3-3}  
                      SURF-64     &   5.97          & 15.32    \\ 
                      SURF-128    &   5.71          & 11.42    \\ 
                      SIFT        &   5.45          & 8.32     \\  
                      \addlinespace
                      U-SURF-64   &   5.32          & 5.52     \\  
                      U-SURF-128  &   5.71          & \textbf{4.86}  \\ 
                      U-SIFT      &   \textbf{4.15} & 8.99     \\  
                      \bottomrule
                    \end{tabular}
                  \end{table}
                \end{column}
                \begin{column}{.39\textwidth}                
                  \vskip-3ex
                  \begin{itemize}
                  \item Manually aligned faces

                      \includegraphics[width=0.9\linewidth]{hanselmann-thesis/slides/figures/alignment_cmu}

                  \item Unaligned faces

                      \includegraphics[width=0.9\linewidth]{hanselmann-thesis/slides/figures/unalignment_cmu}

                  \end{itemize}
                \end{column}
              \end{columns}
            \end{block}
            \vfill
            \begin{block}{Results: Partially Occluded Faces}
              \begin{itemize}
              \item AR-Face: 110 classes, 110 train (``one-shot'' training), 550 test
              \end{itemize}
              \vskip-0.5ex
              \begin{table}
                \small
                \centering
                \begin{tabular}{@{} l @{} r r r r r r@{}}
                  \toprule 
                  Descriptor          & \multicolumn{6}{c @{}}{Error Rates [\%]}                                                                 \\
                  \cmidrule(l){2-7}
                  & \textit{AR1scarf}  & \textit{AR1sun}  & \textit{ARneutral} & \textit{AR2scarf} & \textit{AR2sun}  & Avg. \\
                  \cmidrule(r){1-1}     \cmidrule(lr){2-2}   \cmidrule(lr){3-3} \cmidrule(lr){4-4}   \cmidrule(lr){5-5}  \cmidrule(lr){6-6}  \cmidrule(l){7-7} 
                  SURF-64             & 2.72             & 30.00          & 0.00                & 4.54            & 47.27      & 16.90 \\        
                  SURF-128            & 1.81             & 23.63          & 0.00                & 3.63            & 40.90      & 13.99 \\
                  SIFT                & 1.81             & 24.54          & 0.00                & 2.72            & 44.54      & 14.72 \\
                  \addlinespace
                  U-SURF-64           & 4.54             & 23.63          & 0.00                & 4.54            & 47.27      & 15.99 \\
                  U-SURF-128          & 1.81             & \textbf{20.00} & 0.00                & 3.63            & 41.81      & 13.45 \\
                  U-SIFT              & \textbf{1.81}    & 20.90          & \textbf{0.00}       & \textbf{1.81}   & \textbf{38.18} & \textbf{12.54} \\
                  \cmidrule(r){1-1}     \cmidrule(lr){2-2}   \cmidrule(lr){3-3} \cmidrule(lr){4-4}   \cmidrule(lr){5-5}  \cmidrule(lr){6-6}  \cmidrule(l){7-7}
                  U-SURF-128+R        & 1.81             & 19.09          & 0.00                & 3.63             & 43.63     & 13.63 \\
                  U-SIFT+R            & 2.72             & \textbf{14.54} & 0.00                & \textbf{0.90}    & 35.45     & 10.72 \\
                  U-SURF-128+U-SIFT+R &  \textbf{0.90}   & 16.36          & \textbf{0.00}       & 2.72             & \textbf{32.72} & \textbf{10.54} \\       
                  % \midrule 
                  % \midrule 
                  % DCT \cite{ekenel:facialocclusion:icb2009}, baseline%
                  % & 8.2              & 61.8           & 7.3                & 16.4            & 62.7       & 31.28 \\
                  % DCT \cite{ekenel:facialocclusion:icb2009}, realigned%
                  % & 2.7              & 1.8            & 0.0                & 6.4             & 4.5        & 3.08  \\
                  \bottomrule
                \end{tabular}
              \end{table}
            \end{block}
            \vfill
            \begin{block}{Conclusions}
              \begin{itemize}
              \item Grid-based local feature extraction instead of interest points
              \item Local descriptors:
                \begin{itemize}
                \item upright descriptor versions achieved better results
                \item SURF-128 better than SURF-64
                \end{itemize}
              \item System robustness: manually aligned/unaligned/partially occluded faces
                \begin{itemize}
                \item SURF more robust to illumination
                \item SIFT more robust to changes in viewing conditions
                \end{itemize}
              \item RANSAC-based system combination and outlier removal
              \end{itemize}
            \end{block}
          }
          % ---------------------------------------------------------%
          % end the column
        \end{minipage}
      \end{beamercolorbox}
    \end{column}
    % ---------------------------------------------------------%
    % end the column
  \end{columns}
  \vskip1ex
  %\tiny\hfill\textcolor{ta2gray}{Created with \LaTeX \texttt{beamerposter}  \url{http://www-i6.informatik.rwth-aachen.de/~dreuw/latexbeamerposter.php}}
  \tiny\hfill{Created with \LaTeX \texttt{beamerposter}  \url{http://www-i6.informatik.rwth-aachen.de/~dreuw/latexbeamerposter.php} \hskip1em}
\end{frame}
\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Local Variables: 
%%% mode: latex
%%% TeX-PDF-mode: t
%%% End:

### 2018-03-07 Risk analysis/Poster/Poster STC 2017/poster.tex
Created 3D uncertainty model with structural analysis.
TODO:
- rework structural analysis algorithms so that they work with current version of GemPy (topology module seems to be a problem sitll)
- work on a potser presentation of the thesis

\documentclass{beamer}
\usepackage[orientation=portrait,size=a0,scale=1.4,debug]{beamerposter}
\mode<presentation>{\usetheme{RWTH2}}
\usepackage{chemformula}
\usepackage[utf8]{inputenc}
\usepackage[german, english]{babel} % required for rendering German special characters
\usepackage{siunitx} %pretty measurement unit rendering
\usepackage{hyperref} %enable hyperlink for urls
\usepackage{ragged2e}
\usepackage[font=scriptsize,justification=justified]{caption}
\usepackage{array,booktabs,tabularx}
\usepackage[autostyle=true,german=quotes,english=american]{csquotes}

\usepackage{amsmath, mhchem} % Unterstützung für mathematische und chemische Formeln

\usepackage{floatflt}

\usepackage[authoryear]{natbib}
\bibliographystyle{apalike}

\newcolumntype{Z}{>{\centering\arraybackslash}X} % centered tabularx columns
\sisetup{per=frac,fraction=sfrac}

\title{\ Summer School \enquote{CCS~in~the~High~North}}
%\author{Marco van Veen$^{1}$}
%\institute[RWTH Aachen University]{$^{1}$Institute of Computational Geoscience and Reservoir Engineering, RWTH Aachen University, Germany}
\author{Marco van Veen$^{1}$, Andrew Steadman$^{2}$, Valentin Zuchuat$^{3}$}
\institute[RWTH Aachen University]{$^{1}$Institute of Computational Geoscience and Reservoir Engineering, RWTH Aachen University, Germany\\
$^{2}$School of Earth Sciences, University of Bristol, United Kingdom\\
$^{3}$Department of Geosciences, University of Oslo, Norway
}
\date{\today}

% edit this depending on how tall your header is. We should make this scaling automatic :-/
\newlength{\columnheight}
\setlength{\columnheight}{104cm}

\begin{document}
\begin{frame}

\begin{columns}

	\begin{column}{.382\textwidth}
		\begin{beamercolorbox}[center]{postercolumn}
			\begin{minipage}{.98\textwidth}  % tweaks the width, makes a new \textwidth
				\parbox[t][\columnheight]{\textwidth}{ % must be some better way to set the the height, width and textwidth simultaneously

\begin{myblock}{Introduction}

\textbf{Carbon Capture and Storage (CCS) is one of the major mitigation measures to handle climate change. Even though international organisations agreed on the importance of CCS and the effective Paris Agreement refers to the widely accepted climate scenarios, the technology is not implemented on a large scale beyond pilot projects yet, nor are economic incentives given to foster its deployment. CCS is not a cutting-edge technology, the applied methods are well-known – not only in the oil and gas industry. Buzz words like enhanced oil recovery (EOR) and carbon capture, utilisation and storage (CCUS) stir up hope to realise large-scale projects soon and in an economically viable way.}
						\vspace{0.5em}
						\begin{figure}
							\begin{minipage}{0.95\textwidth}
								\centering\includegraphics[width=1\textwidth]{figures/photos/group.jpg}
								\caption{Students of 2016 at the edge of the world's northernmost swimming pool in the abandoned Russian mining town of Pyramiden (Photo: V. Zuchuat).}
								\label{fig:group}
							\end{minipage}
						\end{figure}

The summer school \enquote{CCS in the High North} is a collaboration between the \textbf{University of Oslo} (UiO), the \textbf{Colorado School of Mines} (CSM) in Golden and the \textbf{University Centre in Svalbard} (UNIS). It offers a suite of lectures, exercises and geological field work on MSc and PhD level, which are credited with \textbf{13 ECTS}. A highly motivated group of international students with diverse backgrounds and specialisations were not only taught, but also engaged themselves in inspiring discussions.

%Unfortunately, in 2017 the course in Golden, Colorado had to be cancelled due to high political uncertainties and visa restrictions for international students.
\end{myblock}\vfill

\begin{myblock}{Course Content}

	\includegraphics[scale=0.1]{figures/icons/eco_law}~\textbf{Economy, Policies \& Risk}
	\begin{itemize}
		\item The need for CCS: Physics, history and challenges of global climate change
		\item Economics and regulations: Incentives, carbon trading, \ce{CO2} utilisation
		\item Risk Assessment: Monitoring, simulation and liability
	\end{itemize}
	\vspace{0.5em}

	\includegraphics[scale=0.1]{figures/icons/geo}~\textbf{Geology \& Reservoir}
	\begin{itemize}
		\item Trapping mechanisms \& storage options: Static vs. dynamic storage
		\item Geological reservoir characterisation: Field observations and core logging
		\item Seal properties: Mechanical and chemical compaction, global data collections and experimental studies
		\item Rock mechanics: Fractures, secondary permeability
		\item Risk of fault and fracture leakage: Sub-seismic resolution, juxtaposition, smearing
	\end{itemize}
	\vspace{0.5em}

	\includegraphics[scale=0.1]{figures/icons/tech}~\textbf{Technology}
	\begin{itemize}
		\item Physical and chemical properties of \ce{CO2}
		\item Separation technologies: Chemical absorption, phase change, polymeric membranes
		\item Enhanced Oil Recovery: Relative permeability \& mobility ratio
		\item Sequestration modelling: Rock \& fluid properties, multi-phase flow, initialisation, events/wells
	\end{itemize}

\end{myblock}

		}\end{minipage}\end{beamercolorbox}
	\end{column}

	\begin{column}{.618\textwidth}
		\begin{beamercolorbox}[center]{postercolumn}
			\begin{minipage}{.98\textwidth} % tweaks the width, makes a new \textwidth
				\parbox[t][\columnheight]{\textwidth}{ % must be some better way to set the the height, width and textwidth simultaneously

\begin{myblock}{Fieldwork \& The Longyearbyen \ce{CO2} Lab}
Highlight of the Summer School were three days of fieldwork on the Arctic archipelago of Svalbard. The aim of the field investigations was to identify potential reservoir formations, evaluate and compare them on their ability to safely store \ce{CO2}. Beside two promising \ce{CO2} storage formations (see figures \ref{fig:fieldwork_adolfbukta} and \ref{fig:fieldwork_skansbukta}), also the targeted units of the Longyearbyen \ce{CO2} Lab were logged in the field close to Deltanesset. Tight sandstones with quartz-overgrowth and pore-filling clay minerals can be observed. Mudstone layers, a thrust décollement and lateral permafrost provide sufficient sealing (compare figure \ref{fig:cross_section}). Injection tests show secondary permeability caused by fractures and the sealing ability is proven by high underpressure in the reservoir \citep{braathen_longyearbyen_2012}.
	\vspace{0.5em}
	\begin{figure}
		\begin{minipage}{0.95\textwidth}
			\begin{minipage}[t]{0.32\textwidth}
				\centering\includegraphics[width=1\textwidth]{figures/photos/arrival_adolfbukta}
				\caption{Going ashore in Adolfbukta: Survival suits, rifles, redundant communications and an emergency backpack are essential HSE equipment during fieldwork on the Arctic archipelago of Svalbard (Photo: M.\,Gutierrez).}
				\label{fig:arrival_adolfbukta}
			\end{minipage}
			\hspace{\fill}
			\begin{minipage}[t]{0.32\textwidth}
				\centering\includegraphics[width=1\textwidth]{figures/photos/fieldwork_adolfbukta}
				\caption{Exposed sandstones of the Carboniferous Ebbadalen Formation north of Adolfbukta. Seven distinctive coarsening-up sequences of thick beds alternating with silty layers are expected to provide high horizontal permeability (Photo: M.\,Gutierrez).}
				\label{fig:fieldwork_adolfbukta}
			\end{minipage}
			\hspace{\fill}
			\begin{minipage}[t]{0.32\textwidth}
				\centering\includegraphics[width=1\textwidth]{figures/photos/fieldwork_skansbukta}
				\caption{Permian limestone beds of the Gipshuken and Kapp Starostin Formations are found at Skansbukta. Bioturbation structures filled with evaporitic material could provide secondary porosity. The picture shows cross-bedding in upper layers (Photo: M.\,Gutierrez).}
				\label{fig:fieldwork_skansbukta}
			\end{minipage}	
		\end{minipage}
	\end{figure}
	\vspace{0.2em}

	\begin{figure}
		\begin{minipage}{.94\textwidth}
			\centering\includegraphics[width=\textwidth]{figures/cross_section}
			\caption{Cross section of the targeted storage formation (Upper Triassic–Middle Jurassic sandstones of the Kapp Toscana Group), the overlying seal (Jurassic mudstones of the Agardfjellet Formation) and the prominent thrust décollement which is providing an additional seal. The indicated well Dh4 in the west is part of the Longyearbyen \ce{CO2} Lab, the logging took place in the east where the layers are outcropping. Modified from \cite{braathen_longyearbyen_2012}.}
			\label{fig:cross_section}
		\end{minipage}
	\end{figure}
%	\vspace{0.4em}

\end{myblock}\vfill

\begin{myblock}{Limited Information \& Seal Uncertainty}

\vspace{0.3em}
\begin{figure}
	\begin{minipage}{0.95\textwidth}
		\begin{minipage}[t]{0.36\textwidth}
			\vspace{0pt}
			\includegraphics[width=\textwidth]{figures/compaction.pdf}
		\end{minipage}\hfill
		\begin{minipage}[t]{0.22\textwidth}
			\vspace{0pt}
			\caption{
			Comparison of porosity–depth trend studies and experimental results for shales and
argillaceous sediments. While for E\&P projects the seal integrity is proven by the abundance of hydrocarbons, for \ce{CO2} storage it remains uncertain. Controlling factors are clay mineral and pore fluid composition, grain sizes, pore pressure and the influence of mechanical \textbf{(MC)} and chemical compaction \textbf{(CC)}. Because core sampling of seal layers is expensive, experimental studies and global databases might be auxiliary. Still, extrapolation of experimental results has limited validity for natural anisotropy and the influence of chemical compaction \citep{mondol_experimental_2007}.
			} \label{fig:compaction}
		\end{minipage}\hfill
		\begin{minipage}[t]{0.38\textwidth}
			\vspace{0pt}
			\includegraphics[width=\textwidth]{figures/leakage.pdf}
			\caption{
			Different leakage scenarios for a \ce{CO2} storage project which are subject to uncertainty. To prevent a spill out of the storage formation \textbf{(1)}, the storage capacity must be estimated. A break of \ce{CO2} through the formation's roof \textbf{(2)} depends on seal properties and knowledge of fracture pressures (see figure \ref{fig:compaction}). Evaluation of leakage along \textbf{(3)} or through \textbf{(4)} a fault requires the analysis of juxtaposition and smearing. Sub-seismic faults represent an additional risk. Still, the highest risk of leakage exists in close proximity to wells (Figure based on lecture notes, lecture by K. Indrevær \& A. Braathen, 15.6.2017).
			} \label{fig:leakage}
		\end{minipage}
	\end{minipage}
\end{figure}

\end{myblock}\vfill

\begin{myblock}{Outlook \& Motivation}
		\begin{minipage}[t]{0.72\textwidth}
			\vspace{0pt}
			Marco studies \textbf{Georesources Management} in Aachen. During his various stays in Norway -- two semesters in Stavanger and summer courses in Oslo and Longyearbyen -- he became passionate about \textbf{risk} and \textbf{decision analysis}. Currently he is writing his master thesis in the field of\textbf{ geologic modelling and uncertainty quantification} in exploration. His focus will be on the application of a \textbf{value of information (VOI)} approach. A specific project to test assumptions and models would be beneficial. The high potential of future \textbf{\ce{CO2} storage projects} provides a suitable case for application. The mentioned methods are believed to contribute to urgently needed \textbf{cost reduction} and \textbf{risk management} of CCS projects.
		\end{minipage}\hfill
		\begin{minipage}[t]{0.23\textwidth}
			\vspace{0em}
			\includegraphics[width=1\textwidth]{figures/author}
			\vspace{0.5em}
		\end{minipage}

%			\vspace{0.5em}

\end{myblock}\vfill

					As georesources manager who majored in geoscience I am curious about the structure and dynamics of our earth, but I like to put “these rocks” into a bigger picture. I am not only eager to understand what is going on below our feet, but also to estimate its value to society or economy and to communicate gained knowledge and accompanying risks to decision makers – in order to enable them to make good decisions.

Currently I am writing my master thesis in the field of geologic modelling and uncertainty quantification in exploration. My focus will be on the application of a value of information (VOI) approach and using open source software to demonstrate light weight solutions for high-cost projects.

I am looking for a specific project or data to test the mentioned methods on, preferably in Carbon Capture and Storage (CCS), which I believe has a great potential mitigating the dramatic consequences of climate change.

\begin{myblock}{References}
	\footnotesize
	\bibliography{./references}
\end{myblock}

		}\end{minipage}\end{beamercolorbox}
	\end{column}
\end{columns}
\end{frame}
\end{document}

### 2018-03-09 Risk analysis/Poster/Poster STC 2017/poster.tex
Poster draft

@@
 
-\title{\ Summer School \enquote{CCS~in~the~High~North}}
+\title{Bayesian Decision Theory in Structural Geological Modeling – How
+Reducing Uncertainties Affects Reservoir Value Estimations}
 %\author{Marco van Veen$^{1}$}
 %\institute[RWTH Aachen University]{$^{1}$Institute of Computational Geoscience and Reservoir Engineering, RWTH Aachen University, Germany}
-\author{Marco van Veen$^{1}$, Andrew Steadman$^{2}$, Valentin Zuchuat$^{3}$}
-\institute[RWTH Aachen University]{$^{1}$Institute of Computational Geoscience and Reservoir Engineering, RWTH Aachen University, Germany\\
-$^{2}$School of Earth Sciences, University of Bristol, United Kingdom\\
-$^{3}$Department of Geosciences, University of Oslo, Norway
-}
+\author{Fabian Antonio Stamm$^{1}$, Miguel de la Varga$^{1}$, Alexander Schaaf$^{1}$, Florian Wellmann$^{1}$}
+\institute[RWTH Aachen University]{$^{1}$Institute of Computational Geoscience and Reservoir Engineering, RWTH Aachen University, Germany}
 \date{\today}
@@
 
-\textbf{Carbon Capture and Storage (CCS) is one of the major mitigation measures to handle climate change.
-Even though international organisations agreed on the importance of CCS and the effective Paris Agreement refers to the widely accepted climate scenarios, the technology is not implemented on a large scale beyond pilot projects yet, nor are economic incentives given to foster its deployment.
-CCS is not a cutting-edge technology, the applied methods are well-known – not only in the oil and gas industry.
-Buzz words like enhanced oil recovery (EOR) and carbon capture, utilisation and storage (CCUS) stir up hope to realise large-scale projects soon and in an economically viable way.}
+Structural geological modeling is of central importance for the assessment of uncertain hydrocarbon accumulations in potential reservoirs.
+Hydrocarbon exploration and production is a high-risk, high-reward sector in which good decision making is indispensable.
+Actors in this field are faced with numerous uncertainties that have to be considered.
+We examine respective decision making from a Bayesian perspective.
+%We examine respective decision making from a Bayesian perspective by considering two main approaches: (1) we treat geological modeling as a Bayesian inference problem, so that additional geological information can be incorporated as likelihood functions linked to prior parameters in a probabilistic framework; (2) we base the decision-making step on value estimation by optimizing a case-specific loss function.
+
+\end{myblock}
+
+\begin{myblock}{Methods}
+\textbf{Construction of a 3D structural geological model:}
+	\begin{itemize}
+	\item Synthetic model of a simple anticline-fault trap in a potential hydrocarbon system (see Figure \ref{fig:unc_lik})
+	\item Inclusion of uncertainties by assigning probability distributions to the positions of layer interfaces in depth
+	\end{itemize}							
 								\vspace{0.5em}
@@
 									\begin{minipage}{0.95\textwidth}
-								\centering\includegraphics[width=1\textwidth]{figures/photos/group.jpg}
-								\caption{Students of 2016 at the edge of the world's northernmost swimming pool in the abandoned Russian mining town of Pyramiden (Photo: V. Zuchuat).}
-								\label{fig:group}
+										\centering\includegraphics[width=1\textwidth]{figures/Uncertainties_Likelihoods}
+										\caption{The structural geological model illustrated as a 2D cross section (A) and a 3D voxel representation (B).
+The inclusion of $z$-positional uncertainties affecting layer depths and fault offset are depicted in (C) and (D).
+Thicknesses of the three middle layers are defined by the distances of boundary points (E) and are thus directly dependent on (C).}
+										\label{fig:unc_lik}
 									\end{minipage}
 								\end{figure}
+\textbf{Integration in a probabilistic modeling framework for Bayesian analysis (see Figure \ref{fig:model_network}):}
+	\begin{itemize}
+	\item Setting up a full probability model taking into account all parameter probability distributions
+	\item Bayesian updating: Conditioning of parameters on additionally observed data via likelihood functions
+	\item Approximation of posterior distributions through the use of MCMC sampling
+	\end{itemize}
+\textbf{Evaluation of modeling results:}
+	\begin{itemize}
+	\item Shannon entropy for uncertainty visualization (after \citet{wellmann2012uncertainties}.
+	\item Implementation of algorithms for structural analysis, trap recognition (see Figure \ref{fig:trap_cond}) and calculation of recoverable oil volumes (ROV)
+	\item Decision making based on the optimization of a case-specific custom loss function that considers differently risk-affine actors
+	\end{itemize}
 
-The summer school \enquote{CCS in the High North} is a collaboration between the \textbf{University of Oslo} (UiO), the \textbf{Colorado School of Mines} (CSM) in Golden and the \textbf{University Centre in Svalbard} (UNIS).
-It offers a suite of lectures, exercises and geological field work on MSc and PhD level, which are credited with \textbf{13 ECTS}.
-A highly motivated group of international students with diverse backgrounds and specialisations were not only taught, but also engaged themselves in inspiring discussions.
+	%We examine respective decision making from a Bayesian perspective by considering two main
+	%approaches: (1) we treat geological modeling as a Bayesian inference problem, so that additional geological
+	%information can be incorporated as likelihood functions linked to prior parameters in a probabilistic framework; (2)
+	%we base the decision-making step on value estimation by optimizing a case-specific loss function. This function is
+	%customized to reflect the decision making of differently risk-affine actors relative to previously computed reservoir
+	%value probability distributions.
+	%We apply our approach to synthetic geological models which are constructed to represent potential hydro-
+	%carbon systems. Markov chain Monte Carlo sampling is used to approximate posterior models of reduced
+	%uncertainty.
 
-%Unfortunately, in 2017 the course in Golden, Colorado had to be cancelled due to high political uncertainties and visa restrictions for international students.
-\end{myblock}\vfill
+	%For the valuation of prior and posterior models, we also develop algorithms for automatic trap
+	%	recognition and volume calculation.
 
-\begin{myblock}{Course Content}
-
-	\includegraphics[scale=0.1]{figures/icons/eco_law}~\textbf{Economy, Policies \& Risk}
-	\begin{itemize}
-		\item The need for CCS: Physics, history and challenges of global climate change
-		\item Economics and regulations: Incentives, carbon trading, \ce{CO2} utilisation
-		\item Risk Assessment: Monitoring, simulation and liability
-	\end{itemize}
 							\vspace{0.5em}
-
-	\includegraphics[scale=0.1]{figures/icons/geo}~\textbf{Geology \& Reservoir}
-	\begin{itemize}
-		\item Trapping mechanisms \& storage options: Static vs. dynamic storage
-		\item Geological reservoir characterisation: Field observations and core logging
-		\item Seal properties: Mechanical and chemical compaction, global data collections and experimental studies
-		\item Rock mechanics: Fractures, secondary permeability
-		\item Risk of fault and fracture leakage: Sub-seismic resolution, juxtaposition, smearing
-	\end{itemize}
-	\vspace{0.5em}
-
-	\includegraphics[scale=0.1]{figures/icons/tech}~\textbf{Technology}
-	\begin{itemize}
-		\item Physical and chemical properties of \ce{CO2}
-		\item Separation technologies: Chemical absorption, phase change, polymeric membranes
-		\item Enhanced Oil Recovery: Relative permeability \& mobility ratio
-		\item Sequestration modelling: Rock \& fluid properties, multi-phase flow, initialisation, events/wells
-	\end{itemize}
+							\begin{figure}
+								\begin{minipage}{0.8\textwidth}
+									\centering\includegraphics[width=1\textwidth]{figures/Trap_Cond_H}
+									\caption{To be recognized as part of a trap, a reservoir voxel (1) has
+									to be positioned in the	footwall (2). Trap closure is defined by the seal shape and the normal fault (3). The maximum trap fill is defined by either the anticlinal spill point
+									(3;a) or a point of leakage across the fault, depending on juxtapositions with layers
+									underlying (3;b) or overlying the seal (3;c). The latter is only relevant if the
+									critical Shale Smear Factor (SSF$_c$) is exceeded, as determined over D and T in (4).}
+									\label{fig:trap_cond}
+								\end{minipage}
+							\end{figure}
 
@@
 
-\begin{myblock}{Fieldwork \& The Longyearbyen \ce{CO2} Lab}
-Highlight of the Summer School were three days of fieldwork on the Arctic archipelago of Svalbard.
-The aim of the field investigations was to identify potential reservoir formations, evaluate and compare them on their ability to safely store \ce{CO2}.
-Beside two promising \ce{CO2} storage formations (see figures \ref{fig:fieldwork_adolfbukta} and \ref{fig:fieldwork_skansbukta}), also the targeted units of the Longyearbyen \ce{CO2} Lab were logged in the field close to Deltanesset.
-Tight sandstones with quartz-overgrowth and pore-filling clay minerals can be observed.
-Mudstone layers, a thrust décollement and lateral permafrost provide sufficient sealing (compare figure \ref{fig:cross_section}).
-Injection tests show secondary permeability caused by fractures and the sealing ability is proven by high underpressure in the reservoir \citep{braathen_longyearbyen_2012}.
+\begin{myblock}{Results}
+		\begin{minipage}[t]{0.48\textwidth}
+					\vspace{0pt}
+					Posterior ROV probability distributions are realized depending on the nature of likelihoods implemented in the Bayesian inference step.
+Applying the custom loss function shows that the various Bayes estimators shift according to the characteristics of this underlying	value distribution.
+While bimodality and overall uncertainty lead to separation, risk-averse and risk-friendly decisions converge and decrease in expected loss given narrower unimodal distributions.
+A decisive factor in our model is the reliability of the sealing across the fault, as it defines the probability of complete trap failure.
+Two examples are summarized in Figures \ref{fig:ML4} and \ref{fig:ML5}.
+				\end{minipage}\hfill
+				\begin{minipage}[t]{0.48\textwidth}
+					\vspace{0em}
+					\begin{figure}
+					\centering
+					\includegraphics[width=1\textwidth]{figures/model_network}
+					\caption{Illustration of the probabilistic model as a hierarchical Bayesian network after \citet{koller2009probabilistic}. Stochastic nodes are represented by ellipses (blue for
+					priors, yellow for likelihoods). Deterministic functions are depicted as triangles in
+					red. Arrows indicate direct connections from parent to child nodes.}
+					\label{fig:model_network}
+					\end{figure}
+					\vspace{0.5em}
+				\end{minipage}
+
 	\vspace{0.5em}
 	\begin{figure}
-		\begin{minipage}{0.95\textwidth}
-			\begin{minipage}[t]{0.32\textwidth}
-				\centering\includegraphics[width=1\textwidth]{figures/photos/arrival_adolfbukta}
-				\caption{Going ashore in Adolfbukta: Survival suits, rifles, redundant communications and an emergency backpack are essential HSE equipment during fieldwork on the Arctic archipelago of Svalbard (Photo: M.\,Gutierrez).}
-				\label{fig:arrival_adolfbukta}
+		\begin{minipage}{0.98\textwidth}
+			\begin{minipage}[t]{0.5\textwidth}
+				\centering\includegraphics[width=1\textwidth]{figures/ML4}
+				\caption{Evaluation of a model conditioned on thickness and SSF likelihoods which assure a reliable trap closure and reinforce the probability of a high ROV.
+Shannon entropy was significantly reduced in the posterior model (b).
+Cross-fault leakage to overlying layers was eliminated as a trap control mechanism (c) and coincides with the disappearance of the probability of failure in the posterior ROV distribution (d).
+The respective loss functions are plotted in (e), showing high decision convergence.}
+				\label{fig:ML4}
 			\end{minipage}
 			\hspace{\fill}
-			\begin{minipage}[t]{0.32\textwidth}
-				\centering\includegraphics[width=1\textwidth]{figures/photos/fieldwork_adolfbukta}
-				\caption{Exposed sandstones of the Carboniferous Ebbadalen Formation north of Adolfbukta.
-Seven distinctive coarsening-up sequences of thick beds alternating with silty layers are expected to provide high horizontal permeability (Photo: M.\,Gutierrez).}
-				\label{fig:fieldwork_adolfbukta}
-			\end{minipage}
-			\hspace{\fill}
-			\begin{minipage}[t]{0.32\textwidth}
-				\centering\includegraphics[width=1\textwidth]{figures/photos/fieldwork_skansbukta}
-				\caption{Permian limestone beds of the Gipshuken and Kapp Starostin Formations are found at Skansbukta.
-Bioturbation structures filled with evaporitic material could provide secondary porosity.
-The picture shows cross-bedding in upper layers (Photo: M.\,Gutierrez).}
-				\label{fig:fieldwork_skansbukta}
+			\begin{minipage}[t]{0.5\textwidth}
+				\centering\includegraphics[width=1\textwidth]{figures/ML5}
+				\caption{Evaluation of a model conditioned on a SSF likelihood which narrows around the critical SSF$_c$ and thus amplifies the duality in the posterior ROV distribution (d).
+Shannon entropies are only slightly diminished (a,b).
+general cross-fault leakage remains a significant trap control mechanism (c).
+The resulting loss function realizations show a high divergence of differently risk-affine decisions.}
+				\label{fig:ML5}
 			\end{minipage}	
@@
 
-	\begin{figure}
-		\begin{minipage}{.94\textwidth}
-			\centering\includegraphics[width=\textwidth]{figures/cross_section}
-			\caption{Cross section of the targeted storage formation (Upper Triassic–Middle Jurassic sandstones of the Kapp Toscana Group), the overlying seal (Jurassic mudstones of the Agardfjellet Formation) and the prominent thrust décollement which is providing an additional seal.
-The indicated well Dh4 in the west is part of the Longyearbyen \ce{CO2} Lab, the logging took place in the east where the layers are outcropping.
-Modified from \cite{braathen_longyearbyen_2012}.}
-			\label{fig:cross_section}
-		\end{minipage}
-	\end{figure}
+%	\vspace{0.5em}
+%	\begin{figure}
+%		\begin{minipage}{0.95\textwidth}
+%			\begin{minipage}[t]{0.32\textwidth}
+%				\centering\includegraphics[width=1\textwidth]{figures/ML2}
+%				\caption{...}
+%				\label{fig:ML2}
+%			\end{minipage}
+%			\hspace{\fill}
+%			\begin{minipage}[t]{0.32\textwidth}
+%				\centering\includegraphics[width=1\textwidth]{figures/ML4}
+%				\caption{....}
+%				\label{fig:ML4}
+%			\end{minipage}
+%			\hspace{\fill}
+%			\begin{minipage}[t]{0.32\textwidth}
+%				\centering\includegraphics[width=1\textwidth]{figures/ML5}
+%				\caption{...}
+%				\label{fig:ML5}
+%			\end{minipage}	
+%		\end{minipage}
+%	\end{figure}
+%	\vspace{0.2em}
+
+	%\begin{figure}
+	%	\begin{minipage}{.94\textwidth}
+	%		\centering\includegraphics[width=\textwidth]{figures/cross_section}
+	%		\caption{...}
+	%		\label{fig:cross_section}
+	%	\end{minipage}
+	%\end{figure}
 %	\vspace{0.4em}
@@
 
-\begin{myblock}{Limited Information \& Seal Uncertainty}
-
-\vspace{0.3em}
-\begin{figure}
-	\begin{minipage}{0.95\textwidth}
-		\begin{minipage}[t]{0.36\textwidth}
-			\vspace{0pt}
-			\includegraphics[width=\textwidth]{figures/compaction.pdf}
-		\end{minipage}\hfill
-		\begin{minipage}[t]{0.22\textwidth}
-			\vspace{0pt}
-			\caption{
-			Comparison of porosity–depth trend studies and experimental results for shales and
-argillaceous sediments.
+\begin{myblock}{Conclusions}
+\begin{itemize}
+\item The degree of decision convergence can be considered a measure for the state of knowledge and its inherent uncertainty at the moment of decision making.
+\item This decisive uncertainty does not change in alignment with model uncertainty but depends on alterations of critical parameters and respective interdependencies.
+\item Actors are affected differently by one set of information, depending on their risk affinity.
+\item It is important to identify the model parameters which are most influential for the final decision in order to
+optimize the decision-making process.
+\end{itemize}
+These results and conclusions refer to a generic hydrocarbon system case study but are transferable to other fields where decisions
+are based on uncertain geological models, for example in hydrogeological or geothermal exploration.
+%\vspace{0.3em}
+%\begin{figure}
+%	\begin{minipage}{0.95\textwidth}
+%		\begin{minipage}[t]{0.36\textwidth}
+%			\vspace{0pt}
+%			\includegraphics[width=\textwidth]{figures/compaction.pdf}
+%		\end{minipage}\hfill
+%		\begin{minipage}[t]{0.22\textwidth}
+%			\vspace{0pt}
+%			\caption{
+%			Comparison of porosity–depth trend studies and experimental results for shales and
+%argillaceous sediments.
 While for E\&P projects the seal integrity is proven by the abundance of hydrocarbons, for \ce{CO2} storage it remains uncertain.
@@
 Still, extrapolation of experimental results has limited validity for natural anisotropy and the influence of chemical compaction \citep{mondol_experimental_2007}.
-			} \label{fig:compaction}
-		\end{minipage}\hfill
-		\begin{minipage}[t]{0.38\textwidth}
-			\vspace{0pt}
-			\includegraphics[width=\textwidth]{figures/leakage.pdf}
-			\caption{
-			Different leakage scenarios for a \ce{CO2} storage project which are subject to uncertainty.
+%			} \label{fig:compaction}
+%		\end{minipage}\hfill
+%		\begin{minipage}[t]{0.38\textwidth}
+%			\vspace{0pt}
+%			\includegraphics[width=\textwidth]{figures/leakage.pdf}
+%			\caption{
+%			Different leakage scenarios for a \ce{CO2} storage project which are subject to uncertainty.
 To prevent a spill out of the storage formation \textbf{(1)}, the storage capacity must be estimated.
@@
 Braathen, 15.6.2017).
-			} \label{fig:leakage}
-		\end{minipage}
-	\end{minipage}
-\end{figure}
+%			} \label{fig:leakage}
+%		\end{minipage}
+%	\end{minipage}
+%\end{figure}
 
 \end{myblock}\vfill
-
-\begin{myblock}{Outlook \& Motivation}
-		\begin{minipage}[t]{0.72\textwidth}
-			\vspace{0pt}
-			Marco studies \textbf{Georesources Management} in Aachen.
-During his various stays in Norway -- two semesters in Stavanger and summer courses in Oslo and Longyearbyen -- he became passionate about \textbf{risk} and \textbf{decision analysis}.
-Currently he is writing his master thesis in the field of\textbf{ geologic modelling and uncertainty quantification} in exploration.
-His focus will be on the application of a \textbf{value of information (VOI)} approach.
-A specific project to test assumptions and models would be beneficial.
-The high potential of future \textbf{\ce{CO2} storage projects} provides a suitable case for application.
-The mentioned methods are believed to contribute to urgently needed \textbf{cost reduction} and \textbf{risk management} of CCS projects.
-		\end{minipage}\hfill
-		\begin{minipage}[t]{0.23\textwidth}
-			\vspace{0em}
-			\includegraphics[width=1\textwidth]{figures/author}
-			\vspace{0.5em}
-		\end{minipage}
-
-%			\vspace{0.5em}
-
-\end{myblock}\vfill
-
-					As georesources manager who majored in geoscience I am curious about the structure and dynamics of our earth, but I like to put “these rocks” into a bigger picture.
-I am not only eager to understand what is going on below our feet, but also to estimate its value to society or economy and to communicate gained knowledge and accompanying risks to decision makers – in order to enable them to make good decisions.
-
-Currently I am writing my master thesis in the field of geologic modelling and uncertainty quantification in exploration.
-My focus will be on the application of a value of information (VOI) approach and using open source software to demonstrate light weight solutions for high-cost projects.
-
-I am looking for a specific project or data to test the mentioned methods on, preferably in Carbon Capture and Storage (CCS), which I believe has a great potential mitigating the dramatic consequences of climate change.
 
@@
 
+\begin{myblock}{Acknowledgements}
+This research poster is based on a master thesis submitted by Fabian Antonio Stamm to the Department of Computational Geoscience and Reservoir Engineering (CGRE) at the RWTH Aachen University in 2017.
+3D geological models were created using GemPy \citep{gmd-2018-61} and integrated into a probabilistic framework using PyMC \citep{salvatier2016pymc3}.
+
+		%\begin{minipage}[t]{0.72\textwidth}
+		%	\vspace{0pt}
+		%	This research poster is based on work conducted as part of a master thesis submitted to the Institute of Computational Geoscience and Reservoir Engineering (CGRE) at the RWTH Aachen University in 2017.
+		%	
+		%	\begin{itemize}
+		%	\item E-mail: fabian.stamm@rwth-aachen.de
+		%	\item Phone: + 49 174 5231499
+		%	\end{itemize}
+
+		%\end{minipage}\hfill
+		%\begin{minipage}[t]{0.23\textwidth}
+		%	\vspace{0em}
+		%	\includegraphics[width=1\textwidth]{figures/author}
+		%	\vspace{0.5em}
+		%\end{minipage}
+
+%			\vspace{0.5em}
+
+\end{myblock}\vfill
+
 		}\end{minipage}\end{beamercolorbox}

### 2018-04-04 Risk analysis/Poster/Poster STC 2017/poster_landscape.tex
Workwing on landscape poster for EGU

--- 2018-03-09 Risk analysis/Poster/Poster STC 2017/poster.tex
@@
 \documentclass{beamer}
-\usepackage[orientation=portrait,size=a0,scale=1.4,debug]{beamerposter}
+\usepackage[orientation=landscape,size=a0,scale=1.4,debug]{beamerposter}
 \mode<presentation>{\usetheme{RWTH2}}
@@
 
-\title{Bayesian Decision Theory in Structural Geological Modeling – How
+\title{Bayesian Decision Theory in Structural Geological Modeling - \\ How
 Reducing Uncertainties Affects Reservoir Value Estimations}
-%\author{Marco van Veen$^{1}$}
+%\author{Fabian Antonio Stamm$^{1}$}
 %\institute[RWTH Aachen University]{$^{1}$Institute of Computational Geoscience and Reservoir Engineering, RWTH Aachen University, Germany}
@@
 
-	\begin{column}{.382\textwidth}
+	\begin{column}{.249\textwidth}
 		\begin{beamercolorbox}[center]{postercolumn}
@@
 
-\begin{myblock}{Introduction}
+\begin{myblock}{1 - Introduction}
 
@@
 
-\begin{myblock}{Methods}
+\begin{myblock}{2.1 - Methods}
 \textbf{Construction of a 3D structural geological model:}
@@
 	\end{itemize}
-\textbf{Evaluation of modeling results:}
-	\begin{itemize}
-	\item Shannon entropy for uncertainty visualization (after \citet{wellmann2012uncertainties}.
-	\item Implementation of algorithms for structural analysis, trap recognition (see Figure \ref{fig:trap_cond}) and calculation of recoverable oil volumes (ROV)
-	\item Decision making based on the optimization of a case-specific custom loss function that considers differently risk-affine actors
-	\end{itemize}
+
+\end{myblock}\vfill
+
+		}\end{minipage}\end{beamercolorbox}
+	\end{column}
+
+\begin{column}{.499\textwidth}
+	\begin{beamercolorbox}[center]{postercolumn}
+		\begin{minipage}{.98\textwidth} % tweaks the width, makes a new \textwidth
+				\parbox[t][\columnheight]{\textwidth}{ % must be some better way to set the the height, width and textwidth simultaneously
+
+\begin{myblock}{2.2 - Methods (continued)}
+%\textbf{Evaluation of modeling results:}
+%	\begin{itemize}
+%	\item Shannon entropy for uncertainty visualization (after \citet{wellmann2012uncertainties}.
+%	\item Implementation of algorithms for structural analysis, trap recognition (see Figure \ref{fig:trap_cond}) and calculation of recoverable oil volumes (ROV)
+%	\item Decision making based on the optimization of a case-specific custom loss function that considers differently risk-affine actors
+%	\end{itemize}
 
@@
 
-							\vspace{0.5em}
+								\begin{minipage}[h]{0.329\textwidth}
+									\textbf{Evaluation of modeling results:}
+										\begin{itemize}
+										\item Shannon entropy for uncertainty visualization (after \citet{wellmann2012uncertainties}.
+										\item Implementation of algorithms for structural analysis, trap recognition (see Figure \ref{fig:trap_cond}) and calculation of recoverable oil volumes (ROV)
+										\item Decision making based on the optimization of a case-specific custom loss function that considers differently risk-affine actors
+										\end{itemize}
+								\end{minipage}
+								%\vspace{0.5em}
+								\begin{minipage}{0.327\textwidth}
 								\begin{figure}
-								\begin{minipage}{0.8\textwidth}
-									\centering\includegraphics[width=1\textwidth]{figures/Trap_Cond_H}
+									\centering\includegraphics[width=0.9\textwidth]{figures/Trap_Cond_H}
 									\caption{To be recognized as part of a trap, a reservoir voxel (1) has
-									to be positioned in the	footwall (2). Trap closure is defined by the seal shape and the normal fault (3). The maximum trap fill is defined by either the anticlinal spill point
-									(3;a) or a point of leakage across the fault, depending on juxtapositions with layers
+									to be positioned in the	footwall (2).
+The maximum trap fill is defined by either the anticlinal spill point (3;a) or a point of leakage across the fault, depending on juxtapositions with layers
 									underlying (3;b) or overlying the seal (3;c). The latter is only relevant if the
@@
 									\label{fig:trap_cond}
+									%Trap closure is defined by the seal shape and the normal fault (3).
+								\end{figure}
 								\end{minipage}
-							\end{figure}
-
-\end{myblock}
-
-		}\end{minipage}\end{beamercolorbox}
-	\end{column}
-
-	\begin{column}{.618\textwidth}
-		\begin{beamercolorbox}[center]{postercolumn}
-			\begin{minipage}{.98\textwidth} % tweaks the width, makes a new \textwidth
-				\parbox[t][\columnheight]{\textwidth}{ % must be some better way to set the the height, width and textwidth simultaneously
-
-\begin{myblock}{Results}
-		\begin{minipage}[t]{0.48\textwidth}
-					\vspace{0pt}
-					Posterior ROV probability distributions are realized depending on the nature of likelihoods implemented in the Bayesian inference step.
-Applying the custom loss function shows that the various Bayes estimators shift according to the characteristics of this underlying	value distribution.
-While bimodality and overall uncertainty lead to separation, risk-averse and risk-friendly decisions converge and decrease in expected loss given narrower unimodal distributions.
-A decisive factor in our model is the reliability of the sealing across the fault, as it defines the probability of complete trap failure.
-Two examples are summarized in Figures \ref{fig:ML4} and \ref{fig:ML5}.
-				\end{minipage}\hfill
-				\begin{minipage}[t]{0.48\textwidth}
-					\vspace{0em}
+								\begin{minipage}{0.327\textwidth}
 								\begin{figure}
@@
 								\end{figure}
-					\vspace{0.5em}
 								\end{minipage}
+\end{myblock}
 
-	\vspace{0.5em}
+\begin{myblock}{3 - Results}
+Posterior ROV probability distributions are realized depending on the nature of likelihoods implemented in the Bayesian inference step.
+Applying the custom loss function shows that the various Bayes estimators shift according to the characteristics of this underlying	value distribution.
+While bimodality and overall uncertainty lead to separation, risk-averse and risk-friendly decisions converge and decrease in expected loss given narrower unimodal distributions.
+A decisive factor in our model is the reliability of the sealing across the fault, as it defines the probability of complete trap failure.
+Two examples are summarized in Figures \ref{fig:ML4} and \ref{fig:ML5}.
+
+	\vspace{0.2em}
 	\begin{figure}
 		\begin{minipage}{0.98\textwidth}
-			\begin{minipage}[t]{0.5\textwidth}
-				\centering\includegraphics[width=1\textwidth]{figures/ML4}
-				\caption{Evaluation of a model conditioned on thickness and SSF likelihoods which assure a reliable trap closure and reinforce the probability of a high ROV.
+			%\hfill
+			\begin{minipage}[t]{0.49\textwidth}
+				\centering\includegraphics[width=0.8\textwidth]{figures/ML4}
+				\caption{Evaluation of a model conditioned on thickness and SSF likelihoods which reinforce the probability of a high ROV and reliable trap closure.
+Posterior Shannon entropy was reduced (b).
+Cross-fault leakage to overlying layers was eliminated as a trap control mechanism (c) and coincides with the disappearance of failure probability in the posterior ROV distribution (d).
+Respective loss functions visualize high decision convergence (e).}
+				%\caption{Evaluation of a model conditioned on thickness and SSF likelihoods which assure a reliable trap closure and reinforce the probability of a high ROV.
 Shannon entropy was significantly reduced in the posterior model (b).
@@
 			\end{minipage}
-			\hspace{\fill}
-			\begin{minipage}[t]{0.5\textwidth}
-				\centering\includegraphics[width=1\textwidth]{figures/ML5}
+			%\hspace{length=0.1\textwidth}
+			\begin{minipage}[t]{0.49\textwidth}
+				\centering\includegraphics[width=0.8\textwidth]{figures/ML5}
 				\caption{Evaluation of a model conditioned on a SSF likelihood which narrows around the critical SSF$_c$ and thus amplifies the duality in the posterior ROV distribution (d).
-Shannon entropies are only slightly diminished (a,b).
-general cross-fault leakage remains a significant trap control mechanism (c).
-The resulting loss function realizations show a high divergence of differently risk-affine decisions.}
+Shannon entropy was only slightly diminished (a,b), general cross-fault leakage remains a significant trap control mechanism (c).
+The resulting loss function realizations illustrate high divergence of differently risk-affine decisions (e).}
 				\label{fig:ML5}
 			\end{minipage}	
+			%\hfill
 		\end{minipage}
 	\end{figure}
-	\vspace{0.2em}
 
@@
 
-\begin{myblock}{Conclusions}
+		}\end{minipage}\end{beamercolorbox}
+	\end{column}
+
+	\begin{column}{.249\textwidth}
+		\begin{beamercolorbox}[center]{postercolumn}
+			\begin{minipage}{.98\textwidth}  % tweaks the width, makes a new \textwidth
+				\parbox[t][\columnheight]{\textwidth}{ % must be some better way to set the the height, width and textwidth simultaneously
+
+\begin{myblock}{4 - Conclusions}
 \begin{itemize}
@@
 
-\end{myblock}\vfill
+\end{myblock}
 
-\begin{myblock}{References}
+\begin{myblock}{5 - References}
 	\footnotesize
@@
 
-\begin{myblock}{Acknowledgements}
+\begin{myblock}{6 - Acknowledgements}
 This research poster is based on a master thesis submitted by Fabian Antonio Stamm to the Department of Computational Geoscience and Reservoir Engineering (CGRE) at the RWTH Aachen University in 2017.
@@
 		\end{column}
+
 \end{columns}

### 2018-04-05 Risk analysis/Poster/Poster STC 2017/poster_landscape_sfont.tex
Progress poster

--- 2018-04-04 Risk analysis/Poster/Poster STC 2017/poster_landscape.tex
@@
 \documentclass{beamer}
-\usepackage[orientation=landscape,size=a0,scale=1.4,debug]{beamerposter}
+%\usepackage{geometry}
+%\geometry{paperwidth=1920mm,paperheight=1000mm}
+\usepackage[orientation=landscape,size=a0,scale=1.3,debug]{beamerposter}
+%width=192,height=108,scale=1.4
 \mode<presentation>{\usetheme{RWTH2}}
@@
 %\institute[RWTH Aachen University]{$^{1}$Institute of Computational Geoscience and Reservoir Engineering, RWTH Aachen University, Germany}
-\author{Fabian Antonio Stamm$^{1}$, Miguel de la Varga$^{1}$, Alexander Schaaf$^{1}$, Florian Wellmann$^{1}$}
-\institute[RWTH Aachen University]{$^{1}$Institute of Computational Geoscience and Reservoir Engineering, RWTH Aachen University, Germany}
+\author{Fabian Antonio Stamm, Miguel de la Varga, Alexander Schaaf, Florian Wellmann}
+%\author{Fabian Antonio Stamm$^{1}$, Miguel de la Varga$^{1}$, Alexander Schaaf$^{1}$, Florian Wellmann$^{1}$}
+\institute[RWTH Aachen University]{Institute of Computational Geoscience and Reservoir Engineering, RWTH Aachen University, Germany}
 \date{\today}
@@
 \begin{myblock}{1 - Introduction}
-
 Structural geological modeling is of central importance for the assessment of uncertain hydrocarbon accumulations in potential reservoirs.
@@
 We examine respective decision making from a Bayesian perspective.
+%...building on recent developments that have transferred this probabilistic approach to address uncertainties in the field of structural geological modeling.
 %We examine respective decision making from a Bayesian perspective by considering two main approaches: (1) we treat geological modeling as a Bayesian inference problem, so that additional geological information can be incorporated as likelihood functions linked to prior parameters in a probabilistic framework; (2) we base the decision-making step on value estimation by optimizing a case-specific loss function.
@@
 	\begin{itemize}
+	%\item Model creation using GemPy (see \citet{gmd-2018-61})
 	\item Synthetic model of a simple anticline-fault trap in a potential hydrocarbon system (see Figure \ref{fig:unc_lik})
@@
 The inclusion of $z$-positional uncertainties affecting layer depths and fault offset are depicted in (C) and (D).
-Thicknesses of the three middle layers are defined by the distances of boundary points (E) and are thus directly dependent on (C).}
+Thicknesses of the three middle layers are defined by the distances of interface points (E).}
 										\label{fig:unc_lik}
@@
 	\begin{itemize}
+	%\item Utilizing PyMC (see \citet{salvatier2016pymc3})
 	\item Setting up a full probability model taking into account all parameter probability distributions
-	\item Bayesian updating: Conditioning of parameters on additionally observed data via likelihood functions
+	\item Bayesian inference: Conditioning of parameters on additionally observed data via likelihood functions
 	\item Approximation of posterior distributions through the use of MCMC sampling
+	\end{itemize}
+\textbf{Evaluation of modeling results:}
+										\begin{itemize}
+										\item Shannon entropy for uncertainty visualization (after \citet{wellmann2012uncertainties})
+										\item Implementation of customized algorithms for structural analysis, trap recognition (see Figure \ref{fig:trap_cond}) and calculation of recoverable oil volumes (ROV)
+										\item Decision making based on the optimization of a case-specific loss function that considers differently risk-affine actors
 										\end{itemize}
@@
 
-								\begin{minipage}[h]{0.329\textwidth}
-									\textbf{Evaluation of modeling results:}
-										\begin{itemize}
-										\item Shannon entropy for uncertainty visualization (after \citet{wellmann2012uncertainties}.
-										\item Implementation of algorithms for structural analysis, trap recognition (see Figure \ref{fig:trap_cond}) and calculation of recoverable oil volumes (ROV)
-										\item Decision making based on the optimization of a case-specific custom loss function that considers differently risk-affine actors
-										\end{itemize}
+								\begin{minipage}[h]{0.598\textwidth} %0.656
+								\textbf{Design of the custom loss function:}
+								%We regard decision making as allocation of resources based on estimation. 
+								We base value estimation on the 
+								\begin{equation}\label{eq:LFR_final}
+																L(\theta,\hat{\theta}) =
+																\begin{cases}
+																|\theta - \hat{\theta}|*r^{-0.5}, & \text{for } 0<\hat{\theta}<\theta  \\
+																|\theta-\hat{\theta}|*a*r, & \text{for } 0<\theta<\hat{\theta} \\
+																|\theta-\hat{\theta}|*b*r, & \text{for } \theta\leq0<\hat{\theta} \\
+																|\theta-\hat{\theta}|*c*r^{-0.5}, & \text{for } \hat{\theta}\leq0<\theta 
+																\end{cases},
+																\text{ with } a,b,c,r \in \mathbb{Q}.
+																\end{equation}
 								\end{minipage}
 								%\vspace{0.5em}
-								\begin{minipage}{0.327\textwidth}
+								\begin{minipage}{0.398\textwidth}
 								\begin{figure}
@@
 								\end{figure}
+								%\begin{figure}
+								%	\centering
+								%	\includegraphics[width=1\textwidth]{figures/model_network}
+								%	\caption{Illustration of the probabilistic model as a hierarchical Bayesian network after \citet{koller2009probabilistic}. Stochastic nodes are represented by ellipses (blue for
+								%	priors, yellow for likelihoods). Deterministic functions are depicted as triangles in
+								%	red. Arrows indicate direct connections from parent to child nodes.}
+								%	\label{fig:model_network}
+								%\end{figure}
 								\end{minipage}
-								\begin{minipage}{0.327\textwidth}
-								\begin{figure}
-									\centering
-									\includegraphics[width=1\textwidth]{figures/model_network}
-									\caption{Illustration of the probabilistic model as a hierarchical Bayesian network after \citet{koller2009probabilistic}. Stochastic nodes are represented by ellipses (blue for
-									priors, yellow for likelihoods). Deterministic functions are depicted as triangles in
-									red. Arrows indicate direct connections from parent to child nodes.}
-									\label{fig:model_network}
-								\end{figure}
-								\end{minipage}
+								%\begin{minipage}{0.327\textwidth}
+
+								%\end{minipage}
 \end{myblock}
 
-\begin{myblock}{3 - Results}
-Posterior ROV probability distributions are realized depending on the nature of likelihoods implemented in the Bayesian inference step.
-Applying the custom loss function shows that the various Bayes estimators shift according to the characteristics of this underlying	value distribution.
-While bimodality and overall uncertainty lead to separation, risk-averse and risk-friendly decisions converge and decrease in expected loss given narrower unimodal distributions.
-A decisive factor in our model is the reliability of the sealing across the fault, as it defines the probability of complete trap failure.
-Two examples are summarized in Figures \ref{fig:ML4} and \ref{fig:ML5}.
-
+\begin{myblock}{3.1 - Results}				
 	\vspace{0.2em}
@@
 			\begin{minipage}[t]{0.49\textwidth}
-				\centering\includegraphics[width=0.8\textwidth]{figures/ML4}
-				\caption{Evaluation of a model conditioned on thickness and SSF likelihoods which reinforce the probability of a high ROV and reliable trap closure.
-Posterior Shannon entropy was reduced (b).
-Cross-fault leakage to overlying layers was eliminated as a trap control mechanism (c) and coincides with the disappearance of failure probability in the posterior ROV distribution (d).
-Respective loss functions visualize high decision convergence (e).}
+				\centering\includegraphics[width=0.99\textwidth]{figures/ML4}
+				\caption{Evaluation of a model conditioned on thickness and SSF likelihoods which reinforce the probability of a high ROV and a reliable trap closure.}
 				%\caption{Evaluation of a model conditioned on thickness and SSF likelihoods which assure a reliable trap closure and reinforce the probability of a high ROV.
@@
 			\begin{minipage}[t]{0.49\textwidth}
-				\centering\includegraphics[width=0.8\textwidth]{figures/ML5}
-				\caption{Evaluation of a model conditioned on a SSF likelihood which narrows around the critical SSF$_c$ and thus amplifies the duality in the posterior ROV distribution (d).
-Shannon entropy was only slightly diminished (a,b), general cross-fault leakage remains a significant trap control mechanism (c).
-The resulting loss function realizations illustrate high divergence of differently risk-affine decisions (e).}
+				\centering\includegraphics[width=0.99\textwidth]{figures/ML5}
+				\caption{Evaluation of a model conditioned on a SSF likelihood which narrows around the critical SSF$_c$ and thus amplifies the duality in the posterior ROV distribution (d).}
 				\label{fig:ML5}
@@
 
+\begin{myblock}{3.2 - Results (continued)}
+Posterior ROV probability distributions are realized depending on the nature of likelihoods implemented in the Bayesian inference step.
+Applying the custom loss function shows that the various Bayes estimators shift according to the characteristics of this underlying	value distribution.
+While bimodality and overall uncertainty lead to separation, risk-averse and risk-friendly decisions converge and decrease in expected loss given narrower unimodal distributions.
+A decisive factor in our model is seal reliability, as it defines the probability of complete trap failure.
+Two examples are summarized in Figures \ref{fig:ML4} and \ref{fig:ML5}.
+\end{myblock}
+
 \begin{myblock}{4 - Conclusions}
+%\vspace{0.5em}
+%								\begin{figure}
+%									\begin{minipage}{0.8\textwidth}
+%									\centering
+%									\includegraphics[width=1\textwidth]{figures/model_network}
+%									\caption{Illustration of the probabilistic model as a hierarchical Bayesian network after \citet{koller2009probabilistic}. Stochastic nodes are represented by ellipses (blue for
+%									priors, yellow for likelihoods). Deterministic functions are depicted as triangles in
+%									red. Arrows indicate direct connections from parent to child nodes.}
+%									\label{fig:model_network}
+%									\end{minipage}
+%								\end{figure}
+
 \begin{itemize}
@@
 
-\begin{myblock}{6 - Acknowledgements}
-This research poster is based on a master thesis submitted by Fabian Antonio Stamm to the Department of Computational Geoscience and Reservoir Engineering (CGRE) at the RWTH Aachen University in 2017.
+\begin{myblock}{6 - Further Information}
+This research poster is based on a master thesis submitted by Fabian Antonio Stamm to the Institute of Computational Geoscience and Reservoir Engineering (CGRE) at the RWTH Aachen University in 2017.
+\\
 3D geological models were created using GemPy \citep{gmd-2018-61} and integrated into a probabilistic framework using PyMC \citep{salvatier2016pymc3}.
@@
 		%\begin{minipage}[t]{0.72\textwidth}	
-		%	\vspace{0pt}
-		%	This research poster is based on work conducted as part of a master thesis submitted to the Institute of Computational Geoscience and Reservoir Engineering (CGRE) at the RWTH Aachen University in 2017.
-		%	
 		%	\begin{itemize}

### 2018-04-06 Risk analysis/Poster/Poster STC 2017/poster_landscape_sfont.tex
Changed figures from PNG to PDF
Small corrections poster

@@
 									\begin{minipage}{0.95\textwidth}
-										\centering\includegraphics[width=1\textwidth]{figures/Uncertainties_Likelihoods}
+										\centering\includegraphics[width=1\textwidth]{figures/Uncertainties_Likelihoods.pdf}
 										\caption{The structural geological model illustrated as a 2D cross section (A) and a 3D voxel representation (B).
@@
 								\end{figure}
-\textbf{Integration in a probabilistic modeling framework for Bayesian analysis (see Figure \ref{fig:model_network}):}
+\textbf{Integration in a probabilistic modeling framework for Bayesian analysis:}
 	\begin{itemize}
@@
 								\begin{figure}
-									\centering\includegraphics[width=0.9\textwidth]{figures/Trap_Cond_H}
+									\centering\includegraphics[width=0.9\textwidth]{figures/Trap_Cond_H.pdf}
 									\caption{To be recognized as part of a trap, a reservoir voxel (1) has
@@
 			\begin{minipage}[t]{0.49\textwidth}
-				\centering\includegraphics[width=0.99\textwidth]{figures/ML4}
+				\centering\includegraphics[width=0.99\textwidth]{figures/ML4.pdf}
 				\caption{Evaluation of a model conditioned on thickness and SSF likelihoods which reinforce the probability of a high ROV and a reliable trap closure.}
@@
 			\begin{minipage}[t]{0.49\textwidth}
-				\centering\includegraphics[width=0.99\textwidth]{figures/ML5}
+				\centering\includegraphics[width=0.99\textwidth]{figures/ML5.pdf}
 				\caption{Evaluation of a model conditioned on a SSF likelihood which narrows around the critical SSF$_c$ and thus amplifies the duality in the posterior ROV distribution (d).}
@@
 \begin{myblock}{6 - Further Information}
-This research poster is based on a master thesis submitted by Fabian Antonio Stamm to the Institute of Computational Geoscience and Reservoir Engineering (CGRE) at the RWTH Aachen University in 2017.
+This research poster is based on a master thesis submitted by Fabian Antonio Stamm to the Institute of Computational Geoscience and Reservoir Engineering (CGRE) at the RWTH Aachen University in 2017 (see \citet{stamm2017}).
 \\

### 2018-04-06 Risk analysis/Poster/Poster STC 2017/poster_landscape_sfont.tex
EGU Poster Final (for me, awaiting review)

@@
 Hydrocarbon exploration and production is a high-risk, high-reward sector in which good decision making is indispensable.
-Actors in this field are faced with numerous uncertainties that have to be considered.
-We examine respective decision making from a Bayesian perspective.
+Utilizing a Bayesian approach, we examine the influence of model uncertainties and respective uncertainty changes on decision making of actors in this field.
+%Actors in this field are faced with numerous uncertainties that have to be considered. We examine respective decision making from a Bayesian perspective. 
 %...building on recent developments that have transferred this probabilistic approach to address uncertainties in the field of structural geological modeling.
@@
 
-\begin{myblock}{2.1 - Methods}
+\begin{myblock}{2 - Methods}
 \textbf{Construction of a 3D structural geological model:}
@@
 The inclusion of $z$-positional uncertainties affecting layer depths and fault offset are depicted in (C) and (D).
-Thicknesses of the three middle layers are defined by the distances of interface points (E).}
+Thicknesses of the three middle layers are defined by the distances of interface points (E) and are thus directly dependent on (C).}
 										\label{fig:unc_lik}
@@
 										\item Implementation of customized algorithms for structural analysis, trap recognition (see Figure \ref{fig:trap_cond}) and calculation of recoverable oil volumes (ROV)
-										\item Decision making based on the optimization of a case-specific loss function that considers differently risk-affine actors
+										\item Decision making based on the optimization of a case-specific loss function that is employed to estimate the ROV as the crucial reservoir value %that considers differently risk-affine actors
 										\end{itemize}
+
+	\vspace{0.1em}
 
@@
 
-\begin{myblock}{2.2 - Methods (continued)}
+\begin{myblock}{2 - Methods (continued)}
+\vspace{0.5em}
 %\textbf{Evaluation of modeling results:}
@@
 
-								\begin{minipage}[h]{0.598\textwidth} %0.656
-								\textbf{Design of the custom loss function:}
-								%We regard decision making as allocation of resources based on estimation. 
-								We base value estimation on the 
-								\begin{equation}\label{eq:LFR_final}
+								\begin{minipage}[h]{0.6\textwidth} %0.656
+								\textbf{Custom loss function design:}
+
+								We regard decision making as the process of estimating the posterior ROV by optimizing a customized loss function that considers the preferences of differently risk-affine actors.
+A decision that minimizes expected loss, according to such a function, is referred to as Bayes action \citep{davidson2015}.
+We extend the standard absolute-error loss function with a risk factor \textit{r} and several other weighting factors to express the expected loss relative to the nature and magnitude of deviation of the estimate $\hat{\theta}$ from the true value $\theta$:
+
+								\begin{equation*} %\label{eq:LFR_final}
 																L(\theta,\hat{\theta}) =
@@
 																\text{ with } a,b,c,r \in \mathbb{Q}.
-																\end{equation}
-								\end{minipage}
+								\end{equation*}
+
+								We define that normal overestimation is 25\% (\textit{a} = 1.25), fatal overestimation 100\% (\textit{b} = 2) and fatal underestimation 50\% (\textit{c} = 1.5) worse than normal underestimation.
+A risk factor \textit{r} = 1 represents risk neutrality, while \textit{r} $<$ 1 expresses risk-friendly, and \textit{r} $>$ 1 risk-averse behavior.
+
+								\end{minipage}\hspace{0.005\textwidth}
 								%\vspace{0.5em}
-								\begin{minipage}{0.398\textwidth}
+								\begin{minipage}{0.39\textwidth}
 								\begin{figure}
-									\centering\includegraphics[width=0.9\textwidth]{figures/Trap_Cond_H.pdf}
+									\centering\includegraphics[width=1\textwidth]{figures/Trap_Cond_H.pdf}
 									\caption{To be recognized as part of a trap, a reservoir voxel (1) has
@@
 									underlying (3;b) or overlying the seal (3;c). The latter is only relevant if the
-									critical Shale Smear Factor (SSF$_c$) is exceeded, as determined over D and T in (4).}
+									critical Shale Smear Factor (SSF$_c$) is exceeded, as determined over D and T in (4). Voxels connect to the model border are discarded.}
 									\label{fig:trap_cond}
@@
 
-\begin{myblock}{3.1 - Results}				
-	\vspace{0.2em}
+\begin{myblock}{3 - Results}				
+	\vspace{0.1em}
 	\begin{figure}
-		\begin{minipage}{0.98\textwidth}
+		\begin{minipage}{0.99\textwidth}
 			%\hfill
 			\begin{minipage}[t]{0.49\textwidth}
-				\centering\includegraphics[width=0.99\textwidth]{figures/ML4.pdf}
-				\caption{Evaluation of a model conditioned on thickness and SSF likelihoods which reinforce the probability of a high ROV and a reliable trap closure.}
+				\centering\includegraphics[width=1\textwidth]{figures/ML4.pdf}
+				\caption{Evaluation of a model conditioned on thickness and SSF likelihoods which assure a reliable trap closure and reinforce the probability of a high ROV.
+Shannon entropy is significantly reduced in the posterior model (b).
+Cross-fault leakage to overlying layers is eliminated as a trap control mechanism (c) and coincides with the disappearance of the probability of failure in the posterior ROV distribution (d).
+The respective loss functions are plotted in (e), showing high decision convergence.}
+				%\caption{Evaluation of a model conditioned on thickness and SSF likelihoods which reinforce the probability of a high ROV and a reliable trap closure.}
 				%\caption{Evaluation of a model conditioned on thickness and SSF likelihoods which assure a reliable trap closure and reinforce the probability of a high ROV.
@@
 				\label{fig:ML4}
-			\end{minipage}
+			\end{minipage}\hfill
 			%\hspace{length=0.1\textwidth}
 			\begin{minipage}[t]{0.49\textwidth}
-				\centering\includegraphics[width=0.99\textwidth]{figures/ML5.pdf}
-				\caption{Evaluation of a model conditioned on a SSF likelihood which narrows around the critical SSF$_c$ and thus amplifies the duality in the posterior ROV distribution (d).}
+				\centering\includegraphics[width=1\textwidth]{figures/ML5.pdf}
+				\caption{Evaluation of a model conditioned on a SSF likelihood which narrows around the critical SSF$_c$ and thus amplifies the duality in the posterior ROV distribution (d).
+Shannon entropy is only slightly diminished (a,b).
+general cross-fault leakage remains a significant trap control mechanism (c).
+The resulting loss function realizations show a high divergence of differently risk-affine decisions.}
+				%\caption{Evaluation of a model conditioned on a SSF likelihood which narrows around the critical SSF$_c$ and thus amplifies the duality in the posterior ROV distribution (d).}
 				\label{fig:ML5}
@@
 	%\end{figure}
-%	\vspace{0.4em}
+	%\vspace{0.2em}
+	%\vfill
 
-\end{myblock}\vfill
+\end{myblock}
 
@@
 
-\begin{myblock}{3.2 - Results (continued)}
+\begin{myblock}{3 - Results (continued)}
 Posterior ROV probability distributions are realized depending on the nature of likelihoods implemented in the Bayesian inference step.
-Applying the custom loss function shows that the various Bayes estimators shift according to the characteristics of this underlying	value distribution.
+Applying the custom loss function shows that the various Bayes actions shift according to the characteristics of this underlying	value distribution.
 While bimodality and overall uncertainty lead to separation, risk-averse and risk-friendly decisions converge and decrease in expected loss given narrower unimodal distributions.
@@
 
-\begin{myblock}{6 - Further Information}
-This research poster is based on a master thesis submitted by Fabian Antonio Stamm to the Institute of Computational Geoscience and Reservoir Engineering (CGRE) at the RWTH Aachen University in 2017 (see \citet{stamm2017}).
-\\
-3D geological models were created using GemPy \citep{gmd-2018-61} and integrated into a probabilistic framework using PyMC \citep{salvatier2016pymc3}.
+\begin{myblock}{6 - Further information}
+\small
+Geological models were created in a Python environment using GemPy \citep{gmd-2018-61} and integrated into a probabilistic framework via PyMC \citep{salvatier2016pymc3}.
+
+This research poster is based on a master thesis submitted to the Institute of Computational Geoscience and Reservoir Engineering (CGRE) at the RWTH Aachen University in 2017 (see \citet{stamm2017}).
 
 		%\begin{minipage}[t]{0.72\textwidth}	
+		%\textbf{Contact:}
 		%	\begin{itemize}
 		%	\item E-mail: fabian.stamm@rwth-aachen.de
-		%	\item Phone: + 49 174 5231499
+			%\item Phone: + 49 174 5231499
 		%	\end{itemize}
@@
 
-%			\vspace{0.5em}
+			%\vspace{0.4em}
+			%\vfill
 

### 2018-04-24 Risk analysis/Poster/Poster STC 2017/poster_landscape_sfont.tex
Poster update + theses correction

@@
 Hydrocarbon exploration and production is a high-risk, high-reward sector in which good decision making is indispensable.
-Utilizing a Bayesian approach, we examine the influence of model uncertainties and respective uncertainty changes on decision making of actors in this field.
+Utilizing a Bayesian approach, we examine the influence of model uncertainties and respective uncertainty changes on the decision making of actors in this field.
 %Actors in this field are faced with numerous uncertainties that have to be considered. We examine respective decision making from a Bayesian perspective. 
@@
 \begin{itemize}
-\item The degree of decision convergence can be considered a measure for the state of knowledge and its inherent uncertainty at the moment of decision making.
-\item This decisive uncertainty does not change in alignment with model uncertainty but depends on alterations of critical parameters and respective interdependencies.
-\item Actors are affected differently by one set of information, depending on their risk affinity.
-\item It is important to identify the model parameters which are most influential for the final decision in order to
-optimize the decision-making process.
+\item The degree of \textbf{decision convergence} can be considered a measure for the state of knowledge and its inherent uncertainty at the moment of decision making.
+\item This decisive uncertainty does not change in alignment with model uncertainty but depends on alterations of \textbf{critical parameters} and respective interdependencies.
+\item Actors are \textbf{affected differently} by one set of information, depending on their risk affinity.
+\item It is important to \textbf{identify the model parameters} which are most influential for the final decision in order to
+\textbf{optimize the decision-making process}.
 \end{itemize}

### 2018-04-24 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Poster update + theses correction

@@
 			\includegraphics[width=1\textwidth]{Figures/LFR_normal2.png}
-			\caption{Plotting of expected loss realizations after including the risk factor $r$ in the loss function (Equation \ref{eq:LFR_final}) for actors with risk affinities ranging from risk-averse ($r$ = 0.5 and 0.75), over risk-neutral ($r$ = 1), to risk-friendly ($r$ = 1.25 and $r$ = 1.5).
+			\caption{Plotting of expected loss realizations after including the risk factor $r$ in the loss function (Equation \ref{eq:LFR_final}) for actors with risk affinities ranging from risk-friendly ($r$ = 0.5 and 0.75), over risk-neutral ($r$ = 1), to risk-averse ($r$ = 1.25 and $r$ = 1.5).
 The loss function was applied the same normal distribution ($\mu=0; \sigma=500$) as in Figure \ref{fig:LF_4steps_normal}.}\label{fig:LFR_normal}
